### 2024-03-05 paper/main.tex
add draft
Former-commit-id: 40ea7082326e50fdda6c121e854ea816768bb999

 \documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{bm}
\usepackage{caption}
\usepackage{subcaption}
\graphicspath{{figures/}}

\title{Molecular transport in human pial perivasculature}

\author[1,x]{Rami Masri}
\author[1,x]{Miroslav Kuchta}
\author[1,x]{Marius Causemann}
\author[1,*]{Marie E. Rognes}
\affil[1]{Department of Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway}
\affil[x]{Author order to be discussed.}
\affil[*]{meg@simula.no}

\newcommand{\rami}[1]{\textcolor{blue}{#1}}

%\affil[+]{these authors contributed equally to this work}

%\keywords{Keyword1, Keyword2, Keyword3}

\begin{abstract}
\end{abstract}
\begin{document}

\flushbottom
\maketitle
% * <john.hammersley@gmail.com> 2015-02-09T12:07:31.197Z:
%
%  Click the title above to edit the author information and abstract
%
\thispagestyle{empty}

\section*{Introduction}

Molecular transport in perivascular spaces (PVSs) is established as a key pathway for human brain clearance and delivery. Previous studies indicate that molecules move rapidly in the subarachnoid space (SAS) and in PVSs surrounding pial arteries. Here, we aim to model and study this transport in a full pial perivascular and/or vascular network embedded in the SAS.

\begin{itemize}
\item Vinje et al~\cite{vinje2021brain} and references therein illustrates the location and characteristics of human pial perivascular spaces.
\item Mestre et al~\cite{mestre2022periarteriolar} study the properties of pial perivascular spaces in detail in mice.
\item Mestre et al~\cite{mestre2018flow} measure and characterize pial perivascular transport and estimate flow magnitudes (in mice).
\item Bollman et al~\cite{bollmann2022imaging} image the human pial vasculature at high resolution (7T).
\item Hornkjøl et al~\cite{hornkjol2022csf} model CSF flow and solute transport in the human SAS and brain parenchyma. 
\item Mardal et al~\cite{mardal2022mathematical} give a succinct introduction to MRI-based computational brain modelling. 
\end{itemize}

\section*{Methods}

\subsection*{Geometries and domains}

The computational domain $\Omega$ is defined by the union of the SAS $\Omega_{\rm sas}$ and brain parenchyma $\Omega_{\rm brain}$. The thin pia membrane encloses the brain, and thus the pial surface (of the brain) defines the SAS-brain interface $\Gamma_{\rm pia}$. A network of blood vessels are embedded in the SAS, potentially very near to or on the pial surface. We assume that a network of idealized (annular or semi-annular) perivascular spaces surround the vasculature. We represent the vasculature and associated perivasculature by a network of one-dimensional curves $\Gamma$ representing the network centerlines equipped with inner $R_1$ and outer $R_2$ radius information.

Surface representations of the outer SAS boundary, pial membrane, and delimiters of other CSF-filled spaces such as the lateral ventricles, third and fourth ventricle, and aqueduct are extracted from MR-images of human subject(s), see e.g. the pipeline described by Hornkjøl et al.~\cite{hornkjol2022csf}. From these surface representations, we create volumetric meshes of the SAS and brain using SVMTK~\cite{mardal2022mathematical}.

\begin{itemize}
    \item To discuss: How to generate subject-specific representations of the pial/SAS vasculature? Can we use contrast MRI? Can we use time-of-flight images? What are the right tools? Do any data sets already exist? 
    \item How to add radius information to vasculature representation? Do we need more information as well?
\end{itemize}

\begin{figure}
\caption{(A) Illustration of the location of pial vasculature and subarachnoid space. (B) Conceptual illustration of the vascular, perivascular, SAS and cortex domains (cross-section and lateral view).}
\label{fig:concept}
\end{figure}

\subsection*{Multi-dimensional transport model}

We consider a time-dependent convection-diffusion model~\cite{masri2023modelling} of a solute concentration $c$ in the network representing the pial vasculature and surrounding perivascular spaces $\Lambda$ and the surrounding three-dimensional SAS-brain domain $\Omega$ (Fig.~\ref{fig:concept}).

\subsection*{Diffusion}
We use the diffusion coefficient $D_{\rm gad}$ of Gadubutrol in free water in the SAS and PVS, and effective diffusion coefficient of Gadubutrol $D_{\rm brain}$ -- accounting for the extracellular space (ECS) porosity $\phi$ and the ECS tortuousity $\lambda$ but ignoring anisotropy  -- in the brain parenchyma~\cite{hornkjol2022csf}.

We solve for two concentration fields $c, \hat c$ in the domain $\Omega$ and in the network $\Lambda$. The 
network $\Lambda$ consists of 1D centerlines of vessels with their radii information and nodes. We consider the following model. 
\begin{alignat}{2}
\partial_t c - \nabla \cdot (D \nabla c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma & = f, && \quad \mathrm{in } \,\, \Omega \\ 
\partial_t (A\hat c) - \partial_s(\hat D A  \partial_s \hat c) + \xi P (\hat c - \overline{c})   & = A \hat f, && \quad \mathrm{in } \,\,  \Lambda 
 \end{alignat}
We split the boundary of $\Omega$ into $\Gamma_{\mathrm{skull}}$ and $\Gamma_{\mathrm{spine}}$,  and we enforce the following boundary conditions. 
\begin{alignat}{2}
D \nabla c \cdot \bm{n} &= 0, && \quad \mathrm{on} \,\, \Gamma_{\mathrm{skull}}, \\ 
D \nabla c \cdot \bm{n} &= g(t),  && \quad \mathrm{on} \,\, \Gamma_{\mathrm{spine}}.
\end{alignat}
The function $g$ is given in 
\begin{equation*}
    g(t) = \max(0, a(T_0 - t))
\end{equation*}
For the network, the boundary nodes are split into $\partial \Lambda^{\mathrm{in}}$  and $\partial \Lambda^{\mathrm{out}}$ based on the specified orientation of the network (\color{blue} to be specified more precisely (with a figure?)\color{black}). We impose the following conditions. 
\begin{alignat}{2}
D A \partial_s \hat c &= 0, && \quad \mathrm{on} \,\, \partial \Lambda^ {\mathrm{in}}, \\ 
D A \partial_s  \hat c &= \hat g(t),  && \quad \mathrm{on} \,\, \partial \Lambda^{\mathrm{out}} .
\end{alignat}
\color{blue} Running different models depending on the choices of $g$ and $\hat g$ \color{black}
 \subsection*{Convection}
We ignore the deformation of the domains (in this study), and thus set the domain velocity $w = 0$. We assume no significant convection in the brain $\Omega_{\rm brain}$, and set $u = u_{\rm brain} = 0$ there as a first approximation. We may consider non-zero convection in the SAS $u_{\rm sas} \not = 0$ e.g.~resulting from steady state CSF production as predicted by~\cite{hornkjol2022csf} in model variations, but may start with $u_{\rm sas} = 0$ as a baseline. In the perivascular network, we set an average axial velocity $u_{\rm pvs} = \langle u_{v, s} \rangle$ of $18.7$ $\mu$m/s as reported by Mestre et al.~\cite{mestre2018flow} in mouse perivascular spaces under normal physiological conditions. We may also consider a reduced perivascular velocity corresponding to hypertension in model variations.
\subsubsection*{Steady state PVS flow model} We solve 1D Darcy equations in the vessel network. The unknowns are the averaged PVS velocity, $u_{\rm pvs} = \langle u_{v,s} \rangle$, and the cross-section pressure $p_{\rm pvs} $ (assumed to be constant in each cross-section). The system is given by  \cite{daversin2022geometrically} 
\begin{alignat}{2}
A \,  u_{\rm pvs}   + \frac{\kappa}{\mu} \, A \, \partial_{s} p_{\rm pvs} & = 0, &&  \quad \rm in  \,\, \Lambda  \\ 
-\partial_s (A \, u_{\rm pvs}) & = f, && \quad \rm in  \,\, \Lambda .  
\end{alignat} 
In the above, $A$ is the area of the PVS, $\mu$ is the dynamic viscosity, and $\kappa$ is derived from the assumptio of Poiseuille
flow in the annular cross-section of the PVS \cite{daversin2022geometrically,tithof2022network}: 
\begin{equation}
\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
\end{equation}
On bifurcations, conservation of fluxes is enforced weakly and continuity of the pressure is enforced strongly (in the choice of the finite element spaces).

\subsubsection*{Case 1} We prescribe zero Dirichlet boundary conditions for the pressure, and vary $f$ \rami{include the reasoning behind having this $f$ as a source or sink term, and include the value that we opted for} to obtain a flow field with a similar magnititude to what is commonly reported in the literature.

\subsection*{Case 2} We prescribe a Robin type boundary condition on the outlet nodes. In particular, we set 
\begin{equation}
A  u_{\rm pvs} =  - \frac{\kappa}{ \mu} A\, \partial_s p_{\rm pvs} = \alpha (p_{\rm pvs} - p_0), \quad  \rm on \,\,\, \partial \Lambda^{\rm out} 
\end{equation}
On $\partial \Lambda^{\rm in}$, we prescribe homogenous Dirichlet data for the pressure (\rami{does this make sense?}). 
\rami{to be completed }
\subsection*{PVS permeability}
We will consider different values for the SAS-PVS permeability $\xi$, including $\xi = 0$ (sealed PVS), $\xi \approx \infty$ (PVS continuous with SAS), as well as reasonable intermediate values. We will ignore PVS-blood exchange, and thus only model the perivascular concentration at the network level (in addition to the SAS/brain concentration in the surroundings).

\subsection*{Boundary and initial conditions}
We set no flux on all boundaries unless otherwise specified. We start with an initial non-zero concentration in the first branch of the perivascular network and surrounding SAS (and zero concentration elsewhere).

\subsection*{Parameter values}

\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|} \hline 
 Parameter& Symbol& Value& Unit&Reference\\ \hline 
         parenchyma app. diffusion&  $d_p$&  $1.2 \times 10^{-4}$& $\text{mm}^2/\text{s}$  & \cite{valnes2020apparent}\\\hline 
         CSF app. diffusion&  $d_{CSF}$&  $3.8 \times 10^{-4}$& $\text{mm}^2/\text{s}$ & \cite{valnes2020apparent}\\\hline 
         PVS diffusion&  $d_{PVS}$&  $3.8 \times 10^{-4}$& $\text{mm}^2/\text{s}$ & \cite{valnes2020apparent}\\\hline 
         extracellular space volume fraction& $\phi$& 0.2& - &\cite{nicholson1981ion} \\\hline 
         PVS-parenchyma permeability&  $\xi_{PVS,P}$& $5 \times 10^{2} - 6 \times 10^{3} $  & $\text{m}^{-1}$ & \cite{koch2023estimates} \\\hline 
         PVS-CSF permeability&  $\xi_{PVS,CSF}$& $\{0, \text{inf}\}$ & $\text{m}^{-1}$ & \\ \hline 
 boundary concentration& $\hat{c}_D$& & &\\ \hline
    \end{tabular}
    \caption{Caption}
    \label{tab:parameters}
\end{table}

\subsection*{Quantities of interest}

\begin{itemize}
    \item Snapshots (e.g. coronal plane) of concentration at relevant times (15 min, 30 min, 1h, 2h, ...) 
    \item Total amount of tracer in the PVS, SAS and parenchyma
     \item Time to reach a threshold (e.g. 30\%) in different "places"
\end{itemize}

\section*{Results}

\begin{table}
\begin{center}
    \begin{tabular}{l|cc}
    \toprule
    Label & Description & Parameter values \\
    Model 0 & Baseline &  \\
    \bottomrule
    \end{tabular}
    \end{center}
    \caption{Overview of models}
\end{table}

\begin{figure}
     \centering
     \begin{subfigure}[b]{0.33\textwidth}
         \centering
         \includegraphics[width=\textwidth]{paper/figures/modelA_total_conc.png}
         \caption{arterial inflow, pure diffusion}
         \label{fig:y equals x}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.33\textwidth}
         \centering
         \includegraphics[width=\textwidth]{paper/figures/modelB_total_conc.png}
         \caption{SAS inflow, pure diffusion}
         \label{fig:three sin x}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.33\textwidth}
         \centering
         \includegraphics[width=\textwidth]{paper/figures/modelC_total_conc.png}
         \caption{arterial inflow, diffusion + PVS convection}
         \label{fig:five over x}
     \end{subfigure}
        \caption{Total tracer amount in parenchyma, CSF, arterial PVS and venous PVS}
        \label{fig:three graphs}
\end{figure}

\subsection*{Effect of interfaces (interface permeability)}

\begin{itemize}
    \item Model 0 (baseline): only diffusion, different diffusion coefficient in the parenchyma and SAS/PVS. No leak to vasculature (zero permeability). $\infty$ $\zeta_0$ at the PVS-SAS interface, $\zeta_1$ at the PVS-SAS interface. Koch et al could be a reference for PVS-parenchyma, maybe Tithof et al (2022, "A network model" ...) could give some ideas for parameters, also see Mestre et al, Nature comms, 2022. 
    \item 
    With altered interface permeabilities in the PVS (Model 1), and or in the parenchyma (Model 2) 
    \end{itemize}

\begin{figure}
    \caption{}
    \label{fig:1}
\end{figure}

\subsection*{Effect of convection}
\begin{itemize}
    \item 
    With convective flow in the PVS (zero flow in the parenchyma for now)
    \begin{itemize}
        \item 
        Model 3: PVS flow equal to 18.7 $\mu$m/s
        \item 
        Model 4: Better PVS flow (distributed, according to conservation of mass, assumptions on PVS area)
    \end{itemize}    
    \item 
    With convective flow in the ECS (out of scope here, for follow-up work?)
\end{itemize}

\begin{figure}
    \caption{}
    \label{fig:2}
\end{figure}
\subsection*{Effect of BBB}

Leakage to the blood, some sink/non-zero permeability for the PVS-blood (BBB) interface (Model N).

\begin{figure}
    \caption{}
    \label{fig:3}
\end{figure}

\subsection*{Effect of pia}

Effect of pia permeability (Model N-1)

\begin{figure}
    \caption{}
    \label{fig:4}
\end{figure}

\section*{Discussion}

\subsection*{Limitations}

\begin{itemize}
    \item Pulsating brain. Natural to consider in follow-up work.
    \item PVS-blood exchange. Natural to consider in model variation here, or follow-up work.
    \item Accounting for brain diffusive anisotropy. Little effect expected. Maybe something to consider in small follow-up work.
\end{itemize}

\bibliography{references}

\newpage
\section*{Acknowledgements (not compulsory)}

Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.

\section*{Author contributions statement}

Must include all authors, identified by initials, for example:
A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.

\section*{Additional information}

To include, in this order: \textbf{Accession codes} (where applicable); \textbf{Competing interests} (mandatory statement).

The corresponding author is responsible for submitting a \href{http://www.nature.com/srep/policies/index.html#competing}{competing interests statement} on behalf of all authors of the paper. This statement must be included in the submitted article file.

\end{document}

### 2024-03-05 scripts/pvs_network_netflow.py
add Maries network flow script
Former-commit-id: d148ae84d7d965a2d76501edc2e16d56124173f7

@@ -1,0 +1,432 @@
+import numpy as np
+import numpy.linalg
+from numpy import pi
+
+def _invR(beta):
+    "Helper function for evaluating \mathcal{R}^{-1}(beta)."
+    
+    val = np.pi/8*(beta**4 - 1 - (beta**2 - 1)**2/np.log(beta))
+    return val
+
+def _R(beta):
+    "Helper function for evaluating \mathcal{R}(beta)."
+    return 1.0/_invR(beta)
+
+def _delta(beta):
+    "Helper function for evaluating Delta(beta)."
+    delt = ((2 - (beta**2-1)/np.log(beta))**2)/(beta**4 - 1 - (beta**2 - 1)**2/np.log(beta))
+    return delt
+    
+def _beta(r_e, r_o):
+    return r_e/r_o
+
+def _alpha(l, P, R):
+    "Helper function for matrix/vector expression."
+    z = 1j
+    A1 = (0.5 - (1 - np.cos(l))/(l**2))
+    A2 = P*(1 - np.exp(z*l))/(2*l**2*R)
+    return (A1 + A2.real)
+
+def _xi(l):
+    "Helper function for matrix/vector expression."
+    z = 1j
+    xi1 = (np.exp(z*l) - 1)/l
+    return xi1
+
+
+
+def get_Q_single(beta, l, epsilon):
+    delta = _delta(beta)
+    return epsilon*delta*(0.5 - (1 - np.cos(l))/(l**2))
+    
+
+def get_Q_tandem(r0, betas, ls, epsilon):
+    '''
+    Get expected flow rate in tandem vessels a and b
+    
+    Args:
+        betas (list): aspect ratios of R1_a vs R2_a and R1_b vs R2_b
+        ls (list): fractional length of vessel a and b
+        epsilon (float): amplitude of vasomotion
+        
+    Returns: <Q> computed using (27)
+    '''
+    
+    Delta_a, Delta_b = [_delta(beta) for beta in betas]
+    Res_a, Res_b = [_R(beta)/r0**4 for beta in betas]
+    la, lb = ls
+    
+    term1 = (Delta_a*Res_a*la + Delta_b*Res_b*lb)/(2*(Res_a*la+Res_b*lb))
+    
[…]

### 2024-03-05 paper/main.tex
add small paragraph on meshing
Former-commit-id: 8fc05466de95b4ab7865a18bec0d741fea27c770

@@
 \section*{Methods}
+
+\subsection*{Meshing}
+
+\citet{hodneland2019new} developed a method to image the brains' microvasculature based on time-of-flight (ToF) and quantitative susceptibility mapping (QSM) to identify arterial and venous networks.
+We derive a detailed tetrahedral mesh of both the parenchyma and the surrounding CSF-filled spaces alongside a network representation of the vasculature from the anatomical data provided therein.
+
+\paragraph{3D domain}
+
+We merge the white and grey matter masks provided by \cite{hodneland2019new} to extract the pial surface.
+As no data on the CSF-filled spaces is available, we smooth and extend the pial surface by about 5\,mm to obtain an approximation of the SAS and include the inner fluid-filled cavities such as the ventricular system.
+Finally, we generate a tetrahedral mesh of both compartments using the fTetWild meshing algorithm \cite{hu2020fast}.
+
+\paragraph{Vessel networks}
+
+We skeletonize the binary masks of both the arterial and venous network using Kiminaro \cite{william_silversmith_2021_5539913} and obtain networks consisting of centerline and radius information of all detected vessels.
+The resulting networks are further smoothened using a spline interpolation technique and the nodes with lowest z-coordinate are identified as the networks' root nodes.
 

### 2024-03-11 paper/main.tex
Make .tex compile outside of Overleaf.

@@
 \usepackage{subcaption}
+%\usepackage{natbib}
 \graphicspath{{figures/}}
@@
 
-\citet{hodneland2019new} developed a method to image the brains' microvasculature based on time-of-flight (ToF) and quantitative susceptibility mapping (QSM) to identify arterial and venous networks.
+\cite{hodneland2019new} developed a method to image the brains' microvasculature based on time-of-flight (ToF) and quantitative susceptibility mapping (QSM) to identify arterial and venous networks.
 We derive a detailed tetrahedral mesh of both the parenchyma and the surrounding CSF-filled spaces alongside a network representation of the vasculature from the anatomical data provided therein.
@@
          \centering
-         \includegraphics[width=\textwidth]{paper/figures/modelA_total_conc.png}
+         \includegraphics[width=\textwidth]{modelA_total_conc.png}
          \caption{arterial inflow, pure diffusion}
@@
          \centering
-         \includegraphics[width=\textwidth]{paper/figures/modelB_total_conc.png}
+         \includegraphics[width=\textwidth]{modelB_total_conc.png}
          \caption{SAS inflow, pure diffusion}
@@
          \centering
-         \includegraphics[width=\textwidth]{paper/figures/modelC_total_conc.png}
+         \includegraphics[width=\textwidth]{modelC_total_conc.png}
          \caption{arterial inflow, diffusion + PVS convection}

### 2024-03-11 paper/main.tex
Add some Bedussi et al references.

@@
 \item Mestre et al~\cite{mestre2018flow} measure and characterize pial perivascular transport and estimate flow magnitudes (in mice).
+\item Bakker and coauthors studied perivascular spaces at the brain surface in a series of papers.
+They found that paravascular spaces (in particular periarterial spaces) represented low-resistance pathways~\cite{bedussi2017paravascular}, in communication with the surrounding CSF-filled subarachnoid spaces~\cite{bedussi2018paravascular}.
 \item Bollman et al~\cite{bollmann2022imaging} image the human pial vasculature at high resolution (7T).

### 2024-03-11 paper/main.tex
Add some more references and summary of context.
Former-commit-id: 42cf4fc6a21e981afbbbea9f9c120c77f42b3755

--- 2024-03-11 paper/main.tex
@@
 \item Mestre et al~\cite{mestre2018flow} measure and characterize pial perivascular transport and estimate flow magnitudes (in mice).
+\item
+  Weller and coauthors conducted pioneering studies of perivascular spaces, flow and transport~\cite{zhang1990interrelationships, zhang1992directional}.
+They identified thin sheaths of pial cells surrounding arteries and arterioles (but not veins or venules) on the brain surface and within the human brain itself~\cite{zhang1990interrelationships}.
+Further, they observed that tracers spread along perivascular (arterial, venous and capillar) spaces in (rat) grey matter, and in the subarachnoid space (to the cribriform plate and nasal lymphatics)..
+\item In a series of papers, Bakker and coauthors study perivascular anatomy and solute transport~\cite{bedussi2017paravascular, bedussi2018paravascular}.
+They find that the subarachnoid space, the cisterns, ventricles and penetrating periarteriolar spaces form a continuous cerebrospinal-fluid filled space surrounding and penetrating the rat brain\cite{bedussi2017paravascular}.
+Moreover, they demonstrate pulsatile and directional (antegrade) flow in perivascular (predominantly periarterial) spaces on the brain surface.
 \item Bollman et al~\cite{bollmann2022imaging} image the human pial vasculature at high resolution (7T).

### 2024-03-11 paper/main.tex
Updates from Overleaf

--- 2024-03-11 paper/main.tex
@@
 \usepackage{subcaption}
-%\usepackage{natbib}
 \graphicspath{{figures/}}
@@
 
-\cite{hodneland2019new} developed a method to image the brains' microvasculature based on time-of-flight (ToF) and quantitative susceptibility mapping (QSM) to identify arterial and venous networks.
+Hodneland et al~\cite{hodneland2019new} developed a method to image the brains' microvasculature based on time-of-flight (ToF) and quantitative susceptibility mapping (QSM) to identify arterial and venous networks.
 We derive a detailed tetrahedral mesh of both the parenchyma and the surrounding CSF-filled spaces alongside a network representation of the vasculature from the anatomical data provided therein.
@@
          \centering
-         \includegraphics[width=\textwidth]{modelA_total_conc.png}
+         \includegraphics[width=\textwidth]{paper/figures/modelA_total_conc.png}
          \caption{arterial inflow, pure diffusion}
@@
          \centering
-         \includegraphics[width=\textwidth]{modelB_total_conc.png}
+         \includegraphics[width=\textwidth]{paper/figures/modelB_total_conc.png}
          \caption{SAS inflow, pure diffusion}
@@
          \centering
-         \includegraphics[width=\textwidth]{modelC_total_conc.png}
+         \includegraphics[width=\textwidth]{paper/figures/modelC_total_conc.png}
          \caption{arterial inflow, diffusion + PVS convection}
@@
 \begin{figure}
-    \caption{}
+    \centering
+    \includegraphics[width = 0.9 \textwidth]{paper/figures/modelB_modelC_overview.png}
+    \caption{Tracer concentrations for a purely diffusive model with tracer inflow via the CSF-filled space (left) and a convective model with Tracer inflow via the arterial PVS (middle) and their difference (right) at 1, 6, 12, 18 and 24 hours after injection.}
     \label{fig:2}

### 2024-03-11 paper/main.tex

--- 2024-03-11 paper/main.tex
@@
 
-\cite{hodneland2019new} developed a method to image the brains' microvasculature based on time-of-flight (ToF) and quantitative susceptibility mapping (QSM) to identify arterial and venous networks.
+Hodneland et al~\cite{hodneland2019new} developed a method to image the brains' microvasculature based on time-of-flight (ToF) and quantitative susceptibility mapping (QSM) to identify arterial and venous networks.
+
 We derive a detailed tetrahedral mesh of both the parenchyma and the surrounding CSF-filled spaces alongside a network representation of the vasculature from the anatomical data provided therein.
@@
 \begin{figure}
-    \caption{}
+    \centering
+    \includegraphics[width = 0.9 \textwidth]{modelB_modelC_overview.png}
+    \caption{Tracer concentrations for a purely diffusive model with tracer inflow via the CSF-filled space (left) and a convective model with Tracer inflow via the arterial PVS (middle) and their difference (right) at 1, 6, 12, 18 and 24 hours after injection.}
     \label{fig:2}

### 2024-03-12 paper/main.tex
More references

@@
 Moreover, they demonstrate pulsatile and directional (antegrade) flow in perivascular (predominantly periarterial) spaces on the brain surface.
+\item
+  The architecture of pial and subarachnoid perivascular spaces in the rat brain was also studied by Thorne and colleagues~\cite{pizzo2018intrathecal, hannocks2018molecular}.
 \item Bollman et al~\cite{bollmann2022imaging} image the human pial vasculature at high resolution (7T).
@@
 \item Mardal et al~\cite{mardal2022mathematical} give a succinct introduction to MRI-based computational brain modelling. 
+\item
+  Rey et al~\cite{rey2023perivascular} (from the group of Sartinoranont) model and study perivascular transport in rat brain networks.
 \end{itemize}

### 2024-03-12 paper/main.tex
Add some notes from reading Eide and Ringstad (2024).

@@
 
+\newpage
+\appendix
+
+\section{Notes}
+
+Various notes associated with \cite{eide2024functional}:
+\begin{itemize}
+  \item
+    Can observations of tracer indicate compartmentalization without
+    the existence of physical barriers but in the presence of
+    e.g.~substantial flow? We made observations in this direction in
+    \cite{vinje2021brain}, relating to the then-active
+    shape-of-perivascular-spaces discussion. We can address this by
+    trying to quantify and evaluate the permeability of the PVS-SAS
+    interface - comparing model predictions with the timings and
+    patterns presented by \cite{eide2024functional}.
+  \item
+    Do the tracer primarily move from the PVS into the SAS or along
+    (more slowly) along the SAS? This is left as an open question by
+    \cite{eide2024functional} (page 3)
+  \item
+    How do altered ICP affect transport? Note that PVSAS transport is
+    impaired (delayed) with increasing mean wave amplitude in the ICP
+    signal and for iNPH patients (with enlarged PVSAS spaces).
+  \item
+    Very useful reference for PVS sizes (10--20 mm$^2$ in reference
+    patients, 15-70 mm$^2$ in iNPH), and average propagation speeds.
+\end{itemize}
+
 \end{document}

### 2024-03-15 paper/main.tex
Updates from Overleaf
Former-commit-id: 8599a4326329a8560609bb3cbb3348554613dd74

@@
 Hodneland et al~\cite{hodneland2019new} developed a method to image the brains' microvasculature based on time-of-flight (ToF) and quantitative susceptibility mapping (QSM) to identify arterial and venous networks.
-
-We derive a detailed tetrahedral mesh of both the parenchyma and the surrounding CSF-filled spaces alongside a network representation of the vasculature from the anatomical data provided therein.
+We derive a detailed tetrahedral mesh of the parenchyma and the surrounding CSF-filled spaces alongside a network representation of the vasculature from the anatomical data provided therein.
 

### 2024-04-04 scripts/map_to_net_flow_field.py
Have a go at describing the roots.
Former-commit-id: 14136d8f6d96c15568728e430ff6b6d626d9d92e

@@ -22,6 +22,11 @@ def map_vasculature(fname):
     # Reduce the vascular network to a bifurcating tree and keep a map
     # from tree element index to vessel segment indices.
 
+    print(mesh.num_vertices())
+    print(mesh.num_cells())
+    exit()
+          
+    
     # Start with the root node with lowest vertex index.
     first_roots = np.where(roots.array() == 2)[0]
     print("Vertices where root == 2: ", first_roots)
@@ -46,6 +51,7 @@ def map_vasculature(fname):
     
 if __name__ == '__main__':
 
-    map_vasculature("../mesh/networks/arteries_smooth.vtk")
+    #map_vasculature("../mesh/networks/arteries_smooth.vtk")
+    map_vasculature("../mesh/networks/venes_smooth.vtk")
 
     exit()

### 2024-04-04 scripts/branch_marking.py
Store output from branch_marking.py to separate directory.
Former-commit-id: 01385f1ab56ba5b871f252918a4079e83a0eddf4

@@ -1,3 +1,4 @@
+import os.path
 from collections import defaultdict
 from itertools import chain, combinations
 import networkx as nx
@@ -198,11 +199,13 @@ if __name__ == '__main__':
     from collections import defaultdict
     from itertools import repeat
 
+    subdir = "branches"
+    
     artery, artery_radii, artery_roots = read_vtk_network("../mesh/networks/arteries_smooth.vtk")
     # Find branch and mark the cells that make it up by unique color
     marking_branch, branch_colors, loop_colors = color_branches(artery)
     # This may be too many colors. An alternative is to minimally color the graph
-    df.File('unique_branch.pvd') << marking_branch
+    df.File(os.path.join(subdir, 'unique_branch.pvd')) << marking_branch
 
     mesh = marking_branch.mesh()
     Q = df.FunctionSpace(mesh, 'DG', 0)
@@ -236,12 +239,12 @@ if __name__ == '__main__':
         
         active_cell_f_array *= 0
     foo.vector().set_local(foo_values)
-    with df.XDMFFile('branch_length_read.xdmf') as xdmf:
+    with df.XDMFFile(os.path.join(subdir, 'branch_length_read.xdmf')) as xdmf:
         xdmf.write_checkpoint(foo, "branch_length")
-    df.File('branch_length.pvd') << foo
+    df.File(os.path.join(subdir, 'branch_length.pvd')) << foo
     
     minimal_marking_branch = minimal_color(marking_branch)
-    df.File('minimal_branch.pvd') << minimal_marking_branch
+    df.File(os.path.join(subdir, 'minimal_branch.pvd')) << minimal_marking_branch
     
     # Getting the network out in different formats
     mesh = marking_branch.mesh()
@@ -303,4 +306,4 @@ if __name__ == '__main__':
     reduced_color_f = df.MeshFunction('size_t', reduced_mesh, 1, 0)
     reduced_color_f.array()[:] = np.fromiter(reduced_edges, dtype='uintp')
 
-    df.File('reduced_branch.pvd') << reduced_color_f
+    df.File(os.path.join(subdir, 'reduced_branch.pvd')) << reduced_color_f

### 2024-04-04 scripts/branch_marking.py
Do not rescale branches to m by default.
Former-commit-id: 862b0547c08cb5d0a24c9021ecf172e627f0f296

@@ -199,9 +199,11 @@ if __name__ == '__main__':
     from collections import defaultdict
     from itertools import repeat
 
+    # Place output in separate subdirectory:
     subdir = "branches"
     
-    artery, artery_radii, artery_roots = read_vtk_network("../mesh/networks/arteries_smooth.vtk")
+    artery, artery_radii, artery_roots = read_vtk_network("../mesh/networks/arteries_smooth.vtk", rescale_mm2m=False)
+
     # Find branch and mark the cells that make it up by unique color
     marking_branch, branch_colors, loop_colors = color_branches(artery)
     # This may be too many colors. An alternative is to minimally color the graph

### 2024-04-15 scripts/map_to_net_flow_field.py
Start also constructing maps of indices between graphs and trees and radius information
Former-commit-id: 77d666ca5cd28022f09cc6c064efa493896451aa

@@ -13,11 +13,11 @@ def compute_shortest_paths(mesh, H, roots, output="_tmp"):
     # Find and mark shortest path from each of the nodes in G to the
     # 'roots'.
 
-    # Helper function to compute path length/weight based on edge data
+    # Helper function to compute path length based on edge data
     def path_length(G, path):
         length = 0.0
         for (i, _) in enumerate(path[:-1]):
-            length += G.get_edge_data(path[i], path[i+1])["weight"]
+            length += G.get_edge_data(path[i], path[i+1])["length"]
         return length
 
     # Create a vertex function that represents whether r0, r1 or r2 is
@@ -42,7 +42,7 @@ def compute_shortest_paths(mesh, H, roots, output="_tmp"):
     # Iterate over these nodes in an attempt to cover whole network
     print("Computing nearest (weighted) supply node for %d nodes " % len(nodes))
     for node in nodes:
-        paths = [nx.dijkstra_path(H, node, r, weight="weight")
+        paths = [nx.dijkstra_path(H, node, r, weight="length")
                  for r in roots]
         path_lengths = np.array([path_length(H, p) for p in paths])
         shortest = np.argmin(path_lengths)
@@ -59,19 +59,35 @@ def compute_shortest_paths(mesh, H, roots, output="_tmp"):
         
     return mf
 
-
 # Recursive function used to reduce full network to minimal set of
 # nodes and edges
-def add_branches(tree, a0, a, a_, Gr):
+def add_branches(tree, a0, a, a_, Gr, indices, radii, index_map, counter):
     degree = tree.degree(a)
 
     # Handle roots and leafs:
     if degree == 1:
+        # At a root node:
         if a0 == a:
             (_, b) = list(tree.edges(a))[0]
-            add_branches(tree, a0, b, a, Gr)
+            indices += [tree.get_edge_data(a, b)["index"]]
+            radii += [tree.get_edge_data(a, b)["radius"]]
+            add_branches(tree, a0, b, a, Gr, indices, radii, index_map, counter)
+        # At a leaf node:
         else:
-            Gr.add_edge(a0, a)
+            # End of branch, add new edge to minimal tree
+            counter += 1
+            Gr.add_edge(a0, a, index=counter)
+
+            # Update maps, and reset path information
[…]

### 2024-04-15 scripts/solver.py

@@ -43,6 +43,8 @@ def read_vtk_network(filename, rescale_mm2m=True):
 
     if rescale_mm2m:
         radii.array()[:] = netw["radius"] * 1e-3 # scale to m
+    else:
+        radii.array()[:] = netw["radius"]
 
     return mesh, radii, roots
 

### 2024-04-17 scripts/extract_surfaces.py
add config for different meshes

@@ -3,6 +3,9 @@ import pyvista as pv
 import numpy as np
 import skimage.morphology as skim
 import os
+import typer
+from plotting_utils import read_config
+from pathlib import Path
 
 def extract_surface(img):
     # img should be a binary 3D np.array
@@ -16,40 +19,50 @@ def binary_smoothing(img, footprint=None):
     openend = skim.binary_opening(img, footprint=footprint)
     return skim.binary_closing(openend, footprint=footprint)
 
-print("Loading data")
-os.makedirs("../mesh/surfaces", exist_ok=True)
-#load white matter data 
-wmdata = nibabel.load("../data/pcbi.1007073.s005.nii.gz")
-wmimg = wmdata.get_fdata() 
-surf = extract_surface(wmimg)
-surf = surf.smooth_taubin(n_iter=20, pass_band=0.05)
-surf.save("../mesh/surfaces/wm.ply")
+def extract_surfaces(configfile : str):
 
-#load grey matter data 
-gmdata = nibabel.load("../data/pcbi.1007073.s006.nii.gz")
-gmimg = gmdata.get_fdata() 
-surf_grey = extract_surface(gmimg)
-smooth_taubin_grey = surf_grey.smooth_taubin(n_iter=20, pass_band=0.05)
-smooth_taubin_grey.save("../mesh/surfaces/gm.ply")
+    config = read_config(configfile)
+    meshname = Path(configfile).stem
 
-# create parenchmya surface
-wmimg = np.pad(wmimg, ((0,0), (0,0), (0,8)))
-img = wmimg + gmimg
-surf_grey = extract_surface(img)
-smooth_taubin_parenchyma = surf_grey.smooth_taubin(n_iter=20, pass_band=0.05)
-smooth_taubin_parenchyma.save("../mesh/surfaces/parenchyma.ply")
+    os.makedirs(f"mesh/{meshname}/surfaces", exist_ok=True)
+    #load white matter data 
+    wmdata = nibabel.load("data/pcbi.1007073.s005.nii.gz")
+    wmimg = wmdata.get_fdata() 
+    surf = extract_surface(wmimg)
+    surf = surf.smooth_taubin(n_iter=20, pass_band=0.05)
+    surf.save(f"mesh/{meshname}/surfaces/wm.ply")
 
-## creating a skull 
-ball = skim.ball(5) 
-img = np.pad(img, (0, 20))
-for i in range(3):
-    img = skim.binary_dilation(img, footprint=ball)
[…]

### 2024-04-17 scripts/sas_flow.py
compute csf flow in largest sas component, add null pivot detection

@@ -7,19 +7,21 @@ from xii import *
 from petsc4py import PETSc
 import sympy as sp
 
-
+PETScOptions.set("mat_mumps_icntl_4", 3)  # mumps verbosity
+PETScOptions.set("mat_mumps_icntl_24", 1)  # null pivot detection
+#PETScOptions.set("mat_mumps_icntl_11", 1)  # error analysis
 
 # get mesh 
 sas = Mesh()
-with XDMFFile('mesh/volmesh/mesh.xdmf') as f:
+with XDMFFile('mesh/mid_mesh/volmesh/mesh.xdmf') as f:
     f.read(sas)
     gdim = sas.geometric_dimension()
-    vol_subdomains = MeshFunction('size_t', sas, gdim, 0)
-    f.read(vol_subdomains, 'label')
+    sas_components = MeshFunction('size_t', sas, gdim, 0)
+    f.read(sas_components, 'sas_components')
     sas.scale(1e-3)  # scale from mm to m
 
 # pick the sas mesh 
-sas_outer = EmbeddedMesh(vol_subdomains, 1) 
+sas_outer = EmbeddedMesh(sas_components, 1) 
 
 # create boundary markers 
 boundary_markers = MeshFunction("size_t", sas, sas.topology().dim() - 1)
@@ -27,9 +29,9 @@ boundary_markers.set_all(0)
 # Sub domain for efflux route (mark whole boundary of the full domain) 
 class Efflux(SubDomain):
     def inside(self, x, on_boundary):
-        return on_boundary
+        return on_boundary and x[2] > 0.2
 efflux = Efflux()
-efflux.mark(boundary_markers,1)
+efflux.mark(boundary_markers, 1)
 
 # translate markers to the sas outer mesh 
 boundary_markers_outer = MeshFunction("size_t", sas_outer, sas_outer.topology().dim() - 1, 0)
@@ -40,13 +42,13 @@ File("fpp.pvd") << boundary_markers_outer
 
 ds = Measure("ds", domain=sas_outer, subdomain_data=boundary_markers_outer)
 
+
 # Define function spaces for velocity and pressure
 cell = sas_outer.ufl_cell()  
 Velm = VectorElement('Lagrange', cell, 2)
 Qelm = FiniteElement('Lagrange', cell, 1) 
[…]

### 2024-04-17 scripts/peristalticflow.py
Add a table with some key characteristics and data for the appropriateness of the approach.

@@ -238,7 +238,7 @@ def graph_to_bifurcations(G, i0, relative_pvs_width):
         r_e[e0] = relative_pvs_width*d["radius"]
         L[e0] = d["length"]
 
-    print("... shortest and longest branch in T (mm):", min(L), max(L))
+        
     return (indices, paths, r_o, r_e, L)
 
 def test_graph_to_bifurcations():
@@ -354,6 +354,9 @@ def compute_pvs_flow(meshfile, output):
     mesh.init()
     DG0 = df.FunctionSpace(mesh, "DG", 0)
 
+    print("Vascular radii r_o (min, max, avg, std) (mm): ", radii.array().max(),
+          radii.array().min(), np.average(radii.array()), np.std(radii.array()))
+    
     # Note that the asymptotic estimate is is derived under the
     # assumption that when k L = 2 pi/lmbda L = O(1) i.e. when lmbda ~
     # 2 pi L.
@@ -369,14 +372,15 @@ def compute_pvs_flow(meshfile, output):
 
     # Slow vasomotion parameters
     f = 0.1                 # VLF
-    lmbda = 80.0            # VLF wave length (mm), 4% of cardiac in mice?
-    varepsilon = 0.1        # Relative wave amplitude of VLF wave?
+    lmbda = 2000.0            # VLF wave length (mm), 4% of cardiac in mice?
+    varepsilon = 0.01        # Relative wave amplitude of VLF wave?
 
     omega = 2*np.pi*f      # Angular frequency (Hz)
     k = 2*np.pi/lmbda      # Wave number (1/mm)
     
     Q = df.Function(DG0)
     u = df.Function(DG0)
+    Ls = []
     for i0 in roots:
         # Read minimal subtree from file. Note that graphml converts
         # ints to strings ...
@@ -386,6 +390,7 @@ def compute_pvs_flow(meshfile, output):
         
         # Map minimal subtree T into PVS net flow data representation
         (indices, paths, r_o, r_e, L) = graph_to_bifurcations(T, i0, beta)
+        Ls += L
         
         # Compute the PVS netflow in T
         network_data = (indices, paths, r_o, r_e, L, k, omega, varepsilon)
@@ -408,7 +413,12 @@ def compute_pvs_flow(meshfile, output):
             if (T_i < UNDEFINED):
                 Q.vector()[i] = avg_Q[index_map[i]]
[…]

### 2024-04-17 scripts/peristalticflow.py
- Update Supplementary methods text to be complete. - Fix computation of branch velocities to be flux/branch area and not flux/average area. - Improve outputs.

@@ -121,7 +121,7 @@ def add_branches(G, a0, a, a_, T, indices, radii, lengths, index_map):
         path_radius = average_radius(radii)
         path_length = sum(lengths)
         T.add_edge(a0, a,
-                    index=counter, radius=path_radius, length=path_length)
+                   index=counter, radius=path_radius, length=path_length)
         
         # Update index map
         index_map.array()[indices] = counter
@@ -145,7 +145,7 @@ def add_branches(G, a0, a, a_, T, indices, radii, lengths, index_map):
         path_radius = average_radius(radii)
         path_length = sum(lengths)
         T.add_edge(a0, a,
-                    index=counter, radius=path_radius, length=path_length)
+                   index=counter, radius=path_radius, length=path_length)
 
         # Update index map
         index_map.array()[indices] = counter
@@ -166,7 +166,7 @@ def add_branches(G, a0, a, a_, T, indices, radii, lengths, index_map):
         path_radius = average_radius(radii)
         path_length = sum(lengths)
         T.add_edge(a0, a,
-                    index=counter, radius=path_radius, length=path_length)
+                   index=counter, radius=path_radius, length=path_length)
 
         # Update index map
         index_map.array()[indices] = counter
@@ -178,7 +178,7 @@ def add_branches(G, a0, a, a_, T, indices, radii, lengths, index_map):
 
         add_branches(G, a, e1[1], a, T, [], [], [], index_map)
         add_branches(G, a, e2[1], a, T, [], [], [], index_map)
-        print("WARNING: Ignoring edge: ", e3)
+        print("WARNING: Ignoring edge: (degree > 3) ", e3)
         #add_branches(G, a, e3[1], a, T, [], [], [], index_map)
 
     if degree > 4:
@@ -344,6 +344,10 @@ def dg0_to_mf(u):
     mf.array()[:] = u.vector().get_local()
     return mf
 
+def print_stats(name, L, unit):
+    print("%s (avg, std, min, max): %.4g, %.4g, %.4g, %.4g (%s)" %
+          (name, np.average(L), np.std(L), L.min(), L.max(), unit))
+
 def compute_pvs_flow(meshfile, output):
 
     # Label the network supply nodes (FIXME: automate)
@@ -354,8 +358,7 @@ def compute_pvs_flow(meshfile, output):
     mesh.init()
[…]

### 2024-04-18 scripts/peristalticflow.py
Write Methods for PVS flow by peristalsis
Former-commit-id: e35a266dfafa5f0970053020eeb3209dac4aa344

@@ -462,9 +462,10 @@ if __name__ == '__main__':
 
     # To run:
     # $ mamba activate pvs_transport_env
-    # $ python3 peristalticflow.py --frequency 0.1 --wavelength 80.0 --amplitude 0.1 --beta 3.0 --recompute
+    # $ python3 peristalticflow.py --frequency 1.0 --wavelength 2000.0 --amplitude 0.01 --beta 3.0 --recompute # Cardiac
+    # $ python3 peristalticflow.py --frequency 0.1 --wavelength 80.0 --amplitude 0.1 --beta 3.0 
     #
-    # (--recompute is only needed upon first run)
+    # --recompute is only needed upon first run)
     
     import argparse
 

### 2024-04-18 scripts/pvs_network_netflow.py
Update schematic.

@@ -501,10 +501,33 @@ def run_murray_tree():
 
     return True
 
+def simple_schematic_data(alpha=1.0):
+    """Data for a single bifurcation test case. Data for network
+    consisting of three edges and a single bifurcation, with varying
+    thickness and length between the branches. Scale length by alpha
+    """
+    indices = [(0, 1, 2),]
+    paths = [(0, 1), (0, 2)]
+    
+    # Peristaltic wave parameters: wave length lmbda and (angular) wave number k
+    lmbda = 10.0
+    f = 1.0                 # frequency (Hz = 1/s)
+    omega = 2*np.pi*f       # Angular frequency (Hz)
+    k = 2*np.pi/lmbda       # wave number (1/mm)
+    varepsilon = 0.1        # AU 
+
+    beta=3.0
+    r_o = [0.3, 0.2, 0.1]     # Inner radii (mm) for each element/edge
+    r_e = [beta*r for r in r_o]  # Outer radii (mm) for each element/edge
+    L = [alpha*l for l in [2.0, 1.0, 2.0]]       # Element lengths (mm)
+        
+    data = (indices, paths, r_o, r_e, L, k, omega, varepsilon)
+    return data
+    
 def test_estimate_net_flow():
 
-    print("Running single bifurcation test case via estimate_net_flow")
-    data = single_bifurcation_data(0.1)
+    print("Running simple schematic test case via estimate_net_flow")
+    data = simple_schematic_data(1.0)
     
     avg_Q_prime, avg_u_prime = estimate_net_flow(data)
 

### 2024-06-07 scripts/mesh_sanity.py
update
Former-commit-id: 5d8bcddb84f1f75639cf9588fb7921d8296dad1c

@@ -1,0 +1,98 @@
+import dolfin as df
+import networkx as nx
+import numpy as np
+
+
+def color_connected_components(mesh):
+    '''
+    Cell function with colors corresponding to connected components of mesh graph. 
+    Largest components first.
+    '''
+    print(f'Coloring components in mesh with {mesh.num_cells()} cells')
+    tdim = mesh.topology().dim()
+    mesh.init(tdim-1, tdim)  #  Facet to cell
+
+    g = nx.Graph()
+    g.add_edges_from(tuple(facet.entities(tdim))
+                     for facet in df.facets(mesh) if len(facet.entities(tdim)) > 1)
+
+    ccs = sorted(list(nx.algorithms.connected_components(g)))
+
+    
+    components = df.MeshFunction('size_t', mesh, tdim, 0)
+    values = components.array()
+    for (color, component) in enumerate(ccs, 1):
+        print(f'Component {color} with {len(component)} cells')
+        values[list(component)] = color
+    return components
+
+# ----------------------------------------------------------------------
+
+if __name__ == '__main__':
+    from slepc4py import SLEPc
+    from petsc4py import PETSc
+    
+    path = '/home/mirok/Downloads/mesh.xdmf'
+    mesh = df.Mesh()
+    with df.XDMFFile(mesh.mpi_comm(), path) as f:
+        f.read(mesh)
+    gdim = mesh.geometry().dim()
+
+    # Extract mesh as largest connected component
+    if mesh.mpi_comm().size == 1:
+        cell_f = color_connected_components(mesh)        
+        ncomps = len(np.unique(cell_f.array()))
+        if ncomps > 1:
+            mesh = df.SubMesh(mesh, cell_f, 1)
+
+    df.set_log_level(10)
+    # Sanity check in terms of vector Poisson. Don't want any zero
+    # eigenvalues
+    V = df.VectorFunctionSpace(mesh, 'CG', 1)
+    print(f'dim(V) = {V.dim()}')
+    bcs = df.DirichletBC(V, df.Constant((0, )*gdim), 'on_boundary')
+    
+    u, v = df.TrialFunction(V), df.TestFunction(V)
+
+    a = df.inner(df.grad(u), df.grad(v))*df.dx
+    m = df.inner(u, v)*df.dx
[…]

### 2024-08-30 scripts/solver.py
fix assertion
Former-commit-id: 214646c22198b7e210d61233259f409c3462c9ff [formerly 214646c22198b7e210d61233259f409c3462c9ff [formerly 0c4974b21641f3d6ad556e22876684c3b324c491]]
Former-commit-id: 154711742751d7f873cc7b50457b9ff35488f3df
Former-commit-id: 28b75353b27e119b68d77c55c89a962f29008d43

@@ -106,7 +106,7 @@ def pcws_constant(subdomains, values):
     u = Function(FunctionSpace(mesh, 'DG', 0))
     for k,v in values.items():
         u.vector().add_local(v*(subdomains.array()==k))
-    assert np.allclose(np.unique(u.vector()), np.sort(list(values.values())))
+    assert np.allclose(np.unique(u.vector()), np.unique(list(map(float, values.values()))))
     return u
 
 def as_P0_function(mesh_f):

### 2024-08-30 scripts/time_dependent.py
add bdm
Former-commit-id: 73d073f73009eff312433191ce8b24dc79e94df3 [formerly 73d073f73009eff312433191ce8b24dc79e94df3 [formerly ea0ab7721b4c83072d06dace8663afef49e20636]]
Former-commit-id: daf88fadc06a02f8ea9b8a55aba216a306b4a2ed
Former-commit-id: 15957bdb6f5c7acd36b5e661fa73f67ecf3aa2d3

@@ -112,13 +112,13 @@ def run_simulation(configfile: str):
                            doms=[PARID])
     File("bm.pvd") << bm
 
-    xi_dict = {0:Constant(0), CSFID: Constant(pvs_csf_permability), 
-               PARID: Constant(pvs_parenchyma_permability)}
-    phi_dict = {CSFID: Constant(1),  PARID: Constant(ecs_share)}
+    xi_dict = {0:0, CSFID: pvs_csf_permability, 
+               PARID: pvs_parenchyma_permability}
+    phi_dict = {CSFID: 1,  PARID: ecs_share}
 
     phi = pcws_constant(vol_subdomains, phi_dict)
-    Ds = pcws_constant(vol_subdomains, {CSFID: Constant(sas_diffusion),  # csf
-                                        PARID: Constant(parenchyma_diffusion) # parenchyma
+    Ds = pcws_constant(vol_subdomains, {CSFID: sas_diffusion,  # csf
+                                        PARID: parenchyma_diffusion # parenchyma
                                         })
     xi_a = pcws_constant(artmarker, xi_dict)
     xi_v = pcws_constant(veinmarker, xi_dict)
@@ -146,11 +146,16 @@ def run_simulation(configfile: str):
         vel_file = config["csf_velocity_file"]
 
         with XDMFFile(vel_file) as file:
-            sm = xii.EmbeddedMesh(vol_subdomains, [CSFID, LVID, V34ID])
-            CG3 = VectorFunctionSpace(mesh, "CG", 3)
-            velocity_csf = Function(CG3)
+            if "bdm" in vel_file:
+                vel_space = VectorFunctionSpace(mesh, "DG", 1)
+            else:
+                vel_space = VectorFunctionSpace(mesh, "CG", 3)
+            velocity_csf = Function(vel_space)
             file.read_checkpoint(velocity_csf, "velocity")
-        
+        if "bdm" in vel_file:
+            dx_s = Measure("dx", mesh, subdomain_data=vol_subdomains)
+            assert abs(assemble(div(velocity_csf)*dx_s(CSFID))) < 1e-14
+
         File("csf_velocity.pvd") << velocity_csf
     else:
         velocity_csf = Constant([0]*gdim)
[…]

### 2024-08-30 scripts/test_map_on_global.py
minor change

@@ -63,6 +63,6 @@ with XDMFFile("uh.xdmf") as file:
     file.write_checkpoint(uhdg, "uh")
 
 uh.set_allow_extrapolation(True)
-uh_global_dg = interpolate(uh_global, VectorFunctionSpace(mesh, "DG", 1))
+uh_global_dg = interpolate(uh, VectorFunctionSpace(mesh, "DG", 1))
 with XDMFFile("uh_gobal.xdmf") as file:
     file.write_checkpoint(uh_global_dg, "uh_global")

### 2024-09-04 scripts/create_diff_plot.py
interpolate for diff on different meshes

@@ -21,6 +21,11 @@ def plot_model_diff(modela: str, modelb: str, t:int, type,  cmax:float=None, fil
     sasb = get_result(modelb, "sas", t)
     artb = get_result(modelb, "artery", t)
     venb = get_result(modelb, "vein", t)
+    if sasa.n_points > sasb.n_points:
+        sasb = sasa.sample(sasb, pass_point_data=False)
+    elif sasa.n_points < sasb.n_points:
+        sasa = sasb.sample(sasa, pass_point_data=False)
+
     sasa["diff"] = sasa[f"c_{t}"] - sasb[f"c_{t}"]
     arta["diff"] = arta[f"c_{t}"] - artb[f"c_{t}"]
     vena["diff"] = vena[f"c_{t}"] - venb[f"c_{t}"]
@@ -28,6 +33,7 @@ def plot_model_diff(modela: str, modelb: str, t:int, type,  cmax:float=None, fil
     arta.set_active_scalars("diff")
     vena.set_active_scalars("diff")
     clim = (-cmax, cmax) if cmax is not None else None
+        
 
     if type=="overview":
         filename = f"{plotdir}/{modela}_{modelb}_diff_{t}.png"

### 2024-09-04 scripts/label_arteries.py
add artery label plot
Former-commit-id: f9eca99915d566797ef954e2c9d06232ba8202f1

@@ -1,0 +1,34 @@
+import pyvista as pv
+filename = "plots/labeled_arteries.png"
+art = pv.read("mesh/networks/arteries_tubes.vtk")
+art["radius"] *= 1e3
+
+pointlabels = [("ACA", (0.0836656391620636, 0.12674154341220856, 0.10104840993881226)),
+               ("ICA",(0.09893926978111267, 0.11347689479589462, 0.03919529169797897)),
+            ("ICA",(0.0687378594571459, 0.1128146693515864, 0.03797117973496493)),
+           ("MCA",(0.11279711872339249, 0.12369046360254288, 0.0656440332531929)),
+           ("MCA",(0.05993403121829033, 0.1203613206744194, 0.06074688211083412)),
+            ("BA",(0.08326485485531022, 0.10635409315410013, 0.038923607855650535)),
+            ("PCA",(0.06962447613477707, 0.057159677147865295, 0.07165350019931793)),
+            ("PCA",(0.09846239537000656, 0.07383400201797485, 0.0722760483622551)),
+            ]
+points = [p for l, p in pointlabels]
+labels = [l for l, p in pointlabels]
+bar_args=dict(title="radius (mm)", vertical=False,
+                    height=0.07, width=0.6, position_x=0.2,
+                    position_y=0.0, title_font_size=24,
+                    label_font_size=24, fmt="%.1f")
+pl = pv.Plotter(off_screen=True)
+pl.add_mesh(art, cmap="coolwarm", scalar_bar_args=bar_args)
+pl.add_point_labels(points, labels, font_size=24,
+ margin=4, shadow=True, shape="rounded_rect",shape_opacity=0.7,
+ render_points_as_spheres=True,
+ shape_color="white",
+ #justification_horizontal="center", 
+ always_visible=True,
+ point_color="red", point_size= 18, show_points=True)
+pl.camera_position = 'zx'
+pl.camera.roll += 90
+pl.camera.azimuth -= 25
+pl.camera.zoom(1.6)
+pl.screenshot(filename, transparent_background=False)

### 2024-09-04 scripts/time_dependent.py
add flush output
Former-commit-id: 5ab96a5e65342c315f4a85003472b1c0da5a7355 [formerly d1895ed2c3001b47e04a0bdb26c2d7d1fb62c514]
Former-commit-id: df571a28f084b4b75d01596a89e3d943380746e8

@@ -344,6 +344,10 @@ def run_simulation(configfile: str):
     xdmfsas = XDMFFile(results_dir + f'{modelname}_sas.xdmf') 
     xdmfart = XDMFFile(results_dir + f'{modelname}_artery.xdmf') 
     xdmfven = XDMFFile(results_dir + f'{modelname}_vein.xdmf') 
+    xdmfsas.parameters["flush_output"] = True
+    xdmfart.parameters["flush_output"] = True
+    xdmfsas.parameters["flush_output"] = True
+
     pvdarteries = File(results_dir + f'{modelname}_artery.pvd') 
     pvdvenes = File(results_dir + f'{modelname}_vein.pvd') 
     xfiles = (xdmfsas, xdmfart, xdmfven)

### 2024-09-04 scripts/uhat_prod.py
add metrics output to yaml

@@ -6,6 +6,7 @@ from xii import *
 from petsc4py import PETSc
 import sympy as sp
 import typer
+import yaml
 
 def compute_avg_flow(csf_flow_model:  str):
 
@@ -59,6 +60,14 @@ def compute_avg_flow(csf_flow_model:  str):
         xdmf.write_checkpoint(pvs_flow_vec, "velocity")
 
     File(f'results/csf_flow/{csf_flow_model}/pvs_flow.pvd') << pvs_flow_vec
+
+
+    length = assemble(1*dx(domain=artery))
+    umean = assemble(pvs_flow*dx) / length
+    umax = norm(pvs_flow.vector(), "linf")
+    metrics = dict(umean=umean,umax=umax)
+    with open(f'results/csf_flow/{csf_flow_model}/pvs_metrics.yml', 'w') as outfile:
+        yaml.dump(metrics, outfile, default_flow_style=False)
 
 if __name__ == "__main__":
     typer.run(compute_avg_flow)

### 2024-09-04 scripts/sas_flow_Hdiv.py
add metrics output to yaml
Former-commit-id: 41d5cf826c66bc12ef4473779b36eda0220e0ab9 [formerly 41d5cf826c66bc12ef4473779b36eda0220e0ab9 [formerly 0a3f596c092b0c4b83bc635a61c117ed7149dc25]]
Former-commit-id: 8d28dad11395024f4d05dff487d1e436700ea510
Former-commit-id: 8aed260829015f9e1c1b6c7290a3db7be24f7cc2

@@ -10,6 +10,7 @@ import numpy_indexed as npi
 from plotting_utils import read_config
 from IPython import embed
 from pathlib import Path
+import yaml
 
 def directsolve(a, L, bcs, a_prec, W):
     wh = Function(W)
@@ -229,7 +230,7 @@ def compute_sas_flow(configfile : str, elm:str = 'BDM'):
                     + assemble(div(uh_global)*div(uh_global)*dxglob(LVID)) \
                     + assemble(div(uh_global)*div(uh_global)*dxglob(V34ID))
 
-    divu = assemble(div(uh)*div(uh)*dx)
+    divu = assemble(sqrt(div(uh)*div(uh))*dx)
 
     assert np.isclose(divu, 0)
 
@@ -248,7 +249,17 @@ def compute_sas_flow(configfile : str, elm:str = 'BDM'):
     assert np.isclose(divu, divu_global_csf)
     #assert np.isclose(assemble(inner(uh_global, uh_global)*dxglob(CSFNOFLOWID)), 0)
 
-    print(f"div u = {divu}")
+    # collect key metrics:
+    metrics
+    sas_vol = assemble(1*dx(domain=sas_outer))
+    umean = assemble(sqrt(inner(uh, uh))*dx) / sas_vol
+    umag = project(sqrt(inner(uh, uh)), FunctionSpace(sas_outer, "CG", 1),
+                   solver_type="cg", preconditioner_type="hypre_amg")
+    umax = norm(umag.vector(), 'linf') 
+    metrics = dict(umean=umean, umax=umax, divu=divu)
+
+    with open(f'{results_dir}/metrics.yml', 'w') as outfile:
+        yaml.dump(metrics, outfile, default_flow_style=False)
 
 if __name__ == "__main__":
     typer.run(compute_sas_flow)

### 2024-09-04 scripts/sas_flow_TH.py
in-core fact
Former-commit-id: 36a6834645f0347c0924feb536327a8468a5cb4a [formerly 56a1ec3fee3112e80b414ab5325fea3a7ea49a60]
Former-commit-id: d36fd2ef00c3ec017303ae0efcad35aa5c1d52e9

@@ -62,7 +62,7 @@ PETScOptions.set("mat_mumps_icntl_4", 3)  # mumps verbosity
 PETScOptions.set("mat_mumps_icntl_35", 1)  # BLR feature
 PETScOptions.set("mat_mumps_cntl_7", 1e-8)  # BLR eps
 #PETScOptions.set("mat_mumps_icntl_32", 1)  # forward elimination during solve (useful, but not passed on by petsc)
-PETScOptions.set("mat_mumps_icntl_22", 1)  # out-of-core to reduce memory
+#PETScOptions.set("mat_mumps_icntl_22", 1)  # out-of-core to reduce memory
 #PETScOptions.set("mat_mumps_icntl_11", 1)  # error analysis
 #PETScOptions.set("mat_mumps_icntl_25", 2)  # turn on null space basis
 

### 2024-10-14 scripts/sas_flow_Hdiv.py
add PVS flow field from production
Former-commit-id: 3ab42c3a2f56f0826d8cf3f6ddb003d3b1494fa3 [formerly b4154cce31c9052dc05fe00a5c7a2386005be509]
Former-commit-id: c57d90e8f1cb31892ea71b1534531d02ea91834e

@@ -59,7 +59,7 @@ CSF_NO_FLOW_CSF_ID = 7
 
 def get_normal_func(mesh, scale=Constant(1)):
     n = FacetNormal(mesh)
-    V = VectorFunctionSpace(mesh, "DG", 1)
+    V = VectorFunctionSpace(mesh, "DG", 0)
     u = TrialFunction(V)
     v = TestFunction(V)
     a = inner(u,v)*ds

### 2024-10-14 scripts/test_map_on_global_coords_shift.py
add PVS flow field from production
Former-commit-id: 3ab42c3a2f56f0826d8cf3f6ddb003d3b1494fa3 [formerly 3ab42c3a2f56f0826d8cf3f6ddb003d3b1494fa3 [formerly b4154cce31c9052dc05fe00a5c7a2386005be509]]
Former-commit-id: c57d90e8f1cb31892ea71b1534531d02ea91834e
Former-commit-id: 0bd95bef475f0497a418dd6690b9ba9541d26e3f

@@ -31,12 +31,12 @@ def map_dg_on_global(uh, uh_global=None, parentmesh=None):
     # shift coords towards the cell center
     C_shift = 0.5*(C + np.repeat(Cc, int(dofs_per_cell/eldim), axis=0))
     P_shift = 0.5*(P + np.repeat(Pc, int(dofs_per_cell/eldim), axis=0))
-
     #idxmap = npi.indices(np.round(P_shift, ndigits), np.round(C_shift, ndigits), axis=0)
     idxmap = map_kdtree(P_shift, C_shift)
     for i in range(eldim):
         uh_global.vector()[idxmap*eldim + i] = uh.vector()[i::eldim]
     return uh_global
+    
 CSFID = 1
 PARID = 2
 LVID = 3

### 2024-10-14 paper/cut.tex

\begin{table}
\begin{center}
  \begin{tabular}{ll|llllll|ll}
    \toprule
    & Description & $R$ & $\xi_{\rm PVS-SAS}$, $\xi_{\rm PVS-PAR}$ & $\beta_{\rm efflux}$ & $D$ & $\hat{u}$ & $\mathbf{u}_{\rm SAS, PAR}$ & $g_{\rm influx}$ & $c_0$ \\
    \midrule
    A & Baseline  & $R_2 = 2 R_1$ & $\xi$, $\tilde \xi$\cite{koch2023estimates} &  $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $D^{\rm gad}$\cite{sykova2008diffusion, valnes2020apparent}  & $\hat{u} = \hat{u}_{\rm prod}$ & $\mathbf{u}_{\rm SAS} = \mathbf{u}_{\rm prod}$, $\mathbf{u}_{\rm PAR} = 0$ & $> 0$ & 0 \\
    B & PVS "sheaths" & $R_2 = 2 R_1$ & $0, 0.5 \xi, \xi, 2 \xi, 1000 \xi \approx \infty$, --\cite{koch2023estimates} & -- & --  &  --  & --  & -- & -- \\
    C & Larger PVS & $R_2 = N R_1$ & -- & -- & --  &  --  & -- & -- & -- \\
    D & PVS I & $R_2 = 2 R_1$ & $\xi$, -- & -- & --  &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow$  & -- & -- & -- \\
    E & PVS II & $R_2 = 2 R_1$ & $\xi$, -- & -- & --  &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow\uparrow$  & -- & -- & -- \\
    F & PVS III & $R_2 = 2 R_1$ & $\xi$, -- & -- & $10 D_{\rm PVS}$ &  $\hat{u} = \hat{u}_{\rm prod}$  & -- & -- & -- \\
    G & Glymphatics & $R_2 = 2 R_1$ & $\xi$, -- & -- & \cite{sykova2008diffusion, valnes2020apparent} &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow$  & $\mathbf{u}_{\rm PAR}$ > 0 & -- & -- \\
    H & Sleep &  &  & &  &  &  & & \\
    I & Pathology I &  &  & &  &  &  & & \\
    J & Pathology II &  &  & &  &  &  & & \\
    \midrule
    X & Clearance & $R_2 = 2 R_1$ & $50\%$, $\xi$ & -- & $D^{\rm gad, \tau, A\beta}$  &  $\hat{u} = \hat{u}_{\rm prod}$  & -- & $0$ & $1$ \\
    Y & (PVS) & $R_2 = 2 R_1$ & $50\%$, $\xi$ & -- & $D^{\rm gad, \tau, A\beta}$  &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow$  & -- & -- & -- \\
    Z & (Glymphatic) & $R_2 = 2 R_1$ & $50\%$, $\xi$ & -- & $D^{\rm gad, \tau, A\beta}$  &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow$  & $\mathbf{u}_{\rm PAR}$ > 0 & -- & -- \\
    \bottomrule
    \end{tabular}
    \end{center}
\caption{Overview of models (-- denotes as immediately above). Model A represents a baseline scenario with semi-permeable barriers between the PVS and SAS ($\xi_{\rm PVS-SAS} \approx 50\%$, \mer{we try to estimate this value by a bit of trial-and-error from the observations of timings of PVS-SAS and SAS from~\cite{eide2024functional}})~\cite{bedussi2017paravascular} and astrocytic endfeet forming barriers within the parenchyma~\cite{koch2023estimates}. $R_1$ and $R_2$ denotes the inner and outer radius of the PVS, with PVS width comparable to vascular diameter ($R_2 - R_1 \approx 2 R_1$)\cite{mestre2018flow} as a baseline. The extracranial solute efflux permeability $\beta_{\rm efflux}$ is uniformly distributed over the outer SAS boundary with a reasonable rate as baseline\cite{hornkjol2022csf, eide2021clinical, ringstad2024glymphatic}. At baseline, the effective diffusion coefficients $D^{\rm x}_{\rm SAS} = D^{\rm x}_{\rm PVS}$ of the SAS and PVS represent the free diffusion coefficient of Gadubutrol (${\rm x} = {\rm gad}$) in CSF (water), and $D^{\rm x}_{\rm PAR}$ the effective diffusion coefficient of human cortical tissue, all at body temperature. At baseline, we consider net flow due to CSF-production in the SAS (and PVS): $\mathbf{u}_{\rm SAS} > 0$ (and $\hat{u} > 0$), but no additional effects from perivascular pulsatility nor bulk flow in the parenchyma $\mathbf{u}_{\rm PAR} = 0$. Due to the timescale for human intracranial tracer transport (minutes to hours) compared to typical (cardiac, respiratory, slow vasomotion) human intracranial pulsatility (0.1 -- 1 Hz), we consider net (constant-in-time) flow fields only in subsequent model variations. Model B represents a PVS-SAS interface configuration with minimal or more permeable structural barriers~\cite{eide2024functional}, while Model C represents enlarged PVS. Models D, E, and F represent three perivascular pathway scenarios with more rapid flow in the PVS induced by vasomotion (PVS I)~\cite{gjerde2023directional} or a net fluid source in the PVS (1D Stokes flow)/net pressure difference between PVS inlets and outlets (PVS II), or with enhanced effective diffusion due to mixing (PVS III)~\cite{hornkjol2022csf, troyetsky2021dispersion, vinje2023human}, respectively. Model G attempts to represent a glymphatic--like convective flow within the parenchyma. \mer{Ideas could be:  (a) network induces arterial and venous sources in a Darcy porous media flow model (reflects the glymphatic hypothesis), (b) Croci-style \cite{croci2019uncertainty}, (c) \cite{vinje2023human}} Model H represents physiological alterations associated with sleep (lower CSF production, enhanced diffusion within the parenchyma, enhanced PVS pulsatility), while Models I--J represents variations associated with pathologies (e.g.~dementia, hypertension, angiopathies, CSF-disorders and/or others). Models X--Z represent version of the aforementioned scenarios but for studying metabolite clearance (instead of solute influx).}
\label{tab:scenarios}
\end{table}

### 2024-10-17 scripts/detail_plot.py
add detail plots
Former-commit-id: d9f3b83273db6d05447885fc85812c0bd68c922c [formerly d9f3b83273db6d05447885fc85812c0bd68c922c [formerly e4cc70747dfb736b3b8e02b10a20838c7201790b]]
Former-commit-id: 3a1f440605e29a4c84671c011ded0672ea476797
Former-commit-id: 64d6b7bfb077d95cbda17a9d5257f64bd4f194a8

@@ -1,0 +1,116 @@
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import ImageGrid
+import numpy as np
+import typer
+from typing import List
+from plotting_utils import time_str, read_config, detail_plot, get_result
+from generate_plot import plot_model
+import matplotlib as mpl
+import cmocean
+from label_arteries import pointlabels
+from test_map_on_global_coords_shift import map_kdtree
+import pyvista as pv
+from generate_synthseg_mesh import PARID
+from mpl_toolkits.axes_grid1.inset_locator import inset_axes
+
+percentile = 95
+
+def get_tangent(v, n):
+    return v - np.outer(np.dot(v,n), n)
+
+def get_normal(netw, pidx):
+    neighors = netw.point_neighbors(pidx)
+    p1, p2 = netw.points[pidx], netw.points[neighors[0]]
+    return (p1 - p2) / np.linalg.norm(p1 - p2)
+
+def get_radius(netw, pidx):
+    return netw.extract_points(pidx)["radius"].max()
+
+def compare_models(modelname:str):
+    times = np.array([1, 3, 6, 12, 24])*3600
+    n_arteries = len(pointlabels)
+    points = np.array([c for n,c in pointlabels[:n_arteries]])
+    artlabels = [n for n,c in pointlabels[:n_arteries]]
+    fig = plt.figure(figsize=(len(artlabels)*4, len(times)*3), frameon=True)
+    grid = ImageGrid(fig, 111, nrows_ncols=(len(artlabels), len(times)),
+                     axes_pad=(0.27, 0.03),
+                    #cbar_location="right",
+                    #cbar_mode="each",
+                   )
+
+    subdomains = pv.read_meshio("mesh/standard/standard.xdmf")
+    par = subdomains.extract_cells(subdomains["label"]==PARID).extract_surface()
+    par.compute_normals(inplace=True, cell_normals=False)
+
+    sas = get_result(modelname, "sas", times)
+    art = get_result(modelname, "artery", times)
+
+    idx = map_kdtree(art.points, points)
+    normals = [get_normal(art, pidx) for pidx in idx]
+    radii = [get_radius(art, pidx) for pidx in idx]
+
+    pvs_radius_ratio = 2
+
+    cmap = cmocean.cm.algae_r
[…]

### 2024-10-17 scripts/label_arteries.py

@@ -1,5 +1,6 @@
 import pyvista as pv
-
+import numpy as np
+import seaborn as sns
 pointlabels = [("ACA", (0.0797341, 0.132915, 0.0937325)),
                ("ICA1",(0.0942801, 0.11788, 0.0505869)),
                ("ICA2",(0.0679531, 0.118951, 0.0485859)),
@@ -13,7 +14,11 @@ pointlabels = [("ACA", (0.0797341, 0.132915, 0.0937325)),
 if __name__=="__main__":
    filename = "plots/labeled_arteries.png"
    art = pv.read("mesh/networks/arteries_tubes.vtk")
+   labelcoords = lambda label: art.bounds[::2]
+
    art["radius"] *= 1e3
+   col = "fuchsia"      
+   arrow_length = 0.03
 
    points = [p for l, p in pointlabels]
    labels = [l for l, p in pointlabels]
@@ -23,15 +28,23 @@ if __name__=="__main__":
                      label_font_size=24, fmt="%.1f")
    pl = pv.Plotter(off_screen=True)
    pl.add_mesh(art, cmap="coolwarm", scalar_bar_args=bar_args)
-   pl.add_point_labels(points, labels, font_size=24,
-   margin=4, shadow=True, shape="rounded_rect",shape_opacity=0.7,
-   render_points_as_spheres=True,
-   shape_color="white",
-   #justification_horizontal="center", 
-   always_visible=True,
-   point_color="red", point_size= 18, show_points=True)
-   pl.camera_position = 'zx'
-   pl.camera.roll += 90
-   pl.camera.azimuth -= 25
+   pl.add_points(np.array(points), render_points_as_spheres=False, 
+                 point_size=20, color=col)
+   center = art.center_of_mass() - np.array([0,0, 0.047])
+   label_coords = []
+   for i, (l, p) in enumerate(pointlabels):
+      dir = np.array(p) - center
+      dir /= np.linalg.norm(dir)
+      arr = pv.Arrow(start=np.array(p), direction=dir, scale=arrow_length,
+                     tip_length=1e-3, tip_radius=5e-3, shaft_radius=1e-2)
+      pl.add_mesh(arr, color=col)
+      label_coords.append(p + arrow_length*dir)
+      
+   pl.add_point_labels(label_coords, labels, show_points=False, 
+      always_visible=True, shape_color=col)
+   pl.camera_position = 'xz'
+   #pl.camera.roll += 10
[…]

### 2024-10-17 scripts/generate_plot.py
add detail plots
Former-commit-id: d9f3b83273db6d05447885fc85812c0bd68c922c [formerly e4cc70747dfb736b3b8e02b10a20838c7201790b]
Former-commit-id: 3a1f440605e29a4c84671c011ded0672ea476797

@@ -5,6 +5,7 @@ import typer
 from typing import Tuple
 from plotting_utils import get_result, time_str, clip_plot, detail_plot, isosurf_plot
 import yaml
+pv.global_theme.allow_empty_mesh = True
 
 CSFID = 1
 PARID = 2
@@ -12,7 +13,7 @@ LVID = 3
 V34ID = 4
 CSFNOFLOWID = 5
 
-def plot_model(modelname: str, t:int, type, cmax:float=None, filename:str=None):
+def plot_model(modelname: str, t:int, type:str, cmax:float=None, filename:str=None):
     plotdir = f"plots/{modelname}"
     os.makedirs(plotdir, exist_ok=True)
 
@@ -31,15 +32,13 @@ def plot_model(modelname: str, t:int, type, cmax:float=None, filename:str=None):
     
     if type=="isosurf":
         title = f"time: {time_str(t)} h"
-        sas = sas.cell_data_to_point_data()
-        sas[f"c_{t}"] *= (1 - (0.8*sas["label"]==PARID))
-        return isosurf_plot(sas, [art, ven], filename, title, clim=clim, cmap="matter",
+        csf = sas.extract_cells(np.isin(sas["label"], [CSFID, LVID, V34ID, CSFNOFLOWID]))
+        par = sas.extract_cells(sas["label"]==PARID)
+        #par[f"c_{t}"] *= 0.2
+        csf.merge(par, inplace=True, merge_points=False)
+        #sas[f"c_{t}"] *= (1 - (0.8*sas["label"]==PARID))
+        return isosurf_plot(sas, [art], filename, title, clim=clim, cmap="tempo",
                             cbar_title="concentration (mmol/l)")
     
-    elif type=="detail":
-        center = (0.2877, 0.17, 0.23)
-        return detail_plot(sas, [art, ven], filename, center, clim=clim, cmap="matter",
-                cbar_title="concentration (mmol/l)")
-
 if __name__ == "__main__":
     typer.run(plot_model)

### 2024-10-17 scripts/mean_concentrations.py
add detail plots

@@ -12,13 +12,15 @@ from IPython import embed
 import pandas as pd
 from matplotlib.markers import MarkerStyle
 import yaml
+from branch_marking import color_branches
+from label_arteries import pointlabels
+from test_map_on_global_coords_shift import map_kdtree
 
 CSFID = 1
 PARID = 2
 LVID = 3
 V34ID = 4
 CSFNOFLOWID = 5
-
 
 def compare_concentrations(modelname:str, times:List[int]):
     sas_conc, subd_marker = get_result_fenics(modelname, "sas", times)
@@ -84,8 +86,6 @@ def compare_concentrations(modelname:str, times:List[int]):
     with open(f'results/{modelname}/mean_concentrations.yml', 'w') as outfile:
         yaml.dump(metrics, outfile, default_flow_style=False)
 
-    times = times[1:]
-
     def ii_project(f, V):
         u = TrialFunction(V)
         v = TestFunction(V)  
@@ -100,10 +100,47 @@ def compare_concentrations(modelname:str, times:List[int]):
 
     artery = art_radii.function_space().mesh()
     DG0a = FunctionSpace(artery, "DG", 0)
+    V = art_conc[0].function_space()
     pvs_radii = Function(DG0a)
     pvs_radii.vector().set_local(pvs_ratio_artery*art_radii.vector().get_local())
+    priority = InterfaceResolution(subdomains=subd_marker,
+                                       resolve_conflicts={(CSFID, PARID): 4})
     pvs_shape = Circle(radius=pvs_radii, degree=40, quad_rule='midpoint')
-    c_averages = [ii_project(Average(c, artery, pvs_shape), DG0a) for c in sas_conc]
+    c_averages = [ii_project(Average(c, artery, pvs_shape, 
+                                     normalize=True, resolve_interfaces=priority), V) for c in sas_conc]
+
+    segments, segids ,_ = color_branches(artery)
+    dxs = Measure("dx", artery, subdomain_data=segments)
+    points = np.array([c for n,c in pointlabels])
+    artlabels = [n for n,c in pointlabels]
+    cellidx = map_kdtree(DG0a.tabulate_dof_coordinates(), points)
+    artsegids = [int(segments.array()[i]) for i in cellidx]
+
+    seglengths = [assemble(1*dxs(si)) for si in artsegids]
+    conc_at_point, avg_conc_around_point= {}, {}
[…]

### 2024-10-17 scripts/plotting_utils.py

@@ -32,6 +32,8 @@ def get_result(modelname, domain, times):
         grid[f"{r.name()}_{t}"] = r.vector()[:]
     if domain == "sas":
         grid["label"] = label.array()[:]
+    else:
+        grid["radius"] = label.vector()[:]
     return grid
 
 def get_result_fenics(modelname, domain, times):
@@ -124,47 +126,57 @@ def clip_plot(csf, par, networks, filename, title, clim, cmap, cbar_title):
 
 def isosurf_plot(sas, networks, filename, title, clim, cmap, cbar_title):
     import pyvista as pv
-    n = 3
+    n = 5
     if clim is not None:
-        contours = sas.contour(np.linspace(clim[0] + clim[1]/n, clim[1], 3))
+        contours = sas.contour(np.linspace(clim[0] + clim[1]/n, clim[1], n))
     else:
         contours = sas.contour(n)
-    #embed()
+        
     pl = pv.Plotter(off_screen=True)
+    pl.add_mesh(sas.outline(), color="white")
     pl.add_mesh(contours, opacity=0.75, clim=clim, cmap=cmap, show_scalar_bar=False)
-    pl.add_mesh(sas.extract_surface(), opacity=0.3, clim=clim, cmap=cmap,
-                scalar_bar_args=dict(title=cbar_title, vertical=False,
-                                    height=0.1, width=0.6, position_x=0.2,
-                                    position_y=0.0, title_font_size=36,
-                                    label_font_size=32))
+    #pl.add_mesh(sas.extract_surface(), opacity=0.3, clim=clim, cmap=cmap,
+    #            scalar_bar_args=dict(title=cbar_title, vertical=False,
+    #                                height=0.1, width=0.6, position_x=0.2,
+    #                                position_y=0.0, title_font_size=36,
+    #                                label_font_size=32))
     for netw in networks:
-        pl.add_mesh(netw, cmap=cmap, clim=clim, show_scalar_bar=False,
+        if clim is None: clim = (0, contours.active_scalars.max())
+        netthres = netw.threshold(clim[1] / n)
+        pl.add_mesh(netthres, cmap=cmap, clim=clim, show_scalar_bar=False,
                     render_lines_as_tubes=True,line_width=5)
 
-    pl.camera_position = 'zx'
-    pl.camera.roll += 90
+    pl.camera_position = 'yz'
+    #pl.camera.roll += 90
     pl.camera.zoom(1.3)
     pl.add_title(title, font_size=12)
[…]

### 2024-10-17 scripts/pvs_flow_prod.py
evaulate on segments

@@ -15,7 +15,11 @@ from matplotlib.ticker import PercentFormatter
 from label_arteries import pointlabels
 import pandas as pd
 import yaml
+from branch_marking import color_branches
 
+m2mm = 1e3
+m2mum = 1e6
+Pa2mPa = 1e3
 
 def map_pressure_on_network(p, pnet):
     idx = map_kdtree(p.function_space().tabulate_dof_coordinates(), 
@@ -116,9 +120,15 @@ def compute_pvs_flow(csf_flow_model:  str):
         xdmf.write(ph, t=0)
     
 
-    umag = uh_mag.vector()[:] *1e6
+    segments, segids ,_ = color_branches(mesh)
+    dxs = Measure("dx", mesh, subdomain_data=segments)
+
+
+    seglengths = np.array([assemble(1*dxs(si)) for si in segids])
+    assert (seglengths > 0).all()
+    umag = np.array([assemble(m2mum*uh_mag*dxs(si))/ l for si,l in zip(segids, seglengths)])
     umagabs = abs(umag)
-    uavg = np.average(umagabs)#, weights=grid["Length"])
+    uavg = np.average(umagabs, weights=seglengths)
     umax = umagabs.max()
     umed = np.median(umagabs)
 
@@ -127,7 +137,7 @@ def compute_pvs_flow(csf_flow_model:  str):
     fig, ax = plt.subplots()
     counts, bins, containers = plt.hist(umag, density=False, bins=50, histtype="bar",
             range=(-0.4, 0.6), edgecolor='black', linewidth=0.4, 
-                    weights=np.ones(len(umag)) / len(umag))
+                    weights=seglengths / seglengths.sum())
     for bar, ca in zip(containers, sns.color_palette("Purples_r", n_colors=20) + 
                                    sns.color_palette("Greens", n_colors=30)):
         bar.set_facecolor(ca)
@@ -142,27 +152,25 @@ def compute_pvs_flow(csf_flow_model:  str):
     ax.yaxis.set_major_formatter(PercentFormatter(1))
     plt.savefig(f"{results_dir}/velocity_histo.png")
 
-
-    cell_mf = Function(FunctionSpace(mesh, "DG", 0))
-    point_mf = Function(FunctionSpace(mesh, "CG", 1))
     points = np.array([c for n,c in pointlabels])
     cellidx = map_kdtree(FunctionSpace(mesh, "DG", 0).tabulate_dof_coordinates(),
[…]

### 2024-10-17 scripts/time_dependent.py
adjust weight
Former-commit-id: 87ab9f0b9038daffe494079df86ff23b466f3e8a

@@ -39,8 +39,6 @@ def get_subdomain_dofs(V, subdomains, subd_id):
     return subd_dofs
 
 
-
-
 def run_simulation(configfile: str):
 
     config = read_config(configfile)
@@ -91,7 +89,7 @@ def run_simulation(configfile: str):
 
     artery_shape = xii.Circle(radius=artery_radii, degree=40, quad_rule='midpoint')
     vein_shape = xii.Circle(radius=vein_radii, degree=40, quad_rule='midpoint')
-    csf_par_weights = {PARID:0.01, CSFID:0.99}
+    csf_par_weights = {PARID:0.2, CSFID:0.8}
     artmarker = volmarker_to_networkmarker(vol_subdomains, artery, artery_shape,
                                          filename=results_dir + "arttagshares.pvd",
                                          weights=csf_par_weights)

### 2024-10-17 scripts/peristalticflow.py
update supply nodes

@@ -12,7 +12,7 @@ from solver import read_vtk_network
 import pvs_network_netflow as pnf
 
 # FIXME: Automate
-SUPPLY_NODES = [3492, 8115, 6743]
+SUPPLY_NODES = [8065, 7173, 4085]
 
 
 def mark_shortest_paths(mesh, G, roots, output, use_precomputed=False):

### 2024-10-21 scripts/peristalticflow.py
Add more on vasomotion - note that directionality is not given.
Former-commit-id: 71c1633fa837907a6fec92a040d33bcaf438dabe [formerly 23d5e66f1b2bb83b8c859fa149a4da05b50614bf]
Former-commit-id: 2f813bff8a68dc79bcc42648eab9f870af7f2993

@@ -490,6 +490,7 @@ def compute_pvs_flow(meshfile, output, args):
     print_stats("<Q'_i>", Q.vector()*1.e9, "mm^3/s")
     print_stats("<u'_i>", u.vector(), "m/s")
     print_stats("<u'_i>", u.vector()*1.e3, "mm/s")
+    print_stats("<u'_i>", u.vector()*1.e6, "mum/s")
 
     fluxfile = os.path.join(output, "pvs_Q.xdmf")
     with df.XDMFFile(mesh.mpi_comm(), fluxfile) as xdmf:
@@ -538,8 +539,8 @@ if __name__ == '__main__':
     # $ mamba activate pvs_transport_env
     # $ pip install git+https://github.com/MiroK/fenics_ii.git@ufl_legacy
     # 
-    # $ python3 peristalticflow.py --frequency 1.0 --wavelength 2000.0 --amplitude 0.01 --beta 3.0 --recompute # Cardiac
-    # $ python3 peristalticflow.py --frequency 0.1 --wavelength 80.0 --amplitude 0.1 --beta 3.0 
+    # $ python3 peristalticflow.py --frequency 1.0 --wavelength 2.0 --amplitude 0.01 --beta 2.0 --recompute # Cardiac ~1 Hz, 2m/s pulse wave speed
+    # $ python3 peristalticflow.py --frequency 0.1 --wavelength 0.02 --amplitude 0.1 --beta 2.0 # 
     #
     # --recompute is only needed upon first run)
     
@@ -549,7 +550,7 @@ if __name__ == '__main__':
     parser.add_argument('--recompute', action="store_true")
     parser.add_argument('--run-tests', action="store_true", help="Just run basic test.")
     parser.add_argument('--frequency', action="store", default=1.0, help="Vascular wave frequency (Hz)", type=float)
-    parser.add_argument('--wavelength', action="store", default=2000.0, help="Vascular wave wavelength (m)", type=float)
+    parser.add_argument('--wavelength', action="store", default=2.0, help="Vascular wave wavelength (m)", type=float)
     parser.add_argument('--amplitude', action="store", default=0.01, help="Vascular wave relative amplitude (relative to (inner) vascular radius)", type=float)
     parser.add_argument('--beta', action="store", default=2.0, help="Ratio outer-to-inner vessel radio (PVS width + 1)", type=float)
     parser.add_argument('--tag', action="store", default="tmp", help="Tag")

### 2024-11-01 scripts/generate_plot.py
update plots

@@ -2,8 +2,8 @@ import pyvista as pv
 import numpy as np
 import os
 import typer
-from typing import Tuple
-from plotting_utils import get_result, time_str, clip_plot, detail_plot, isosurf_plot
+from typing import List
+from plotting_utils import get_result, time_str, clip_plot, timesurf_plot, isosurf_plot
 import yaml
 pv.global_theme.allow_empty_mesh = True
 
@@ -13,16 +13,18 @@ LVID = 3
 V34ID = 4
 CSFNOFLOWID = 5
 
-def plot_model(modelname: str, t:int, type:str, cmax:float=None, filename:str=None):
+def plot_model(modelname: str, times:List[int], currenttime:int, type:str, cmax:float=None, filename:str=None):
     plotdir = f"plots/{modelname}"
     os.makedirs(plotdir, exist_ok=True)
 
-    sas = get_result(modelname, "sas", t)
-    art = get_result(modelname, "artery", t)
-    ven = get_result(modelname, "vein", t)
+    sas = get_result(modelname, "sas", times)
+    art = get_result(modelname, "artery", times)
+    ven = get_result(modelname, "vein", times)
 
     clim = (0, cmax) if cmax is not None else None
     if type=="overview":
+        assert len(times) == 1
+        t = times[0]
         title = f"time: {time_str(t)} h"
         csf = sas.extract_cells(np.isin(sas["label"], [CSFID, LVID, V34ID, CSFNOFLOWID]))
         par = sas.extract_cells(sas["label"]==PARID)
@@ -31,14 +33,16 @@ def plot_model(modelname: str, t:int, type:str, cmax:float=None, filename:str=No
               cbar_title="concentration (mmol/l)")
     
     if type=="isosurf":
+        assert len(times) == 1
+        t = times[0]
         title = f"time: {time_str(t)} h"
         csf = sas.extract_cells(np.isin(sas["label"], [CSFID, LVID, V34ID, CSFNOFLOWID]))
         par = sas.extract_cells(sas["label"]==PARID)
-        #par[f"c_{t}"] *= 0.2
+        par[f"c_{t}"] *= 0.2
         csf.merge(par, inplace=True, merge_points=False)
         #sas[f"c_{t}"] *= (1 - (0.8*sas["label"]==PARID))
-        return isosurf_plot(sas, [art], filename, title, clim=clim, cmap="tempo",
+        return isosurf_plot(csf, [art], filename, title, clim=clim, cmap="tempo",
[…]

### 2024-11-01 scripts/compare_models.py
update plots
Former-commit-id: b4befa3a5a8582d3f3f586ad22ba20b6c2e7f810 [formerly 1430a4fd8cb911768f5528c7bd110f3a46157745]
Former-commit-id: 785196394576f278f44440dc79e4e94526ccfa2c

@@ -3,18 +3,23 @@ from mpl_toolkits.axes_grid1 import ImageGrid
 import numpy as np
 import typer
 from typing import List
-from plotting_utils import time_str, read_config, compute_ranges, compute_diff_ranges
-from generate_plot import plot_model
-from create_diff_plot import plot_model_diff
+from plotting_utils import (time_str, read_config, compute_ranges,
+                    get_result, clip_plot, isosurf_plot, timesurf_plot,
+                    compute_diff_ranges)
+import pyvista as pv
+from generate_synthseg_mesh import CSFID, CSFNOFLOWID, PARID, LVID, V34ID
+import matplotlib as mpl
 
 crop_dict = {"overview":lambda img : img[100:-10,140:-140],
              "overview_bottom":lambda img : img[100:-100,140:-140],
              "detail":lambda img : img,
             "isosurf":lambda img : img[100:-10,140:-140],
+            "timesurf":lambda img : img[100:-10,140:-140],
             "detail_bottom":lambda img : img,
             "isosurf_bottom":lambda img : img[100:-100,140:-140],
+            "timesurf_bottom":lambda img : img[100:-100,140:-140],
              }
-padding = {"overview":0, "detail":0.05, "isosurf":0}
+padding = {"overview":0, "detail":0.05, "isosurf":0, "timesurf":0}
 percentile = 95
 
 def compare_models(modela:str, modelb:str, type:str, cmax:float, diffmax:float, times:List[int]):
@@ -30,21 +35,62 @@ def compare_models(modela:str, modelb:str, type:str, cmax:float, diffmax:float, 
         rmin = np.min([r[0] for r in rangediff.values()])
         rmax = np.max([r[1] for r in rangediff.values()])
         diffmax = max(abs(rmin), abs(rmax))
-    for i, t in enumerate(times):
-        for j,m in enumerate([modela, modelb,"diff"]):
-            if m=="diff":
-                img = plot_model_diff(modela, modelb, t, type, cmax=diffmax)
-            else:
-                img = plot_model(m, t, type, cmax=cmax)
-            ax = grid.axes_row[i][j]
-            ax.axis('off')
-            if i + 1==len(times):
-                ax.imshow(crop_dict[type](img))
-            else:
-                ax.imshow(crop_dict[type + "_bottom"](img))
 
-        ax = grid.axes_row[i][0]
[…]

### 2024-11-01 scripts/compare_models_horizontal.py

@@ -3,44 +3,85 @@ from mpl_toolkits.axes_grid1 import ImageGrid
 import numpy as np
 import typer
 from typing import List
-from plotting_utils import time_str, read_config, compute_ranges, compute_diff_ranges
-from generate_plot import plot_model
-import matplotlib as mpl
+from plotting_utils import (time_str, read_config, compute_ranges,
+                    get_result, clip_plot, isosurf_plot, timesurf_plot)
 import cmocean
+from generate_synthseg_mesh import CSFID, CSFNOFLOWID, PARID, LVID, V34ID
+import pyvista as pv
+from cmap import Colormap
+import nibabel
+from extract_vessels import np2pv
+import matplotlib as mpl
 
 crop_dict = {"overview":lambda img : img[80:-100,200:-200],
              "overview_bottom":lambda img : img[100:-100,140:-140],
              "detail":lambda img : img,
             "isosurf":lambda img : img[100:-10,140:-140],
+            "timesurf":lambda img : img[100:-10,140:-140],
             "detail_bottom":lambda img : img,
             "isosurf_bottom":lambda img : img[100:-100,140:-140],
+            "timesurf_bottom":lambda img : img[100:-100,140:-140],
              }
-padding = {"overview":0, "detail":0.05, "isosurf":0}
 percentile = 95
 
 def compare_models(modela:str, modelb:str, type:str, cmax:float, times:List[int]):
 
     fig = plt.figure(figsize=(len(times)*3, len(times)*3), frameon=True)
-    grid = ImageGrid(fig, 111, nrows_ncols=(2, len(times)), axes_pad=padding[type],
+    grid = ImageGrid(fig, 111, nrows_ncols=(2, len(times)),
                      cbar_location="right",
                      cbar_mode="single",
                      cbar_size="4%",
                      cbar_pad=0.1,)
+    
     if not cmax:
         rangea = compute_ranges(modela, times, percentile,)
         rangeb = compute_ranges(modelb, times, percentile)
         cmax = np.max([r[1] for r in rangea.values()] + [r[1] for r in rangeb.values()])
-
-    for i, t in enumerate(times):
-        for j,m in enumerate([modela, modelb]):
-            img = plot_model(m, t, type, cmax=cmax)
+    for j,m in enumerate([modela, modelb]):
+        conf = read_config(f"configfiles/{m}.yml")
[…]

### 2024-11-01 scripts/compute_dispersion_field.py
add R histo
Former-commit-id: 008a86642886908cfc73ceab84ec487f153f2c76 [formerly fe189b76ed43ba82b562c40176ea35444d28ba6a]
Former-commit-id: dda9cff24e35c1e76bed123e2c093be9b3ddd6b2

@@ -3,6 +3,44 @@ import numpy as np
 import typer
 from pathlib import Path
 from vtk.util.numpy_support import numpy_to_vtk
+from plot_subdomains import get_camera
+from plotting_utils import set_plotting_defaults
+from generate_synthseg_mesh import CSFID, LVID, V34ID
+import matplotlib.pyplot as plt
+import seaborn as sns
+from matplotlib.ticker import PercentFormatter
+
+def R_histo(sm, R, filename):
+    set_plotting_defaults()
+    DG0 = sm.function_space()
+    cellvol = assemble(TestFunction(DG0)*dx)
+    ids = [CSFID, LVID, V34ID]
+    indices = [sm.vector()[:]==i for i in ids]  
+    fig, axes = plt.subplots(ncols=3, figsize=(7,3))
+    nbins = 20
+    colors = sns.color_palette("crest", n_colors=3)
+    for ax, name, idx, col in zip(axes, ["SAS", "LV", "V3 & V4"], indices, colors):
+        Rvals = R.vector()[idx]
+        xlims = (0, np.percentile(Rvals, 80))
+        bins = np.logspace(start=np.log10(1e-1), stop=np.log10(200), num=nbins)
+        n, bins,_ = ax.hist(Rvals, bins=bins, weights=cellvol[idx] / cellvol[idx].sum(),
+                color=col, alpha=0.7, label=name)
+        ax.set_xscale("log")
+        mean = np.average(Rvals, weights=cellvol[idx])
+        ax.axvline(mean, 
+                   color="black",ymax=0.6, label=f"{mean:.1f} \n (mean)")
+        #ax.axvline(np.max(Rvals), ls="dotted",
+        #           color="black",ymax=0.6, label=f"{max(Rvals):.1f}")
+        ax.set_xlabel(f"R")
+        ax.set_ylabel("frequency")
+        ax.set_xticks([0.1, 1, 10, 100])
+        plt.tight_layout()
+        #ax.set_xlim(xlims)
+        ax.set_ylim((0, max(n)*1.5))
+        ax.legend(frameon=False, loc="lower left", alignment="right",
+                  borderaxespad=0, handlelength=1.3, bbox_to_anchor=[0.2, 0.7])
+        ax.yaxis.set_major_formatter(PercentFormatter(1, decimals=0))
+    plt.savefig(filename, bbox_inches='tight')
 
 
[…]

### 2024-11-01 scripts/evaluate_pvs_flow.py
minor changes
Former-commit-id: eb0658fe530dcbce1fdb496b3e3bd82358e0c59b [formerly 74f5e3df2984e7593a8d51c45a90f926cc12889e]
Former-commit-id: 9f913219dbf78c0c2ed3d7fa2e13a78359723f9a

@@ -39,8 +39,9 @@ def compute_pvs_flow(pvs_flow_file):
         p, uh_mag, artery_radii = map(Function, [p_space, umag_space, r_space])
         for u,n in zip([p, uh_mag, artery_radii], ["p", "umag", "radii"]):
             f.read(u, n)
-
-    plot_dir = f"plots/{'/'.join(pvs_flow_file.split('/')[1:-1])}"
+    modelstr = pvs_flow_file.split('/')[1:-1]
+    model = modelstr[-1]
+    plot_dir = f"plots/{'/'.join(modelstr)}"
     os.makedirs(plot_dir, exist_ok=True)
     segments, segids ,_ = color_branches(mesh)
     dxs = Measure("dx", mesh, subdomain_data=segments)
@@ -60,7 +61,7 @@ def compute_pvs_flow(pvs_flow_file):
         uavg = np.average(uvals, weights=lengths)
         umax = abs(uvals).max()
         umed = np.median(abs(uvals))
-        range = np.round(minmax(uvals, percentile=95), 1)
+        range = np.round(minmax([uvals], percentile=99.9), 1)
         nbins = 50
         fig, ax = plt.subplots()
         counts, bins, containers = plt.hist(uvals, density=False, bins=nbins, histtype="bar",
@@ -74,15 +75,15 @@ def compute_pvs_flow(pvs_flow_file):
                                                        n_colors=int(pos_share * nbins))):
             bar.set_facecolor(ca)
         plt.axvline(uavg, color="black", linestyle=":", label=f"avg: {uavg:.2f} µm/s")
-        plt.axvline(umed, color="black", linestyle="-.", label=f"median: {umed:.2f} µm/s")
+        #plt.axvline(umed, color="black", linestyle="-.", label=f"median: {umed:.2f} µm/s")
         plt.axvline(umax, color="black", linestyle="--", label=f"max: {umax:.2f} µm/s")
         plt.xlabel("PVS flow velocity (µm/s)")
         plt.ylabel("frequency")
         plt.tight_layout()
-        #plt.xlim(range)
+        plt.xlim(range)
         plt.legend(frameon=False)
         ax.yaxis.set_major_formatter(PercentFormatter(1))
-        plt.savefig(f"{plot_dir}/velocity_histo_{variant}.png")
[…]

### 2024-11-01 scripts/extract_vessels.py
add __main__
Former-commit-id: 847062e29d24d402b5c40e70c9cac777c805aeee

@@ -179,44 +179,46 @@ def smooth_coords(netw):
 scale = 1.5
 const = 3.0
 
-os.makedirs("mesh/networks", exist_ok=True)
-os.makedirs("plots", exist_ok=True)
 
-files = ["data/pcbi.1007073.s007.nii.gz", "data/pcbi.1007073.s008.nii.gz"]
-names = ["arteries", "venes"]
-nroots = [3, 50]
+if __name__=="__main__":
+    os.makedirs("mesh/networks", exist_ok=True)
+    os.makedirs("plots", exist_ok=True)
 
-for file, name, nr in zip(files, names, nroots):
-    data = nibabel.load(file)
-    resolution = data.header["pixdim"][1:4]
-    print(f"resolution: {resolution}")
-    img = data.get_fdata().astype(int)
-    skel = skeletonize(img, resolution)
-    if name=="arteries":
-        skel = skel.crop(Bbox([0, 0, 37], img.shape))
-    G = as_networkx(skel, nroots=nr)
-    orig_netw = nx_to_pv(G)
-    orig_netw["radius"] *= mm2m
-    orig_netw.scale(mm2m, inplace=True)
-    remove_duplicate_cells(orig_netw)
-    orig_netw.save(f"mesh/networks/{name}.vtk")
-    splines = get_splines(orig_netw, point_ratio=3)
-    lines = [pv.lines_from_points(spl.points) for spl in splines]
-    for l, spl in zip(lines, splines):
-        l["radius"] = spl["radius"]
-        l["root"] = spl["root"]
-    mean_radius = [l["radius"].mean() for l in lines]
-    order_by_radius = np.flip(np.argsort(mean_radius))
-    smooth_netw = pv.merge([lines[i] for i in order_by_radius],
-                            merge_points=True).cast_to_unstructured_grid()
-    for i in range(3):
-        smooth_radius(smooth_netw)
-    for i in range(3):
-        smooth_coords(smooth_netw)
-    smooth_netw.save(f"mesh/networks/{name}_smooth.vtk")
-    netw_tubes = get_tubes(smooth_netw)
-    netw_tubes.save(f"mesh/networks/{name}_tubes.vtk")
-    print(smooth_netw.n_points)
-    #netw_tubes.plot(off_screen=True, screenshot=f"plots/{name}_network.png", zoom=1.6)
[…]

### 2024-11-01 scripts/label_arteries.py
add new labels
Former-commit-id: 24ae015614ad71394f6b66ad08aca827fa702199 [formerly 24ae015614ad71394f6b66ad08aca827fa702199 [formerly 1f81cd295477754ecc54a375ab75e8d7daacff94]]
Former-commit-id: 39f516d250f2bac1f6f51ae8c46b986ec994f71b
Former-commit-id: 757ccdc8c86f5481f674386173b0724c8cbdf641

@@ -10,7 +10,12 @@ pointlabels = [("BA",(0.084971, 0.104599, 0.0424848)),
                ("MCA2-L",(0.0360319, 0.114267, 0.0717319)),
                ("PCA-R",(0.100641, 0.0843866, 0.0609018)),
                ("PCA-L",(0.0686598, 0.0831056, 0.0568451)),
-               ("ACA", (0.0795614, 0.132763, 0.0892396)),
+               ("ACA-A1-R", (0.0864598, 0.112447, 0.0623659)),
+               ("ACA-A1-L", (0.0746942, 0.109032, 0.0643898)),
+               ("ACA-A2", (0.0805165,0.115209,0.0781812)),
+               ("ACA-A3", (0.0795565, 0.133436, 0.0905888)),
+               ("ACA-A4", (0.0815724, 0.111484, 0.107046)),
+               ("ACA-A5", (0.0815178, 0.0984657, 0.167766)),
                ("PER-R", (0.123073, 0.101526, 0.135256)),
                ("PER-L", (0.0401122, 0.0939796, 0.13324))
             ]

### 2024-11-01 scripts/mean_concentrations.py
add timing and nearby values
Former-commit-id: f49f1be724c659d1685ffb1bb839390804281d2a

@@ -22,11 +22,13 @@ LVID = 3
 V34ID = 4
 CSFNOFLOWID = 5
 
-def compare_concentrations(modelname:str, times:List[int]):
+def compare_concentrations(modelname:str):
+    config = read_config(f"configfiles/{modelname}.yml")
+    dt , T= config["dt"], config["T"]
+    times = np.arange(0, T + dt, 3*dt*config["output_frequency"])
     sas_conc, subd_marker = get_result_fenics(modelname, "sas", times)
     art_conc, art_radii = get_result_fenics(modelname, "artery", times)
     ven_conc, ven_radii = get_result_fenics(modelname, "vein", times)
-    config = read_config(f"configfiles/{modelname}.yml")
     pvs_ratio_artery = config["pvs_ratio_artery"]
     pvs_ratio_vein = config["pvs_ratio_venes"]
     subd_marker.array()[np.isin(subd_marker.array(),[LVID, V34ID, CSFNOFLOWID])] = CSFID
@@ -47,9 +49,7 @@ def compare_concentrations(modelname:str, times:List[int]):
     art_tot = np.array([assemble(area_artery*c*dxa) for c in art_conc])
     ven_tot = np.array([assemble(area_vein*c*dxv) for c in ven_conc])
     tot = par_tot + csf_tot + art_tot + ven_tot
-    times = np.array(times)
-    dt , T= config["dt"], config["T"]
-    alltimes = np.arange(0, times[-1], dt)
+
     #inflow = np.cumsum([k*max(0.0, t0 - t) for t in alltimes]) * dt
     set_plotting_defaults()
     sns.set_palette("magma_r", n_colors=4)
@@ -106,9 +106,13 @@ def compare_concentrations(modelname:str, times:List[int]):
     priority = InterfaceResolution(subdomains=subd_marker,
                                        resolve_conflicts={(CSFID, PARID): 4})
     pvs_shape = Circle(radius=pvs_radii, degree=40, quad_rule='midpoint')
+    proximity_dist = 2e-3
+    nearby_shape = Circle(radius=project(pvs_radii + proximity_dist, DG0a), degree=40, quad_rule='midpoint')
     c_averages = [ii_project(Average(c, artery, pvs_shape, 
-                                     normalize=True, resolve_interfaces=priority), V) for c in sas_conc]
-
+                                     normalize=True), V) for c in sas_conc]
[…]

### 2024-11-01 scripts/plot_boundaries.py
update plots
Former-commit-id: fd682abad4eecc4a93fec0e0144fcd585d46ddd2 [formerly fbf90e44530728f0007cc425baebdb52923cf94a]
Former-commit-id: feec8144020462dffa701b4f3a3a435f4d9c6a7a

@@ -5,6 +5,7 @@ LOWER_SKULL_ID, UPPER_SKULL_ID, SPINAL_CORD_ID)
 import seaborn as sns
 import numpy as np
 from cmap import Colormap
+from plot_subdomains import get_camera
 
 grid = pv.read_meshio("mesh/standard/standard_facets.xdmf")
 dirname = "plots/meshplots/"
@@ -37,15 +38,7 @@ for i, (n, bid) in enumerate(bds.items()):
     meshes.append(m)
 
 
-pl = pv.Plotter(off_screen=True, window_size=(1600, 1600))
-pl.add_mesh(grid.clip())
-pl.camera_position = 'yz'
-pl.camera.roll += 0
-pl.camera.azimuth += 30
-#pl.camera.elevation += 10
-pl.camera.zoom(1.2)
-camera = pl.camera.copy()
-
+camera = get_camera(grid)
 
 pl = pv.Plotter(off_screen=True, window_size=(1600, 1600))
 pl.add_mesh(pv.merge(meshes), scalars="l", show_scalar_bar=True, 

### 2024-11-01 scripts/plotting_utils.py

@@ -3,6 +3,8 @@ import yaml
 import collections
 from typing import List
 from IPython import embed
+from cmap import Colormap
+from tqdm import tqdm
 
 def set_plotting_defaults():
     import seaborn as sns
@@ -53,8 +55,9 @@ def get_result_fenics(modelname, domain, times):
         mesh, radii, _ = read_vtk_network("mesh/networks/venes_smooth.vtk", rescale_mm2m=False)
         V = FunctionSpace(mesh, "CG", 1)
     results = []
+    print(f"reading results from {modelname} at")
     with XDMFFile(filename) as f:
-        for t in times:
+        for t in tqdm(times):
             i = np.where(alltimes==t)[0][0]
             c = Function(V)
             c.rename("c","c")
@@ -90,8 +93,6 @@ def compute_diff_ranges(modela:str, modelb:str, times: List[int],percentile=95):
 
     return ranges
 
-
-
 def time_str(sec):
     m, s = divmod(sec, 60)
     h, m = divmod(m, 60)
@@ -105,6 +106,8 @@ def read_config(configfile):
 
 def clip_plot(csf, par, networks, filename, title, clim, cmap, cbar_title):
     import pyvista as pv
+    from extract_vessels import get_tubes
+
     csf_clipped = csf.clip(normal="y")
     par_clipped = par.clip(normal="y")
     pl = pv.Plotter(off_screen=True)
@@ -115,7 +118,7 @@ def clip_plot(csf, par, networks, filename, title, clim, cmap, cbar_title):
                                     label_font_size=32))
     pl.add_mesh(par_clipped, cmap=cmap, clim=clim, show_scalar_bar=False)
     for netw in networks:
-        pl.add_mesh(netw, cmap=cmap, clim=clim, show_scalar_bar=False,
+        pl.add_mesh(get_tubes(netw), cmap=cmap, clim=clim, show_scalar_bar=False,
                     render_lines_as_tubes=True,line_width=5)
 
     pl.camera_position = 'zx'
@@ -124,8 +127,9 @@ def clip_plot(csf, par, networks, filename, title, clim, cmap, cbar_title):
     pl.add_title(title, font_size=12)
     return pl.screenshot(filename, transparent_background=True, return_img=True)
 
-def isosurf_plot(sas, networks, filename, title, clim, cmap, cbar_title):
+def isosurf_plot(sas, networks, bg, filename, title, clim, cmap, cbar_title):
     import pyvista as pv
[…]

### 2024-11-01 scripts/plot_csf_flow.py
update plots
Former-commit-id: fbf90e44530728f0007cc425baebdb52923cf94a

@@ -12,10 +12,12 @@ from generate_synthseg_mesh import CSFID, LVID, V34ID
 import matplotlib.pyplot as plt
 from matplotlib.ticker import PercentFormatter
 import seaborn as sns
+from plot_subdomains import get_camera
+from vtk.util.numpy_support import numpy_to_vtk
 
 m2mum = 1e6
 
-def velocity_histo(sm, u, filename):
+def velocity_histo(sm, u, filename, vscale, vunit):
     set_plotting_defaults()
     DG0 = sm.function_space()
     umag = project(sqrt(inner(u,u)), DG0,
@@ -27,33 +29,38 @@ def velocity_histo(sm, u, filename):
     nbins = 20
     colors = sns.color_palette("crest", n_colors=3)
     for ax, name, idx, col in zip(axes, ["SAS", "LV", "V3 & V4"], indices, colors):
-        uvals = umag.vector()[idx] * m2mum
+        uvals = umag.vector()[idx] * vscale
         xlims = (0, np.percentile(uvals, 92))
         ax.hist(uvals, bins=nbins, weights=cellvol[idx] / cellvol[idx].sum(),
                  range=xlims,
                   color=col, alpha=0.7, label=name)
         mean = np.average(uvals, weights=cellvol[idx])
         ax.axvline(mean, 
-                   color="black",ymax=0.6, label=f"{mean:.1f} µm/s \n (mean)")
+                   color="black",ymax=0.6, label=f"{mean:.1f} {vunit} \n (mean)")
         #ax.axvline(np.max(uvals), ls="dotted",
         #           color="black",ymax=0.6, label=f"{max*1e-3:.1f} mm/s")
-        ax.set_xlabel("velocity (µm/s)")
+        ax.set_xlabel(f"velocity ({vunit})")
         ax.set_ylabel("frequency")
         plt.tight_layout()
         ax.set_xlim(xlims)
         ax.legend(frameon=False, loc="lower left", alignment="right",
                   borderaxespad=0, handlelength=1.3, bbox_to_anchor=[0.2, 0.7])
         ax.yaxis.set_major_formatter(PercentFormatter(1, decimals=0))
-    plt.savefig(filename)
+    plt.savefig(filename, bbox_inches='tight')
 
 def from_k3d(colorlist):
     cm = np.array(colorlist).reshape(-1, 4)[:, 1:]
     return matplotlib.colors.LinearSegmentedColormap.from_list("", cm)
 
[…]

### 2024-11-01 scripts/plot_subdomains.py

@@ -4,40 +4,45 @@ from generate_synthseg_mesh import PARID, CSFID, LVID,V34ID
 import seaborn as sns
 import numpy as np
 
-grid = pv.read_meshio("mesh/standard/standard.xdmf")
-dirname = "plots/meshplots/"
-os.makedirs(dirname, exist_ok=True)
+def get_camera(grid):
+    pl = pv.Plotter(off_screen=True, window_size=(1600, 1600))
+    pl.add_mesh(grid)
+    pl.camera_position = 'yz'
+    pl.camera.roll += 0
+    pl.camera.azimuth += 30
+    #pl.camera.elevation += 10
+    pl.camera.zoom(1.4)
+    return pl.camera.copy()
 
-PAR = grid.extract_cells(grid["label"]==PARID)
-CSF = grid.extract_cells(grid["label"]==CSFID)
-V34 = grid.extract_cells(grid["label"]==V34ID)
-LV = grid.extract_cells(grid["label"]==LVID)
 
-CSF["l"] = np.full(CSF["label"].shape, " SAS", dtype='<U10')
-V34["l"] = np.full(V34["label"].shape, "V3 & V4", dtype='<U10')
-LV["l"] = np.full(LV["label"].shape, "LV", dtype='<U10')
+if __name__=="__main__":
+    grid = pv.read_meshio("mesh/standard/standard.xdmf")
+    dirname = "plots/meshplots/"
+    os.makedirs(dirname, exist_ok=True)
 
-comb = pv.merge([CSF.clip(), V34, LV])
+    PAR = grid.extract_cells(grid["label"]==PARID)
+    CSF = grid.extract_cells(grid["label"]==CSFID)
+    V34 = grid.extract_cells(grid["label"]==V34ID)
+    LV = grid.extract_cells(grid["label"]==LVID)
 
-bar_args=dict(title="", vertical=False, height=0.08, width=0.6, position_x=0.2,
-                        position_y=-0.05, title_font_size=52,
-                        bold=False, font_family="times",
-                        label_font_size=44)
+    CSF["l"] = np.full(CSF["label"].shape, " SAS", dtype='<U10')
+    V34["l"] = np.full(V34["label"].shape, "V3 & V4", dtype='<U10')
+    LV["l"] = np.full(LV["label"].shape, "LV", dtype='<U10')
 
-pl = pv.Plotter(off_screen=True, window_size=(1600, 1600))
-pl.add_mesh(CSF.clip())
-pl.camera_position = 'yz'
-pl.camera.roll += 0
[…]

### 2024-11-01 scripts/time_dependent.py
add root permeabiltiy increase
Former-commit-id: e78034a5231a9ac962a15c1a691b59faac911e0b

@@ -7,7 +7,7 @@ import numpy as np
 import typer
 from pathlib import Path
 from plotting_utils import read_config
-from test_map_on_global_coords_shift import map_dg_on_global
+from test_map_on_global_coords_shift import map_dg_on_global, map_kdtree
 from IPython import embed
 import yaml
 
@@ -118,8 +118,10 @@ def run_simulation(configfile: str):
                                         PARID: parenchyma_diffusion # parenchyma
                                         })
     art_xi_root = Function(FunctionSpace(artery, "CG", 1))
-    art_xi_root.vector()[:] = np.where(artery_roots.array()[:] > 0, root_xi_factor, 1)
-    xi_a = pcws_constant(artmarker, art_xi_dict)*art_xi_root
+    idx = map_kdtree(artery.coordinates(), FunctionSpace(artery, "CG", 1).tabulate_dof_coordinates())
+    art_xi_root.vector()[:] = np.where(artery_roots.array()[:] > 0, root_xi_factor, 1)[idx]
+    xi_a = pcws_constant(artmarker, art_xi_dict)
+    xi_a.vector()[:] *= interpolate(art_xi_root, xi_a.function_space()).vector()[:]
     xi_v = pcws_constant(veinmarker, ven_xi_dict)
 
     Da = Constant(arterial_pvs_diffusion) 
@@ -328,6 +330,7 @@ def run_simulation(configfile: str):
     u_i.rename("c", "c")
     pa_i.rename("c", "c")
     pv_i.rename("c", "c")
+    xi_a.rename("xi_a", "xi_a")
 
     xdmfsas = XDMFFile(results_dir + f'{modelname}_sas.xdmf') 
     xdmfart = XDMFFile(results_dir + f'{modelname}_artery.xdmf') 
@@ -350,7 +353,7 @@ def run_simulation(configfile: str):
 
     #write((u_i, pa_i, pv_i), pvdfiles, 0.0, hdffile=hdffile)
     write((artmarker, veinmarker), pvdfiles, 0.0)
-    #write((xi_a, xi_v), pvdfiles, 0.0)
+    write((xi_a, xi_v), pvdfiles, 0.0)
     #write([phi], [pvdsas], 0.0)
 
     wh = xii.ii_Function(W)

### 2024-11-01 scripts/uhat_prod.py
remove old file
Former-commit-id: b420d7f58abd24fb27759d8e3ebd7754945d8cd4 [formerly cdb0c9ebefec181b49c6c5e2c65334d8744dcbf1]
Former-commit-id: be2438ebc6b39c98fb7084c2f5084a3a55fb1d9a

@@ -1,77 +1,0 @@
-from dolfin import *
-from solver import * 
-import os
-import numpy as np
-from xii import *
-from petsc4py import PETSc
-import sympy as sp
-import typer
-import yaml
-
-def compute_avg_flow(csf_flow_model:  str):
-
-    mesh = Mesh()
-    filename = f"results/csf_flow/{csf_flow_model}/flow.hdf"
-    with HDF5File(MPI.comm_world, filename,'r') as f:
-        f.read(mesh, "mesh", False)
-        DG = VectorFunctionSpace(mesh, "DG", 2)
-        velocity_sas = Function(DG)
-        f.read(velocity_sas, "velocity")
-
-
-    #cell = sas.ufl_cell()
-    #VP1_elem = VectorElement('Lagrange', cell, 1)
-    #VP1 = FunctionSpace(sas, VP1_elem)
-    #velocity_sas = Function(VP1) 
-    #velocity_sas = interpolate(velocity_sas_ , VP1)
-    # read the arterial network and mesh 
-    pvs_ratio_artery  = 2.0 
-    artery, artery_radii, artery_roots = read_vtk_network("mesh/networks/arteries_smooth.vtk", rescale_mm2m=False)
-    artery_radii = as_Pk_function(artery_radii, k=3)
-
-    # artery_radii.set_allow_extrapolation(True)
-    pvs_radii = Function(artery_radii.function_space())
-    pvs_radii.vector().set_local(pvs_ratio_artery*artery_radii.vector().get_local())
-    pvs_radii.set_allow_extrapolation(True)
-    artery_radii.set_allow_extrapolation(True)
-
-    # define velocity average using Fenics_ii average functionality 
-    disk_pvs  = Disk(radius    = pvs_radii, degree =11)
-    disk_artery  = Disk(radius = artery_radii, degree =11)
-    velocity_sas_averaged = 1 / (pvs_radii**2 - artery_radii**2) * (
-        pvs_radii **2 * Average(velocity_sas, artery, disk_pvs) -
-        artery_radii **2 * Average(velocity_sas, artery, disk_artery))
-
-    Qa = FunctionSpace(artery, 'CG', 1)
-    tau = TangentCurve(artery)
-
-    #  project velocity_sas_averaged \cdot tangent vector onto the function space over the network 
-
[…]

### 2024-11-11 paper/cut.tex
Just commit changes
Former-commit-id: b0716566f8c908e3bf54eb6f60b870a3e40dabcb
More shuffling in Methods. Short snippets targeting general audience in Methods; complete version targeting math readers in the Supplementary Methods.

@@
+We consider non-zero convection in the CSF space $u_{\rm CSF} \not = 0$ in \eqref{eq:multi_transport_3d} resulting from steady state CSF production.
+Recall that $\Omega_{\mathrm{CSF}}$ represents the CSF space and that $\Gamma_{\mathrm{skull^{u/l}}}$ and $\Gamma_{\mathrm{pia}}$, $\Gamma_{\mathrm{LV}}$ represent the boundaries facing the upper and lower parts of the dura, the pia and the lateral ventricles, respectively.
+Assuming CSF production in the choroid plexus located at the lateral ventricle wall, we solve the  steady state Stokes equations for the velocity and the pressure $(\bm u_{\mathrm{CSF}}, p_{\mathrm{CSF}})$ in the CSF domain:
+\begin{subequations}
+    \begin{alignat}{2}
+ - \mu \Delta \bm u_{\mathrm{CSF}} + \nabla p_{\rm{CSF}} & =  0 \quad && \mathrm{in} \,\,  \Omega_{\rm CSF}, \label{eq:momnetum_equation}  \\ 
+ \nabla \cdot  \bm u_{\mathrm{CSF}} & = 0 \quad && \mathrm{in} \,\,   \Omega_{\rm CSF}, \label{eq:divergence_equation}  \\ 
+\mu \nabla \bm u_{\mathrm{CSF}} \cdot \bm{n} -  p \bm n  &  = -R_0 ( \bm u \cdot \bm n ) \bm n\,\,   && \mathrm{on}  \label{eq:efflux_condition} \,\, \Gamma_{\mathrm{skull^u}}, \\ 
+\bm u_{\mathrm{CSF}} & = 0 && \mathrm{on} \,\, \Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull^l}} \Gamma_{\rm{SSAS}}  \\
+\bm u_{\mathrm{CSF}} \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm \tau = 0 \quad && \mathrm{on} \,\, \Gamma_{\rm{LV}} .  
+\end{alignat}
+\end{subequations}
+
+Here, $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the CSF viscosity, and $\bm u_{\rm in}$ is the steady CSF flow across the lateral ventricle wall, see \Cref{tab:pvs:parameters}.
+Condition \eqref{eq:efflux_condition} models an efflux site with positive resistance $R_0 \geq 0$.
+The Stokes system is solved with the finite element software $FEniCS$, stored and then subsequently read for all the relevant solute transport simulations.
+See Figure~\ref{fig:csf_flow_cardiac} for a streamline visualization of the obtained velocity profile and \Cref{sec:details_numerical_method} for details on the numerical algorithm employed.
+
+%% \begin{figure}
+%%      \centering
+%%      \begin{subfigure}[b]{0.45\textwidth}
+%%          \centering
+%%          \includegraphics[width=\textwidth]{figures/Brain-PVS-callouts.pdf}
+%%          \label{fig:intracranial_domains}
+%%      \end{subfigure}
+%%      \hfill
+%%      \begin{subfigure}[b]{0.3\textwidth}
+%%          \centering
+%%          \includegraphics[width=\textwidth]{figures/geometry_concept.png}
+%%          \caption{The arterial (red) and venous (blue) networks (top left), the parenchyma and CSF space (top right), combined (bottom row)}
+%%          \label{fig:three sin x}
+%%      \end{subfigure}
+%%      \hfill
+%%      \begin{subfigure}[b]{0.24\textwidth}
+%%          \centering
+%%          \includegraphics[trim={4cm 0 4cm 0},clip,width=\textwidth]{figures/labeled_arteries.png}
+%%          \caption{The arterial network with the main cerebral arteries: internal carotid artery (ICA), middle cerebral artery (MCA), anterior cerebral artery (ACA) and posterior cerebral artery (PCA)}
+%%          \label{fig:five over x}
+%%      \end{subfigure}
+%%      %\caption{@Marius: Starts adding parts of this.
+\mer{MER: Concept figure illustrating the compartments and story line.
+(A) Concept sketch of the SAS, PVS and the brain; (B) Visualization of geometry domains (brain, sas and vasculature; think about colors and associations; please don't make blood blue and water red e.g.) ; (C) Zoom-in/detail on vasculature, which are which arteries (ACA, MCA, PCA, CoW)  and veins (XXX); (D) Zoom-in 2 on vascular position versus SAS and brain; (E) Zoom-in 3 on where PVS ends; (F) Illustration of boundaries?
+(G) Concept illustration of the mathematical model (diffusion, convection, exchange, boundaries).}}
+%% \label{fig:concept}
+%% \end{figure}
+
 \begin{table}

### 2024-11-12 paper/supplementary_main.tex
Add supplementary_main.tex
Former-commit-id: f5966ce73d40f0f2989861844bd9d2ddaf8fb249 [formerly f5966ce73d40f0f2989861844bd9d2ddaf8fb249 [formerly 49c575915cf10ea4372752097ecaae050c401d8f]]
Former-commit-id: afb79e000b043a29137335afab62a471c14ab085
Former-commit-id: 55cf4239e4a3693f6300c1e6450e2bcceabcd2c2

\section{Supplementary methods}

\subsection{Mathematical model definitions and further model specifications}
\label{sec:more_maths}

In these sections, we provide a complete definition of the
mathematical models considered.

\subsubsection{Notation and definitions}

\subsubsection{Solute transport and exchange in intracranial domains (3D) and perivascular networks (1D)}

In this section, we describe the mathematical model~\eqref{eq:multi_transport} and associated conditions in further detail.

At bifurcation nodes in $\Gamma_a, \Gamma_v$, we impose flux conservation and continuity of the concentration $\hat c$. We refer to \eqref{eq:flux_conservation} in \Cref{sec:details_numerical_method} for more details. For the perivascular network, the boundary nodes are collected into $\partial \Lambda$. We assume no tracer outflow on all boundary nodes and prescribe a homogeneous Neumann condition: 
\begin{equation}
    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \partial \Lambda. 
\end{equation}

Here $\phi$ is the fluid volume fraction (or porosity) of the brain (where $\phi \ll 1$), of the SAS (where $\phi = 1$) and of the PVS (where $0 \ll \phi \leq 1$); $D$ is the effective diffusion coefficient of the relevant solute in the respective media~\cite{sykova2008diffusion} which takes different values over the SAS and brain,  $\bm u$ and $\hat u$ are convective velocity fields in $\Omega$ and $\Lambda$ respectively, and $\xi$ is a transfer coefficient across the PVS-CSF and the PVS-Brain interface

The notation $\overline{c}$ denotes the lateral average of the concentration over the outer perivascular boundary \footnote{
The lateral average is defined as 
\[ \overline{c}(s) = \frac{1}{A(s)} \int_{\Theta(s)} c ,\]
where $\Theta(s)$ is a cross-section of the PVS at the point on $\bm{\lambda}(s)$ on $\Lambda$ where $\bm \lambda$ is a given parameterization of $\Lambda$ and $A(s)$ is the area of $\Theta(s)$. 
}. The (Dirac delta) term $\delta_\Gamma$ is a source term concentrated on the outer boundary of the PVS (the union of the annular domains surrounding $\Lambda^i$) that models the solute exchange between the PVSs and their surroundings \footnote{The term $\delta_{\Gamma}$ is given for any function $v$, see \Cref{sec:details_numerical_method} for more details
\[\xi  (\overline{c} - \hat c ) \delta_\Gamma (v) 
= \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} 
\]
}.

\[
%\phi =  \begin{cases}
%   1  & \mathrm{in} \;  \Omega_{\rm{csf}} \\ 
%   \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
%   \end{cases}, \; 
%D= \begin{cases}
%   D_{\rm{csf}} & \mathrm{in} \;  \Omega_{\rm{csf}} \\ 
%   D_{\rm{PAR}} (1+ R) & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
%   \quad \bm u  = \begin{cases}
%   \bm{u}_{\rm{csf}} & \mathrm{in} \; \Omega_{\rm{csf}} \\ 
%   \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
%\end{cases}, \; 
\xi = \begin{cases}
\xi_{\rm{PVS}, \rm{PAR}} & \mathrm{if} |\Lambda^i \cap \Omega_{\rm{PAR}}| \neq 0 \\ 
 \xi_{\rm{PVS}, \rm{csf}} & \mathrm{if} |\Lambda^i \cap \Omega_{\rm{csf}}| \neq 0
\end{cases} . 
\]
In the above,  $|\Lambda^i \cap \Omega_{\rm{PAR}}|$ (resp. $|\Lambda^i \cap \Omega_{\rm{csf}}|$) is nonzero whenever the surrounding PVS around the line $\Lambda^i$ intersects $\Omega_{\rm{PAR}}$ (resp. $\Omega_{\rm{csf}}$). If the surrounding PVS intersects both, then we set $\xi = \xi_{\mathrm{PVS}, \rm{PAR}}$ (resp $\xi_{\mathrm{PVS}, \rm{CSF}}$) if the PVS lies mostly ($80$ percent) in $\Omega_{\rm{PAR}}$ (resp. $\Omega_{\rm{CSF}}$) and the interaction with $\Omega_{\rm{PVS}}$ (resp. $\Omega_{\rm{CSF}}$ is ignored.

Further, $P$ and $A$ are the perimeter and area of the PVS which can vary along the centerline network, respectively.  Finally, we remark that $\hat u$ takes different values on each $\Lambda_i$.  Both $\Omega$ and $\Lambda$ are fixed in time and we assume constant-in- time radii (despite varying radii to drive peristaltic flow, see \cref{sec:sup:peristalsis}). Finally, we note that the model assumes that there is no interaction between the inner perivascular boundary and the blood vessel.

\paragraph*{Interface conditions}
On the interface $\Gamma_{\mathrm{pia}}$ between the CSF and brain, we prescribe the following condition modeling a semi--permeable pial interface 
\begin{alignat}{2}
(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{PAR}} &= - (- D \nabla c_{\rm{CSF}} + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n}_{\rm{PAR}} &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}} ,  \\  
(-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{PAR}} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad  \pia \cup \Gamma_{\mathrm{LV}} .  
\end{alignat} 
Here, $\beta_{\mathrm{pia}}$ is a permeability constant between the brain and the CSF:

\paragraph*{Boundary conditions}
On $\Gamma_{\mathrm{skull}}$, the interface towards the dura and skull (\Cref{fig:concept}), we consider a constant uptake (or efflux) rate, represented by the boundary condition 
\begin{alignat}{2}
(- D \nabla ( c_{\mathrm{CSF}}) + \bm u_{\mathrm{CSF}} c_{\mathrm{CSF}}) \cdot \bm{n}   & = \beta c_{\mathrm{CSF}}  && \quad \mathrm{on} \quad \Gamma_{\mathrm{skull}^{\rm u}},\\ 
(-D \nabla c_{\rm CSF} + \bm u c_{\rm CSF}) \cdot \bm{n} & =  0   &&  \quad  \mathrm{on} \quad \Gamma_{\rm skull^{l}} , 
\end{alignat}
where $\beta$ represents a molecular outflow resistance. Here, we take $\beta = 10^{-4} \,\, \mathrm{mm}^2/\mathrm{s}$\cite{hornkjol2022csf}.

\subsubsection{Incompressible Stokes flow of CSF in the SAS and ventricular system}

We consider non-zero convection in the CSF space $u_{\rm CSF} \not = 0$ in \eqref{eq:multi_transport_3d} resulting from steady state CSF production.  Recall that $\Omega_{\mathrm{CSF}}$ represents the CSF space and that $\Gamma_{\mathrm{skull^{u/l}}}$ and $\Gamma_{\mathrm{pia}}$, $\Gamma_{\mathrm{LV}}$ represent the boundaries facing the upper and lower parts of the dura, the pia and the lateral ventricles, respectively.  Assuming CSF production in the choroid plexus located at the lateral ventricle wall, we solve the  steady state Stokes equations for the velocity and the pressure $(\bm u_{\mathrm{CSF}}, p_{\mathrm{CSF}})$ in the CSF domain: 
\begin{subequations}
    \begin{alignat}{2}
 - \mu \Delta \bm u_{\mathrm{CSF}} + \nabla p_{\rm{CSF}} & =  0 \quad && \mathrm{in} \,\,  \Omega_{\rm CSF}, \label{eq:momnetum_equation}  \\ 
 \nabla \cdot  \bm u_{\mathrm{CSF}} & = 0 \quad && \mathrm{in} \,\,   \Omega_{\rm CSF}, \label{eq:divergence_equation}  \\ 
\mu \nabla \bm u_{\mathrm{CSF}} \cdot \bm{n} -  p \bm n  &  = -R_0 ( \bm u \cdot \bm n ) \bm n\,\,   && \mathrm{on}  \label{eq:efflux_condition} \,\, \Gamma_{\mathrm{skull^u}}, \\ 
\bm u_{\mathrm{CSF}} & = 0 && \mathrm{on} \,\, \Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull^l}} \Gamma_{\rm{SSAS}}  \\
\bm u_{\mathrm{CSF}} \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm \tau = 0 \quad && \mathrm{on} \,\, \Gamma_{\rm{LV}} .  
\end{alignat}
\end{subequations}

Here, $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the CSF viscosity, and $\bm u_{\rm in}$ is the steady CSF flow across the lateral ventricle wall, see \Cref{tab:pvs:parameters}.
Condition \eqref{eq:efflux_condition} models an efflux site with positive resistance $R_0 \geq 0$. The Stokes system is solved with the finite element software $FEniCS$, stored and then subsequently read for all the relevant solute transport simulations. See Figure~\ref{fig:csf_flow_cardiac} for a streamline visualization of the obtained velocity profile and \Cref{sec:details_numerical_method} for details on the numerical algorithm employed.

We do not explicitly model convection within the brain parenchyma $\Omega_{\rm PAR}$, and thus set ${\bm u} = {\bm u}_{\rm PAR} = 0$ there.

\subsubsection*{Dispersion in the CSF spaces}

In addition, CSF pulsates in the SAS and ventricular system with each
cardiac and respiratory cycle.

Taking the large spatial differences of pulsatile CSF flow into account, we compute the local diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac induced CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}.

Sharp et al \cite{keith2019dispersion} estimated the enhancement of diffusion in the spinal subarachnoid space due to dispersive effects to average 5.8 times that of purely molecular diffusion with a theoretical maximum of $10^6$.

Similar to the case of CSF flow driven by production introduced in section \ref{sec:csf_fluid_vel}, we employ a Stokes flow model to predict CSF velocities at peak arterial dilation. In particular, we enforce a uniformly distributed inflow of $10$\,ml/s and $2$\,ml/s across the outer skull ($\Gamma_{\rm skull}$) and lateral ventricular surfaces ($\Gamma_{\rm LV}$) respectively, mimicking the effect of reduced CSF volume due to blood inflow. We prescribe a no-slip boundary condition on the pial surface $\Gamma_{\rm Pia}$ and zero-traction on the spinal boundary $\Gamma_{\rm SSAS}$, making the spinal compartment the only outflow route of CSF. Thus, we compute the cardiac-driven flow field by augmenting the Stokes flow model (eq. \ref{eq:momnetum_equation} and \ref{eq:divergence_equation}) with the following boundary conditions:

\begin{equation}\label{eq:cardiac_flow_bcs}
    \bm u = 0 \text{ on } \Gamma_{\rm Pia}; \quad 
    \mu \nabla \bm u \cdot \bm n - p \bm n = 0 \text{ on } \Gamma_{\rm SSAS}; \quad 
    \bm u = \frac{0.31 \cdot \bm n}{|\Gamma_{\rm LV}|} \,\text{ml/s} \text{ on } \Gamma_{\rm LV}; \quad
    \bm u = \frac{8 \cdot \bm n}{|\Gamma_{\rm skull}|} \,\text{ml/s} \text{ on } \Gamma_{\rm skull}.
\end{equation}

The obtained pressure field only account for viscous forces, albeit cardiac induced CSF flow is known to be unsteady and substantially impacted by transient inertial forces. Assuming an oscillatory flow with the heartbeat frequency $\omega = 2 \pi$, a mean gap width of the CSF-filled spaces of $2h=3$\,mm and CSF density $\rho=10^3\,\text{kg/m}^3$, the Womersley number $\alpha$ becomes $\alpha^2 = \frac{h^2 \omega \rho}{\mu} = 20.2$, which is unsteady. Based on theoretical considerations on the ratio of oscillatory flow to steady flow impedances in a tube, we account for inertial forces in the pulsatile flow by upscaling the computed viscous pressure field by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter. 4.3).

Further assuming unsteady dispersion, we follow Sharp et al \cite{keith2019dispersion} and calculate the enhancement factor $R$ from the non-dimensionalized pressure gradient $dP=\frac{|\nabla p|}{\omega \mu / h}$ as $R\approx dP^2 / \alpha^3$. Combining the computed pressure gradient with this scaling law yields a pointwise approximation of the dispersion factor. Finally, we account for the non-local nature of the dispersion mechanism by applying a smoothing technique to obtain a spatially varying, but locally averaged diffusion enhancement factor $R$ (see the visualization in \cref{fig:visualize_R} and \cref{sec:dispersion_details} for additional details).

\subsubsection{Perivascular fluid flow induced by pressure gradients}

The production of CSF in the choroid plexus generates a small, but consistent pressure gradient within the intracranial space. To compute the resulting CSF flow within the PVS, we solve 1D Darcy equations in the PVS network. The unknowns are the averaged PVS velocity, $u_{\rm pvs} = \langle u_{v,s} \rangle$, and the cross-section pressure $p_{\rm pvs} $ (assumed to be constant in each cross-section). The system is given by  \cite{daversin2022geometrically, gjerde2024directional} 
\begin{alignat}{2}
A \,  u_{\rm pvs}   + \frac{\kappa}{\mu} \, A \, \partial_{s} p_{\rm pvs} & = 0, &&  \quad \rm in  \,\, \Lambda  \\ 
-\partial_s (A \, u_{\rm pvs}) & = f, && \quad \rm in  \,\, \Lambda .  
\end{alignat} 
In the above, $A$ is the area of the PVS, $\mu$ is the dynamic viscosity, and $\kappa$ is derived from the assumption of Poiseuille
flow in the annular cross-section of the PVS \cite{daversin2022geometrically,tithof2022network}: 
\begin{equation}
\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
\end{equation}
On bifurcations, conservation of fluxes is enforced weakly and continuity of the pressure is enforced strongly (in the choice of the finite element spaces).

Since the pressure gradient the caused by the steady production-driven CSF flow (see \cref{sec:csf_fluid_vel}), we impose the precomputed fluid pressure $p_{\rm CSF}$ on the networks root nodes:

\begin{equation}
    p_{\rm PVS} = p_{\rm CSF} \quad \text{ on } \Lambda^{\rm out}
\end{equation}

In case that the root does not lie within the CSF-space but the parenchyma, the CSF pressure is not yet computed. For that reason, we apply a simple extension technique to map the CSF pressure field into $\Omega_{\rm PAR}$. In particular, we solve a Laplace equation on $\Omega_{\rm PAR}$ with the CSF pressure on the Pial membrane ($\Gamma_{\rm Pia}$) as a boundary condition.

\subsection{Estimating net perivascular flow induced by peristaltic waves}
\label{sec:sup:peristalsis}

We use the theoretical framework introduced by Gjerde et al.~\cite{gjerde2023directional} to compute an analytic estimate of the time-averaged (or \emph{net}) flow rates $\langle Q'_i \rangle$ (mm$^3$/s) induced by peristaltic pumping in a perivascular network$\Lambda = \cup_i \Lambda_i$. The motion of the (inner) vascular wall is described by a periodic traveling (peristaltic) wave of relative amplitude $\varepsilon$, wave length $\lambda$ (mm) and frequency $f$, acting normal to the wall. By definition, the wave number is $k = 2 \pi/\lambda$ and the angular frequency is $\omega = 2 \pi f$. Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L$, baseline inner radius $r_{o_i}'$, fixed outer radius $r_{e_i}'$, and outer-to-inner ratio $\beta_i = r_{e_i}'/r_{o_i}'$ (see also \Cref{tab:pvs:parameters}). These geometric parameters and the assumption of annular cylindrical PVS segments yield hydraulic resistances $\mathcal{R}_{o, i}$ and additional characteristic parameters, see \cite{gjerde2023directional} for the complete definitions and schematics. Since the analytical estimate is derived under the assumption that $k L \approx \mathcal{O} 1)$~\cite{gjerde2023directional} and has been verified against numerical simulations for $k L$ of the order $10^{-1}-10^2$~\cite[Table I]{gjerde2023directional}, we consider it applicable for the wave lengths and vascular network data considered here in which $k L$ ranges from \mer{$X$ to $Y$.}

\begin{table}
  \small
  \begin{tabular}{llrc}
    \toprule
    Parameter & Description & Value(s)  & Ref.\\ 
    \midrule
    Relative PVS size $\beta$ & $\beta = \beta_i = r_{e_i} / r_{o_i}$ & 2 & $\approx$\cite{eide2024functional} \\
    Wave frequency $f$ & Traveling wave frequency of vascular motion & $0.1-1.0$ Hz & \discuss{\cite{gjerde2023directional}} \\
    Wave length $\lambda$ & Traveling wave length of vascular motion & $20-2000$ mm & \discuss{\cite{broggini2024long, gjerde2023directional}} \\
    Wave amplitude & Relative amplitude of inner wall motion & $1-10\%$ & \discuss{\cite{gjerde2023directional}} \\
    Vascular radii $\{r_{o_i}\}$ & Vascular network radii data (avg $\pm$ std, $\min$, $\max$) & $X$, $X$, $X$  mm & \cite{hodneland2019new} \\
    Vascular branch lengths $\{L_{i}\}$ & Path length between vasc. junctions (avg $\pm$ std, $\min$, $\max$) & $X \pm Y$, $X$, $X$  mm & \discuss{--} \\
    \bottomrule
  \end{tabular}
  \caption{Perivascular flow induced by vascular wall motion: parameters and key data characteristics.}
  \label{tab:pvs:parameters}
\end{table}
% Vascular radii r_o (min, max, avg, std) (mm):  ....
% Vascular branch lengths L (avg, std, min, max) (mm): ...

\input{tikz/fig_pvs_flow}
This theoretical formalism~\cite{gjerde2023directional} is defined
relative to a network in the form of a directed, bifurcating tree with
a single supply node/root $i_0$. To extend to a network of cerebral
arteries with multiple supply nodes (such as the basilar and two
internal carotid arteries in the current data
set~\cite{hodneland2019new}), we separate the network $\Lambda$ into
edge-disjoint subnetworks $\Lambda^1, \Lambda^2, \Lambda^3$, one for
each of the supply nodes (\Cref{fig:sup:peristalsis}). Each node is assigned to the subnetwork associated with the nearest supply node, and edges between nodes are preserved. Next, we compute a minimal, bifurcating and directed tree representation of each subnetwork: $\mathcal{T}^1, \mathcal{T}^2, \mathcal{T}^3$. Each tree $\mathcal{T}^j = \cup E_n^j$ consists of the subset of the nodes from $\Lambda^j$ that have degree 1 (are leaf or root nodes) or degree $3$ (are true bifurcation points), and each path between nodes with degree $\not = 2$ in $\Lambda^j$ is represented in $\mathcal{T}^j$ by an edge $E^j$ with edge length $L$ corresponding to the total length of the original path and edge radius $r_o$ as the average of the path radii. For the few (2) nodes in $\Lambda$ with degree $>3$; i.e.~junctions where a branch splits into more than two downstream branches, only the two longest downstream paths are included in $\mathcal{T}^j$.

For each subtree $\mathcal{T}^j = \cup_n E_n^j$, we compute the
time-averaged downstream flow rate $\langle Q'_n \rangle$ induced by
the vascular wall motion for each edge $n$ via~\cite[eq.~(5),
  (34)]{gjerde2023directional}. We next uniquely assign the flow rate
$\langle Q'_n \rangle$ to each of the branches $\Lambda_i$ that form
the path $E_n^j$, thus yielding $\langle Q'_i \rangle$ for each
perivascular segment $\Lambda_i$ (except those ignored above, for
which we set a flow rate of zero). Finally, we define the mean
longitudinal perivascular velocity $\langle u^{\rm PVS}_i \rangle$ by
dividing the flow rate $\langle Q'_i \rangle$ by the cross-section
area $A_i = \pi (r_{e_i}^2 - r_{o_i}^2) = \pi (\beta_i^2 - 1) r_{o_i}^2$

\subsection{Numerical Method} \label{sec:details_numerical_method}
Here we provide details on the numerical method used to solve the coupled system. We are given a mesh $\mathcal{T}_h $, consisting of elements denoted by $E$, of $\Omega= \brain \cup \sas$.  The collection of all interior facets to $\Omega_{\rm PAR}$ (resp.  to $\Omega_{\mathrm{CSF}}$) is  denoted by $\mathcal{F}_{h,b}$ (resp. $\mathcal{F}_{h,c}$).  The union of all interior facets $\mathcal{F}_i = \mathcal{F}_{h,b} \cup \mathcal{F}_{h,c}$. Facets lying on the pial interface are denoted by $\mathcal{F}_{\pia}$, on the skull by $\mathcal{F}_{\skull}$, and on the spinal canal by $\mathcal{F}_{\spinal}$. Note that $\mathcal{F}_i \cap \mathcal{F}_{\pia} = \emptyset $. The union of all the facets is denoted by $\mathcal{F}_h$. The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is denoted by $\mathcal{E}_{h}$. 
\subsubsection{CSF Flow}
We recall the system of equations that we approximate. Find $(\bm u, p)$ such that 
\begin{subequations}
    \begin{alignat}{2}
 - \mu \Delta \bm u + \nabla p & =  \bm 0 \quad && \mathrm{in} \,\,  \Omega_{\rm CSF}, \label{eq:momnetum_equation}  \\ 
 \nabla \cdot  \bm u & = 0 \quad && \mathrm{in} \,\,   \Omega_{\rm CSF}, \label{eq:divergence_equation}  \\ 
\mu \nabla \bm u \cdot \bm{n} -  p \bm n  &  = -R_0 ( \bm u \cdot \bm n ) \bm n, \,\,  \quad \bm u \cdot \bm \tau = 0 \quad  && \mathrm{on}  \label{eq:efflux_condition} \,\, \Gamma_{\mathrm{skull^u}}, \\ 
\bm u & = 0 && \mathrm{on} \,\, \Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull^l}} \cup \Gamma_{\rm{SSAS}} \\
\bm u \cdot  \bm n   =  \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}  , & \; \; \bm u \cdot \bm \tau  = 0  && \mathrm{on} \;  \Gamma_{\mathrm{LV}},
\end{alignat}
\end{subequations}
We follow \cite{hong2016robust} and approximate the velocity field $u$ and the pressure $p$ with the following finite element spaces. 
\begin{align}
\bm V_h  & = \{ \bm v  \in H(\mathrm{div}; \Omega_{\mathrm{CSF}}): \; \bm v \vert_{E} \in \mathbb{BDM}^2 (E),  \; E \in \mathcal{E}_h; \;  \bm v \cdot \bm n = 0 \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull^l}} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} \text{ on } \Gamma_{\mathrm{LV}} \} \\ 
S_h  & = \{q \in L^2(\Omega):  \;  q \vert_E \in \mathbb{P}^1(E),  \; E \in \mathcal{E}_h \}. 
\end{align}
Here, $H(\mathrm{div};\Omega)$ is the space of $L^2$ vector fields with $L^2$ divergence,  $\mathbb{BDM}^2$ is the Brezzi-Douglas--Marini element \cite{brezzi1987mixed} of order 2, and  $\bm n$ (resp. $\bm \tau$) is the unit outward normal (resp. tangent) vector.  
Given any vector $\bm v$, the normal and tangential components on each facet  are denoted by 
\[ 
\bm v_n = (\bm v \cdot \bm n)\bm n , \quad \bm v_t = (\bm v - \bm v_n). 
\]
Since $V_h \subset H(\mathrm{div}, \Omega_{\rm{CSF}})$, then $[\bm v_n] = 0 $ on $\mathcal{F}_{h,c}$. Continuity in the tangential component is enforced weakly via interior penalization. For notational simplicity, denote by $\mathcal{F}_{\Gamma}$ the facets lying on $\Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull}} \cup \Gamma_{\rm{SSAS}} \cup \Gamma_{\mathrm{LV}}$. Define the form 
\begin{multline}
\mathcal{A}_h (\bm u, \bm v) = \sum_{E \in \mathcal{E}_h} \int_{E} \mu \nabla u : \nabla v - \sum_{
F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma} 
} \int_{F} \mu \{\nabla u\}\cdot \bm n_F \cdot [\bm v_t] 
\\ - \sum_{
F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma} 
} \int_{F} \mu \{\nabla v\}\cdot \bm n_F \cdot [\bm u_t]
+ \sum_{
F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma} 
} \frac{\sigma \mu}{h} \int_{F} [\bm u_t ] \cdot [\bm v_t].   
\end{multline}
We set $\sigma =20$. The discretization for the CSF flow, is then to find $(\bm u_h, p_h) \in \bm V_h \times S_h$ such  that 
\begin{align}
\mathcal{A}_h(u_h, v_h)  + \sum_{F \in \mathcal{F}_{\Gamma_{\mathrm{skull}}}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n) - \int_{\Omega} \nabla \cdot \bm v_h \, p_h  & = 0, \quad \forall \bm v_h \in \bm V_h   \\ 
(\nabla \cdot \bm u_h , q_h ) & = 0, \quad \forall q_h \in S_h.
\end{align}
\subsubsection{PVS flow}

Recall that we compute pressure-driven flow within the PVS by solving a 1D Darcy equations in the PVS network. The unknowns are the averaged PVS velocity, $u_{\rm pvs} = \langle u_{v,s} \rangle$, and the cross-section pressure $p_{\rm pvs} $ (assumed to be constant in each cross-section). The system is given by  \cite{daversin2022geometrically, 
gjerde2024directional}

\begin{alignat}{2}
A \,  u_{\rm pvs}   + \frac{\kappa}{\mu} \, A \, \partial_{s} p_{\rm pvs} & = 0, &&  \quad \rm in  \,\, \Lambda  \\ 
-\partial_s (A \, u_{\rm pvs}) & = f, && \quad \rm in  \,\, \Lambda .  
\end{alignat}

where $A$ is the area of the PVS, $\mu$ is the dynamic viscosity, and $\kappa$ is derived from the assumption of Poiseuille flow in the annular cross-section of the PVS  \cite{daversin2022geometrically,tithof2022network}: 
\begin{equation}
\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
\end{equation}

As boundary conditions, we impose the precomputed fluid pressure $p_{\rm CSF}$ on the networks root nodes:

\begin{equation}
    p_{\rm PVS} = p_{\rm CSF} \quad \text{ on } \Lambda^{\rm out}
\end{equation}

In case that the root does not lie within the CSF-space but the parenchyma, the CSF pressure is not yet computed. For that reason, we apply a simple extension technique to map the CSF pressure field into $\Omega_{\rm PAR}$.
In particular, we solve the Laplace equation
\begin{equation}
    - \Delta p = 0; \quad p = p_{\rm CSF} \text{ on } \Gamma_{\rm Pia}
\end{equation}
to obtain the extension of $p_{\rm CSF}$ onto $\Omega_{\rm PAR}$.

To enforce conservation of flux and continuity of pressure on bifurcation points, we approximate the velocity with a piecewise discontinuous polynomials and the pressure with continuous piecewise polynomials, respectively. The finite element spaces thus are:

\begin{align}
   V_h^{\rm PVS} = \{v_h \in L^2(\Lambda): v_h \vert_{\Lambda_i} \in
   \mathbb{P}^{k -1}(\Lambda_i)\} \text{ and }
   Q_h^{\rm PVS} = \{ q_h \in C^0(\Lambda),  q_h \vert_{\Lambda^i} \in \mathbb{P}^k(\Lambda_i), q_h \vert_{\Gamma_{\rm Pia}} = p_{\rm CSF} \}
\end{align}

The weak form reads: Find $(u_{\rm PVS}, p_{\rm PVS}) \in V_h^{\rm PVS} \times Q_h^{\rm PVS}$ such that

\begin{align}
(A u_{\rm PVS}, v) + \left(v, \frac{A \kappa}{\mu} \partial_s p_{\rm PVS}  \right) &= 0 \quad \forall v \in V_h^{\rm PVS} \\
(A u_{\rm PVS} , \partial_s q ) &= 0 \quad  \forall q \in  Q_h^{\rm PVS}
\end{align}

where $\tau$ is the tangent vector of the network. Here we employ the lowest order discretization and set $k=1$. For discussion of stability of the discrete model we refer to \cite{gjerde2024directional}.

\subsubsection{Solute transport} For the 3D concentration, we use broken polynomial spaces and interior penalty discontinuous Galerkin (DG) methods with upwinding in order to capture the sharp boundary layers.  For DG, we utilize the discrete space 
\begin{equation}
V_h^k = \{v_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
\end{equation}
Recall that the 3D concentration $c$ is defined locally in each subdomain of $\Omega$ and satisfies
\begin{equation}
     \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma = f \quad \quad \mathrm{in} \quad \Omega_{\rm CSF} \cup \Omega_{\rm{PAR}}.  \label{eq:3d_pde}
\end{equation}
The above is supplemented with the following boundary conditions. 
\begin{alignat}{2}
(D \nabla (\phi c) - \bm u c ) \cdot \bm{n}_{\rm SAS} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \spinal  \\ 
% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\spinal} &= 0 &&  \quad  \mathrm{on} \quad \spinal^{\mathrm{out}}  \\ 
(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\rm skull} & =   \beta c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm skull^{u}}  \\ 
(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\rm skull} & =   0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm skull^{l}}  \\ 
(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{Pia}} &= - (-D \nabla  c_{\rm{CSF}} + \bm u_{\rm CSF} c_{\rm CSF} ) \cdot \bm{n}_{\rm{Pia}} &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}}  \\ 
(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{Pia}} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}}   .  
\end{alignat}

% \mar{I think we can simplify the BCs here a bit, mostly since $\bm u \cdot \bm n$ is zero everywhere except for $\Gamma_{\rm LV}$ and $\Gamma_{\rm skull}$. So I suggest we remove all terms related to this? And we probably also don't need to split $\Gamma_{SC}$ in $\Gamma_{SC}^{in}$ and  $\Gamma_{SC}^{out}$?}

In the above, note that $\bm u \cdot \bm n  \neq 0$ on $\Gamma_{\rm LV} \cup \Gamma_{\rm skull^u}$ and $\bm u \cdot \bm n = 0$ on $\Gamma_{\rm SSAS} \cup \Gamma_{\rm skull^l} \cup \Gamma_{\rm Pia}$. To discretize the diffusion term, we utilize symmetric weighted interior penalty dG bilinear forms. We refer to \cite{ern2009discontinuous} and \cite[Section 4.5.2.3]{di2011mathematical} for details on this method.  We define 
\begin{multline}
a_h (c_h, v_h) = \sum_{E \in \mathcal{T}_h } \int_{E} D \nabla (\phi c_h) \cdot \nabla v_h + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c_h][v_h] 
- \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c_h \}_w \cdot \bm n_F [v_h]  + \{D \phi\nabla v_h \}_w \cdot \bm n_F [c_h] \right).
 \nonumber
 \end{multline}
In the above, for $F\in \mathcal{F}_h^i$ with a normal $n_F$ pointing from $E_F^1$ to $E_F^2$,  the jump is given by $[v_h] = v_h{}_{\vert_{E_F^1}} - v_h{}_{\vert_{E_{F}^2}}$, and the weighed average is defined as follows.
\begin{equation}
\{v_h\}_{w} = \frac{ \kappa_{\vert_{E_F^2}}}{\kappa_{\vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}} v_h{}_{\vert_{E_{F}^1}} + \frac{\kappa_{\vert_{E_F^1}}}{\kappa_{ \vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}} v_h{}_{\vert_{E_{F}^2}}, \quad \text{where} \quad \kappa = D \phi.  
\end{equation}
The parameter $\gamma_{D,F}$ is the harmonic mean of the diffusion coefficient given by 
$$
\gamma_{D,F} = \frac{ \kappa_{\vert_{E_F^1}}\kappa_{\vert_{E_F^2}}}{\kappa_{\vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}},
$$
and $\eta$ is a user specified penalty to be determined later.

For the convection term, we use upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the references therein. Define 
\begin{equation*}
a_h^{\mathrm{up}}(c_h, v_h) =- \sum_{E\in\mathcal{T}_h } \int_E \bm{u} c_h \nabla v_h  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c_h\} \cdot \bm{n}_F [v_h]  + \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F |}{2} [c_h ] [v_h ] 
\end{equation*}
The discrete formulation for \eqref{eq:3d_pde} then reads 
\begin{multline}
\int_{\Omega}\frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v_h + a_h(c_h^n, v_h) + a_h^{\mathrm{up}}(c_h^n,v_h) + \int_{\Gamma_{\mathrm{pia}} } \beta_{\mathrm{pia}}[c_h^n][v_h]  \\ + \int_{\Gamma_{\rm skull^u}} \beta c_h^n v_h +  \int_{\Lambda} \xi  P (\overline{c_h^n} - \hat c_h^n) \overline{v_h}  = \int_{\Omega} f^n v_h + \int_{\spinal} \gin v_h , \quad \forall v_h \in V_h^k.
\end{multline}
For the 1D concentration $\hat c$, we recall that $\hat c = \hat c_i $ on each segment of the network $\Lambda^i$.  We enforce continuity of the concentration on the bifurcation points and conservation of fluxes. We let $\mathcal{V}$ denote the set of all vertices in $\Lambda$ and $ 
\mathcal{E}(\mathsf{v})$ denote the set of segments sharing a vertex $\mathsf{v}$.  For a given edge $\Lambda^i= (\mathsf{v}_\mathrm{in}^i, \mathsf{v}_{\mathrm{out}}^{i})$, we define the function $n_i: \mathcal{V} \rightarrow \{-1,0,1\}$ with $$
n_i(\mathsf{v}_{\mathrm{in}}^i) = 1 , \,\, n _i (\mathsf{v}_{\mathrm{out}}^i) = -1 \,\, \mathrm{and} \,\,\, n_i(\mathsf{v}) = 0, \quad \forall \mathsf{v} \in 
\mathcal{V} \,  \backslash \,  \{ \mathsf{v}_{\mathrm{in}}^\mathsf e,\mathsf{v}_{\mathrm{out}}^\mathsf e \}.  
$$
The collection of bifurcation points is denoted by $\mathcal{B} = \{ \mathsf{v} \in \mathcal{V}, \,\,  \mathrm{card}(\mathcal{E}(\mathsf{v})) \geq  3\}$. For any $\mathsf{v} \in \mathcal{B}$, we have 
\begin{equation}
\sum_{ i \in \mathcal{E}(\mathsf{v})} (\hat D A_i \phi \partial_s \hat{u}_i(\mathsf{v})  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n_i(\mathsf{v}) = 0  \label{eq:flux_conservation}
\end{equation}
For approximating the solution of the 1D equations in the network, we use the space 
\[ \hat V_h = \{ v_h \in C^0(\Lambda), 
\; v_h \vert_{\Lambda^i} \in V_h(\Lambda^i)
\}, 
\]
where $V_h(\Lambda^i)$ is the space of continuous polynomials of degree $1$ on each $\Lambda^i$,

\subsection{Cardiac Dispersion Factor Estimation}\label{sec:dispersion_details}

To compute a spatially varying estimate of the diffusion enhancement factor $R$, we consider the Stokes flow problem (\cref{eq:momnetum_equation} and \cref{eq:divergence_equation}) equipped with the boundary conditions \cref{eq:cardiac_flow_bcs} describing cardiac induced CSF flow at peak arterial pulsation. We use the numerical method described in \cref{sec:details_numerical_method} to solve the Stokes system and compute the non-dimensionalized local pressure difference $dP=\frac{|\nabla p| \rho}{\omega \mu / h}$ pointwise. Here, $\mu$ is the CSF viscosity, $\omega = 2 \pi$ is the oscillatory heart beat frequency, $2h=3$\,mm the assumed gap width and $\rho=10^3\,\text{kg/m}^3$ the density of the CSF.

In addition, we account for inertial forces in the pulsatile flow by upscaling the obtained pressure gradient by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter. 4.3), where $\alpha^2 = \frac{h^2 \omega \rho}{\mu} = 20.2$ is the square of the Womersley number for an oscillatory flow.

Finally, we smooth the resulting field by solving a Helmholtz problem with $dP^2$ as the right-hand side

\begin{equation}
- 10^{-4} \Delta R + R = dP^2 \quad \text{ on } \Omega_{\rm CSF} \qquad \text{ and } \quad \nabla R \cdot \bm n = 0 \quad \text{ on } \partial \Omega_{\rm CSF}
\end{equation}

to obtain the locally averaged, spatially varying diffusion enhancement factor $R$.

\subsection{Numerical verification}

We assess the accuracy and numerical convergence of our simulation results by performing a series of experiments with different spatial and temporal resolutions. Specifically, we generate a sequence of three meshes with an increasing number of points and cells, solve all relevant simulation steps (CSF flow and intracranial transport computations) for standard model A, and compare the results with respect to a set of key quantities of interest. Similarly, we investigate the effect of the time step size by solving the intracranial transport model with varying time steps $dt \in \{60, 120, 240 \}$\,s.

Considering the mean tracer concentrations in the CSF, parenchyma and arterial and venous PVS domain over the first 24\,h after injection, we observe negligible changes with both mesh and time refinement (\cref{fig:mesh_convergence_concentrations} and 
\ref{fig:time_convergence_concentrations}). In addition, we compute the mean and maximum enhancement factor, the maximum CSF pressure and velocity in both the cardiac-driven and CSF production-induced flow fields, and mean concentrations at 3, 6, 12 and 24\,h for all mesh resolutions and time steps (\cref{fig:mesh_convergence} and \cref{fig:time_convergence}). While the maximum dispersion factor increases by about 60\,\% from the coarse to the standard mesh, it stabilizes with the next refinement step. All other quantities change with less than 10\,\% with mesh refinement, and less than 1\,\% with time refinement. We thus conclude that the standard resolution mesh and a time step of $dt=120\,$s offers sufficient accuracy for our simulations.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width = 1 \linewidth]{figures/mesh_refinement_par_csf_mean.png}
        \caption{Mean tracer concentration in CSF and parenchyma under mesh refinement}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
     \includegraphics[width= 1 \linewidth]{figures/mesh_refinement_art_ven_mean.png}
         \caption{Mean tracer concentration in PVSs around (veins and) arteries  under mesh refinement}
    \end{subfigure}
    \caption{Mean tracer concentrations over the first 24\,h after injection on the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the LowRes, standard and HighRes meshes with a timestep of $dt=120$\,s.}
    \label{fig:mesh_convergence_concentrations}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \includegraphics[trim={0.5cm 1cm 0.05cm 0.8cm}, clip, width = 1.05 \linewidth]{figures/mesh_refinement.pdf}
        \caption{Convergence of key quantities of interest with mesh refinement}
        \label{fig:mesh_convergence}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
     \includegraphics[trim={0.5cm 1cm 0.05cm 0.8cm}, clip,width= 1.05 \linewidth]{figures/time_refinement.pdf}
        \caption{Convergence of key quantities of interest with time refinement}
        \label{fig:time_convergence}
    \end{subfigure}
    \caption{minimal ($\rm h_{min}$) and maximal ($\rm h_{max}$) mesh cell sizes (computed as cell circumradius $\times 2$) (A); number of points and tetrahedral cells in each mesh (B); mean dispersion enhancement factor $R$ (C), maximum dispersion enhancement factor $R$ (D); maximum pressure in steady CSF production flow (E), maximum pressure in cardiac-driven CSF flow (F); maximum CSF velocity in steady CSF production flow (G); maximum CSF velocity in cardiac-driven CSF flow (H);  mean tracer concentration in the CSF, parenchyma, arterial PVS and venous PVS at 3,6,12 and 24\,h (I, J, K and L).}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width = 1 \linewidth]{figures/time_refinement_par_csf_mean.png}
        \caption{Mean tracer concentration in CSF and parenchyma under time step refinement}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
     \includegraphics[width= 1\linewidth]{figures/time_refinement_art_ven_mean.png}
         \caption{Mean tracer concentration in PVSs around (veins and) arteries under time step refinement}
    \end{subfigure}
    \caption{Mean tracer concentrations over the first 24\,h after injection on the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the standard resolution mesh for timesteps $dt \in \{60, 120, 240 \}$\,s.}    \label{fig:time_convergence_concentrations}
\end{figure}

\newpage
\section{Notes}

Various notes associated with \cite{eide2024functional}:
\begin{itemize}
  \item
    Can observations of tracer indicate compartmentalization without
    the existence of physical barriers but in the presence of
    e.g.~substantial flow? We made observations in this direction in
    \cite{vinje2021brain}, relating to the then-active
    shape-of-perivascular-spaces discussion. We can address this by
    trying to quantify and evaluate the permeability of the PVS-SAS
    interface - comparing model predictions with the timings and
    patterns presented by \cite{eide2024functional}.
  \item
    Do the tracer primarily move from the PVS into the SAS or along
    (more slowly) along the SAS? This is left as an open question by
    \cite{eide2024functional} (page 3)
  \item
    How do altered ICP affect transport? Note that PVSAS transport is
    impaired (delayed) with increasing mean wave amplitude in the ICP
    signal and for iNPH patients (with enlarged PVSAS spaces).
  \item
    Very useful reference for PVS sizes (10--20 mm$^2$ in reference
    patients, 15-70 mm$^2$ in iNPH), and average propagation speeds.
\end{itemize}

### 2024-11-12 paper/supplementary_main.tex
Clean-ups
Former-commit-id: 386f212cee0a5a09e05ddfa15b8e59d054bcc2c3 [formerly 3aaecf5b8769e98b3e4aa00676e63a48fee3ba03]
Former-commit-id: 7152b8a813ccaba34a88f72b36c2592d6eed8397

@@
 
-\subsection{Mathematical model definitions and further model specifications}
+These sections provide more detailed descriptions of the mathematical
+models and numerical approximations considered.
+
+\subsection{Mathematical models and definitions}
 \label{sec:more_maths}
 
-In these sections, we provide a complete definition of the
-mathematical models considered.
+\subsubsection{Notation}
 
-\subsubsection{Notation and definitions}
+In terms of domains, interfaces and boundaries, we consider the
+parenchyma $\Omega_{\rm PAR} \subset \R^3$ and CSF spaces $\Omega_{\rm
+  CSF} \subset \R^3$ with $\Omega = \Omega_{\rm PAR} \cup \Omega_{\rm
+  CSF}$. The interface between the parenchyma and CSF spaces is
+decomposed
+
+\subsection*{Notation and left-overs}
+
+\paragraph*{To be moved: notation and labels}
+The thin pia membrane encloses the brain, and we hence denote the CSF-brain interface with $\Gamma_{\rm pia}$, except for the separately marked surface of the lateral ventricles $\Gamma_{\rm LV}$ (thus $\partial \Omega_{\rm CSF} \bigcap \partial \Omega_{\rm PAR} = \Gamma_{\rm LV} \cup \Gamma_{\rm Pia}$).
+
+We denote the different parts of the outer boundary of $\Omega_{\rm
+  CSF}$ with $\Gamma_{\mathrm{skull}}$ facing the dura and skull and
+$\Gamma_{\rm SSAS}$ representing the lower interface towards the
+spinal subarachnoid space (SSAS). $\Gamma_{\mathrm{skull}}$ is further
+subdivided into its lower and upper parts: $\Gamma_{\mathrm{skull^l}}$
+and $\Gamma_{\mathrm{skull^u}}$ (compare \Cref{fig:concept}). The
+boundary of the spinal cord is denoted by $\Gamma_{\rm{SC}}$, which is
+given by $\Gamma_{\rm{SC}} = \partial \Omega_{\rm PAR} \setminus
+\partial \Omega_{\rm CSF}$ .
+
+%% We denote the domain defined by the connected centerlines by
+%% $\Lambda$.The vascular domain is then represented as the union of
+%% cylindrical vessels of radius $R_1^i$ surrounding the centerlines
+%% $\Lambda^i$. Moreover, we consider the surrounding perivascular spaces
+%% as the union of annular cylinders of inner radius $R_1^i$ and outer
+%% radius $R_2^i = \beta R_1^i$ of width $R_2^i - R_1^i = (\beta - 1)
+%% R_1^i$.
 
@@
 In this section, we describe the mathematical model~\eqref{eq:multi_transport} and associated conditions in further detail.
+
+Recall that the 3D concentration $c$ is defined locally in each subdomain of $\Omega$ and satisfies
+\begin{equation}
+     \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma = f \quad \quad \mathrm{in} \quad \Omega_{\rm CSF} \cup \Omega_{\rm{PAR}}.
+\label{eq:3d_pde}
+\end{equation}
+The above is supplemented with the following boundary conditions. 
+\begin{alignat}{2}
+(D \nabla (\phi c) - \bm u c ) \cdot \bm{n}_{\rm SAS} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \spinal  \\ 
+% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\spinal} &= 0 &&  \quad  \mathrm{on} \quad \spinal^{\mathrm{out}}  \\ 
+(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\rm skull} & =   \beta c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm skull^{u}}  \\ 
+(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\rm skull} & =   0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm skull^{l}}  \\ 
+(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{Pia}} &= - (-D \nabla  c_{\rm{CSF}} + \bm u_{\rm CSF} c_{\rm CSF} ) \cdot \bm{n}_{\rm{Pia}} &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}}  \\
+(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{Pia}} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}}   .
+\end{alignat}
+
+% \mar{I think we can simplify the BCs here a bit, mostly since $\bm u \cdot \bm n$ is zero everywhere except for $\Gamma_{\rm LV}$ and $\Gamma_{\rm skull}$.
+So I suggest we remove all terms related to this?
+And we probably also don't need to split $\Gamma_{SC}$ in $\Gamma_{SC}^{in}$ and  $\Gamma_{SC}^{out}$?}
+
+In the above, note that $\bm u \cdot \bm n  \neq 0$ on $\Gamma_{\rm LV} \cup \Gamma_{\rm skull^u}$ and $\bm u \cdot \bm n = 0$ on $\Gamma_{\rm SSAS} \cup \Gamma_{\rm skull^l} \cup \Gamma_{\rm Pia}$.
 
@@
 
-\subsection{Estimating net perivascular flow induced by peristaltic waves}
-\label{sec:sup:peristalsis}
-
-We use the theoretical framework introduced by Gjerde et al.~\cite{gjerde2023directional} to compute an analytic estimate of the time-averaged (or \emph{net}) flow rates $\langle Q'_i \rangle$ (mm$^3$/s) induced by peristaltic pumping in a perivascular network$\Lambda = \cup_i \Lambda_i$.
-The motion of the (inner) vascular wall is described by a periodic traveling (peristaltic) wave of relative amplitude $\varepsilon$, wave length $\lambda$ (mm) and frequency $f$, acting normal to the wall.
-By definition, the wave number is $k = 2 \pi/\lambda$ and the angular frequency is $\omega = 2 \pi f$.
-Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L$, baseline inner radius $r_{o_i}'$, fixed outer radius $r_{e_i}'$, and outer-to-inner ratio $\beta_i = r_{e_i}'/r_{o_i}'$ (see also \Cref{tab:pvs:parameters}).
-These geometric parameters and the assumption of annular cylindrical PVS segments yield hydraulic resistances $\mathcal{R}_{o, i}$ and additional characteristic parameters, see \cite{gjerde2023directional} for the complete definitions and schematics.
-Since the analytical estimate is derived under the assumption that $k L \approx \mathcal{O} 1)$~\cite{gjerde2023directional} and has been verified against numerical simulations for $k L$ of the order $10^{-1}-10^2$~\cite[Table I]{gjerde2023directional}, we consider it applicable for the wave lengths and vascular network data considered here in which $k L$ ranges from \mer{$X$ to $Y$.}
-
-\begin{table}
-  \small
-  \begin{tabular}{llrc}
-    \toprule
-    Parameter & Description & Value(s)  & Ref.\\ 
-    \midrule
-    Relative PVS size $\beta$ & $\beta = \beta_i = r_{e_i} / r_{o_i}$ & 2 & $\approx$\cite{eide2024functional} \\
-    Wave frequency $f$ & Traveling wave frequency of vascular motion & $0.1-1.0$ Hz & \discuss{\cite{gjerde2023directional}} \\
-    Wave length $\lambda$ & Traveling wave length of vascular motion & $20-2000$ mm & \discuss{\cite{broggini2024long, gjerde2023directional}} \\
-    Wave amplitude & Relative amplitude of inner wall motion & $1-10\%$ & \discuss{\cite{gjerde2023directional}} \\
-    Vascular radii $\{r_{o_i}\}$ & Vascular network radii data (avg $\pm$ std, $\min$, $\max$) & $X$, $X$, $X$  mm & \cite{hodneland2019new} \\
-    Vascular branch lengths $\{L_{i}\}$ & Path length between vasc. junctions (avg $\pm$ std, $\min$, $\max$) & $X \pm Y$, $X$, $X$  mm & \discuss{--} \\
-    \bottomrule
-  \end{tabular}
-  \caption{Perivascular flow induced by vascular wall motion: parameters and key data characteristics.}
-  \label{tab:pvs:parameters}
-\end{table}
-% Vascular radii r_o (min, max, avg, std) (mm):  ....
-% Vascular branch lengths L (avg, std, min, max) (mm): ...
-
-\input{tikz/fig_pvs_flow}
-This theoretical formalism~\cite{gjerde2023directional} is defined
-relative to a network in the form of a directed, bifurcating tree with
-a single supply node/root $i_0$. To extend to a network of cerebral
-arteries with multiple supply nodes (such as the basilar and two
-internal carotid arteries in the current data
-set~\cite{hodneland2019new}), we separate the network $\Lambda$ into
-edge-disjoint subnetworks $\Lambda^1, \Lambda^2, \Lambda^3$, one for
-each of the supply nodes (\Cref{fig:sup:peristalsis}).
-Each node is assigned to the subnetwork associated with the nearest supply node, and edges between nodes are preserved.
-Next, we compute a minimal, bifurcating and directed tree representation of each subnetwork: $\mathcal{T}^1, \mathcal{T}^2, \mathcal{T}^3$.
-Each tree $\mathcal{T}^j = \cup E_n^j$ consists of the subset of the nodes from $\Lambda^j$ that have degree 1 (are leaf or root nodes) or degree $3$ (are true bifurcation points), and each path between nodes with degree $\not = 2$ in $\Lambda^j$ is represented in $\mathcal{T}^j$ by an edge $E^j$ with edge length $L$ corresponding to the total length of the original path and edge radius $r_o$ as the average of the path radii.
-For the few (2) nodes in $\Lambda$ with degree $>3$; i.e.~junctions where a branch splits into more than two downstream branches, only the two longest downstream paths are included in $\mathcal{T}^j$.
-
-For each subtree $\mathcal{T}^j = \cup_n E_n^j$, we compute the
-time-averaged downstream flow rate $\langle Q'_n \rangle$ induced by
-the vascular wall motion for each edge $n$ via~\cite[eq.~(5),
-  (34)]{gjerde2023directional}. We next uniquely assign the flow rate
-$\langle Q'_n \rangle$ to each of the branches $\Lambda_i$ that form
-the path $E_n^j$, thus yielding $\langle Q'_i \rangle$ for each
-perivascular segment $\Lambda_i$ (except those ignored above, for
-which we set a flow rate of zero). Finally, we define the mean
-longitudinal perivascular velocity $\langle u^{\rm PVS}_i \rangle$ by
-dividing the flow rate $\langle Q'_i \rangle$ by the cross-section
-area $A_i = \pi (r_{e_i}^2 - r_{o_i}^2) = \pi (\beta_i^2 - 1) r_{o_i}^2$
-
-\subsection{Numerical Method} \label{sec:details_numerical_method}
+\subsection{Numerical Method}
+\label{sec:details_numerical_method}
 Here we provide details on the numerical method used to solve the coupled system.
@@
 
-\subsubsection{Solute transport} For the 3D concentration, we use broken polynomial spaces and interior penalty discontinuous Galerkin (DG) methods with upwinding in order to capture the sharp boundary layers.
+\subsubsection{Solute transport}
+For the 3D concentration, we use broken polynomial spaces and interior penalty discontinuous Galerkin (DG) methods with upwinding in order to capture the sharp boundary layers.
 For DG, we utilize the discrete space
@@
 \end{equation}
-Recall that the 3D concentration $c$ is defined locally in each subdomain of $\Omega$ and satisfies
-\begin{equation}
-     \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma = f \quad \quad \mathrm{in} \quad \Omega_{\rm CSF} \cup \Omega_{\rm{PAR}}.
-\label{eq:3d_pde}
-\end{equation}
-The above is supplemented with the following boundary conditions. 
-\begin{alignat}{2}
-(D \nabla (\phi c) - \bm u c ) \cdot \bm{n}_{\rm SAS} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \spinal  \\ 
-% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\spinal} &= 0 &&  \quad  \mathrm{on} \quad \spinal^{\mathrm{out}}  \\ 
-(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\rm skull} & =   \beta c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm skull^{u}}  \\ 
-(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\rm skull} & =   0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm skull^{l}}  \\ 
-(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{Pia}} &= - (-D \nabla  c_{\rm{CSF}} + \bm u_{\rm CSF} c_{\rm CSF} ) \cdot \bm{n}_{\rm{Pia}} &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}}  \\
-(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{Pia}} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}}   .
-\end{alignat}
-
-% \mar{I think we can simplify the BCs here a bit, mostly since $\bm u \cdot \bm n$ is zero everywhere except for $\Gamma_{\rm LV}$ and $\Gamma_{\rm skull}$.
-So I suggest we remove all terms related to this?
-And we probably also don't need to split $\Gamma_{SC}$ in $\Gamma_{SC}^{in}$ and  $\Gamma_{SC}^{out}$?}
-
-In the above, note that $\bm u \cdot \bm n  \neq 0$ on $\Gamma_{\rm LV} \cup \Gamma_{\rm skull^u}$ and $\bm u \cdot \bm n = 0$ on $\Gamma_{\rm SSAS} \cup \Gamma_{\rm skull^l} \cup \Gamma_{\rm Pia}$.
 To discretize the diffusion term, we utilize symmetric weighted interior penalty dG bilinear forms.
@@
 to obtain the locally averaged, spatially varying diffusion enhancement factor $R$.
+
+\subsection{Estimating net perivascular flow induced by peristaltic waves}
+\label{sec:sup:peristalsis}
+
+We use the theoretical framework introduced by Gjerde et al.~\cite{gjerde2023directional} to compute an analytic estimate of the time-averaged (or \emph{net}) flow rates $\langle Q'_i \rangle$ (mm$^3$/s) induced by peristaltic pumping in a perivascular network$\Lambda = \cup_i \Lambda_i$.
+The motion of the (inner) vascular wall is described by a periodic traveling (peristaltic) wave of relative amplitude $\varepsilon$, wave length $\lambda$ (mm) and frequency $f$, acting normal to the wall.
+By definition, the wave number is $k = 2 \pi/\lambda$ and the angular frequency is $\omega = 2 \pi f$.
+Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L$, baseline inner radius $r_{o_i}'$, fixed outer radius $r_{e_i}'$, and outer-to-inner ratio $\beta_i = r_{e_i}'/r_{o_i}'$ (see also \Cref{tab:pvs:parameters}).
+These geometric parameters and the assumption of annular cylindrical PVS segments yield hydraulic resistances $\mathcal{R}_{o, i}$ and additional characteristic parameters, see \cite{gjerde2023directional} for the complete definitions and schematics.
+Since the analytical estimate is derived under the assumption that $k L \approx \mathcal{O} 1)$~\cite{gjerde2023directional} and has been verified against numerical simulations for $k L$ of the order $10^{-1}-10^2$~\cite[Table I]{gjerde2023directional}, we consider it applicable for the wave lengths and vascular network data considered here in which $k L$ ranges from \mer{$X$ to $Y$.}
+
+\begin{table}
+  \small
+  \begin{tabular}{llrc}
+    \toprule
+    Parameter & Description & Value(s)  & Ref.\\ 
+    \midrule
+    Relative PVS size $\beta$ & $\beta = \beta_i = r_{e_i} / r_{o_i}$ & 2 & $\approx$\cite{eide2024functional} \\
+    Wave frequency $f$ & Traveling wave frequency of vascular motion & $0.1-1.0$ Hz & \discuss{\cite{gjerde2023directional}} \\
+    Wave length $\lambda$ & Traveling wave length of vascular motion & $20-2000$ mm & \discuss{\cite{broggini2024long, gjerde2023directional}} \\
+    Wave amplitude & Relative amplitude of inner wall motion & $1-10\%$ & \discuss{\cite{gjerde2023directional}} \\
+    Vascular radii $\{r_{o_i}\}$ & Vascular network radii data (avg $\pm$ std, $\min$, $\max$) & $X$, $X$, $X$  mm & \cite{hodneland2019new} \\
+    Vascular branch lengths $\{L_{i}\}$ & Path length between vasc. junctions (avg $\pm$ std, $\min$, $\max$) & $X \pm Y$, $X$, $X$  mm & \discuss{--} \\
+    \bottomrule
+  \end{tabular}
+  \caption{Perivascular flow induced by vascular wall motion: parameters and key data characteristics.}
+  \label{tab:pvs:parameters}
+\end{table}
+% Vascular radii r_o (min, max, avg, std) (mm):  ....
+% Vascular branch lengths L (avg, std, min, max) (mm): ...
+
+\input{tikz/fig_pvs_flow}
+This theoretical formalism~\cite{gjerde2023directional} is defined
+relative to a network in the form of a directed, bifurcating tree with
+a single supply node/root $i_0$. To extend to a network of cerebral
+arteries with multiple supply nodes (such as the basilar and two
+internal carotid arteries in the current data
+set~\cite{hodneland2019new}), we separate the network $\Lambda$ into
+edge-disjoint subnetworks $\Lambda^1, \Lambda^2, \Lambda^3$, one for
+each of the supply nodes (\Cref{fig:sup:peristalsis}).
+Each node is assigned to the subnetwork associated with the nearest supply node, and edges between nodes are preserved.
+Next, we compute a minimal, bifurcating and directed tree representation of each subnetwork: $\mathcal{T}^1, \mathcal{T}^2, \mathcal{T}^3$.
+Each tree $\mathcal{T}^j = \cup E_n^j$ consists of the subset of the nodes from $\Lambda^j$ that have degree 1 (are leaf or root nodes) or degree $3$ (are true bifurcation points), and each path between nodes with degree $\not = 2$ in $\Lambda^j$ is represented in $\mathcal{T}^j$ by an edge $E^j$ with edge length $L$ corresponding to the total length of the original path and edge radius $r_o$ as the average of the path radii.
+For the few (2) nodes in $\Lambda$ with degree $>3$; i.e.~junctions where a branch splits into more than two downstream branches, only the two longest downstream paths are included in $\mathcal{T}^j$.
+
+For each subtree $\mathcal{T}^j = \cup_n E_n^j$, we compute the
+time-averaged downstream flow rate $\langle Q'_n \rangle$ induced by
+the vascular wall motion for each edge $n$ via~\cite[eq.~(5),
+  (34)]{gjerde2023directional}. We next uniquely assign the flow rate
+$\langle Q'_n \rangle$ to each of the branches $\Lambda_i$ that form
+the path $E_n^j$, thus yielding $\langle Q'_i \rangle$ for each
+perivascular segment $\Lambda_i$ (except those ignored above, for
+which we set a flow rate of zero). Finally, we define the mean
+longitudinal perivascular velocity $\langle u^{\rm PVS}_i \rangle$ by
+dividing the flow rate $\langle Q'_i \rangle$ by the cross-section
+area $A_i = \pi (r_{e_i}^2 - r_{o_i}^2) = \pi (\beta_i^2 - 1) r_{o_i}^2$
 

### 2024-11-12 paper/supplementary_main.tex
Edit notation
Former-commit-id: 5b20f535011df2fea88ac2135027bc51cd2ba7e4

@@
 
-In terms of domains, interfaces and boundaries, we consider the
-parenchyma $\Omega_{\rm PAR} \subset \R^3$ and CSF spaces $\Omega_{\rm
-  CSF} \subset \R^3$ with $\Omega = \Omega_{\rm PAR} \cup \Omega_{\rm
-  CSF}$. The interface between the parenchyma and CSF spaces is
-decomposed
-
-\subsection*{Notation and left-overs}
-
-\paragraph*{To be moved: notation and labels}
-The thin pia membrane encloses the brain, and we hence denote the CSF-brain interface with $\Gamma_{\rm pia}$, except for the separately marked surface of the lateral ventricles $\Gamma_{\rm LV}$ (thus $\partial \Omega_{\rm CSF} \bigcap \partial \Omega_{\rm PAR} = \Gamma_{\rm LV} \cup \Gamma_{\rm Pia}$).
-
-We denote the different parts of the outer boundary of $\Omega_{\rm
-  CSF}$ with $\Gamma_{\mathrm{skull}}$ facing the dura and skull and
-$\Gamma_{\rm SSAS}$ representing the lower interface towards the
-spinal subarachnoid space (SSAS). $\Gamma_{\mathrm{skull}}$ is further
-subdivided into its lower and upper parts: $\Gamma_{\mathrm{skull^l}}$
-and $\Gamma_{\mathrm{skull^u}}$ (compare \Cref{fig:concept}). The
-boundary of the spinal cord is denoted by $\Gamma_{\rm{SC}}$, which is
-given by $\Gamma_{\rm{SC}} = \partial \Omega_{\rm PAR} \setminus
-\partial \Omega_{\rm CSF}$ .
+In terms of geometrical domains, we consider the parenchyma
+$\Omega_{\rm PAR} \subset \R^3$ and CSF spaces $\Omega_{\rm CSF}
+\subset \R^3$ with $\Omega = \Omega_{\rm PAR} \cup \Omega_{\rm CSF}$
+(see also \Cref{fig:intracranial_domains}). The interface between the
+parenchyma and CSF spaces is given by $\partial \Omega_{\rm PAR} \cap
+\partial \Omega_{\rm CSF}$ and we separate it as two parts: the
+surface of the lateral ventricles $\Gamma_{\rm LV}$ and the remaining
+pial interface $\Gamma_{\rm pia}$ (thus $\partial \Omega_{\rm CSF}
+\bigcap \partial \Omega_{\rm PAR} = \Gamma_{\rm LV} \cup \Gamma_{\rm
+  pia}$). The remaining, outer, boundary of $\Omega_{\rm CSF}$ is
+again separated into two parts: $\Gamma_{\rm SSAS}$ represents the
+lower interface towards the spinal subarachnoid space (SSAS), while
+$\Gamma_{\rm AM}$ is the outer interface towards the arachnoid and
+dura membranes. $\Gamma_{\rm AM}$ is further subdivided into its lower
+and upper parts: $\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$. The
+boundary of the spinal cord is denoted by $\Gamma_{\rm{SC}}$ and given
+by $\Gamma_{\rm{SC}} = \partial \Omega_{\rm PAR} \setminus \partial
+\Omega_{\rm CSF}$ .
 

### 2024-11-12 paper/supplementary_main.tex
Stopped for today.
Former-commit-id: e8071970fb905a277e20eef1c778c57f560e9cb6

@@
 \subset \R^3$ with $\Omega = \Omega_{\rm PAR} \cup \Omega_{\rm CSF}$
-(see also \Cref{fig:intracranial_domains}). The interface between the
-parenchyma and CSF spaces is given by $\partial \Omega_{\rm PAR} \cap
-\partial \Omega_{\rm CSF}$ and we separate it as two parts: the
-surface of the lateral ventricles $\Gamma_{\rm LV}$ and the remaining
-pial interface $\Gamma_{\rm pia}$ (thus $\partial \Omega_{\rm CSF}
-\bigcap \partial \Omega_{\rm PAR} = \Gamma_{\rm LV} \cup \Gamma_{\rm
-  pia}$). The remaining, outer, boundary of $\Omega_{\rm CSF}$ is
-again separated into two parts: $\Gamma_{\rm SSAS}$ represents the
-lower interface towards the spinal subarachnoid space (SSAS), while
-$\Gamma_{\rm AM}$ is the outer interface towards the arachnoid and
-dura membranes. $\Gamma_{\rm AM}$ is further subdivided into its lower
-and upper parts: $\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$. The
-boundary of the spinal cord is denoted by $\Gamma_{\rm{SC}}$ and given
-by $\Gamma_{\rm{SC}} = \partial \Omega_{\rm PAR} \setminus \partial
-\Omega_{\rm CSF}$ .
+(see also \Cref{fig:intracranial_domains}). The coordinates in these
+3D domains is given by $x$. The interface between the parenchyma and
+CSF spaces is given by $\partial \Omega_{\rm PAR} \cap \partial
+\Omega_{\rm CSF}$ and we separate it as two parts: the surface of the
+lateral ventricles $\Gamma_{\rm LV}$ and the remaining pial interface
+$\Gamma_{\rm pia}$ (thus $\partial \Omega_{\rm CSF} \bigcap \partial
+\Omega_{\rm PAR} = \Gamma_{\rm LV} \cup \Gamma_{\rm pia}$). The
+remaining, outer, boundary of $\Omega_{\rm CSF}$ is again separated
+into two parts: $\Gamma_{\rm SSAS}$ represents the lower interface
+towards the spinal subarachnoid space (SSAS), while $\Gamma_{\rm AM}$
+is the outer interface towards the arachnoid and dura
+membranes. $\Gamma_{\rm AM}$ is further subdivided into its lower and
+upper parts: $\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$. The boundary
+of the spinal cord is denoted by $\Gamma_{\rm{SC}}$ and given by
+$\Gamma_{\rm{SC}} = \partial \Omega_{\rm PAR} \setminus \partial
+\Omega_{\rm CSF}$.
 
-%% We denote the domain defined by the connected centerlines by
-%% $\Lambda$.The vascular domain is then represented as the union of
-%% cylindrical vessels of radius $R_1^i$ surrounding the centerlines
-%% $\Lambda^i$. Moreover, we consider the surrounding perivascular spaces
-%% as the union of annular cylinders of inner radius $R_1^i$ and outer
-%% radius $R_2^i = \beta R_1^i$ of width $R_2^i - R_1^i = (\beta - 1)
-%% R_1^i$.
+In addition, we consider two sets of perivascular networks represented
+as graphs: a periarterial network $\Lambda_a$ represented by the
+(connected) centerlines $\Lambda^i_{(a)}$ of the arterial tree, and a
+perivenous network $\Lambda_v$ corresponding to the centerlines
+$\Lambda^i_{(v)}$ of the veins. Recall that we represent the vascular
+domains as the union of cylindrical vessels of radius $R_1^i$
+surrounding the centerlines $\Lambda^i$. Moreover, we consider the
+surrounding perivascular spaces as the union of annular cylinders
+$\Omega^i$ of inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$
+and thus of width $R_2^i - R_1^i$. We denote the outer lateral surface
+of the periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$, respectively. We
+omit the subscript $a$ or $v$ when referring to any such network,
+vessel segment or perivascular outer surface. We assume that $\Gamma$
+is parametrized by the coordinate $s$, and with a minor abuse of
+notation, simply write $s \in \Lambda$ to represent the point ${\bm
+  \lambda}(s)$ on $\Lambda$ corresponding to $s$.
 
@@
 
-In this section, we describe the mathematical model~\eqref{eq:multi_transport} and associated conditions in further detail.
-
-Recall that the 3D concentration $c$ is defined locally in each subdomain of $\Omega$ and satisfies
+In this section, we describe the mathematical
+model~\eqref{eq:multi_transport}, the associated definitions,
+interface and boundary conditions in further detail. We refer to the
+main text for explicit parameter values and give the general, abstract
+form here. Recall that we model a concentration field $c = c(x, t)$
+for $x \in \Omega$, $t > 0$ defined in $\Omega_{\rm PAR}$ and
+$\Omega_{\rm CSF}$ separately. In each of these domains, $c$ satisfies
 \begin{equation}
-     \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma = f \quad \quad \mathrm{in} \quad \Omega_{\rm CSF} \cup \Omega_{\rm{PAR}}.
+  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma = f \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
   \label{eq:3d_pde}
 \end{equation}
+In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$).
+In the parenchyma, $\phi$ represents the extracellular space volume fraction.
+Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$.
+$\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivacular networks $\Lambda_a, \Lambda_v$.
+To summarize,
+\[
+\phi =  \begin{cases}
+  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
+  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
+  \end{cases}, \; 
+D = \begin{cases}
+  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
+  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
+  \quad \bm u  = \begin{cases}
+  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
+  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
+\end{cases}, \; 
+\xi = \begin{cases}
+  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
+  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
+\end{cases} . 
+\]
+In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp.
+$|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp.
+$\Omega_{\rm{CSF}}$).
+If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$  if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$ in which case the interaction with $\Omega_{\rm{CSF}}$ (resp.
+$\Omega_{\rm{PAR}}$) is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.
+
+\fixme{Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by}
+\begin{equation}
+  \overline{c}(s) = \frac{1}{A(s)} \int_{\Theta(s)} c 
+\end{equation}
+\fixme{where $\Theta(s)$ is a cross-section of the PVS $\Omega^i$ at $s$ and $A(s)$ is the area of $\Theta(s)$.} For both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the (Dirac delta) term $\delta_\Gamma$ is a distribution concentrated on the outer lateral surfaces of the PVSs.
+\discuss{More precisely, it is defined in terms of its action on any sufficiently smooth function $v : \Omega \rightarrow \R$ and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$ by}
+\begin{equation}
+  %\xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
+  \delta_\Gamma (v) = \int_{\Gamma} v = \sum_{i} \int_{\Gamma^i} v = \sum_i \int_{\Lambda^i} P \, \overline{v} = \int_{\Lambda} P  \, \overline{v} .
+\end{equation}
+where $P = P(s)$ is defined as the perimeter of each cross-section $\Theta(s)$ for $s \in \Lambda^i$.
+
+\mer{MER: Stopped here.
+Feel free to check the above.
+I thought the overline was defined as the cross-section perimeter average and not the cross-section average.
+This is more compatible with the definition of $\delta_{\Gamma}$ as well.}
+
+We refer to the reference~\cite{masri2023modelling} for a complete model derivation and analysis.
+
 The above is supplemented with the following boundary conditions. 
@@
 
-% \mar{I think we can simplify the BCs here a bit, mostly since $\bm u \cdot \bm n$ is zero everywhere except for $\Gamma_{\rm LV}$ and $\Gamma_{\rm skull}$.
-So I suggest we remove all terms related to this?
-And we probably also don't need to split $\Gamma_{SC}$ in $\Gamma_{SC}^{in}$ and  $\Gamma_{SC}^{out}$?}
+%When convenient, we also denote $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$
+%and $c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$.
 
@@
 \end{equation}
-
-Here $\phi$ is the fluid volume fraction (or porosity) of the brain (where $\phi \ll 1$), of the SAS (where $\phi = 1$) and of the PVS (where $0 \ll \phi \leq 1$); $D$ is the effective diffusion coefficient of the relevant solute in the respective media~\cite{sykova2008diffusion} which takes different values over the SAS and brain,  $\bm u$ and $\hat u$ are convective velocity fields in $\Omega$ and $\Lambda$ respectively, and $\xi$ is a transfer coefficient across the PVS-CSF and the PVS-Brain interface
-
-The notation $\overline{c}$ denotes the lateral average of the concentration over the outer perivascular boundary \footnote{
-The lateral average is defined as 
-\[ \overline{c}(s) = \frac{1}{A(s)} \int_{\Theta(s)} c ,\]
-where $\Theta(s)$ is a cross-section of the PVS at the point on $\bm{\lambda}(s)$ on $\Lambda$ where $\bm \lambda$ is a given parameterization of $\Lambda$ and $A(s)$ is the area of $\Theta(s)$. 
-}.
-The (Dirac delta) term $\delta_\Gamma$ is a source term concentrated on the outer boundary of the PVS (the union of the annular domains surrounding $\Lambda^i$) that models the solute exchange between the PVSs and their surroundings \footnote{The term $\delta_{\Gamma}$ is given for any function $v$, see \Cref{sec:details_numerical_method} for more details
-\[\xi  (\overline{c} - \hat c ) \delta_\Gamma (v) 
-= \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} 
-\]
-}.
-
-\[
-%\phi =  \begin{cases}
-%   1  & \mathrm{in} \;  \Omega_{\rm{csf}} \\ 
-%   \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
-%   \end{cases}, \; 
-%D= \begin{cases}
-%   D_{\rm{csf}} & \mathrm{in} \;  \Omega_{\rm{csf}} \\ 
-%   D_{\rm{PAR}} (1+ R) & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
-%   \quad \bm u  = \begin{cases}
-%   \bm{u}_{\rm{csf}} & \mathrm{in} \; \Omega_{\rm{csf}} \\ 
-%   \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
-%\end{cases}, \; 
-\xi = \begin{cases}
-\xi_{\rm{PVS}, \rm{PAR}} & \mathrm{if} |\Lambda^i \cap \Omega_{\rm{PAR}}| \neq 0 \\ 
- \xi_{\rm{PVS}, \rm{csf}} & \mathrm{if} |\Lambda^i \cap \Omega_{\rm{csf}}| \neq 0
-\end{cases} . 
-\]
-In the above,  $|\Lambda^i \cap \Omega_{\rm{PAR}}|$ (resp.
-$|\Lambda^i \cap \Omega_{\rm{csf}}|$) is nonzero whenever the surrounding PVS around the line $\Lambda^i$ intersects $\Omega_{\rm{PAR}}$ (resp.
-$\Omega_{\rm{csf}}$).
-If the surrounding PVS intersects both, then we set $\xi = \xi_{\mathrm{PVS}, \rm{PAR}}$ (resp $\xi_{\mathrm{PVS}, \rm{CSF}}$) if the PVS lies mostly ($80$ percent) in $\Omega_{\rm{PAR}}$ (resp.
-$\Omega_{\rm{CSF}}$) and the interaction with $\Omega_{\rm{PVS}}$ (resp.
-$\Omega_{\rm{CSF}}$ is ignored.
 
@@
 \end{subequations}
-
 Here, $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the CSF viscosity, and $\bm u_{\rm in}$ is the steady CSF flow across the lateral ventricle wall, see \Cref{tab:pvs:parameters}.

### 2024-11-13 paper/supplementary_main.tex
Updates from Overleaf
Former-commit-id: b5385f5a5991d9c255a83e9fcca9525d22a4afbc

@@
 \begin{equation}
-  \overline{c}(s) = \frac{1}{A(s)} \int_{\Theta(s)} c 
+  \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta(s)} c 
 \end{equation}
-\fixme{where $\Theta(s)$ is a cross-section of the PVS $\Omega^i$ at $s$ and $A(s)$ is the area of $\Theta(s)$.} For both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the (Dirac delta) term $\delta_\Gamma$ is a distribution concentrated on the outer lateral surfaces of the PVSs.
+\rami{where $\partial \Theta(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the area of $\Theta(s)$.} For both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the (Dirac delta) term $\delta_\Gamma$ is a distribution concentrated on the outer lateral surfaces of the PVSs.
 \discuss{More precisely, it is defined in terms of its action on any sufficiently smooth function $v : \Omega \rightarrow \R$ and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$ by}

### 2024-11-13 paper/supplementary_main.tex
More work on supplementary description of 3D-1D
Former-commit-id: 25f33f36b50624376d77049c2fed17f7f41f272d [formerly 25f33f36b50624376d77049c2fed17f7f41f272d [formerly 581f10dae6a3590eed2499a0e0d71e5f004fbea8]]
Former-commit-id: b831c57a115c61eadc6c5a60de64c7351765073b
Former-commit-id: f16d224d3238198aa07b629ebedd87897205a6ab

--- 2024-11-12 paper/supplementary_main.tex
@@
 (see also \Cref{fig:intracranial_domains}). The coordinates in these
-3D domains is given by $x$. The interface between the parenchyma and
+3D domains is denoted by $x$. The interface between the parenchyma and
 CSF spaces is given by $\partial \Omega_{\rm PAR} \cap \partial
@@
 upper parts: $\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$. The boundary
-of the spinal cord is denoted by $\Gamma_{\rm{SC}}$ and given by
+towards the spinal cord is denoted by $\Gamma_{\rm{SC}}$ and given by
 $\Gamma_{\rm{SC}} = \partial \Omega_{\rm PAR} \setminus \partial
@@
 (connected) centerlines $\Lambda^i_{(a)}$ of the arterial tree, and a
-perivenous network $\Lambda_v$ corresponding to the centerlines
+perivenous network $\Lambda_v$ associated with the centerlines
 $\Lambda^i_{(v)}$ of the veins. Recall that we represent the vascular
@@
 and thus of width $R_2^i - R_1^i$. We denote the outer lateral surface
-of the periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$, respectively. We
-omit the subscript $a$ or $v$ when referring to any such network,
-vessel segment or perivascular outer surface. We assume that $\Gamma$
-is parametrized by the coordinate $s$, and with a minor abuse of
-notation, simply write $s \in \Lambda$ to represent the point ${\bm
-  \lambda}(s)$ on $\Lambda$ corresponding to $s$.
+of the periarterial and perivenous spaces by $\Gamma_a$ and
+$\Gamma_v$, respectively. We omit the subscript $a$ or $v$ when
+referring to any such network, vessel segment or perivascular outer
+surface. We assume that $\Lambda$ is parametrized by the coordinate
+$s$, and with a minor abuse of notation, simply write $s \in \Lambda$
+to represent the point ${\bm \lambda}(s)$ on $\Lambda$ corresponding
+to $s$.
 
@@
 interface and boundary conditions in further detail. We refer to the
-main text for explicit parameter values and give the general, abstract
+main text for explicit parameter values while giving the general, abstract
 form here. Recall that we model a concentration field $c = c(x, t)$
-for $x \in \Omega$, $t > 0$ defined in $\Omega_{\rm PAR}$ and
+for $x \in \Omega$ and $t > 0$ defined in $\Omega_{\rm PAR}$ and
 $\Omega_{\rm CSF}$ separately. In each of these domains, $c$ satisfies
@@
 $\Omega_{\rm{CSF}}$).
-If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$  if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$ in which case the interaction with $\Omega_{\rm{CSF}}$ (resp.
-$\Omega_{\rm{PAR}}$) is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.
+If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.
 
-\fixme{Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by}
+Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by
 \begin{equation}
-  \overline{c}(s) = \frac{1}{A(s)} \int_{\Theta(s)} c 
+  \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta_2^i(s)} c 
 \end{equation}
-\fixme{where $\Theta(s)$ is a cross-section of the PVS $\Omega^i$ at $s$ and $A(s)$ is the area of $\Theta(s)$.} For both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the (Dirac delta) term $\delta_\Gamma$ is a distribution concentrated on the outer lateral surfaces of the PVSs.
-\discuss{More precisely, it is defined in terms of its action on any sufficiently smooth function $v : \Omega \rightarrow \R$ and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$ by}
+where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$.
+Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the (Dirac delta) term $\delta_\Gamma$ is a distribution concentrated on the outer lateral surfaces of the PVSs.
+More precisely, it is defined in terms of its \mer{integrated} action on any sufficiently smooth function $v : \Omega \rightarrow \R$ and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$ such that
 \begin{equation}
-  %\xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
-  \delta_\Gamma (v) = \int_{\Gamma} v = \sum_{i} \int_{\Gamma^i} v = \sum_i \int_{\Lambda^i} P \, \overline{v} = \int_{\Lambda} P  \, \overline{v} .
+  \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
+  \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
 \end{equation}
-where $P = P(s)$ is defined as the perimeter of each cross-section $\Theta(s)$ for $s \in \Lambda^i$.
+under the assumption that $\xi$ is constant on each $\partial
+\Theta_2^i$.
+%% , since:
+%% \begin{multline}
+%%   \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
+%%   \int_{\Gamma} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
+%%   \sum_i \int_{\Gamma^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = 
+%%   \sum_i \int_{\Lambda^i} \int_{\partial \Theta_2^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \\
+%%   \sum_i \int_{\Lambda^i} \left (  \xi (\overline{c} - \hat c ) \int_{\partial \Theta_2^i} \delta_\Gamma (v) \right ) =
+%%   \sum_i \int_{\Lambda^i} P \xi (\overline{c} - \hat c ) \overline{v} =
+%%   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v} .
+%% \end{multline}
+We refer to the reference~\cite{masri2023modelling} for a complete model derivation and analysis.
+%% \mer{MER: The previous was:
+%% \begin{equation}
+%%   \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
+%% \end{equation}
+%% }
 
-\mer{MER: Stopped here.
-Feel free to check the above.
-I thought the overline was defined as the cross-section perimeter average and not the cross-section average.
-This is more compatible with the definition of $\delta_{\Gamma}$ as well.}
-
-We refer to the reference~\cite{masri2023modelling} for a complete model derivation and analysis.
+In addtion, we model the concentration $\hat{c} = \hat{c}(s, t)$ in
+the periarterial and perivenous networks ($s \in \Lambda_a, s \in
+\Lambda_v$) which represents the concentration in the perivascular
+space averaged over each perivascular cross-section and satisfies:
+\begin{equation}
+\partial_t (A  \hat c) - \partial_s(\hat D A \partial_s ( \hat c)) +\partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0  \qquad \quad \mathrm{in} \quad  \Lambda_a, \Lambda_v ,
+\end{equation}
+where $A = A(s)$ is defined for each $s \in \Lambda^i$ and each
+centerline $\Lambda^i$ with PVS $\Omega^i$ as the area of the PVS
+cross-section $\Theta^i(s)$.
 

### 2024-11-13 paper/supplementary_main.tex
Updates, still WIP.
Former-commit-id: 2f4a6b4897e53e79fdcd9489a6b92ad8b364a5c4 [formerly 2f4a6b4897e53e79fdcd9489a6b92ad8b364a5c4 [formerly bd39ac695ec46dc46e73a39431bad712a1273d11]]
Former-commit-id: 2c0e453cab63dcc64971d3615f8a0264574b7ab4
Former-commit-id: d40b4ce2e1dd33f3ebb5a6f5b16573d9a1175119

@@
 
-In addition, we consider two sets of perivascular networks represented
-as graphs: a periarterial network $\Lambda_a$ represented by the
-(connected) centerlines $\Lambda^i_{(a)}$ of the arterial tree, and a
-perivenous network $\Lambda_v$ associated with the centerlines
-$\Lambda^i_{(v)}$ of the veins. Recall that we represent the vascular
-domains as the union of cylindrical vessels of radius $R_1^i$
-surrounding the centerlines $\Lambda^i$. Moreover, we consider the
-surrounding perivascular spaces as the union of annular cylinders
-$\Omega^i$ of inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$
-and thus of width $R_2^i - R_1^i$. We denote the outer lateral surface
-of the periarterial and perivenous spaces by $\Gamma_a$ and
-$\Gamma_v$, respectively. We omit the subscript $a$ or $v$ when
-referring to any such network, vessel segment or perivascular outer
-surface. We assume that $\Lambda$ is parametrized by the coordinate
-$s$, and with a minor abuse of notation, simply write $s \in \Lambda$
-to represent the point ${\bm \lambda}(s)$ on $\Lambda$ corresponding
-to $s$.
+In addition, we consider two sets of perivascular networks: a
+periarterial network $\Lambda_a$ represented by the (connected)
+centerlines $\Lambda^i_{(a)}$ of the arterial tree, and a perivenous
+network $\Lambda_v$ associated with the centerlines $\Lambda^i_{(v)}$
+of the veins. Recall that we represent the vascular domains as the
+union of cylindrical vessels of radius $R_1^i$ surrounding the
+centerlines $\Lambda^i$. Moreover, we consider the surrounding
+perivascular spaces as the union of annular cylinders $\Omega^i$ of
+inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$ and thus of
+width $R_2^i - R_1^i$. We denote the outer lateral surface of the
+periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$,
+respectively. We omit the subscript $a$ or $v$ when referring to any
+such network, vessel segment or perivascular outer surface. We assume
+that $\Lambda$ is parametrized by the coordinate $s$, and with a minor
+abuse of notation, simply write $s \in \Lambda$ to represent the point
+${\bm \lambda}(s)$ on $\Lambda$ corresponding to $s$. Finally, we view
+each perivascular network both as a geometric domain and as a directed
+graph with the centerlines $\{ \Lambda^i \}$ as oriented edges and the
+connections as nodes $\mathcal{V}$.
 
@@
 interface and boundary conditions in further detail. We refer to the
-main text for explicit parameter values while giving the general, abstract
-form here. Recall that we model a concentration field $c = c(x, t)$
-for $x \in \Omega$ and $t > 0$ defined in $\Omega_{\rm PAR}$ and
-$\Omega_{\rm CSF}$ separately. In each of these domains, $c$ satisfies
+main text for explicit parameter values while giving the general,
+abstract form here, and to the reference~\cite{masri2023modelling} for
+the derivation and analysis of this 3D-1D model. Recall that we model
+a concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
+defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
+each of these domains, $c$ satisfies
 \begin{equation}
@@
 \begin{equation}
-  \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
+  \mer{\int_{\Omega}} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
@@
 %% \end{multline}
-We refer to the reference~\cite{masri2023modelling} for a complete model derivation and analysis.
 %% \mer{MER: The previous was:
@@
 
-In addtion, we model the concentration $\hat{c} = \hat{c}(s, t)$ in
-the periarterial and perivenous networks ($s \in \Lambda_a, s \in
-\Lambda_v$) which represents the concentration in the perivascular
-space averaged over each perivascular cross-section and satisfies:
+The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
+perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
+in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
+space averaged over each perivascular cross-section and is governed by 
 \begin{equation}
-\partial_t (A  \hat c) - \partial_s(\hat D A \partial_s ( \hat c)) +\partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0  \qquad \quad \mathrm{in} \quad  \Lambda_a, \Lambda_v ,
+  \partial_t (A  \hat c) - \partial_s(\hat D A \partial_s ( \hat c)) +\partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
+  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
+\label{eq:1d_pde}
 \end{equation}
-where $A = A(s)$ is defined for each $s \in \Lambda^i$ and each
-centerline $\Lambda^i$ with PVS $\Omega^i$ as the area of the PVS
-cross-section $\Theta^i(s)$.
+where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
+of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
+PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
+associated PVS $\Omega^i$, the area of the cross-section
+$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
+and $\hat{u}$ is a convective velocity representing an (average) CSF
+flow velocity in the axial direction of the perivascular spaces. To
+complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
+internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
+boundary conditions at the end nodes. To this end, define the set of
+internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
+that are connected to two or more edges and the set of end nodes as
+$\mathcal{T}$, write $\hat{c}_i = \hat{c}|_{\Lambda^i}$, and let
+$\mathcal{E}(y)$ denote the set of (one, two or three) edges in
+$\{\Lambda^i\}$ sharing the node $y$. To impose continuity of the
+concentrations, we set that the concentrations when viewed from each
+edge must match at nodes:
+\begin{equation}
+  \hat{c}_i (y) = \hat{c}_y \qquad \foralls \Lambda^i \in \mathcal{E}(y),
+  \quad \foralls y \in \mathcal{B} .
+\end{equation}
+Moreover, to ensure conservation of flux, we set that the flux going
+in and out at nodes should add to zero, or more precisely, that for
+each $y \in \mathcal{B}$:
+\begin{equation}
+  \sum_{i \in \mathcal{E}(y)} (\hat{D} A_i \partial_s \hat{c}_i (y)  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n_i(\mathsf{v}) = 0
+  \label{eq:flux_conservation}
+\end{equation}
+
+For a given edge $\Lambda^i= (\mathsf{v}_\mathrm{in}^i, \mathsf{v}_{\mathrm{out}}^{i})$, we define the function $n_i: \mathcal{V} \rightarrow \{-1,0,1\}$ with $$
+n_i(\mathsf{v}_{\mathrm{in}}^i) = 1 , \,\, n _i (\mathsf{v}_{\mathrm{out}}^i) = -1 \,\, \mathrm{and} \,\,\, n_i(\mathsf{v}) = 0, \quad \forall \mathsf{v} \in 
+\mathcal{V} \,  \backslash \,  \{ \mathsf{v}_{\mathrm{in}}^\mathsf e,\mathsf{v}_{\mathrm{out}}^\mathsf e \}.  
+$$
+The collection of bifurcation points is denoted by $\mathcal{B} = \{ \mathsf{v} \in \mathcal{V}, \,\,  \mathrm{card}(\mathcal{E}(\mathsf{v})) \geq  3\}$. For any $\mathsf{v} \in \mathcal{B}$, we have 
+\begin{equation}
+\sum_{ i \in \mathcal{E}(\mathsf{v})} (\hat D A_i \phi \partial_s \hat{u}_i(\mathsf{v})  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n_i(\mathsf{v}) = 0  \label{eq:flux_conservation}
+\end{equation}
+
+\paragraph{Boundary and interface conditions}
+
+On the interface $\Gamma_{\mathrm{pia}}$ between the CSF and brain, we prescribe the following interface condition modeling a semi--permeable interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and $c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
+\begin{subequations}
+\begin{alignat}{2}
+(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{PAR}} &= - (- D \nabla c_{\rm{CSF}} + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n}_{\rm{PAR}} &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}} ,  \\
+(-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{PAR}} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad  \pia \cup \Gamma_{\mathrm{LV}} .
+\end{alignat} 
+\end{subequations}
+Here, $\beta_{\mathrm{pia}}$ is a permeability constant between the
+brain and the CSF. \mer{MER: Is the inclusion of $\Gamma_{\rm LV}$
+  correct in the equation here? If so, text needs minor edits.}
+
+For the perivascular network, the boundary nodes are collected into
+$\partial \Lambda$. We assume no tracer outflow on all boundary nodes
+and prescribe a homogeneous Neumann condition:
+\begin{equation}
+    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \partial \Lambda. 
+\end{equation}
+
+%% Both $\Omega$ and $\Lambda$ are fixed in time and we assume
+%% constant-in- time radii (despite varying radii to drive peristaltic
+%% flow, see \cref{sec:sup:peristalsis}).
 
@@
 
-At bifurcation nodes in $\Gamma_a, \Gamma_v$, we impose flux conservation and continuity of the concentration $\hat c$.
-We refer to \eqref{eq:flux_conservation} in \Cref{sec:details_numerical_method} for more details.
-For the perivascular network, the boundary nodes are collected into $\partial \Lambda$.
-We assume no tracer outflow on all boundary nodes and prescribe a homogeneous Neumann condition:
-\begin{equation}
-    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \partial \Lambda. 
-\end{equation}
-
-Further, $P$ and $A$ are the perimeter and area of the PVS which can vary along the centerline network, respectively.
-Finally, we remark that $\hat u$ takes different values on each $\Lambda_i$.
-Both $\Omega$ and $\Lambda$ are fixed in time and we assume constant-in- time radii (despite varying radii to drive peristaltic flow, see \cref{sec:sup:peristalsis}).
-Finally, we note that the model assumes that there is no interaction between the inner perivascular boundary and the blood vessel.
-
 \paragraph*{Interface conditions}
-On the interface $\Gamma_{\mathrm{pia}}$ between the CSF and brain, we prescribe the following condition modeling a semi--permeable pial interface 
-\begin{alignat}{2}
-(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{PAR}} &= - (- D \nabla c_{\rm{CSF}} + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n}_{\rm{PAR}} &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}} ,  \\
-(-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{PAR}} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad  \pia \cup \Gamma_{\mathrm{LV}} .
-\end{alignat} 
-Here, $\beta_{\mathrm{pia}}$ is a permeability constant between the brain and the CSF:
 
@@
 We do not explicitly model convection within the brain parenchyma $\Omega_{\rm PAR}$, and thus set ${\bm u} = {\bm u}_{\rm PAR} = 0$ there.
-
-\subsubsection*{Dispersion in the CSF spaces}
-
-In addition, CSF pulsates in the SAS and ventricular system with each
-cardiac and respiratory cycle.
-
-Taking the large spatial differences of pulsatile CSF flow into account, we compute the local diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac induced CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}.
-
-Sharp et al \cite{keith2019dispersion} estimated the enhancement of diffusion in the spinal subarachnoid space due to dispersive effects to average 5.8 times that of purely molecular diffusion with a theoretical maximum of $10^6$.
-
-Similar to the case of CSF flow driven by production introduced in section \ref{sec:csf_fluid_vel}, we employ a Stokes flow model to predict CSF velocities at peak arterial dilation.
-In particular, we enforce a uniformly distributed inflow of $10$\,ml/s and $2$\,ml/s across the outer skull ($\Gamma_{\rm skull}$) and lateral ventricular surfaces ($\Gamma_{\rm LV}$) respectively, mimicking the effect of reduced CSF volume due to blood inflow.
-We prescribe a no-slip boundary condition on the pial surface $\Gamma_{\rm Pia}$ and zero-traction on the spinal boundary $\Gamma_{\rm SSAS}$, making the spinal compartment the only outflow route of CSF.
-Thus, we compute the cardiac-driven flow field by augmenting the Stokes flow model (eq.
-\ref{eq:momnetum_equation} and \ref{eq:divergence_equation}) with the following boundary conditions:
-
-\begin{equation}\label{eq:cardiac_flow_bcs}
-    \bm u = 0 \text{ on } \Gamma_{\rm Pia}; \quad 
-    \mu \nabla \bm u \cdot \bm n - p \bm n = 0 \text{ on } \Gamma_{\rm SSAS}; \quad 
-    \bm u = \frac{0.31 \cdot \bm n}{|\Gamma_{\rm LV}|} \,\text{ml/s} \text{ on } \Gamma_{\rm LV}; \quad
-    \bm u = \frac{8 \cdot \bm n}{|\Gamma_{\rm skull}|} \,\text{ml/s} \text{ on } \Gamma_{\rm skull}.
-\end{equation}
-
-The obtained pressure field only account for viscous forces, albeit cardiac induced CSF flow is known to be unsteady and substantially impacted by transient inertial forces.
-Assuming an oscillatory flow with the heartbeat frequency $\omega = 2 \pi$, a mean gap width of the CSF-filled spaces of $2h=3$\,mm and CSF density $\rho=10^3\,\text{kg/m}^3$, the Womersley number $\alpha$ becomes $\alpha^2 = \frac{h^2 \omega \rho}{\mu} = 20.2$, which is unsteady.
-Based on theoretical considerations on the ratio of oscillatory flow to steady flow impedances in a tube, we account for inertial forces in the pulsatile flow by upscaling the computed viscous pressure field by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter.
-4.3).
-
-Further assuming unsteady dispersion, we follow Sharp et al \cite{keith2019dispersion} and calculate the enhancement factor $R$ from the non-dimensionalized pressure gradient $dP=\frac{|\nabla p|}{\omega \mu / h}$ as $R\approx dP^2 / \alpha^3$.
-Combining the computed pressure gradient with this scaling law yields a pointwise approximation of the dispersion factor.
-Finally, we account for the non-local nature of the dispersion mechanism by applying a smoothing technique to obtain a spatially varying, but locally averaged diffusion enhancement factor $R$ (see the visualization in \cref{fig:visualize_R} and \cref{sec:dispersion_details} for additional details).
 
@@
 \end{multline}
-For the 1D concentration $\hat c$, we recall that $\hat c = \hat c_i $ on each segment of the network $\Lambda^i$.
-We enforce continuity of the concentration on the bifurcation points and conservation of fluxes.
-We let $\mathcal{V}$ denote the set of all vertices in $\Lambda$ and $
-\mathcal{E}(\mathsf{v})$ denote the set of segments sharing a vertex $\mathsf{v}$.
-For a given edge $\Lambda^i= (\mathsf{v}_\mathrm{in}^i, \mathsf{v}_{\mathrm{out}}^{i})$, we define the function $n_i: \mathcal{V} \rightarrow \{-1,0,1\}$ with $$
-n_i(\mathsf{v}_{\mathrm{in}}^i) = 1 , \,\, n _i (\mathsf{v}_{\mathrm{out}}^i) = -1 \,\, \mathrm{and} \,\,\, n_i(\mathsf{v}) = 0, \quad \forall \mathsf{v} \in 
-\mathcal{V} \,  \backslash \,  \{ \mathsf{v}_{\mathrm{in}}^\mathsf e,\mathsf{v}_{\mathrm{out}}^\mathsf e \}.  
-$$
-The collection of bifurcation points is denoted by $\mathcal{B} = \{ \mathsf{v} \in \mathcal{V}, \,\,  \mathrm{card}(\mathcal{E}(\mathsf{v})) \geq  3\}$. For any $\mathsf{v} \in \mathcal{B}$, we have 
-\begin{equation}
-\sum_{ i \in \mathcal{E}(\mathsf{v})} (\hat D A_i \phi \partial_s \hat{u}_i(\mathsf{v})  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n_i(\mathsf{v}) = 0  \label{eq:flux_conservation}
-\end{equation}
 For approximating the solution of the 1D equations in the network, we use the space 
@@
 
-\subsection{Cardiac Dispersion Factor Estimation}\label{sec:dispersion_details}
+\subsection{Cardiac Dispersion Factor Estimation}
+
+\subsubsection*{Dispersion in the CSF spaces}
+
+In addition, CSF pulsates in the SAS and ventricular system with each
+cardiac and respiratory cycle.
+
+Taking the large spatial differences of pulsatile CSF flow into account, we compute the local diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac induced CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}.
+
+Sharp et al \cite{keith2019dispersion} estimated the enhancement of diffusion in the spinal subarachnoid space due to dispersive effects to average 5.8 times that of purely molecular diffusion with a theoretical maximum of $10^6$.
+
+Similar to the case of CSF flow driven by production introduced in section \ref{sec:csf_fluid_vel}, we employ a Stokes flow model to predict CSF velocities at peak arterial dilation.
+In particular, we enforce a uniformly distributed inflow of $10$\,ml/s and $2$\,ml/s across the outer skull ($\Gamma_{\rm skull}$) and lateral ventricular surfaces ($\Gamma_{\rm LV}$) respectively, mimicking the effect of reduced CSF volume due to blood inflow.
+We prescribe a no-slip boundary condition on the pial surface $\Gamma_{\rm Pia}$ and zero-traction on the spinal boundary $\Gamma_{\rm SSAS}$, making the spinal compartment the only outflow route of CSF.
+Thus, we compute the cardiac-driven flow field by augmenting the Stokes flow model (eq.
+\ref{eq:momnetum_equation} and \ref{eq:divergence_equation}) with the following boundary conditions:
+
+\begin{equation}\label{eq:cardiac_flow_bcs}
+    \bm u = 0 \text{ on } \Gamma_{\rm Pia}; \quad 
+    \mu \nabla \bm u \cdot \bm n - p \bm n = 0 \text{ on } \Gamma_{\rm SSAS}; \quad 
+    \bm u = \frac{0.31 \cdot \bm n}{|\Gamma_{\rm LV}|} \,\text{ml/s} \text{ on } \Gamma_{\rm LV}; \quad
+    \bm u = \frac{8 \cdot \bm n}{|\Gamma_{\rm skull}|} \,\text{ml/s} \text{ on } \Gamma_{\rm skull}.
+\end{equation}
+
+The obtained pressure field only account for viscous forces, albeit cardiac induced CSF flow is known to be unsteady and substantially impacted by transient inertial forces.
+Assuming an oscillatory flow with the heartbeat frequency $\omega = 2 \pi$, a mean gap width of the CSF-filled spaces of $2h=3$\,mm and CSF density $\rho=10^3\,\text{kg/m}^3$, the Womersley number $\alpha$ becomes $\alpha^2 = \frac{h^2 \omega \rho}{\mu} = 20.2$, which is unsteady.
+Based on theoretical considerations on the ratio of oscillatory flow to steady flow impedances in a tube, we account for inertial forces in the pulsatile flow by upscaling the computed viscous pressure field by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter.
+4.3).
+
+Further assuming unsteady dispersion, we follow Sharp et al \cite{keith2019dispersion} and calculate the enhancement factor $R$ from the non-dimensionalized pressure gradient $dP=\frac{|\nabla p|}{\omega \mu / h}$ as $R\approx dP^2 / \alpha^3$.
+Combining the computed pressure gradient with this scaling law yields a pointwise approximation of the dispersion factor.
+Finally, we account for the non-local nature of the dispersion mechanism by applying a smoothing technique to obtain a spatially varying, but locally averaged diffusion enhancement factor $R$ (see the visualization in \cref{fig:visualize_R} and \cref{sec:dispersion_details} for additional details).
+
+\subsubsection{Cardiac Dispersion Factor Estimation}\label{sec:dispersion_details}
 

### 2024-11-13 paper/supplementary_main.tex
More updates in Supplementary, describe bifurcation conditions.

@@
 \begin{equation}
-  \partial_t (A  \hat c) - \partial_s(\hat D A \partial_s ( \hat c)) +\partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
+  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s (\hat c) ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
   \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
@@
 that are connected to two or more edges and the set of end nodes as
-$\mathcal{T}$, write $\hat{c}_i = \hat{c}|_{\Lambda^i}$, and let
+$\mathcal{T}$, write $\hat{c}^i = \hat{c}|_{\Lambda^i}$, and let
 $\mathcal{E}(y)$ denote the set of (one, two or three) edges in
@@
 \begin{equation}
-  \hat{c}_i (y) = \hat{c}_y \qquad \foralls \Lambda^i \in \mathcal{E}(y),
+  \hat{c}^i (y) = \hat{c}_y \qquad \foralls \Lambda^i \in \mathcal{E}(y),
   \quad \foralls y \in \mathcal{B} .
@@
 \begin{equation}
-  \sum_{i \in \mathcal{E}(y)} (\hat{D} A_i \partial_s \hat{c}_i (y)  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n_i(\mathsf{v}) = 0
+  \sum_{i \in \mathcal{E}(y)}
+  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c^i}(y) \right ) n_i(y)
+  = 0
   \label{eq:flux_conservation}
 \end{equation}
+where the function $n_i : \mathcal{V} \rightarrow \{-1, 0, 1\}$
+defines an orientation of each vertex relative to the
+edges. Specifically, for an edge $\Lambda^i$ from $y_{\rm in}$ to $y_{\rm out}$
+\begin{equation}
+  n_i(y_{\rm in}) = 1, \quad
+  n_i(y_{\rm{out}}) = -1, \quad
+  n_i(y) = 0, \quad \forall y \in 
+  \mathcal{V} \
+  \,  \backslash \,  \{ y_{\rm in}, y_{\rm out} \}.  
+\end{equation}
 
-For a given edge $\Lambda^i= (\mathsf{v}_\mathrm{in}^i, \mathsf{v}_{\mathrm{out}}^{i})$, we define the function $n_i: \mathcal{V} \rightarrow \{-1,0,1\}$ with $$
-n_i(\mathsf{v}_{\mathrm{in}}^i) = 1 , \,\, n _i (\mathsf{v}_{\mathrm{out}}^i) = -1 \,\, \mathrm{and} \,\,\, n_i(\mathsf{v}) = 0, \quad \forall \mathsf{v} \in 
-\mathcal{V} \,  \backslash \,  \{ \mathsf{v}_{\mathrm{in}}^\mathsf e,\mathsf{v}_{\mathrm{out}}^\mathsf e \}.  
-$$
-The collection of bifurcation points is denoted by $\mathcal{B} = \{ \mathsf{v} \in \mathcal{V}, \,\,  \mathrm{card}(\mathcal{E}(\mathsf{v})) \geq  3\}$. For any $\mathsf{v} \in \mathcal{B}$, we have 
-\begin{equation}
-\sum_{ i \in \mathcal{E}(\mathsf{v})} (\hat D A_i \phi \partial_s \hat{u}_i(\mathsf{v})  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n_i(\mathsf{v}) = 0  \label{eq:flux_conservation}
-\end{equation}
+\mer{MER: Stopped here.}
+
+%% For a given edge $\Lambda^i= (\mathsf{v}_\mathrm{in}^i, \mathsf{v}_{\mathrm{out}}^{i})$, we define the function $n_i: \mathcal{V} \rightarrow \{-1,0,1\}$ with $$
+%% $$
+%% The collection of bifurcation points is denoted by $\mathcal{B} = \{ \mathsf{v} \in \mathcal{V}, \,\,  \mathrm{card}(\mathcal{E}(\mathsf{v})) \geq  3\}$.
+For any $\mathsf{v} \in \mathcal{B}$, we have
+%% \begin{equation}
+%% \sum_{ i \in \mathcal{E}(\mathsf{v})} (\hat D A_i \phi \partial_s \hat{u}_i(\mathsf{v})  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n_i(\mathsf{v}) = 0  \label{eq:flux_conservation}
+%% \end{equation}
 
@@
 In the above, note that $\bm u \cdot \bm n  \neq 0$ on $\Gamma_{\rm LV} \cup \Gamma_{\rm skull^u}$ and $\bm u \cdot \bm n = 0$ on $\Gamma_{\rm SSAS} \cup \Gamma_{\rm skull^l} \cup \Gamma_{\rm Pia}$.
-
-\paragraph*{Interface conditions}
 

### 2024-11-14 paper/supplementary_main.tex
Complete section A.1.2

@@
 where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$.
-Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the (Dirac delta) term $\delta_\Gamma$ is a distribution concentrated on the outer lateral surfaces of the PVSs.
-More precisely, it is defined in terms of its \mer{integrated} action on any sufficiently smooth function $v : \Omega \rightarrow \R$ and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$ such that
+Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the term $\delta_\Gamma$ is concentrated on the outer lateral surfaces of the PVSs, and defined in terms of its action on any sufficiently smooth function $v : \Omega \rightarrow \R$
+%and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$
+such that
 \begin{equation}
-  \mer{\int_{\Omega}} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
+  \langle \xi (\overline{c} - \hat c ) \delta_\Gamma, v \rangle =
   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
@@
 under the assumption that $\xi$ is constant on each $\partial
-\Theta_2^i$. 
+\Theta_2^i$, see also~\cite{masri2024discontinuous}.
 %% , since:
@@
   n_i(y_{\rm{out}}) = -1, \quad
-  n_i(y) = 0, \quad \forall y \in 
-  \mathcal{V} \
-  \,  \backslash \,  \{ y_{\rm in}, y_{\rm out} \}.  
+  n_i(y) = 0 \quad \foralls y \in 
+  \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} .  
 \end{equation}
-
-\mer{MER: Stopped here.}
 
@@
 
-\paragraph{Boundary and interface conditions}
+\paragraph{Interface conditions between the brain and CSF spaces}
 
-On the interface $\Gamma_{\mathrm{pia}}$ between the CSF and brain, we prescribe the following interface condition modeling a semi--permeable interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and $c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
+On the interface between the brain and the CSF spaces
+$\Gamma_{\mathrm{pia}} \cup \Gamma_{\rm LV}$, we prescribe the
+following interface conditions, which represent a semi--permeable
+interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and
+$c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
+%% \begin{subequations}
+%% \begin{alignat}{2}
+%% (-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} &= - (- D (\phi \nabla c_{\rm{CSF}}) + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n} &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\
+%% (-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,
+%% \end{alignat} 
+%% \end{subequations}
 \begin{subequations}
   \begin{alignat}{2}
-(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{PAR}} &= - (- D \nabla c_{\rm{CSF}} + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n}_{\rm{PAR}} &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}} ,  \\
-(-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{PAR}} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad  \pia \cup \Gamma_{\mathrm{LV}} .
+  (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n}
+  &= - (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm CSF}} \cdot \bm{n}
+  &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
+  (- D \nabla (\phi  c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
 \end{alignat} 
-\end{subequations}
-Here, $\beta_{\mathrm{pia}}$ is a permeability constant between the
-brain and the CSF. \mer{MER: Is the inclusion of $\Gamma_{\rm LV}$
-  correct in the equation here? If so, text needs minor edits.}
+  \label{eq:interface:3d}%
+\end{subequations}%
+where $\bm{n} = \bm{n}_{\rm PAR}$ is the normal vector field defined
+over the interface, outward-pointing when viewed from $\Omega_{\rm
+  PAR}$, and $\beta_{\mathrm{pia}} \geqslant 0$ is a permeability
+constant.
 
-For the perivascular network, the boundary nodes are collected into
-$\partial \Lambda$. We assume no tracer outflow on all boundary nodes
-and prescribe a homogeneous Neumann condition:
+\paragraph{Boundary conditions: outer SAS interfaces and perivascular ends}
+
+We supplement~\eqref{eq:3d_pde} and~\eqref{eq:interface:3d} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
+\begin{alignat}{2}
+(D \nabla (\phi c) - \bm u c ) \cdot \bm{n} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \Gamma_{\rm SSAS},  \\ 
+% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\Gamma_{\rm SAS}} &= 0 &&  \quad  \mathrm{on} \quad \Gamma_{\rm SAS}^{\mathrm{out}}  \\ 
+(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = \beta c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-U},  \\ 
+(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = 0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-L} \discuss{\cup \Gamma_{\rm SC}}  .
+\end{alignat} 
+where again $\bm n$ denotes the outward-pointing boundary normal.
+
+%In the above, note that $\bm u \cdot \bm n  \neq 0$ on $\Gamma_{\rm LV} \cup \Gamma_{\rm skull^u}$ and $\bm u \cdot \bm n = 0$ on $\Gamma_{\rm SSAS} \cup \Gamma_{\rm skull^l} \cup \Gamma_{\rm Pia}$.
+
+For the perivascular networks, we assume no molecular efflux at all
+end nodes $\mathcal{T}$ via a homogeneous Neumann condition
+augmenting~\eqref{eq:1d_pde}:
 \begin{equation}
-    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \partial \Lambda. 
+    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{T}.
 \end{equation}
@@
 
-The above is supplemented with the following boundary conditions. 
-\begin{alignat}{2}
-(D \nabla (\phi c) - \bm u c ) \cdot \bm{n}_{\rm SAS} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \spinal  \\ 
-% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\spinal} &= 0 &&  \quad  \mathrm{on} \quad \spinal^{\mathrm{out}}  \\ 
-(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\rm skull} & =   \beta c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm skull^{u}}  \\ 
-(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\rm skull} & =   0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm skull^{l}}  \\ 
-(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{Pia}} &= - (-D \nabla  c_{\rm{CSF}} + \bm u_{\rm CSF} c_{\rm CSF} ) \cdot \bm{n}_{\rm{Pia}} &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}}  \\
-(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{Pia}} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}}   .
-\end{alignat}
-
-%When convenient, we also denote $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$
-%and $c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$.
-
-In the above, note that $\bm u \cdot \bm n  \neq 0$ on $\Gamma_{\rm LV} \cup \Gamma_{\rm skull^u}$ and $\bm u \cdot \bm n = 0$ on $\Gamma_{\rm SSAS} \cup \Gamma_{\rm skull^l} \cup \Gamma_{\rm Pia}$.
-
-\paragraph*{Boundary conditions}
-On $\Gamma_{\mathrm{skull}}$, the interface towards the dura and skull (\Cref{fig:concept}), we consider a constant uptake (or efflux) rate, represented by the boundary condition 
-\begin{alignat}{2}
-(- D \nabla ( c_{\mathrm{CSF}}) + \bm u_{\mathrm{CSF}} c_{\mathrm{CSF}}) \cdot \bm{n}   & = \beta c_{\mathrm{CSF}}  && \quad \mathrm{on} \quad \Gamma_{\mathrm{skull}^{\rm u}},\\ 
-(-D \nabla c_{\rm CSF} + \bm u c_{\rm CSF}) \cdot \bm{n} & =  0   &&  \quad  \mathrm{on} \quad \Gamma_{\rm skull^{l}} , 
-\end{alignat}
-where $\beta$ represents a molecular outflow resistance. Here, we take $\beta = 10^{-4} \,\, \mathrm{mm}^2/\mathrm{s}$\cite{hornkjol2022csf}.
+\mer{MER: Stopped here.}
 
@@
 Here we provide details on the numerical method used to solve the coupled system.
-We are given a mesh $\mathcal{T}_h $, consisting of elements denoted by $E$, of $\Omega= \brain \cup \sas$.
+We are given a mesh $\mathcal{T}_h $, consisting of elements denoted by $E$, of $\Omega= \Omega_{\rm PAR} \cup \Omega_{\rm CSF}$.
 The collection of all interior facets to $\Omega_{\rm PAR}$ (resp.
@@
 The union of all interior facets $\mathcal{F}_i = \mathcal{F}_{h,b} \cup \mathcal{F}_{h,c}$.
-Facets lying on the pial interface are denoted by $\mathcal{F}_{\pia}$, on the skull by $\mathcal{F}_{\skull}$, and on the spinal canal by $\mathcal{F}_{\spinal}$.
-Note that $\mathcal{F}_i \cap \mathcal{F}_{\pia} = \emptyset $.
+Facets lying on the pial interface are denoted by $\mathcal{F}_{\Gamma_{\rm pia}}$, on the skull by $\mathcal{F}_{\Gamma_{\rm skull}}$, and on the spinal canal by $\mathcal{F}_{\Gamma_{\rm SAS}}$.
+Note that $\mathcal{F}_i \cap \mathcal{F}_{\Gamma_{\rm pia}} = \emptyset $.
 The union of all the facets is denoted by $\mathcal{F}_h$.
@@
 As boundary conditions, we impose the precomputed fluid pressure $p_{\rm CSF}$ on the networks root nodes:
-
 \begin{equation}
@@
 \begin{multline}
-\int_{\Omega}\frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v_h + a_h(c_h^n, v_h) + a_h^{\mathrm{up}}(c_h^n,v_h) + \int_{\Gamma_{\mathrm{pia}} } \beta_{\mathrm{pia}}[c_h^n][v_h]  \\ + \int_{\Gamma_{\rm skull^u}} \beta c_h^n v_h +  \int_{\Lambda} \xi  P (\overline{c_h^n} - \hat c_h^n) \overline{v_h}  = \int_{\Omega} f^n v_h + \int_{\spinal} \gin v_h , \quad \forall v_h \in V_h^k.
+\int_{\Omega}\frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v_h + a_h(c_h^n, v_h) + a_h^{\mathrm{up}}(c_h^n,v_h) + \int_{\Gamma_{\mathrm{pia}} } \beta_{\mathrm{pia}}[c_h^n][v_h]  \\ + \int_{\Gamma_{\rm skull^u}} \beta c_h^n v_h +  \int_{\Lambda} \xi  P (\overline{c_h^n} - \hat c_h^n) \overline{v_h}  = \int_{\Omega} f^n v_h + \int_{\Gamma_{\rm SAS}} \gin v_h , \quad \forall v_h \in V_h^k.
 \end{multline}

### 2024-11-18 paper/supplementary_main.tex
Updates from Overleaf

@@
 
-Further assuming unsteady dispersion, we follow Sharp et al \cite{keith2019dispersion} and calculate the enhancement factor $R$ from the non-dimensionalized pressure gradient $dP=\frac{|\nabla p|}{\omega \mu / h}$ as $R\approx dP^2 / \alpha^3$.
+Further assuming unsteady dispersion, we follow Sharp et al \cite{keith2019dispersion} and calculate the enhancement factor $R$ from the non-dimensionalized pressure gradient $dP=\frac{|\nabla p|}{\omega \mu / h}$ \rami{Can we refer to the exact equation here from Sharp et al ?
+i find that they write $P = (\partial \tilde p / \partial \tilde x) / (\rho \omega \mu_e/h)$ ...
+so i am a bit confused
+}as $R\approx dP^2 / \alpha^3$.
 Combining the computed pressure gradient with this scaling law yields a pointwise approximation of the dispersion factor.

### 2024-11-18 paper/supplementary_main.tex
Updates in CSF flow model description
Former-commit-id: 423b4a85d4efc35597c4c1cb144d9047fdea5bbb [formerly 4f0e343c77b5e54c0f7fef108d821346b20faa9e]
Former-commit-id: 8ac31944effe4346db0c8a30d960464fda95b132

@@
 We supplement~\eqref{eq:3d_pde} and~\eqref{eq:interface:3d} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
+\begin{subequations}
 \begin{alignat}{2}
@@
 (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = \beta c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-U},  \\ 
-(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = 0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-L} \discuss{\cup \Gamma_{\rm SC}}  .
+(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = 0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-L} \cup \Gamma_{\rm SC}  .
 \end{alignat} 
+\end{subequations}
 where again $\bm n$ denotes the outward-pointing boundary normal.
@@
 
-\mer{MER: Stopped here.}
+\subsubsection{Modelling CSF flow via the incompressible Stokes equations}
 
-\subsubsection{Incompressible Stokes flow of CSF in the SAS and ventricular system}
+The time-independent, incompressible Stokes equations model the flow
+of an incompressible Newtonian fluid at low Reynolds numbers, and read
+as follows: over the domain $\Omega_{\rm CSF} \subset \R^3$, find the
+velocity vector field $\bm u : \Omega_{\rm CSF} \rightarrow \R^3$ and
+the pressure $p : \Omega_{\rm CSF} \rightarrow \R$ such that
+\begin{subequations}
+  \begin{alignat}{2}
+    - \mu \Delta \bm u  + \nabla p &=  0 \quad && \text{ in } \Omega_{\rm CSF},
+    \label{eq:momnetum_equation} \\ 
+    \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF},
+    \label{eq:divergence_equation} .
+  \end{alignat}
+\end{subequations}
+In addition, we impose the following boundary conditions to model flow
+induced by CSF production in the choroid plexus with $\Gamma_{\rm AM-U}$ as the main efflux site:
+\begin{subequations}
+  \begin{alignat}{2}
+    \mu \nabla \bm u \cdot \bm{n} - p \bm n &= -R_0 ( \bm u \cdot \bm n ) \bm n
+    && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
+    \bm u &= 0 && \text{ on } \Gamma_{\mathrm{AM-L}}, \\
+    \bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
+    \bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
+    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm \tau = 0 \quad && \text{ on } \Gamma_{\rm{LV}} ,  
+  \end{alignat}
+\end{subequations}
+where $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$.
+We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:supp:dispersion}.
 
-We consider non-zero convection in the CSF space $u_{\rm CSF} \not = 0$ in \eqref{eq:multi_transport_3d} resulting from steady state CSF production.
+%% We consider non-zero convection in the CSF space $u_{\rm CSF} \not = 0$ in \eqref{eq:multi_transport_3d} resulting from steady state CSF production.
 Recall that $\Omega_{\mathrm{CSF}}$ represents the CSF space and that $\Gamma_{\mathrm{skull^{u/l}}}$ and $\Gamma_{\mathrm{pia}}$, $\Gamma_{\mathrm{LV}}$ represent the boundaries facing the upper and lower parts of the dura, the pia and the lateral ventricles, respectively.
 Assuming CSF production in the choroid plexus located at the lateral ventricle wall, we solve the  steady state Stokes equations for the velocity and the pressure $(\bm u_{\mathrm{CSF}}, p_{\mathrm{CSF}})$ in the CSF domain:
-\begin{subequations}
-    \begin{alignat}{2}
- - \mu \Delta \bm u_{\mathrm{CSF}} + \nabla p_{\rm{CSF}} & =  0 \quad && \mathrm{in} \,\,  \Omega_{\rm CSF}, \label{eq:momnetum_equation}  \\ 
- \nabla \cdot  \bm u_{\mathrm{CSF}} & = 0 \quad && \mathrm{in} \,\,   \Omega_{\rm CSF}, \label{eq:divergence_equation}  \\ 
-\mu \nabla \bm u_{\mathrm{CSF}} \cdot \bm{n} -  p \bm n  &  = -R_0 ( \bm u \cdot \bm n ) \bm n\,\,   && \mathrm{on}  \label{eq:efflux_condition} \,\, \Gamma_{\mathrm{skull^u}}, \\ 
-\bm u_{\mathrm{CSF}} & = 0 && \mathrm{on} \,\, \Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull^l}} \Gamma_{\rm{SSAS}}  \\
-\bm u_{\mathrm{CSF}} \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm \tau = 0 \quad && \mathrm{on} \,\, \Gamma_{\rm{LV}} .  
-\end{alignat}
-\end{subequations}
-Here, $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the CSF viscosity, and $\bm u_{\rm in}$ is the steady CSF flow across the lateral ventricle wall, see \Cref{tab:pvs:parameters}.
-Condition \eqref{eq:efflux_condition} models an efflux site with positive resistance $R_0 \geq 0$.
-The Stokes system is solved with the finite element software $FEniCS$, stored and then subsequently read for all the relevant solute transport simulations.
-See Figure~\ref{fig:csf_flow_cardiac} for a streamline visualization of the obtained velocity profile and \Cref{sec:details_numerical_method} for details on the numerical algorithm employed.
 
-We do not explicitly model convection within the brain parenchyma $\Omega_{\rm PAR}$, and thus set ${\bm u} = {\bm u}_{\rm PAR} = 0$ there.
+\subsubsection{Flow in perivascular networks induced by pressure gradients}
 
-\subsubsection{Perivascular fluid flow induced by pressure gradients}
-
-The production of CSF in the choroid plexus generates a small, but consistent pressure gradient within the intracranial space.
-To compute the resulting CSF flow within the PVS, we solve 1D Darcy equations in the PVS network.
-The unknowns are the averaged PVS velocity, $u_{\rm pvs} = \langle u_{v,s} \rangle$, and the cross-section pressure $p_{\rm pvs} $ (assumed to be constant in each cross-section).
-The system is given by  \cite{daversin2022geometrically, gjerde2024directional}
+To model flow induced by pressure differences between end nodes in a
+perivascular network $\Lambda$, with segments $\Lambda^i$, internal
+nodes $\mathcal{B}$ and end nodes $\mathcal{T}$, we consider the
+following system of hydraulic network
+equations~\cite{daversin2022geometrically, gjerde2024directional}. The
+unknowns are the PVS velocity $u_{\rm pvs}$ and pressure $p_{\rm
+  pvs}$, which represent the fluid velocity in the axial direction
+averaged across the PVS cross-sections and the cross-section-average
+fluid pressure, respectively, solving
+\mer{MER: Stopped here.}
 \begin{alignat}{2}
@@
 The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is denoted by $\mathcal{E}_{h}$.
+
+\subsubsection{Solute transport}
+For the 3D concentration, we use broken polynomial spaces and interior penalty discontinuous Galerkin (DG) methods with upwinding in order to capture the sharp boundary layers.
+For DG, we utilize the discrete space
+\begin{equation}
+V_h^k = \{v_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
+\end{equation}
+To discretize the diffusion term, we utilize symmetric weighted interior penalty dG bilinear forms.
+We refer to \cite{ern2009discontinuous} and \cite[Section 4.5.2.3]{di2011mathematical} for details on this method.
+We define
+\begin{multline}
+a_h (c_h, v_h) = \sum_{E \in \mathcal{T}_h } \int_{E} D \nabla (\phi c_h) \cdot \nabla v_h + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c_h][v_h] 
+- \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c_h \}_w \cdot \bm n_F [v_h]  + \{D \phi\nabla v_h \}_w \cdot \bm n_F [c_h] \right).
+ \nonumber
+ \end{multline}
+In the above, for $F\in \mathcal{F}_h^i$ with a normal $n_F$ pointing from $E_F^1$ to $E_F^2$,  the jump is given by $[v_h] = v_h{}_{\vert_{E_F^1}} - v_h{}_{\vert_{E_{F}^2}}$, and the weighed average is defined as follows.
+\begin{equation}
+\{v_h\}_{w} = \frac{ \kappa_{\vert_{E_F^2}}}{\kappa_{\vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}} v_h{}_{\vert_{E_{F}^1}} + \frac{\kappa_{\vert_{E_F^1}}}{\kappa_{ \vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}} v_h{}_{\vert_{E_{F}^2}}, \quad \text{where} \quad \kappa = D \phi.
+\end{equation}
+The parameter $\gamma_{D,F}$ is the harmonic mean of the diffusion coefficient given by 
+$$
+\gamma_{D,F} = \frac{ \kappa_{\vert_{E_F^1}}\kappa_{\vert_{E_F^2}}}{\kappa_{\vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}},
+$$
+and $\eta$ is a user specified penalty to be determined later.
+
+For the convection term, we use upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the references therein. Define 
+\begin{equation*}
+a_h^{\mathrm{up}}(c_h, v_h) =- \sum_{E\in\mathcal{T}_h } \int_E \bm{u} c_h \nabla v_h  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c_h\} \cdot \bm{n}_F [v_h]  + \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F |}{2} [c_h ] [v_h ]
+\end{equation*}
+The discrete formulation for \eqref{eq:3d_pde} then reads 
+\begin{multline}
+\int_{\Omega}\frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v_h + a_h(c_h^n, v_h) + a_h^{\mathrm{up}}(c_h^n,v_h) + \int_{\Gamma_{\mathrm{pia}} } \beta_{\mathrm{pia}}[c_h^n][v_h]  \\ + \int_{\Gamma_{\rm skull^u}} \beta c_h^n v_h +  \int_{\Lambda} \xi  P (\overline{c_h^n} - \hat c_h^n) \overline{v_h}  = \int_{\Omega} f^n v_h + \int_{\Gamma_{\rm SAS}} \gin v_h , \quad \forall v_h \in V_h^k.
+\end{multline}
+For approximating the solution of the 1D equations in the network, we use the space 
+\[ \hat V_h = \{ v_h \in C^0(\Lambda), 
+\; v_h \vert_{\Lambda^i} \in V_h(\Lambda^i)
+\}, 
+\]
+where $V_h(\Lambda^i)$ is the space of continuous polynomials of degree $1$ on each $\Lambda^i$,
+
 \subsubsection{CSF Flow}
-We recall the system of equations that we approximate. Find $(\bm u, p)$ such that 
-\begin{subequations}
-    \begin{alignat}{2}
- - \mu \Delta \bm u + \nabla p & =  \bm 0 \quad && \mathrm{in} \,\,  \Omega_{\rm CSF}, \label{eq:momnetum_equation}  \\ 
- \nabla \cdot  \bm u & = 0 \quad && \mathrm{in} \,\,   \Omega_{\rm CSF}, \label{eq:divergence_equation}  \\ 
-\mu \nabla \bm u \cdot \bm{n} -  p \bm n  &  = -R_0 ( \bm u \cdot \bm n ) \bm n, \,\,  \quad \bm u \cdot \bm \tau = 0 \quad  && \mathrm{on}  \label{eq:efflux_condition} \,\, \Gamma_{\mathrm{skull^u}}, \\
-\bm u & = 0 && \mathrm{on} \,\, \Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull^l}} \cup \Gamma_{\rm{SSAS}} \\
-\bm u \cdot  \bm n   =  \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}  , & \; \; \bm u \cdot \bm \tau  = 0  && \mathrm{on} \;  \Gamma_{\mathrm{LV}},
-\end{alignat}
-\end{subequations}
+%% We recall the system of equations that we approximate. Find $(\bm u, p)$ such that 
+%% \begin{subequations}
+%%     \begin{alignat}{2}
+%%  - \mu \Delta \bm u + \nabla p & =  \bm 0 \quad && \mathrm{in} \,\,  \Omega_{\rm CSF}, \label{eq:momnetum_equation}  \\ 
+%%  \nabla \cdot  \bm u & = 0 \quad && \mathrm{in} \,\,   \Omega_{\rm CSF}, \label{eq:divergence_equation}  \\ 
+%% \mu \nabla \bm u \cdot \bm{n} -  p \bm n  &  = -R_0 ( \bm u \cdot \bm n ) \bm n, \,\,  \quad \bm u \cdot \bm \tau = 0 \quad  && \mathrm{on}  \label{eq:efflux_condition} \,\, \Gamma_{\mathrm{skull^u}}, \\
+%% \bm u & = 0 && \mathrm{on} \,\, \Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull^l}} \cup \Gamma_{\rm{SSAS}} \\
+%% \bm u \cdot  \bm n   =  \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}  , & \; \; \bm u \cdot \bm \tau  = 0  && \mathrm{on} \;  \Gamma_{\mathrm{LV}},
+%% \end{alignat}
+%% \end{subequations}
+
 We follow \cite{hong2016robust} and approximate the velocity field $u$ and the pressure $p$ with the following finite element spaces. 
@@
 
-\subsubsection{Solute transport}
-For the 3D concentration, we use broken polynomial spaces and interior penalty discontinuous Galerkin (DG) methods with upwinding in order to capture the sharp boundary layers.
-For DG, we utilize the discrete space
-\begin{equation}
-V_h^k = \{v_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
-\end{equation}
-To discretize the diffusion term, we utilize symmetric weighted interior penalty dG bilinear forms.
-We refer to \cite{ern2009discontinuous} and \cite[Section 4.5.2.3]{di2011mathematical} for details on this method.
-We define
-\begin{multline}
-a_h (c_h, v_h) = \sum_{E \in \mathcal{T}_h } \int_{E} D \nabla (\phi c_h) \cdot \nabla v_h + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c_h][v_h] 
-- \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c_h \}_w \cdot \bm n_F [v_h]  + \{D \phi\nabla v_h \}_w \cdot \bm n_F [c_h] \right).
- \nonumber
- \end{multline}
-In the above, for $F\in \mathcal{F}_h^i$ with a normal $n_F$ pointing from $E_F^1$ to $E_F^2$,  the jump is given by $[v_h] = v_h{}_{\vert_{E_F^1}} - v_h{}_{\vert_{E_{F}^2}}$, and the weighed average is defined as follows.
-\begin{equation}
-\{v_h\}_{w} = \frac{ \kappa_{\vert_{E_F^2}}}{\kappa_{\vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}} v_h{}_{\vert_{E_{F}^1}} + \frac{\kappa_{\vert_{E_F^1}}}{\kappa_{ \vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}} v_h{}_{\vert_{E_{F}^2}}, \quad \text{where} \quad \kappa = D \phi.
-\end{equation}
-The parameter $\gamma_{D,F}$ is the harmonic mean of the diffusion coefficient given by 
-$$
-\gamma_{D,F} = \frac{ \kappa_{\vert_{E_F^1}}\kappa_{\vert_{E_F^2}}}{\kappa_{\vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}},
-$$
-and $\eta$ is a user specified penalty to be determined later.
-
-For the convection term, we use upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the references therein. Define 
-\begin{equation*}
-a_h^{\mathrm{up}}(c_h, v_h) =- \sum_{E\in\mathcal{T}_h } \int_E \bm{u} c_h \nabla v_h  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c_h\} \cdot \bm{n}_F [v_h]  + \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F |}{2} [c_h ] [v_h ]
-\end{equation*}
-The discrete formulation for \eqref{eq:3d_pde} then reads 
-\begin{multline}
-\int_{\Omega}\frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v_h + a_h(c_h^n, v_h) + a_h^{\mathrm{up}}(c_h^n,v_h) + \int_{\Gamma_{\mathrm{pia}} } \beta_{\mathrm{pia}}[c_h^n][v_h]  \\ + \int_{\Gamma_{\rm skull^u}} \beta c_h^n v_h +  \int_{\Lambda} \xi  P (\overline{c_h^n} - \hat c_h^n) \overline{v_h}  = \int_{\Omega} f^n v_h + \int_{\Gamma_{\rm SAS}} \gin v_h , \quad \forall v_h \in V_h^k.
-\end{multline}
-For approximating the solution of the 1D equations in the network, we use the space 
-\[ \hat V_h = \{ v_h \in C^0(\Lambda), 
-\; v_h \vert_{\Lambda^i} \in V_h(\Lambda^i)
-\}, 
-\]
-where $V_h(\Lambda^i)$ is the space of continuous polynomials of degree $1$ on each $\Lambda^i$,
-
-\subsection{Cardiac Dispersion Factor Estimation}
-
-\subsubsection*{Dispersion in the CSF spaces}
+\subsection{Dispersion in the CSF spaces}
+\label{sec:supp:dispersion}
 
@@
 
-\subsubsection{Cardiac Dispersion Factor Estimation}\label{sec:dispersion_details}
+%\subsubsection{Cardiac Dispersion Factor Estimation}
+%\label{sec:dispersion_details}
 

### 2024-11-18 paper/supplementary_main.tex
Reached the numerics - woohoo!

@@
 
-These sections provide more detailed descriptions of the mathematical
-models and numerical approximations considered.
+These sections provide more detailed descriptions of the mathematical models and numerical approximations considered.
 
-\subsection{Mathematical models and definitions}
+\subsection{Mixed-dimensional transport and flow equations}
 \label{sec:more_maths}
@@
 \begin{equation}
-  \hat{c}^i (y) = \hat{c}_y \qquad \foralls \Lambda^i \in \mathcal{E}(y),
-  \quad \foralls y \in \mathcal{B} .
+  \hat{c}^i (y) = \hat{c}^j(y)
+  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
+  \qquad \foralls y \in \mathcal{B} .
 \end{equation}
-Moreover, to ensure conservation of flux, we set that the flux going
-in and out at nodes should add to zero, or more precisely, that for
-each $y \in \mathcal{B}$:
+Moreover, to ensure \discuss{mass conservation}, we set that the flux
+going in and out at nodes should add to zero, or more precisely, that
+for each $y \in \mathcal{B}$:
 \begin{equation}
-  \sum_{i \in \mathcal{E}(y)}
-  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c^i}(y) \right ) n_i(y)
+  \sum_{\Lambda^i \in \mathcal{E}(y)}
+  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n_i(y)
   = 0
@@
 \end{subequations}
-where $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$.
+where $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the (dynamic) CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$.
 We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:supp:dispersion}.
@@
 
-To model flow induced by pressure differences between end nodes in a
-perivascular network $\Lambda$, with segments $\Lambda^i$, internal
-nodes $\mathcal{B}$ and end nodes $\mathcal{T}$, we consider the
-following system of hydraulic network
-equations~\cite{daversin2022geometrically, gjerde2024directional}. The
-unknowns are the PVS velocity $u_{\rm pvs}$ and pressure $p_{\rm
-  pvs}$, which represent the fluid velocity in the axial direction
-averaged across the PVS cross-sections and the cross-section-average
-fluid pressure, respectively, solving
-\mer{MER: Stopped here.}
+To model steady flow induced by pressure differences between end nodes
+in a perivascular network $\Lambda$, with edges $\{\Lambda^i\}_i$,
+internal nodes $\mathcal{B}$ and end nodes $\mathcal{T}$, we consider
+the following system of hydraulic network
+equations~\cite{tithof2022network, daversin2022geometrically,
+  gjerde2024directional}. The unknowns are the PVS flux $\hat{q} :
+\Lambda \rightarrow \R$ and pressure $\hat{p} : \Lambda
+\rightarrow \R$, which represent the fluid flux and the average
+pressure across cross-sections of the PVSs, respectively. These are
+defined for each $\Lambda^i$ by solving:
+\begin{subequations}
 \begin{alignat}{2}
-A \,  u_{\rm pvs}   + \frac{\kappa}{\mu} \, A \, \partial_{s} p_{\rm pvs} & = 0, &&  \quad \rm in  \,\, \Lambda  \\ 
--\partial_s (A \, u_{\rm pvs}) & = f, && \quad \rm in  \,\, \Lambda .  
+  \hat{q} + \frac{A \kappa}{\mu} \partial_{s} \hat{p} & = 0 &&
+  \quad \text{ in } \Lambda^i , \\ 
+  - \partial_s \hat{q} & = 0 &&
+  \quad \text{ in } \Lambda^i ,  
 \end{alignat} 
-In the above, $A$ is the area of the PVS, $\mu$ is the dynamic viscosity, and $\kappa$ is derived from the assumption of Poiseuille
-flow in the annular cross-section of the PVS \cite{daversin2022geometrically,tithof2022network}: 
+\label{eq:1d_flow}%
+\end{subequations}%
+where $A$ is the area of the PVS cross-sections, $\mu$ is the CSF viscosity as before, and $\kappa$ is derived from an assumption of Poiseuille flow in the annular cross-section of the PVS as~\cite{tithof2022network}:
 \begin{equation}
@@
 \end{equation}
-On bifurcations, conservation of fluxes is enforced weakly and continuity of the pressure is enforced strongly (in the choice of the finite element spaces).
-
-Since the pressure gradient the caused by the steady production-driven CSF flow (see \cref{sec:csf_fluid_vel}), we impose the precomputed fluid pressure $p_{\rm CSF}$ on the networks root nodes:
-
+The PVS flux $\hat{q}$ and the associated average PVS velocity in the
+axial direction $\hat{u}$ are directly related by 
+area
 \begin{equation}
-    p_{\rm PVS} = p_{\rm CSF} \quad \text{ on } \Lambda^{\rm out}
+  \hat{q}^i = A^i \hat{u}^i \qquad \foralls \Lambda^i.
 \end{equation}
 
-In case that the root does not lie within the CSF-space but the parenchyma, the CSF pressure is not yet computed.
-For that reason, we apply a simple extension technique to map the CSF pressure field into $\Omega_{\rm PAR}$.
-In particular, we solve a Laplace equation on $\Omega_{\rm PAR}$ with the CSF pressure on the Pial membrane ($\Gamma_{\rm Pia}$) as a boundary condition.
+In addition, to complete~\eqref{eq:1d_flow}, we impose continuity of
+the fluid pressure and conservation of flux at each internal node. Write
+$\hat{p}^i = \hat{p}|_{\Lambda^i}$. These conditions then read as: for
+each $y \in \mathcal{B}$ with connected edges $\mathcal{E}(y)$:
+\begin{subequations}
+\begin{align}
+  \hat{p}^i(y) &= \hat{p}^j(y) \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y), \\
+  \sum_{\Lambda^i \in \mathcal{E}(y)} \hat{q}^i n^i(y) &= 0  .
+\end{align}
+\end{subequations}
+Finally, to drive this flow, we impose given fluid pressures $p_0$ at the
+network end nodes:
+\begin{equation}
+  \hat{p}(y) = p_0(y) \qquad y \in \mathcal{T} .
+\end{equation}
+In the case where modelling flow in the perivascular networks
+$\Lambda_a, \Lambda_v$ induced by CSF production, we impose the fluid
+pressure $p_{\rm CSF}$ computed in the CSF spaces $\Omega_{\rm CSF}$
+(and its harmonic extensions into $\Omega_{\rm PAR}$) as the given
+fluid pressures $p_0$.
 
-\subsection{Numerical Method}
+\mer{MER: Stopped here.}
+
+\subsection{Numerical solution of the transport and flow equations}
 \label{sec:details_numerical_method}
+
 Here we provide details on the numerical method used to solve the coupled system.
@@
 \end{align}
+
 \subsubsection{PVS flow}
-
-Recall that we compute pressure-driven flow within the PVS by solving a 1D Darcy equations in the PVS network.
-The unknowns are the averaged PVS velocity, $u_{\rm pvs} = \langle u_{v,s} \rangle$, and the cross-section pressure $p_{\rm pvs} $ (assumed to be constant in each cross-section).
-The system is given by  \cite{daversin2022geometrically,
-gjerde2024directional}
-
-\begin{alignat}{2}
-A \,  u_{\rm pvs}   + \frac{\kappa}{\mu} \, A \, \partial_{s} p_{\rm pvs} & = 0, &&  \quad \rm in  \,\, \Lambda  \\ 
--\partial_s (A \, u_{\rm pvs}) & = f, && \quad \rm in  \,\, \Lambda .  
-\end{alignat}
-
-where $A$ is the area of the PVS, $\mu$ is the dynamic viscosity, and $\kappa$ is derived from the assumption of Poiseuille flow in the annular cross-section of the PVS  \cite{daversin2022geometrically,tithof2022network}:
-\begin{equation}
-\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
-\end{equation}
-
-As boundary conditions, we impose the precomputed fluid pressure $p_{\rm CSF}$ on the networks root nodes:
-\begin{equation}
-    p_{\rm PVS} = p_{\rm CSF} \quad \text{ on } \Lambda^{\rm out}
-\end{equation}
-
-In case that the root does not lie within the CSF-space but the parenchyma, the CSF pressure is not yet computed.
-For that reason, we apply a simple extension technique to map the CSF pressure field into $\Omega_{\rm PAR}$.
-In particular, we solve the Laplace equation
-\begin{equation}
-    - \Delta p = 0; \quad p = p_{\rm CSF} \text{ on } \Gamma_{\rm Pia}
-\end{equation}
-to obtain the extension of $p_{\rm CSF}$ onto $\Omega_{\rm PAR}$.
 
@@
 
-The weak form reads: Find $(u_{\rm PVS}, p_{\rm PVS}) \in V_h^{\rm PVS} \times Q_h^{\rm PVS}$ such that
+The weak form reads: Find $(\hat{u}, \hat{p}) \in V_h^{\rm PVS} \times Q_h^{\rm PVS}$ such that
 
 \begin{align}
-(A u_{\rm PVS}, v) + \left(v, \frac{A \kappa}{\mu} \partial_s p_{\rm PVS}  \right) &= 0 \quad \forall v \in V_h^{\rm PVS} \\
-(A u_{\rm PVS} , \partial_s q ) &= 0 \quad  \forall q \in  Q_h^{\rm PVS}
+(A \hat{u}, v) + \left(v, \frac{A \kappa}{\mu} \partial_s \hat{p}  \right) &= 0 \quad \forall v \in V_h^{\rm PVS} \\
+(A \hat{u}, \partial_s q ) &= 0 \quad  \forall q \in  Q_h^{\rm PVS}
 \end{align}
@@
 
-\subsection{Dispersion in the CSF spaces}
+\subsection{Modelling dispersion in the CSF spaces}
 \label{sec:supp:dispersion}

### 2024-11-20 paper/supplementary_main.tex
Simplify notation.
Former-commit-id: 97a165d7f0321b6c335f5b019593e033c356b3ef

@@
   \qquad \foralls y \in \mathcal{B} .
+  \label{eq:conc_continuity}
 \end{equation}
@@
 \end{equation}
-where the function $n_i : \mathcal{V} \rightarrow \{-1, 0, 1\}$
+where the function $n^i : \mathcal{V} \rightarrow \{-1, 0, 1\}$
 defines an orientation of each vertex relative to the
@@
 \begin{equation}
-  n_i(y_{\rm in}) = 1, \quad
-  n_i(y_{\rm{out}}) = -1, \quad
-  n_i(y) = 0 \quad \foralls y \in 
+  n^i(y_{\rm in}) = 1, \quad
+  n^i(y_{\rm{out}}) = -1, \quad
+  n^i(y) = 0 \quad \foralls y \in 
   \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} .  
+  \label{eq:def:sign}
 \end{equation}
 
-%% For a given edge $\Lambda^i= (\mathsf{v}_\mathrm{in}^i, \mathsf{v}_{\mathrm{out}}^{i})$, we define the function $n_i: \mathcal{V} \rightarrow \{-1,0,1\}$ with $$
+%% For a given edge $\Lambda^i= (\mathsf{v}_\mathrm{in}^i, \mathsf{v}_{\mathrm{out}}^{i})$, we define the function $n^i: \mathcal{V} \rightarrow \{-1,0,1\}$ with $$
 %% $$
@@
 %% \begin{equation}
-%% \sum_{ i \in \mathcal{E}(\mathsf{v})} (\hat D A_i \phi \partial_s \hat{u}_i(\mathsf{v})  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n_i(\mathsf{v}) = 0  \label{eq:flux_conservation}
+%% \sum_{ i \in \mathcal{E}(\mathsf{v})} (\hat D A_i \phi \partial_s \hat{u}_i(\mathsf{v})  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n^i(\mathsf{v}) = 0  \label{eq:flux_conservation}
 %% \end{equation}
@@
 \end{alignat} 
-  \label{eq:interface:3d}%
+  \label{eq:3d_interface}%
 \end{subequations}%
@@
 
-We supplement~\eqref{eq:3d_pde} and~\eqref{eq:interface:3d} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
+We supplement~\eqref{eq:3d_pde} and~\eqref{eq:3d_interface} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
 \begin{subequations}
@@
 \end{alignat} 
+\label{eq:3d_bcs}
 \end{subequations}
@@
     \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{T}.
+    \label{eq:1d_bcs}
 \end{equation}
@@
 
-\subsubsection{Flow in perivascular networks induced by pressure gradients}
+\subsubsection{Steady flow in perivascular networks -- induced by pressure differences}
 
@@
 The PVS flux $\hat{q}$ and the associated average PVS velocity in the
-axial direction $\hat{u}$ are directly related by 
-area
+axial direction $\hat{u}$ are directly related by the cross-sectional area
 \begin{equation}
@@
 \end{subequations}
-Finally, to drive this flow, we impose given fluid pressures $p_0$ at the
+where $n^i$ is as defined by~\eqref{eq:def:sign}. Finally, to drive this flow, we impose given fluid pressures $p_0$ at the
 network end nodes:
@@
 
-\mer{MER: Stopped here.}
-
 \subsection{Numerical solution of the transport and flow equations}
@@
 
-Here we provide details on the numerical method used to solve the coupled system.
-We are given a mesh $\mathcal{T}_h $, consisting of elements denoted by $E$, of $\Omega= \Omega_{\rm PAR} \cup \Omega_{\rm CSF}$.
-The collection of all interior facets to $\Omega_{\rm PAR}$ (resp.
-to $\Omega_{\mathrm{CSF}}$) is  denoted by $\mathcal{F}_{h,b}$ (resp.
-$\mathcal{F}_{h,c}$).
-The union of all interior facets $\mathcal{F}_i = \mathcal{F}_{h,b} \cup \mathcal{F}_{h,c}$.
-Facets lying on the pial interface are denoted by $\mathcal{F}_{\Gamma_{\rm pia}}$, on the skull by $\mathcal{F}_{\Gamma_{\rm skull}}$, and on the spinal canal by $\mathcal{F}_{\Gamma_{\rm SAS}}$.
-Note that $\mathcal{F}_i \cap \mathcal{F}_{\Gamma_{\rm pia}} = \emptyset $.
-The union of all the facets is denoted by $\mathcal{F}_h$.
-The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is denoted by $\mathcal{E}_{h}$.
+We solve the models presented in~\Cref{sec:more_maths} numerically
+using finite element methods. The following subsections provide more
+details, model-by-model.
 
-\subsubsection{Solute transport}
-For the 3D concentration, we use broken polynomial spaces and interior penalty discontinuous Galerkin (DG) methods with upwinding in order to capture the sharp boundary layers.
-For DG, we utilize the discrete space
+\subsubsection{Computational mesh entities: definitions and notation}
+
+We consider a mesh $\mathcal{T} = \{ E \}$ of $\Omega = \Omega_{\rm
+  PAR} \cup \Omega_{\rm CSF}$, consisting of tetrahedral mesh cells
+$E$, and conforming to the domains $\Omega_{\rm PAR}$ and $\Omega_{\rm
+  CSF}$ and to the CSF-brain interface $\Gamma_{\rm pia} \cup
+\Gamma_{\rm LV}$. We denote the restriction of $\mathcal{T}$ to
+$\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ by $\mathcal{T}_{\rm PAR}$
+and $\mathcal{T}_{\rm CSF}$, respectively. The collection of all
+interior facets (faces of the mesh cells) in $\mathcal{T}_{\rm PAR}$
+(resp. in $\mathcal{T}_{\rm CSF}$) is denoted by $\mathcal{F}_{i, \rm
+  PAR}$ (resp. $\mathcal{F}_{i, \rm CSF}$). We define the union of
+facets interior both to $\mathcal{T}_{i, \rm PAR}$ and
+$\mathcal{T}_{i, \rm CSF}$ as $\mathcal{F}_i = \mathcal{F}_{i, \rm
+  PAR} \cup \mathcal{F}_{i, \rm CSF}$. Facets lying on the pial
+interface $\Gamma_{\rm pia}$ are denoted by $\mathcal{F}_{\rm pia}$,
+on the outer (arachnoid) boundary $\Gamma_{\rm AM}$ by
+$\mathcal{F}_{\rm AM}$, on the interface toward the spinal cord
+$\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$ and on the interface
+towards the spinal SAS $\Gamma_{\rm SSAS}$ by $\mathcal{F}_{\rm
+  SSAS}$. The union of all facets in $\mathcal{T}$ is denoted by
+$\mathcal{F}$.
+
+%Note in particular that $\mathcal{F}_i \cap \mathcal{F}_{pia}} =
+%  \emptyset$.
+
+%The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is
+%denoted by $\mathcal{E}_{h}$.
+
+\subsubsection{Finite element solution of coupled 3D-1D solute transport equations }
+
+We consider the system of coupled 3D-1D solute transport equations
+given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
+conditions~\eqref{eq:3d_interface}, the bifurcation
+conditions~\eqref{eq:conc_continuity}
+and~\eqref{eq:flux_conservation}, and the boundary
+conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
+these equations using an implicit finite difference scheme in time and
+a discontinuous Galerkin (DG) finite element method with upwinding in
+space to capture the sharp boundary layers.
+
+We first consider the discretization of~\eqref{eq:3d_pde} and
+introduce the discrete space
 \begin{equation}
-V_h^k = \{v_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
+  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
+  V_h^k = \{c \in L^2(\discuss{\mathcal{T}}): \;  c \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}\} ,
 \end{equation}
-To discretize the diffusion term, we utilize symmetric weighted interior penalty dG bilinear forms.
-We refer to \cite{ern2009discontinuous} and \cite[Section 4.5.2.3]{di2011mathematical} for details on this method.
-We define
-\begin{multline}
-a_h (c_h, v_h) = \sum_{E \in \mathcal{T}_h } \int_{E} D \nabla (\phi c_h) \cdot \nabla v_h + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c_h][v_h] 
-- \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c_h \}_w \cdot \bm n_F [v_h]  + \{D \phi\nabla v_h \}_w \cdot \bm n_F [c_h] \right).
- \nonumber
- \end{multline}
-In the above, for $F\in \mathcal{F}_h^i$ with a normal $n_F$ pointing from $E_F^1$ to $E_F^2$,  the jump is given by $[v_h] = v_h{}_{\vert_{E_F^1}} - v_h{}_{\vert_{E_{F}^2}}$, and the weighed average is defined as follows.
+where $L^2(\mathcal{T})$ is the space of square-integrable functions
+defined relative to $\mathcal{T}$ and $\mathbb{P}^k(E)$ denotes the
+space of polynomials of degree $\leqslant k$ defined over the
+tetrahedra $E$. To discretize the diffusion term in~\eqref{eq:3d_pde},
+we use a symmetric weighted interior penalty DG formulation, referring
+to~\cite{ern2009discontinuous} and \cite[Section
+  4.5.2.3]{di2011mathematical} for details on this
+method. \discuss{Recall that $\phi$ is constant in each domain
+  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
+  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.}
+Define
 \begin{equation}
-\{v_h\}_{w} = \frac{ \kappa_{\vert_{E_F^2}}}{\kappa_{\vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}} v_h{}_{\vert_{E_{F}^1}} + \frac{\kappa_{\vert_{E_F^1}}}{\kappa_{ \vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}} v_h{}_{\vert_{E_{F}^2}}, \quad \text{where} \quad \kappa = D \phi.
+  a_h (c, v)
+  =
+  \sum_{E \in \mathcal{T}} \int_{E} D \discuss{\phi} \nabla c \cdot \nabla v
+  + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
+  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
+  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right).
+  \label{eq:dg_sip}
 \end{equation}
-The parameter $\gamma_{D,F}$ is the harmonic mean of the diffusion coefficient given by 
-$$
-\gamma_{D,F} = \frac{ \kappa_{\vert_{E_F^1}}\kappa_{\vert_{E_F^2}}}{\kappa_{\vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}},
-$$
-and $\eta$ is a user specified penalty to be determined later.
+In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
+between cells $E^1$ and $E^2$ with facet normal vector $\bm n_F$
+pointing from $E^1$ to $E^2$, the jump $[\cdot]$ is given by $[v] =
+v|_{E^1} - v|_{E^2}$, and the weighed average $\{ \cdot \}_w$ is
+defined as:
+\begin{equation}
+  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2},
+  \quad \text{where} \quad \kappa_i = \phi D|_{E^i}.  
+\end{equation}
+The parameter $\gamma_{D,F}$ is the harmonic mean of the \discuss{porosity-weighted} diffusion coefficient given by 
+\begin{equation}
+  \gamma_{D,F} = \frac{\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
+\end{equation}
+and $\eta$ is a user-specified penalty parameter.
 
-For the convection term, we use upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the references therein. Define 
+\mer{MER. Stopped here.}
+
+For the convection term, we use upwinding, see \cite[Section
+  2.3.1]{di2011mathematical} and the references therein. For discrete
+functions $c, v \in V_h^k$, define
 \begin{equation*}

### 2024-11-20 paper/supplementary_main.tex
More work on supplementary
Former-commit-id: 85df86cb99d6072149c58ac2ed723651489a50af [formerly 1da9eedc456c75b36c5208f6be310588e50a307c]
Former-commit-id: 9081a0be04f951961e32ba35b3252367260989cf

@@
 \subsubsection{Solute transport and exchange in intracranial domains (3D) and perivascular networks (1D)}
+\label{sec:app:transport_eqs}
 
@@
 \begin{equation}
-  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma = f \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
+  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma
+  %= f
+   = 0
+  \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
   \label{eq:3d_pde}
@@
 To summarize,
-\[
+\begin{equation*}
 \phi =  \begin{cases}
@@
 \end{cases} . 
-\]
+\end{equation*}
 In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp.
@@
 Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by
-\begin{equation}
+\begin{equation*}
   \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta_2^i(s)} c 
-\end{equation}
+\end{equation*}
 where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$.
@@
 such that
-\begin{equation}
+\begin{equation*}
   \langle \xi (\overline{c} - \hat c ) \delta_\Gamma, v \rangle =
   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
-\end{equation}
+\end{equation*}
 under the assumption that $\xi$ is constant on each $\partial
@@
 
+On the interface between the brain and the CSF spaces
+$\Gamma_{\mathrm{pia}} \cup \Gamma_{\rm LV}$, we prescribe the
+following interface conditions, which represent a semi--permeable
+interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and
+$c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
+%% \begin{subequations}
+%% \begin{alignat}{2}
+%% (-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} &= - (- D (\phi \nabla c_{\rm{CSF}}) + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n} &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\
+%% (-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,
+%% \end{alignat} 
+%% \end{subequations}
+\begin{subequations}
+  \begin{alignat}{2}
+  (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n}
+  &= - (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm CSF}} \cdot \bm{n}
+  &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
+  (- D \nabla (\phi  c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
+\end{alignat} 
+  \label{eq:3d_interface}%
+\end{subequations}%
+where $\bm{n} = \bm{n}_{\rm PAR}$ is the normal vector field defined
+over the interface, outward-pointing when viewed from $\Omega_{\rm
+  PAR}$, and $\beta_{\mathrm{pia}} \geqslant 0$ is a permeability
+constant.
+
+We supplement~\eqref{eq:3d_pde} and~\eqref{eq:3d_interface} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
+\begin{subequations}
+\begin{alignat}{2}
+(D \nabla (\phi c) - \bm u c ) \cdot \bm{n} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \Gamma_{\rm SSAS},  \\ 
+% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\Gamma_{\rm SAS}} &= 0 &&  \quad  \mathrm{on} \quad \Gamma_{\rm SAS}^{\mathrm{out}}  \\ 
+(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = \beta_{\rm exit} c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-U},  \\ 
+(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = 0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-L} \cup \Gamma_{\rm SC}  .
+\end{alignat} 
+\label{eq:3d_bcs}
+\end{subequations}
+where again $\bm n$ denotes the outward-pointing boundary normal.
+
 The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
@@
 and $\hat{u}$ is a convective velocity representing an (average) CSF
-flow velocity in the axial direction of the perivascular spaces. To
-complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
+flow velocity in the axial direction of the perivascular spaces.
+
+To complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
 internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
@@
 
-%% For a given edge $\Lambda^i= (\mathsf{v}_\mathrm{in}^i, \mathsf{v}_{\mathrm{out}}^{i})$, we define the function $n^i: \mathcal{V} \rightarrow \{-1,0,1\}$ with $$
-%% $$
-%% The collection of bifurcation points is denoted by $\mathcal{B} = \{ \mathsf{v} \in \mathcal{V}, \,\,  \mathrm{card}(\mathcal{E}(\mathsf{v})) \geq  3\}$.
-For any $\mathsf{v} \in \mathcal{B}$, we have
-%% \begin{equation}
-%% \sum_{ i \in \mathcal{E}(\mathsf{v})} (\hat D A_i \phi \partial_s \hat{u}_i(\mathsf{v})  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n^i(\mathsf{v}) = 0  \label{eq:flux_conservation}
-%% \end{equation}
-
-\paragraph{Interface conditions between the brain and CSF spaces}
-
-On the interface between the brain and the CSF spaces
-$\Gamma_{\mathrm{pia}} \cup \Gamma_{\rm LV}$, we prescribe the
-following interface conditions, which represent a semi--permeable
-interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and
-$c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
-%% \begin{subequations}
-%% \begin{alignat}{2}
-%% (-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} &= - (- D (\phi \nabla c_{\rm{CSF}}) + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n} &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\
-%% (-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,
-%% \end{alignat} 
-%% \end{subequations}
-\begin{subequations}
-  \begin{alignat}{2}
-  (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n}
-  &= - (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm CSF}} \cdot \bm{n}
-  &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
-  (- D \nabla (\phi  c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
-\end{alignat} 
-  \label{eq:3d_interface}%
-\end{subequations}%
-where $\bm{n} = \bm{n}_{\rm PAR}$ is the normal vector field defined
-over the interface, outward-pointing when viewed from $\Omega_{\rm
-  PAR}$, and $\beta_{\mathrm{pia}} \geqslant 0$ is a permeability
-constant.
-
-\paragraph{Boundary conditions: outer SAS interfaces and perivascular ends}
-
-We supplement~\eqref{eq:3d_pde} and~\eqref{eq:3d_interface} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
-\begin{subequations}
-\begin{alignat}{2}
-(D \nabla (\phi c) - \bm u c ) \cdot \bm{n} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \Gamma_{\rm SSAS},  \\ 
-% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\Gamma_{\rm SAS}} &= 0 &&  \quad  \mathrm{on} \quad \Gamma_{\rm SAS}^{\mathrm{out}}  \\ 
-(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = \beta c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-U},  \\ 
-(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = 0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-L} \cup \Gamma_{\rm SC}  .
-\end{alignat} 
-\label{eq:3d_bcs}
-\end{subequations}
-where again $\bm n$ denotes the outward-pointing boundary normal.
-
-%In the above, note that $\bm u \cdot \bm n  \neq 0$ on $\Gamma_{\rm LV} \cup \Gamma_{\rm skull^u}$ and $\bm u \cdot \bm n = 0$ on $\Gamma_{\rm SSAS} \cup \Gamma_{\rm skull^l} \cup \Gamma_{\rm Pia}$.
-
-For the perivascular networks, we assume no molecular efflux at all
-end nodes $\mathcal{T}$ via a homogeneous Neumann condition
-augmenting~\eqref{eq:1d_pde}:
+Finally, we assume no molecular efflux at all end nodes $\mathcal{T}$
+via a homogeneous Neumann condition augmenting~\eqref{eq:1d_pde}:
 \begin{equation}
@@
 velocity vector field $\bm u : \Omega_{\rm CSF} \rightarrow \R^3$ and
-the pressure $p : \Omega_{\rm CSF} \rightarrow \R$ such that
+the pressure field $p : \Omega_{\rm CSF} \rightarrow \R$ such that
 \begin{subequations}
@@
 a discontinuous Galerkin (DG) finite element method with upwinding in
-space to capture the sharp boundary layers.
+space to accurately capture sharp boundary layers.
 
@@
   that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.}
-Define
+Define for $c, v \in V_h^k$:
 \begin{equation}
@@
 pointing from $E^1$ to $E^2$, the jump $[\cdot]$ is given by $[v] =
-v|_{E^1} - v|_{E^2}$, and the weighed average $\{ \cdot \}_w$ is
-defined as:
+v|_{E^1} - v|_{E^2}$, and the unweighted average $\{ \cdot \}$ and
+weighed average $\{ \cdot \}_w$ are defined as:
 \begin{equation}
-  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2},
+  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
+  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
   \quad \text{where} \quad \kappa_i = \phi D|_{E^i}.  
 \end{equation}
-The parameter $\gamma_{D,F}$ is the harmonic mean of the \discuss{porosity-weighted} diffusion coefficient given by 
+The parameter $\gamma_{D,F}$ is the harmonic mean of the
+\discuss{porosity-weighted} diffusion coefficient given by
 \begin{equation}
@@
 \end{equation}
-and $\eta$ is a user-specified penalty parameter.
+and $\eta$ is a user-specified penalty parameter. To discretize the convection term in~\eqref{eq:3d_pde}, we use
+upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
+references therein. For $c, v \in V_h^k$, define
+\begin{equation}
+  a_h^{\mathrm{up}}(c, v)
+  = - \sum_{E \in \mathcal{T}} \int_E \bm{u} c \nabla v
+  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
+  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
+\end{equation}
+Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, find $c_h^n \in V_h^k$ such that, for all $v \in V_h^k$:
+\begin{equation}
+  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
+  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
+  + \int_{\Gamma_{\rm pia} \discuss{\cup \Gamma_{\rm LV}} } \beta_{\rm pia}[c_h^n][v] 
+  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
+  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
+  = 
+  %\int_{\Omega} f^n v +
+  \int_{\Gamma_{\rm \discuss{S}SAS}} g_{\rm influx} v
+  %, \quad \forall v \in V_h^k
+  .
+\end{equation}
 
-\mer{MER. Stopped here.}
-
-For the convection term, we use upwinding, see \cite[Section
-  2.3.1]{di2011mathematical} and the references therein. For discrete
-functions $c, v \in V_h^k$, define
-\begin{equation*}
-a_h^{\mathrm{up}}(c_h, v_h) =- \sum_{E\in\mathcal{T}_h } \int_E \bm{u} c_h \nabla v_h  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c_h\} \cdot \bm{n}_F [v_h]  + \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F |}{2} [c_h ] [v_h ]
-\end{equation*}
-The discrete formulation for \eqref{eq:3d_pde} then reads 
-\begin{multline}
-\int_{\Omega}\frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v_h + a_h(c_h^n, v_h) + a_h^{\mathrm{up}}(c_h^n,v_h) + \int_{\Gamma_{\mathrm{pia}} } \beta_{\mathrm{pia}}[c_h^n][v_h]  \\ + \int_{\Gamma_{\rm skull^u}} \beta c_h^n v_h +  \int_{\Lambda} \xi  P (\overline{c_h^n} - \hat c_h^n) \overline{v_h}  = \int_{\Omega} f^n v_h + \int_{\Gamma_{\rm SAS}} \gin v_h , \quad \forall v_h \in V_h^k.
-\end{multline}
-For approximating the solution of the 1D equations in the network, we use the space 
-\[ \hat V_h = \{ v_h \in C^0(\Lambda), 
-\; v_h \vert_{\Lambda^i} \in V_h(\Lambda^i)
-\}, 
-\]
-where $V_h(\Lambda^i)$ is the space of continuous polynomials of degree $1$ on each $\Lambda^i$,
+To discretize~\eqref{eq:1d_pde} with the bifurcation
+conditions~\eqref{eq:conc_continuity}
+and~\eqref{eq:flux_conservation}, and the boundary
+conditions~\eqref{eq:1d_bcs}, we use the space of continuous piecewise
+linear polynomials defined over $\Lambda$:
+\begin{equation}
+  \hat V_h
+  = \{ v \in C^0(\Lambda), \; v \vert_{\Lambda^i} \in \mathbb{P}^1(\Lambda^i) \}, 
+\end{equation}
+where $\mathbb{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$.
 

### 2024-11-20 paper/supplementary_main.tex
Updates in Stokes FEM. Check definitions of facets once again.

@@
   \end{alignat}
+  \label{eq:stokes}
 \end{subequations}
@@
   \end{alignat}
+  \label{eq:stokes_bcs}
 \end{subequations}
@@
 on the outer (arachnoid) boundary $\Gamma_{\rm AM}$ by
-$\mathcal{F}_{\rm AM}$, on the interface toward the spinal cord
-$\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$ and on the interface
-towards the spinal SAS $\Gamma_{\rm SSAS}$ by $\mathcal{F}_{\rm
-  SSAS}$. The union of all facets in $\mathcal{T}$ is denoted by
-$\mathcal{F}$.
+$\mathcal{F}_{\rm AM}$, on the inner ventricular interface
+$\Gamma_{\rm LV}$ by $\mathcal{F}_{\rm LV}$, on the boundary toward
+the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$ and on the
+boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
+$\mathcal{F}_{\rm SSAS}$. \discuss{For convenience, we denote $\mathcal{F}_e =
+\mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV} \cup \mathcal{F}_{\rm
+  AM-U} \cup \mathcal{F}_{\rm SSAS}$, and the union of all facets in
+$\mathcal{T}$ is denoted by $\mathcal{F} = \mathcal{F}_i \cup
+\mathcal{F}_e \discuss{\cup \mathcal{F}_{\rm SC} \cup \mathcal{F}_{\rm AM-L}}$.}
 
@@
 between cells $E^1$ and $E^2$ with facet normal vector $\bm n_F$
-pointing from $E^1$ to $E^2$, the jump $[\cdot]$ is given by $[v] =
-v|_{E^1} - v|_{E^2}$, and the unweighted average $\{ \cdot \}$ and
-weighed average $\{ \cdot \}_w$ are defined as:
+pointing from $E^1$ to $E^2$, \discuss{$h_F$ denotes the facet size},
+the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
+unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
+are defined as:
 \begin{equation}
@@
 \end{equation}
-Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, find $c_h^n \in V_h^k$ such that, for all $v \in V_h^k$:
+Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h^k$ such that for all $v \in V_h^k$:
 \begin{equation}
@@
 
-\subsubsection{CSF Flow}
-%% We recall the system of equations that we approximate. Find $(\bm u, p)$ such that 
-%% \begin{subequations}
-%%     \begin{alignat}{2}
-%%  - \mu \Delta \bm u + \nabla p & =  \bm 0 \quad && \mathrm{in} \,\,  \Omega_{\rm CSF}, \label{eq:momnetum_equation}  \\ 
-%%  \nabla \cdot  \bm u & = 0 \quad && \mathrm{in} \,\,   \Omega_{\rm CSF}, \label{eq:divergence_equation}  \\ 
-%% \mu \nabla \bm u \cdot \bm{n} -  p \bm n  &  = -R_0 ( \bm u \cdot \bm n ) \bm n, \,\,  \quad \bm u \cdot \bm \tau = 0 \quad  && \mathrm{on}  \label{eq:efflux_condition} \,\, \Gamma_{\mathrm{skull^u}}, \\
-%% \bm u & = 0 && \mathrm{on} \,\, \Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull^l}} \cup \Gamma_{\rm{SSAS}} \\
-%% \bm u \cdot  \bm n   =  \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}  , & \; \; \bm u \cdot \bm \tau  = 0  && \mathrm{on} \;  \Gamma_{\mathrm{LV}},
-%% \end{alignat}
-%% \end{subequations}
+\subsubsection{Finite element solution of the incompressible Stokes equations}
 
-We follow \cite{hong2016robust} and approximate the velocity field $u$ and the pressure $p$ with the following finite element spaces. 
+We consider an H(div)-based finite element approximation of the incompressible Stokes equations~\eqref{eq:stokes} defined over $\Omega_{\rm CSF}$ with the boundary conditions~\eqref{eq:stokes_bcs}.
+Following\cite{hong2016robust}, we approximate the velocity field $\bm u$ and the pressure field $p$ with the following finite element spaces:
+\begin{align*}
+  \bm V_h &= \{ \bm v  \in H(\mathrm{div}, \Omega_{\mathrm{CSF}}): \;
+  \bm v \vert_{E} \in \mathrm{BDM}^2(E), \; E \in \mathcal{T}_{\rm CSF}; \;
+  \bm v \cdot \bm n = 0
+  \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\rm AM-L} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} \text{ on } \Gamma_{\mathrm{LV}} \}  \\ 
+  Q_h  &= \{q \in L^2(\Omega_{\rm CSF}): \; q \vert_E \in \mathbb{P}^1(E),
+  \; E \in \mathcal{T}_{\rm CSF} \}. 
+\end{align*}
+Here, $H(\mathrm{div}, \Omega)$ is the space of $L^2(\Omega)$ vector fields with $L^2(\Omega)$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, and $\bm n$ and $\bm \tau$ are the unit outward normal and unit tangent vectors to each facet, respectively.
+Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by
+\begin{equation}
+\bm v_n = (\bm v \cdot \bm n) \bm n, \quad \bm v_t = \bm v - \bm v_n. 
+\end{equation}
+Since $\bm V_h \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm v_n] = 0 $ on $\mathcal{F}_{\rm CSF}$.
+Continuity in the tangential component is enforced weakly via interior penalization.
+Define the form \mer{MER: Check $\Gamma_{\rm AM}$ part.}
+\begin{multline}
+  \mathcal{A}_h(\bm u, \bm v)
+  = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
+  - \sum_{F \in \discuss{\mathcal{F}_{\rm CSF} \cup \mathcal{F}_{e}}} \left (
+  \int_{F} \mu \{\nabla \bm u\} \cdot \bm n_F \cdot [\bm v_t] 
+  - %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
+  \int_{F} \mu \{\nabla \bm v\}\cdot \bm n_F \cdot [\bm u_t]
+  + %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
+  \frac{\sigma \mu}{h} \int_{F} [\bm u_t ] \cdot [\bm v_t] \right ),   
+\end{multline}
+where \discuss{$h$ denotes the element size}.
+We set the penalty parameter for the tangential continuity to be $\sigma = 20$.
+Our finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_h \times Q_h$ such that
+\begin{subequations}
 \begin{align}
-\bm V_h  & = \{ \bm v  \in H(\mathrm{div}; \Omega_{\mathrm{CSF}}): \; \bm v \vert_{E} \in \mathbb{BDM}^2 (E),  \; E \in \mathcal{E}_h; \;  \bm v \cdot \bm n = 0 \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull^l}} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} \text{ on } \Gamma_{\mathrm{LV}} \} \\
-S_h  & = \{q \in L^2(\Omega):  \;  q \vert_E \in \mathbb{P}^1(E),  \; E \in \mathcal{E}_h \}. 
+  \mathcal{A}_h(\bm u_h, \bm v_h)
+  + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
+  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0, \qquad \foralls \bm v_h \in \bm V_h, \\ 
+\int_{\Omega_{\rm CSF}} \nabla \cdot \bm u_h \, q_h  &= 0, \qquad \foralls q_h \in Q_h.
 \end{align}
-Here, $H(\mathrm{div};\Omega)$ is the space of $L^2$ vector fields with $L^2$ divergence,  $\mathbb{BDM}^2$ is the Brezzi-Douglas--Marini element \cite{brezzi1987mixed} of order 2, and  $\bm n$ (resp.
-$\bm \tau$) is the unit outward normal (resp.
-tangent) vector.
-Given any vector $\bm v$, the normal and tangential components on each facet  are denoted by 
-\[ 
-\bm v_n = (\bm v \cdot \bm n)\bm n , \quad \bm v_t = (\bm v - \bm v_n). 
-\]
-Since $V_h \subset H(\mathrm{div}, \Omega_{\rm{CSF}})$, then $[\bm v_n] = 0 $ on $\mathcal{F}_{h,c}$.
-Continuity in the tangential component is enforced weakly via interior penalization.
-For notational simplicity, denote by $\mathcal{F}_{\Gamma}$ the facets lying on $\Gamma_{\rm{pia}} \cup \Gamma_{\mathrm{skull}} \cup \Gamma_{\rm{SSAS}} \cup \Gamma_{\mathrm{LV}}$.
-Define the form
-\begin{multline}
-\mathcal{A}_h (\bm u, \bm v) = \sum_{E \in \mathcal{E}_h} \int_{E} \mu \nabla u : \nabla v - \sum_{
-F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma} 
-} \int_{F} \mu \{\nabla u\}\cdot \bm n_F \cdot [\bm v_t] 
-\\ - \sum_{
-F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma} 
-} \int_{F} \mu \{\nabla v\}\cdot \bm n_F \cdot [\bm u_t]
-+ \sum_{
-F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma} 
-} \frac{\sigma \mu}{h} \int_{F} [\bm u_t ] \cdot [\bm v_t].   
-\end{multline}
-We set $\sigma =20$. The discretization for the CSF flow, is then to find $(\bm u_h, p_h) \in \bm V_h \times S_h$ such  that 
-\begin{align}
-\mathcal{A}_h(u_h, v_h)  + \sum_{F \in \mathcal{F}_{\Gamma_{\mathrm{skull}}}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n) - \int_{\Omega} \nabla \cdot \bm v_h \, p_h  & = 0, \quad \forall \bm v_h \in \bm V_h   \\
-(\nabla \cdot \bm u_h , q_h ) & = 0, \quad \forall q_h \in S_h.
-\end{align}
+\end{subequations}
 

### 2024-11-27 paper/supplementary_main.tex
Updates from Overleaf
Former-commit-id: 4979054c206f9013549cfe20f7a45416e5c88b29

@@
   \sum_{\Lambda^i \in \mathcal{E}(y)}
-  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n_i(y)
+  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) ni(y)
   = 0
@@
 \subsubsection{Modelling CSF flow via the incompressible Stokes equations}
-
+\label{sec:csf_flow}
 The time-independent, incompressible Stokes equations model the flow
@@
     && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
-    \bm u &= 0 && \text{ on } \Gamma_{\mathrm{AM-L}}, \\
-    \bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
-    \bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
+    \bm u &= 0 && \text{ on } \Gamma_{\mathrm{AM-L}}\cup\Gamma_{\rm{pia}}\cup\Gamma_{\rm{SSAS}}, \\
+    %\bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
+    %\bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
     \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm \tau = 0 \quad && \text{ on } \Gamma_{\rm{LV}} ,  
@@
 between cells $E^1$ and $E^2$ with facet normal vector $\bm n_F$
-pointing from $E^1$ to $E^2$, \discuss{$h_F$ denotes the facet size},
+pointing from $E^1$ to $E^2$, \discuss{$h_F$ denotes the facet diameter},
 the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
@@
 Continuity in the tangential component is enforced weakly via interior penalization.
-Define the form \mer{MER: Check $\Gamma_{\rm AM}$ part.}
+Define the form \mer{MER: Check $\Gamma_{\rm AM}$ part.
+26.11.24: Rami will check $\mathcal{F}_e$ versus definitions of the boundaries and where it appears.}
 \begin{multline}
@@
 
-In addition, CSF pulsates in the SAS and ventricular system with each
-cardiac and respiratory cycle.
+CSF pulsates in the SAS and ventricular system with each cardiac and respiratory cycle.
+Taking the large spatial differences of pulsatile CSF flow into account, we estimate a local diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac- and respiration-induced pulsatile CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}.
 
-Taking the large spatial differences of pulsatile CSF flow into account, we compute the local diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac induced CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}.
+\rami{I rewrote the above paragraph and some aspects of this section}
 
+\rami{  CSF pulsates in the SAS and ventricular system with each cardiac and respiratory cycle with large spatial differences.  This leads to dispersion effects. For example, 
 Sharp et al \cite{keith2019dispersion} estimated the enhancement of diffusion in the spinal subarachnoid space due to dispersive effects to average 5.8 times that of purely molecular diffusion with a theoretical maximum of $10^6$.
+This phenomena is accounted for in our model by scaling the free diffusion coeffiecient $D^{\mathrm{Gad}}$ by the formula $D^{\mathrm{eff}} = (1+R) D^{\mathrm{Gad}}$ where $R$ represents a local diffusion enhancement in the CSF-filled space $\Omega_{\rm CSF}$.}
 
-Similar to the case of CSF flow driven by production introduced in section \ref{sec:csf_fluid_vel}, we employ a Stokes flow model to predict CSF velocities at peak arterial dilation.
-In particular, we enforce a uniformly distributed inflow of $10$\,ml/s and $2$\,ml/s across the outer skull ($\Gamma_{\rm skull}$) and lateral ventricular surfaces ($\Gamma_{\rm LV}$) respectively, mimicking the effect of reduced CSF volume due to blood inflow.
+\rami{ Note that since there is no data for the local values of $R$, we estimate $R$ by combining computational results (solving Stokes flow equations, see below) and explicit estimates for dispersion enhancement that are derived under simplifying assumptions on the problem geometry, the flow field and concentration profiles \cite{watson1983diffusion, keith2019dispersion}.
+Therefore such an estimate is heuristic; however,  our computational results qualitatively agree with observed data.
+A key feature in our estimate of $R$ is that it is spatially varying, including local effects depending on a  computed flow field.
+}
+
+Similar to the case of CSF flow driven by production introduced in \Cref{sec:csf_flow}, we employ a Stokes flow model to predict CSF velocities corresponding to the peak cardiac and respiratory blood inflow rates.
+In particular, we enforce a uniformly distributed inflow of $u^{\rm in}_{\rm AM}$ and $u^{\rm in}_{\rm LV}$ across the arachnoid mater ($\Gamma_{\rm AM}$) and lateral ventricular surfaces ($\Gamma_{\rm LV}$) respectively, mimicking the effect of reduced CSF volume due to blood inflow.
+We set $u^{\rm in}_{\rm AM} = 6\,$ml/s and $u^{\rm in}_{\rm LV}=0.31\,$ml/s in the cardiac case, and $u^{\rm{in}}_{\rm{AM}} = 1\,$ml/s and $u^{\rm in}_{\rm{LV}}=0.121\,$ml/s in the respiratory case, respectively.
 We prescribe a no-slip boundary condition on the pial surface $\Gamma_{\rm Pia}$ and zero-traction on the spinal boundary $\Gamma_{\rm SSAS}$, making the spinal compartment the only outflow route of CSF.
-Thus, we compute the cardiac-driven flow field by augmenting the Stokes flow model (eq.
+Thus, we compute the pulsatile flow fields by augmenting the Stokes flow model (eq.
 \ref{eq:momnetum_equation} and \ref{eq:divergence_equation}) with the following boundary conditions:
-
 \begin{equation}\label{eq:cardiac_flow_bcs}
@@
     \mu \nabla \bm u \cdot \bm n - p \bm n = 0 \text{ on } \Gamma_{\rm SSAS}; \quad 
-    \bm u = \frac{0.31 \cdot \bm n}{|\Gamma_{\rm LV}|} \,\text{ml/s} \text{ on } \Gamma_{\rm LV}; \quad
-    \bm u = \frac{8 \cdot \bm n}{|\Gamma_{\rm skull}|} \,\text{ml/s} \text{ on } \Gamma_{\rm skull}.
+    \bm u = \frac{u^{\rm in}_{\rm LV} \cdot \bm n}{|\Gamma_{\rm LV}|} \text{ on } \Gamma_{\rm LV}; \quad
+    \bm u = \frac{u^{\rm in}_{\rm AM} \cdot \bm n}{|\Gamma_{\rm skull}|} \text{ on } \Gamma_{\rm skull}.
 \end{equation}
-
-The obtained pressure field only account for viscous forces, albeit cardiac induced CSF flow is known to be unsteady and substantially impacted by transient inertial forces.
-Assuming an oscillatory flow with the heartbeat frequency $\omega = 2 \pi$, a mean gap width of the CSF-filled spaces of $2h=3$\,mm and CSF density $\rho=10^3\,\text{kg/m}^3$, the Womersley number $\alpha$ becomes $\alpha^2 = \frac{h^2 \omega \rho}{\mu} = 20.2$, which is unsteady.
+The obtained pressure fields $p_{\rm cardiac}$ and $p_{\rm resp}$ only account for viscous forces, albeit cardiac and respiratory induced CSF flow is known to be unsteady and substantially impacted by transient inertial forces.
+Assuming an oscillatory flow with either the heartbeat frequency $\omega_{\rm cardiac} = 2 \pi$ or respiratory frequency $\omega_{\rm resp} = 0.5 \pi$, a mean gap width of the CSF-filled spaces of $2h=3$\,mm and CSF density $\rho=10^3\,\text{kg/m}^3$, the Womersley numbers $\alpha$ become $\alpha_{\rm cardiac}^2 = \frac{h^2 \omega_{\rm cardiac} \rho}{\mu} = 20.2$ and $\alpha_{\rm resp}^2=5.05$, which is unsteady.
 Based on theoretical considerations on the ratio of oscillatory flow to steady flow impedances in a tube, we account for inertial forces in the pulsatile flow by upscaling the computed viscous pressure field by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter.
 4.3).
+Thus, the upscaled pressure fields become $p^{\rm up}_{\rm cardiac}= (1 + \alpha_{\rm cardiac}^2 / 8)p$ and $p^{\rm up}_{\rm resp}= (1 + \alpha_{\rm resp}^2 / 8)p$.
 
-Further assuming unsteady dispersion, we follow Sharp et al \cite{keith2019dispersion} and calculate the enhancement factor $R$ from the non-dimensionalized pressure gradient $dP=\frac{|\nabla p|}{\omega \mu / h}$ \rami{Can we refer to the exact equation here from Sharp et al ?
-i find that they write $P = (\partial \tilde p / \partial \tilde x) / (\rho \omega \mu_e/h)$ ...
-so i am a bit confused
-}as $R\approx dP^2 / \alpha^3$.
-Combining the computed pressure gradient with this scaling law yields a pointwise approximation of the dispersion factor.
-Finally, we account for the non-local nature of the dispersion mechanism by applying a smoothing technique to obtain a spatially varying, but locally averaged diffusion enhancement factor $R$ (see the visualization in \cref{fig:visualize_R} and \cref{sec:dispersion_details} for additional details).
+Further assuming unsteady dispersion, we follow Sharp et al \cite{keith2019dispersion} and estimate each enhancement factor $R_{\rm cardiac}$ and $R_{\rm respiratory}$ from the non-dimensionalized pressure gradient
+$$R_{i} \approx dP_{i}^2 / \alpha_{i}^3, \quad  \text{where} \quad  dP=\frac{|\nabla p^{\rm up}_{i}|}{\omega_{i} \mu / h} \quad \text{for} \quad i = \{\mathrm{cardiac}, \mathrm{resp}\}.$$
+Finally, we account for the non-local nature of the dispersion mechanism and smooth the resulting field by solving a Helmholtz problem with $dP+i^2/\alpha_i^3$ as the right-hand side
+\begin{equation}
+- 10^{-4} \Delta R_i + R_i = dP_i^2/\alpha_i^3 \quad \text{ on } \Omega_{\rm CSF} \qquad \text{ and } \quad \nabla R_i \cdot \bm n = 0, \quad \text{ on } \partial \Omega_{\rm CSF}
+\end{equation}
+to obtain the locally averaged, spatially varying diffusion enhancement factor $R_i$.
+The diffusion enhancement factor $R$ in $D^{\mathrm{eff}} = D^{\mathrm{Gad}}( 1 +R)$ is then estimated by $R \approx R_{\rm{cardiac}} + R_{\rm{resp}}$.
+% %\subsubsection{Cardiac Dispersion Factor Estimation}
+% %\label{sec:dispersion_details}
 
-%\subsubsection{Cardiac Dispersion Factor Estimation}
-%\label{sec:dispersion_details}
-
-To compute a spatially varying estimate of the diffusion enhancement factor $R$, we consider the Stokes flow problem (\cref{eq:momnetum_equation} and \cref{eq:divergence_equation}) equipped with the boundary conditions \cref{eq:cardiac_flow_bcs} describing cardiac induced CSF flow at peak arterial pulsation.
+% To compute a spatially varying estimate of the diffusion enhancement factor $R$, we consider the Stokes flow problem (\cref{eq:momnetum_equation} and \cref{eq:divergence_equation}) equipped with the boundary conditions \cref{eq:cardiac_flow_bcs} describing cardiac induced CSF flow at peak arterial pulsation.
 We use the numerical method described in \cref{sec:details_numerical_method} to solve the Stokes system and compute the non-dimensionalized local pressure difference $dP=\frac{|\nabla p| \rho}{\omega \mu / h}$ pointwise.
@@
 
-In addition, we account for inertial forces in the pulsatile flow by upscaling the obtained pressure gradient by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter.
+% In addition, we account for inertial forces in the pulsatile flow by upscaling the obtained pressure gradient by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter.
 4.3), where $\alpha^2 = \frac{h^2 \omega \rho}{\mu} = 20.2$ is the square of the Womersley number for an oscillatory flow.
 
-Finally, we smooth the resulting field by solving a Helmholtz problem with $dP^2$ as the right-hand side
+% Finally, we smooth the resulting field by solving a Helmholtz problem with $dP^2/alpha^3$ as the right-hand side
 
-\begin{equation}
-- 10^{-4} \Delta R + R = dP^2 \quad \text{ on } \Omega_{\rm CSF} \qquad \text{ and } \quad \nabla R \cdot \bm n = 0 \quad \text{ on } \partial \Omega_{\rm CSF}
-\end{equation}
+% \begin{equation}
+% - 10^{-4} \Delta R + R = dP^2/\alpha^3 \quad \text{ on } \Omega_{\rm CSF} \qquad \text{ and } \quad \nabla R \cdot \bm n = 0 \quad \text{ on } \partial \Omega_{\rm CSF}
+% \end{equation}
 
-to obtain the locally averaged, spatially varying diffusion enhancement factor $R$.
+% to obtain the locally averaged, spatially varying diffusion enhancement factor $R$.
 

### 2024-11-27 paper/supplementary_main.tex
Minor corrections in notation in Section A.3
Former-commit-id: 1f75dc2b2672d3ba1bfe2bde2a096988001a3724

@@
   \end{alignat}
-  \label{eq:stokes}
-\end{subequations}
+  \label{eq:stokes}%
+\end{subequations}%
 In addition, we impose the following boundary conditions to model flow
@@
   \end{alignat}
-  \label{eq:stokes_bcs}
-\end{subequations}
+  \label{eq:stokes_bcs}%
+\end{subequations}%
 where $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the (dynamic) CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$.
@@
 and $\mathcal{T}_{\rm CSF}$, respectively. The collection of all
-interior facets (faces of the mesh cells) in $\mathcal{T}_{\rm PAR}$
-(resp. in $\mathcal{T}_{\rm CSF}$) is denoted by $\mathcal{F}_{i, \rm
-  PAR}$ (resp. $\mathcal{F}_{i, \rm CSF}$). We define the union of
-facets interior both to $\mathcal{T}_{i, \rm PAR}$ and
-$\mathcal{T}_{i, \rm CSF}$ as $\mathcal{F}_i = \mathcal{F}_{i, \rm
-  PAR} \cup \mathcal{F}_{i, \rm CSF}$. Facets lying on the pial
-interface $\Gamma_{\rm pia}$ are denoted by $\mathcal{F}_{\rm pia}$,
-on the outer (arachnoid) boundary $\Gamma_{\rm AM}$ by
-$\mathcal{F}_{\rm AM}$, on the inner ventricular interface
-$\Gamma_{\rm LV}$ by $\mathcal{F}_{\rm LV}$, on the boundary toward
-the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$ and on the
-boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
-$\mathcal{F}_{\rm SSAS}$. \discuss{For convenience, we denote $\mathcal{F}_e =
-\mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV} \cup \mathcal{F}_{\rm
-  AM-U} \cup \mathcal{F}_{\rm SSAS}$, and the union of all facets in
-$\mathcal{T}$ is denoted by $\mathcal{F} = \mathcal{F}_i \cup
-\mathcal{F}_e \discuss{\cup \mathcal{F}_{\rm SC} \cup \mathcal{F}_{\rm AM-L}}$.}
+interior facets (i.e.~triangular faces of the tetrahedral mesh cells)
+in $\mathcal{T}_{\rm PAR}$ and $\mathcal{T}_{\rm CSF}$ are denoted by
+$\mathcal{F}_{i, \rm PAR}$ and $\mathcal{F}_{i, \rm CSF}$,
+respectively. We define the union of facets interior both to
+$\mathcal{T}_{i, \rm PAR}$ and $\mathcal{T}_{i, \rm CSF}$ as
+$\mathcal{F}_i = \mathcal{F}_{i, \rm PAR} \cup \mathcal{F}_{i, \rm
+  CSF}$. Facets lying on the pial interface $\Gamma_{\rm pia}$ are
+denoted by $\mathcal{F}_{\rm pia}$, on the outer (arachnoid) boundary
+$\Gamma_{\rm AM}$ by $\mathcal{F}_{\rm AM}$, on the inner ventricular
+interface $\Gamma_{\rm LV}$ by $\mathcal{F}_{\rm LV}$, on the boundary
+toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$ and
+on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
+$\mathcal{F}_{\rm SSAS}$. \discuss{For convenience, we denote
+  $\mathcal{F}_e = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
+  \cup \mathcal{F}_{\rm AM-U} \cup \mathcal{F}_{\rm SSAS}$, and the
+  union of all facets in $\mathcal{T}$ is denoted by $\mathcal{F} =
+  \mathcal{F}_i \cup \mathcal{F}_e \discuss{\cup \mathcal{F}_{\rm SC}
+    \cup \mathcal{F}_{\rm AM-L}}$.}
 
@@
   %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
-  V_h^k = \{c \in L^2(\discuss{\mathcal{T}}): \;  c \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}\} ,
+  V_h^k = \{c \in L^2(\Omega): \;  c \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}\} ,
 \end{equation}
-where $L^2(\mathcal{T})$ is the space of square-integrable functions
-defined relative to $\mathcal{T}$ and $\mathbb{P}^k(E)$ denotes the
-space of polynomials of degree $\leqslant k$ defined over the
-tetrahedra $E$. To discretize the diffusion term in~\eqref{eq:3d_pde},
-we use a symmetric weighted interior penalty DG formulation, referring
+where $L^2(\Omega)$ is the space of square-integrable functions
+defined over $\Omega$ and $\mathbb{P}^k(E)$ denotes the space of
+polynomials of degree $\leqslant k$ defined over the tetrahedra
+$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
+symmetric weighted interior penalty DG formulation, referring
 to~\cite{ern2009discontinuous} and \cite[Section
@@
 \end{equation}
-Since $\bm V_h \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm v_n] = 0 $ on $\mathcal{F}_{\rm CSF}$.
+Since $\bm V_h \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$.
 Continuity in the tangential component is enforced weakly via interior penalization.
@@
   = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
-  - \sum_{F \in \discuss{\mathcal{F}_{\rm CSF} \cup \mathcal{F}_{e}}} \left (
+  - \sum_{F \in \discuss{\mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}}} \left (
   \int_{F} \mu \{\nabla \bm u\} \cdot \bm n_F \cdot [\bm v_t] 

### 2024-11-27 paper/supplementary_main.tex
Move some facet notation definitions to where they are needed. Still need to sort out the details for the Stokes flow
Former-commit-id: 61d724ce8109ac4d1afc5a4ef0eed3ea7fcf6a6e [formerly 61d724ce8109ac4d1afc5a4ef0eed3ea7fcf6a6e [formerly 9a035bc844e88182c4f65a712129b9f1871bd6aa]]
Former-commit-id: 8f1e5a788ed7ec5874b9287bcee44aabf43b5fe1
Former-commit-id: 9d49eeb9c59ddc949f4d612bb9fbbf2348985446

@@
 
-\subsubsection{Computational mesh entities: definitions and notation}
+\subsubsection{Computational mesh entities: definitions and common notation}
 
@@
 $\mathcal{F}_i = \mathcal{F}_{i, \rm PAR} \cup \mathcal{F}_{i, \rm
-  CSF}$. Facets lying on the pial interface $\Gamma_{\rm pia}$ are
-denoted by $\mathcal{F}_{\rm pia}$, on the outer (arachnoid) boundary
-$\Gamma_{\rm AM}$ by $\mathcal{F}_{\rm AM}$, on the inner ventricular
-interface $\Gamma_{\rm LV}$ by $\mathcal{F}_{\rm LV}$, on the boundary
-toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$ and
-on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
-$\mathcal{F}_{\rm SSAS}$. \discuss{For convenience, we denote
-  $\mathcal{F}_e = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
-  \cup \mathcal{F}_{\rm AM-U} \cup \mathcal{F}_{\rm SSAS}$, and the
-  union of all facets in $\mathcal{T}$ is denoted by $\mathcal{F} =
-  \mathcal{F}_i \cup \mathcal{F}_e \discuss{\cup \mathcal{F}_{\rm SC}
-    \cup \mathcal{F}_{\rm AM-L}}$.}
+  CSF}$.
 
@@
 Continuity in the tangential component is enforced weakly via interior penalization.
-Define the form \mer{MER: Check $\Gamma_{\rm AM}$ part.
-26.11.24: Rami will check $\mathcal{F}_e$ versus definitions of the boundaries and where it appears.}
+Facets lying on the pial interface $\Gamma_{\rm pia}$ are
+denoted by $\mathcal{F}_{\rm pia}$, on the outer (arachnoid) boundary
+$\Gamma_{\rm AM}$ by $\mathcal{F}_{\rm AM}$, on the inner ventricular
+interface $\Gamma_{\rm LV}$ by $\mathcal{F}_{\rm LV}$, on the boundary
+toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$ and
+on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
+$\mathcal{F}_{\rm SSAS}$.
+\discuss{For convenience, we denote
+  $\mathcal{F}_e = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
+  \cup \mathcal{F}_{\rm AM-U} \cup \mathcal{F}_{\rm SSAS}$, and the
+  union of all facets in $\mathcal{T}$ is denoted by $\mathcal{F} =
+  \mathcal{F}_i \cup \mathcal{F}_e \discuss{\cup \mathcal{F}_{\rm SC}
+    \cup \mathcal{F}_{\rm AM-L}}$.}
+Now, define the form 
 \begin{multline}
@@
   \mathcal{A}_h(\bm u_h, \bm v_h)
-  + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
+  + \sum_{F \in \discuss{\mathcal{F}_{\rm AM-U}}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
   - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0, \qquad \foralls \bm v_h \in \bm V_h, \\ 
@@
 \end{subequations}
+
+\mer{MER: Stopped here.}
 

### 2024-11-27 paper/supplementary_main.tex
Update notation and correct AM-U til AM-L in sum in (29)
Former-commit-id: e62a1fa2212d20cc17982642b75f944865b720d3

--- 2024-11-27 paper/supplementary_main.tex
@@
 \end{align*}
-Here, $H(\mathrm{div}, \Omega)$ is the space of $L^2(\Omega)$ vector fields with $L^2(\Omega)$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, and $\bm n$ and $\bm \tau$ are the unit outward normal and unit tangent vectors to each facet, respectively.
+\mer{MER: How about the homogenized version of $\bm V_h$?
+- to distinguish for test space?} Here, $H(\mathrm{div}, \Omega)$ is the space of $L^2(\Omega)$ vector fields with $L^2(\Omega)$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, and $\bm n$ and $\bm \tau$ are the unit outward normal and unit tangent vectors to each facet, respectively.
 Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by
@@
 \end{equation}
-Since $\bm V_h \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$.
-Continuity in the tangential component is enforced weakly via interior penalization.
-Define the form \mer{MER: Check $\Gamma_{\rm AM}$ part.
-26.11.24: Rami will check $\mathcal{F}_e$ versus definitions of the boundaries and where it appears.}
+Since $\bm V_h \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm
+  v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$. Continuity in the
+tangential component is enforced weakly via interior penalization. \mer{For
+convenience, we collect all facets exterior to the CSF space but not
+on the (Neumann) boundary $\Gamma_{\rm AM-U}$ in the set
+$\mathcal{F}_{e} = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
+\cup \mathcal{F}_{\rm AM-L} \cup \mathcal{F}_{\rm SSAS}$, where facets
+lying on the pial interface $\Gamma_{\rm pia}$ are denoted by
+$\mathcal{F}_{\rm pia}$, on the lower and upper outer (arachnoid) boundary
+$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$, on the inner ventricular
+interface $\Gamma_{\rm LV}$ by $\mathcal{F}_{\rm LV}$,
+%on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$
+and on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
+$\mathcal{F}_{\rm SSAS}$.} Now, define the form 
 \begin{multline}
@@
   = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
-  - \sum_{F \in \discuss{\mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}}} \left (
+  - \sum_{F \in \mer{\mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}}} \left (
   \int_{F} \mu \{\nabla \bm u\} \cdot \bm n_F \cdot [\bm v_t] 
@@
 \end{multline}
-where \discuss{$h$ denotes the element size}.
+where \discuss{$h$ denotes the element size} and on exterior facets \mer{the average and jump operators take the one-sided values}.
 We set the penalty parameter for the tangential continuity to be $\sigma = 20$.
@@
 
-\rami{  CSF pulsates in the SAS and ventricular system with each cardiac and respiratory cycle with large spatial differences.  This leads to dispersion effects. For example, 
+\rami{CSF pulsates in the SAS and ventricular system with each cardiac and respiratory cycle with large spatial differences.  This leads to dispersion effects. For example, 
 Sharp et al \cite{keith2019dispersion} estimated the enhancement of diffusion in the spinal subarachnoid space due to dispersive effects to average 5.8 times that of purely molecular diffusion with a theoretical maximum of $10^6$.

### 2024-11-27 paper/supplementary_main.tex
Wrap up in Supplementary Methods for now.
Former-commit-id: bc37098f781cc5c6edb3804605a3dc8712f1bbb6

@@
 that are connected to two or more edges and the set of end nodes as
-$\mathcal{T}$, write $\hat{c}^i = \hat{c}|_{\Lambda^i}$, and let
-$\mathcal{E}(y)$ denote the set of (one, two or three) edges in
-$\{\Lambda^i\}$ sharing the node $y$. To impose continuity of the
-concentrations, we set that the concentrations when viewed from each
-edge must match at nodes:
+$\mathcal{N} \subset \mathcal{V}$, write $\hat{c}^i =
+\hat{c}|_{\Lambda^i}$, and let $\mathcal{E}(y)$ denote the set of
+(one, two or three) edges in $\{\Lambda^i\}$ sharing the node $y$. To
+impose continuity of the concentrations, we set that the
+concentrations when viewed from each edge must match at nodes:
 \begin{equation}
@@
 
-Finally, we assume no molecular efflux at all end nodes $\mathcal{T}$
+Finally, we assume no molecular efflux at all end nodes $\mathcal{N}$
 via a homogeneous Neumann condition augmenting~\eqref{eq:1d_pde}:
 \begin{equation}
-    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{T}.
+    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{N}.
     \label{eq:1d_bcs}
@@
 in a perivascular network $\Lambda$, with edges $\{\Lambda^i\}_i$,
-internal nodes $\mathcal{B}$ and end nodes $\mathcal{T}$, we consider
+internal nodes $\mathcal{B}$ and end nodes $\mathcal{N}$, we consider
 the following system of hydraulic network
@@
 \end{align}
+\label{eq:1d_flow:bifurcations}
 \end{subequations}
@@
 \begin{equation}
-  \hat{p}(y) = p_0(y) \qquad y \in \mathcal{T} .
+  \hat{p}(y) = p_0(y) \qquad y \in \mathcal{N} .
+  \label{eq:1d_flow:bcs}
 \end{equation}
@@
   %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
-  V_h^k = \{c \in L^2(\Omega): \;  c \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}\} ,
+  V_h^k = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^k(E), \; E \in \mathcal{T}\} ,
 \end{equation}
 where $L^2(\Omega)$ is the space of square-integrable functions
-defined over $\Omega$ and $\mathbb{P}^k(E)$ denotes the space of
+defined over $\Omega$ and $\mathrm{P}^k(E)$ denotes the space of
 polynomials of degree $\leqslant k$ defined over the tetrahedra
@@
   \hat V_h
-  = \{ v \in C^0(\Lambda), \; v \vert_{\Lambda^i} \in \mathbb{P}^1(\Lambda^i) \}, 
+  = \{ v \in C^0(\Lambda), \; v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda^i) \}, 
 \end{equation}
-where $\mathbb{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$.
+where $\mathrm{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$.
 
@@
   \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\rm AM-L} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} \text{ on } \Gamma_{\mathrm{LV}} \}  \\ 
-  Q_h  &= \{q \in L^2(\Omega_{\rm CSF}): \; q \vert_E \in \mathbb{P}^1(E),
+  Q_h  &= \{q \in L^2(\Omega_{\rm CSF}): \; q \vert_E \in \mathrm{P}^1(E),
   \; E \in \mathcal{T}_{\rm CSF} \}. 
@@
   + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
-  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0, \qquad \foralls \bm v_h \in \bm V_h, \\ 
-\int_{\Omega_{\rm CSF}} \nabla \cdot \bm u_h \, q_h  &= 0, \qquad \foralls q_h \in Q_h.
+  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0 \qquad \foralls \bm v_h \in \bm V_h, \\ 
+\int_{\Omega_{\rm CSF}} \nabla \cdot \bm u_h \, q_h  &= 0 \qquad \foralls q_h \in Q_h.
 \end{align}
@@
 
+\subsubsection{Finite element solution of the perivascular network equations}
+
+We consider meshes $\mathcal{I}_a, \mathcal{I}_v$ representing a
+conforming subdivision of each of the perivascular networks
+$\Lambda_a$, $\Lambda_v$. Relative to each mesh $\mathcal{I}$, we define the
+space of (discontinuous) piecewise constants $\hat{Z}_h(\mathcal{I})$ and define
+spaces of continuous piecewise linears $\hat{V}_{h, g}(\mathcal{I})$ with
+prescribed boundary node values (on $\mathcal{N}$) given by $g$:
+\begin{align}
+  \hat{Z}_h(\mathcal{I}) &= \{z \in L^2(\Lambda) \, | \, z \vert_{\Lambda_i} \in \mathrm{P}^{0}(\Lambda_i) \, \foralls \Lambda^i \in \mathcal{I}\}, \\
+  \hat{V}_{h, g}(\mathcal{I}) &= \{ v \in C^0(\Lambda) \, | \, v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda_i), \, v(x) = g(x) \, \foralls x \in \mathcal{N} \} .
+  %Q_h^{\rm PVS} &= \{ q_h \in C^0(\Lambda),  q_h \vert_{\Lambda^i} \in \mathrm{P}^k(\Lambda_i), q_h \vert_{\Gamma_{\rm Pia}} = p_{\rm CSF} \}
+\end{align}
+To discretize~\eqref{eq:1d_flow} with the bifurcation
+conditions~\eqref{eq:1d_flow:bifurcations} and boundary
+conditions~\eqref{eq:1d_flow:bcs}, we use the space of continuous
+functions $\hat{Z}_h(\mathcal{I})$ to enforce the continuity of the pressure at
+bifurcation points, while the conservation of flux is enforced
+(weakly) through the variational formulation.
+The discrete variational form of the equations then reads: Find $(\hat{q}, \hat{p}) \in \hat{Z}_h(\mathcal{I}) \times \hat{V}_{h, p_0}(\mathcal{I})$ such that
+\begin{subequations}
+\begin{align}
+  \langle \hat{q}, z \rangle +  \langle z, \frac{A \kappa}{\mu} \partial_s \hat{p}  \rangle &= 0 \quad \forall z \in \hat{Z}_h(\mathcal{I}) \\
+  \langle \hat{q}, \partial_s w \rangle &= 0 \quad  \forall w \in \hat{V}_{h, 0}(\mathcal{I}).
+\end{align}
+\end{subequations}%
+where $\langle \cdot, \cdot \rangle$ denotes the $L^2(\mathcal{I})$-inner product and defined segment-wise.
+For a stability and convergence analysis of this discrete model, we refer to the reference\cite{gjerde2024directional}.
+
 \mer{MER: Stopped here.}
-
-\subsubsection{PVS flow}
-
-To enforce conservation of flux and continuity of pressure on bifurcation points, we approximate the velocity with a piecewise discontinuous polynomials and the pressure with continuous piecewise polynomials, respectively.
-The finite element spaces thus are:
-
-\begin{align}
-   V_h^{\rm PVS} = \{v_h \in L^2(\Lambda): v_h \vert_{\Lambda_i} \in
-   \mathbb{P}^{k -1}(\Lambda_i)\} \text{ and }
-   Q_h^{\rm PVS} = \{ q_h \in C^0(\Lambda),  q_h \vert_{\Lambda^i} \in \mathbb{P}^k(\Lambda_i), q_h \vert_{\Gamma_{\rm Pia}} = p_{\rm CSF} \}
-\end{align}
-
-The weak form reads: Find $(\hat{u}, \hat{p}) \in V_h^{\rm PVS} \times Q_h^{\rm PVS}$ such that
-
-\begin{align}
-(A \hat{u}, v) + \left(v, \frac{A \kappa}{\mu} \partial_s \hat{p}  \right) &= 0 \quad \forall v \in V_h^{\rm PVS} \\
-(A \hat{u}, \partial_s q ) &= 0 \quad  \forall q \in  Q_h^{\rm PVS}
-\end{align}
-
-where $\tau$ is the tangent vector of the network.
-Here we employ the lowest order discretization and set $k=1$.
-For discussion of stability of the discrete model we refer to \cite{gjerde2024directional}.
 
@@
 \label{sec:sup:peristalsis}
+
+\mer{MER: This needs a pass to update to new notation and new style.}
 

### 2024-11-27 paper/ideas.tex
Clean-up and shuffle.
Former-commit-id: 5fdf3caa5f44975b8abfa821b108e6866cf9bb64

\subsection*{Intracranial solute influx and clearance during sleep}
\begin{itemize}
\item
  Sleep is known to affect a number of the transport characteristics including: increased extracellular volume fraction, altered vascular pulsatility and a-forteriori perivascular flow and mixing, reduced CSF production, reduced glymphatic transport.
\end{itemize}

\subsection*{Clearance and the brain's waterscape}
\begin{itemize}
\item 
  Modelling metabolite clearance rather than solute influx (Model
  C). Consider several diffusion coefficients corresponding to
  gadubutrol, dextran, tau and amyloid-beta, otherwise identical flow
  set-ups.
\end{itemize}

\subsection*{Influx and clearance in pathologies}
\begin{itemize}
\item
  A number of neurological and neurodegenerative disorders yield mechanical changes in the intracranial environment (increased arterial stiffness, cerebral arterial angiopathies, perivascular flow changes due to hypertension, altered CSF flow patterns, altered diffusion properties in cancer, BBB leakage). One key clinical aspect is increased PVS width in e.g. iNPH. 
\end{itemize}

\subsection*{Effect of BBB}

Leakage to the blood, some sink/non-zero permeability for the PVS-blood (BBB) interface (Model N).

### 2024-11-27 paper/main.tex
Clean-up and shuffle.
Former-commit-id: 5fdf3caa5f44975b8abfa821b108e6866cf9bb64

--- 2024-03-12 paper/main.tex
@@
 \usepackage{subcaption}
-%\usepackage{natbib}
+% \captionsetup[subfigure]{justification=justified,singlelinecheck=false}
+\usepackage{tikz}
+\usepackage{newtxmath} % MER: Prettier math
+\usepackage{cleveref} % MER: Allows \Cref (but use Section X.Y not Subsection X.Y)
+\usepackage{booktabs} % MER: Prettier tables
+
+\usepackage{graphicx}
+\usepackage{subcaption}
+\usepackage{tikz}
+\usepackage{float}
+\usepackage{amsmath}
+\usepackage{siunitx}
+\usepackage{placeins}
+\usetikzlibrary{quotes}
+\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,positioning,fit,petri}
+
 \graphicspath{{figures/}}
 
-\title{Molecular transport in human pial perivasculature}
+\captionsetup[subfigure]{justification=justified,singlelinecheck=false}
 
-\author[1,x]{Rami Masri}
-\author[1,x]{Miroslav Kuchta}
-\author[1,x]{Marius Causemann}
-\author[1,*]{Marie E. Rognes}
+\title{In-silico molecular enrichment and clearance of the human intracranial space}
+\author[1]{Marius Causemann}
+\author[1]{Miroslav Kuchta}
+\author[2]{Rami Masri}
+\author[1,3,*]{Marie E. Rognes }
 \affil[1]{Department of Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway}
-\affil[x]{Author order to be discussed.}
+\affil[2]{Brown University, Providence, Rhode Island, USA}
+\affil[3]{K. G. Jebsen Centre for Brain Fluid Research}
 \affil[*]{meg@simula.no}
@@
 \newcommand{\rami}[1]{\textcolor{blue}{#1}}
+\newcommand{\mer}[1]{\textcolor{magenta}{#1}}
+\newcommand{\mar}[1]{\textcolor{violet}{#1}}
+\newcommand{\discuss}[1]{\textcolor{red}{#1}}
+\newcommand{\fixme}[1]{\textcolor{red}{#1}}
+\newcommand{\draft}[1]{\textcolor{gray}{#1}}
+%\newcommand{\brain}{\Omega_{\rm{PAR}}} 
+%\newcommand{\sas}{\Omega_{\rm CSF}}
+%\newcommand{\spinal}{\Gamma_{\rm SAS}}
+%\newcommand{\skull}{\Gamma_{\rm{skull}}}
+%\newcommand{\gin}{g_{\rm{influx}}}
+
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\foralls}{\forall \,}
 
@@
 
+\mer{MER: Target length: 5000 words, 4-6 figures: 600 words Intro. 1400 words Methods. 2000 words Results. 1000 words Discussion.}
+
 \section*{Introduction}
@@
 
+Motivation: as indicated by~\cite{iliff2012paravascular}, referring to
+Cserr, Phys Rev., 1971, bulk convective flow may be more important for
+larger species (larger distances) and larger molecules.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Results}
+\mar{~ 2000 words and 7 subsections - around 300 words each?}
+\begin{enumerate}
+\item
+  ``A digital twin'' for molecular enrichment and clearance in human intracranial spaces
+\item
+  CSF flow, the role of production versus no/cardiac/respiratory
+  dispersive effects on tracer trajectories, flow bifurcation. With and without tentorium cerebellum.
+\item
+  ``The role'' of PVS flow speed
+\item
+  ``The role'' of PVS sheaths and permeability - does functional separation entail structural separation?
+\item
+  ``The role'' of PVS size (motivated by pathologies)
+\item
+  Clearance from the CSF spaces - with and without low pressure veins, but keeping the intrathecal source of solute. 
+\item
+  Clearance from the parenchyma: $c_0 = 1.0$. Different substances?
+\end{enumerate}
+
+\begin{figure}
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{figures/Brain-PVS-callouts.png}
+     \end{subfigure}
+     \hfill
+     \caption{a) Illustration of the different intracranial domains
+       considered with labels;}
+     \label{fig:intracranial_domains}
+\end{figure}
+
+\subsection*{High-fidelity in-silico predictions of intracranial solute transport and exchange}
+
 \begin{itemize}
-\item Vinje et al~\cite{vinje2021brain} and references therein illustrates the location and characteristics of human pial perivascular spaces.
-\item Mestre et al~\cite{mestre2022periarteriolar} study the properties of pial perivascular spaces in detail in mice.
-\item Mestre et al~\cite{mestre2018flow} measure and characterize pial perivascular transport and estimate flow magnitudes (in mice).
 \item
-  Weller and coauthors conducted pioneering studies of perivascular spaces, flow and transport~\cite{zhang1990interrelationships, zhang1992directional}.
-They identified thin sheaths of pial cells surrounding arteries and arterioles (but not veins or venules) on the brain surface and within the human brain itself~\cite{zhang1990interrelationships}.
-Further, they observed that tracers spread along perivascular (arterial, venous and capillar) spaces in (rat) grey matter, and in the subarachnoid space (to the cribriform plate and nasal lymphatics)..
-\item In a series of papers, Bakker and coauthors study perivascular anatomy and solute transport~\cite{bedussi2017paravascular, bedussi2018paravascular}.
-They find that the subarachnoid space, the cisterns, ventricles and penetrating periarteriolar spaces form a continuous cerebrospinal-fluid filled space surrounding and penetrating the rat brain\cite{bedussi2017paravascular}.
-Moreover, they demonstrate pulsatile and directional (antegrade) flow in perivascular (predominantly periarterial) spaces on the brain surface.
+  Present mathematical and computational model and geometries (\Cref{tab:scenarios}). 
 \item
-  The architecture of pial and subarachnoid perivascular spaces in the rat brain was also studied by Thorne and colleagues~\cite{pizzo2018intrathecal, hannocks2018molecular}.
-\item Bollman et al~\cite{bollmann2022imaging} image the human pial vasculature at high resolution (7T).
-\item Hornkjøl et al~\cite{hornkjol2022csf} model CSF flow and solute transport in the human SAS and brain parenchyma. 
-\item Mardal et al~\cite{mardal2022mathematical} give a succinct introduction to MRI-based computational brain modelling. 
+  Highlight computational model availability, open source code.
 \item
-  Rey et al~\cite{rey2023perivascular} (from the group of Sartinoranont) model and study perivascular transport in rat brain networks.
+  Provides an interactive in-silico platform for studying intracranial solute transport -- distribute such that anyone can run (possibly try to recruit MinRK if advantageous).
+\item
+  \mer{MER: Remember to report on how much concentration enters into the brain?}
 \end{itemize}
 
-\section*{Methods}
+\subsection*{CSF flow and dispersion shifts enrichment patterns}
 
-\subsection*{Meshing}
+Flow of CSF through the SAS and the ventricular system is a crucial driver of transport in the human brain environment.
+While cardiac- and respiratory induced oscillatory flow increases mixing, the production of CSF in the choroid plexus creates a steady advective flow within the CSF-filled spaces of the cranium \cite{hornkjol2022csf}.
+We construct a 3D model of the SAS and the ventricular system (\Cref{fig:csf}A,B and C) and compute CSF flow fields for three different scenarios: CSF production in the choroid plexus, CSF flow at systolic peak blood inflow and CSF flow in response to free breathing.
 
-Hodneland et al~\cite{hodneland2019new} developed a method to image the brains' microvasculature based on time-of-flight (ToF) and quantitative susceptibility mapping (QSM) to identify arterial and venous networks.
+% production flow simulation
+To simulate the production-driven flow, we impose an inflow of $400\,$ml/day \cite{nilsson1992circadian} across the surface of the lateral ventricle and allow a resistive outflow across the upper convexity of the arachnoid mater, leading to a total pressure drop of $26\,$mPa between the production and outflow sites (\Cref{fig:csf}D).
+A maximum flow velocity of $1.8\,$mm/s is attained in the aqueduct, while the mean velocity in the SAS is $2.6\,\mu$m/s (\Cref{fig:csf} E).
 
-We derive a detailed tetrahedral mesh of both the parenchyma and the surrounding CSF-filled spaces alongside a network representation of the vasculature from the anatomical data provided therein.
+%The Stokes system is solved with the finite element software $FEniCS$, stored and then subsequently read for all the relevant solute transport simulations.
+See Figure~\ref{fig:csf} for a streamline visualization of the obtained velocity profile and \Cref{sec:details_numerical_method} for details on the numerical algorithm employed.
 
-\paragraph{3D domain}
+% cardiac pulsatility
+Assessing the effect of cardiac-driven dispersion, we compute a steady CSF flow field at peak systolic blood inflow.
+We mimic the resulting reduction of CSF space volume by imposing an inflow of $6\,$ml/s \cite{baledent2014imaging, causemann2022human} and $0.31\,$ml/s \cite{vinje2019respiratory} across the upper arachnoid mater ($\Gamma_{\rm AM-U}$) and the lateral ventricle ($\Gamma_{\rm VL}$) surface, respectively.
+We allow free outflow into the spinal SAS ($\Gamma_{\rm SSAS}$).
+In addition, we upscale the obtained pressure field to account for the lack of pulsatile inertial effects in our model (see \ref{sec:dispersion_details}).
+In this scenario, we observe a pressure drop of $10\,$Pa between lateral ventricles and the spinal SAS and a maximum flow velocity of $19.8\,$cm/s in the aqueduct (\Cref{fig:csf}F and J).
+Next, we estimate a spatially varying dispersion enhancement factor $R$ from cardiac-induced pulsatility by combining the computed pressure field with the theory of shear-augmented dispersion (see \Cref{sec:dispersion_details}).
+Assuming a cardiac frequency of $1\,$Hz, find that cardiac-induced pulsatile CSF flow increases the effective diffusion by up to two orders of magnitude in the aqueduct and around the cisterna magna, but has little effect in most of the SAS ($R < 1$) (\Cref{fig:csf} G and K).
 
-We merge the white and grey matter masks provided by \cite{hodneland2019new} to extract the pial surface.
-As no data on the CSF-filled spaces is available, we smooth and extend the pial surface by about 5\,mm to obtain an approximation of the SAS and include the inner fluid-filled cavities such as the ventricular system.
-Finally, we generate a tetrahedral mesh of both compartments using the fTetWild meshing algorithm \cite{hu2020fast}.
+% respiratory pulsatility
+Employing the same methodology, we estimate the respiratory contribution to dispersive mixing by computing the CSF flow field driven by an inflow of $0.121\,$ml/s \cite{liu2024using} and $1\,$ml/s across the lateral ventricular wall and the arachnoid mater, respectively.
+This amounts to a respiratory peak flow volume of $1.121\,$ ml/s at the craniocervical junction \cite{gutierrez2022effect}.
+While the resulting flow velocities are only about one fourth of their cardiac-induced counterparts (see \Cref{fig:csf}H and L), respiratory dispersive mixing reaches a factor of up to 350 due the lower respiratory frequency ($0.25\,$Hz) (\Cref{fig:csf}I and M).
 
-\paragraph{Vessel networks}
+% tracer trajectory of the standard model
+Next, we use the computed advective velcity field and dispersion enhancement factors to predict the spreading of $0.5\,$mmol intrathecally injected Gadobutrol in CSF-filled spaces, the parenchymal tissue and PVSs around arteries and veines.
+One hour post-injection, tracer moves upwards in the SAS frontally of the brainstem,  and quickly reaches supratentorial regions (\Cref{fig:csf}N).
+Here, it spreads both posteriorly through the longitudinal fissure and anteriorly through the outer SAS, reaching the top of the cerebral cortex after around 12 h.
+After 24h, the tracer covers most of the brains surface (with exception of some posterior regions) and has penetrated several centimeters into the parenchymal tissue.
 
-We skeletonize the binary masks of both the arterial and venous network using Kiminaro \cite{william_silversmith_2021_5539913} and obtain networks consisting of centerline and radius information of all detected vessels.
-The resulting networks are further smoothened using a spline interpolation technique and the nodes with lowest z-coordinate are identified as the networks' root nodes.
-
-\subsection*{Geometries and domains}
-
-The computational domain $\Omega$ is defined by the union of the SAS $\Omega_{\rm sas}$ and brain parenchyma $\Omega_{\rm brain}$.
-The thin pia membrane encloses the brain, and thus the pial surface (of the brain) defines the SAS-brain interface $\Gamma_{\rm pia}$.
-A network of blood vessels are embedded in the SAS, potentially very near to or on the pial surface.
-We assume that a network of idealized (annular or semi-annular) perivascular spaces surround the vasculature.
-We represent the vasculature and associated perivasculature by a network of one-dimensional curves $\Gamma$ representing the network centerlines equipped with inner $R_1$ and outer $R_2$ radius information.
-
-Surface representations of the outer SAS boundary, pial membrane, and delimiters of other CSF-filled spaces such as the lateral ventricles, third and fourth ventricle, and aqueduct are extracted from MR-images of human subject(s), see e.g.
-the pipeline described by Hornkjøl et al.~\cite{hornkjol2022csf}.
-From these surface representations, we create volumetric meshes of the SAS and brain using SVMTK~\cite{mardal2022mathematical}.
+% tracer trajectory of model variations
 
 \begin{itemize}
-    \item To discuss: How to generate subject-specific representations of the pial/SAS vasculature?
-Can we use contrast MRI?
-Can we use time-of-flight images?
-What are the right tools?
-Do any data sets already exist?
-    \item How to add radius information to vasculature representation? Do we need more information as well?
+    \item only dispersion model: No production flow.
+Tracer remains in the subtentorial regions, in cerebellum, brain stem and surrounding CSF.
+In contrast to other models, it travels up the ventricular system - similar to what is observed in iNPH patients?
+    \item Low dispersion model: account for uncertainty (or pathologies with impaired pulsatile motion?) - tracer exclusively spreads in the frontal half of the brain.
+Bifurcation of CSF production flow field: flow bifurcates around the midbrain in two directions: a) through the quadrigeminal cistern into the longitudinal fissure and b) frontally around the brain through the outer SAS.
+Only models with "enough" dispersion enable tracer to follow trajectory a).
+    \item without tentorium: 
 \end{itemize}
 
-\begin{figure}
-\caption{(A) Illustration of the location of pial vasculature and subarachnoid space.
-(B) Conceptual illustration of the vascular, perivascular, SAS and cortex domains (cross-section and lateral view).}
-\label{fig:concept}
+\begin{figure}[h!]
+\centering 
+\includegraphics[width = 0.9\linewidth]{figures/figure2.png}
+\caption{
+A) Illustration of 3D-domains; 
+B) Illustration of relevant boundaries; 
+C) schematic illustration of the model, its doamins and boundaries; 
+D) CSF flow induced by CSF production; 
+E) production-induced CSF velocities in the SAS and third and fourth ventricle (V3 \& V4); 
+F) CSF flow induced by peak systolic expansion; 
+G) cardiac dispersion enhancement factor;
+H) CSF flow induced by respiration; 
+I) respiratory dispersion enhancement factor;
+J) cardiac-induced flow velocities;
+K) cardiac-induced dispersion enhancement factor;
+L) respiratory flow velocities;
+M) respiratory dispersion enhancement factor;
+N) sagittal, coronal and axial planes of tracer concentration after 1, 6, 12 and 24$\,$h for model A. The outer model boundaries is shown in cyan, the pial membrane in purple and the arteries in red.;
+O) model X,Y and Z after 12h;
+}
+\label{fig:csf}
 \end{figure}
 
-\subsection*{Multi-dimensional transport model}
+\subsection*{The significance of perivascular spaces as flow highways}
+\begin{itemize}
+\item
+  Perivascular flow and transport has been identified as a key
+  mechanism and potential target for enhancing brain drug delivery and
+  metabolic waste clearance -- what are the global (intracranial)
+  effects of such intracranial highways? 
+\item
+  Do solutes move from PVS to the SAS, or both along the PVS and along
+  the SAS?
+\end{itemize}
+\begin{figure}
+    \centering
+    \includegraphics[height=\textheight]{figures/figure3.png}
+    \caption{\mer{MER: Put vasculature network illustration here.
+Many aspects actually.
+Marking of segments.
+Location relative to 3D.
+Radius etc.
+Marius will evaluate and suggest.
+Results: Group segments by logical order.
+For key arterial segments: average radius and length (can be moved to Supplementary if need be); pressure values; flow velocities; and then histograph with flow velocities for each mesh cell.
+Make sure to minorly note that ACA refers to a segment of the ACA and not the whole thing e.g.}}
+    \label{fig:pvs}
+\end{figure}
+\Cref{fig:pvs}.
 
-We consider a time-dependent convection-diffusion model~\cite{masri2023modelling} of a solute concentration $c$ in the network representing the pial vasculature and surrounding perivascular spaces $\Lambda$ and the surrounding three-dimensional SAS-brain domain $\Omega$ (Fig.~\ref{fig:concept}).
-
-\subsection*{Diffusion}
-We use the diffusion coefficient $D_{\rm gad}$ of Gadubutrol in free water in the SAS and PVS, and effective diffusion coefficient of Gadubutrol $D_{\rm brain}$ -- accounting for the extracellular space (ECS) porosity $\phi$ and the ECS tortuousity $\lambda$ but ignoring anisotropy  -- in the brain parenchyma~\cite{hornkjol2022csf}.
-
-We solve for two concentration fields $c, \hat c$ in the domain $\Omega$ and in the network $\Lambda$. The 
-network $\Lambda$ consists of 1D centerlines of vessels with their radii information and nodes. We consider the following model. 
-\begin{alignat}{2}
-\partial_t c - \nabla \cdot (D \nabla c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma & = f, && \quad \mathrm{in } \,\, \Omega \\ 
-\partial_t (A\hat c) - \partial_s(\hat D A  \partial_s \hat c) + \xi P (\hat c - \overline{c})   & = A \hat f, && \quad \mathrm{in } \,\,  \Lambda 
- \end{alignat}
-We split the boundary of $\Omega$ into $\Gamma_{\mathrm{skull}}$ and $\Gamma_{\mathrm{spine}}$,  and we enforce the following boundary conditions. 
-\begin{alignat}{2}
-D \nabla c \cdot \bm{n} &= 0, && \quad \mathrm{on} \,\, \Gamma_{\mathrm{skull}}, \\ 
-D \nabla c \cdot \bm{n} &= g(t),  && \quad \mathrm{on} \,\, \Gamma_{\mathrm{spine}}.
-\end{alignat}
-The function $g$ is given in 
-\begin{equation*}
-    g(t) = \max(0, a(T_0 - t))
-\end{equation*}
-For the network, the boundary nodes are split into $\partial \Lambda^{\mathrm{in}}$  and $\partial \Lambda^{\mathrm{out}}$ based on the specified orientation of the network (\color{blue} to be specified more precisely (with a figure?)\color{black}).
-We impose the following conditions.
-\begin{alignat}{2}
-D A \partial_s \hat c &= 0, && \quad \mathrm{on} \,\, \partial \Lambda^ {\mathrm{in}}, \\ 
-D A \partial_s  \hat c &= \hat g(t),  && \quad \mathrm{on} \,\, \partial \Lambda^{\mathrm{out}} .
-\end{alignat}
-\color{blue} Running different models depending on the choices of $g$ and $\hat g$ \color{black}
- \subsection*{Convection}
-We ignore the deformation of the domains (in this study), and thus set the domain velocity $w = 0$.
-We assume no significant convection in the brain $\Omega_{\rm brain}$, and set $u = u_{\rm brain} = 0$ there as a first approximation.
-We may consider non-zero convection in the SAS $u_{\rm sas} \not = 0$ e.g.~resulting from steady state CSF production as predicted by~\cite{hornkjol2022csf} in model variations, but may start with $u_{\rm sas} = 0$ as a baseline.
-In the perivascular network, we set an average axial velocity $u_{\rm pvs} = \langle u_{v, s} \rangle$ of $18.7$ $\mu$m/s as reported by Mestre et al.~\cite{mestre2018flow} in mouse perivascular spaces under normal physiological conditions.
-We may also consider a reduced perivascular velocity corresponding to hypertension in model variations.
-\subsubsection*{Steady state PVS flow model} We solve 1D Darcy equations in the vessel network.
-The unknowns are the averaged PVS velocity, $u_{\rm pvs} = \langle u_{v,s} \rangle$, and the cross-section pressure $p_{\rm pvs} $ (assumed to be constant in each cross-section).
-The system is given by  \cite{daversin2022geometrically}
-\begin{alignat}{2}
-A \,  u_{\rm pvs}   + \frac{\kappa}{\mu} \, A \, \partial_{s} p_{\rm pvs} & = 0, &&  \quad \rm in  \,\, \Lambda  \\ 
--\partial_s (A \, u_{\rm pvs}) & = f, && \quad \rm in  \,\, \Lambda .  
-\end{alignat} 
-In the above, $A$ is the area of the PVS, $\mu$ is the dynamic viscosity, and $\kappa$ is derived from the assumptio of Poiseuille
-flow in the annular cross-section of the PVS \cite{daversin2022geometrically,tithof2022network}: 
-\begin{equation}
-\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
-\end{equation}
-On bifurcations, conservation of fluxes is enforced weakly and continuity of the pressure is enforced strongly (in the choice of the finite element spaces).
-
-\subsubsection*{Case 1} We prescribe zero Dirichlet boundary conditions for the pressure, and vary $f$ \rami{include the reasoning behind having this $f$ as a source or sink term, and include the value that we opted for} to obtain a flow field with a similar magnititude to what is commonly reported in the literature.
-
-\subsection*{Case 2} We prescribe a Robin type boundary condition on the outlet nodes. In particular, we set 
-\begin{equation}
-A  u_{\rm pvs} =  - \frac{\kappa}{ \mu} A\, \partial_s p_{\rm pvs} = \alpha (p_{\rm pvs} - p_0), \quad  \rm on \,\,\, \partial \Lambda^{\rm out} 
-\end{equation}
-On $\partial \Lambda^{\rm in}$, we prescribe homogenous Dirichlet data for the pressure (\rami{does this make sense?}). 
-\rami{to be completed }
-\subsection*{PVS permeability}
-We will consider different values for the SAS-PVS permeability $\xi$, including $\xi = 0$ (sealed PVS), $\xi \approx \infty$ (PVS continuous with SAS), as well as reasonable intermediate values.
-We will ignore PVS-blood exchange, and thus only model the perivascular concentration at the network level (in addition to the SAS/brain concentration in the surroundings).
-
-\subsection*{Boundary and initial conditions}
-We set no flux on all boundaries unless otherwise specified.
-We start with an initial non-zero concentration in the first branch of the perivascular network and surrounding SAS (and zero concentration elsewhere).
-
-\subsection*{Parameter values}
-
-\begin{table}
-    \centering
-    \begin{tabular}{|c|c|c|c|c|} \hline 
- Parameter& Symbol& Value& Unit&Reference\\ \hline 
-         parenchyma app. diffusion&  $d_p$&  $1.2 \times 10^{-4}$& $\text{mm}^2/\text{s}$  & \cite{valnes2020apparent}\\\hline 
-         CSF app. diffusion&  $d_{CSF}$&  $3.8 \times 10^{-4}$& $\text{mm}^2/\text{s}$ & \cite{valnes2020apparent}\\\hline 
-         PVS diffusion&  $d_{PVS}$&  $3.8 \times 10^{-4}$& $\text{mm}^2/\text{s}$ & \cite{valnes2020apparent}\\\hline 
-         extracellular space volume fraction& $\phi$& 0.2& - &\cite{nicholson1981ion} \\\hline 
-         PVS-parenchyma permeability&  $\xi_{PVS,P}$& $5 \times 10^{2} - 6 \times 10^{3} $  & $\text{m}^{-1}$ & \cite{koch2023estimates} \\\hline 
-         PVS-CSF permeability&  $\xi_{PVS,CSF}$& $\{0, \text{inf}\}$ & $\text{m}^{-1}$ & \\ \hline 
- boundary concentration& $\hat{c}_D$& & &\\ \hline
-    \end{tabular}
-    \caption{Caption}
-    \label{tab:parameters}
-\end{table}
-
-\subsection*{Quantities of interest}
+\subsection*{Structural versus functional compartmentalization of perivascular spaces}
 
 \begin{itemize}
-    \item Snapshots (e.g. coronal plane) of concentration at relevant times (15 min, 30 min, 1h, 2h, ...) 
-    \item Total amount of tracer in the PVS, SAS and parenchyma
-     \item Time to reach a threshold (e.g. 30\%) in different "places"
+\item
+  Are perivascular spaces functionally or structurally
+compartmentalized? Human and rodent observations indicate that tracers
+concentrate in perivascular spaces surrounding the pial and
+subarachnoid vasculature. Must these spaces be structural i.e.~defined
+by semi-permeable structural
+membrane~\cite{zhang1990interrelationships, zhang1992directional,
+  mestre2018flow, eide2024functional}, or can such
+compartmentalization be a result of regions of enhanced flow or
+mixing~\cite{bedussi2017paravascular, vinje2021brain}? See also
+references in Background.
+\item
+  Vary the PVS permeability $\xi$.
 \end{itemize}
 
-\section*{Results}
-
-\begin{table}
-\begin{center}
-    \begin{tabular}{l|cc}
-    \toprule
-    Label & Description & Parameter values \\
-    Model 0 & Baseline &  \\
-    \bottomrule
-    \end{tabular}
-    \end{center}
-    \caption{Overview of models}
-\end{table}
-
-\begin{figure}
-     \centering
-     \begin{subfigure}[b]{0.33\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{modelA_total_conc.png}
-         \caption{arterial inflow, pure diffusion}
-         \label{fig:y equals x}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.33\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{modelB_total_conc.png}
-         \caption{SAS inflow, pure diffusion}
-         \label{fig:three sin x}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.33\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{modelC_total_conc.png}
-         \caption{arterial inflow, diffusion + PVS convection}
-         \label{fig:five over x}
-     \end{subfigure}
-        \caption{Total tracer amount in parenchyma, CSF, arterial PVS and venous PVS}
-        \label{fig:three graphs}
-\end{figure}
-
-\subsection*{Effect of interfaces (interface permeability)}
+\subsection*{Enlarged perivascular spaces yield ...}
 
 \begin{itemize}
-    \item Model 0 (baseline): only diffusion, different diffusion coefficient in the parenchyma and SAS/PVS.
-No leak to vasculature (zero permeability).
-$\infty$ $\zeta_0$ at the PVS-SAS interface, $\zeta_1$ at the PVS-SAS interface.
-Koch et al could be a reference for PVS-parenchyma, maybe Tithof et al (2022, "A network model" ...) could give some ideas for parameters, also see Mestre et al, Nature comms, 2022.
-    \item 
-    With altered interface permeabilities in the PVS (Model 1), and or in the parenchyma (Model 2) 
+  \item Take $R_2 = 3 R_1$ instead of $2 R_1$.
 \end{itemize}
 
-\begin{figure}
-    \caption{}
-    \label{fig:1}
-\end{figure}
+\subsection*{The perivenous contribution to molecular clearance of CSF tracers}
 
-\subsection*{Effect of convection}
 \begin{itemize}
 \item
-    With convective flow in the PVS (zero flow in the parenchyma for now)
+  Look at clearance phase 12h-24h-48h-72h.
+\item
+  Look at concentration in the veins
+\item
+  Look at concentration in the parenchyma
+\item
+  Consider low-pressure sink for the perivenous spaces.
+\end{itemize}
+
+\subsection*{Clearance of intraparenchymal solutes}
+
 \begin{itemize}
 \item
-        Model 3: PVS flow equal to 18.7 $\mu$m/s
-        \item 
-        Model 4: Better PVS flow (distributed, according to conservation of mass, assumptions on PVS area)
-    \end{itemize}    
-    \item 
-    With convective flow in the ECS (out of scope here, for follow-up work?)
+  Clearance of a molecular concentration in the parenchyma, different
+  molecular sizes - ranging from Gadubutrol to amyloid-$\beta$/$\tau$.
 \end{itemize}
 
-\begin{figure}
-    \centering
-    \includegraphics[width = 0.9 \textwidth]{modelB_modelC_overview.png}
-    \caption{Tracer concentrations for a purely diffusive model with tracer inflow via the CSF-filled space (left) and a convective model with Tracer inflow via the arterial PVS (middle) and their difference (right) at 1, 6, 12, 18 and 24 hours after injection.}
-    \label{fig:2}
-\end{figure}
-\subsection*{Effect of BBB}
+\bigskip
 
-Leakage to the blood, some sink/non-zero permeability for the PVS-blood (BBB) interface (Model N).
+\FloatBarrier
+\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Discussion}
 
-\begin{figure}
-    \caption{}
-    \label{fig:3}
-\end{figure}
+\paragraph{CSF flow model validation}
 
-\subsection*{Effect of pia}
+Hornkjøl et al \cite{hornkjol2022csf} model CSF flow induced by production in the choroid plexus and report a peak velocity of $8.9\,$mm/s in the aqueduct.
+Given that similar flow rates were enforced, the deviation from our value of $2.0\,$mm/s is likely due to differences in diameter of the reconstructed aqueduct.
+Additionally, their peak velocities in the SAS range between 20 - 50 $\mu\,$m/s, which is compatible with our mean velocity of $4.4\,\mu$m/s (\Cref{fig:csf_flow_histo}).
 
-Effect of pia permeability (Model N-1)
+The aqueductal ($0.31\,$ml/s) and craniocervical ($8.31\,$ml/s) flow rates predicted by our model align well with previously reported values of $0.31\,$ml/s peak aqueductal flow rate \cite{vinje2019respiratory} and about $9\,$ml/s peak net blood inflow into the cranium \cite{baledent2014imaging}.
+Further, Linninger et al \cite{linninger2007cerebrospinal} model cardiac-induced CSF flow and predict a maximum transmantle pressure difference of $10\,$Pa, which is in good agreement with our estimate of $8.5\,$Pa as well as clinically measured mean pressure differences of $11.5\,$Pa between sensors placed subdurally and in the lateral ventricle \cite{vinje2019respiratory}.
+% 10.7 cm/s CSF flow at the craniovertebral junction in control \cite{bapuraj2016cerebrospinal}, area difference?
 
-\begin{figure}
-    \caption{}
-    \label{fig:4}
-\end{figure}
-
-\section*{Discussion}
+Sharp et al used the theory of dispersive mixing to predict a dispersion enhancement factor of $91.8$ in the (open) spinal SAS, which lies well within the range of our spatially varying estimate (ranging from 0 to 160).
+Conversely, \cite{ray2021quantitative} report up to 10,000 times greater diffusion in perivascular regions, but their modelling lumps diffusive, dispersive and advective transport in a single enhancement factor and is thus not directly comparable.
 
@@
 \begin{itemize}
-    \item Pulsating brain. Natural to consider in follow-up work.
-    \item PVS-blood exchange. Natural to consider in model variation here, or follow-up work.
-    \item Accounting for brain diffusive anisotropy. Little effect expected. Maybe something to consider in small follow-up work.
+\item
+  \mer{MER/MC: How about dispersion in the PVS?}
+\item 
+  Inaccuracies in vascular geometry representation - and consequences downstream for e.g. numerical integration.
+\item
+  PVS geometry assumptions and approximations. Humans different from mice. Limitations in resolution.
+\item
+  No vasculature in the cerebellum, and of course no vasculature below certain resolution.
+\item
+  Only partially accounting for pulsatility. 
+\item
+  Accounting for brain diffusive anisotropy. Little effect expected. Maybe something to consider in small follow-up work.
+\item
+  Permeability and stiffness of the pial membrane
+\item
+  PVS-blood exchange and BBB leakage.
+Natural to consider in model variation here, or follow-up work.
+We here ignore PVS-blood exchange, and thus only model the perivascular concentration at the network level (in addition to the SAS/brain concentration in the surroundings).
+
 \end{itemize}
+
+\subsection*{Outlook}
+
+\begin{itemize}
+\item
+  Considering the (maybe) low diffusion in CSF spaces, maybe pure
+  advection would be relevant? Both computational and physiological
+  implications/interest.
+\end{itemize}
+
+\newpage
+\section*{Methods}
+
+%\mer{MER: I recommend writing each (sub)section of Methods as a nearly stand-alone paragraph.
+Let's see if we can avoid subsubsections/paragraphs.
+If you need to add more detail than you think is appropriate, put it in a Supplementary Methods section in the Appendix.}
+
+\subsection*{Intracranial compartments: ventricular system, SAS, and brain parenchyma}
+
+From T1-weighted MR images of a human subject~\cite{hodneland2019new},
+we first automatically segment the brain parenchyma and CSF spaces,
+including the ventricular system and SAS, using
+Synthseg~\cite{billot2023robust,billot2023synthseg}. We next manually
+adjust the segmentation to accurately represent the connections
+between CSF spaces (aqueduct, median aperture), smoothen, and finally
+extract surface representations of the outer (arachnoid) boundary, the
+pial membrane and other interfaces. Conforming to these surface and
+interface representations, we generate a tetrahedral mesh $\Omega$
+representing the complete intracranial volume as the union of the
+parenchyma $\Omega_{\rm PAR}$ and CSF spaces $\Omega_{\rm CSF}$, using
+fTetWild~\cite{hu2020fast}. The resulting computational mesh consists
+of $153\,729$ vertices and $851\,904$ mesh cells, which vary between
+$0.4$ mm and $1.2$ cm in mesh cell diameter. The parenchyma has a
+total volume of $1386$ ml, whereas the ventricles and the outer SAS
+contribute $28$ ml and $277$ ml, respectively, to a total CSF volume
+of $305$ml, in agreement with recent
+estimates~\cite{hladky2024regulation}.
+
+%% In addition, we slightly modify the segmentation to ensure correct
+%% connectivity of the ventricular system. In particular, we elongate
+%% the aqueduct and enforce no communication between the ventricles
+%% and the surrounding CSF spaces except for the median aperture.
+%% After morphological smoothing of the segmented image, we extract
+%% surface representations of the outer SAS boundary, the pial
+%% membrane and the components of the ventricular system.
+
+%% Note that the widely reported value of $150$\,ml of intracranial
+%% CSF volume is likely an underestimate. Recent estimates indicate
+%% that a value of $200$ to $300$\,ml is more accurate (see
+%% \cite{hladky2024regulation} and references therein
+
+\subsection*{Periarterial and perivenous spaces}
+
+To represent surface networks of periarterial and perivenous spaces,
+we use Kiminaro~\cite{william_silversmith_2021_5539913} to separately
+skeletonize time-of-flight angiography (ToF) and quantitative
+susceptibility mapping (QSM) images from the same human
+subject~\cite{hodneland2019new}. This technique yields networks of
+one-dimensional curves, each curve $\Lambda^i$ indicating the
+centerline of a blood vessel segment and labeled with its lumen radius
+$R_1^i$. We also assign an outer radius $R_2^i > R_1^i$ to each
+segment $i$, form the annular cylinder ensheathing $\Lambda^i$ and
+define the union of these as the PVS. The periarterial network graph
+consists of $12\,708$ edges connected at $12\,562$ inner nodes and
+with $147$ end nodes, and the associated domain is denoted by
+$\Lambda_a$. We identify three of the end nodes -- corresponding to
+the two internal carotid arteries (ICAs) and the basilar artery (BA)
+-- as root nodes, and designate the other ends as leaf nodes. The
+resulting perivenous network graph consists of $\fixme{N}$ edges
+connected at $\fixme{N}$ inner nodes and with $\fixme{N}$ end nodes,
+and the associated domain is denoted by $\Lambda_v$. \fixme{Finally, we label
+the outer surface of the PVS associated with $\Lambda_a$ and
+$\Lambda_v$, by $\Gamma_a$ and $\Gamma_v$ respectively.}
+
+%% The resulting vascular networks are further smoothened using a
+%% spline interpolation technique, and represented as graphs with
+%% nodes connected by edges.
+
+\subsection*{Molecular transport equations}
+
+We model diffusion, convection, and exchange of a molecular solute in
+the PVS networks $\Lambda_a$ and $\Lambda_v$, in the CSF spaces
+$\Omega_{\rm CSF}$ and in the brain parenchyma $\Omega_{\rm PAR}$ via
+a mixed-dimensional transport model~\cite{masri2023modelling} over a
+timescale of minutes to hours. Specifically, for $t > 0$, we solve for
+a concentration $c = c(x, t)$ in the 3D intracranial compartments ($x
+\in \Omega_{\rm CSF}, x \in \Omega_{\rm PAR}$) and for a cross-section
+averaged concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial
+and perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$) such that
+the following equations hold:
+\begin{subequations}
+\begin{alignat}{2}
+  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma & = 0 && \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}},
+  \label{eq:multi_transport_3d}
+  \\ 
+  \partial_t (A  \hat c) - \partial_s(\hat D A \partial_s ( \hat c)) +\partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c})  &= 0 && \quad \quad \mathrm{in} \quad  \Lambda_a, \Lambda_v .
+  \label{eq:multi_transport_1d}
+ \end{alignat}%
+\label{eq:multi_transport}%
+\end{subequations}%
+In~\eqref{eq:multi_transport}, $0 < \phi \leqslant 1$ is the fluid
+volume fraction, $D$ and $\hat{D}$ are effective diffusion
+coefficients, and $\bm u$ and $\hat{u}$ are fluid (CSF/ISF)
+velocities, all defined in the respective compartments; $P = 2 \pi
+R_2$ is the perimeter and $A = \pi (R_2^2 - R_1^2)$ the area of the
+PVS cross-sections, $\delta_{\Gamma}$ is the Dirac delta function
+concentrated on the interfaces $\Gamma_a, \Gamma_v$, the overlined
+$\overline{c}$ denotes the average of $c$ over cross-sections of
+$\Gamma_a, \Gamma_v$, and $\xi$ is a permeability allowing for
+transfer/exchange over the interfaces $\Gamma_a, \Gamma_v$ between the
+periarterial and perivenous networks and their surroundings.
+Moreover, we model the membrane between the parenchyma and the CSF
+spaces as a semi-permeable membrane with permeability cofficient
+$\beta_{\rm pia}$. For a complete description of this mathematical
+model, see \Cref{sec:more_maths}.
+
+\subsection*{Intracranial influx after intrathecal injection}
+To represent molecular influx into the intracranial compartments
+following intrathecal injection, without modelling the spinal
+compartment explicitly, we prescribe a time-dependent molecular influx
+over the interface towards the spinal compartment with the condition
+$(D \nabla c - \bm u c ) \cdot \bm{n} = g_{\mathrm{influx}}$ on
+$\Gamma_{\mathrm{SSAS}}$. We set
+\begin{equation}
+  g_\mathrm{influx}(t) = \frac{m_{\rm tot}}{T_{\rm max}^2 |\Gamma_{\rm SSAS}|} \max(0, T_{\rm max} - |t - T_{\rm max}|)
+\end{equation}
+which expresses a hat-shaped influx function with a peak at $T_{\rm max}$ and a total tracer injection of $m_{\rm tot}$ (\Cref{tab:parameters}).
+
+\subsection*{Molecular clearance from the intracranial space and perivascular network ends}
+%\subsection*{Boundary conditions: extracranial efflux and perivascular parenchymal influx}
+
+We model molecular clearance, into the meningeal lymphatics or other
+pathways, across the upper, outer (arachnoid) boundary
+(\Cref{fig:intracranial_domains}), proportional to the concentration
+in the SAS with a rate constant $\beta_{\rm exit}$. We set a no-flux
+condition at the end nodes of the perivascular networks, not allowing
+for direct molecular clearance there.
+
+\subsection*{CSF flow in the SAS and ventricular system}
+\label{sec:csf_fluid_vel}
+
+We model CSF as an incompressible Newtonian fluid at low Reynolds and
+Womersley numbers via the Stokes equations with viscosity $\mu$, and
+account for the convective contribution of the flow induced by CSF
+production in the choroid plexus. Specifically, we compute the
+velocity $\bm u_{\rm CSF}$ and associated pressure $p_{\rm CSF}$ in
+the SAS and ventricular system induced by a constant production at a
+rate of $u_{\rm in}$ across the lateral ventricle walls and with
+efflux across the upper, outer (arachnoid) boundary with efflux
+resistance $R_0$; and then set $\bm u = \bm u_{\rm CSF}$ in
+$\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport}
+(\Cref{tab:parameters}, \Cref{fig:intracranial_domains},
+\Cref{fig:csf_flow_prod}, see also \Cref{sec:more_maths}). We do not
+directly model convection within the brain tissue (beyond the
+perivascular spaces), and thus set $\bm u = 0$ in $\Omega_{\rm PAR}$
+in~\eqref{eq:multi_transport}.
+
+\subsection*{Dispersion in the CSF spaces}
+
+The pulsatile flow of CSF in the SAS and ventricular system,
+associated with the cardiac and respiratory cycles, substantially
+enhances molecular diffusion through
+dispersion~\cite{taylor1953dispersion, watson1983diffusion,
+  asgari2016glymphatic, keith2019dispersion, ray2021quantitative,
+  troyetsky2021dispersion}. To account for these dispersive effects,
+we compute a spatially-varying dispersion factor $R$ in $\Omega_{\rm
+  CSF}$ (\Cref{fig:csf_flow_R}) by combining a computational model of
+the CSF pressure gradient at peak arterial dilation with theoretical
+estimates for shear-augmented (Taylor) dispersion
+\cite{taylor1953dispersion, watson1983diffusion, keith2019dispersion}. \mer{Include 1-2 sentences from Supplementary Methods here -- but which? And also revisit references with MC.}
+
+\mer{MER: Stopped here. WIP in Supplementary Methods.}
+
+\subsection*{Dispersion in the PVS}
+
+Similar to the CSF spaces, dispersion due to cardiac and respiratory pulsation has been hypothesized to enhance transport in the perivascular spaces.
+However, recent estimates indicate that the effect of dispersion does not exceed a factor of 2 compared to diffusion alone \cite{asgari2016glymphatic,keith2019dispersion,bojarskaite2023sleep,asgari2016glymphatic,troyetsky2021dispersion}, and we thus do not model dispersive effects in the PVS.
+\mar{Maybe we should just take a factor of 2 as the upper bound?}
+
+% \cite{keith2019dispersion}: dispersion in PVS is negligiable in porous PVS, but might be relevant in open PVS (R=220) (theoretical estimates)
+%  \cite{bojarskaite2023sleep} numerical simulations show that the enhancement factor is likely in the range of 10 - 50 \%
+%\cite{asgari2016glymphatic} numerical simulations show increase of 16% to 50%  in distribution lengths
+%\cite{troyetsky2021dispersion}:   maximum enhancement of diffusion of only about 5%
+
+\subsection*{Perivascular fluid flow induced by CSF pressure differences}
+
+Since the pressure gradient the caused by the steady production-driven CSF flow (see \cref{sec:csf_fluid_vel}), we impose the precomputed fluid pressure $p_{\rm CSF}$ on the networks root nodes:
+\begin{equation}
+    \hat{p} = p_{\rm CSF} \quad \text{ on } \Lambda^{\rm out}
+\end{equation}
+In case that the root does not lie within the CSF-space but the parenchyma, the CSF pressure is not yet computed.
+For that reason, we apply a simple extension technique to map the CSF pressure field into $\Omega_{\rm PAR}$.
+In particular, we solve a Laplace equation on $\Omega_{\rm PAR}$ with the CSF pressure on the Pial membrane ($\Gamma_{\rm Pia}$) as a boundary condition.
+
+\subsection*{Perivascular fluid flow induced by vascular wall motion}
+
+Using the theoretical framework introduced by Gjerde et
+al.~\cite{gjerde2023directional}, we compute analytic estimates for
+the time-average perivascular flow rate $\langle Q' \rangle$ induced
+by rhythmic vascular wall motions for different frequencies $f$,
+amplitudes $\varepsilon$ and wave lengths $\lambda$ in the arterial
+network $\Lambda^a$ (see also Supplementary
+Methods~\ref{sec:sup:peristalsis}). The corresponding contribution to
+the convective velocity $\hat{u}_{\rm x}$ is defined vessel-wise as
+$\hat{u}_{\rm x}|_{\Lambda_i} = \langle Q'_i \rangle/A_i$ where $A_i$
+is the cross-section area of the PVS segment ($A_i = \pi (R_2^i -
+R_1^i$)) and $\langle Q_i' \rangle$ is its mean flow rate. Two wall
+motion patterns are considered: \emph{cardiac} pulsations with $f =
+1.0$ (Hz), $\lambda = 2000$ (mm) and $\varepsilon = 1\%$, and very
+low-frequency \emph{vasomotion} with $f = 0.1$ (Hz), $\lambda = 20$
+(mm), and $\varepsilon = 10\%$. \mer{The corresponding net convective
+velocities (in the antegrade direction) are for \emph{vasomotion}:
+$9.345 \pm 6.695$ $\mu$m/s, range: $(-65.81, 63.96) \mu$m/s, and for
+\emph{cardiac}: $0.475 \pm 0.5115 \mu$m/s, range: $(- 3.825, 4.861)
+\mu$m/s, where negative velocities correspond to retrograde flow
+(\Cref{fig:pvs}\mer{X}).} No perivascular fluid flow induced by wall
+motion is considered for the venous network.
+
+\subsection*{Model and material parameters}
+
+Model and material parameters are summarized in \Cref{tab:parameters}.
+We assume that the porosity equals the extracellular space volume
+fraction $\phi = \phi_{\rm PAR}$ within the parenchyma and $\phi = 1$
+elsewhere. In influx scenarios, we model the concentration of
+Gadubutrol with effective diffusion coefficient equal to that of
+Gadubutrol in free water in the PVS networks: $\hat{D} = D^{\rm Gad}$;
+weighted by a dispersion factor $R$ in the CSF spaces:
+$D|_{\Omega_{\rm CSF}} = ( 1 + R) D^{\rm Gad}$; and modulated by the
+porosity and tortuousity ($\lambda = 1.78$) in the parenchyma:
+$D|_{\Omega_{\rm PAR}} = D^{\rm Gad}_{\rm PAR}$. We set $\xi$
+depending on whether the PVS segment is fully embedded within the CSF
+spaces ($\xi = \xi_{\rm CSF}$) or parenchyma ($\xi = \xi_{\rm EF}$),
+or neither (\Cref{sec:more_maths}). Additionally, we increase the
+permeability $\xi$ ($100\times$) near the leaf nodes of $\Lambda_a$ to
+account for the transition from explicitly resolved PVSs to their
+unresolved continuation in the CSF spaces and parenchyma.
+
+\begin{table}
+  \begin{center}
+    \begin{tabular}{ll|ccc}
+      \toprule
+      Parameter& Symbol & Value & Unit& Reference\\
+      \midrule
+      Extracellular space volume fraction & $\phi_{\rm PAR}$ & 0.2 & -- & \cite{nicholson1981ion} \\
+      Free diffusion coefficient, Gadubutrol & $D^{\rm Gad}$ & $3.8 \times 10^{-4}$& $\text{mm}^2/\text{s}$ & \cite{valnes2020apparent}\\
+      Parenchymal diffusion coefficient, Gadubutrol & $D^{\rm Gad}_{\rm PAR}$ & $1.2 \times 10^{-4}$ & $\text{mm}^2/\text{s}$  & \cite{hornkjol2022csf} \\
+      Permeability coefficient, pia & $\beta_{\rm pia}$ & $2.6 \cdot 10^{-8}$ & m/s & \cite{riseth2025twocompartment} \\
+      Permeability coefficient, endfoot sheath & $\xi_{\rm EF}$ & $3.8\cdot 10^{-7}$  & m/s & \cite{koch2023estimates} \\
+      Molecular efflux rate & $\beta_{\rm exit}$ & $10^{-4}$ & mm$^2$/s & \cite{hornkjol2022csf} \\
+      Total molecular volume & $m_{\rm tot}$ & $0.5$ & mmol & \cite{eide2024functional} \\
+      Molecular influx, time-to-peak & $T_{\rm max}$ & $1$ & h & \fixme{$\ast$} \\
+      CSF production rate & $u_{\mathrm{in}}$ & $0.63$  & l / day & \cite{nilsson1992circadian} \\
+      CSF viscosity & $\mu$ & $0.7$ & mPa$ \cdot $s & \cite{bloomfield1998effects} \\ 
+      CSF outflow resistance & $R_0$ & $10^{-5}$  & Pa/(mm$\cdot$s) & \cite{hornkjol2022csf} \\ 
+      Dispersion factor, CSF spaces & $R$ & $\dagger$ & -- & $\ast$ \\
+      \bottomrule
+    \end{tabular}
+  \end{center}
+  \caption{Model parameters. $\dagger$/$\ast$ denotes computed/estimated within this work.}
+  \label{tab:parameters}
+\end{table}
+
+\subsection*{Initial conditions}
+
+We set the initial concentrations $c_0, \hat{c}_0$ in the
+three-dimensional domain $\Omega$ and networks $\Lambda_a, \Lambda_v$
+to be zero for all influx scenarios. For parenchymal
+clearance scenarios, we set $c_0 = 1, \hat{c}_0 = 0$.
+
+\subsection*{Numerical approximation of multi-dimensional transport and CSF flow}
+
+\mer{@Rami: Could you draft one paragraph here describing the essential bits of the numerical scheme?
+I recommend ``describing with words'' to reach relevant audiences.
+Refer to the Supplementary Methods for details.
+Mention theoretical expected convergence order.}
+
+In this section, we provide a brief overview on the employed numerical discretization.  
+We refer to \Cref{sec:details_numerical_method} for details. 
+Our implementation relies on the FEniCS finite element framework \cite{logg2012automated}; the 3D-1D coupling is handled by its extension FEniCS\textsubscript{ii} \cite{kuchta2020assembly}.
+The numerical scheme for the CSF flow equations utilizes polynomial spaces that preserve the divergence free condition \eqref{eq:divergence_equation} on the discrete level \cite{hong2016robust}.
+This requirement is important for stability in the transport equation, see for example \cite{cesmelioglu2022compatible}.
+The spaces we used for the velocity are continuous along the normal direction; continuity of the tangential component is enforced via interior penalty approaches \cite{hong2016robust}.
+If the exact solution of \eqref{eq:momnetum_equation}--\eqref{eq:divergence_equation} is smooth, then one expects optimal error convergence rates for this approximation.
+
+For the 3D transport problem, we use the interior penalty discontinuous Galerkin (DG) method with weighted averages and upwinding for the convection term \cite{ern2009discontinuous}.
+We remark that the choice of DG in the 3D domain is necessary here since  continuous Galerkin methods even when supplemented with artificial diffusion stabilization terms were unstable in our experiments.
+For the 1D transport problem, we use continuous Galerkin spaces defined over the entire network with additional artificial diffusion silization when the 1D velocity is non--zero.
+DG methods for the coupled 3D--1D problem have been recently analyzed  where convergence of the method for the elliptic case is established \cite{masri2024discontinuous}.
+While the analysis is for DG methods for both the 3D and the 1D problem, we expect similar arguments to hold here.
+However, note that the 3D--1D coupling induces a singularity in the 3D solution and one can only expect suboptimal error rates near the 1D network.
+In subdomains excluding the 1D lines, almost optimal rates can be expected for both DG methods \cite{masri2023discontinuous} and CG methods \cite{koppl2016local}.
+Extension of the convergence analysis to convection--diffusion problems is an interesting future research direction.
+
+\begin{table}
+  \begin{center}
+  \begin{tabular}{ll|llllll|ll}
+    \toprule
+    & Description & $R_1, R_2$ & $\xi_{\rm PVS-SAS}$ &  $\beta_{\rm efflux}$ & $D$ & $\hat{u}$ & $\mathbf{u}_{SAS}$ & $c_0$ \\
+    \midrule
+    A & Baseline with sheaths & $R_2 = 2 R_1$ & $\xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
+    B1 & Shealth permeability & $R_2 = 2 R_1$ & $0, \xi, 10 \xi, 100 \xi, 1000 \xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
+    B2 & Shealth permeability & $R_2 = 2 R_1$ & $0, \xi, 10 \xi, 100 \xi, 1000 \xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} +  \hat{u}_{\rm cardiac} + \hat{u}_{\rm vaso}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
+    %C & Larger PVS & $R_2 = N R_1$ & -- & -- & --  &  --  & -- &  -- \\
+    C & Higher PVS flow & -- & $\xi$ & -- & --  & $\hat{u}_{\rm prod} +  \hat{u}_{\rm cardiac} + \hat{u}_{\rm vaso}$  & -- & --  \\
+    AA & PVS size & $R_2 = x \beta R_1$ & $\xi$ & -- & -- & -- &  -- & 0 \\ 
+    %C & Larger PVS & $R_2 = N R_1$ & -- & -- & --  &  --  & -- &  -- \\
+    B & PVS flow & -- & $\xi$ & -- & --  & $\hat{u}_{\rm prod} + \hat{u}_{\rm vaso}$  & -- & --  \\
+    C & Clearance & -- & -- & -- & $\alpha D^{\rm gad, \tau, A\beta}$  & $\hat{u}_{\rm prod}$ & -- & 1 \\
+    %E & PVS II & $R_2 = 2 R_1$ & $\xi$, -- & -- & --  &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow\uparrow$  & -- & --  \\
+    %F & PVS III & $R_2 = 2 R_1$ & $\xi$, -- & -- & $10 D_{\rm PVS}$ &  $\hat{u} = \hat{u}_{\rm prod}$  & -- & --  \\
+    %G & Glymphatics & $R_2 = 2 R_1$ & $\xi$, -- & -- & \cite{sykova2008diffusion, valnes2020apparent} &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow$  & $\mathbf{u}_{\rm PAR}$ > 0 & --  \\
+    \bottomrule
+    \end{tabular}
+    \end{center}
+  \caption{Overview of models (-- denotes as immediately above). Each
+    PVS cross-section is modelled as a concentric annulus with inner
+    radius $R_1$ and outer radius $R_2$ and such that $R_2 = \beta
+    R_1$ with $\beta = 2.0$ based on~\cite{eide2024functional}. Let
+    $\xi = \dots$ denote the permeability estimated by Koch et
+    al.~\cite{koch2023estimates}. Model A and its variations aims to
+    quantify the effect of PVS sheaths by varying the permeability
+    $\xi$ between the PVS and its surroundings. We consider a fixed
+    permeability $\xi_{\rm PVS-PAR}$ based
+    on~\cite{koch2023estimates}. \mer{The permeability between the PVS
+      and the SAS $\xi_{\rm PVS-SAS}$ must likely be estimated by a
+      bit of trial-error from the observations of timings of PVS-SAS
+      and SAS from~\cite{eide2024functional}.} \discuss{The
+      extracranial solute efflux permeability $\beta_{\rm efflux}$ is
+      uniformly distributed over the outer SAS boundary with a
+      reasonable rate as baseline\cite{hornkjol2022csf,
+        eide2021clinical, ringstad2024glymphatic}.} At baseline, the
+    effective diffusion coefficients $D^{\rm x}_{\rm SAS} = D^{\rm
+      x}_{\rm PVS}$ of the SAS and PVS represent the free diffusion
+    coefficient of Gadubutrol (${\rm x} = {\rm gad}$) in CSF (water),
+    and $D^{\rm x}_{\rm PAR}$ the effective diffusion coefficient of
+    human cortical tissue, all at body temperature; weighted by a
+    dispersion factor $\alpha$ (see Methods). At baseline, we consider
+    net flow due to CSF-production in the SAS and PVS:
+    $\mathbf{u}_{\rm SAS} > 0$ and $\hat{u} > 0$ (see Methods) but no
+    bulk flow in the parenchyma $\mathbf{u}_{\rm PAR} = 0$. Due to the
+    timescale for human intracranial tracer transport (minutes to
+    hours) compared to typical (cardiac, respiratory, slow vasomotion)
+    human intracranial pulsatility (0.1 -- 1 Hz), we consider net
+    (constant-in-time) flow fields only in subsequent model variations
+    (but see dispersion factor above).  Model B represents
+    perivascular pathway scenarios with more rapid flow in the PVS
+    induced by vasomotion (PVS I)~\cite{gjerde2023directional}.
+    Models C represents a version of the aforementioned scenarios but
+    for studying metabolite clearance (instead of solute influx).}
+\label{tab:scenarios}
+\end{table}
 
@@
 \newpage
-\section*{Acknowledgements (not compulsory)}
+\section*{Acknowledgments (not compulsory)}
 
-Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.
+Acknowledgments should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.
 
@@
 
+\newpage
+\appendix
+\input{supplementary_main.tex}
+
+\input{background.tex}
+
 \end{document}

### 2024-11-27 paper/main.tex
Update on lengthgit status -mar
Former-commit-id: 2cfaaa6f5d37908d108703d045f2b1895cc0f6a3

@@
 \begin{abstract}
+``The abstract — which should be no more than 200 words long and contain
+no references — should serve both as a general introduction to the
+topic and as a brief, non-technical summary of the main results and
+their implications.'' 
 \end{abstract}
@@
 
-\mer{MER: Target length: 5000 words, 4-6 figures: 600 words Intro. 1400 words Methods. 2000 words Results. 1000 words Discussion.}
+\mer{MER: Target length: 5000 words (Intro, Results, Discussion) plus <3000 words Methods: 800 words Introduction. 2800 words Results. 1800 words Discussion.}
 
@@
 \section*{Results}
-\mar{~ 2000 words and 7 subsections - around 300 words each?}
+\mer{Update: ~2800 words and 7 subsections - around 300-450 words each?}
 \begin{enumerate}

### 2024-11-27 scripts/compute_dispersion_field.py
black and white plots / remove LV from histo
Former-commit-id: ba4a98fe2bf20e8744032cd27268910d5bb17a8b [formerly ba4a98fe2bf20e8744032cd27268910d5bb17a8b [formerly 6fc92bd183090fb714c0546ff662d6b8c29ed464]]
Former-commit-id: dd38e66de6cdf567fb5894668d60c58ae930053e
Former-commit-id: 35f4e7273eafad4db1870f77462691c5fb138b1d

@@ -9,17 +9,20 @@ from generate_synthseg_mesh import CSFID, LVID, V34ID
 import matplotlib.pyplot as plt
 import seaborn as sns
 from matplotlib.ticker import PercentFormatter
+import seaborn as sns
 
 def R_histo(sm, R, filename):
     set_plotting_defaults()
+    sns.set_context('paper', font_scale=1.8)
     DG0 = sm.function_space()
     cellvol = assemble(TestFunction(DG0)*dx)
-    ids = [CSFID, LVID, V34ID]
-    indices = [sm.vector()[:]==i for i in ids]  
-    fig, axes = plt.subplots(ncols=3, figsize=(7,3))
+    domains = [[CSFID], [V34ID],]
+    names = ["SAS", "V3 & V4"]
+    indices = [np.isin(sm.vector()[:], dom) for dom in domains]      
+    fig, axes = plt.subplots(ncols=len(domains), figsize=(7,3))
     nbins = 20
-    colors = sns.color_palette("crest", n_colors=3)
-    for ax, name, idx, col in zip(axes, ["SAS", "LV", "V3 & V4"], indices, colors):
+    colors = sns.color_palette("crest", n_colors=len(domains))
+    for ax, name, idx, col in zip(axes, names, indices, colors):
         Rvals = R.vector()[idx]
         xlims = (0, np.percentile(Rvals, 80))
         bins = np.logspace(start=np.log10(1e-1), stop=np.log10(200), num=nbins)
@@ -34,7 +37,7 @@ def R_histo(sm, R, filename):
         ax.set_xlabel(f"R")
         ax.set_ylabel("frequency")
         ax.set_xticks([0.1, 1, 10, 100])
-        plt.tight_layout()
+        #plt.tight_layout()
         #ax.set_xlim(xlims)
         ax.set_ylim((0, max(n)*1.5))
         ax.legend(frameon=False, loc="lower left", alignment="right",
@@ -53,32 +56,74 @@ def plot_R(R, filename):
     grid = pv.UnstructuredGrid(topology, cell_types, x)
     grid["R"] = R.vector()[:]
 
-    Rmin, Rmax = grid.clip()["R"].min(), grid.clip()["R"].max()
+    Rmin, Rmax = grid["R"].min(), grid["R"].max()
     bg = pv.read_meshio("mesh/standard/standard.xdmf")
-    pl = pv.Plotter(off_screen=True, window_size=(1600, 1600))
[…]

### 2024-11-27 scripts/evaluate_pvs_flow.py
prettify plots
Former-commit-id: afd810a943df022e04744075eb5bc50c6605effa [formerly afd810a943df022e04744075eb5bc50c6605effa [formerly 3c5497585c1de8afe7d1f72126dd7fec7b55f4ac]]
Former-commit-id: e6d9544d1b33972900adb6d62ec6ba814e5a07bd
Former-commit-id: f76f9d2b99cfbc3bcabe6e5a555ae5f59983f64e

@@ -63,7 +63,7 @@ def compute_pvs_flow(pvs_flow_file):
         umed = np.median(abs(uvals))
         range = np.round(minmax([uvals], percentile=99.9), 1)
         nbins = 50
-        fig, ax = plt.subplots()
+        fig, ax = plt.subplots(figsize=(4,3))
         counts, bins, containers = plt.hist(uvals, density=False, bins=nbins, histtype="bar",
                 range=range, edgecolor='black', linewidth=0.4, 
                         weights=lengths / lengths.sum())
@@ -83,7 +83,7 @@ def compute_pvs_flow(pvs_flow_file):
         plt.xlim(range)
         plt.legend(frameon=False)
         ax.yaxis.set_major_formatter(PercentFormatter(1))
-        plt.savefig(f"{plot_dir}/{model}_velocity_histo_{variant}.png")
+        plt.savefig(f"{plot_dir}/{model}_velocity_histo_{variant}.png", dpi=300)
 
     points = np.array([c for n,c in pointlabels])
     cellidx = map_kdtree(FunctionSpace(mesh, "DG", 0).tabulate_dof_coordinates(),
@@ -102,35 +102,35 @@ def compute_pvs_flow(pvs_flow_file):
         radii.append(assemble(m2mm*artery_radii*dxs(si))/ length)
         lengths.append(length*m2mm)
     df = pd.DataFrame({"p":pressures, "u":velocities, "loc":artlabels, "L":lengths,
-                        "R":radii}).sort_values(by="loc")
+                        "R":radii})
 
-    plt.figure()
-    ax = sns.barplot(df, x="loc", y="p", palette="crest")
+    plt.figure(figsize=(4,3))
+    ax = sns.barplot(df, x="loc", y="p",)# palette="crest")
     ax.set_xlabel("")
     ax.set_ylabel("pressure (mPa)")
-    plt.xticks(rotation=45); plt.tight_layout()
-    plt.savefig(f"{plot_dir}/{model}_arteries_labels_pressure.png")
+    plt.xticks(rotation=45,ha='right', rotation_mode='anchor', size="10");plt.tight_layout()
+    plt.savefig(f"{plot_dir}/{model}_pressure.png")
 
-    plt.figure()
-    ax = sns.barplot(df, x="loc", y="u", palette="flare")
+    plt.figure(figsize=(4,3))
[…]

### 2024-11-27 scripts/generate_comparison_barplots.py
minor changes
Former-commit-id: 56ed8902b9c07f721bd8dfaed39990345a3d5b1f

@@ -7,25 +7,10 @@ import seaborn as sns
 import os
 from pathlib import Path
 from plotting_utils import time_str, set_plotting_defaults
-from compute_dispersion_field import alpha
+from compute_dispersion_field import alpha_cardiac, alpha_respiratory
 import typer
 from typing import List
 
-csf_flow_model = {"LowRes":"sas_flow_LowRes",
-                "standard":"sas_flow",
-                "HighRes":"sas_flow_HighRes",
-                "dt=240s":"sas_flow",
-                "dt=120s":"sas_flow",
-                "dt=60s":"sas_flow",
-                "vasomotion":"sas_flow"}
-
-transport_model = {"LowRes":"modelALowRes",
-                "standard":"modelA",
-                "HighRes":"modelAHighRes",
-                "dt=120s":"modelA",
-                "dt=240s":"modelAlargedt",
-                "dt=60s":"modelAsmalldt",
-                "vasomotion":"modelAB"}
 
 mesh_refinement_models = ["standard", "vasomotion"]
 time_refinement_models = ["dt=240s", "dt=120s", "dt=60s"]
@@ -50,8 +35,11 @@ def compare_models(models: List[str]):
         # get PVS flow stats
         #md.update(read_config(f"results/csf_flow/{csfm}/pvs_metrics.yml"))
         # get cardiac PVS flow stats
-        cardiac_csf = read_config(f"results/csf_flow/cardiac_{csfm}/metrics.yml")
-        cardiac_csf["pmax"] *= (1 + alpha**2 / 8)
+        dispersion_file = tmconfig["csf_dispersion_file"]
+        if isinstance(dispersion_file, list): dispersion_file=dispersion_file[0]
+        csfm_cardiac = dispersion_file.split("/")[-2]
+        cardiac_csf = read_config(f"results/csf_flow/{csfm_cardiac}/metrics.yml")
+        cardiac_csf["pmax"] *= (1 + alpha_cardiac ** 2 / 8)
         md.update({f"cardiac_{k}":v for k,v in cardiac_csf.items()})
         # get mean concentrations
         md.update(read_config(f"results/{tm}/mean_concentrations.yml"))
@@ -106,7 +94,7 @@ def compare_models(models: List[str]):
 
     nfigs = len(barplot_groups)
     fig_width = [ 1  + int(len(bp["vars"]) / 4) for bp in barplot_groups]
-
+    from IPython import embed; embed()
[…]

### 2024-11-27 scripts/label_arteries.py
minor update
Former-commit-id: a286b2f55344b94b1b4f2045e401886d64a09789 [formerly a286b2f55344b94b1b4f2045e401886d64a09789 [formerly cb941dc1dd9eabd1eef8dc7a600b0e281fcef1d4]]
Former-commit-id: aad090353d6b2bef936d970ad46889876c29be67
Former-commit-id: 9bb1df1f3e7c54287ecc382b2f4b30aba4de1933

@@ -15,7 +15,7 @@ pointlabels = [("BA",(0.084971, 0.104599, 0.0424848)),
                ("ACA-A2", (0.0805165,0.115209,0.0781812)),
                ("ACA-A3", (0.0795565, 0.133436, 0.0905888)),
                ("ACA-A4", (0.0815724, 0.111484, 0.107046)),
-               ("ACA-A5", (0.0815178, 0.0984657, 0.167766)),
+               #("ACA-A5", (0.0815178, 0.0984657, 0.167766)),
                ("PER-R", (0.123073, 0.101526, 0.135256)),
                ("PER-L", (0.0401122, 0.0939796, 0.13324))
             ]
@@ -24,7 +24,7 @@ def get_tangent(v, n):
     return v - np.dot(np.dot(v,n), n)
 
 if __name__=="__main__":
-   filename = "plots/labeled_arteries.png"
+   filename = "plots/meshplots/labeled_arteries.png"
    art = pv.read("mesh/networks/arteries_tubes.vtk")
    labelcoords = lambda label: art.bounds[::2]
 
@@ -39,7 +39,7 @@ if __name__=="__main__":
                      position_y=0.0, title_font_size=24,
                      label_font_size=24, fmt="%.1f")
    pl = pv.Plotter(off_screen=True)
-   pl.add_mesh(art, cmap="coolwarm", scalar_bar_args=bar_args)
+   pl.add_mesh(art, cmap="coolwarm", scalar_bar_args=bar_args, show_scalar_bar=True)
    pl.add_points(np.array(points), render_points_as_spheres=False, 
                  point_size=20, color=col)
    center = art.center_of_mass() - np.array([0,0, 0.047])
@@ -60,6 +60,6 @@ if __name__=="__main__":
    pl.camera.azimuth += 160
    pl.camera.elevation += 7
    pl.camera.zoom(1.6)
-   pl.screenshot(filename, transparent_background=False)
+   pl.screenshot(filename, transparent_background=True)
    print(pl.camera.direction)
    pl.export_html(filename.replace("png", "html"))

### 2024-11-27 scripts/make_tubes.py
generate tubes from line

@@ -1,0 +1,72 @@
+
+from plotting_utils import get_result, read_config
+import numpy as np
+from extract_vessels import get_tubes
+import typer
+import pyvista as pv
+import meshio
+from meshio._vtk_common import vtk_to_meshio_type
+
+
+def get_cells(mesh):
+    if not isinstance(mesh, pv.UnstructuredGrid):
+        mesh = mesh.cast_to_unstructured_grid()
+
+    # Copy useful arrays to avoid repeated calls to properties
+    vtk_offset = mesh.offset
+    vtk_cells = mesh.cells
+    vtk_cell_type = mesh.celltypes
+
+    # Check that meshio supports all cell types in input mesh
+    pixel_voxel = {8, 11}  # Handle pixels and voxels
+    for cell_type in np.unique(vtk_cell_type):
+        if cell_type not in vtk_to_meshio_type.keys() and cell_type not in pixel_voxel:
+            raise TypeError(f'meshio does not support VTK type {cell_type}.')
+
+    # Get cells
+    cells = []  # type: ignore[var-annotated]
+    c = 0
+    for i, (offset, cell_type) in enumerate(zip(vtk_offset, vtk_cell_type)):
+        if cell_type == 42:
+            cell_ = mesh.get_cell(i)
+            cell = [face.point_ids for face in cell_.faces]
+            cell_type = f'polyhedron{cell_.n_points}'
+
+        else:
+            numnodes = vtk_cells[offset + c]
+            cell = vtk_cells[offset + 1 + c : offset + 1 + c + numnodes]
+            c += 1
+            cell = (
+                cell
+                if cell_type not in pixel_voxel
+                else cell[[0, 1, 3, 2]]  # type: ignore[call-overload]
+                if cell_type == 8
+                else cell[[0, 1, 3, 2, 4, 5, 7, 6]]  # type: ignore[call-overload]
+            )
+            cell_type = cell_type if cell_type not in pixel_voxel else cell_type + 1
+            cell_type = vtk_to_meshio_type[cell_type]
+
+        if len(cells) > 0 and cells[-1][0] == cell_type:
+            cells[-1][1].append(cell)
+        else:
+            cells.append((cell_type, [cell]))
+    return cells
+        
+
+
+def make_tubes(model:str):
+    config = read_config(f"configfiles/{model}.yml")
+    dt,T = config["dt"], config["T"]
+    times = np.arange(0, T + dt, dt*config["output_frequency"])
[…]

### 2024-11-27 scripts/make_video.py
generate movie

@@ -1,0 +1,70 @@
+from plotting_utils import get_result, read_config
+import pyvista as pv
+import typer
+import numpy as np
+from pathlib import Path
+from tqdm import tqdm
+from plot_csf_flow import from_k3d
+import k3d.colormaps.paraview_color_maps as pcm
+from extract_vessels import get_tubes
+
+def make_movie(modelname:str):
+    config = read_config(f"configfiles/{modelname}.yml")
+    dt , T= config["dt"], config["T"]
+    cmap = from_k3d(pcm.Yellow___Gray___Blue).reversed()
+    clim = (0,2)
+    times = np.arange(0, T + dt, dt*config["output_frequency"])
+    scalarbar_args = dict(color="white", title="concentration (mmol/l)", 
+                          vertical=False, height=0.08, width=0.6,
+                          position_x=0.2, position_y=0.05, 
+                          title_font_size=36, use_opacity=False,
+                          label_font_size=28, fmt="%.1f")
+    meshfile = Path(config["mesh"])
+    skull = pv.read(meshfile.parent / "surfaces/skull.ply")
+    ventricles = pv.read(meshfile.parent / "surfaces/ventricles.ply")
+    art = get_result(modelname, "artery", times).ctp()
+    sas = get_result(modelname, "sas", times)
+    arteries = get_tubes(art)
+    arteries["c"] = arteries["c_0"]
+    for t in times:
+        sas[f"c_{t}"] = np.where(sas[f"c_{t}"] > clim[1], 
+                                 clim[1], sas[f"c_{t}"])
+    img = pv.create_grid(sas, dimensions=(300, 300, 300))
+    img = img.sample(sas, progress_bar=True)
+
+    pl = pv.Plotter(off_screen=True, window_size=(1200, 1200))
+    pl.set_background("black")
+    pl.open_movie(f"plots/{modelname}/{modelname}.mp4", framerate=24,
+                  quality=8)
+    pl.add_mesh(skull,style='points',point_size=0.1, opacity=0.6, show_scalar_bar=False,
+                color="silver", specular=1, render_points_as_spheres=True)
+    pl.add_mesh(ventricles, style='points',point_size=0.8,
+                show_scalar_bar=False,specular=1,
+                render_points_as_spheres=True,
+                color="silver")
+    pl.add_mesh(arteries, scalars="c", cmap=cmap, 
+                clim=clim, show_scalar_bar=False,)
[…]

### 2024-11-27 scripts/mark_and_refine_mesh.py
fix ordering issue
Former-commit-id: 659148585a49138c57026690a34905d9239733d6 [formerly 659148585a49138c57026690a34905d9239733d6 [formerly f48a73285e34e69390519b1e9dc8345227b664eb]]
Former-commit-id: a662cec98943069b8be2a76b1f60f9d55045db08
Former-commit-id: 0c702911ce8aabb4cc9021f54fc373fbcb093379

@@ -95,10 +95,10 @@ def mark_boundaries(label):
     bm = df.MeshFunction("size_t", mesh, 2, 0)
 
     # set upper and lower skull
-    upper_skull_z = 0.05
+    upper_skull_z = 0.00
     lower_skull = df.CompiledSubDomain("on_boundary")
     lower_skull.mark(bm, LOWER_SKULL_ID)
-    upper_skull = df.CompiledSubDomain("on_boundary && x[2] > m",
+    upper_skull = df.CompiledSubDomain("on_boundary && (x[2] - 0.8*x[1]) > m",
                                      m=upper_skull_z)
     upper_skull.mark(bm, UPPER_SKULL_ID)
 
@@ -155,7 +155,7 @@ def mark_and_refine(configfile : str):
         crit = abs(get_surface_dist(sas, parenchyma_surf)) < 0.003
         print(sas.num_cells())
         sas, label = refine_sphere(sas, bottom_refine_coords, 0.08, label, 
-                                    criterion=crit, min_size=config["epsilon"] * 4)
+                                    criterion=crit, min_size=config["edge_length_r"] / 3)
         print(sas.num_cells())
 
     # refine inlet area
@@ -167,7 +167,17 @@ def mark_and_refine(configfile : str):
         print(sas.num_cells())
         crit = abs(get_surface_dist(sas, parenchyma_surf)) < 0.002
         sas, label = refine_sphere(sas, inlet_coords, 0.01, label, criterion=crit,
-                                    min_size=config["epsilon"]*2)
+                                    min_size=config["edge_length_r"] / 8)
+        print(sas.num_cells())
+
+    # refine AQ area
+    print("======================")
+    print("refining AQ area...")
+    aq_coords = (0.0815, 0.082, 0.0665)
+    for i in range(1):
+        print("======================")
+        print(sas.num_cells())
+        sas, label = refine_sphere(sas, aq_coords, 0.005, label)
         print(sas.num_cells())
 
     # refine interpeduncular
@@ -194,7 +204,10 @@ def mark_and_refine(configfile : str):
     for i,m in enumerate(cell_f.array()):
         colors.array()[cellmap[i]] = m
 
-    label.array()[colors.array()[:] >=2]  = CSFNOFLOWID
+    cols,  color_counts = np.unique(colors.array(), return_counts=True)
+
+    label.array()[np.isin(colors.array()[:],
[…]

### 2024-11-27 scripts/mean_concentrations.py
update plots
Former-commit-id: a64c6d58ec0524d48c27aa24d75c659979b3fd01 [formerly 0054ed41d7efd956a4d3c4e975d81a101844bc4c]
Former-commit-id: 9e2ad76cd0c7818fa7d7812ab36225116a7cbfcc

@@ -5,30 +5,41 @@ import typer
 from typing import List
 import matplotlib.pyplot as plt
 from plotting_utils import set_plotting_defaults, read_config, minmax
-from solver import pcws_constant, read_vtk_network, as_P0_function
+from solver import pcws_constant
 import numpy as np
 import seaborn as sns
 from IPython import embed
 import pandas as pd
-from matplotlib.markers import MarkerStyle
 import yaml
 from branch_marking import color_branches
 from label_arteries import pointlabels
 from test_map_on_global_coords_shift import map_kdtree
+from peristalticflow import mesh_to_weighted_graph, nx
+from generate_synthseg_mesh import CSFID, PARID, LVID, V34ID, CSFNOFLOWID
+import joypy
+from cmap import Colormap
+from scipy.stats import binned_statistic
 
-CSFID = 1
-PARID = 2
-LVID = 3
-V34ID = 4
-CSFNOFLOWID = 5
+def ii_project(f, V):
+    u = TrialFunction(V)
+    v = TestFunction(V)  
+    dx_a = Measure('dx', domain=V.mesh())
+    a  = inner(u,v)*dx_a  
+    L  = inner(f, v)*dx_a
+    A,b = map(ii_assemble, (a,L)) 
+    A,b = map(ii_convert, (A,b)) 
+    sol = Function(V) 
+    solve(A, sol.vector(), b) 
+    return sol
 
 def compare_concentrations(modelname:str):
     config = read_config(f"configfiles/{modelname}.yml")
     dt , T= config["dt"], config["T"]
     times = np.arange(0, T + dt, 3*dt*config["output_frequency"])
+    assert 3600 in times
     sas_conc, subd_marker = get_result_fenics(modelname, "sas", times)
-    art_conc, art_radii = get_result_fenics(modelname, "artery", times)
-    ven_conc, ven_radii = get_result_fenics(modelname, "vein", times)
+    art_conc, art_radii, art_roots = get_result_fenics(modelname, "artery", times, getroots=True)
+    ven_conc, ven_radii, ven_roots = get_result_fenics(modelname, "vein", times, getroots=True)
     pvs_ratio_artery = config["pvs_ratio_artery"]
     pvs_ratio_vein = config["pvs_ratio_venes"]
     subd_marker.array()[np.isin(subd_marker.array(),[LVID, V34ID, CSFNOFLOWID])] = CSFID
[…]

### 2024-11-27 scripts/overview_plot.py
generate plot with T1 background

@@ -1,0 +1,121 @@
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import ImageGrid
+import numpy as np
+import typer
+from typing import List
+from plotting_utils import (time_str, read_config,
+                    get_result, T1_clip_plot)
+from generate_synthseg_mesh import CSFID, CSFNOFLOWID, PARID, LVID, V34ID
+import pyvista as pv
+from cmap import Colormap
+import nibabel
+from extract_vessels import np2pv
+import matplotlib as mpl
+from extract_vessels import get_tubes
+from pyvista.core import _vtk_core as _vtk
+from pyvista.core.filters import _get_output, _update_alg
+from pyvista.core.utilities.helpers import generate_plane
+from mpl_toolkits.axes_grid1.inset_locator import inset_axes
+import seaborn as sns
+
+ndict = {"x":(1,0,0), "y":(0,-1,0), "z":(0,0,1)}
+
+percentile = 95
+
+def clip_closed_surface(surf, normal='x', origin=None, tolerance=1e-06,
+                        transfer_data=False):
+    if isinstance(normal, str): normal = np.array(ndict[normal])
+    if origin is None: origin = surf.center_of_mass()
+    plane = generate_plane(normal, origin)
+    collection = _vtk.vtkPlaneCollection()
+    collection.AddItem(plane)
+    alg = _vtk.vtkClipClosedSurface()
+    alg.SetGenerateFaces(True)
+    alg.SetInputDataObject(surf)
+    alg.SetTolerance(tolerance)
+    alg.SetClippingPlanes(collection)
+    _update_alg(alg, False, 'Clipping Closed Surface')
+    result = _get_output(alg)
+    if transfer_data: return result.sample(surf)
+    return result
+
+def generate_overview_plot(model:str, 
+                           times:List[int] = [3600, 21600, 43200, 86400]):
+    cbar_title = "concentration (mmol/l)"
+    m = model
+    conf = read_config(f"configfiles/{m}.yml")
+    mesh = pv.read_meshio(conf["mesh"])
+    t1data = nibabel.load("data/T1.nii.gz")
+    res = np.array(t1data.header["pixdim"][1:4])*1e-3
+    Timg = np2pv(t1data.get_fdata()[:,:-50, :], res, 
+                origin=(0,0,0))
+    sas = get_result(m, "sas", times)
+    art = get_result(m, "artery", times).ctp()
+    ven = get_result(m, "vein", times).ctp()
+    arteries = get_tubes(art)
[…]

### 2024-11-27 scripts/plot_boundaries.py
prettify plot

@@ -11,16 +11,16 @@ grid = pv.read_meshio("mesh/standard/standard_facets.xdmf")
 dirname = "plots/meshplots/"
 os.makedirs(dirname, exist_ok=True)
 
-bds = {"SK-U":UPPER_SKULL_ID,"SK-L": LOWER_SKULL_ID,
+bds = {"AM-U":UPPER_SKULL_ID,"AM-L": LOWER_SKULL_ID,
        "SSAS":SPINAL_OUTLET_ID,
         #"SC": SPINAL_CORD_ID, 
-        "LV":LV_INTERF_ID,  "Pia":PIA_ID,}
-clips = ["Pia", "SK-L", "SK-U"]
+        " LV":LV_INTERF_ID,  " Pia":PIA_ID,}
+clips = [" Pia", "AM-L", "AM-U"]
 
-bar_args=dict(title="", vertical=False, height=0.08, width=0.8, position_x=0.1,
-                        position_y=-0.05, title_font_size=52,
+bar_args=dict(title="", vertical=True, height=0.8, width=0.06, position_x=-0.035,
+                        position_y=0.1, title_font_size=52,
                         bold=False, font_family="times",
-                        label_font_size=44)
+                        label_font_size=72)
 print(grid.array_names)
 bounds = list(grid.bounds)
 bounds[0] = (bounds[0] + bounds[1]) * 0.5

### 2024-11-27 scripts/plot_csf_flow.py

@@ -1,4 +1,5 @@
 
+from xml import dom
 import pyvista as pv
 import typer
 from vtk_adapter import create_vtk_structures
@@ -6,7 +7,7 @@ import numpy as np
 from fenics import *
 import matplotlib 
 import k3d.colormaps.paraview_color_maps as pcm
-from compute_dispersion_field import alpha
+from compute_dispersion_field import alpha_cardiac, alpha_respiratory
 from plotting_utils import set_plotting_defaults, minmax
 from generate_synthseg_mesh import CSFID, LVID, V34ID
 import matplotlib.pyplot as plt
@@ -14,21 +15,26 @@ from matplotlib.ticker import PercentFormatter
 import seaborn as sns
 from plot_subdomains import get_camera
 from vtk.util.numpy_support import numpy_to_vtk
+import seaborn as sns
 
 m2mum = 1e6
 
 def velocity_histo(sm, u, filename, vscale, vunit):
     set_plotting_defaults()
+    sns.set_context('paper', font_scale=1.8)
+
     DG0 = sm.function_space()
     umag = project(sqrt(inner(u,u)), DG0,
                    solver_type="cg", preconditioner_type="hypre_amg")
     cellvol = assemble(TestFunction(DG0)*dx)
-    ids = [CSFID, LVID, V34ID]
-    indices = [sm.vector()[:]==i for i in ids]  
-    fig, axes = plt.subplots(ncols=3, figsize=(7,3))
+    domains = [[CSFID], [V34ID],]
+    names = ["SAS", "V3 & V4"]
+    indices = [np.isin(sm.vector()[:], dom) for dom in domains]  
+    #from IPython import embed; embed()
+    fig, axes = plt.subplots(ncols=len(domains), figsize=(7,3))
     nbins = 20
-    colors = sns.color_palette("crest", n_colors=3)
-    for ax, name, idx, col in zip(axes, ["SAS", "LV", "V3 & V4"], indices, colors):
+    colors = sns.color_palette("crest", n_colors=len(domains))
+    for ax, name, idx, col in zip(axes,names , indices, colors):
         uvals = umag.vector()[idx] * vscale
         xlims = (0, np.percentile(uvals, 92))
         ax.hist(uvals, bins=nbins, weights=cellvol[idx] / cellvol[idx].sum(),
@@ -41,10 +47,10 @@ def velocity_histo(sm, u, filename, vscale, vunit):
         #           color="black",ymax=0.6, label=f"{max*1e-3:.1f} mm/s")
[…]

### 2024-11-27 scripts/pvs_flow_prod.py

@@ -35,12 +35,13 @@ def get_blood_flow_orientation(tau, radius_f, root_marker):
     return uh_sign
 # --------------------------------------------------------------------
 
-def compute_pvs_flow(csf_flow_model:  str):
+def compute_pvs_flow(csf_flow_model:  str, network: str):
 
     radius_ratio = 2
-    mesh, artery_radii, artery_roots = read_vtk_network("mesh/networks/arteries_smooth.vtk", rescale_mm2m=False)
+    networkfile = f"mesh/networks/{network}_smooth.vtk"
+    mesh, artery_radii, artery_roots = read_vtk_network(networkfile, rescale_mm2m=False)
     pressure_file = f"results/csf_flow/{csf_flow_model}/flow.hdf"
-    results_dir = f"results/pvs_flow_prod/{csf_flow_model}/"
+    results_dir = f"results/pvs_flow_prod/{csf_flow_model}-{network}/"
     csf_flow_config = read_config(f"configfiles/{csf_flow_model}.yml")
     radius_f = as_P0_function(artery_radii)
     

### 2024-11-27 scripts/plot_subdomains.py
prettify plot
Former-commit-id: bd3d563c985127b137609962ff91657c9d7045d2 [formerly 91a780012743cedb522c94cadc72a5399c4f310f]
Former-commit-id: f9bbb08f0a2a07968f54bf72ee207f0cc466244f

@@ -26,15 +26,15 @@ if __name__=="__main__":
     LV = grid.extract_cells(grid["label"]==LVID)
 
     CSF["l"] = np.full(CSF["label"].shape, " SAS", dtype='<U10')
-    V34["l"] = np.full(V34["label"].shape, "V3 & V4", dtype='<U10')
+    V34["l"] = np.full(V34["label"].shape, "V3&4", dtype='<U10')
     LV["l"] = np.full(LV["label"].shape, "LV", dtype='<U10')
 
     comb = pv.merge([CSF.clip(), V34, LV])
 
-    bar_args=dict(title="", vertical=False, height=0.08, width=0.6, position_x=0.2,
-                            position_y=-0.05, title_font_size=52,
+    bar_args=dict(title="", vertical=True, height=0.8, width=0.06, position_x=-0.035,
+                            position_y=0.1, title_font_size=52,
                             bold=False, font_family="times",
-                            label_font_size=44)
+                            label_font_size=72)
 
     pl = pv.Plotter(off_screen=True, window_size=(1600, 1600))
     colors = sns.color_palette("Blues", n_colors=3)

### 2024-11-27 scripts/plotting_utils.py
prettify plot
Former-commit-id: 91a780012743cedb522c94cadc72a5399c4f310f

@@ -38,7 +38,7 @@ def get_result(modelname, domain, times):
         grid["radius"] = label.vector()[:]
     return grid
 
-def get_result_fenics(modelname, domain, times):
+def get_result_fenics(modelname, domain, times, getroots=False):
     from fenics import XDMFFile, FunctionSpace, Function
     from solver  import get_mesh, read_vtk_network, as_P0_function
     filename = f"results/{modelname}/{modelname}_{domain}.xdmf"
@@ -49,10 +49,10 @@ def get_result_fenics(modelname, domain, times):
         mesh, vol_subdomains = get_mesh(config["mesh"])
         V = FunctionSpace(mesh, "DG", 1)
     if domain == "artery":
-        mesh, radii, _ = read_vtk_network("mesh/networks/arteries_smooth.vtk", rescale_mm2m=False)
+        mesh, radii, roots = read_vtk_network("mesh/networks/arteries_smooth.vtk", rescale_mm2m=False)
         V = FunctionSpace(mesh, "CG", 1)
     if domain == "vein":
-        mesh, radii, _ = read_vtk_network("mesh/networks/venes_smooth.vtk", rescale_mm2m=False)
+        mesh, radii, roots = read_vtk_network("mesh/networks/venes_smooth.vtk", rescale_mm2m=False)
         V = FunctionSpace(mesh, "CG", 1)
     results = []
     print(f"reading results from {modelname} at")
@@ -64,7 +64,10 @@ def get_result_fenics(modelname, domain, times):
             f.read_checkpoint(c, "c", i)
             results.append(c)
     if domain in ["artery", "vein"]:
-        return results, as_P0_function(radii)
+        if getroots:
+            return results, as_P0_function(radii), roots
+        else:
+            return results, as_P0_function(radii)
     elif domain=="sas":
         return results, vol_subdomains
 
@@ -107,6 +110,7 @@ def read_config(configfile):
 def clip_plot(csf, par, networks, filename, title, clim, cmap, cbar_title):
     import pyvista as pv
     from extract_vessels import get_tubes
+    cmap = Colormap(cmap).to_matplotlib()
 
     csf_clipped = csf.clip(normal="y")
     par_clipped = par.clip(normal="y")
@@ -126,6 +130,20 @@ def clip_plot(csf, par, networks, filename, title, clim, cmap, cbar_title):
     pl.camera.zoom(1.3)
     pl.add_title(title, font_size=12)
[…]

### 2024-11-27 scripts/time_dependent.py
minor updates
Former-commit-id: 1a161e57a869b0909d0fcddb1d451aa9e9421aa6

@@ -38,6 +38,19 @@ def get_subdomain_dofs(V, subdomains, subd_id):
         [dm.cell_dofs(c.index()) for c in SubsetIterator(subdomains, subd_id)]))
     return subd_dofs
 
+def read_pvs_velocity(vel_file, netw):
+    from evaluate_pvs_flow import sig_to_space
+    print(f"reading arterial PVS velocity from {vel_file}")
+    with HDF5File(MPI.comm_world, vel_file,'r') as f:
+        sig = f.attributes("/u").to_dict()["signature"]
+        vel = Function(sig_to_space(sig, netw))
+        f.read(vel, "u")
+    # make sure the velocity field is tangential to the network
+    tau = xii.TangentCurve(netw)
+    un = vel - tau*inner(vel, tau)
+    assert assemble(inner(un,un)*dx) < 1e-16
+    return vel 
+
 
 def run_simulation(configfile: str):
 
@@ -65,6 +78,7 @@ def run_simulation(configfile: str):
     beta = config["molecular_outflow_resistance"]
     beta_csf_par = config["par_csf_permability"]
     root_xi_factor = config["root_pvs_permability_factor"]
+    c_init = config["initial_concentration"]
 
     vein, vein_radii, vein_roots = read_vtk_network("mesh/networks/venes_smooth.vtk", rescale_mm2m=False)
     vein_radii = as_P0_function(vein_radii)
@@ -117,34 +131,29 @@ def run_simulation(configfile: str):
     Ds = pcws_constant(vol_subdomains, {CSFID: sas_diffusion,  # csf
                                         PARID: parenchyma_diffusion # parenchyma
                                         })
-    art_xi_root = Function(FunctionSpace(artery, "CG", 1))
-    idx = map_kdtree(artery.coordinates(), FunctionSpace(artery, "CG", 1).tabulate_dof_coordinates())
-    art_xi_root.vector()[:] = np.where(artery_roots.array()[:] > 0, root_xi_factor, 1)[idx]
-    xi_a = pcws_constant(artmarker, art_xi_dict)
-    xi_a.vector()[:] *= interpolate(art_xi_root, xi_a.function_space()).vector()[:]
-    xi_v = pcws_constant(veinmarker, ven_xi_dict)
+    
+    def get_xi(marker, xi_dict, roots):
+        netw = marker.mesh()
+        netw_xi_root =  Function(FunctionSpace(netw, "CG", 1))
[…]

### 2024-11-27 paper/cut.tex

@@
+\mer{MER: Old, revisit to see if we have forgotten any of these interesting ideas.}
+\subsection*{Effect of interfaces (interface permeability)}
+
+\begin{itemize}
+    \item Model 0 (baseline): only diffusion, different diffusion coefficient in the parenchyma and SAS/PVS.
+No leak to vasculature (zero permeability).
+$\infty$ $\zeta_0$ at the PVS-SAS interface, $\zeta_1$ at the PVS-SAS interface.
+Koch et al could be a reference for PVS-parenchyma, maybe Tithof et al (2022, "A network model" ...) could give some ideas for parameters, also see Mestre et al, Nature comms, 2022.
+    \item 
+    With altered interface permeabilities in the PVS (Model 1), and or in the parenchyma (Model 2) 
+    \end{itemize}
+
+\subsection*{Effect of convection}
+\begin{itemize}
+    \item 
+    With convective flow in the PVS (zero flow in the parenchyma for now)
+    \begin{itemize}
+        \item 
+        Model 3: PVS flow equal to 18.7 $\mu$m/s
+        \item 
+        Model 4: Better PVS flow (distributed, according to conservation of mass, assumptions on PVS area)
+    \end{itemize}    
+    \item 
+    With convective flow in the ECS (out of scope here, for follow-up work?)
+\end{itemize}
+
+\begin{figure}
+    \centering
+    \includegraphics[width = 0.9 \textwidth]{modelB_modelC_overview.png}
+    \caption{Tracer concentrations for a purely diffusive model with tracer inflow via the CSF-filled space (left) and a convective model with Tracer inflow via the arterial PVS (middle) and their difference (right) at 1, 6, 12, 18 and 24 hours after injection.}
+    \label{fig:2}
+\end{figure}
+
+\subsection*{Effect of pia}
+
+Effect of pia permeability (Model N-1)
+
+\begin{figure}
+     \centering
+     \begin{subfigure}[b]{0.33\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{modelA_total_conc.png}
+         \caption{arterial inflow, pure diffusion}
+         \label{fig:y equals x}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.33\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{modelB_total_conc.png}
+         \caption{SAS inflow, pure diffusion}
+         \label{fig:three sin x}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.33\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{modelC_total_conc.png}
+         \caption{arterial inflow, diffusion + PVS convection}
+         \label{fig:five over x}
+     \end{subfigure}
+        \caption{Total tracer amount in parenchyma, CSF, arterial PVS and venous PVS}
+        \label{fig:three graphs}
+\end{figure}
+
+\begin{figure}
+\begin{subfigure}{0.5 \linewidth}
+    \centering
+    \includegraphics[width=1\linewidth]{figures/modelA_conc_at_label.png}
+    \caption{B1 - $\xi$}
+    \includegraphics[width=1\linewidth]{figures/modelB1-10_conc_at_label.png}
+    \caption{B1 - $10 \xi$}
+    \includegraphics[width=1\linewidth]{figures/modelB1-100_conc_at_label.png}
+    \caption{B1 - $100 \xi$}
+    \includegraphics[width=1\linewidth]{figures/modelB1-1000_conc_at_label.png}
+    \caption{B1 - $1000 \xi$}
+\end{subfigure}
+\begin{subfigure}{0.5 \linewidth}
+    \centering
+    \includegraphics[width=1\linewidth]{figures/modelB2-1_conc_at_label.png}
+    \caption{B2 - $\xi$}
+    \includegraphics[width=1\linewidth]{figures/modelB2-10_conc_at_label.png}
+    \caption{B2 - $10 \xi$}
+    \includegraphics[width=1\linewidth]{figures/modelB2-100_conc_at_label.png}
+    \caption{B2 - $100 \xi$}
+    \includegraphics[width=1\linewidth]{figures/modelB2-1000_conc_at_label.png}
+    \caption{B2 - $1000 \xi$}
+\end{subfigure}
+    \label{fig:modelB_xi_variations}
+\end{figure}
+
+\section{Foo}
+
 We consider non-zero convection in the CSF space $u_{\rm CSF} \not = 0$ in \eqref{eq:multi_transport_3d} resulting from steady state CSF production.

### 2024-11-27 paper/supplementary_main.tex

@@
 In particular, we enforce a uniformly distributed inflow of $u^{\rm in}_{\rm AM}$ and $u^{\rm in}_{\rm LV}$ across the arachnoid mater ($\Gamma_{\rm AM}$) and lateral ventricular surfaces ($\Gamma_{\rm LV}$) respectively, mimicking the effect of reduced CSF volume due to blood inflow.
-We set $u^{\rm in}_{\rm AM} = 6\,$ml/s and $u^{\rm in}_{\rm LV}=0.31\,$ml/s in the cardiac case, and $u^{\rm{in}}_{\rm{AM}} = 1\,$ml/s and $u^{\rm in}_{\rm{LV}}=0.121\,$ml/s in the respiratory case, respectively.
+We set $u^{\rm in}_{\rm AM} = 6\,$ml/s \cite{causemann2022human,baledent2014imaging} and $u^{\rm in}_{\rm LV}=0.31\,$ml/s \cite{vinje2019respiratory} in the cardiac case, and $u^{\rm{in}}_{\rm{AM}} = 1\,$ml/s \cite{gutierrez2022effect} and $u^{\rm in}_{\rm{LV}}=0.121\,$ml/s \cite{liu2024using} in the respiratory case, respectively.
 We prescribe a no-slip boundary condition on the pial surface $\Gamma_{\rm Pia}$ and zero-traction on the spinal boundary $\Gamma_{\rm SSAS}$, making the spinal compartment the only outflow route of CSF.
@@
     \bm u = \frac{u^{\rm in}_{\rm LV} \cdot \bm n}{|\Gamma_{\rm LV}|} \text{ on } \Gamma_{\rm LV}; \quad
-    \bm u = \frac{u^{\rm in}_{\rm AM} \cdot \bm n}{|\Gamma_{\rm skull}|} \text{ on } \Gamma_{\rm skull}.
+    \bm u = \frac{u^{\rm in}_{\rm AM} \cdot \bm n}{|\Gamma_{\rm skull}|} \text{ on } \Gamma_{\rm AM}.
 \end{equation}
 The obtained pressure fields $p_{\rm cardiac}$ and $p_{\rm resp}$ only account for viscous forces, albeit cardiac and respiratory induced CSF flow is known to be unsteady and substantially impacted by transient inertial forces.
-Assuming an oscillatory flow with either the heartbeat frequency $\omega_{\rm cardiac} = 2 \pi$ or respiratory frequency $\omega_{\rm resp} = 0.5 \pi$, a mean gap width of the CSF-filled spaces of $2h=3$\,mm and CSF density $\rho=10^3\,\text{kg/m}^3$, the Womersley numbers $\alpha$ become $\alpha_{\rm cardiac}^2 = \frac{h^2 \omega_{\rm cardiac} \rho}{\mu} = 20.2$ and $\alpha_{\rm resp}^2=5.05$, which is unsteady.
+Assuming an oscillatory flow with either the heartbeat frequency $\omega_{\rm cardiac} = 2 \pi$ or respiratory frequency $\omega_{\rm resp} = 0.5 \pi$, a mean gap width of the CSF-filled spaces of $2h=3$\,mm and CSF density $\rho=10^3\,\text{kg/m}^3$, the squared Womersley numbers $\alpha_i^2 = \frac{h^2 \omega_{i} \rho}{\mu}$ become $\alpha_{\rm cardiac}^2 = 20.2$ and $\alpha_{\rm resp}^2=5.05$, which is unsteady.
 Based on theoretical considerations on the ratio of oscillatory flow to steady flow impedances in a tube, we account for inertial forces in the pulsatile flow by upscaling the computed viscous pressure field by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter.
 4.3).
-Thus, the upscaled pressure fields become $p^{\rm up}_{\rm cardiac}= (1 + \alpha_{\rm cardiac}^2 / 8)p$ and $p^{\rm up}_{\rm resp}= (1 + \alpha_{\rm resp}^2 / 8)p$.
+Thus, the upscaled pressure fields become $p^{\rm up}_{i}= (1 + \alpha_{i}^2 / 8)p_i \text{ for } i = \{\mathrm{cardiac}, \mathrm{resp}\}$.
 
-Further assuming unsteady dispersion, we follow Sharp et al \cite{keith2019dispersion} and estimate each enhancement factor $R_{\rm cardiac}$ and $R_{\rm respiratory}$ from the non-dimensionalized pressure gradient
+Further assuming unsteady dispersion, we follow Sharp et al \cite{keith2019dispersion} and estimate each enhancement factor $R_{\rm cardiac}$ and $R_{\rm resp}$ from the non-dimensionalized pressure gradient
 $$R_{i} \approx dP_{i}^2 / \alpha_{i}^3, \quad  \text{where} \quad  dP=\frac{|\nabla p^{\rm up}_{i}|}{\omega_{i} \mu / h} \quad \text{for} \quad i = \{\mathrm{cardiac}, \mathrm{resp}\}.$$

### 2024-11-27 paper/background.tex

\section{Background}

\mer{Point from Miro: Goriely and scaling from mice to men.}

\begin{enumerate}
\item
  Communication between perivascular spaces and the subarachnoid space
  \begin{itemize}
  \item
    Rennels et al~\cite{rennels1985evidence} infused tracer into the lateral cerebral ventricles and subarachnoid space of cats (and one dog) and observed that the tracer distributed in a perivascular pattern. The tracer distributed more rapidly around arterioles than around capillaries and veins. Importantly, the rapid paravascular influx was modulated by arterial pulsatility and in particular reduced by reduced vessel pulsatility. They directly postulate that ``fluid circulation through the central nervous system occurs via paravascular pathways'', and discuss both mixing and pumping as pulsatility-associated mechanisms for transport. 
  \item
    Weller and coauthors conducted pioneering studies of perivascular spaces, flow and transport~\cite{zhang1990interrelationships, zhang1992directional, weller2005microscopic}. They identified thin sheaths of pial cells surrounding arteries and arterioles (but not veins or venules) on the brain surface and within the human brain itself~\cite{zhang1990interrelationships}. Further, they observed that tracers spread along perivascular (arterial, venous and capillar) spaces in (rat) grey matter, and in the subarachnoid space (to the cribriform plate and nasal lymphatics).   
  \item
    Ichimura et al~\cite{ichimura1991distribution} study the distribution of large molecular weight tracers in the subarachnoid and surface perivascular spaces and within the cortex in rats. They observed rapid spread in perivascular spaces, and that tracers injected into the CSF of the subarachnoid space spread also into subarachnoid and cortical perivascular spaces. In conclusion, they highlight the presence of an extensive perivascular network, and report of some bulk flow, but slow and of variable directionality - in contrast to e.g Rennels et al~\cite{rennels1985evidence}. 
  \item In a series of papers, Bakker and coauthors study perivascular anatomy and solute transport~\cite{bedussi2017paravascular, bedussi2018paravascular}. They find that the subarachnoid space, the cisterns, ventricles and penetrating periarteriolar spaces form a continuous cerebrospinal fluid-filled space surrounding and penetrating into the murine brain \cite{bedussi2017paravascular}. Moreover, they demonstrate pulsatile and directional (antegrade) flow in perivascular (predominantly periarterial) spaces on the brain surface \cite{bedussi2018paravascular}. 
  \item
    Iliff et al~\cite{iliff2012paravascular} observed rapid movement of tracer (injected in the cisterna magna) along the outside of surface arteries and penetrating arterioles in mice, and state this is in agreement with the presence of perivascular sheaths as reviewed by Weller~\cite{weller2005microscopic}.
  \item
    The architecture of perivascular spaces in and around the rat brain was also studied in detail by Thorne and colleagues~\cite{pizzo2018intrathecal, hannocks2018molecular}. Hannocks et al~\cite{hannocks2018molecular} characterize the molecular composition of different perivascular compartments. Pizzo et al~\cite{pizzo2018intrathecal} emphasizes the presence of stomata or pores present on the interface between blood vessels and the CSF in the subarachnoid space. Neither seem to quantitatively report on PVS widths.
  \item Mestre et al~\cite{mestre2022periarteriolar} study the properties of pial perivascular spaces in detail in mice, and in particular the structure and coverage of pial cells and pial layers surrounding leptomeningeal arteries (and arterioles). They report of pial cells forming sheaths for larger ($> 10000 \mu$m$^2$ cross-section lumen area) surface arteries and partially coverage for smaller surface arteries, with higher coverage in ventral SAS regions. They find that these pial layers do not form an impermeable barrier to small molecules. Importantly, they also study the pial coverage in aged and Alzheimer's model mice, and find significant and complex changes in these pial structures.
  \item
    In humans, Eide and Ringstad investigate cerebrospinal fluid tracer transport in the human subarachnoid space after intrathecal injection~\cite{eide2024functional}. They observe tracer enrichment antegrade along the major cerebral arteries, and enrichment of the nearly cerebral cortex. They also consider variations in different patient cohorts and find impaired transport in iNPH patients with increased perivascular space size. They report of tracer enrichment in donut-shaped perivascular spaces. The timings here may be the best estimate of human PVS flow speeds. 
  \item
    Yamamoto et al~\cite{yamamoto2024perivascular} also report of contrast-enhancement in human PVS.
  \end{itemize}
\item
  Shapes and sizes of the perivascular spaces\footnote{The area $A$ of
  a circle with inner radius $R_1$ is $A = \pi R_1^2$. The area $A_2$
  of an annulus with inner radius $R_1$ and outer radius $R_2$ is $A_2
  = \pi (R_2^2 - R_1^2)$. The width of this annulus is $R_2 -
  R_1$. The ratio PVS/lumen area can thus be estimated as $A_2/A =
  (R_2^2 - R_1^2)/R_1^2$. For example, if $R_2 = n R_1$ ($n = 2
  \rightarrow$ PVS width equal to lumen radius, $n = 3 \rightarrow$
  PVS width equal to lumen diameter), then $A_2/A = n^2 - 1$. However,
  if the annulus has an elliptic outer boundary with radii $R_a$ and
  $R_b$, then its area is $A_e = \pi (R_a R_b - R_1^2)$, and so in the
  extreme case where $R_a > R_b = R_1$, the area is $\pi R_1 (R_a -
  R_1)$, and the PVS/lumen area ratio is $A_e/A = (R_a -
  R_1)/R_1$. And so, for $R_a = n R_1$, then $A_e/A = n-1$.}. Note
  that representing the surface PVS as annular cylinders is a
  simplification, and several studies emphasize the asymmetric nature
  of the PVS shape~\cite{mestre2018flow, tithof2019hydraulic,
    vinje2021brain, raicevic2023sizes}. Moreover, do not forget that
  the sizes of mice and humans are orders of magnitude apart.
  \begin{itemize}
  \item
    Ichimura et al~\cite{ichimura1991distribution} report on a typical
    perivascular width of $1-10\mu$m, but broader around vessel
    bifurcations in the rat subarachnoid space (fixed specimens).
  \item Foley et al~\cite{foley2012realtime} report on perivascular
    space width within rat cortex of $8-10\mu$m for arterioles of
    $\approx$30$\mu$m in diameter.
  \item
    Schain et al~\cite{schain2017cortical} report on the sizes of pial
    arteries ($6.0-11 \mu$m in diameter\footnote{Schain et
    al~\cite{schain2017cortical} state diameter rather, but the
    diameter does not match with the given area, so perhaps this
    should be the radius}) and veins ($9.4-21\mu$m) in mice, with an
    average PVS/lumen area ratio of $1.26$ for arteries and $0.13$ for
    veins. 
  \item Mestre et al~\cite{mestre2018flow} measure and characterize
    pial perivascular transport and estimate flow magnitudes in
    mice. They report a PVS/lumen area ratio of $1.4 \pm 0.1$ and a
    PVS width of $\approx 40 \mu$m which described as comparable to
    the adjacent artery (diameter). 
\item
  Raicevic et al~\cite{raicevic2023sizes} study the sizes and shapes
  of perivascular spaces surrounding pial arteries in mice in detail,
  in particular the relationship between periarterial space and lumen
  (cross-section) area and its variation with vessel area and
  location. They remark that the variation in PVS area is larger
  between PVS segments than along a single PVS segment. The PVS area
  seems to increase with the vessel area, but an affine approximation
  gives a poor fit, and from inspecting the distribution, the
  relationship seems superlinear. The peak distribution value of
  segmented PVS/lumen area is $1.12$.
  \item
    In terms of size, Bedussi et al.~\cite{bedussi2018paravascular} report a PVS width of $\approx$20$\mu$m surrounding surface arteries branching from the MCA (of diameters $45 \pm 7$ $\mu$m). Note that they include both data from mice and OCT data from humans. They report of human PVS size linearly related with the vessel size (and in the range 0.1--0.35 mm for vessels of diameter 0.1--0.5 mm).
  \item Vinje et al~\cite{vinje2021brain} and references therein illustrates the location and characteristics of human pial perivascular spaces, using human OCT data from the Bakker lab.
  \item Bollman et al~\cite{bollmann2022imaging} image the human pial vasculature at high resolution (7T), and report of pial arterial diameters in the range $50-280 \mu$m.
\item
  Smets et al~\cite{smets2024perivascular} study the size of
  perivascular spaces surrounding pial arteries and veins in mice
  (in-vivo), with particular emphasis on the properties of perivenous
  spaces. They observe the PVS as two ``triangular spaces'' adjacent
  to both arteries and veins, and separated from the subarachnoid
  space by a membrane (see also references
  to~\cite{zhang1990interrelationships, pizzo2018intrathecal,
    mollgard2023mesothelium}). They found that the cross-section area
  of the PVS correlates with the lumen area both for pial arteries
  and veins, but not for penetrating vessels. They report a
  PVS-to-lumen-area ratio of 0.43 for arteries and 0.35 for veins.
\item
  Personal communication (Ringstad, Oct 15 2024)~\cite{eide2024functional}: a typical blood vessel in the M2-segment of the MCA is 1.56 mm in diameter ($2 R_1$) with an annular PVS of width ($R_2 - R_1$) 1.23 mm and total PVS diameter $2 R_2$ of 3.93. In iNPH patients, extreme outliers are observed; e.g. still with annular PVS, but not concentric (shifted vessel), for vessel diameter of 2 mm and PVS width from 1.1 mm to 6.0 mm.
  \end{itemize}
\item
  Perivascular hydraulic resistances. The hydraulic resistance of
  perivascular spaces has been studied extensively, with key
  contributions from Thomas, Kelley, Tithof and collaborators~\cite{tithof2019hydraulic}.
  \begin{itemize}
  \item Boster et al.~\cite{boster2024hydraulic} study methods of estimating the hydraulic resistance of perivascular spaces via in-vivo imaging-based 3D reconstructions of pial perivascular spaces in mice; they report of hydraulic resistances of the order $1.7 \times 10^6 - 8.9 \times 10^5$ mmHg·min/mL/m).
  \end{itemize}
\item
  Perivascular flow characteristics: magnitudes and directionality?
  \begin{itemize}
  \item
    Rennels et al~\cite{rennels1985evidence} highlight in their
    Discussion that
    \begin{quote}
      This evidence of fluid displacement in both directions through
      the perivascular spaces suggest either that influx and efflux
      occur intermittently through the same perivascular spaces or
      that these fluid movements occur via separate populations of
      perivascular spaces.
    \end{quote}
  \item
    Ichimura et al~\cite{ichimura1991distribution} found that albumin in the subarachnoid periarterial space moved 1.5 mm away from the injection site in about 7 minutes, thus a net speed of around $3.6 \mu$m/s. 
  \item
    Hadaczek et al~\cite{hadaczek2006perivascular} studied the role of arterial pulsations in more detail, following up on~\cite{rennels1985evidence}.
  \item Foley et al~\cite{foley2012realtime} study perivascular transport characteristics of nanoparticles injected in the cerebral cortex of rats, though in the context of convection-enhanced delivery.
  \item
    In terms of movement, Bedussi et al~\cite{bedussi2018paravascular} found a net movement of particles in surface PVS (in mice) in the antegrade (same direction as blood flow) direction with an average velocity of $17 \pm 2\mu$m/s, and a mean amplitude of the pulsatile motion of $14 \pm 2 \mu$m.
\item Mestre et al~\cite{mestre2018flow} report a mean (averaged over time and space) flow speed of $18.7 \mu$m/s (root-mean-square velocity?) with net transport in the antegrade direction (near the MCA, in mice).
\item
  Asgari et al~\cite{asgari2016glymphatic} model tracer transport in axial periarterial spaces, highlight the negligible net (directional, bulk) flow induced by arterial wall pulsations at physiologically realistic wavelengths, but also the contributions from wall pulsations to dispersion. 
\item
  Rey and Sartinoranont~\cite{rey2018pulsatile} examine pulsatility as a driver for net flow in perivascular spaces within the brain parenchyma via a hydraulic network models, again highlighting the negligible net flow induced by vascular wall movements, but significant pulsatile flow and thus potential for dispersive effects~\cite{watson1983diffusion, asgari2016glymphatic}.
\item
  Rey et al~\cite{rey2023perivascular} consider a comprehensive ex-vivo perivascular network segmentation in the rat brain following in-vivo intraventricular contrast infusion. They highlight the numerous connections between the ventricular system and the perivascular spaces, including long perivascular segments extending from the surface and deep into the brain. 
\item
  In the context of dispersion, see also~\cite{asgari2016glymphatic, keith2019dispersion, troyetsky2021dispersion}
\item
  Wright et al~\cite{wright2024coupled} study the dynamics (waveforms over time) of human arterial blood flow (with resting-state fMRI) and perivascular CSF flow (with dynDWI, apparent diffusion coefficient (mm$^2$/s)). The waveforms seem coupled, with the arterial peak preceding the CSF peak by around 0.06 (of a cardiac cycle). The report ADC in the PVS on the order $2.5-5.5 \times 10^{-3}$ mm$^2$/s, with alterations in the coupling with age.
\item
  Boster and coauthors~\cite{boster2023artificial, toscano2024infeering} estimate perivascular flow parameters in mice via physics-informed neural networks (e.g.~pressure gradients, hydraulic resistances, shear stresses). Boster et al~\cite{boster2023artificial} estimate PVS pressure gradients on the order of $0.7-5 \times 10^{-4} \unit{Pa/\mu m}$\footnote{Note that 
  \begin{equation}
    1 \times 10^{-4} \, \unit{Pa/\mu m} = 1 \times 10^{2} \, \unit{Pa/m} \approx 100/133 \unit{mmHg/m} \approx 0.75 \unit{mmHg/m}
  \end{equation}
   }, downstream velocity component of $12.75 \pm 6.25 \unit{\mu m/s}$. 
  \end{itemize}
\item
  Broggini et al~\cite{broggini2024long} describe long-wavelength traveling waves of vasomotion (0.1 Hz) in pial arterioles in mice. They find waves traveling in both directions, with waves exiting the surface of the cortex more often than chance (85 of 140 exit). ``The wavelength of the observed pial waves is long, typically 20 mm or twice the length of the mouse cortex (inset in Figure 4C).'', with relative amplitudes in the range 0.02--0.12 in pial arterioles.
\item
  Kedarasetti et al~\cite{kedarasetti2020functional} model functional hyperemia waves (aka stimulus-induced vasomotion, see also van Veluw et al~\cite{vanveluw2020vasomotion}). Note that the directionality of vasomotion is classically thought to be in the opposite direction of the blood flow.
\item
  Note that Gokina et al~\cite{gokina1996electrical} report of spontaneous rhythmic contraction in human pial arteries.
\end{enumerate}
Cerebrospinal fluid flow
\begin{itemize}
\item
  Sweetman and Linninger~\cite{sweetman2011cerebrospinal} model the flow of cerebrospinal fluid in the cranial and spinal subarachnoid space of a healthy human driven by pulsatile displacement of the upper walls of the lateral ventricles representing volumetric changes of the vasculature, with model predictions in good agreement with clinical observations of CSF velocity magnitude and stroke volume. 
\item Causemann et al.~\cite{causemann2022human}
\item
  Vinje et al.~\cite{vinje2019respiratory} consider clinical measurements of intracranial pressure differences and human CSF-space geometries together with computational modelling to study CSF flow patterns in the aqueduct. They estimate the pressure gradient across the cerebral aqueduct resulting from a CSF production of 0.5L/day to be $0.009 \pm 0.006$ mmHg/m.
\item Hornkjøl et al~\cite{hornkjol2022csf} model CSF flow and solute transport in the human SAS and brain parenchyma. 
\end{itemize}

### 2024-11-27 paper/main.tex
Fix simple addition ;-)
Former-commit-id: 040a8657f636cd392eb5e965529b88b89ed5f0a3 [formerly 0c67e9ac2b4024d14c6260c7dd13c93919b42c76]
Former-commit-id: 89abed48071cb04d72d2068e47f53a6d0e9e6269
Updates from Overleaf

@@
 
-\mer{MER: Target length: 5000 words (Intro, Results, Discussion) plus <3000 words Methods: 800 words Introduction. 2800 words Results. 1800 words Discussion.}
+\mer{MER: Target length: 5000 words (Intro, Results, Discussion) plus <3000 words Methods: 700 words Introduction. 2800 words Results. 1500 words Discussion.}
 
@@
 \end{itemize}
+
+\paragraph{PVS Flow velocities}
+\begin{itemize}
+    \item we consider three different PVS flow fields: PVS flow driven by production field pressure differences, cardiac peristaltic flow and vasomotion peristaltic flow. 
+\end{itemize}
+
 \begin{figure}
@@
 adjust the segmentation to accurately represent the connections
-between CSF spaces (aqueduct, median aperture), smoothen, and finally
+between CSF spaces (aqueduct, median aperture, tentorium cerebelli), smoothen, and finally
 extract surface representations of the outer (arachnoid) boundary, the
@@
 fTetWild~\cite{hu2020fast}. The resulting computational mesh consists
-of $153\,729$ vertices and $851\,904$ mesh cells, which vary between
-$0.4$ mm and $1.2$ cm in mesh cell diameter. The parenchyma has a
-total volume of $1386$ ml, whereas the ventricles and the outer SAS
-contribute $28$ ml and $277$ ml, respectively, to a total CSF volume
-of $305$ml, in agreement with recent
-estimates~\cite{hladky2024regulation}.
+of $221\,988$ vertices and $1\,225\,774$ mesh cells, which vary between
+$0.24$ mm and $1.29$ cm in mesh cell diameter. The parenchyma has a
+total volume of $1318$ ml, whereas the ventricles and the outer SAS
+contribute $32$ ml and $328$ ml, respectively, to a total CSF volume
+of $360$ml, in agreement with recent estimates~\cite{hladky2024regulation}.
 

### 2024-11-27 paper/supplementary_main.tex
Updates from Overleaf

@@
 
-CSF pulsates in the SAS and ventricular system with each cardiac and respiratory cycle.
-Taking the large spatial differences of pulsatile CSF flow into account, we estimate a local diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac- and respiration-induced pulsatile CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}.
+CSF pulsates in the SAS and ventricular system with each cardiac and respiratory cycle, leading to an increase in effective solute diffusivity via a process known as Taylor dispersion \cite{taylor1953dispersion, watson1983diffusion}.
+Previous estimates of the magnitude of this effect in the CSF-filled space range from an enhancement factor of 5.8 \cite{keith2019dispersion} up to $10^4$ \cite{ray2021quantitative}.
+The large discrepancies in reported estimates can (at least partly) be attributed to different assumptions on representative values for local width, pressure differences and fluid velocities in the SAS, which reflect the diverse flow characteristics in the different regions of the intracranial CSF-filled space.
 
-\rami{I rewrote the above paragraph and some aspects of this section}
-
-\rami{CSF pulsates in the SAS and ventricular system with each cardiac and respiratory cycle with large spatial differences.  This leads to dispersion effects. For example, 
-Sharp et al \cite{keith2019dispersion} estimated the enhancement of diffusion in the spinal subarachnoid space due to dispersive effects to average 5.8 times that of purely molecular diffusion with a theoretical maximum of $10^6$.
-This phenomena is accounted for in our model by scaling the free diffusion coeffiecient $D^{\mathrm{Gad}}$ by the formula $D^{\mathrm{eff}} = (1+R) D^{\mathrm{Gad}}$ where $R$ represents a local diffusion enhancement in the CSF-filled space $\Omega_{\rm CSF}$.}
-
-\rami{ Note that since there is no data for the local values of $R$, we estimate $R$ by combining computational results (solving Stokes flow equations, see below) and explicit estimates for dispersion enhancement that are derived under simplifying assumptions on the problem geometry, the flow field and concentration profiles \cite{watson1983diffusion, keith2019dispersion}.
+Taking the large spatial differences of pulsatile CSF flow into account, we estimate a spatially varying diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac- and respiration-induced pulsatile CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}.
+These estimates are are derived under simplifying assumptions on the problem geometry, the flow field and concentration profiles \cite{watson1983diffusion, keith2019dispersion}.
 Therefore such an estimate is heuristic; however, our computational results qualitatively agree with observed data.
-A key feature in our estimate of $R$ is that it is spatially varying, including local effects depending on a  computed flow field.
-}
 
@@
 The obtained pressure fields $p_{\rm cardiac}$ and $p_{\rm resp}$ only account for viscous forces, albeit cardiac and respiratory induced CSF flow is known to be unsteady and substantially impacted by transient inertial forces.
-Assuming an oscillatory flow with either the heartbeat frequency $\omega_{\rm cardiac} = 2 \pi$ or respiratory frequency $\omega_{\rm resp} = 0.5 \pi$, a mean gap width of the CSF-filled spaces of $2h=3$\,mm and CSF density $\rho=10^3\,\text{kg/m}^3$, the squared Womersley numbers $\alpha_i^2 = \frac{h^2 \omega_{i} \rho}{\mu}$ become $\alpha_{\rm cardiac}^2 = 20.2$ and $\alpha_{\rm resp}^2=5.05$, which is unsteady.
+Assuming an oscillatory flow with either the heartbeat frequency $\omega_{\rm cardiac} = 2 \pi$ or respiratory frequency $\omega_{\rm resp} = 0.5 \pi$, a mean gap width of the CSF-filled spaces of $2h=3$\,mm and CSF density $\rho=10^3\,\text{kg/m}^3$, the squared Womersley number $\alpha_i^2 = \frac{h^2 \omega_{i} \rho}{\mu}$ becomes $\alpha_{\rm cardiac}^2 = 20.2$ and $\alpha_{\rm resp}^2=5.05$, which is unsteady.
 Based on theoretical considerations on the ratio of oscillatory flow to steady flow impedances in a tube, we account for inertial forces in the pulsatile flow by upscaling the computed viscous pressure field by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter.

### 2024-11-27 paper/main.tex

@@
 % respiratory pulsatility
-Employing the same methodology, we estimate the respiratory contribution to dispersive mixing by computing the CSF flow field driven by an inflow of $0.121\,$ml/s \cite{liu2024using} and $1\,$ml/s across the lateral ventricular wall and the arachnoid mater, respectively.
+Employing the same methodology, we estimate the respiratory contribution to dispersive mixing by computing the CSF flow field driven by an inflow of $0.121\,$ml/s \cite{liu2024using} and $1\,$ml/s \cite{gutierrez2022effect} across the lateral ventricular wall and the arachnoid mater, respectively.
 This amounts to a respiratory peak flow volume of $1.121\,$ ml/s at the craniocervical junction \cite{gutierrez2022effect}.

### 2024-11-28 paper/main.tex
Push some rough stuff for the introduction. WIP.

@@
 
-Molecular transport in perivascular spaces (PVSs) is established as a key pathway for human brain clearance and delivery.
-Previous studies indicate that molecules move rapidly in the subarachnoid space (SAS) and in PVSs surrounding pial arteries.
-Here, we aim to model and study this transport in a full pial perivascular and/or vascular network embedded in the SAS.
+\section*{Introduction (rough ideas)}
 
-Motivation: as indicated by~\cite{iliff2012paravascular}, referring to
+% What is the problem? (Big picture)
+% - Big picture/importance sentence.
+% - More in-depth info
+% - "XXX is unknown".
+
+Transport of molecules within the spaces of the central nervous system
+(CSF spaces, parenchyma) is important. Both in the context of
+intracranial drug delivery with potential for personalized treatment
+of cancer, diagnostics and treatment of CSF-related disorders, and in
+the context of understanding clearance in connection with
+neurodegenerative disease and for future treatment and diagnostics of
+dementia disorders. Understanding and predicting molecular enrichment
+is important for treatment and diagnostics of (brain) cancer, for
+diagnostics of CSF-related disorders like normal pressure
+hydrocephalus, subarachnoid hemorrage. Moreover, altered transport and
+clearance is believed to be key factor in neurodegenerative diseases
+like Alzheimer's disease and Parkinson's disease, and in sleep. CSF
+flow and PVS established to play a key role. Clinical studies have
+revealed large individual variation and variation between
+cohorts/disease groups, therefore important in particular in the
+context of personalized medicine. However, we are far from
+understanding (the mechanisms behind) these variations.
+
+% What we already know about the problem
+Our current understanding has primarily been established via
+experimental medical research (cats, rats, mice, pigs). Big open
+questions and controversy remain, lack holistic understanding of these
+physiological process remains, hard to translate experimental studies
+to humans. In the last decade, clinical studies. Clinical studies have
+revealed large individual variation and variation between
+cohorts/disease groups. However, we are far from understanding the
+mechanisms behind these variations and integrating and translating
+experimental observations to humans.  In-silico modelling is starting
+to play an integral role, evaluating hypothesis and pinpointing
+mechanisms, but most efforts have been focused on local aspects and
+lacking a holistic model. Hard to quantify the overall contribution of
+individual effects.
+
+PVS flow is important as shown by experimental studies. This is being
+corroborated by clinical studies in humans (Eide and Ringstad (2024),
+and the new one). However, a number of open questions remain: to what
+extent do the PVS constitute an ensheathed space, how fast is the flow
+in human PVS, what is the effect of dilated PVS as seen in disease
+cases. Also, open questions on transport characteristics of the CSF
+spaces - what gives the individual variations in tracer enhancements?
+
+What is known is that fluid motion (in the ventricular system and SAS,
+in perivascular spaces, and possibly within the parenchyma) and
+pulsatility play key
+roles. Glymphatics~\cite{iliff2012paravascular}. Diffusion, PVS flow
+and dispersion, CSF flow. Molecular transport in perivascular spaces
+(PVSs) is established as a key pathway for human brain clearance and
+delivery. Previous studies indicate that molecules move rapidly in the
+subarachnoid space (SAS) and in PVSs surrounding pial arteries. CSF
+production in the choroid plexus and possibly in the parenchyma. CSF
+efflux pathways, meningeal lympahtics, nerve routes. See also to
 Cserr, Phys Rev., 1971, bulk convective flow may be more important for
-larger species (larger distances) and larger molecules.
+larger species (larger distances) and larger molecules. Contribution
+both from net flow (convection) and pulsatility (mixing) both in the
+CSF spaces and in PVSs. Some aspects are better established, such as;
+while other factors are more disputed such as ... contribution of
+vasomotion. Moreover, the role of local effects in the bigger picture
+has remained hard to evaluate.
 
+% What are our research questions and how do we answer them
+Recent developments on 3D-1D modelling of transport in PVS and
+tissue. ``Digital twin for molecular enrichment and clearance in human
+intracranial spaces''. Can we model the interplay between
+molecular transport in the intracranial compartments, CSF flow and PVS
+flow and transport to holistically quantify the contributions in humans?
+
+% Main findings and outlook
+Here, we aim to model and study this transport in a full pial
+perivascular and/or vascular network embedded in the SAS.
+
+\iffalse
+\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\fi
+
 \bibliography{references}
 
+\iffalse
 \newpage
@@
 
+\fi
+
 \end{document}

### 2024-11-28 paper/main.tex
Uncomment iffalse!

@@
 
-\section*{Introduction (rough ideas)}
-
 % What is the problem? (Big picture)
@@
 % - "XXX is unknown".
+
+% What we already know about the problem
+
+% What are our research questions and how do we answer them
+
+% Main findings and outlook
+
+\newpage
+\section*{Introduction (rough ideas)}
+
+\begin{itemize}
+\item
+  Need better understanding of transport of solutes in intracranial spaces.
+\item
+  Need to integrate separate observations e.g. on diffusion in
+  neuropil, PVS flow, net flow, effect of dispersion
+\item
+ Need to transfer and integrate insights from animal models,
+  computational models and theoretical insights into the human case. 
+\item
+ Controversies need to be resolved.
+\end{itemize}
 
@@
 and the new one). However, a number of open questions remain: to what
-extent do the PVS constitute an ensheathed space, how fast is the flow
-in human PVS, what is the effect of dilated PVS as seen in disease
-cases. Also, open questions on transport characteristics of the CSF
-spaces - what gives the individual variations in tracer enhancements?
+extent are PVSs compartmentalized, how fast is the flow in human PVS,
+what is the effect of dilated PVS as seen in disease cases etc.  Also,
+open questions on transport characteristics of the CSF spaces - what
+gives the individual variations in tracer enhancements?
 
@@
 
-\iffalse
 \newpage
@@
 
-\fi
-
 \bibliography{references}
 
-\iffalse
 \newpage
@@
 
-\fi
-
 \end{document}

### 2024-11-28 scripts/figure1.py
update plots

@@ -4,16 +4,25 @@ from pathlib import Path
 
 units = Units.cm
 xstart = 0.1
-textargs = {"x":fs/2, "y":0.4, "size":8, "anchor":"middle",
+textargs = {"x":2, "y":0.4, "size":8, "anchor":"middle",
              "weight":"normal", "font":"DejaVu Sans"}
 textargs_histo = dict(textargs, y=0)
 y0 = 0.4
+y1 = 7.4
+w = 2.65
+
 
 class Figure2_Panels(PanelsSpec):
     a: Panel = Panel(
+            RasterImage("paper/figures/Brain-PVS-callouts.png",
+                         ElemSize(9, 7, units),),
+            Location(xstart, y0),
+            content_offset=Location(0,-0.5),
+        )
+    b: Panel = Panel(
             RasterImage("plots/modelA/modelA_overview.png",
                          ElemSize(13, 8, units), crop=(0,100,0,0)),
-            Location(xstart, y0),
+            Location(xstart, y1),
             content_offset=Location(0,0),
             text=(Text(" 1:00 h", **dict(textargs, y=-0.1, x=w/2)),
                   Text(" 6:00 h", **dict(textargs, y=-0.1, x=3*w/2 )),
@@ -23,7 +32,7 @@ class Figure2_Panels(PanelsSpec):
 
 
 class Figure2(FigureSpec):
-    figure_size = ElemSize(18.2, 24, units)
+    figure_size = ElemSize(18.2, 16, units)
     output_file = Path.cwd() / "plots" / "figures" / "figure1.png"
     auto_label_options = AutoLabelOptions(
         first_char:=Text("A", 0, 0, size=12, weight="bold"))

### 2024-11-28 scripts/figure3.py

@@ -26,12 +26,17 @@ class Figure2_Panels(PanelsSpec):
             content_offset=Location(-1,-1.3),
             text=Text("Arterial network", **dict(textargs, x=4.5, y=1.2))
         )
+    a2: Panel = Panel(
+            RasterImage("paper/figures/peristaltic_flow.png", ElemSize(5, 2, units)),
+            Location(xstart + 8, y0),
+            text=Text("Peristaltic flow", **dict(textargs, x=2, y=0))
+        )
     b: Panel = Panel(
         RasterImage("plots/pvs_flow_prod/sas_flow-arteries/sas_flow-arteries_radius.png",
                      ElemSize(6, 6, units)),
-        Location(xstart + 8, y0),
-        #content_offset=Location(-1,-1.3),
-        text=Text("radii of selected arteries", **dict(textargs, y=0.3))
+        Location(xstart + 8, 2.5),
+        content_offset=Location(-1,-0.8),
+        text=Text("radii of selected arteries", **dict(textargs, y=1.2, x=3))
     )
     c: Panel = Panel(
         RasterImage("plots/pvs_flow_prod/sas_flow-arteries/sas_flow-arteries_velocity_histo_cell.png",

### 2024-11-28 scripts/plot_mean_concentrations.py

@@ -1,0 +1,80 @@
+from fenics import *
+from xii import *
+from plotting_utils import get_result_fenics
+import typer
+from typing import List
+import matplotlib.pyplot as plt
+from plotting_utils import set_plotting_defaults, read_config, minmax
+from solver import pcws_constant
+import numpy as np
+import seaborn as sns
+from IPython import embed
+import pandas as pd
+import yaml
+from branch_marking import color_branches
+from label_arteries import pointlabels
+from test_map_on_global_coords_shift import map_kdtree
+from peristalticflow import mesh_to_weighted_graph, nx
+from generate_synthseg_mesh import CSFID, PARID, LVID, V34ID, CSFNOFLOWID
+import joypy
+from cmap import Colormap
+from scipy.stats import binned_statistic
+
+mol2mmol = 1e3
+
+def compare_concentrations(modelname:str):
+    config = read_config(f"configfiles/{modelname}.yml")
+    dt , T= config["dt"], config["T"]
+    times = np.arange(0, T + dt, 3*dt*config["output_frequency"])
+
+    data = read_config(f"results/{modelname}/mean_concentrations.yml")
+
+    labels = ["CSF", "parenchyma", "PVS artery", "PVS vein"]
+    tot_values = [[data[f"ctot_{l}_{t}"] for t in times] for l in labels]
+    concentrations = [[data[f"cmean_{l}_{t}"] for t in times] for l in labels]
+    labels = ["CSF", "PAR", "PVS artery", "PVS vein"]
+
+    set_plotting_defaults()
+    #sns.set_palette("BrBG", n_colors=4)
+    sns.set_palette(["#0a9396","#94d2bd", "#e9d8a6", "#ee9b00"])
+
+    fig, ax = plt.subplots(figsize=(5,4))
+    tot = np.zeros_like(times)
+
+    for q, l in zip(tot_values, labels):
+        new_tot = tot + np.array(q)*mol2mmol
+        fill = ax.fill_between(times/ (60*60), new_tot, tot, 
+                               alpha=0.7, label=l)
+        tot = new_tot
+    
+    plt.xlabel("time (h)")
+    #ax2.set_ylabel('mean tracer concentration (mmol/l)')
+    ax.set_ylabel("total tracer content (mmol)")
+    ax.legend(loc='upper center', bbox_to_anchor=(0.5, 1.2), ncol=4, 
[…]

### 2024-11-28 scripts/make_video.py
change cmap

@@ -7,10 +7,12 @@ from tqdm import tqdm
 from plot_csf_flow import from_k3d
 import k3d.colormaps.paraview_color_maps as pcm
 from extract_vessels import get_tubes
+import seaborn as sns
 
 def make_movie(modelname:str):
     config = read_config(f"configfiles/{modelname}.yml")
     dt , T= config["dt"], config["T"]
+    #cmap = sns.blend_palette(["lightsteelblue", "orange", "darkviolet"],as_cmap=True)
     cmap = from_k3d(pcm.Yellow___Gray___Blue).reversed()
     clim = (0,2)
     times = np.arange(0, T + dt, dt*config["output_frequency"])
@@ -34,7 +36,7 @@ def make_movie(modelname:str):
 
     pl = pv.Plotter(off_screen=True, window_size=(1200, 1200))
     pl.set_background("black")
-    pl.open_movie(f"plots/{modelname}/{modelname}.mp4", framerate=24,
+    pl.open_movie(f"plots/{modelname}/{modelname}_violet.mp4", framerate=24,
                   quality=8)
     pl.add_mesh(skull,style='points',point_size=0.1, opacity=0.6, show_scalar_bar=False,
                 color="silver", specular=1, render_points_as_spheres=True)

### 2024-11-28 paper/main.tex
Updates from Overleaf

@@
 \begin{figure}
-     \begin{subfigure}[b]{0.45\textwidth}
     \centering
-         \includegraphics[width=\textwidth]{figures/Brain-PVS-callouts.png}
-     \end{subfigure}
-     \hfill
+    \includegraphics[width=\textwidth]{paper/figures/figure1.png}
      \caption{a) Illustration of the different intracranial domains
@@
 M) respiratory dispersion enhancement factor;
-N) sagittal, coronal and axial planes of tracer concentration after 1, 6, 12 and 24$\,$h for model A. The outer model boundaries is shown in cyan, the pial membrane in purple and the arteries in red.;
-O) model X,Y and Z after 12h;
+N) sagittal, coronal and axial planes of tracer concentration after 12h in different model variations.
+The outer model boundaries is shown in cyan, the pial membrane in purple and the arteries in red.;
+O) total content of tracer in each compartment;
+P) mean tracer concentration in each compartment;
 }

### 2024-12-01 scripts/label_arteries.py
refine annotations

@@ -1,6 +1,11 @@
 import pyvista as pv
 import numpy as np
 import seaborn as sns
+from cmap import Colormap
+import pandas as pd
+from PIL import Image, ImageFont, ImageDraw  
+from test_map_on_global_coords_shift import map_kdtree
+
 pointlabels = [("BA",(0.084971, 0.104599, 0.0424848)),
                ("ICA-R",(0.0965484, 0.112456,0.0413785)),
                ("ICA-L",(0.0695824, 0.111343, 0.0415065)),
@@ -25,41 +30,95 @@ def get_tangent(v, n):
 
 if __name__=="__main__":
    filename = "plots/meshplots/labeled_arteries.png"
-   art = pv.read("mesh/networks/arteries_tubes.vtk")
-   labelcoords = lambda label: art.bounds[::2]
+   art = pv.read("mesh/networks/arteries_tubes.vtk").triangulate()
 
    art["radius"] *= 1e3
-   col = "fuchsia"      
+   col = "cyan"   
+   cmap = Colormap("curl_pink_r")   
    arrow_length = 0.015
+   window_size = (1200, 850)
    camera_direction = (-0.256, -0.958, -0.122)
-   points = [p for l, p in pointlabels]
-   labels = [l for l, p in pointlabels]
+   #pointlabels.sort(key=lambda pl: pl[1][2], reverse=True)
+   #from IPython import embed; embed()
+   idx = map_kdtree(np.array(art.points), 
+                    np.array([p for l,p in pointlabels]).astype(np.float32),
+                     distance_upper_bound=3e-3, k=100)
+
+   ex = art.extract_points(np.unique(idx))
+   #from IPython import embed; embed()
    bar_args=dict(title="radius (mm)", vertical=False,
                      height=0.07, width=0.6, position_x=0.2,
-                     position_y=0.0, title_font_size=24,
-                     label_font_size=24, fmt="%.1f")
-   pl = pv.Plotter(off_screen=True)
-   pl.add_mesh(art, cmap="coolwarm", scalar_bar_args=bar_args, show_scalar_bar=True)
-   pl.add_points(np.array(points), render_points_as_spheres=False, 
-                 point_size=20, color=col)
-   center = art.center_of_mass() - np.array([0,0, 0.047])
-   label_coords = []
-   for i, (l, p) in enumerate(pointlabels):
-      dir = np.array(p) - center
[…]

### 2024-12-01 scripts/test_map_on_global_coords_shift.py

@@ -4,9 +4,9 @@ import numpy as np
 import numpy_indexed as npi
 from pykdtree.kdtree import KDTree
 
-def map_kdtree(data_points, query_points):
+def map_kdtree(data_points, query_points, **kwargs):
     tree = KDTree(data_points)
-    dist, idx = tree.query(query_points, k=1)
+    dist, idx = tree.query(query_points, **kwargs)
     return idx
 
 def get_global_space(uh, mesh):

### 2024-12-01 scripts/compare_PVS_timing.py
change colors

@@ -40,12 +40,13 @@ def plot_timings(models:List[str]):
     namedict = {"modelA": "base", "modelA-strongVM":"base + VM",
                 "modelA-PVS-disp": "base + 10x disp",}
     print(models)
+    colors = [ "#3a0ca3","#f72585", "#4cc9f0"]
 
     for t, (pvsdf, outerdf) in zip(["fta", "peaktime"] , [(ftas, outerftas), (pts, otps)]):
-        fig,axes = plt.subplots(ncols=len(art_groups), figsize=(12,5.5))
+        fig, axes = plt.subplots(ncols=len(art_groups), figsize=(12,5.5))
 
         for i, (ax, (agn, ag))  in enumerate(zip(axes, art_groups.items())):
-            for m,c in zip(models,["olive", "purple", "blue"]):
+            for m,c in zip(models, colors):
                 group = pvsdf.loc[ag]
                 outergroup = outerdf.loc[ag]
                 ax.plot(group["dist"], group[m] / 3600, color=c, marker="o", markersize=8, ls="dashed",

### 2024-12-01 scripts/figure1.py
small update

@@ -20,6 +20,18 @@ class Figure2_Panels(PanelsSpec):
             content_offset=Location(0,-0.5),
         )
     b: Panel = Panel(
+            RasterImage("paper/figures/geometry_front.png",
+                         ElemSize(4.5, 4.5, units),),
+            Location(xstart + 9, y0),
+            content_offset=Location(0,-0.5),
+        )
+    c: Panel = Panel(
+            RasterImage("paper/figures/geometry_side.png",
+                         ElemSize(4.5, 4.5, units),),
+            Location(xstart + 13.5, y0),# auto_label=False,
+            content_offset=Location(0,-0.5),
+        )
+    d: Panel = Panel(
             RasterImage("plots/modelA/modelA_overview.png",
                          ElemSize(13, 8, units), crop=(0,100,0,0)),
             Location(xstart, y1),

### 2024-12-01 scripts/figure3.py

@@ -21,21 +21,21 @@ y7 = 44
 
 class Figure2_Panels(PanelsSpec):
     a: Panel = Panel(
-            RasterImage("plots/meshplots/labeled_arteries.png", ElemSize(9, 8, units)),
+            RasterImage("plots/meshplots/labeled_arteries.png", ElemSize(10.8, 7, units)),
             Location(xstart, y0),
-            content_offset=Location(-1,-1.3),
-            text=Text("Arterial network", **dict(textargs, x=4.5, y=1.2))
+            content_offset=Location(0,-0.5),
+            text=Text("Arterial network", **dict(textargs, x=4.5, y=0))
         )
     a2: Panel = Panel(
             RasterImage("paper/figures/peristaltic_flow.png", ElemSize(5, 2, units)),
-            Location(xstart + 8, y0),
+            Location(xstart + 11, y0),
             text=Text("Peristaltic flow", **dict(textargs, x=2, y=0))
         )
     b: Panel = Panel(
         RasterImage("plots/pvs_flow_prod/sas_flow-arteries/sas_flow-arteries_radius.png",
                      ElemSize(6, 6, units)),
-        Location(xstart + 8, 2.5),
-        content_offset=Location(-1,-0.8),
+        Location(xstart + 11, 2.5),
+        content_offset=Location(-0.5,-0.8),
         text=Text("radii of selected arteries", **dict(textargs, y=1.2, x=3))
     )
     c: Panel = Panel(
@@ -88,9 +88,16 @@ class Figure2_Panels(PanelsSpec):
         text=Text("model A total PVS tracer content", **dict(textargs, y=-0.3))
     )
     j: Panel = Panel(
-        RasterImage("plots/modelA-strongVM/modelA-strongVM_ridgeline_total_smoothed.png",
+        RasterImage("plots/modelA-PVS-disp/modelA-PVS-disp_ridgeline_total_smoothed.png",
                      ElemSize(6, 6, units)),
         Location(xstart + 6, y3),
+        #content_offset=Location(-1,-1.3),
+        text=Text("model A + PVS disp (10x) - total PVS tracer content", **dict(textargs, y=-0.3))
+    )
+    j2: Panel = Panel(
+        RasterImage("plots/modelA-strongVM/modelA-strongVM_ridgeline_total_smoothed.png",
+                     ElemSize(6, 6, units)),
+        Location(xstart + 12, y3),
[…]

### 2024-12-06 scripts/detail_plot.py
update plots

@@ -12,8 +12,19 @@ from test_map_on_global_coords_shift import map_kdtree
 import pyvista as pv
 from generate_synthseg_mesh import PARID
 from mpl_toolkits.axes_grid1.inset_locator import inset_axes
+import argparse
 
 percentile = 95
+
+def shrink_cbar(ax, shrink=0.9):
+    b = ax.get_position()
+    new_w = b.width*shrink
+    pad = (b.width-new_w)/2.
+    new_x0 = b.x0 + pad
+    new_x1 = b.x1 - pad
+    b.x0 = new_x0
+    b.x1 = new_x1
+    ax.set_position(b)
 
 def get_tangent(v, n):
     return v - np.outer(np.dot(v,n), n)
@@ -26,16 +37,50 @@ def get_normal(netw, pidx):
 def get_radius(netw, pidx):
     return netw.extract_points(pidx)["radius"].max()
 
-def compare_models(modelname:str):
-    times = np.array([1, 3, 6, 12, 24])*3600
-    n_arteries = len(pointlabels)
-    points = np.array([c for n,c in pointlabels[:n_arteries]])
-    artlabels = [n for n,c in pointlabels[:n_arteries]]
-    fig = plt.figure(figsize=(len(artlabels)*4, len(times)*3), frameon=True)
+def get_slice(art, sas, par, p, pvs_radius_ratio=2):
+        pidx = map_kdtree(art.points, np.array([p]))[0]
+        r = get_radius(art, pidx)
+        n = get_normal(art, pidx)
+        fp1, fp2, fp3 = p + 1e-4*n, p + 2e-4*n, p + 3e-4*n
+        vessel = pv.SolidSphere(center=fp3, direction=n, outer_radius=r).slice(origin=fp3, normal=n)
+        vessel.add_field_data(name="color", array="red")
+        pvs = pv.SolidSphere(center=fp2, direction=n, outer_radius=r*pvs_radius_ratio).slice(origin=fp2, normal=n)
+        pvs_surf = pv.Sphere(center=fp2, radius=r*pvs_radius_ratio).slice(origin=fp2, normal=n)
+        pvs_surf.add_field_data(name="color", array="yellow")
+        pvs_surf.add_field_data(name="line_width", array=5)
+        par["n"] = get_tangent(par.point_data["Normals"], n)
+        par["n"] /= np.linalg.norm(par["n"], axis=1, keepdims=True)
+        pia = par.warp_by_vector("n", factor=-2e-4).slice(origin=fp1, normal=n)
[…]

### 2024-12-06 scripts/evaluate_pvs_flow.py

@@ -18,6 +18,7 @@ import yaml
 from branch_marking import color_branches
 from plotting_utils import minmax
 import ufl_legacy as ufl
+from cmap import Colormap 
 m2mm = 1e3
 m2mum = 1e6
 Pa2mPa = 1e3
@@ -27,7 +28,8 @@ def sig_to_space(sig, mesh):
     return FunctionSpace(mesh, elem)
 
 def compute_pvs_flow(pvs_flow_file):
-
+    artcolor = "#94221F"
+    color = "#95AAD3"
     mesh = Mesh()
 
     with HDF5File(MPI.comm_world, pvs_flow_file,'r') as f:
@@ -63,16 +65,14 @@ def compute_pvs_flow(pvs_flow_file):
         umed = np.median(abs(uvals))
         range = np.round(minmax([uvals], percentile=99.9), 1)
         nbins = 50
-        fig, ax = plt.subplots(figsize=(4,3))
+        fig, ax = plt.subplots(figsize=(3.5,3.5))
         counts, bins, containers = plt.hist(uvals, density=False, bins=nbins, histtype="bar",
                 range=range, edgecolor='black', linewidth=0.4, 
                         weights=lengths / lengths.sum())
         pos_share = max([0, range[1] / (range[1] - range[0])])
         print(pos_share)
-        for bar, ca in zip(containers, sns.color_palette("Purples_r",
-                                                         n_colors=int((1 - pos_share) * nbins)) + 
-                                       sns.color_palette("Greens",
-                                                       n_colors=int(pos_share * nbins))):
+        for bar, ca in zip(containers, list(Colormap("balance_blue").iter_colors(int((1 - pos_share) * nbins))) + 
+                                       list(Colormap("curl_pink_r").iter_colors(int(pos_share * nbins)))):
             bar.set_facecolor(ca)
         plt.axvline(uavg, color="black", linestyle=":", label=f"avg: {uavg:.2f} µm/s")
         #plt.axvline(umed, color="black", linestyle="-.", label=f"median: {umed:.2f} µm/s")
@@ -83,7 +83,8 @@ def compute_pvs_flow(pvs_flow_file):
         plt.xlim(range)
         plt.legend(frameon=False)
[…]

### 2024-12-06 scripts/figure2.py

@@ -129,11 +129,11 @@ class Figure2_Panels(PanelsSpec):
             text=Text("low dispersion (12 h)", **dict(textargs, y=-0.1, x=1.5))
         )
     n4: Panel = Panel(
-            RasterImage("plots/modelA-OnlyDispersion/modelA-OnlyDispersion_overview.png",
+            RasterImage("plots/modelA-HighD/modelA-HighD_overview.png",
                          ElemSize(3, hn, units), crop=(1420, 100, 1350, 0)),
             Location(xstart + 3*compwidth, y3), auto_label=False,
             content_offset=Location(0,0),
-            text=Text("placeholder", **dict(textargs, y=-0.1, x=1.5))
+            text=Text("high dispersion (12 h)", **dict(textargs, y=-0.1, x=1.5))
         )
     n5: Panel = Panel(
             RasterImage("plots/modelA-OnlyDispersion/modelA-OnlyDispersion_overview.png",

### 2024-12-06 paper/cut.tex

@@
+ \section{Introduction}
+It remains an open question however how these physical and
+physiological mechanisms, though well-characterized when viewed in
+isolation, contribute to enhance human intracranial molecular
+transport.
+
+PVS flow is important as shown by experimental studies. This is being
+corroborated by clinical studies in humans (Eide and Ringstad (2024),
+and the new one). However, a number of open questions remain: to what
+extent are PVSs compartmentalized, how fast is the flow in human PVS,
+what is the effect of dilated PVS as seen in disease cases etc.  Also,
+open questions on transport characteristics of the CSF spaces - what
+gives the individual variations in tracer enhancements?
+
+What is known is that fluid motion (in the ventricular system and SAS,
+in perivascular spaces, and possibly within the parenchyma) and
+pulsatility play key
+roles. Glymphatics~\cite{iliff2012paravascular}. Diffusion, PVS flow
+and dispersion, CSF flow. Molecular transport in perivascular spaces
+(PVSs) is established as a key pathway for human brain clearance and
+delivery. Previous studies indicate that molecules move rapidly in the
+subarachnoid space (SAS) and in PVSs surrounding pial arteries. CSF
+production in the choroid plexus and possibly in the parenchyma. CSF
+efflux pathways, meningeal lympahtics, nerve routes. See also to
+Cserr, Phys Rev., 1971, bulk convective flow may be more important for
+larger species (larger distances) and larger molecules. Contribution
+both from net flow (convection) and pulsatility (mixing) both in the
+CSF spaces and in PVSs. Some aspects are better established, such as;
+while other factors are more disputed such as ... contribution of
+vasomotion. Moreover, the role of local effects in the bigger picture
+has remained hard to evaluate.
+
+\section{Results}
+
 \mer{MER: Old, revisit to see if we have forgotten any of these interesting ideas.}

### 2024-12-06 paper/supplementary_main.tex

@@
   \sum_{\Lambda^i \in \mathcal{E}(y)}
-  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) ni(y)
+  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n^i(y)
   = 0
@@
 respectively. We define the union of facets interior both to
-$\mathcal{T}_{i, \rm PAR}$ and $\mathcal{T}_{i, \rm CSF}$ as
+$\mathcal{T}_{ \rm PAR}$ and $\mathcal{T}_{ \rm CSF}$ as
 $\mathcal{F}_i = \mathcal{F}_{i, \rm PAR} \cup \mathcal{F}_{i, \rm
@@
   %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
-  V_h^k = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^k(E), \; E \in \mathcal{T}\} ,
+  V_h = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^1(E), \; E \in \mathcal{T}\} ,
 \end{equation}
 where $L^2(\Omega)$ is the space of square-integrable functions
-defined over $\Omega$ and $\mathrm{P}^k(E)$ denotes the space of
-polynomials of degree $\leqslant k$ defined over the tetrahedra
+defined over $\Omega$ and $\mathrm{P}^1(E)$ denotes the space of
+polynomials of total degree $\leqslant 1$ defined over the tetrahedra
 $E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
@@
   4.5.2.3]{di2011mathematical} for details on this
-method. \discuss{Recall that $\phi$ is constant in each domain
+method. Recall that $\phi$ is constant in each domain
   $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
-  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.}
+  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.
 Define for $c, v \in V_h^k$:
@@
   =
-  \sum_{E \in \mathcal{T}} \int_{E} D \discuss{\phi} \nabla c \cdot \nabla v
+  \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
   + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
@@
 between cells $E^1$ and $E^2$ with facet normal vector $\bm n_F$
-pointing from $E^1$ to $E^2$, \discuss{$h_F$ denotes the facet diameter},
+pointing from $E^1$ to $E^2$, $h_F$ denotes the facet diameter,
 the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
@@
   \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
-  \quad \text{where} \quad \kappa_i = \phi D|_{E^i}.  
+  \quad \text{where} \quad \kappa_i = (D \phi) |_{E^i}.  
 \end{equation}
 The parameter $\gamma_{D,F}$ is the harmonic mean of the
-\discuss{porosity-weighted} diffusion coefficient given by
+porosity-weighted diffusion coefficient given by
 \begin{equation}
@@
 \end{equation}
-Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h^k$ such that for all $v \in V_h^k$:
-\begin{equation}
+Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h$ such that for all $v \in V_h$:
+\begin{equation} \label{eq:3d_dg_disc}
   \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
   + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
-  + \int_{\Gamma_{\rm pia} \discuss{\cup \Gamma_{\rm LV}} } \beta_{\rm pia}[c_h^n][v] 
+  + \int_{\Gamma_{\rm pia} \cup \Gamma_{\rm LV}} \beta_{\rm pia}[c_h^n][v] 
   + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
@@
   %\int_{\Omega} f^n v +
-  \int_{\Gamma_{\rm \discuss{S}SAS}} g_{\rm influx} v
+  \int_{\Gamma_{\rm SSAS}} g_{\rm influx} v
   %, \quad \forall v \in V_h^k
@@
 where $\mathrm{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$.
+The discrete formulation then reads: for $n = 1,2,\ldots, $ find $\hat c_h^n \in \hat V_h$ such that for all $\hat v \in \hat V_h^k$:
+\begin{align}\label{eq:1d_dg_disc}
+\int_{\Lambda}  \frac{1}{\tau}(A  \hat c_h^n - A \hat c_h^{n-1}) \hat v  + \int_{\Lambda} \hat D A \partial_s \hat c_h^n  \partial_s \hat v^n - \int_{\Lambda} A \hat u \hat c_h^n \partial_s \hat v   +  \int_{\Lambda} \xi P (\hat c_h^n - \overline{c_h^n}) +a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) = 0.
+\end{align}
+ In the above, if $\hat u$ is non--zero on $\Lambda^i$, then the artificial diffusion stabilization term, for more details see for e.g. \cite[Section 12.6]{quarteroni2009numerical}: 
+ \begin{align}
+    a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) =  \frac{1}{\|\hat u\|_{L^2(\Lambda^i)}} 
+ h_{\Lambda^i} \int_{\Lambda^i} A (\hat u \partial_s \hat c_h  ) (\hat u \partial_s \hat v),  
+ \end{align}
+ where $h_{\Lambda^i}$ is the mesh--size of $\Lambda^i$.
 
+ \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} comlemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0$ is the following:
+
+ For $n = 1,2, \ldots $, find  $c_h^n \in V_h$ and $\hat c^n_h \in \hat V_h$ such that the coupled equations \eqref{eq:3d_dg_disc} and \eqref{eq:1d_dg_disc} hold for all $v \in V_h$ and for all $\hat v \in \hat V_h$.
 \subsubsection{Finite element solution of the incompressible Stokes equations}
@@
   + %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
-  \frac{\sigma \mu}{h} \int_{F} [\bm u_t ] \cdot [\bm v_t] \right ),   
+\int_{F}   \frac{\sigma \mu}{h_F} [\bm u_t ] \cdot [\bm v_t] \right ),   
 \end{multline}
-where \discuss{$h$ denotes the element size} and on exterior facets \mer{the average and jump operators take the one-sided values}.
+where on exterior facets \mer{the average and jump operators take the one-sided values}.
 We set the penalty parameter for the tangential continuity to be $\sigma = 20$.

### 2024-12-07 paper/supplementary_main.tex

@@
 
+\mk{What is justification of \eqref{eq:1d_bcs}?}
 %% Both $\Omega$ and $\Lambda$ are fixed in time and we assume
@@
 \end{subequations}%
-where $A$ is the area of the PVS cross-sections, $\mu$ is the CSF viscosity as before, and $\kappa$ is derived from an assumption of Poiseuille flow in the annular cross-section of the PVS as~\cite{tithof2022network}:
+where $A$ is the area of the PVS cross-sections, $\mu$ is the CSF viscosity as before, and $\kappa$ is derived from an assumption of Poiseuille flow in the annular cross-section of the PVS as~\cite{tithof2022network}:  \rami{Here,  we consider the steady state  case and we assume that the PVS domain is fixed and not pulsating  }
 \begin{equation}

### 2024-12-13 scripts/detail_plot.py
add and update plots

@@ -65,10 +65,11 @@ def generate_plot(
     times:List[int],
     cmax: List[float],
 ):
-
+    if len(cmax)==1 and cmax[0]==0:
+        cmax = 0
     times = np.array(times)*3600
     pointdict = {l:p for l,p in pointlabels}
-    if selected_artlabels is None:
+    if selected_artlabels[0]=="all":
         artlabels = list(pointdict.keys())
     else:
         artlabels = selected_artlabels
@@ -130,14 +131,10 @@ def generate_plot(
         ax = grid.axes_row[j][0]
         ax.text( -0.13, 0.5, label, va="center",
          transform=ax.transAxes, fontsize=fontsize, rotation=90)
-    tstr = "-".join([f"{t/3600:.0f}" for t in times])
-    if selected_artlabels:
-        artstr = "-".join(selected_artlabels) + "-"
-    else: 
-        artstr = ""
-    if cmax:
-        cmstr = "-".join([str(cm) for cm in cmax])+"_"
-    else: cmstr=""
+    tstr = "+".join([f"{t/3600:.0f}" for t in times])
+    artstr = "+".join(selected_artlabels) + "_"
+    artstr = ""
+    cmstr = "+".join([str(cm) for cm in cmax])+"_"
 
     plt.savefig(f"plots/{modelname}/{modelname}_{tstr}_{artstr}{cmstr}details.png",
                  dpi=200, transparent=True,bbox_inches="tight")
@@ -150,8 +147,8 @@ def main():
     parser.add_argument("modelname", type=str, help="The name of the model")
     # Optional arguments
     parser.add_argument("--times", type=int, nargs='+', default=[1,3,6,12,24], help="List of times as integers")
-    parser.add_argument("--cmax", type=float, nargs='+', default=None, help="List of cmax values as floats")
-    parser.add_argument("--artlabels", type=str, nargs='+', default=None, help="List of art labels as strings")
+    parser.add_argument("--cmax", type=float, nargs='+', default=0, help="List of cmax values as floats")
+    parser.add_argument("--artlabels", type=str, nargs='+', default="all", help="List of art labels as strings")
     
     args = parser.parse_args()
     

### 2024-12-13 scripts/generate_comparison_barplots.py

@@ -55,16 +55,16 @@ def compare_models(models: List[str]):
         results[m] = md
 
     df = pd.DataFrame(results)
-    dfstr = df.applymap('{:,.3g}'.format)
-    fig, ax = plt.subplots(figsize=(6,22))
-    ax.axis('tight')
-    ax.axis('off')
-    the_table = ax.table(cellText=dfstr.values,
-                            rowLabels=dfstr.index,
-                            colLabels=dfstr.columns,
-                            loc="center")
-    fig.tight_layout()
-    fig.savefig(f"plots/comparisons/{v}/table.png")
+    #dfstr = df.applymap('{:,.3g}'.format)
+    #fig, ax = plt.subplots(figsize=(6,22))
+    #ax.axis('tight')
+    #ax.axis('off')
+    #the_table = ax.table(cellText=dfstr.values,
+    #                        rowLabels=dfstr.index,
+    #                        colLabels=dfstr.columns,
+    #                        loc="center")
+    #fig.tight_layout()
+    #fig.savefig(f"plots/comparisons/{v}/table.png")
 
     set_plotting_defaults()
 
@@ -94,7 +94,6 @@ def compare_models(models: List[str]):
 
     nfigs = len(barplot_groups)
     fig_width = [ 1  + int(len(bp["vars"]) / 4) for bp in barplot_groups]
-    from IPython import embed; embed()
     nrows, ncols = 4,4
     fig = plt.figure(figsize=(7, 8))
     gs = plt.GridSpec(nrows, ncols, figure=fig)

### 2024-12-13 scripts/label_arteries.py

@@ -29,7 +29,8 @@ def get_tangent(v, n):
     return v - np.dot(np.dot(v,n), n)
 
 if __name__=="__main__":
-   filename = "plots/meshplots/labeled_arteries.png"
+   filename = "plots/meshplots/arteries.png"
+   filename_labeled = "plots/meshplots/labeled_arteries.png"
    art = pv.read("mesh/networks/arteries_tubes.vtk").triangulate()
 
    art["radius"] *= 1e3
@@ -118,7 +119,7 @@ if __name__=="__main__":
    add_labels2d(ax, df[(df["left"]==False) & (df["y"] < 120)], xpos=20, cs ="angle,angleA=0,angleB=70,rad=50", offset=-40)
 
    ax.set_xmargin(0);ax.set_ymargin(0)
-   plt.savefig(filename, transparent=True, pad_inches=0.0,bbox_inches='tight', dpi=300)
+   plt.savefig(filename_labeled, transparent=True, pad_inches=0.0,bbox_inches='tight', dpi=300)
 
    print(pl.camera.direction)
    pl.export_html(filename.replace("png", "html"))

### 2024-12-13 scripts/make_3D_FTA.py

@@ -1,0 +1,95 @@
+from plotting_utils import get_result, read_config
+import pyvista as pv
+import typer
+import numpy as np
+from pathlib import Path
+from tqdm import tqdm
+from plot_csf_flow import from_k3d
+import k3d.colormaps.paraview_color_maps as pcm
+from extract_vessels import get_tubes
+import seaborn as sns
+import os
+from cmap import Colormap
+
+def make_3D_time_view(modelname:str):
+    config = read_config(f"configfiles/{modelname}.yml")
+    os.makedirs(f"plots/{modelname}/3D_FTA/", exist_ok=True)
+    nres = 300
+    times = np.array([1, 3, 6, 9, 12]) * 3600
+    clim = (0, times[-1] / 3600)
+    scalarbar_args = dict(color="black", title="First-time arrival (h)", 
+                          vertical=False, height=0.08, width=0.6,
+                          position_x=0.2, position_y=0.05, 
+                          title_font_size=36, use_opacity=False,
+                          label_font_size=28, fmt="%.1f")
+    meshfile = Path(config["mesh"])
+    skull = pv.read(meshfile.parent / "surfaces/skull.ply")
+    ventricles = pv.read(meshfile.parent / "surfaces/ventricles.ply")
+    art = get_result(modelname, "artery", times).ctp()
+    ven = get_result(modelname, "vein", times).ctp()
+    sas = get_result(modelname, "sas", times)
+    arteries = get_tubes(art)
+    veins = get_tubes(ven)
+
+    img = pv.ImageData(pv.create_grid(sas, dimensions=[nres]*3))
+    img = img.sample(sas, progress_bar=True)
+    img["FTA"] = np.full(img[f"c_{times[0]}"].shape, times[-1] + 10)
+    arteries["FTA"] = np.full(arteries[f"c_{times[0]}"].shape, np.nan)
+    veins["FTA"] = np.full(veins[f"c_{times[0]}"].shape, np.nan)
+    for t in times[::-1]:
+        print(t)
+        img["FTA"] = np.where(img[f"c_{t}"] > 0.1, t/3600, img["FTA"])
+        arteries["FTA"] = np.where(arteries[f"c_{t}"] > 0.1, t/3600, arteries["FTA"])
+        veins["FTA"] = np.where(veins[f"c_{t}"] > 0.1, t/3600, veins["FTA"])
+
+    pl = pv.Plotter(off_screen=True, window_size=(1200, 1200))
+    pl.set_background("white")
[…]

### 2024-12-13 scripts/make_3D_timeview.py

@@ -1,0 +1,76 @@
+from plotting_utils import get_result, read_config
+import pyvista as pv
+import typer
+import numpy as np
+from pathlib import Path
+from tqdm import tqdm
+from plot_csf_flow import from_k3d
+import k3d.colormaps.paraview_color_maps as pcm
+from extract_vessels import get_tubes
+import seaborn as sns
+import os
+
+def make_3D_time_view(modelname:str):
+    config = read_config(f"configfiles/{modelname}.yml")
+    os.makedirs(f"plots/{modelname}/timeview3D/", exist_ok=True)
+    colors = [ "#ff595e", "#ffca3a", "#8ac926", "#1982c4", "#6a4c93"]
+    getcmap = lambda color: sns.blend_palette(["white", color],as_cmap=True)
+    clim = (0,2)
+    nres = 350
+    times = np.array([1, 6, 12, 24]) * 3600
+    
+    meshfile = Path(config["mesh"])
+    skull = pv.read(meshfile.parent / "surfaces/skull.ply")
+    ventricles = pv.read(meshfile.parent / "surfaces/ventricles.ply")
+    art = get_result(modelname, "artery", times).ctp()
+    sas = get_result(modelname, "sas", times)
+    arteries = get_tubes(art)
+    for t in times:
+        sas[f"c_{t}"] = np.where(sas[f"c_{t}"] > clim[1], 
+                                 clim[1], sas[f"c_{t}"])
+    img = pv.ImageData(pv.create_grid(sas, dimensions=[nres]*3))
+    img = img.sample(sas, progress_bar=True)
+
+    for type in ["volume", "contour"]:
+        pl = pv.Plotter(off_screen=True, window_size=(1200, 1200))
+        pl.set_background("white")
+        pl.add_mesh(skull,style='points',point_size=0.1, opacity=0.6, show_scalar_bar=False,
+                    color="silver", specular=1, render_points_as_spheres=True)
+        pl.add_mesh(ventricles, style='points',point_size=0.8,
+                    show_scalar_bar=False,specular=1,
+                    render_points_as_spheres=True,
+                    color="silver")
+        #pl.add_mesh(arteries, color="silver",
+        #            clim=clim, show_scalar_bar=False,)
+        pl.camera_position = 'xz'
+        pl.camera.azimuth += 140
+        pl.camera.zoom(1.5)
+
+        # generate frames
+        for t, color in zip(times, colors):
[…]

### 2024-12-13 scripts/plot_conc_percentiles.py

@@ -113,7 +113,7 @@ def plot_conc_percentiles(models:List[str]):
                     xis.append(results[m]["config"]["arterial_pvs_csf_permability"])
                     r.append(results[m][q] / (3600 if "(h)" in ylabeldict[q] else 1))
                 ax.scatter(xis,r, color=col, label=labeld[gr[0]], marker=markerd[gr[0]],
-                          s=5)
+                          s=50)
                 ax.legend(loc='upper center', bbox_to_anchor=(0.5, 1.15), ncol=2, 
                         columnspacing=0.3, frameon=False)        
             plt.xlabel("PVS-CSF permeability (m/s)")

### 2024-12-15 scripts/plot_conc_percentiles.py
minor change

@@ -17,7 +17,7 @@ flip = lambda items, ncol: itertools.chain(*[items[i::ncol] for i in range(ncol)
 def plot_conc_percentiles(models:List[str]):
     dpi = 500
     group_a = ["modelA", ]#"modelB1-10", "modelB1-100", "modelB1-1000"]
-    group_b = ["modelA-strongVM", "modelB2-10",]# "modelB2-100", "modelB2-1000"]
+    group_b = ["modelA-strongVM", "modelB2-10", "modelB2-100",]# "modelB2-1000"]
     blues = sns.color_palette("Blues_d", n_colors=len(group_a))
     oranges = sns.color_palette("Oranges_d", n_colors=len(group_b))
     artlabels = [l for l,p in pointlabels]
@@ -95,30 +95,32 @@ def plot_conc_percentiles(models:List[str]):
             plt.close()
             
     qois = [f"_fta", f"_lag", f"_pvs_peak_time", f"_c_peak", f"_delta_c_peak"]
-    for q in qois:
-        fig, ax = plt.subplots(figsize=(4,3))
-        # plot qois over xi
-        for art in ["BA","MCA-R", "MCA-L", "ACA-A2", "ACA-A3"]:
-            ylabeldict = {f"{art}_fta" : f"{art} FTA (h)", f"{art}_lag" : f"{art} Δt (h)",
-                        f"{art}_pvs_peak_time": f"{art} peak time (h)", 
-                        f"{art}_c_peak": f"{art} " + r"$c_{\rm peak}$ (mmol/l)",
-                        f"{art}_delta_c_peak":f"{art} " + r"$\Delta c_{\rm peak}$  (mmol/l)"}
-            labeld = {group_a[0]: "baseline", group_b[0]:"baseline + VM"}
-            markerd = {group_a[0]: "d", group_b[0]:"o"}
+    ylabeldict = {f"_fta" : f"FTA (h)", f"_lag" : f"Δt (h)",
+                            f"_pvs_peak_time": f"peak time (h)", 
+                            f"_c_peak": r"$c_{\rm peak}$ (mmol/l)",
+                            f"_delta_c_peak":r"$\Delta c_{\rm peak}$  (mmol/l)"}
+    for gr in [group_b]:
+        va = "_".join(gr)
+        os.makedirs(f"plots/comparisons/{va}/", exist_ok=True)
+        for q in qois:
+            sns.set_palette("tab10")
+            fig, ax = plt.subplots(figsize=(4,3))
[…]

### 2024-12-15 paper/supplementary_main.tex
WIP 2

@@
 each of these domains, $c$ satisfies
+\mer{MER: Rami asks if it should be a $\phi$ in the advective term here.}
 \begin{equation}
@@
 \begin{align*}
-  \bm V_h &= \{ \bm v  \in H(\mathrm{div}, \Omega_{\mathrm{CSF}}): \;
+  \bm V_{h,g} &= \{ \bm v  \in H(\mathrm{div}, \Omega_{\mathrm{CSF}}): \;
   \bm v \vert_{E} \in \mathrm{BDM}^2(E), \; E \in \mathcal{T}_{\rm CSF}; \;
   \bm v \cdot \bm n = 0
-  \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\rm AM-L} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} \text{ on } \Gamma_{\mathrm{LV}} \}  \\ 
+  \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\rm AM-L} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n =  g %\frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} 
+  \text{ on } \Gamma_{\mathrm{LV}} \}  \\ 
   Q_h  &= \{q \in L^2(\Omega_{\rm CSF}): \; q \vert_E \in \mathrm{P}^1(E),
@@
 \end{align*}
-\mer{MER: How about the homogenized version of $\bm V_h$?
-- to distinguish for test space?} Here, $H(\mathrm{div}, \Omega)$ is the space of $L^2(\Omega)$ vector fields with $L^2(\Omega)$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, and $\bm n$ and $\bm \tau$ are the unit outward normal and unit tangent vectors to each facet, respectively.
+Here, $H(\mathrm{div}, \Omega_{\rm{CSF}})$ is the space of $L^2(\Omega_{\rm{CSF}})$ vector fields with $L^2(\Omega_{\rm{CSF}})$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, and $\bm n$ and $\bm \tau$ are the unit outward normal and unit tangent vectors to each facet, respectively.
 Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by
@@
 \end{equation}
-Since $\bm V_h \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm
+Since $\bm V_{h,g} \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm
   v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$. Continuity in the
-tangential component is enforced weakly via interior penalization. \mer{For
+tangential component is enforced weakly via interior penalization. For
 convenience, we collect all facets exterior to the CSF space but not
@@
 and on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
-$\mathcal{F}_{\rm SSAS}$.} Now, define the form 
+$\mathcal{F}_{\rm SSAS}$. Now, define the form 
 \begin{multline}
@@
   = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
-  - \sum_{F \in \mer{\mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}}} \left (
+  - \sum_{F \in \mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}} \left (
   \int_{F} \mu \{\nabla \bm u\} \cdot \bm n_F \cdot [\bm v_t] 
@@
 \end{multline}
-where on exterior facets \mer{the average and jump operators take the one-sided values}.
+where on exterior facets the average and jump operators take the one-sided values.
 We set the penalty parameter for the tangential continuity to be $\sigma = 20$.
-Our finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_h \times Q_h$ such that
+Our finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_{h,g} \times Q_h$ with $g  = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}$  such that
 \begin{subequations}
@@
   + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
-  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0 \qquad \foralls \bm v_h \in \bm V_h, \\ 
+  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0 \qquad \foralls \bm v_h \in \bm V_{h,0} \\ 
 \int_{\Omega_{\rm CSF}} \nabla \cdot \bm u_h \, q_h  &= 0 \qquad \foralls q_h \in Q_h.

### 2024-12-16 paper/supplementary_main.tex

@@
 \begin{equation}
-  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s (\hat c) ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
+  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s \hat c  ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
   \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
@@
 \begin{align}\label{eq:1d_dg_disc}
-\int_{\Lambda}  \frac{1}{\tau}(A  \hat c_h^n - A \hat c_h^{n-1}) \hat v  + \int_{\Lambda} \hat D A \partial_s \hat c_h^n  \partial_s \hat v^n - \int_{\Lambda} A \hat u \hat c_h^n \partial_s \hat v   +  \int_{\Lambda} \xi P (\hat c_h^n - \overline{c_h^n}) +a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) = 0.
+\int_{\Lambda}  \frac{1}{\tau}(A  \hat c_h^n - A \hat c_h^{n-1}) \hat v  + \int_{\Lambda} \hat D A \partial_s \hat c_h^n  \partial_s \hat v^n - \int_{\Lambda} A \hat u \hat c_h^n \partial_s \hat v   +  \int_{\Lambda} \xi P (\hat c_h^n - \overline{c_h^n}) \hat v +a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) = 0.
 \end{align}
- In the above, if $\hat u$ is non--zero on $\Lambda^i$, then the artificial diffusion stabilization term, for more details see for e.g. \cite[Section 12.6]{quarteroni2009numerical}: 
+ In the above, if $\hat u$ is non--zero on $\Lambda^i$, then the artificial diffusion $a_h^{\mathrm{stab}}$ stabilization term is nonzero  and is given below, for more details see for  \cite[Section 12.6]{quarteroni2009numerical}.
  \begin{align}
@@
 
- \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} comlemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0$ is the following:
+ \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} comlemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0 = 0 $ is the following:
 

### 2024-12-19 paper/supplementary_main.tex
Clean-up Dispersion in Methods - moving to the Supplementary.

@@
 
+\mer{MER: Moved the below from Results.}
+\mar{To examine cardiac-driven dispersion, we compute a steady CSF flow field at peak systolic blood inflow.
+We mimic the resulting reduction of CSF space volume by imposing an inflow of $6\,$ml/s \cite{baledent2014imaging, causemann2022human} and $0.31\,$ml/s \cite{vinje2019respiratory} across the upper arachnoid mater ($\Gamma_{\rm AM-U}$) and the lateral ventricle ($\Gamma_{\rm LV}$) surface, respectively.
+We allow free outflow into the spinal SAS ($\Gamma_{\rm SSAS}$).
+In addition, we upscale the obtained pressure field to account for the lack of pulsatile inertial effects in our model (see \Cref{sec:csf_dispersion}).}
+
+% respiratory pulsatility
+\mar{Employing the same methodology, we estimate the respiratory contribution to dispersive mixing by computing the CSF flow field driven by an inflow of $0.121\,$ml/s \cite{liu2024using} and $1\,$ml/s \cite{gutierrez2022effect} across the lateral ventricular wall and the arachnoid mater, respectively.}
+
 CSF pulsates in the SAS and ventricular system with each cardiac and respiratory cycle, leading to an increase in effective solute diffusivity via a process known as Taylor dispersion \cite{taylor1953dispersion, watson1983diffusion}.
@@
 The diffusion enhancement factor $R$ in $D^{\mathrm{eff}} = D^{\mathrm{Gad}}( 1 +R)$ is then estimated by $R \approx R_{\rm{cardiac}} + R_{\rm{resp}}$.
+
 % %\subsubsection{Cardiac Dispersion Factor Estimation}

### 2024-12-19 paper/supplementary_main.tex
Add a justification and discussion of the 1D no-flux condition.

@@
 \subset \R^3$ with $\Omega = \Omega_{\rm PAR} \cup \Omega_{\rm CSF}$
-(see also \Cref{fig:intracranial_domains}). The coordinates in these
+(\Cref{fig:results1}A, \Cref{fig:csf}A--C). The coordinates in these
 3D domains is denoted by $x$. The interface between the parenchyma and
@@
 inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$ and thus of
-width $R_2^i - R_1^i$. We denote the outer lateral surface of the
+width $R_2^i - R_1^i$. \discuss{We denote the outer lateral surface of the
 periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$,
-respectively. We omit the subscript $a$ or $v$ when referring to any
+respectively.} We omit the subscript $a$ or $v$ when referring to any
 such network, vessel segment or perivascular outer surface. We assume
@@
 each of these domains, $c$ satisfies
-\mer{MER: Rami asks if it should be a $\phi$ in the advective term here.}
+%\mer{MER: Rami asks if it should be a $\phi$ in the advective term here.
+We can include it, or drop it, but in the latter case, perhaps we should clarify that u is an average/superficial/Darcy velocity.}
 \begin{equation}
@@
 In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$).
-In the parenchyma, $\phi$ represents the extracellular space volume fraction.
+In the parenchyma, $\phi$ represents the extracellular space volume fraction, and thus $c$ here generally represents the intrinsic (in contrast to the superficial) concentration~\cite{sykova2008diffusion}.
 Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$.
@@
 
-Finally, we assume no molecular efflux at all end nodes $\mathcal{N}$
-via a homogeneous Neumann condition augmenting~\eqref{eq:1d_pde}:
+Finally, we impose a homogogenous Neumann condition at all network end
+nodes to augment~\eqref{eq:1d_pde}:
 \begin{equation}
@@
 \end{equation}
+This no-flux condition states that the network end nodes represent
+barriers for molecular efflux into or out of the perivascular network,
+and thus that all solute exchange between the PVSs and their
+surroundings takes place via the lateral outer PVS surface and is
+regulated by the exchange parameter $\xi$ cf.~\eqref{eq:3d_pde}
+and~\eqref{eq:1d_pde}. Larger particles have been observed to
+accumulate within the PVS as the surface arteries penetrate into the
+brain parenchyma~\cite{mestre2018flow, bedussi2018paravascular},
+and~\eqref{eq:1d_bcs} is appropriate to represent such
+behavior. However, to represent a continuously extending PVSs also
+along penetrating vessels, this condition should be revisited.
 
-\mk{What is justification of \eqref{eq:1d_bcs}?}
+%\mk{What is justification of \eqref{eq:1d_bcs}?}
 %% Both $\Omega$ and $\Lambda$ are fixed in time and we assume

### 2024-12-19 paper/supplementary_main.tex
Fix a bunch of labels and refs.

@@
 \subsection{Mixed-dimensional transport and flow equations}
-\label{sec:more_maths}
+\label{sec:app:more_maths}
 
@@
 \end{equation}
-Moreover, to ensure \discuss{mass conservation}, we set that the flux
+Moreover, to ensure mass conservation, we set that the flux
 going in and out at nodes should add to zero, or more precisely, that
@@
     - \mu \Delta \bm u  + \nabla p &=  0 \quad && \text{ in } \Omega_{\rm CSF},
-    \label{eq:momnetum_equation} \\ 
+    \label{eq:momentum_equation} \\ 
     \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF},
@@
 where $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the (dynamic) CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$.
-We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:supp:dispersion}.
+We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:app:dispersion}.
 
@@
 
-We solve the models presented in~\Cref{sec:more_maths} numerically
+We solve the models presented in~\Cref{sec:app:more_maths} numerically
 using finite element methods. The following subsections provide more
@@
 \subsection{Modelling dispersion in the CSF spaces}
-\label{sec:supp:dispersion}
+\label{sec:app:dispersion}
 
@@
 We allow free outflow into the spinal SAS ($\Gamma_{\rm SSAS}$).
-In addition, we upscale the obtained pressure field to account for the lack of pulsatile inertial effects in our model (see \Cref{sec:csf_dispersion}).}
+In addition, we upscale the obtained pressure field to account for the lack of pulsatile inertial effects in our model.}
 
@@
 We prescribe a no-slip boundary condition on the pial surface $\Gamma_{\rm Pia}$ and zero-traction on the spinal boundary $\Gamma_{\rm SSAS}$, making the spinal compartment the only outflow route of CSF.
-Thus, we compute the pulsatile flow fields by augmenting the Stokes flow model (eq.
-\ref{eq:momnetum_equation} and \ref{eq:divergence_equation}) with the following boundary conditions:
+Thus, we compute the pulsatile flow fields by augmenting the Stokes flow model (\eqref{eq:momentum_equation} and \ref{eq:divergence_equation}) with the following boundary conditions:
 \begin{equation}\label{eq:cardiac_flow_bcs}

### 2024-12-19 paper/supplementary_main.tex
Fix in A.1.4.

@@
 inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$ and thus of
-width $R_2^i - R_1^i$. \discuss{We denote the outer lateral surface of the
+width $R_2^i - R_1^i$. We denote the outer lateral surface of the
 periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$,
-respectively.} We omit the subscript $a$ or $v$ when referring to any
+respectively. We omit the subscript $a$ or $v$ when referring to any
 such network, vessel segment or perivascular outer surface. We assume
@@
 
-%\mk{What is justification of \eqref{eq:1d_bcs}?}
-%% Both $\Omega$ and $\Lambda$ are fixed in time and we assume
-%% constant-in- time radii (despite varying radii to drive peristaltic
-%% flow, see \cref{sec:sup:peristalsis}).
-
 \subsubsection{Modelling CSF flow via the incompressible Stokes equations}
@@
 
-\subsubsection{Steady flow in perivascular networks -- induced by pressure differences}
+\subsubsection{Steady flow in perivascular networks induced by pressure differences}
 
@@
 \end{subequations}%
-where $A$ is the area of the PVS cross-sections, $\mu$ is the CSF viscosity as before, and $\kappa$ is derived from an assumption of Poiseuille flow in the annular cross-section of the PVS as~\cite{tithof2022network}:  \rami{Here,  we consider the steady state  case and we assume that the PVS domain is fixed and not pulsating  }
+where $A$ is the area of the PVS cross-sections, $\mu$ is the CSF viscosity as before, and $\kappa$ is derived from an assumption of Poiseuille flow in the annular cross-section of the PVS as~\cite{tithof2022network}:
 \begin{equation}
@@
 (and its harmonic extensions into $\Omega_{\rm PAR}$) as the given
-fluid pressures $p_0$.
+fluid pressures $p_0$. Note that we here consider steady PVS flow in a
+non-moving domain; the effect of domain motion is addressed
+in~\Cref{sec:app:peristalsis}.
 
@@
 \subsection{Estimating net perivascular flow induced by peristaltic waves}
-\label{sec:sup:peristalsis}
+\label{sec:app:peristalsis}
 
@@
 edge-disjoint subnetworks $\Lambda^1, \Lambda^2, \Lambda^3$, one for
-each of the supply nodes (\Cref{fig:sup:peristalsis}).
+each of the supply nodes (\Cref{fig:app:peristalsis}).
 Each node is assigned to the subnetwork associated with the nearest supply node, and edges between nodes are preserved.

### 2024-12-19 paper/supplementary_main.tex
Make description of dispersion very structured.

@@
 
-\mer{MER: Stopped here.}
-
-\subsection{Modelling dispersion in the CSF spaces}
+\subsection{Estimating dispersion factors from pulsatile CSF flow}
 \label{sec:app:dispersion}
 
-\mer{MER: Moved the below from Results.}
+The cardiac and respiratory cycles (typically 0.25--1.0 Hz in humans)
+induce pulsatile flow of CSF in the ventricular system and in the
+cranial and spinal SAS. Pulsatile flow leads to dispersion which in
+turn may enhance molecular transport~\cite{taylor1953dispersion,
+  watson1983diffusion, asgari2016glymphatic, keith2019dispersion,
+  ray2021quantitative, troyetsky2021dispersion}. To account for the
+dispersive effects over a longer time scale (hours to days) and with
+spatial differences, we propose the following four-step approach for
+the cardiac and respiratory contributions separately.
+\begin{enumerate}%[\label=\roman]
+\item
+  To account for viscous forces, we compute spatially-varying CSF
+  pressure fields $p^0_c, p^0_r$ in $\Omega_{\rm CSF}$ corresponding to the
+  Stokes flow induced by the peak volumetric reduction of the CSF
+  space in the respective cycle. This step is described in further
+  detail below.
+\item
+  Next, we estimate the Womersley numbers $\alpha_c, \alpha_r$
+  associated with the cardiac and respiratory flow patterns by the
+  definition
+  \begin{equation}
+    \alpha^2_i = \frac{h^2 \omega_i \rho}{\mu} \qquad i \in \{c, r\},
+  \end{equation}
+  with CSF density $\rho = 10^3$ kg/m$^3$, a mean CSF space width $h =
+  1.5$mm, and CSF viscosity $\mu$ given in~\Cref{tab:parameters}.  For
+  the cardiac cycle, we consider an angular frequency $\omega_c = 2
+  \pi$, while for the respiratory cycle, we set $\omega_r = 0.5
+  \pi$. The resulting (square) Womersley numbers are $\alpha_c^2 = 20.2$
+  for the cardiac cycle and $\alpha_r^2 = 5.05$ for the respiratory
+  cycle.
+\item
+  To account for inertial forces in addition to the viscous forces, we
+  use the respective Womersley number $\alpha_i$ to calculate an
+  upscaled pressure field $p^1$ from $p^0$:
+  \begin{equation}
+    p^1_i(x) = \left ( 1 + \frac{\alpha_i^2}{8} \right ) p_i^0(x)
+    \qquad \text{for } x \in \Omega_{\rm CSF} .
+  \end{equation}
+\item
+  Further, assuming unsteady dispersion, we follow Sharp et
+  al.~\cite{keith2019dispersion} to estimate local enhancement factors
+  $S_i$ from the non-dimensionalized pressure gradients:
+  \begin{equation}
+    S_i = \frac{1}{\alpha_i^3} \frac{1}{\omega_i \mu/h} | \nabla p_i^1 |.
+  \end{equation}
+\item
+  Finally, we account for the non-local nature of dispersion by
+  smoothing ...
+\end{enumerate}
+
 \mar{To examine cardiac-driven dispersion, we compute a steady CSF flow field at peak systolic blood inflow.
@@
 \mar{Employing the same methodology, we estimate the respiratory contribution to dispersive mixing by computing the CSF flow field driven by an inflow of $0.121\,$ml/s \cite{liu2024using} and $1\,$ml/s \cite{gutierrez2022effect} across the lateral ventricular wall and the arachnoid mater, respectively.}
-
-CSF pulsates in the SAS and ventricular system with each cardiac and respiratory cycle, leading to an increase in effective solute diffusivity via a process known as Taylor dispersion \cite{taylor1953dispersion, watson1983diffusion}.
-Previous estimates of the magnitude of this effect in the CSF-filled space range from an enhancement factor of 5.8 \cite{keith2019dispersion} up to $10^4$ \cite{ray2021quantitative}.
-The large discrepancies in reported estimates can (at least partly) be attributed to different assumptions on representative values for local width, pressure differences and fluid velocities in the SAS, which reflect the diverse flow characteristics in the different regions of the intracranial CSF-filled space.
 

### 2024-12-19 paper/supplementary_main.tex
Finalize dispersion clean-up in supplementary.

--- 2024-12-19 paper/supplementary_main.tex
@@
 
-\mer{MER: Stopped here.}
-
-\subsection{Modelling dispersion in the CSF spaces}
+\subsection{Estimating dispersion factors from pulsatile CSF flow}
 \label{sec:app:dispersion}
 
-\mer{MER: Moved the below from Results.}
-\mar{To examine cardiac-driven dispersion, we compute a steady CSF flow field at peak systolic blood inflow.
-We mimic the resulting reduction of CSF space volume by imposing an inflow of $6\,$ml/s \cite{baledent2014imaging, causemann2022human} and $0.31\,$ml/s \cite{vinje2019respiratory} across the upper arachnoid mater ($\Gamma_{\rm AM-U}$) and the lateral ventricle ($\Gamma_{\rm LV}$) surface, respectively.
-We allow free outflow into the spinal SAS ($\Gamma_{\rm SSAS}$).
-In addition, we upscale the obtained pressure field to account for the lack of pulsatile inertial effects in our model.}
-
-% respiratory pulsatility
-\mar{Employing the same methodology, we estimate the respiratory contribution to dispersive mixing by computing the CSF flow field driven by an inflow of $0.121\,$ml/s \cite{liu2024using} and $1\,$ml/s \cite{gutierrez2022effect} across the lateral ventricular wall and the arachnoid mater, respectively.}
-
-CSF pulsates in the SAS and ventricular system with each cardiac and respiratory cycle, leading to an increase in effective solute diffusivity via a process known as Taylor dispersion \cite{taylor1953dispersion, watson1983diffusion}.
-Previous estimates of the magnitude of this effect in the CSF-filled space range from an enhancement factor of 5.8 \cite{keith2019dispersion} up to $10^4$ \cite{ray2021quantitative}.
-The large discrepancies in reported estimates can (at least partly) be attributed to different assumptions on representative values for local width, pressure differences and fluid velocities in the SAS, which reflect the diverse flow characteristics in the different regions of the intracranial CSF-filled space.
-
-Taking the large spatial differences of pulsatile CSF flow into account, we estimate a spatially varying diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac- and respiration-induced pulsatile CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}.
-These estimates are are derived under simplifying assumptions on the problem geometry, the flow field and concentration profiles \cite{watson1983diffusion, keith2019dispersion}.
-Therefore such an estimate is heuristic; however, our computational results qualitatively agree with observed data.
-
-Similar to the case of CSF flow driven by production introduced in \Cref{sec:csf_flow}, we employ a Stokes flow model to predict CSF velocities corresponding to the peak cardiac and respiratory blood inflow rates.
-In particular, we enforce a uniformly distributed inflow of $u^{\rm in}_{\rm AM}$ and $u^{\rm in}_{\rm LV}$ across the arachnoid mater ($\Gamma_{\rm AM}$) and lateral ventricular surfaces ($\Gamma_{\rm LV}$) respectively, mimicking the effect of reduced CSF volume due to blood inflow.
-We set $u^{\rm in}_{\rm AM} = 6\,$ml/s \cite{causemann2022human,baledent2014imaging} and $u^{\rm in}_{\rm LV}=0.31\,$ml/s \cite{vinje2019respiratory} in the cardiac case, and $u^{\rm{in}}_{\rm{AM}} = 1\,$ml/s \cite{gutierrez2022effect} and $u^{\rm in}_{\rm{LV}}=0.121\,$ml/s \cite{liu2024using} in the respiratory case, respectively.
-We prescribe a no-slip boundary condition on the pial surface $\Gamma_{\rm Pia}$ and zero-traction on the spinal boundary $\Gamma_{\rm SSAS}$, making the spinal compartment the only outflow route of CSF.
-Thus, we compute the pulsatile flow fields by augmenting the Stokes flow model (\eqref{eq:momentum_equation} and \ref{eq:divergence_equation}) with the following boundary conditions:
-\begin{equation}\label{eq:cardiac_flow_bcs}
+The cardiac ($\sim$1Hz) and respiratory ($\sim$0.25Hz) cycles induce
+pulsatile flow of CSF in the ventricular system and in the cranial and
+spinal SAS. Pulsatile flow leads to dispersion which in turn may
+enhance molecular transport~\cite{taylor1953dispersion,
+  watson1983diffusion, asgari2016glymphatic, keith2019dispersion,
+  ray2021quantitative, troyetsky2021dispersion}. To account for the
+dispersive effects over a longer time scale (hours to days), and in
+the absence of measurements or estimates of dispersion coefficients in
+human CSF spaces, we adapt existing theoretical
+estimates~\cite{watson1983diffusion, keith2019dispersion}. More
+specifically, we compute spatially-varying dispersion enhancement
+fields $R_c$ and $R_r$ (\Cref{fig:csf}G, I), associated with the
+cardiac and respiratory cycles respectively, via the algorithm
+presented below. These fields then contribute to the diffusion
+coefficient in $\Omega_{\rm CSF}$ in~\eqref{eq:3d_pde} as $D = (1 +
+R_c + R_r) D^{\rm Gad}$.
+\begin{enumerate}[label=\roman*)]
+\item
+  To account for viscous forces, we compute spatially-varying CSF
+  pressure fields $p^0_c, p^0_r$ in $\Omega_{\rm CSF}$ corresponding
+  to the Stokes flow induced by the peak volumetric reduction of the
+  CSF space in the respective cycle (\Cref{fig:csf}F, H). More
+  precisely, we numerically solve the incompressible Stokes
+  equations~\eqref{eq:stokes} equipped with the following boundary
+  conditions mimicking a dilation of the brain parenchyma with the
+  spinal SAS as the only route for CSF efflux:
+  \begin{equation}\label{eq:pulsatility:bcs}
     \bm u = 0 \text{ on } \Gamma_{\rm Pia}; \quad 
@@
 \end{equation}
-The obtained pressure fields $p_{\rm cardiac}$ and $p_{\rm resp}$ only account for viscous forces, albeit cardiac and respiratory induced CSF flow is known to be unsteady and substantially impacted by transient inertial forces.
+  In the cardiac cycle case, we set $u^{\rm in}_{\rm AM} = 6\,$ml/s \cite{causemann2022human,baledent2014imaging} and $u^{\rm in}_{\rm LV}=0.31\,$ml/s \cite{vinje2019respiratory} to solve for $p^0_c$.
+In the respiratory cycle case, we set $u^{\rm{in}}_{\rm{AM}} = 1\,$ml/s \cite{gutierrez2022effect} and $u^{\rm in}_{\rm{LV}}=0.121\,$ml/s \cite{liu2024using} to solve for $p^0_r$.
+\item
+  We also estimate the Womersley numbers $\alpha_c, \alpha_r$
+  associated with the cardiac and respiratory flow patterns,
+  respectively, by the definition
+  \begin{equation}
+    \alpha^2_i = \frac{h^2 \omega_i \rho}{\mu} \qquad i \in \{c, r\},
+  \end{equation}
+  with CSF density $\rho = 10^3$ kg/m$^3$, a mean CSF space width $h =
+  1.5$mm, and CSF viscosity $\mu$ given in~\Cref{tab:parameters}.  For
+  the cardiac cycle, we consider an angular frequency $\omega_c = 2
+  \pi$, while for the respiratory cycle, we set $\omega_r = 0.5
+  \pi$. The resulting (square) Womersley numbers are $\alpha_c^2 = 20.2$
+  for the cardiac cycle and $\alpha_r^2 = 5.05$ for the respiratory
+  cycle.
+\item
+  To account for inertial forces in addition to the viscous forces, we
+  use the Womersley numbers $\alpha_c, \alpha_r$ to calculate upscaled
+  pressure fields $p^1_c, p^1_r$ from $p^0_c, p^0_r$ as:
+  \begin{equation}
+    p^1_i(x) = \left ( 1 + \frac{\alpha_i^2}{8} \right ) p_i^0(x)
+    \qquad \text{for } x \in \Omega_{\rm CSF}, \qquad i \in \{c, r\}.
+  \end{equation}
+  Note that this scaling is based on theoretical considerations on the
+  ratio of oscillatory flow to steady flow impedances in a
+  tube~\cite[Chap.~4.3.]{van1998cardiovascular}.
+\item
+  Further, assuming unsteady dispersion, we follow Sharp et
+  al.~\cite{keith2019dispersion} to estimate local enhancement factors
+  $S_c, S_r$ from the non-dimensionalized pressure gradients:
+  \begin{equation}
+    S_i(x) = \frac{1}{\alpha_i^3} \frac{1}{\omega_i \mu/h} | \nabla p_i^1(x) |
+    \qquad i \in \{c, r\}.
+  \end{equation}
+\item
+  Finally, we define the cardiac and respiratory dispersion
+  enhancement factors $R_c$ and $R_r$ by smoothing $S_c$ and $S_r$,
+  respectively, to account for the non-local nature of
+  dispersion. Specifically, for $i \in \{c, r\}$, we define $R_i :
+  \Omega_{\rm CSF} \rightarrow \R^{+}$ by solving a heuristic weighted
+  Helmholtz problem over $\Omega_{\rm CSF}$ with $S_i$ as the
+  right-hand side:
+  \begin{equation}
+    - 10^{-4} \Delta R_i + R_i
+    = S_i \quad \text{ on } \Omega_{\rm CSF}
+    \qquad \text{ and } \quad
+    \nabla R_i \cdot \bm n = 0, \quad \text{ on } \partial \Omega_{\rm CSF} .
+\end{equation}
+\end{enumerate}
+Considering the uncertainty associated with the validity of
+simplifying assumptions, the resulting estimates of the cardiac and
+respiratory dispersion factors $R_c$ and $R_r$ 
+should be viewed as heuristic rather than absolute.
+
+%% \mar{To examine cardiac-driven dispersion, we compute a steady CSF flow field at peak systolic blood inflow.
+We mimic the resulting reduction of CSF space volume by imposing an inflow of $6\,$ml/s \cite{baledent2014imaging, causemann2022human} and $0.31\,$ml/s \cite{vinje2019respiratory} across the upper arachnoid mater ($\Gamma_{\rm AM-U}$) and the lateral ventricle ($\Gamma_{\rm LV}$) surface, respectively.
+We allow free outflow into the spinal SAS ($\Gamma_{\rm SSAS}$).
+In addition, we upscale the obtained pressure field to account for the lack of pulsatile inertial effects in our model.}
+
+%% % respiratory pulsatility
+%% \mar{Employing the same methodology, we estimate the respiratory contribution to dispersive mixing by computing the CSF flow field driven by an inflow of $0.121\,$ml/s \cite{liu2024using} and $1\,$ml/s \cite{gutierrez2022effect} across the lateral ventricular wall and the arachnoid mater, respectively.}
+
+%% Taking the large spatial differences of pulsatile CSF flow into account, we estimate a spatially varying diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac- and respiration-induced pulsatile CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}.
+These estimates are are derived under simplifying assumptions on the problem geometry, the flow field and concentration profiles \cite{watson1983diffusion, keith2019dispersion}.
+Therefore such an estimate is heuristic; however, our computational results qualitatively agree with observed data.
+
+%% Similar to the case of CSF flow driven by production introduced in \Cref{sec:csf_flow}, we employ a Stokes flow model to predict CSF velocities corresponding to the peak cardiac and respiratory blood inflow rates.
+In particular, we enforce a uniformly distributed inflow of $u^{\rm in}_{\rm AM}$ and $u^{\rm in}_{\rm LV}$ across the arachnoid mater ($\Gamma_{\rm AM}$) and lateral ventricular surfaces ($\Gamma_{\rm LV}$) respectively, mimicking the effect of reduced CSF volume due to blood inflow.
+We set $u^{\rm in}_{\rm AM} = 6\,$ml/s \cite{causemann2022human,baledent2014imaging} and $u^{\rm in}_{\rm LV}=0.31\,$ml/s \cite{vinje2019respiratory} in the cardiac case, and $u^{\rm{in}}_{\rm{AM}} = 1\,$ml/s \cite{gutierrez2022effect} and $u^{\rm in}_{\rm{LV}}=0.121\,$ml/s \cite{liu2024using} in the respiratory case, respectively.
+%% We prescribe a no-slip boundary condition on the pial surface $\Gamma_{\rm Pia}$ and zero-traction on the spinal boundary $\Gamma_{\rm SSAS}$, making the spinal compartment the only outflow route of CSF.
+Thus, we compute the pulsatile flow fields by augmenting the Stokes flow model (\eqref{eq:momentum_equation} and \ref{eq:divergence_equation}) with the following boundary conditions:
+%% \begin{equation}\label{eq:cardiac_flow_bcs}
+%%     \bm u = 0 \text{ on } \Gamma_{\rm Pia}; \quad 
+%%     \mu \nabla \bm u \cdot \bm n - p \bm n = 0 \text{ on } \Gamma_{\rm SSAS}; \quad 
+%%     \bm u = \frac{u^{\rm in}_{\rm LV} \cdot \bm n}{|\Gamma_{\rm LV}|} \text{ on } \Gamma_{\rm LV}; \quad
+%%     \bm u = \frac{u^{\rm in}_{\rm AM} \cdot \bm n}{|\Gamma_{\rm skull}|} \text{ on } \Gamma_{\rm AM}.
+%% \end{equation}
+
+%% The obtained pressure fields $p_{\rm cardiac}$ and $p_{\rm resp}$ only account for viscous forces, albeit cardiac and respiratory induced CSF flow is known to be unsteady and substantially impacted by transient inertial forces.
 Assuming an oscillatory flow with either the heartbeat frequency $\omega_{\rm cardiac} = 2 \pi$ or respiratory frequency $\omega_{\rm resp} = 0.5 \pi$, a mean gap width of the CSF-filled spaces of $2h=3$\,mm and CSF density $\rho=10^3\,\text{kg/m}^3$, the squared Womersley number $\alpha_i^2 = \frac{h^2 \omega_{i} \rho}{\mu}$ becomes $\alpha_{\rm cardiac}^2 = 20.2$ and $\alpha_{\rm resp}^2=5.05$, which is unsteady.
@@
 
-Further assuming unsteady dispersion, we follow Sharp et al \cite{keith2019dispersion} and estimate each enhancement factor $R_{\rm cardiac}$ and $R_{\rm resp}$ from the non-dimensionalized pressure gradient
-$$R_{i} \approx dP_{i}^2 / \alpha_{i}^3, \quad  \text{where} \quad  dP=\frac{|\nabla p^{\rm up}_{i}|}{\omega_{i} \mu / h} \quad \text{for} \quad i = \{\mathrm{cardiac}, \mathrm{resp}\}.$$
-Finally, we account for the non-local nature of the dispersion mechanism and smooth the resulting field by solving a Helmholtz problem with $dP+i^2/\alpha_i^3$ as the right-hand side
-\begin{equation}
-- 10^{-4} \Delta R_i + R_i = dP_i^2/\alpha_i^3 \quad \text{ on } \Omega_{\rm CSF} \qquad \text{ and } \quad \nabla R_i \cdot \bm n = 0, \quad \text{ on } \partial \Omega_{\rm CSF}
-\end{equation}
-to obtain the locally averaged, spatially varying diffusion enhancement factor $R_i$.
+%% Further assuming unsteady dispersion, we follow Sharp et al \cite{keith2019dispersion} and estimate each enhancement factor $R_{\rm cardiac}$ and $R_{\rm resp}$ from the non-dimensionalized pressure gradient
+%% $$R_{i} \approx dP_{i}^2 / \alpha_{i}^3, \quad  \text{where} \quad  dP=\frac{|\nabla p^{\rm up}_{i}|}{\omega_{i} \mu / h} \quad \text{for} \quad i = \{\mathrm{cardiac}, \mathrm{resp}\}.$$
+%% Finally, we account for the non-local nature of the dispersion mechanism and smooth the resulting field by solving a Helmholtz problem with $dP+i^2/\alpha_i^3$ as the right-hand side
+%% \begin{equation}
+%% - 10^{-4} \Delta R_i + R_i = dP_i^2/\alpha_i^3 \quad \text{ on } \Omega_{\rm CSF} \qquad \text{ and } \quad \nabla R_i \cdot \bm n = 0, \quad \text{ on } \partial \Omega_{\rm CSF}
+%% \end{equation}
+%% to obtain the locally averaged, spatially varying diffusion enhancement factor $R_i$.
 The diffusion enhancement factor $R$ in $D^{\mathrm{eff}} = D^{\mathrm{Gad}}( 1 +R)$ is then estimated by $R \approx R_{\rm{cardiac}} + R_{\rm{resp}}$.

### 2024-12-19 paper/cut.tex
Move cut text from Results 2 and 1 to cut.tex and map out word count.
Clean-up in peristalsis in Methods

@@
+\section{Notes}
+
+Various notes associated with \cite{eide2024functional}:
+\begin{itemize}
+  \item
+    Can observations of tracer indicate compartmentalization without
+    the existence of physical barriers but in the presence of
+    e.g.~substantial flow? We made observations in this direction in
+    \cite{vinje2021brain}, relating to the then-active
+    shape-of-perivascular-spaces discussion. We can address this by
+    trying to quantify and evaluate the permeability of the PVS-SAS
+    interface - comparing model predictions with the timings and
+    patterns presented by \cite{eide2024functional}.
+  \item
+    Do the tracer primarily move from the PVS into the SAS or along
+    (more slowly) along the SAS? This is left as an open question by
+    \cite{eide2024functional} (page 3)
+  \item
+    How do altered ICP affect transport? Note that PVSAS transport is
+    impaired (delayed) with increasing mean wave amplitude in the ICP
+    signal and for iNPH patients (with enlarged PVSAS spaces).
+  \item
+    Very useful reference for PVS sizes (10--20 mm$^2$ in reference
+    patients, 15-70 mm$^2$ in iNPH), and average propagation speeds.
+\end{itemize}
+
+%Flow of CSF through the SAS and the ventricular system is a crucial driver of transport in the human brain environment.
+%While cardiac- and respiratory induced oscillatory flow increases mixing, the production of CSF in the choroid plexus creates a steady advective flow within the CSF-filled spaces of the cranium \cite{hornkjol2022csf}.
+%We construct a 3D model of the SAS and the ventricular system (\Cref{fig:csf}A,B and C) and compute CSF flow fields for three different scenarios: CSF production in the choroid plexus, CSF flow at systolic peak blood inflow and CSF flow in response to free breathing.
+
+% production flow simulation
+%% To simulate the production-driven flow, we impose an inflow of $400\,$ml/day across the surface of the lateral ventricle and allow a resistive outflow across the upper convexity of the arachnoid mater, leading to a total pressure drop of $26\,$mPa between the production and outflow sites (\Cref{fig:csf}D).
+A maximum flow velocity of $1.8\,$mm/s is attained in the aqueduct, while the mean velocity in the SAS is $2.6\,\mu$m/s (\Cref{fig:csf} E).
+
+%% % cardiac pulsatility
+%% Assessing the effect of cardiac-driven dispersion, we compute a steady CSF flow field at peak systolic blood inflow.
+We mimic the resulting reduction of CSF space volume by imposing an inflow of $6\,$ml/s \cite{baledent2014imaging, causemann2022human} and $0.31\,$ml/s \cite{vinje2019respiratory} across the upper arachnoid mater ($\Gamma_{\rm AM-U}$) and the lateral ventricle ($\Gamma_{\rm VL}$) surface, respectively.
+We allow free outflow into the spinal SAS ($\Gamma_{\rm SSAS}$).
+In addition, we upscale the obtained pressure field to account for the lack of pulsatile inertial effects in our model (see \Cref{sec:csf_dispersion}).
+In this scenario, we observe a pressure drop of $10\,$Pa between lateral ventricles and the spinal SAS and a maximum flow velocity of $19.8\,$cm/s in the aqueduct (\Cref{fig:csf}F and J).
+%% Next, we estimate a spatially varying dispersion enhancement factor $R$ from cardiac-induced pulsatility by combining the computed pressure field with the theory of shear-augmented dispersion (see \Cref{sec:csf_dispersion}).
+Assuming a cardiac frequency of $1\,$Hz, we find that cardiac-induced pulsatile CSF flow increases the effective diffusion by more than two orders of magnitude in the aqueduct and around the cisterna magna, but has little effect in most of the SAS ($R < 1$) (\Cref{fig:csf} G and K).
+
+%% % respiratory pulsatility
+%% Employing the same methodology, we estimate the respiratory contribution to dispersive mixing by computing the CSF flow field driven by an inflow of $0.121\,$ml/s \cite{liu2024using} and $1\,$ml/s \cite{gutierrez2022effect} across the lateral ventricular wall and the arachnoid mater, respectively.
+This amounts to a respiratory peak flow volume of $1.121\,$ ml/s at the craniocervical junction.
+While the resulting flow velocities are only about one fourth of their cardiac-induced counterparts (see \Cref{fig:csf}H and L), respiratory dispersive mixing reaches a factor of up to 350 due the lower respiratory frequency of $0.25\,$Hz (\Cref{fig:csf}I and M).
+
+%% % tracer trajectory of the standard model
+%% \mer{MER: I moved this to Results 1 and condensed it a little bit}
+%% \mar{Next, we use the computed advective velocity field and dispersion enhancement factors to predict the spreading of $0.5\,$mmol intrathecally injected Gadobutrol in CSF-filled spaces, the parenchymal tissue and PVSs around arteries and veines.
+We assume pressure-driven flow in arterial and venous PVSs, and cardiac peristaltic flow in the arterial network (\Cref{sec:pvs_flow_results}).
+One hour post-injection, tracer moves upwards in the SAS frontally of the brainstem,  and quickly reaches supratentorial regions (\Cref{fig:csf}N).
+Here, it spreads both posteriorly through quadrigeminal cistern and longitudinal fissure, and anteriorly through the outer SAS, reaching the top of the cerebral cortex after around 12 h.
+After 24h, the tracer covers most of the brains surface (with exception of some posterior regions) and has penetrated several centimeters into the parenchymal tissue.
+%% This pattern of transport is reflected in the mean tracer concentrations in each compartment, where the CSF space reaches its peak of $1.4\,$mmol/l after 2 hours, followed by the the arterial PVS concentration peak ($2.9\,$mmol/l after 4 hours) (\Cref{fig:csf}P).
+While the mean concentrations in these compartments fall quickly soon after their peaks, the parenchymal tissue slowly enriches with tracer over the first 24 hours, with a final value of $0.3\,$ mmol/l.
+%% Overall, about 40\,\% of the total amount of tracer remains in the cranium 24\,h post-injection, with the largest share in the CSF (53\,\%), followed by the parenchyma (35\,\%), the venous PVS (8\,\%) and the arterial PVS (5\,\%).
+(\cref{fig:csf}O).}
+
  \section{Introduction}

### 2024-12-20 paper/supplementary_main.tex

@@
 These sections provide more detailed descriptions of the mathematical models and numerical approximations considered.
+
+\subsection*{Overview of in-silico scenarios and experiments considered}
+
+See~\Cref{tab:scenarios} for an overview of the different in-silico
+scenarios and computational models considered.
+\begin{table}
+  \begin{center}
+  \begin{tabular}{ll|llllll|ll}
+    \toprule
+    & Description & $R_1, R_2$ & $\xi_{\rm PVS-SAS}$ &  $\beta_{\rm efflux}$ & $D$ & $\hat{u}$ & $\mathbf{u}_{SAS}$ & $c_0$ \\
+    \midrule
+    A & Baseline with sheaths & $R_2 = 2 R_1$ & $\xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
+    B1 & Sheath permeability & $R_2 = 2 R_1$ & $0, \xi, 10 \xi, 100 \xi, 1000 \xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
+    B2 & Health permeability & $R_2 = 2 R_1$ & $0, \xi, 10 \xi, 100 \xi, 1000 \xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} +  \hat{u}_{\rm cardiac} + \hat{u}_{\rm vaso}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
+    %C & Larger PVS & $R_2 = N R_1$ & -- & -- & --  &  --  & -- &  -- \\
+    C & Higher PVS flow & -- & $\xi$ & -- & --  & $\hat{u}_{\rm prod} +  \hat{u}_{\rm cardiac} + \hat{u}_{\rm vaso}$  & -- & --  \\
+    AA & PVS size & $R_2 = x \beta R_1$ & $\xi$ & -- & -- & -- &  -- & 0 \\ 
+    %C & Larger PVS & $R_2 = N R_1$ & -- & -- & --  &  --  & -- &  -- \\
+    B & PVS flow & -- & $\xi$ & -- & --  & $\hat{u}_{\rm prod} + \hat{u}_{\rm vaso}$  & -- & --  \\
+    C & Clearance & -- & -- & -- & $\alpha D^{\rm gad, \tau, A\beta}$  & $\hat{u}_{\rm prod}$ & -- & 1 \\
+    %E & PVS II & $R_2 = 2 R_1$ & $\xi$, -- & -- & --  &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow\uparrow$  & -- & --  \\
+    %F & PVS III & $R_2 = 2 R_1$ & $\xi$, -- & -- & $10 D_{\rm PVS}$ &  $\hat{u} = \hat{u}_{\rm prod}$  & -- & --  \\
+    %G & Glymphatics & $R_2 = 2 R_1$ & $\xi$, -- & -- & \cite{sykova2008diffusion, valnes2020apparent} &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow$  & $\mathbf{u}_{\rm PAR}$ > 0 & --  \\
+    \bottomrule
+    \end{tabular}
+    \end{center}
+  \caption{Overview of models (-- denotes as immediately above). Each
+    PVS cross-section is modelled as a concentric annulus with inner
+    radius $R_1$ and outer radius $R_2$ and such that $R_2 = \beta
+    R_1$ with $\beta = 2.0$ based on~\cite{eide2024functional}. Let
+    $\xi = \dots$ denote the permeability estimated by Koch et
+    al.~\cite{koch2023estimates}. Model A and its variations aims to
+    quantify the effect of PVS sheaths by varying the permeability
+    $\xi$ between the PVS and its surroundings. We consider a fixed
+    permeability $\xi_{\rm PVS-PAR}$ based
+    on~\cite{koch2023estimates}. \mer{The permeability between the PVS
+      and the SAS $\xi_{\rm PVS-SAS}$ must likely be estimated by a
+      bit of trial-error from the observations of timings of PVS-SAS
+      and SAS from~\cite{eide2024functional}.} \discuss{The
+      extracranial solute efflux permeability $\beta_{\rm efflux}$ is
+      uniformly distributed over the outer SAS boundary with a
+      reasonable rate as baseline\cite{hornkjol2022csf,
+        eide2021clinical, ringstad2024glymphatic}.} At baseline, the
+    effective diffusion coefficients $D^{\rm x}_{\rm SAS} = D^{\rm
+      x}_{\rm PVS}$ of the SAS and PVS represent the free diffusion
+    coefficient of Gadubutrol (${\rm x} = {\rm gad}$) in CSF (water),
+    and $D^{\rm x}_{\rm PAR}$ the effective diffusion coefficient of
+    human cortical tissue, all at body temperature; weighted by a
+    dispersion factor $\alpha$ (see Methods). At baseline, we consider
+    net flow due to CSF-production in the SAS and PVS:
+    $\mathbf{u}_{\rm SAS} > 0$ and $\hat{u} > 0$ (see Methods) but no
+    bulk flow in the parenchyma $\mathbf{u}_{\rm PAR} = 0$. Due to the
+    timescale for human intracranial tracer transport (minutes to
+    hours) compared to typical (cardiac, respiratory, slow vasomotion)
+    human intracranial pulsatility (0.1 -- 1 Hz), we consider net
+    (constant-in-time) flow fields only in subsequent model variations
+    (but see dispersion factor above).  Model B represents
+    perivascular pathway scenarios with more rapid flow in the PVS
+    induced by vasomotion (PVS I)~\cite{gjerde2023directional}.
+    Models C represents a version of the aforementioned scenarios but
+    for studying metabolite clearance (instead of solute influx).}
+\label{tab:scenarios}
+\end{table}
 
@@
 \subsubsection{Steady flow in perivascular networks induced by pressure differences}
+\label{sec:app:hydraulic_network}
 
@@
 \end{figure}
-
-\newpage
-\section{Notes}
-
-Various notes associated with \cite{eide2024functional}:
-\begin{itemize}
-  \item
-    Can observations of tracer indicate compartmentalization without
-    the existence of physical barriers but in the presence of
-    e.g.~substantial flow? We made observations in this direction in
-    \cite{vinje2021brain}, relating to the then-active
-    shape-of-perivascular-spaces discussion. We can address this by
-    trying to quantify and evaluate the permeability of the PVS-SAS
-    interface - comparing model predictions with the timings and
-    patterns presented by \cite{eide2024functional}.
-  \item
-    Do the tracer primarily move from the PVS into the SAS or along
-    (more slowly) along the SAS? This is left as an open question by
-    \cite{eide2024functional} (page 3)
-  \item
-    How do altered ICP affect transport? Note that PVSAS transport is
-    impaired (delayed) with increasing mean wave amplitude in the ICP
-    signal and for iNPH patients (with enlarged PVSAS spaces).
-  \item
-    Very useful reference for PVS sizes (10--20 mm$^2$ in reference
-    patients, 15-70 mm$^2$ in iNPH), and average propagation speeds.
-\end{itemize}

### 2024-12-20 paper/supplementary_main.tex
Tidy up supplementary PVS peristalsis.

@@
     \midrule
-    A & Baseline with sheaths & $R_2 = 2 R_1$ & $\xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
+    A & Baseline & $R_2 = 2 R_1$ & $\xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
     B1 & Sheath permeability & $R_2 = 2 R_1$ & $0, \xi, 10 \xi, 100 \xi, 1000 \xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
@@
     \end{center}
-  \caption{Overview of models (-- denotes as immediately above). Each
+  \caption{Overview of models. (-- denotes as immediately above). Each
     PVS cross-section is modelled as a concentric annulus with inner
@@
 
-\mer{MER: This needs a pass to update to new notation and new style.}
-
-We use the theoretical framework introduced by Gjerde et al.~\cite{gjerde2023directional} to compute an analytic estimate of the time-averaged (or \emph{net}) flow rates $\langle Q'_i \rangle$ (mm$^3$/s) induced by peristaltic pumping in a perivascular network$\Lambda = \cup_i \Lambda_i$.
+We use the theoretical framework previously introduced by Gjerde et al.~\cite{gjerde2023directional} to compute an analytic estimate of the time-averaged (or \emph{net}) flow rates $\langle Q'_i \rangle$ (mm$^3$/s) induced by peristaltic pumping in a perivascular network $\Lambda = \cup_i \Lambda_i$.
 The motion of the (inner) vascular wall is described by a periodic traveling (peristaltic) wave of relative amplitude $\varepsilon$, wave length $\lambda$ (mm) and frequency $f$, acting normal to the wall.
 By definition, the wave number is $k = 2 \pi/\lambda$ and the angular frequency is $\omega = 2 \pi f$.
-Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L$, baseline inner radius $r_{o_i}'$, fixed outer radius $r_{e_i}'$, and outer-to-inner ratio $\beta_i = r_{e_i}'/r_{o_i}'$ (see also \Cref{tab:pvs:parameters}).
+Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L$, baseline inner radius $R_1^i$, fixed outer radius $R_2^i$, and outer-to-inner ratio $\beta_i = R_2^i/R_1^i$.
 These geometric parameters and the assumption of annular cylindrical PVS segments yield hydraulic resistances $\mathcal{R}_{o, i}$ and additional characteristic parameters, see \cite{gjerde2023directional} for the complete definitions and schematics.
-Since the analytical estimate is derived under the assumption that $k L \approx \mathcal{O} 1)$~\cite{gjerde2023directional} and has been verified against numerical simulations for $k L$ of the order $10^{-1}-10^2$~\cite[Table I]{gjerde2023directional}, we consider it applicable for the wave lengths and vascular network data considered here in which $k L$ ranges from \mer{$X$ to $Y$.}
-
+Since the analytical estimate is derived under the assumption that $k L \approx \mathcal{O} (1)$~\cite{gjerde2023directional} and has been verified against numerical simulations for $k L$ of the order $10^{-1}-10^2$~\cite[Table I]{gjerde2023directional}, we consider it applicable for the wave lengths and vascular network data considered here in which $k L$ range from 0.15 to 30 for the strong vasomotion and 0.0015 to 0.30 for the cardiac waves (\Cref{tab:pvs:parameters}).
 \begin{table}
@@
     \midrule
-    Relative PVS size $\beta$ & $\beta = \beta_i = r_{e_i} / r_{o_i}$ & 2 & $\approx$\cite{eide2024functional} \\
+    Relative PVS size $\beta$ & $\beta = \beta_i = R_2^i / R_1^i$ & 2 & $\approx$\cite{eide2024functional} \\
     Wave frequency $f$ & Traveling wave frequency of vascular motion & $0.1-1.0$ Hz & \discuss{\cite{gjerde2023directional}} \\
-    Wave length $\lambda$ & Traveling wave length of vascular motion & $20-2000$ mm & \discuss{\cite{broggini2024long, gjerde2023directional}} \\
+    Wave length $\lambda$ & Traveling wave length of vascular motion & $0.02-2$ mm & \discuss{\cite{broggini2024long, gjerde2023directional}} \\
     Wave amplitude & Relative amplitude of inner wall motion & $1-10\%$ & \discuss{\cite{gjerde2023directional}} \\
-    Vascular radii $\{r_{o_i}\}$ & Vascular network radii data (avg $\pm$ std, $\min$, $\max$) & $X$, $X$, $X$  mm & \cite{hodneland2019new} \\
-    Vascular branch lengths $\{L_{i}\}$ & Path length between vasc. junctions (avg $\pm$ std, $\min$, $\max$) & $X \pm Y$, $X$, $X$  mm & \discuss{--} \\
+    \midrule
+    Cardiac frequency $f_c$ & Frequency of human cardiac pulse wave  & $1.0$ Hz & \cite{jung2021novel} \\
+    Cardiac wave length $\lambda_c$ & Wave length of human cardiac pulse wave & $2.0$m & \\
+    Cardiac amplitude $\varepsilon_c$ & Wall displacement amplitude of cardiac pulse wave & 1\%  & \\
+    \midrule
+    Vasomotion frequency $f_v$ & Frequency of slow vasomotion wave  & $0.1$ Hz & \cite{jung2021novel} \\
+    Vasomotion wave length $\lambda_v$ & Wave length of vasomotion wave & $0.02$ m & \cite{broggini2024long, gjerde2023directional} \\
+    Vasomotion amplitude $\varepsilon_v$ & Wall displacement amplitude of vasomotion wave & 10\%  & \cite{broggini2024long, gjerde2023directional}\\
+    %Vascular radii $\{r_{o_i}\}$ & Vascular network radii data (avg $\pm$ std, $\min$, $\max$) & $X$, $X$, $X$  mm & \cite{hodneland2019new} \\
+    %Vascular branch lengths $\{L_{i}\}$ & Path length between vasc. junctions (avg $\pm$ std, $\min$, $\max$) & $X \pm Y$, $X$, $X$  mm & \discuss{--} \\
     \bottomrule
   \end{tabular}
-  \caption{Perivascular flow induced by vascular wall motion: parameters and key data characteristics.}
+  \caption{Perivascular flow induced by vascular wall motion: overview of parameters.}
   \label{tab:pvs:parameters}
@@
 \input{tikz/fig_pvs_flow}
+
 This theoretical formalism~\cite{gjerde2023directional} is defined
@@
 edge-disjoint subnetworks $\Lambda^1, \Lambda^2, \Lambda^3$, one for
-each of the supply nodes (\Cref{fig:app:peristalsis}).
-Each node is assigned to the subnetwork associated with the nearest supply node, and edges between nodes are preserved.
-Next, we compute a minimal, bifurcating and directed tree representation of each subnetwork: $\mathcal{T}^1, \mathcal{T}^2, \mathcal{T}^3$.
-Each tree $\mathcal{T}^j = \cup E_n^j$ consists of the subset of the nodes from $\Lambda^j$ that have degree 1 (are leaf or root nodes) or degree $3$ (are true bifurcation points), and each path between nodes with degree $\not = 2$ in $\Lambda^j$ is represented in $\mathcal{T}^j$ by an edge $E^j$ with edge length $L$ corresponding to the total length of the original path and edge radius $r_o$ as the average of the path radii.
-For the few (2) nodes in $\Lambda$ with degree $>3$; i.e.~junctions where a branch splits into more than two downstream branches, only the two longest downstream paths are included in $\mathcal{T}^j$.
+each of the supply nodes (\Cref{fig:results1}A, C,
+\Cref{fig:app:peristalsis}). Each node is assigned to the subnetwork
+associated with the nearest supply node, and edges between nodes are
+preserved. Next, we compute a minimal, bifurcating and directed tree
+representation of each subnetwork: $\mathcal{T}^1, \mathcal{T}^2,
+\mathcal{T}^3$. Each tree $\mathcal{T}^j = \cup E_n^j$ consists of the
+subset of the nodes from $\Lambda^j$ that have degree 1 (are leaf or
+root nodes) or degree $3$ (are true bifurcation points), and each path
+between nodes with degree $\not = 2$ in $\Lambda^j$ is represented in
+$\mathcal{T}^j$ by an edge $E^j$ with edge length $L$ corresponding to
+the total length of the original path and edge radius $R^j$ as the
+average of the path radii. \discuss{For the few (2) nodes in $\Lambda$ with
+degree $>3$; i.e.~junctions where a branch splits into more than two
+downstream branches, only the two longest downstream paths are
+included in $\mathcal{T}^j$.}
 
@@
 which we set a flow rate of zero). Finally, we define the mean
-longitudinal perivascular velocity $\langle u^{\rm PVS}_i \rangle$ by
-dividing the flow rate $\langle Q'_i \rangle$ by the cross-section
-area $A_i = \pi (r_{e_i}^2 - r_{o_i}^2) = \pi (\beta_i^2 - 1) r_{o_i}^2$
+longitudinal perivascular velocity induced by the peristaltic wave
+$\langle \hat{u}^x_i \rangle$ by dividing the flow rate $\langle Q'_i
+\rangle$ by the cross-section area $A_i = \pi ({R_2^i}^2 - {R_1^i}^2)
+= \pi (\beta_i^2 - 1) (R_1^i)^2$.
 

### 2024-12-20 paper/supplementary_main.tex
Remove obsolete comment.
Update Table 2 and put it back in with Methods.

@@
 These sections provide more detailed descriptions of the mathematical models and numerical approximations considered.
-
-\subsection*{Overview of in-silico scenarios and experiments considered}
-
-See~\Cref{tab:scenarios} for an overview of the different in-silico
-scenarios and computational models considered.
-\begin{table}
-  \begin{center}
-  \begin{tabular}{ll|llllll|ll}
-    \toprule
-    & Description & $R_1, R_2$ & $\xi_{\rm PVS-SAS}$ &  $\beta_{\rm efflux}$ & $D$ & $\hat{u}$ & $\mathbf{u}_{SAS}$ & $c_0$ \\
-    \midrule
-    A & Baseline & $R_2 = 2 R_1$ & $\xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
-    B1 & Sheath permeability & $R_2 = 2 R_1$ & $0, \xi, 10 \xi, 100 \xi, 1000 \xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
-    B2 & Health permeability & $R_2 = 2 R_1$ & $0, \xi, 10 \xi, 100 \xi, 1000 \xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} +  \hat{u}_{\rm cardiac} + \hat{u}_{\rm vaso}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
-    %C & Larger PVS & $R_2 = N R_1$ & -- & -- & --  &  --  & -- &  -- \\
-    C & Higher PVS flow & -- & $\xi$ & -- & --  & $\hat{u}_{\rm prod} +  \hat{u}_{\rm cardiac} + \hat{u}_{\rm vaso}$  & -- & --  \\
-    AA & PVS size & $R_2 = x \beta R_1$ & $\xi$ & -- & -- & -- &  -- & 0 \\ 
-    %C & Larger PVS & $R_2 = N R_1$ & -- & -- & --  &  --  & -- &  -- \\
-    B & PVS flow & -- & $\xi$ & -- & --  & $\hat{u}_{\rm prod} + \hat{u}_{\rm vaso}$  & -- & --  \\
-    C & Clearance & -- & -- & -- & $\alpha D^{\rm gad, \tau, A\beta}$  & $\hat{u}_{\rm prod}$ & -- & 1 \\
-    %E & PVS II & $R_2 = 2 R_1$ & $\xi$, -- & -- & --  &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow\uparrow$  & -- & --  \\
-    %F & PVS III & $R_2 = 2 R_1$ & $\xi$, -- & -- & $10 D_{\rm PVS}$ &  $\hat{u} = \hat{u}_{\rm prod}$  & -- & --  \\
-    %G & Glymphatics & $R_2 = 2 R_1$ & $\xi$, -- & -- & \cite{sykova2008diffusion, valnes2020apparent} &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow$  & $\mathbf{u}_{\rm PAR}$ > 0 & --  \\
-    \bottomrule
-    \end{tabular}
-    \end{center}
-  \caption{Overview of models. (-- denotes as immediately above). Each
-    PVS cross-section is modelled as a concentric annulus with inner
-    radius $R_1$ and outer radius $R_2$ and such that $R_2 = \beta
-    R_1$ with $\beta = 2.0$ based on~\cite{eide2024functional}. Let
-    $\xi = \dots$ denote the permeability estimated by Koch et
-    al.~\cite{koch2023estimates}. Model A and its variations aims to
-    quantify the effect of PVS sheaths by varying the permeability
-    $\xi$ between the PVS and its surroundings. We consider a fixed
-    permeability $\xi_{\rm PVS-PAR}$ based
-    on~\cite{koch2023estimates}. \mer{The permeability between the PVS
-      and the SAS $\xi_{\rm PVS-SAS}$ must likely be estimated by a
-      bit of trial-error from the observations of timings of PVS-SAS
-      and SAS from~\cite{eide2024functional}.} \discuss{The
-      extracranial solute efflux permeability $\beta_{\rm efflux}$ is
-      uniformly distributed over the outer SAS boundary with a
-      reasonable rate as baseline\cite{hornkjol2022csf,
-        eide2021clinical, ringstad2024glymphatic}.} At baseline, the
-    effective diffusion coefficients $D^{\rm x}_{\rm SAS} = D^{\rm
-      x}_{\rm PVS}$ of the SAS and PVS represent the free diffusion
-    coefficient of Gadubutrol (${\rm x} = {\rm gad}$) in CSF (water),
-    and $D^{\rm x}_{\rm PAR}$ the effective diffusion coefficient of
-    human cortical tissue, all at body temperature; weighted by a
-    dispersion factor $\alpha$ (see Methods). At baseline, we consider
-    net flow due to CSF-production in the SAS and PVS:
-    $\mathbf{u}_{\rm SAS} > 0$ and $\hat{u} > 0$ (see Methods) but no
-    bulk flow in the parenchyma $\mathbf{u}_{\rm PAR} = 0$. Due to the
-    timescale for human intracranial tracer transport (minutes to
-    hours) compared to typical (cardiac, respiratory, slow vasomotion)
-    human intracranial pulsatility (0.1 -- 1 Hz), we consider net
-    (constant-in-time) flow fields only in subsequent model variations
-    (but see dispersion factor above).  Model B represents
-    perivascular pathway scenarios with more rapid flow in the PVS
-    induced by vasomotion (PVS I)~\cite{gjerde2023directional}.
-    Models C represents a version of the aforementioned scenarios but
-    for studying metabolite clearance (instead of solute influx).}
-\label{tab:scenarios}
-\end{table}
 
@@
 the total length of the original path and edge radius $R^j$ as the
-average of the path radii. \discuss{For the few (2) nodes in $\Lambda$ with
-degree $>3$; i.e.~junctions where a branch splits into more than two
-downstream branches, only the two longest downstream paths are
-included in $\mathcal{T}^j$.}
+average of the path radii.
 
@@
 \rangle$ by the cross-section area $A_i = \pi ({R_2^i}^2 - {R_1^i}^2)
-= \pi (\beta_i^2 - 1) (R_1^i)^2$.
+= \pi (\beta_i^2 - 1) {R_1^i}^2$.
 

### 2024-12-22 scripts/compare_PVS_timing.py
make figure 5

@@ -39,16 +39,24 @@ def plot_timings(models:List[str]):
         }
     
     namedict = {"modelA": "baseline", "modelA-strongVM":"high PVS flow",
-                "modelA-PVS-disp": "high PVS dispersion",}
+                "modelA-PVS-disp": "high PVS dispersion",
+                "LargePVS":"high PVS flow (dilated)", "LargePVSA":"baseline (dilated)"}
     markerdict = {"modelA": "d", "modelA-strongVM":"s",
-                "modelA-PVS-disp": "o",}
+                "modelA-PVS-disp": "o", "LargePVS":"o", "LargePVSA":"*"}
     print(models)
-    coldict = {"modelA-strongVM":"#6610f2","modelA":"#0b774d", "modelA-PVS-disp":"#f7b801"}
-
+    coldict = {"modelA-strongVM":"#6610f2","modelA":"#0b774d", "modelA-PVS-disp":"#f7b801",
+               "LargePVS":"#f7b801", "LargePVSA":"darkred"}
+    annotate_model = "LargePVSA" if "LargePVSA" in models else "modelA"
     for t, (pvsdf, outerdf) in zip(["fta", "peaktime"] , [(ftas, outerftas), (pts, otps)]):
-        fig, axes = plt.subplots(ncols=len(art_groups), figsize=(9,4))
-
-        for i, (ax, (agn, ag))  in enumerate(zip(axes, art_groups.items())):
+        fig, axes = plt.subplots(ncols=len(art_groups) + 1, figsize=(9,4),
+                                  width_ratios=[0]+[3]*len(art_groups), sharey=True)
+        ax = axes[0]
+        if t=="fta":ax.set_ylabel("first-time arrival (h)"); ax.set_ylim((0.0, 6))
+        elif t=="peaktime":ax.set_ylabel("time-of-peak (h)"); ax.set_ylim((2, 9.1))
+        ax.get_xaxis().set_ticks([])
+        ax.spines['bottom'].set_visible(False)
+        #ax.set_xlim((-0.05, 0.1))
+        for i, (ax, (agn, ag))  in enumerate(zip(axes[1:], art_groups.items())):
             for m in models:
                 print(m)
                 group = pvsdf.loc[ag]
@@ -65,20 +73,13 @@ def plot_timings(models:List[str]):
                 #ax.legend(loc='upper center', bbox_to_anchor=(0.5, 1.17), ncol=2, 
                 #        columnspacing=0.3, frameon=False)
[…]

### 2024-12-22 scripts/compare_mean_concentrations.py

@@ -1,0 +1,79 @@
+import typer
+import matplotlib.pyplot as plt
+from plotting_utils import set_plotting_defaults, read_config
+import numpy as np
+import seaborn as sns
+import os
+from typing import List
+
+mol2mmol = 1e3
+
+def compare_concentrations(models:List[str]):
+
+    config = read_config(f"configfiles/{models[0]}.yml")
+    v = '_'.join(models)
+    os.makedirs(f"plots/comparisons/{v}", exist_ok=True)
+
+    dt , T= config["dt"], config["T"]
+    times = np.arange(0, T + dt, 3*dt*config["output_frequency"])
+    labels = ["CSF", "parenchyma", "PVS artery", "PVS vein"]
+    results = dict()
+    for m in models:
+        print(f"reading model {m}")
+        data = read_config(f"results/{m}/mean_concentrations.yml")
+        tot_values = {l:[data[f"ctot_{l}_{t}"] for t in times] for l in labels}
+        concentrations = {l:[data[f"cmean_{l}_{t}"] for t in times] for l in labels}
+        results[m] = dict(conc=concentrations, tot=tot_values)
+
+    set_plotting_defaults()
+    #sns.set_palette("BrBG", n_colors=4)
+    #sns.set_palette(["#0a9396","#94d2bd", "#e9d8a6", "#ee9b00"])
+    colors = ["#0a9396","#81b8a5", "#ee6700", "#eea700"]
+    coldict = {"modelA-strongVM":"#6610f2","modelA":"#0b774d", "modelA-PVS-disp":"#f7b801",
+               "LargePVS":"#f7b801", "LargePVSA":"darkred"}
+    namedict = {"modelA": "baseline", "modelA-strongVM":"high PVS flow",
+                "LargePVS":"high PVS flow (dilated)", "LargePVSA":"baseline (dilated)"}
+    for t in ["tot", "conc"]:
+        fig, axes = plt.subplots(2,2,figsize=(9,4),sharex=True, sharey=True)
+        for i, ax, l in zip(range(len(labels)), axes.flatten(), labels):
+            for m in models:
+                ax.plot(times/3600, results[m][t][l], label=namedict[m] if i==0 else None, color=coldict[m])
+            ax.set_title(l)
+        plt.figlegend(loc='upper center', bbox_to_anchor=(0.5, 1.05), ncol=len(models)*2,
+                     columnspacing=0.3, frameon=False)
[…]

### 2024-12-22 scripts/compare_pvs_flow.py

@@ -1,0 +1,58 @@
+from plotting_utils import read_config, set_plotting_defaults
+import typer
+from typing import List
+import os
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+m2mum = 1e6
+
+def compare_pvs_flow(flow_model:List[str]):
+
+    v = "_".join([f.replace("/","+") for f in flow_model])
+    os.makedirs(f"plots/comparisons/{v}/", exist_ok=True)
+
+    results = dict()
+    for m in flow_model:
+        print(f"reading {m}")
+        md = dict()
+        md.update(read_config(f"plots/{m}/pvs_metrics.yml"))
+        results[m] = md
+    print(results)
+    df = pd.DataFrame(results).transpose() * m2mum
+    df["modelstr"] = df.index.astype("str")
+    df["enlarged"] = "control"
+    df["enlarged"][df.modelstr.str.contains("large") + df.modelstr.str.contains("3")] = "dilated"
+    df["type"] = "none"
+    df["type"][df.modelstr.str.contains("vasomotion") ] = "vasomotion"
+    df["type"][df.modelstr.str.contains("cardiac") ] = "cardiac"
+    df["type"][df.modelstr.str.contains("sas") ] = "pressure-driven"
+
+    df["vm"] = df.modelstr.str.contains("vasomotion") 
+    df["cardiac"] = df.modelstr.str.contains("cardiac") 
+    df["pressure"] = df.modelstr.str.contains("sas") 
+
+    set_plotting_defaults()
+    sns.set_palette(["#2ec4b6", "#124559"])
+    for var, ylabel in zip(["pvsumax", "pvsumean"],
+                           ["max PVS velocity (μm/s)", "mean PVS velocity (μm/s)"]):
+        plt.figure(figsize=(4.0,2.3))
+        ax = sns.barplot(df, y=var, x="type", hue="enlarged", order=["pressure-driven", "cardiac", "vasomotion"])
+        for p in ax.patches:
+            h, w, x = p.get_height(), p.get_width(), p.get_x()
+            if h>0:
+                xy = (x + w / 2., h)
+                text = f'{h:0.2f}'
+                plt.annotate(text=text, xy=xy, ha='center', va='bottom')
+        plt.ylabel(ylabel)
+        plt.xlabel("")
+        ax.legend_.set_title("")
[…]

### 2024-12-22 scripts/detail_plot.py

@@ -69,12 +69,14 @@ def generate_plot(
         cmax = 0
     times = np.array(times)*3600
     pointdict = {l:p for l,p in pointlabels}
+    config = read_config(f"configfiles/{modelname}.yml")
     if selected_artlabels[0]=="all":
         artlabels = list(pointdict.keys())
     else:
         artlabels = selected_artlabels
     points = np.array([pointdict[l] for l in artlabels])
     fontsize = 16
+
 
     fig = plt.figure(figsize=(len(times)*5, len(artlabels)*3), frameon=True)
     grid = ImageGrid(fig, 111, nrows_ncols=(len(artlabels), len(times)),
@@ -90,8 +92,8 @@ def generate_plot(
                          non_manifold_traversal=False)
     sas = get_result(modelname, "sas", times)
     art = get_result(modelname, "artery", times)
+    pvs_radius_ratio = config["pvs_ratio_artery"]
 
-    pvs_radius_ratio = 2
     cmap = cmocean.cm.algae_r
     for j, (p, l) in enumerate(zip(points, artlabels)):
         slice, pvs, vessel, pvs_surf, pia, pia_inner, n = get_slice(art, sas, par, p, pvs_radius_ratio=pvs_radius_ratio)

### 2024-12-22 scripts/evaluate_pvs_flow.py

@@ -1,9 +1,6 @@
 from dolfin import *
-import xii
-from solver import read_vtk_network, as_P0_function, get_mesh
 import numpy as np
 import os
-from pvs_flow import pvs_flow_system
 from test_map_on_global_coords_shift import map_kdtree, map_dg_on_global
 from subdomain_ids import CSFID, V34ID, PARID, LVID, CSFNOFLOWID
 import typer
@@ -17,11 +14,38 @@ import pandas as pd
 import yaml
 from branch_marking import color_branches
 from plotting_utils import minmax
-import ufl_legacy as ufl
 from cmap import Colormap 
+from vtk_adapter import create_vtk_structures
+from extract_vessels import get_tubes
+import ufl_legacy as ufl
+
 m2mm = 1e3
 m2mum = 1e6
 Pa2mPa = 1e3
+
+def plot_pvs_velocity(mesh, uh_mag, artery_radii, filename):
+    CG1 = FunctionSpace(mesh, "CG", 1)
+    bar_args=dict(title="velocity (μm/s)", vertical=False,
+                    height=0.07, width=0.6, position_x=0.2,
+                    position_y=0.0, title_font_size=28, font_family="arial",
+                    label_font_size=28, fmt="%.0f")
+    topology, cell_types, x = create_vtk_structures(CG1)
+    grid = pv.UnstructuredGrid(topology, cell_types, x)
+    grid["u"] = uh_mag.vector()[:] *m2mum
+    umax = abs(grid["u"]).max() 
+    grid["radius"] = artery_radii.vector()[:]
+    tubes = get_tubes(grid.ctp())
+    pl = pv.Plotter()
+    pl = pv.Plotter(off_screen=True, window_size=(1200, 800))
+    pl.add_mesh(tubes,scalars="u", cmap="curl", clim=(-umax, umax),
+                scalar_bar_args=bar_args)
+    pl.camera_position = 'xz'
+    #pl.camera.roll += 5
+    pl.camera.azimuth += 165
+    pl.camera.elevation += 10
+    pl.camera.zoom(1.85)
+    pl.camera.focal_point = np.array(pl.camera.focal_point) + np.array([0,0,-0.008])
+    img = pl.screenshot(filename, transparent_background=True)
 
 def sig_to_space(sig, mesh):
     elem = eval(sig.replace("Cell", "ufl.Cell"))
@@ -38,13 +62,18 @@ def compute_pvs_flow(pvs_flow_file):
[…]

### 2024-12-22 scripts/plot_conc_at_label.py

@@ -26,7 +26,7 @@ def plot_conc_at_label(model:str):
     #from IPython import embed; embed()
     artlabels = sum(artlabelsets, [])
 
-    coldict = {"modelA-strongVM":"#6610f2","modelA":"#0b774d"}
+    coldict = {"modelA-strongVM":"#6610f2","modelA":"#0b774d", "LargePVS":"green"}
     col = coldict[model]
     pvspeaktimedict = {n:results[f"{n}_pvs_peak_time"] for n in artlabels}
     pvsftadict = {n:results[f"{n}_fta"] for n in artlabels}

### 2024-12-22 scripts/pvs_flow_prod.py

@@ -110,6 +110,8 @@ def compute_pvs_flow(csf_flow_model:  str, network: str, radius_ratio:str):
         f.write(uh_mag, "umag")
         f.write(ph, "p")
         f.write(radius_f, "radii")
+        radius_f.vector()[:] *= radius_ratio
+        f.write(radius_f, "pvs_radii")
 
     with XDMFFile(f'{results_dir}/pvs_flow_vis.xdmf') as xdmf:
         xdmf.write(pvs_flow_vec, t=0)

### 2024-12-22 scripts/peristalticflow.py
make figure 5

@@ -531,7 +531,10 @@ def compute_pvs_flow(meshfile, output, args):
         f.write(mesh, "mesh")
         f.write(u_directed, "u")
         f.write(u, "umag")
+
         f.write(mf_to_dg(radii), "radii")
+        radii.array()[:] *= beta
+        f.write(mf_to_dg(radii), "pvs_radii")
         f.write(df.Function(u.function_space()), "p")
             
 def run_all_tests():

### 2024-12-23 scripts/mark_and_refine_mesh.py
resurect mesh convergence

@@ -127,7 +127,7 @@ def mark_and_refine(configfile : str):
     # of the CSF space from the Stokes computation (disconnected domains)
 
     sas = df.Mesh()
-    with df.XDMFFile(f'mesh/{meshname}/volmesh/mesh.xdmf') as f:
+    with df.XDMFFile(config["basemesh"]) as f:
         f.read(sas)
         gdim = sas.geometric_dimension()
         label = df.MeshFunction('size_t', sas, gdim, 0)
@@ -137,11 +137,15 @@ def mark_and_refine(configfile : str):
         # compute distance to interface for later refinement
         parenchyma_surf = pv.read(f"mesh/{meshname}/surfaces/parenchyma_incl_ventr.ply")
 
+        # refine all
+        AQ_coords = (0.0837, 0.08, 0.065) 
+        sas, label = refine_sphere(sas, AQ_coords, 1, label, 
+                                    criterion=None, min_size=config["edge_length_r"] * 0.6) 
+
         # refine V3 and V4
         sas, label = refine_region(sas, label, labelids=[V34ID])
 
         # refine AQ
-        #AQ_coords = (0.0837, 0.08, 0.065) 
         #sas, label = refine_sphere(sas, AQ_coords, 0.007, label)
 
         # refine Par boundary

### 2025-01-05 paper/supplementary_main.tex
Updates from Overleaf

@@
 main text for explicit parameter values while giving the general,
-abstract form here, and to the reference~\cite{masri2023modelling} for
+abstract form here, and to the reference~\cite{masri2024modelling} for
 the derivation and analysis of this 3D-1D model. Recall that we model
@@
     \centering
+    \includegraphics[width=0.9\linewidth]{paper/figures/mesh_refinement.png}
+    \caption{Illustration of the three different meshes: low resolution, standard resolution, high resolution (from left to right).}
+    \label{fig:mesh_refinement}
+\end{figure}
+
+\begin{figure}
+    \centering
     \begin{subfigure}[b]{0.45\textwidth}

### 2025-01-06 paper/supplementary_main.tex
fix figure paths

@@
 We thus conclude that the standard resolution mesh and a time step of $dt=120\,$s offers sufficient accuracy for our simulations.
+In addition, we compute the standard model without allowing tracer transport across the models outer boundary, confirming mass conservation of our numerical scheme (\cref{fig:mass_conservation}).
 
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{paper/figures/mesh_refinement.png}
+    \includegraphics[width=0.9\linewidth]{figures/mesh_refinement.png}
     \caption{Illustration of the three different meshes: low resolution, standard resolution, high resolution (from left to right).}
@@
         \centering
-        \includegraphics[width = 1 \linewidth]{figures/mesh_refinement_par_csf_mean.png}
-        \caption{Mean tracer concentration in CSF and parenchyma under mesh refinement}
+        \includegraphics[width = 1 \linewidth]{figures/csf_minmax.png}
+        \caption{Minimum and maximum tracer concentration in  the CSF under mesh refinement}
     \end{subfigure}
@@
         \centering
-     \includegraphics[width= 1 \linewidth]{figures/mesh_refinement_art_ven_mean.png}
-         \caption{Mean tracer concentration in PVSs around (veins and) arteries  under mesh refinement}
+     \includegraphics[width= 1 \linewidth]{figures/par_minmax.png}
+        \caption{Minimum and maximum tracer concentration in the parenchyma under mesh refinement}
     \end{subfigure}
-    \caption{Mean tracer concentrations over the first 24\,h after injection on the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the LowRes, standard and HighRes meshes with a timestep of $dt=120$\,s.}
+    \begin{subfigure}[b]{0.45\textwidth}
+        \centering
+        \includegraphics[width = 1 \linewidth]{figures/art_minmax.png}
+        \caption{Minimum and maximum tracer concentration in PVSs around arteries under mesh refinement}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.45\textwidth}
+        \centering
+     \includegraphics[width= 1 \linewidth]{figures/ven_minmax.png}
+         \caption{Minimum and maximum tracer concentration in PVSs around veins under mesh refinement}
+    \end{subfigure}
+    \caption{Minumum and maximum tracer concentrations over the first 24\,h after injection on the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the LowRes, standard and HighRes meshes with a timestep of $dt=120$\,s.}
     \label{fig:mesh_convergence_concentrations}
@@
         \centering
-        \includegraphics[trim={0.5cm 1cm 0.05cm 0.8cm}, clip, width = 1.05 \linewidth]{figures/mesh_refinement.pdf}
+        \includegraphics[trim={0.5cm 1cm 0.05cm 0.8cm}, clip, width = 1.05 \linewidth]{figures/modelALowRes_modelA_modelAHighRes.pdf}
         \caption{Convergence of key quantities of interest with mesh refinement}
@@
         \centering
-     \includegraphics[trim={0.5cm 1cm 0.05cm 0.8cm}, clip,width= 1.05 \linewidth]{figures/time_refinement.pdf}
+     \includegraphics[trim={0.5cm 1cm 0.05cm 0.8cm}, clip,width= 1.05 \linewidth]{figures/modelAsmalldt_modelA_modelAlargedt.pdf}
         \caption{Convergence of key quantities of interest with time refinement}
@@
     \end{subfigure}
-    \caption{minimal ($\rm h_{min}$) and maximal ($\rm h_{max}$) mesh cell sizes (computed as cell circumradius $\times 2$) (A); number of points and tetrahedral cells in each mesh (B); mean dispersion enhancement factor $R$ (C), maximum dispersion enhancement factor $R$ (D); maximum pressure in steady CSF production flow (E), maximum pressure in cardiac-driven CSF flow (F); maximum CSF velocity in steady CSF production flow (G); maximum CSF velocity in cardiac-driven CSF flow (H);  mean tracer concentration in the CSF, parenchyma, arterial PVS and venous PVS at 3,6,12 and 24\,h (I, J, K and L).}
+    \caption{minimal ($\rm h_{min}$) and maximal ($\rm h_{max}$) mesh cell sizes (computed as cell circumradius $\times 2$) (A); number of points and tetrahedral cells in each mesh (B); mean cardiac dispersion enhancement factor $R$ (C), maximum cardiac dispersion enhancement factor $R$ (D); maximum pressure in steady CSF production flow (E), maximum pressure in cardiac-driven CSF flow (F); maximum CSF velocity in steady CSF production flow (G); maximum CSF velocity in cardiac-driven CSF flow (H);  mean tracer concentration in the CSF, parenchyma, arterial PVS and venous PVS at 3,6,12 and 24\,h (I, J, K and L).}
 \end{figure}
@@
     \end{subfigure}
-    \caption{Mean tracer concentrations over the first 24\,h after injection on the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the standard resolution mesh for timesteps $dt \in \{60, 120, 240 \}$\,s.}    \label{fig:time_convergence_concentrations}
+    \caption{Mean tracer concentrations over the first 24\,h after injection on the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the standard resolution mesh for timesteps $dt \in \{60, 120, 240 \}$\,s.}
+    \label{fig:time_convergence_concentrations}
 \end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\linewidth]{figures/modelAMassConservation_total_conc.png}
+    \caption{total tracer content in CSF, parenchyma and arterial and venous PVS for the standard model without tracer outflow, illustrating numerical mass conservation.}
+    \label{fig:mass_conservation}
+\end{figure}

### 2025-01-13 paper/supplementary_main.tex
Updates from Overleaf

@@
 In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
-between cells $E^1$ and $E^2$ with facet normal vector $\bm n_F$
-pointing from $E^1$ to $E^2$, $h_F$ denotes the facet diameter,
+between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
+pointing from $E^1$ to $E^2$. The facet diamter is denoted by $h_F$,
 the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
@@
 We assess the accuracy and numerical convergence of our simulation results by performing a series of experiments with different spatial and temporal resolutions.
-Specifically, we generate a sequence of three meshes with an increasing number of points and cells, solve all relevant simulation steps (CSF flow and intracranial transport computations) for standard model A, and compare the results with respect to a set of key quantities of interest.
+Specifically, we generate a sequence of three meshes, see \Cref{fig:mesh_refinement}, with an increasing number of points and cells, solve all relevant simulation steps (CSF flow and intracranial transport computations) for standard model A, and compare the results with respect to a set of key quantities of interest.
+The mesh refinement employed here is adapative in the sense that the coarser mesh is refined near expected sharp concentration gradients.
+This allows us to better capture dynamics and reduce undershoots, see \Cref{fig:mesh_convergence_concentrations}.
 Similarly, we investigate the effect of the time step size by solving the intracranial transport model with varying time steps $dt \in \{60, 120, 240 \}$\,s.

### 2025-01-23 paper/main.tex
Shuffle parts of the Discussion.

--- 2024-11-28 paper/main.tex
@@
 \newcommand{\mar}[1]{\textcolor{violet}{#1}}
+\newcommand{\mk}[1]{\textcolor{orange}{#1}}
 \newcommand{\discuss}[1]{\textcolor{red}{#1}}
 \newcommand{\fixme}[1]{\textcolor{red}{#1}}
-\newcommand{\draft}[1]{\textcolor{gray}{#1}}
+\newcommand{\draft}[1]{\textcolor{lightgray}{#1}}
 %\newcommand{\brain}{\Omega_{\rm{PAR}}} 
@@
 
+%% The problem we are addressing here is that we would like to be able to
+%% understand, integrate and quantify the role of different mechanisms -
+%% including effects associated with individual variations - on molecular
+%% enrichment and clearance of human intracranial spaces. And we think of
+%% this as a step on the way to predict tracer enrichment after
+%% intrathecal injection in a personalized medicine context. Which is
+%% important because of drug delivery/treatment in cancer, diagnostics
+%% for CSF and dementia disorders, and clearance in neurodegenerative
+%% disease.
+
 \begin{abstract}
-``The abstract — which should be no more than 200 words long and contain
-no references — should serve both as a general introduction to the
-topic and as a brief, non-technical summary of the main results and
-their implications.'' 
+  The mechanisms of intracranial solute transport are fundamental to
+  human brain health and function, with alterations often linked to
+  disease and functional impairment, and with distinct opportunities
+  for personalized diagnostics and treatment. However, our
+  understanding of these mechanisms and their interplay remains
+  incomplete, in part due to the complexity of integrating insights
+  across scales, between species and from different modalities. Here,
+  we combine geometric model reduction and mixed-dimensional
+  modelling, multi-modal magnetic resonance images, and high
+  performance computing to construct and explore a high-fidelity
+  in-silico model of human intracranial molecular enrichment and
+  clearance. This model predicts the temporal and spatial spreading of
+  a solute within an image-derived geometric representation of the
+  subarachnoid space, ventricular system and brain parenchyma,
+  including branching networks of surface perivascular spaces
+  (PVSs). Our findings highlight the significant impact of
+  cerebrospinal fluid (CSF) production and intracranial pulsatility on
+  molecular enrichment following intrathecal tracer injection. We
+  demonstrate that low-frequency vasomotion induces moderate CSF flow
+  rates in surface PVS networks that substantially enhance tracer
+  enrichment, and that impaired enrichment is a natural direct
+  consequence of enlarged PVSs. This openly available technology
+  platform thus provides an opportunity for integrating separate
+  observations on diffusion in neuropil, vascular dynamics,
+  intracranial pulsatility, CSF production, and efflux, and for
+  exploring questions relating to drug delivery and clearance in the
+  human brain environment.
 \end{abstract}
@@
 
-\mer{MER: Target length: 5000 words (Intro, Results, Discussion) plus <3000 words Methods: 700 words Introduction. 2800 words Results. 1500 words Discussion.}
+%\mer{MER: Target length: 5000 words (Intro, Results, Discussion) plus <3000 words Methods: 700 words Introduction. 3100 words Results. 1200 words Discussion.}
+
+\iffalse
+\paragraph*{Abbreviations}
+\begin{itemize}
+\item
+  Subarachnoid space (SAS)
+\item
+  Cerebrospinal fluid (CSF)
+\item
+  Perivascular space (PVS), perivascular spaces (PVSs)
+\end{itemize}
+\fi
 
@@
 
-\section*{Introduction (rough ideas)}
+% What is the big picture problem and why is it important? 
+The mechanisms underlying molecular transport within the intracranial
+space are fundamental to human brain health and
+function. Neurodegenerative diseases such as Alzheimer's, Parkinson's,
+and Huntington's disease are all associated with abnormal accumulation
+of protein aggregates together with alterations in transport and
+clearance characteristics~\cite{rasmussen2018glymphatic,
+  harrison2020impaired, eide2023plasma, liu2024glymphatic}. Moreover,
+sleep and conversely sleep-deprivation play a definite yet enigmatic
+role in modulating molecular enrichment and
+clearance~\cite{xie2013sleep, eide2021sleep, eide2022altered,
+  miao2024brain, hauglund2025norepinephrine}. In the last decade,
+established theories have been challenged by new findings revealing a
+greater degree of molecular movement and
+exchange~\cite{iliff2012paravascular, ringstad2017glymphatic,
+  louveau2017understanding, proulx2021cerebrospinal,
+  bohr2022glymphatic}, as well as substantial variability in
+enrichment and clearance between individuals and between patient
+cohorts~\cite{ringstad2018brain, eide2021direction, eide2021impaired,
+  eide2022altered}. These observations provide distinct opportunities
+(and challenges) for personalized medicine e.g.~for tailored
+intrathecal delivery of chemotherapy~\cite{lohela2022glymphatic} and
+for early diagnostics of impaired brain
+clearance~\cite{eide2021clinical, van2024human}. In spite of their
+importance, our understanding of these mechanisms is incomplete with
+open challenges and significant debate -- in part relating to the
+translation of knowledge between scales, species, experimental
+protocols, clinical cohorts, and individuals.
 
-% What is the problem? (Big picture)
-% - Big picture/importance sentence.
-% - More in-depth info
-% - "XXX is unknown".
-
-Transport of molecules within the spaces of the central nervous system
-(CSF spaces, parenchyma) is important. Both in the context of
-intracranial drug delivery with potential for personalized treatment
-of cancer, diagnostics and treatment of CSF-related disorders, and in
-the context of understanding clearance in connection with
-neurodegenerative disease and for future treatment and diagnostics of
-dementia disorders. Understanding and predicting molecular enrichment
-is important for treatment and diagnostics of (brain) cancer, for
-diagnostics of CSF-related disorders like normal pressure
-hydrocephalus, subarachnoid hemorrage. Moreover, altered transport and
-clearance is believed to be key factor in neurodegenerative diseases
-like Alzheimer's disease and Parkinson's disease, and in sleep. CSF
-flow and PVS established to play a key role. Clinical studies have
-revealed large individual variation and variation between
-cohorts/disease groups, therefore important in particular in the
-context of personalized medicine. However, we are far from
-understanding (the mechanisms behind) these variations.
-
-% What we already know about the problem
-Our current understanding has primarily been established via
-experimental medical research (cats, rats, mice, pigs). Big open
-questions and controversy remain, lack holistic understanding of these
-physiological process remains, hard to translate experimental studies
-to humans. In the last decade, clinical studies. Clinical studies have
-revealed large individual variation and variation between
-cohorts/disease groups. However, we are far from understanding the
-mechanisms behind these variations and integrating and translating
-experimental observations to humans.  In-silico modelling is starting
-to play an integral role, evaluating hypothesis and pinpointing
-mechanisms, but most efforts have been focused on local aspects and
-lacking a holistic model. Hard to quantify the overall contribution of
-individual effects.
-
-PVS flow is important as shown by experimental studies. This is being
-corroborated by clinical studies in humans (Eide and Ringstad (2024),
-and the new one). However, a number of open questions remain: to what
-extent do the PVS constitute an ensheathed space, how fast is the flow
-in human PVS, what is the effect of dilated PVS as seen in disease
-cases. Also, open questions on transport characteristics of the CSF
-spaces - what gives the individual variations in tracer enhancements?
-
-What is known is that fluid motion (in the ventricular system and SAS,
-in perivascular spaces, and possibly within the parenchyma) and
-pulsatility play key
-roles. Glymphatics~\cite{iliff2012paravascular}. Diffusion, PVS flow
-and dispersion, CSF flow. Molecular transport in perivascular spaces
-(PVSs) is established as a key pathway for human brain clearance and
-delivery. Previous studies indicate that molecules move rapidly in the
-subarachnoid space (SAS) and in PVSs surrounding pial arteries. CSF
-production in the choroid plexus and possibly in the parenchyma. CSF
-efflux pathways, meningeal lympahtics, nerve routes. See also to
-Cserr, Phys Rev., 1971, bulk convective flow may be more important for
-larger species (larger distances) and larger molecules. Contribution
-both from net flow (convection) and pulsatility (mixing) both in the
-CSF spaces and in PVSs. Some aspects are better established, such as;
-while other factors are more disputed such as ... contribution of
-vasomotion. Moreover, the role of local effects in the bigger picture
-has remained hard to evaluate.
+Perivascular pathways along the brain surface and within the brain
+parenchyma have long been hypothesized to serve a designated role in
+this context~\cite{rennels1985evidence, zhang1990interrelationships,
+  ichimura1991distribution, carare2008solutes, iliff2012paravascular,
+  foley2012realtime, hannocks2018molecular, van2024caa}. Recently, Eide and
+Ringstad~\cite{eide2024functional} and Yamamoto et
+al~\cite{yamamoto2024perivascular} demonstrated that perivascular
+spaces (PVSs) define preferential pathways for molecular transport in
+humans, with delayed periarterial enrichment in dementia
+subtypes~\cite{eide2024functional}. Perivascular flow of cerebrospinal
+fluid (CSF) clearly contributes to this transport, and is inherently
+associated with vascular pulsations~\cite{hadaczek2006perivascular,
+  iliff2013cerebral, bedussi2018paravascular, mestre2018flow,
+  boster2023artificial, hirschler2024region}. Fluid mechanics
+considerations point at intracranial pressure differences and shorter
+wavelength vascular wall pulsations as drivers of directional net flow
+and convection in the PVS~\cite{bilston2003arterial, rey2018pulsatile,
+  daversin2020mechanisms, kedarasetti2020functional,
+  gjerde2023directional, nozaleda2024arterial}, while longer waves
+such as the pulse wave primarily contribute to oscillatory flow and
+dispersion~\cite{asgari2016glymphatic, sharp2019dispersion, thomas2019fluid,
+  kedarasetti2020arterial, troyetsky2021dispersion, 
+  martinac2021phase}. In the bigger picture, CSF is produced by the
+choroid plexus~\cite{damkier2013cerebrospinal,
+  steffensen2018cotransporter, liu2020direct}, pulsates through the
+ventricular system, cisterns, and subarachnoid space (SAS) in
+synchrony with cardiac, respiratory, and neural
+waves~\cite{greitz1993pulsatile, wagshul2011pulsating,
+  sweetman2011cerebrospinal, fultz2019coupled, vinje2019respiratory,
+  eide2021direction, causemann2022human, williams2023neural,
+  zimmermann2023total}, and drains via the dural sinuses, meningeal
+lymphatics, cranial nerves, or other efflux
+pathways~\cite{proulx2021cerebrospinal}. However, how these
+physiological factors and physical mechanisms integrate to enhance or
+impair human intracranial molecular transport over larger spatial
+scales and longer time scales remain unknown. A related key question
+is to what extent surface PVSs are separated from the SAS by
+structural barriers~\cite{zhang1990interrelationships,
+  weller2005microscopic, bedussi2017paravascular,
+  pizzo2018intrathecal, mestre2022periarteriolar,
+  mollgard2023mesothelium, smets2024perivascular, eide2024functional},
+and in turn to what extent such structural compartmentalization is a
+prerequisite for effective perivascular transport.
 
 % What are our research questions and how do we answer them
-Recent developments on 3D-1D modelling of transport in PVS and
-tissue. ``Digital twin for molecular enrichment and clearance in human
-intracranial spaces''. Can we model the interplay between
-molecular transport in the intracranial compartments, CSF flow and PVS
-flow and transport to holistically quantify the contributions in humans?
+In this study, by leveraging geometric model reduction and
+mixed-dimensional modelling~\cite{masri2024modelling}, structural
+magnetic resonance (MR) images~\cite{hodneland2019new}, and high
+performance computing, we introduce an integrated computational model
+of intracranial molecular enrichment and clearance. Focusing on the
+interplay between perivascular pathways, pulsatility and CSF flow
+dynamics, the model predicts the temporal evolution and spatial
+distribution of a solute concentration within a detailed geometric
+representation of the human SAS and ventricular system, networks of
+surface PVSs and the brain parenchyma. In terms of transport dynamics,
+we account for heterogeneous diffusion, dispersive mixing induced by
+cardiac and respiratory pulsatility in the CSF spaces and PVSs,
+convective fluid flow driven by CSF production and peristaltic
+pumping, as well as solute exchange and clearance across
+semi-permeable membranes. Comparing with glymphatic MRI
+studies~\cite{ringstad2017glymphatic, ringstad2018brain,
+  watts2019measuring}, the in-silico predictions accurately represent
+molecular enrichment patterns, timing and intercompartmental
+distributions. This open platform~\cite{ZENODO} thus provides a
+technological opportunity for qualitatively and quantitatively
+exploring key open questions relating to molecular movement within the
+human brain environment such as the role of pulsatility, perivascular
+pathways, structural compartmentalization, or morphological
+pathologies.
 
 % Main findings and outlook
-Here, we aim to model and study this transport in a full pial
-perivascular and/or vascular network embedded in the SAS.
+By exploring this high-dimensional parameter space, we propose that
+the balance between CSF production and intracranial pulsatility is key
+to shaping the large-scale features of intracranial enrichment
+patterns, with the potential to span a wide range of individual and
+cohort variability. Moreover, we predict that CSF production, cardiac-
+and respiratory pulsatility is not sufficient to explain early
+perivascular enrichment, but that fluid flow induced by low-frequency
+vasomotion in surface periarterial spaces (on the order of 13 \textmu
+m/s) is sufficient, even in the absence of a structural
+compartmentalization of the PVS. Conversely, enlarged PVSs in the SAS
+will cause a substantial reduction in cardiac- and vasomotion-driven
+flow velocities, strongly delay perivascular transport, and thus
+impair intracranial enrichment. These findings transfer, reconcile,
+integrate and extend insights from clinical, experimental, and
+theoretical studies, and lay a new foundation for in-silico studies of
+personalized intrathecal drug delivery and brain clearance.
 
-\iffalse
-\newpage
+%% \begin{itemize}
+%% \item
+%%   Need better understanding of transport of solutes in intracranial spaces.
+%% \item
+%%   Need to integrate separate observations e.g. on diffusion in
+%%   neuropil, PVS flow, net flow, effect of dispersion
+%% \item
+%%  Need to transfer and integrate insights from animal models,
+%%   computational models and theoretical insights into the human case. 
+%% \item
+%%  Controversies need to be resolved.
+%% \end{itemize}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Results}
-\mer{Update: ~2800 words and 7 subsections - around 300-450 words each?}
-\begin{enumerate}
-\item
-  ``A digital twin'' for molecular enrichment and clearance in human intracranial spaces
-\item
-  CSF flow, the role of production versus no/cardiac/respiratory
-  dispersive effects on tracer trajectories, flow bifurcation. With and without tentorium cerebellum.
-\item
-  ``The role'' of PVS flow speed
-\item
-  ``The role'' of PVS sheaths and permeability - does functional separation entail structural separation?
-\item
-  ``The role'' of PVS size (motivated by pathologies)
-\item
-  Clearance from the CSF spaces - with and without low pressure veins, but keeping the intrathecal source of solute. 
-\item
-  Clearance from the parenchyma: $c_0 = 1.0$. Different substances?
-\end{enumerate}
+%% \mer{MER: Target ~3100 words in 6 subsections}
 
+%% \mer{MER: Results #1 is about 500 words}
+\subsection*{In-silico predictions of intracranial molecular enrichment and clearance after intrathecal injection}
 \begin{figure}
-     \begin{subfigure}[b]{0.45\textwidth}
     \centering
-         \includegraphics[width=\textwidth]{figures/Brain-PVS-callouts.png}
-     \end{subfigure}
-     \hfill
-     \caption{a) Illustration of the different intracranial domains
-       considered with labels;}
-     \label{fig:intracranial_domains}
+    \includegraphics[width=0.95\textwidth]{figures/figure1.png}
+     \caption{A) Illustration of the spatial domains of the model: CSF-filled space (ventricles and SAS) $\Omega_{\rm CSF}$, parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and the veins $\Lambda_{\rm vein}$ and their interfaces and boundaries (see Methods); B) 3D rendering of the actual geometry used in this study parenchyma, CSF (clipped, blue), arterial network (red), venous network (blue)); C) arterial network (red) and venous network (blue); D) illustration of the modelled clinical tracer injection procedure; E) time of simulated first tracer arrival after intrathecal injection (concentration exceeds $0.1\,$mol/l); F) time of first tracer arrival in the arterial PVS (top) and venous PVS (bottom); G) sagittal, coronal and axial view of computed tracer concentrations over T1-image 1, 6, 12 and 24\,h after injection (low concentrations transparent, pial surface in pink, arachnoid membrane in cyan, and arteries in dark red); H) total amount of tracer in each cranial compartment over the first 24\,h; I) average tracer concentration in each compartment over the first 24\,h.
+}
+     \label{fig:results1}
 \end{figure}
+Using previously published multi-modal magnetic resonance imaging
+(MRI) data~\cite{hodneland2019new, deistung2009tof,
+  schweser2012quantitative, reichenbach2012future,
+  deistung2017overview}, we construct a multiscale computational
+representation of the human intracranial compartments consisting of
+the CSF spaces and brain parenchyma as three-dimensional (3D) domains
+and with the PVSs surrounding major surface arteries and veins as
+embedded networks of topologically one-dimensional (1D) curves
+(\Cref{fig:results1}A--C). We consider a solute concentration field,
+varying in space and time, in the 3D domains and in the PVS networks,
+and assume that the solute can cross between these compartments
+through semi-permeable membranes. As the drivers and modes of
+intracranial transport are under substantial
+debate\cite{smith2019going, proulx2021cerebrospinal,
+  bohr2022glymphatic, hladky2022glymphatic, betsholtz2024advances},
+our first target is to establish a baseline model accounting for a
+reasonably conservative set of mechanisms and their integrated effect
+over a timescale of several minutes to a few days. To this end, we
+assume that the solute will (i) diffuse within all compartments, with
+diffusivity depending on the effective properties of the relevant
+medium\cite{sykova2008diffusion}; (ii) experience significant
+dispersive effects due to the pulsatile flow of CSF induced by the
+cardiac and respiratory cycles\cite{vinje2019respiratory,
+  sharp2019dispersion, ray2021quantitative, troyetsky2021dispersion};
+and (iii) be convected by a (small) net flow of CSF resulting from
+production in the choroid plexus with CSF efflux across the upper
+convexity\cite{hornkjol2022csf} and from the peristaltic pumping
+effect of pulse wave pulsations in surface periarterial
+spaces~\cite{mestre2018flow, gjerde2023directional}
+(\Cref{fig:results1}D). Mathematically, this model is represented by a
+mixed-dimensional system of coupled time-dependent partial
+differential equations\cite{masri2024modelling}, which we solve
+numerically with high accuracy using a mass-conserving finite element
+scheme and the FEniCS finite element software\cite{alnaes2015fenics,
+  kuchta2020assembly} (see Methods). The computational framework and
+associated software are all openly available~\cite{ZENODO}.
 
-\subsection*{High-fidelity in-silico predictions of intracranial solute transport and exchange}
+Simulating a glymphatic MRI protocol~\cite{ringstad2017glymphatic,
+  ringstad2018brain, eide2024functional}, we then predict the
+spreading of $0.5$ mmol intrathecally injected Gadobutrol after its
+appearance at the craniocervical junction
+(\Cref{fig:results1}E--G). After one hour, the in-silico
+tracer moves upwards in the SAS frontally of the brainstem, and
+quickly reaches the supratentorial regions. Here, it spreads both
+posterior through the quadrigeminal cistern and the longitudinal
+fissure, and anteriorly through the outer SAS, reaching the top of the
+cerebral cortex after around 12 hours. After 24 hours, the tracer
+covers most of the brain surface (with the exception of some posterior
+regions) and has penetrated several centimeters into the parenchymal
+tissue. This pattern is reflected in the mean tracer concentrations in
+each compartment, where the CSF space reaches its peak of
+$1.4\,$mmol/l after 2 hours, followed by the arterial PVS
+concentration peak ($2.9\,$mmol/l after 4 hours)
+(\Cref{fig:results1}I). While the mean concentrations in these
+compartments drop soon after peaking, the parenchymal tissue
+slowly enriches with tracer over the first 24 hours, with a final
+value of $0.3$ mmol/l. Overall, about 40\% of the total amount of
+tracer remains in the cranium 24 hours post-injection, with the largest
+share in the CSF (53\%), followed by the parenchyma (35\%), the
+venous PVS (8\%) and the arterial PVS
+(5\%) (\Cref{fig:results1}H).
 
-\begin{itemize}
-\item
-  Present mathematical and computational model and geometries (\Cref{tab:scenarios}). 
-\item
-  Highlight computational model availability, open source code.
-\item
-  Provides an interactive in-silico platform for studying intracranial solute transport -- distribute such that anyone can run (possibly try to recruit MinRK if advantageous).
-\item
-  \mer{MER: Remember to report on how much concentration enters into the brain?}
-\end{itemize}
-
-\subsection*{CSF flow and dispersion shifts enrichment patterns}
-
-Flow of CSF through the SAS and the ventricular system is a crucial driver of transport in the human brain environment.
-While cardiac- and respiratory induced oscillatory flow increases mixing, the production of CSF in the choroid plexus creates a steady advective flow within the CSF-filled spaces of the cranium \cite{hornkjol2022csf}.
-We construct a 3D model of the SAS and the ventricular system (\Cref{fig:csf}A,B and C) and compute CSF flow fields for three different scenarios: CSF production in the choroid plexus, CSF flow at systolic peak blood inflow and CSF flow in response to free breathing.
-
-% production flow simulation
-To simulate the production-driven flow, we impose an inflow of $400\,$ml/day \cite{nilsson1992circadian} across the surface of the lateral ventricle and allow a resistive outflow across the upper convexity of the arachnoid mater, leading to a total pressure drop of $26\,$mPa between the production and outflow sites (\Cref{fig:csf}D).
-A maximum flow velocity of $1.8\,$mm/s is attained in the aqueduct, while the mean velocity in the SAS is $2.6\,\mu$m/s (\Cref{fig:csf} E).
-
-%The Stokes system is solved with the finite element software $FEniCS$, stored and then subsequently read for all the relevant solute transport simulations.
-See Figure~\ref{fig:csf} for a streamline visualization of the obtained velocity profile and \Cref{sec:details_numerical_method} for details on the numerical algorithm employed.
-
-% cardiac pulsatility
-Assessing the effect of cardiac-driven dispersion, we compute a steady CSF flow field at peak systolic blood inflow.
-We mimic the resulting reduction of CSF space volume by imposing an inflow of $6\,$ml/s \cite{baledent2014imaging, causemann2022human} and $0.31\,$ml/s \cite{vinje2019respiratory} across the upper arachnoid mater ($\Gamma_{\rm AM-U}$) and the lateral ventricle ($\Gamma_{\rm VL}$) surface, respectively.
-We allow free outflow into the spinal SAS ($\Gamma_{\rm SSAS}$).
-In addition, we upscale the obtained pressure field to account for the lack of pulsatile inertial effects in our model (see \ref{sec:dispersion_details}).
-In this scenario, we observe a pressure drop of $10\,$Pa between lateral ventricles and the spinal SAS and a maximum flow velocity of $19.8\,$cm/s in the aqueduct (\Cref{fig:csf}F and J).
-Next, we estimate a spatially varying dispersion enhancement factor $R$ from cardiac-induced pulsatility by combining the computed pressure field with the theory of shear-augmented dispersion (see \Cref{sec:dispersion_details}).
-Assuming a cardiac frequency of $1\,$Hz, find that cardiac-induced pulsatile CSF flow increases the effective diffusion by up to two orders of magnitude in the aqueduct and around the cisterna magna, but has little effect in most of the SAS ($R < 1$) (\Cref{fig:csf} G and K).
-
-% respiratory pulsatility
-Employing the same methodology, we estimate the respiratory contribution to dispersive mixing by computing the CSF flow field driven by an inflow of $0.121\,$ml/s \cite{liu2024using} and $1\,$ml/s \cite{gutierrez2022effect} across the lateral ventricular wall and the arachnoid mater, respectively.
-This amounts to a respiratory peak flow volume of $1.121\,$ ml/s at the craniocervical junction \cite{gutierrez2022effect}.
-While the resulting flow velocities are only about one fourth of their cardiac-induced counterparts (see \Cref{fig:csf}H and L), respiratory dispersive mixing reaches a factor of up to 350 due the lower respiratory frequency ($0.25\,$Hz) (\Cref{fig:csf}I and M).
-
-% tracer trajectory of the standard model
-Next, we use the computed advective velcity field and dispersion enhancement factors to predict the spreading of $0.5\,$mmol intrathecally injected Gadobutrol in CSF-filled spaces, the parenchymal tissue and PVSs around arteries and veines.
-One hour post-injection, tracer moves upwards in the SAS frontally of the brainstem,  and quickly reaches supratentorial regions (\Cref{fig:csf}N).
-Here, it spreads both posteriorly through the longitudinal fissure and anteriorly through the outer SAS, reaching the top of the cerebral cortex after around 12 h.
-After 24h, the tracer covers most of the brains surface (with exception of some posterior regions) and has penetrated several centimeters into the parenchymal tissue.
-
-% tracer trajectory of model variations
-
-\begin{itemize}
-    \item only dispersion model: No production flow.
-Tracer remains in the subtentorial regions, in cerebellum, brain stem and surrounding CSF.
-In contrast to other models, it travels up the ventricular system - similar to what is observed in iNPH patients?
-    \item Low dispersion model: account for uncertainty (or pathologies with impaired pulsatile motion?) - tracer exclusively spreads in the frontal half of the brain.
-Bifurcation of CSF production flow field: flow bifurcates around the midbrain in two directions: a) through the quadrigeminal cistern into the longitudinal fissure and b) frontally around the brain through the outer SAS.
-Only models with "enough" dispersion enable tracer to follow trajectory a).
-    \item without tentorium: 
-\end{itemize}
-
-\begin{figure}[h!]
+% \mer{MER: Results #2 is about 600 words}
+\subsection*{Reduced CSF pulsatility strongly shifts intracranial enrichment patterns}
+\begin{figure}%[h!]
 \centering 
-\includegraphics[width = 0.9\linewidth]{figures/figure2.png}
-\caption{
-A) Illustration of 3D-domains; 
-B) Illustration of relevant boundaries; 
-C) schematic illustration of the model, its doamins and boundaries; 
-D) CSF flow induced by CSF production; 
-E) production-induced CSF velocities in the SAS and third and fourth ventricle (V3 \& V4); 
-F) CSF flow induced by peak systolic expansion; 
-G) cardiac dispersion enhancement factor;
-H) CSF flow induced by respiration; 
-I) respiratory dispersion enhancement factor;
-J) cardiac-induced flow velocities;
-K) cardiac-induced dispersion enhancement factor;
-L) respiratory flow velocities;
-M) respiratory dispersion enhancement factor;
-N) sagittal, coronal and axial planes of tracer concentration after 1, 6, 12 and 24$\,$h for model A. The outer model boundaries is shown in cyan, the pial membrane in purple and the arteries in red.;
-O) model X,Y and Z after 12h;
+\includegraphics[width = 0.95\textwidth]{figures/figure2.png}
+\caption{\textbf{The balance between CSF pulsatility and production shapes molecular enrichment patterns.}
+A) CSF spaces: the third and fourth ventricle (V3\&4), lateral ventricles (LV) and SAS; 
+B) Boundaries and interfaces: the spinal subarachnoid space (SSAS), upper and lower arachnoid membrane (AM-U and AM-L), pia (Pia) and lateral ventricles (LV); 
+C) Schematic of the CSF flow model; 
+D) CSF flow induced by CSF production with outflow allowed across the
+upper convexity (AM-U); 
 }
@@
 \end{figure}
+\begin{figure}
+\ContinuedFloat
+\caption{(\textbf{cont.})
+E) Histograms of the production-induced CSF velocities in the SAS and third and fourth ventricle (V3 \& V4) in terms of the relative frequency of the mean flow speed in each computational cell weighted by its volume;
+F) CSF flow induced at peak systolic blood inflow; 
+G) Cardiac-induced dispersion factor $R_c$;
+H) CSF flow induced at peak respiratory expansion; 
+I) Respiration-induced dispersion factor $R_r$, $D = (1 + R_c + R_r) D^{\rm Gad}$; J--M): Histograms of the cardiac-induced flow velocities (J),
+cardiac-induced dispersion factors (K),
+respiratory flow velocities (L), respiratory dispersion factor (M);
+N--Q): sagittal, coronal and axial planes of tracer concentration after 12h for different model variations: baseline (N), no CSF production (O), low dispersion (P), and high dispersion (Q).
+}
+\end{figure}
 
-\subsection*{The significance of perivascular spaces as flow highways}
-\begin{itemize}
-\item
-  Perivascular flow and transport has been identified as a key
-  mechanism and potential target for enhancing brain drug delivery and
-  metabolic waste clearance -- what are the global (intracranial)
-  effects of such intracranial highways? 
-\item
-  Do solutes move from PVS to the SAS, or both along the PVS and along
-  the SAS?
-\end{itemize}
+The enrichment patterns observed clinically after intrathecal
+injection of contrast differ substantially between subjects and
+between pathological conditions~\cite{ringstad2018brain,
+  eide2021direction, eide2021impaired, eide2022altered}. These
+neurological conditions are also associated with alterations in the
+pulsatile flow of CSF in the ventricular system and
+SAS~\cite{eide2021direction}. Clearly, key physiological factors such
+as cardiac pulsatility and respiration easily differ between
+individuals and between both pathological and physiological states. We
+therefore next asked to what extent -- and how -- variations in CSF
+pulsatility would affect the intracranial enrichment characteristics.
 
-\paragraph{PVS Flow velocities}
-\begin{itemize}
-    \item we consider three different PVS flow fields: PVS flow driven by production field pressure differences, cardiac peristaltic flow and vasomotion peristaltic flow. 
-\end{itemize}
+For the CSF dynamics in our baseline model, we combine the net
+contribution to CSF flow induced by CSF production with the integrated
+dispersive effects of cardiac and respiratory pulsatile CSF flow
+(\Cref{fig:csf}A--C). Imposing a constant CSF production of
+$400\,$ml/day across the surface of the lateral ventricles, while
+allowing for efflux across the upper convexity, yields a total CSF
+pressure drop of $26\,$mPa (0.00020 mmHg) (\Cref{fig:csf}D) with a
+maximum flow velocity of $1.85\,$mm/s (in the aqueduct) and a mean
+velocity in the SAS of $2.6$\textmu m/s (\Cref{fig:csf}E). To examine the
+dispersive effects of pulsatile CSF flow, we employ the theory of
+shear-augmented dispersion together with computational fluid dynamics
+to determine a dispersion factor $R$ enhancing the effective solute
+diffusivity (see Methods). For the cardiac contribution, we compute
+CSF pressure and flow fields at peak systolic blood inflow,
+corresponding to a reduction in CSF space volume at a total rate of
+$6\,$ml/s~\cite{baledent2014imaging, causemann2022human} in the SAS
+and $0.31\,$ml/s across the lateral ventricle
+surface\cite{vinje2019respiratory}. This scenario sets up a pressure
+drop of $10\,$Pa ($0.075$ mmHg) between the lateral ventricles and the
+spinal SAS and a maximum flow velocity of $19.8\,$cm/s
+(\Cref{fig:csf}F, J). Assuming a cardiac frequency of $1\,$Hz, we
+infer that this cardiac-induced pulsatile CSF flow increases the
+effective diffusion by more than two orders of magnitude in the
+aqueduct and near the cisterna magna, but has little effect ($R <
+1$) in most of the SAS (\Cref{fig:csf}G, K). For the respiratory
+contribution, we employ the same methodology, but with a total rate of
+$1\,$ml/s \cite{gutierrez2022effect} in the SAS and $0.121\,$ml/s
+\cite{liu2024using} in the lateral ventricles, yielding a respiratory
+peak flow volume of $1.121\,$ ml/s at the craniocervical
+junction. While the resulting flow velocities are only about one
+fourth of their cardiac-induced counterparts (see \Cref{fig:csf}H, L),
+respiratory dispersive mixing reaches a factor of up to 320 due the
+lower respiratory frequency of $0.25\,$Hz (\Cref{fig:csf}I, M).
 
+Now, to examine how CSF pulsatility affects the enrichment patterns,
+we consider three variations of the baseline: (i) no CSF production,
+(ii) reduced pulsatility and thus decreased dispersion ($0.1 \times
+R$), and (iii) higher pulsatility with increased dispersion ($10
+\times R$). Without CSF production, transport is considerably delayed
+(\Cref{fig:csf}N, O). Even after $12$ hours, tracer remains in the
+subtentorial regions around the cerebellum, the brain stem and in the
+surrounding CSF. Interestingly, the lack of CSF production instead
+allows tracer to travel upwards through the ventricular system,
+reaching the third ventricle after around $12$ hours. On the other
+hand, if the CSF pulsatility is reduced, we observe rapid transport
+towards the upper convexity of the cranium, as in the baseline model,
+but the tracer spreads exclusively within the anterior regions
+(\Cref{fig:csf}P). This feature can be attributed to the CSF flow
+bifurcation posterior to the ambient cistern (\Cref{fig:csf}D):
+without sufficient diffusion, the tracer is unable to cross into the
+posterior SAS. Indeed, with higher dispersion, the tracer moves
+through the quadrigeminal cistern and further upwards into the
+longitudinal fissure, with also enrichment of the cerebellum
+(\Cref{fig:csf}Q).
+
+% \mer{MER: Results #3 is about 800 words}
+\subsection*{Perivascular flow shapes and accelerates molecular enrichment}
+\label{sec:pvs_flow_results}
+
+The PVSs are recognized across species as critical pathways for solute
+transport in and around the brain\cite{}, and thus as potential
+targets for enhancing brain drug delivery and metabolic waste
+clearance\cite{}. However, whether CSF flows more rapidly in PVSs and
+what the forces and mechanisms required to drive such flow are, remain
+as key points of debate~\cite{bohr2022glymphatic,
+  van2024caa}. Motivated by experimental observations in animal
+models~\cite{iliff2012paravascular, iliff2013cerebral, mestre2018flow,
+  bedussi2018paravascular}, there is now a remarkable body of
+literature on modelling perivascular fluid flow and
+transport~\cite{bilston2003arterial, asgari2016glymphatic,
+  rey2018pulsatile, daversin2020mechanisms, sharp2019dispersion,
+  thomas2019fluid, kedarasetti2020functional, kedarasetti2020arterial,
+  troyetsky2021dispersion, martinac2021phase, gjerde2023directional,
+  nozaleda2024arterial}. Here, we ask how these proposed mechanisms
+would translate from idealized geometries to human vascular networks
+and moreover evaluate their integrated effect in the context of
+intracranial solute transport.
+
+Our periarterial network extends from the internal carotid arteries
+and basilar artery through up to 18 bifurcations to reach upstream
+network ends located within the SAS or up to 6 mm inside the
+parenchyma (\Cref{fig:results1}A--C, \Cref{fig:pvs}A) and includes
+major surface arteries of radius 0.5--1.4 mm. Imposing the pressure
+field induced by CSF production at the periarterial network ends
+induces ultraslow steady CSF flow of variable direction in these PVSs,
+with an average velocity of 0.08 \textmu m/s (antegrade) and a maximum
+velocity of 0.55 \textmu m/s (\Cref{fig:pvs}B). On the other hand,
+traveling waves of arterial wall motion (\Cref{fig:pvs}C), such as the
+pulse wave or other vasomotion~\cite{vanveluw2020vasomotion,
+  munting2023spontaneous, bojarskaite2023sleep, broggini2024long,hauglund2025norepinephrine},
+also induce net directional flow in the
+PVS~\cite{kedarasetti2020arterial, kedarasetti2020functional,
+  coenen2021lubrication, gjerde2023directional, nozaleda2024arterial}
+-- of magnitude and direction depending on the amplitude, frequency,
+and length of the waves and the characteristics of the perivascular
+network~\cite{gjerde2023directional}. Applying a
+semi-analytic model of the net flow induced by peristalsis in
+perivascular networks~\cite{gjerde2023directional} (see Methods), we
+estimate that the cardiac pulse wave alone, traveling at a frequency
+of 1 Hz with a wavelength of 2.0 m and a 1\% wall
+displacement~\cite{jung2021novel}, will induce mainly antegrade PVS
+flow with an average net velocity of 0.92 \textmu m/s while reaching up
+to 7.31 \textmu m/s near larger, ventral vessels such as the MCA
+(\Cref{fig:pvs}D). The same theory predicts that strong ultraslow
+vasomotions, if traveling antegrade at 0.1Hz with a wavelength of 0.02
+m and a 10\% displacement~\cite{broggini2024long}, will induce both
+retrograde and antegrade net PVS flow with an average velocity of
+13.05 \textmu m/s and maximum velocity 54.44 \textmu m/s
+(\Cref{fig:pvs}E).
+
+We study the effect of such rapid PVS flow on intracranial molecular
+enrichment by comparing the more conservative baseline model and a
+high PVS flow model, where the former still includes the net flow
+contributions from CSF production and the cardiac pulse wave, while
+the latter additionally includes the ultraslow vasomotion
+contribution. In both models, tracers were first observed at the basal
+artery after 48 minutes (\Cref{fig:pvs}F), defined by their first-time
+arrival (FTA), the time at which the concentration first exceeds $0.1$
+mmol/l. At all upstream locations along the middle cerebral arteries
+(MCAs) and anterior cerebral artery (ACA), we found substantially
+reduced FTAs with higher PVS flow (\Cref{fig:pvs}G), up to 2.5 hours
+earlier in the left M2 segment of the MCA. PVS concentrations peaked
+before the surrounding tissue (0:48h vs 1:12h at the MCA and 1:00h vs
+1:48h at MCA2) with high PVS flow, while these peaks occurred nearly
+simultaneously in the baseline model. The earlier appearance and the
+delay in time-to-peak between the PVS and surrounding tissue clearly
+indicates the directionality of tracer enrichment -- it first arrives
+in the PVS and subsequently spreads into the tissue and SAS
+(\Cref{fig:pvs}H, I), and especially so with higher PVS flow. These
+observations also hold on an aggregated level: computing the total
+amount of tracer in the PVS as a function of the distance to the
+arterial network roots, we find accelerated tracer transport with
+higher PVS flow, especially at earlier time points (2--9 hours)
+(\Cref{fig:pvs}J). Moreover, differences in tracer transport along the
+PVS translate to altered enrichment patterns on the whole-organ
+scale. For instance, after 4-6 hours, the faster-moving tracer in the
+PVS is clearly visible in the adjacent spaces to the ACA, MCA, and
+other arteries (\Cref{fig:pvs}L). \mer{One more sentence on effect of
+  high PVS flow on tracer enrichment outside the PVS.} In contrast,
+there are no clear signs of early enrichment surrounding the PVS in
+the baseline model (\Cref{fig:pvs}K).
+
+A natural question is whether early PVS enrichment could be the result
+of increased dispersion rather than net flow in the
+PVS\cite{asgari2016glymphatic,sharp2019dispersion,bojarskaite2023sleep,asgari2016glymphatic,troyetsky2021dispersion}. Interestingly,
+even increasing the dispersion factor by $100 \times$ in the PVS,
+induced only minor changes in the global spreading rate compared to the
+baseline (\Cref{fig:pvs}G, M), thus indicating a negative answer to
+this question.
 \begin{figure}
     \centering
-    \includegraphics[height=\textheight]{figures/figure3.png}
-    \caption{\mer{MER: Put vasculature network illustration here.
-Many aspects actually.
-Marking of segments.
-Location relative to 3D.
-Radius etc.
-Marius will evaluate and suggest.
-Results: Group segments by logical order.
-For key arterial segments: average radius and length (can be moved to Supplementary if need be); pressure values; flow velocities; and then histograph with flow velocities for each mesh cell.
-Make sure to minorly note that ACA refers to a segment of the ACA and not the whole thing e.g.}}
+    \includegraphics[width=0.95\textwidth]{figures/figure3.png}
+    \caption{\textbf{Perivascular flow shapes and accelerates molecular enrichment.}}
     \label{fig:pvs}
 \end{figure}
-\Cref{fig:pvs}.
+\begin{figure}
+\ContinuedFloat
+  \caption{\textbf{(cont.)}
+    A) Surface arterial network colored by arterial radius with vessel locations labeled;
+    B) Net PVS flow induced by CSF production (cf.~\Cref{fig:csf}D) -- histograms show the relative frequencies of local periarterial velocities.
+Positive values indicate antegrade PVS flow, while negative values indicate retrograde PVS flow;
+    C) Illustration of the concept of peristaltic pumping;
+    D) Estimated net PVS flow induced by pulse wave peristaltic pumping;
+    E) Estimated net PVS flow induced by vasomotion/slow wave peristaltic pumping;
+    F) Comparison of mean tracer concentration in the PVS and over the outer PVS surface at key locations of the arterial tree, for the baseline (lower) and high PVS flow (upper) models, with labels: ``peak'': time-to-peak (h), ``$\Delta t$'': time difference between peaks (h), and ``FTA'': first-time tracer arrival (h).
+    G) first-time arrival for the main trunks of the periarterial network over the distance from the closest network root node (BA, ICA-R or ICA-L for the baseline, the high PVS flow and the high PVS dispersion models);
+    H, I) 2D slices showing zoom-in on the region surrounding the MCA-R and MCA2-R for the baseline (upper) and high PVS flow (lower) models.
+Innermost circle indicates the artery (blue), the surrounding annulus is the PVS (color represents concentration value), the surroundings are parenchyma and/or CSF spaces, separated by the pia (blue: towards the parenchyma, pink: towards the CSF));
+    J) Total amount of tracer in the periarterial spaces as a function of the distance to the nearest network root over time for the baseline (left) and the high PVS flow (right) models;
+    K--M) sagittal (upper) and coronal (lower) views of the simulated tracer concentrations overlayed on the T1-weighted MR image after 4 and 6h for the baseline (K), high PVS flow (L) and high PVS dispersion (M) models (concentrations \discuss{below threshold masked}, pial surface outlined in pink, arachnoid membrane in cyan, and arteries in dark red).}
+\end{figure}
 
+% \mer{MER: Results #4 is about 500 words}
 \subsection*{Structural versus functional compartmentalization of perivascular spaces}
 
-\begin{itemize}
-\item
-  Are perivascular spaces functionally or structurally
-compartmentalized? Human and rodent observations indicate that tracers
-concentrate in perivascular spaces surrounding the pial and
-subarachnoid vasculature. Must these spaces be structural i.e.~defined
-by semi-permeable structural
-membrane~\cite{zhang1990interrelationships, zhang1992directional,
-  mestre2018flow, eide2024functional}, or can such
-compartmentalization be a result of regions of enhanced flow or
-mixing~\cite{bedussi2017paravascular, vinje2021brain}? See also
-references in Background.
-\item
-  Vary the PVS permeability $\xi$.
-\end{itemize}
+Human and rodent observations indicate that tracers concentrate in perivascular spaces surrounding the pial and subarachnoid vasculature \cite{zhang1990interrelationships,zhang1992directional, bedussi2017paravascular, mestre2018flow, bedussi2018paravascular, eide2024functional}.
+However, it remains unclear whether such enrichment patterns necessitate a structural barrier, such as a membrane with limited permeability, or if the patterns could result from enhanced flow or mixing in these areas alone.
+We therefore next investigate how the permeability of the interface between the PVS and the surrounding CSF affects tracer enrichment around the major arterial trunks (\Cref{fig:compartmentalization}A--B).
+To this end, we compare models with high and low  permeabilities, representing a highly permeable PVS-CSF interface (functional compartmentalization) and a less permeable interface (structural compartmentalization), respectively.
+For the low permeability model, we set the permeability to $3.8 \cdot 10^{-7}\,$m/s, consistent with previous estimates for the endfoot sheath surrounding penetrating arterioles \cite{koch2023estimates}, which we consider to be a lower bound for the surface PVS-CSF permeability.
+For the high permeability model, we increase this permeability by a factor of 100.
+For both scenarios, we consider the high PVS flow regime as examined in the previous section.
 
-\subsection*{Enlarged perivascular spaces yield ...}
+At the basal artery (BA), tracer appears within one hour in the surrounding CSF in both models (\Cref{fig:compartmentalization}C, E).
+With a more permeable PVS-CSF interface, tracer quickly crosses from the CSF into the PVS, resulting in similar peak concentrations of 25\,mmol/l in the PVS and at its outer surface, while values up to 50\,mmol/l are attained in the vicinity.
+In contrast, the low permeability model exhibits substantially lower perivascular tracer enrichment, with a maximum concentration of 10\,mmol/l after approximately two hours.
+%The high permeability model allows for more tracer crossing into the PVS, leading to higher concentrations further downstream.
+Notably, crossover into the parenchyma is also observed, most clearly at 4 hours.
+For the middle cerebral arteries (MCA-R and MCA-L), tracer appears
+first in the PVS in both models (\Cref{fig:compartmentalization}D, E),
+though with higher concentrations in the high permeability model (up
+to 24 mmol/l). We also observe that the higher permeability allows the
+tracer to leak out of the PVS and spread within the Sylvian fissure to
+a greater extent. \discuss{With a small delay, more tracer appears via
+  the CSF pathway (\Cref{fig:compartmentalization}F, peak time
+  difference).} At the anterior cerebral artery (ACA-A2), the first tracer arrival in the PVS occurs after 3 hours, with a peak at 5 hours.
+The concentration peak in the PVS is tailed by a higher peak in the CSF (\Cref{fig:compartmentalization}E), indicating tracer arrival through a second pathway; i.e., directly through the CSF-filled space outside of the PVS.
+Similar patterns are observed for ACA-A3.
 
-\begin{itemize}
-  \item Take $R_2 = 3 R_1$ instead of $2 R_1$.
-\end{itemize}
+Summarizing and quantifying these observations
+(\Cref{fig:compartmentalization}F), we find that neither the time of
+first arrival nor the time-to-peak are substantially affected by the
+PVS-CSF permeability at any of the periarterial segments considered
+(BA, MCAs, ACAs). However, the concentration difference between the
+PVS and its surrounding CSF quickly decreases with higher
+permeability, as does the time lag between the concentration peaks in
+these domains. Thus, fast transport along the PVS is not contingent
+upon a structural compartmentalization of the PVS, whereas a sharp
+concentration gradient between PVS and CSF is unlikely without a
+restricting barrier.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.95\textwidth]{figures/figure4.png}
+    \caption{\textbf{Does early perivascular enrichment rely on a structural compartmentalization of the PVS?}
+    A) The PVS-CSF membrane permeability regulates the exchange of solutes between the PVS and surrounding CSF spaces; 
+    B) The arterial network with highlighted regions of interest; 
+    C) 2D slices showing zoom-in to the region surrounding the basal artery after 1, 2, 3, and 4 hours for a low (above) and high (below) PVS-CSF permeability model.
+Innermost circle indicates the artery (blue), the surrounding annulus is the PVS (color represents concentration value), the surroundings are parenchyma and/or CSF spaces, separated by the pia (blue: towards the parenchyma, pink: towards the CSF));
+    D) As in (C) for regions surrounding (from upper to lower) the ACA-A3, ACA-A2, MCA-R and MCA-L segments;}
+    \label{fig:compartmentalization}
+\end{figure}
+\begin{figure}
+\ContinuedFloat
+\caption{(\textbf{cont.})
+  E) Comparison of mean tracer concentration in the PVS (dashed) and over the
+outer PVS surface (dotted) for the BA, MCA-R, MCA-L, ACA-2, and ACA3 segments over time for the high (purple) and low permeability (yellow) models;
+  F) Effect of varying the PVS-CSF permeability on first-time-of-arrival (FTA), time difference ($\Delta t$) between concentration peaks in the PVS and surrounding tissue, time of PVS peak concentration, PVS peak concentration and difference in peak concentration between PVS and surrounding tissue.}
+\end{figure}
 
-\subsection*{The perivenous contribution to molecular clearance of CSF tracers}
+% \mer{MER: Target Results 5 300 words}
+\subsection*{Enlarged PVSs delay periarterial and intracranial molecular enrichment}
 
-\begin{itemize}
-\item
-  Look at clearance phase 12h-24h-48h-72h.
-\item
-  Look at concentration in the veins
-\item
-  Look at concentration in the parenchyma
-\item
-  Consider low-pressure sink for the perivenous spaces.
-\end{itemize}
+Enlarged PVSs are associated with impaired periarterial tracer
+transport and more generally with cognitive decline as well as
+certain subtypes of
+dementia~\cite{bown2022physiology,eide2024functional}. A key
+question is thus whether enlarged PVSs alone leads to delayed perivascular
+and intracranial tracer enrichment. To address this question from a
+fluid and transport dynamics perspective, we ask our in-silico model
+to predict the integrated effect of perivascular dilation on PVS flow
+velocities and tracer enrichment patterns (\Cref{fig:enlargement}A).
 
-\subsection*{Clearance of intraparenchymal solutes}
+Again we compute CSF production-, cardiac-, and vasomotion-induced net PVS flow velocities, but now in a perivascular geometry with an increased outer radius ($R^{\rm dilated}_2 = 3 R_1$) compared to the previous ($R^{\rm control}_2 = 2 R_1$), which corresponds to a $2.67 \times$ increase in PVS cross-section area.
+%% \begin{align*}
+%%   A_1 = \pi (R_2^2 - R_1^2) = \pi (2^2 R_1^2 - R_1^2) = 3 \pi R_1^2 \\
+%%   A_2 = \pi (\hat{R}_2^2 - R_1^2) = \pi (3^2 R_1^2 - R_1^2) = 8 \pi R_1^2 \\
+%%   A_2 / A_1 = 8/3 = 2.76
+%% \end{align*}
+The enlargement yields notable and non-uniform changes in the flow
+dynamics. On the one hand, the pressure-induced flow velocities
+increase from a mean of 0.08 \textmu m/s to 0.33 \textmu m/s due to
+the reduced effective resistance of the PVS. However, the peristaltic
+pumping becomes considerably less effective: the cardiac and
+vasomotion-induced net CSF flow velocities drop from a mean of 0.92
+\textmu m/s to 0.17 \textmu m/s and from 13.05 \textmu m/s to 2.33
+\textmu m/s, respectively (\Cref{fig:enlargement}B--D). In total, the
+net CSF velocity reduces, which in turn alters the tracer enrichment
+within the dilated PVS. Notably, we observe later tracer arrival in
+all MCA segments, with up to \discuss{120} min later arrival in the
+MCA2s for the high PVS flow scenario (\Cref{fig:enlargement}E, F). For
+the baseline model, which omits the vasomotion-induced net PVS flow,
+the effect is less pronounced in the MCA but clearly persists and is
+evident in the MCA2s (\Cref{fig:enlargement}F). Finally, we note that
+the larger volume of the dilated PVS results in a greater total
+accumulation of tracer in the PVS from 3--24 hours and reduced
+enrichment of the parenchyma at 6h, while the mean \discuss{PVS}
+concentration drops slightly (\Cref{fig:enlargement}G,H).
+\begin{figure}
+    \centering
+    \includegraphics[width=1 \textwidth]{figures/figure5.png}
+    \caption{\textbf{PVS enlargement reduces net PVS flow and delays tracer transport}
+    A) Modelling dilated PVS by extending the outer PVS boundaries;
+    B) mean and max PVS flow velocities with normal/control and dilated PVS;
+    C--D) 3D rendering of vasomotion-induced PVS flow in the normal (C) and dilated (D) periarterial networks; 
+    E) 2D slices showing zoom-in to the region surrounding the MCA-R after 1, 2, 3, and 4\,h with normal (lower) and dilated (upper) PVS.
+Innermost circle indicates the artery (blue), the surrounding annulus is the PVS (color represents concentration value), the surroundings are parenchyma and/or CSF spaces, separated by the pia (blue: towards the parenchyma, pink: towards the CSF));
+    F) First-time arrival for different branches of the arterial tree for the baseline model (with pressure-driven and cardiac-induced net PVS flow) and the high PVS flow model (also including vasomotion-induced net PVS flow) for normal and dilated PVS;
+    G) total tracer content in the CSF, parenchyma and arterial and venous PVS at 1,3,6,12 and 24\,h for the four models from F;
+    H) mean concentration in the CSF, parenchyma and arterial and venous PVS over time for the four models from F.
+    }
+    \label{fig:enlargement}
+\end{figure}
 
-\begin{itemize}
-\item
-  Clearance of a molecular concentration in the parenchyma, different
-  molecular sizes - ranging from Gadubutrol to amyloid-$\beta$/$\tau$.
-\end{itemize}
+% \mer{MER: Target Results 6 400 words}
+%\subsection*{The perivenous contribution to molecular clearance of CSF tracers}
 
-\bigskip
+%\begin{itemize}
+%\item Look at clearance phase 12h-24h-48h-72h.
+%\item Look at concentration in the veins
+%\item Look at concentration in the parenchyma
+%\item Consider low-pressure sink for the perivenous spaces.
+%\end{itemize}
 
-\FloatBarrier
-\newpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% \subsection*{Clearance of intra-parenchymal solutes}
+
+%% \begin{itemize}
+%% \item
+%%   Clearance of a molecular concentration in the parenchyma, different
+%%   molecular sizes - ranging from Gadubutrol to amyloid-$\beta$/$\tau$.
+%% \end{itemize}
+
 \section*{Discussion}
 
-\paragraph{CSF flow model validation}
+%% Summary of findings
+We have presented a high-fidelity in-silico model of molecular
+enrichment and clearance in human intracranial spaces, enabling
+tailored predictions of the influence of CSF space and (peri)vascular
+morphology, physiological factors such as cardiac and respiratory
+pulsatility, as well as of pathological conditions such as enlarged
+PVSs. A key observation is that the balance between CSF production and
+intracranial pulsatility significantly shapes the large-scale features
+of molecular enrichment, with e.g.~ventricular tracer reflux after
+intrathecal injection in the absence of CSF production, and preference
+towards anterior brain regions with reduced pulsatility. Indeed, the
+simulated distribution patterns cover a substantial range of the
+individual variations observed clinically \cite{ringstad2018brain}. In
+terms of perivascular pathways, we find that even moderate CSF flow on
+the order of 0.1--1.0 \textmu m/s in the surface PVSs results in
+earlier tracer enrichment around major cerebral arteries. This effect
+is more pronounced -- with up to a three-hour difference in time of
+arrival -- with PVS flow rates of approximately 10 \textmu m/s. Our
+models of CSF flow in perivascular networks, based on first principles
+and asymptotic analysis, predict net flow of such magnitudes and both
+antegrade and retrograde flow within the perivascular network,
+occurring with or without pressure differences at the network
+ends. The PVS may thus function as highways facilitating rapid
+transport, even in the absence of structural barriers, though a sharp
+concentration difference between the PVS and surrounding tissues is
+unlikely without such barriers. Finally, dilated PVS will cause a
+substantial reduction in cardiac- and vasomotion-driven flow
+velocities, which in turn impedes perivascular transport.
 
-Hornkjøl et al \cite{hornkjol2022csf} model CSF flow induced by production in the choroid plexus and report a peak velocity of $8.9\,$mm/s in the aqueduct.
-Given that similar flow rates were enforced, the deviation from our value of $2.0\,$mm/s is likely due to differences in diameter of the reconstructed aqueduct.
-Additionally, their peak velocities in the SAS range between 20 - 50 $\mu\,$m/s, which is compatible with our mean velocity of $4.4\,\mu$m/s (\Cref{fig:csf_flow_histo}).
+%% Comparison with others
 
-The aqueductal ($0.31\,$ml/s) and craniocervical ($8.31\,$ml/s) flow rates predicted by our model align well with previously reported values of $0.31\,$ml/s peak aqueductal flow rate \cite{vinje2019respiratory} and about $9\,$ml/s peak net blood inflow into the cranium \cite{baledent2014imaging}.
-Further, Linninger et al \cite{linninger2007cerebrospinal} model cardiac-induced CSF flow and predict a maximum transmantle pressure difference of $10\,$Pa, which is in good agreement with our estimate of $8.5\,$Pa as well as clinically measured mean pressure differences of $11.5\,$Pa between sensors placed subdurally and in the lateral ventricle \cite{vinje2019respiratory}.
-% 10.7 cm/s CSF flow at the craniovertebral junction in control \cite{bapuraj2016cerebrospinal}, area difference?
+% Tracer enrichment predictions
+While molecular enrichment of the brain and surrounding CSF spaces
+have been extensively studied in animal models, especially in murine
+models in connection with studies of the glymphatic system, there are
+fewer reports of contrast enrichment in human subjects over 0--24h. We
+therefore mainly compare our in-silico predictions with the series of
+papers by Eide and colleagues~\cite{ringstad2017glymphatic,
+  ringstad2018brain, eide2021sleep, eide2024functional} and Watts et
+al~\cite{watts2019measuring}. Ringstad et al.~\cite{ringstad2018brain}
+report of tracer enrichment in a centripetal pattern, primarily in
+regions near large cerebral surface arteries, peaking in the CSF
+spaces between 6 and 24 hours, with parenchymal tracer content still
+increasing after 24 hours, but with large individual
+differences. Watts et al~\cite{watts2019measuring} report of a similar
+enrichment pattern and a concentration peak in the SAS of $\approx$0.5
+mmol/L occuring after 10 to 15 hours. Our in-silico enrichment
+patterns agree with these observations, though with an earlier
+\discuss{$\sim$5 + plus spinal transit time)} rather than later peak
+in the CSF spaces. In particular, we note that the substantial
+individual variation observed clinically is comparable to the
+variation in in-silico enrichment patterns associated with reduced CSF
+production or reduced intracranial pulsatility. \mer{One sentence on
+  tracer content comparison.} In terms of perivascular tracer
+enrichment, Eide and Ringstad~\cite{eide2024functional} report of
+first-time tracer appearance in the PVS of the MCA after 37.8 $\pm$
+47.0 (M1), 53.1 $\pm$ 50.5 (M2) and 82.6 $\pm$ 61.5 (M3) min, thus
+with just a 15.3 min delay between M2 and M1. Comparing with the our
+FTAs in the MCA segments, we note that the baseline model predicts a
+more-than 3 hour difference in FTA between the M1 and M2. Instead, the
+high PVS flow scenario predicts a 24 min delay between these
+segments. Moreover, the high PVS flow scenario predicts FTAs of
+84--108 min (M1), 84--146 min (M2) which overlaps with the previous
+ranges \discuss{modulo spinal transit time}. We interpret these
+results as indicative of a net CSF flow in human surface PVS beyond
+what would be induced by CSF production and cardiac peristalsis alone.
+% M2 - M1 (mean) ~ 15.3 min
 
-Sharp et al used the theory of dispersive mixing to predict a dispersion enhancement factor of $91.8$ in the (open) spinal SAS, which lies well within the range of our spatially varying estimate (ranging from 0 to 160).
-Conversely, \cite{ray2021quantitative} report up to 10,000 times greater diffusion in perivascular regions, but their modelling lumps diffusive, dispersive and advective transport in a single enhancement factor and is thus not directly comparable.
+% PVS flow predictions. 
+The existence, magnitude, and directionality of flow in surface and
+parenchymal PVSs have been the subject of active debate for
+decades~\cite{rennels1985evidence, bilston2003arterial,
+  hadaczek2006perivascular, carare2008solutes, iliff2012paravascular,
+  iliff2013cerebral, bakker2016lymphatic, mestre2018flow,
+  bedussi2018paravascular, rey2018pulsatile, thomas2019fluid,
+  daversin2020mechanisms, kedarasetti2020functional,
+  martinac2021phase, vanveluw2020vasomotion, bohr2022glymphatic,
+  boster2023artificial, gjerde2023directional,
+  nozaleda2024arterial}. In mice, Mestre et al~\cite{mestre2018flow}
+observe CSF flow in surface perivascular regions surrounding branches
+of the MCA with average flow speeds of 18.7 \textmu m/s and net flow
+in the antegrade direction, but also regions with retrograde
+flow. Similarly, Bedussi et al.~\cite{bedussi2018paravascular} report
+of oscillatory flow, with an average (net) CSF velocity of 17 $\pm$ 2
+\textmu m/s in the antegrade direction. Using physics-informed neural
+networks, Boster et al~\cite{boster2023artificial} estimate PVS
+velocities of $12.75 \pm 6.25$ \textmu m/s. Less is known about the
+magnitude and directionality of CSF flow in surface PVSs in
+humans. Here, assuming that human surface PVS follow the major surface
+vessels and are CSF-filled but otherwise open regions of width
+proportional to the vessel radius, our estimates indicate that the CSF
+pressure differences due to CSF production, cardiac peristaltic
+pumping and low-frequency vasomotion all yield spatially-varying net
+CSF flow with both antegrade and retrograde PVS segments. This
+observation is thus in agreement with the experimental
+observations~\cite{mestre2018flow, bedussi2018paravascular}, and
+inherently supports the original notion that antegrade and retrograde
+solute transport along PVSs may coexist~\cite{rennels1985evidence}. In
+terms of magnitude, the contribution from CSF production and cardiac
+wall motion are small (0.08 \textmu m/s and 0.92 \textmu m/s on
+average), but intriguingly, our estimates of the contribution from
+vasomotion (13.05 \textmu m/s) are comparable to the experimental
+reports. We do note that this estimate is based on the presence of
+rhythmic waves of vasomotion at a frequency of 0.1Hz, a wavelength of
+20 mm, and a wall displacement of 10\% -- in the antegrade direction,
+all values currently associated with significant
+uncertainty~\cite{gokina1996electrical, vanveluw2020vasomotion,
+  broggini2024long}.
 
-\subsection*{Limitations}
+% Compartmentalization of the PVS
+To what extent are surface PVSs in communication with the surrounding
+SAS and to what extent are these structurally separated compartments?
+Originally, Weller and coauthors~\cite{zhang1990interrelationships,
+  zhang1992directional, weller2005microscopic} identified thin sheaths
+of pial cells surrounding human surface arteries and penetrating
+arterioles (but not veins or venules). On the other hand, Bakker and
+coauthors report that the subarachnoid space, the cisterns, ventricles
+and penetrating periarteriolar spaces form a continuous CSF
+space~\cite{bedussi2017paravascular}. Thorne and colleagues study the
+architectures of rat perivascular spaces in detail, emphasizing the
+presence of openings (stomata) on the interface between the
+vasculature and the CSF within the SAS~\cite{pizzo2018intrathecal,
+  hannocks2018molecular}. Further, Mestre et
+al~\cite{mestre2022periarteriolar} study the properties of pial
+perivascular spaces in mice, and report of pial cells forming sheaths
+for larger surface arteries and partially cover smaller surface
+arteries, with higher coverage in ventral SAS regions. Our results
+demonstrate that the PVSs may act as rapid transport pathways even
+with only a partial barrier between the PVS and surrounding CSF, thus
+contributing to quantifying the properties of the human PVS-CSF
+interface~\cite{eide2024functional}.
 
-\begin{itemize}
-\item
-  \mer{MER/MC: How about dispersion in the PVS?}
-\item 
-  Inaccuracies in vascular geometry representation - and consequences downstream for e.g. numerical integration.
-\item
-  PVS geometry assumptions and approximations. Humans different from mice. Limitations in resolution.
-\item
-  No vasculature in the cerebellum, and of course no vasculature below certain resolution.
-\item
-  Only partially accounting for pulsatility. 
-\item
-  Accounting for brain diffusive anisotropy. Little effect expected. Maybe something to consider in small follow-up work.
-\item
-  Permeability and stiffness of the pial membrane
-\item
-  PVS-blood exchange and BBB leakage.
-Natural to consider in model variation here, or follow-up work.
-We here ignore PVS-blood exchange, and thus only model the perivascular concentration at the network level (in addition to the SAS/brain concentration in the surroundings).
+% Variations with sleep, exercise
+Lifestyle factors such as sleep~\cite{xie2013sleep, miao2024brain,
+  bojarskaite2023sleep, eide2021sleep, vinje2023human,
+  hauglund2025norepinephrine, larsen2024sleep},
+exercise~\cite{holstein2018voluntary, yildiz2022immediate}, and
+alcohol intake~\cite{lundgaard2018beneficial} affect molecular
+enrichment and clearance of the brain. In particular, pial arteries
+display higher-amplitude low-frequency vasomotion during NREM and
+intermediate sleep states, while REM sleep is associated with reduced
+PVS width~\cite{bojarskaite2023sleep}. Peristaltic pumping is more
+effective under both of these configurations, resulting in higher
+estimated net PVS flow velocities~\cite{gjerde2023directional}. Even
+higher PVS velocities would amplify the PVS flow effects simulated
+here, with earlier arrival times along the PVSs, clear enrichment in
+adjacent spaces, and accelerated molecular enrichment
+overall. Moreover, sleep is linked with low-frequency oscillations in
+human CSF flow~\cite{fultz2019coupled}. As we have shown here, the
+level of dispersion in the CSF spaces strongly shapes molecular
+enrichment patterns. While we, at baseline, account for the dispersion
+induced by cardiac and respiratory pulsatility, the high dispersion
+scenarios considered can be viewed as representative of additional
+pulsatility induced by e.g.~neural waves~\cite{fultz2019coupled,
+  williams2023neural} or other respiration patterns depending on
+activity level. More broadly, we consider this as a key strength of
+the current in-silico framework: allowing for integrating and
+exploring the mechanistic macroscale implications of changes in
+pulsatility at smaller spatial or temporal scales.
 
-\end{itemize}
+% Variations with disease states
+%cohort variations, individual variations, physiological factor
+%variations. Enlarged PVS reduce transport at the whole cerebral scale.
 
-\subsection*{Outlook}
+In terms of limitations, our primary constraint modelling-wise is the
+geometric representation of the vascular (and therefore also
+perivascular) networks. Concretely, we account for no vessels below
+the original MRI resolution limits, and thus only a part of the
+cerebral surface vasculature and effectively not the cortical
+vessels. In addition, and again due to lack of resolution, there is
+uncertainty associated with the location of the vessel segments
+relative to the CSF spaces and parenchyma, with some vessels
+surrounded by both domains and some vessels completely entering the
+parenchyma before returning to the SAS. Higher resolution MRI data
+(T1w and e.g. time-of-flight, QSM or other modalities) in more
+subjects would immediately allow for more detailed studies of this
+aspect. As a result of the lack of an accurate representation of the
+cortical and subcortical white matter vasculature, we have considered
+a conservative transport model within the parenchyma, modelling
+extracellular diffusion alone. To compensate, we have focused on
+reporting detailed predictions in the CSF spaces and surface PVS after
+not more than 24 hours. In particular, we note that, we observe tracer
+accumulation at the perivascular network ends after \fixme{24} hours,
+similar to that previously observed for
+microspheres~\cite{bedussi2018paravascular, mestre2018flow}, which
+here can be interpreted as an artifact of the incomplete network
+creating up a barrier for tracer movement. Previously, we have found
+that enrichment within the parenchyma is well-represented by $\approx
+3\times$ enhanced diffusion augmented by local clearance e.g.~across
+the blood-brain barrier or diffusion-convection with a complex pattern
+of spatially-varying convective velocities of 1-10 \textmu
+m/min~\cite{vinje2023human}. We will incorporate such transport
+mechanisms within the parenchyma in future work. Finally, we also note
+that we simulate the intracranial distribution of a tracer
+concentration appearing at the craniocervical junction rather than as
+an intrathecal injection. We therefore overestimate the amount of
+tracer entering into the intracranial spaces; however, due to the
+linearity of the mathematical model, it is valid to directly interpret
+the in-silico concentrations relative to the total amount of tracer.
 
-\begin{itemize}
-\item
-  Considering the (maybe) low diffusion in CSF spaces, maybe pure
-  advection would be relevant? Both computational and physiological
-  implications/interest.
-\end{itemize}
-
-\newpage
 \section*{Methods}
-
-%\mer{MER: I recommend writing each (sub)section of Methods as a nearly stand-alone paragraph.
-Let's see if we can avoid subsubsections/paragraphs.
-If you need to add more detail than you think is appropriate, put it in a Supplementary Methods section in the Appendix.}
 
@@
 including the ventricular system and SAS, using
-Synthseg~\cite{billot2023robust,billot2023synthseg}. We next manually
-adjust the segmentation to accurately represent the connections
-between CSF spaces (aqueduct, median aperture, tentorium cerebelli), smoothen, and finally
-extract surface representations of the outer (arachnoid) boundary, the
-pial membrane and other interfaces. Conforming to these surface and
+Synthseg~\cite{billot2023robust,billot2023synthseg}
+(\Cref{fig:results1}A--B). We next manually adjust the segmentation to
+accurately represent the connections between CSF spaces (aqueduct,
+median aperture, tentorium cerebelli), smoothen, and finally extract
+surface representations of the outer (arachnoid) boundary, the pial
+membrane and other interfaces. Conforming to these surface and
 interface representations, we generate a tetrahedral mesh $\Omega$
@@
 parenchyma $\Omega_{\rm PAR}$ and CSF spaces $\Omega_{\rm CSF}$, using
-fTetWild~\cite{hu2020fast}. The resulting computational mesh consists
-of $221\,988$ vertices and $1\,225\,774$ mesh cells, which vary between
-$0.24$ mm and $1.29$ cm in mesh cell diameter. The parenchyma has a
-total volume of $1318$ ml, whereas the ventricles and the outer SAS
-contribute $32$ ml and $328$ ml, respectively, to a total CSF volume
-of $360$ml, in agreement with recent estimates~\cite{hladky2024regulation}.
+fTetWild~\cite{hu2020fast} (\Cref{fig:csf}A--B). The resulting
+computational mesh consists of $233\,592$ vertices and $1\,290\,131$
+mesh cells, which vary between $0.22$ mm and $8.9$ mm in mesh cell
+diameter. The parenchyma has a total volume of $1318$ ml, whereas the
+ventricles and the outer SAS contribute $32$ ml and $329$ ml,
+respectively, to a total CSF volume of $361$ml, in agreement with
+recent estimates~\cite{hladky2024regulation}.
 
@@
 centerline of a blood vessel segment and labeled with its lumen radius
-$R_1^i$. We also assign an outer radius $R_2^i > R_1^i$ to each
-segment $i$, form the annular cylinder ensheathing $\Lambda^i$ and
-define the union of these as the PVS. The periarterial network graph
-consists of $12\,708$ edges connected at $12\,562$ inner nodes and
-with $147$ end nodes, and the associated domain is denoted by
-$\Lambda_a$. We identify three of the end nodes -- corresponding to
-the two internal carotid arteries (ICAs) and the basilar artery (BA)
--- as root nodes, and designate the other ends as leaf nodes. The
-resulting perivenous network graph consists of $\fixme{N}$ edges
-connected at $\fixme{N}$ inner nodes and with $\fixme{N}$ end nodes,
-and the associated domain is denoted by $\Lambda_v$. \fixme{Finally, we label
-the outer surface of the PVS associated with $\Lambda_a$ and
-$\Lambda_v$, by $\Gamma_a$ and $\Gamma_v$ respectively.}
+$R_1^i$ (\Cref{fig:results1}C, \Cref{fig:pvs}A). We also assign an
+outer radius $R_2^i > R_1^i$ to each segment $i$, form the annular
+cylinder ensheathing $\Lambda^i$ and define the union of these as the
+PVS. The periarterial network graph consists of $12\,708$ edges
+connected at $12\,562$ inner nodes and with $147$ end nodes, and the
+associated domain is denoted by $\Lambda_a$. We identify three of the
+end nodes -- corresponding to the two internal carotid arteries (ICAs)
+and the basilar artery (BA) -- as root nodes, and designate the other
+ends as leaf nodes. The resulting perivenous network graph consists of
+$\fixme{N}$ edges connected at $\fixme{N}$ inner nodes and with
+$\fixme{N}$ end nodes, and the associated domain is denoted by
+$\Lambda_v$. \discuss{Finally, we label the outer surface of the PVSs
+  associated with $\Lambda_a$ and $\Lambda_v$, by $\Gamma_a$ and
+  $\Gamma_v$ respectively.}
 
@@
 $\Omega_{\rm CSF}$ and in the brain parenchyma $\Omega_{\rm PAR}$ via
-a mixed-dimensional transport model~\cite{masri2023modelling} over a
-timescale of minutes to hours. Specifically, for $t > 0$, we solve for
-a concentration $c = c(x, t)$ in the 3D intracranial compartments ($x
+a mixed-dimensional transport model~\cite{masri2024modelling,laurino2019derivation} over a
+timescale of minutes to hours (\Cref{fig:results1}A--C,
+\Cref{fig:csf}A--C). Specifically, for $t > 0$, we solve for a
+concentration $c = c(x, t)$ in the 3D intracranial compartments ($x
 \in \Omega_{\rm CSF}, x \in \Omega_{\rm PAR}$) and for a cross-section
@@
 R_2$ is the perimeter and $A = \pi (R_2^2 - R_1^2)$ the area of the
-PVS cross-sections, $\delta_{\Gamma}$ is the Dirac delta function
-concentrated on the interfaces $\Gamma_a, \Gamma_v$, the overlined
-$\overline{c}$ denotes the average of $c$ over cross-sections of
-$\Gamma_a, \Gamma_v$, and $\xi$ is a permeability allowing for
-transfer/exchange over the interfaces $\Gamma_a, \Gamma_v$ between the
-periarterial and perivenous networks and their surroundings.
-Moreover, we model the membrane between the parenchyma and the CSF
-spaces as a semi-permeable membrane with permeability cofficient
-$\beta_{\rm pia}$. For a complete description of this mathematical
-model, see \Cref{sec:more_maths}.
+PVS cross-sections, $\delta_{\Gamma}$ is a Dirac term concentrated on
+the interfaces $\Gamma_a, \Gamma_v$, the overlined $\overline{c}$
+denotes the average of $c$ over cross-sections of $\Gamma_a,
+\Gamma_v$, and $\xi$ is a permeability allowing for transfer/exchange
+over the interfaces $\Gamma_a, \Gamma_v$ between the periarterial and
+perivenous networks and their surroundings.  Moreover, we model the
+membrane between the parenchyma and the CSF spaces as a semi-permeable
+membrane with permeability coefficient $\beta_{\rm pia}$. For a
+complete description of this mathematical model and simulation
+scenarios, see \Cref{sec:app:more_maths}. The initial concentrations
+in the three-dimensional domain $\Omega$ and networks $\Lambda_a,
+\Lambda_v$ are all set to zero.
 
@@
 compartment explicitly, we prescribe a time-dependent molecular influx
-over the interface towards the spinal compartment with the condition
-$(D \nabla c - \bm u c ) \cdot \bm{n} = g_{\mathrm{influx}}$ on
-$\Gamma_{\mathrm{SSAS}}$. We set
-\begin{equation}
-  g_\mathrm{influx}(t) = \frac{m_{\rm tot}}{T_{\rm max}^2 |\Gamma_{\rm SSAS}|} \max(0, T_{\rm max} - |t - T_{\rm max}|)
-\end{equation}
+over the interface towards the spinal CSF compartment with the condition
+\begin{equation*}
+  (D \nabla c - \bm u c ) \cdot \bm{n} = g_{\mathrm{influx}},  \quad \mathrm{on}  \quad \Gamma_{\mathrm{SSAS}}.
+\end{equation*}
+Here we set
+\begin{equation*}
+  g_\mathrm{influx}(t) = \frac{m_{\rm tot}}{T_{\rm max}^2 |\Gamma_{\rm SSAS}|} \max(0, T_{\rm max} - |t - T_{\rm max}|), 
+\end{equation*}
 which expresses a hat-shaped influx function with a peak at $T_{\rm max}$ and a total tracer injection of $m_{\rm tot}$ (\Cref{tab:parameters}).
@@
 \subsection*{Molecular clearance from the intracranial space and perivascular network ends}
-%\subsection*{Boundary conditions: extracranial efflux and perivascular parenchymal influx}
 
@@
 pathways, across the upper, outer (arachnoid) boundary
-(\Cref{fig:intracranial_domains}), proportional to the concentration
-in the SAS with a rate constant $\beta_{\rm exit}$. We set a no-flux
-condition at the end nodes of the perivascular networks, not allowing
-for direct molecular clearance there.
+(\Cref{fig:results1}A, \Cref{fig:csf}C), proportional to the
+concentration in the SAS with a rate constant $\beta_{\rm exit}$. We
+set a no-flux condition at the end nodes of the perivascular networks,
+not allowing for direct molecular clearance there.
 
 \subsection*{CSF flow in the SAS and ventricular system}
-\label{sec:csf_fluid_vel}
+%\label{sec:csf_fluid_vel}
 
@@
 $\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport}
-(\Cref{tab:parameters}, \Cref{fig:intracranial_domains},
-\Cref{fig:csf_flow_prod}, see also \Cref{sec:more_maths}). We do not
-directly model convection within the brain tissue (beyond the
-perivascular spaces), and thus set $\bm u = 0$ in $\Omega_{\rm PAR}$
+(\Cref{tab:parameters}, \Cref{fig:csf}A--E, see also
+\Cref{sec:app:more_maths}). We do not model bulk flow within the brain
+parenchyma, except within the perivascular network segments extending
+below the pial surface, and thus set $\bm u = 0$ in $\Omega_{\rm PAR}$
 in~\eqref{eq:multi_transport}.
@@
 \subsection*{Dispersion in the CSF spaces}
+%\label{sec:csf_dispersion}
+%% MER: I'm commenting out labels as long as there is no numbering
 
@@
 dispersion~\cite{taylor1953dispersion, watson1983diffusion,
-  asgari2016glymphatic, keith2019dispersion, ray2021quantitative,
+  asgari2016glymphatic, sharp2019dispersion, ray2021quantitative,
   troyetsky2021dispersion}. To account for these dispersive effects,
-we compute a spatially-varying dispersion factor $R$ in $\Omega_{\rm
-  CSF}$ (\Cref{fig:csf_flow_R}) by combining a computational model of
-the CSF pressure gradient at peak arterial dilation with theoretical
-estimates for shear-augmented (Taylor) dispersion
-\cite{taylor1953dispersion, watson1983diffusion, keith2019dispersion}. \mer{Include 1-2 sentences from Supplementary Methods here -- but which? And also revisit references with MC.}
+while bridging from the pulsatile flow time scale of seconds to a
+molecular transport time scale of hours, we compute cardiac and
+respiratory dispersion factors $R_c$ and $R_r$, both
+spatially-varying, in $\Omega_{\rm CSF}$ (\Cref{fig:csf}A) by
+combining computational estimates of the respective peak CSF pressure
+gradients with theoretical estimates for shear-augmented (Taylor)
+dispersion \cite{taylor1953dispersion, watson1983diffusion,
+  sharp2019dispersion} (see \Cref{sec:app:dispersion} for a complete
+description). Combining these effects, we then set $R = R_c + R_r$ and
+define the effective diffusion coefficient $D = (1 + R) D^{\rm Gad}$
+in $\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport} in the baseline model.
 
-\mer{MER: Stopped here. WIP in Supplementary Methods.}
+\subsection*{Perivascular fluid flow induced by CSF pressure differences}
 
-\subsection*{Dispersion in the PVS}
+To estimate the contribution also to perivascular fluid flow from CSF
+production, we impose the fluid pressure $p_{\rm CSF}$ induced by CSF
+production, computed in the SAS and ventricular system, at the end
+nodes of the periarterial network located within the SAS. For the
+network end nodes located within the parenchyma, we compute a harmonic
+extension of the CSF pressure field (by solving a Laplace equation in
+$\Omega_{\rm PAR}$ with $p_{\rm CSF}$ as the boundary value), and set
+the corresponding value at these end nodes. We then numerically solve
+a system of hydraulic network equations to compute $\hat{u}^p_a$
+defined over $\Lambda_a$; i.e., the pressure-difference induced CSF
+flow velocity in the periarterial network (see
+also~\Cref{sec:app:hydraulic_network}). We repeat this procedure to
+compute a corresponding velocity $\hat{u}^p_v$ in the perivenous
+network $\Lambda_v$.
 
-Similar to the CSF spaces, dispersion due to cardiac and respiratory pulsation has been hypothesized to enhance transport in the perivascular spaces.
-However, recent estimates indicate that the effect of dispersion does not exceed a factor of 2 compared to diffusion alone \cite{asgari2016glymphatic,keith2019dispersion,bojarskaite2023sleep,asgari2016glymphatic,troyetsky2021dispersion}, and we thus do not model dispersive effects in the PVS.
-\mar{Maybe we should just take a factor of 2 as the upper bound?}
+%% \subsection*{Dispersion in the PVS}
 
-% \cite{keith2019dispersion}: dispersion in PVS is negligiable in porous PVS, but might be relevant in open PVS (R=220) (theoretical estimates)
+%% Similar to the CSF spaces, dispersion due to cardiac and respiratory pulsation has been hypothesized to enhance transport in the perivascular spaces.
+However, recent estimates indicate that the effect of dispersion does not exceed a factor of 2 compared to diffusion alone \cite{asgari2016glymphatic,sharp2019dispersion,bojarskaite2023sleep,asgari2016glymphatic,troyetsky2021dispersion}.
+
+% \cite{sharp2019dispersion}: dispersion in PVS is negligiable in porous PVS, but might be relevant in open PVS (R=220) (theoretical estimates)
 %  \cite{bojarskaite2023sleep} numerical simulations show that the enhancement factor is likely in the range of 10 - 50 \%
@@
 
-\subsection*{Perivascular fluid flow induced by CSF pressure differences}
-
-Since the pressure gradient the caused by the steady production-driven CSF flow (see \cref{sec:csf_fluid_vel}), we impose the precomputed fluid pressure $p_{\rm CSF}$ on the networks root nodes:
-\begin{equation}
-    \hat{p} = p_{\rm CSF} \quad \text{ on } \Lambda^{\rm out}
-\end{equation}
-In case that the root does not lie within the CSF-space but the parenchyma, the CSF pressure is not yet computed.
-For that reason, we apply a simple extension technique to map the CSF pressure field into $\Omega_{\rm PAR}$.
-In particular, we solve a Laplace equation on $\Omega_{\rm PAR}$ with the CSF pressure on the Pial membrane ($\Gamma_{\rm Pia}$) as a boundary condition.
-
-\subsection*{Perivascular fluid flow induced by vascular wall motion}
+\subsection*{Perivascular fluid flow induced by arterial wall motion}
 
 Using the theoretical framework introduced by Gjerde et
-al.~\cite{gjerde2023directional}, we compute analytic estimates for
-the time-average perivascular flow rate $\langle Q' \rangle$ induced
-by rhythmic vascular wall motions for different frequencies $f$,
-amplitudes $\varepsilon$ and wave lengths $\lambda$ in the arterial
-network $\Lambda^a$ (see also Supplementary
-Methods~\ref{sec:sup:peristalsis}). The corresponding contribution to
-the convective velocity $\hat{u}_{\rm x}$ is defined vessel-wise as
-$\hat{u}_{\rm x}|_{\Lambda_i} = \langle Q'_i \rangle/A_i$ where $A_i$
-is the cross-section area of the PVS segment ($A_i = \pi (R_2^i -
-R_1^i$)) and $\langle Q_i' \rangle$ is its mean flow rate. Two wall
-motion patterns are considered: \emph{cardiac} pulsations with $f =
-1.0$ (Hz), $\lambda = 2000$ (mm) and $\varepsilon = 1\%$, and very
-low-frequency \emph{vasomotion} with $f = 0.1$ (Hz), $\lambda = 20$
-(mm), and $\varepsilon = 10\%$. \mer{The corresponding net convective
-velocities (in the antegrade direction) are for \emph{vasomotion}:
-$9.345 \pm 6.695$ $\mu$m/s, range: $(-65.81, 63.96) \mu$m/s, and for
-\emph{cardiac}: $0.475 \pm 0.5115 \mu$m/s, range: $(- 3.825, 4.861)
-\mu$m/s, where negative velocities correspond to retrograde flow
-(\Cref{fig:pvs}\mer{X}).} No perivascular fluid flow induced by wall
-motion is considered for the venous network.
+al.~\cite{gjerde2023directional}, we also compute analytic estimates
+for the time-average perivascular flow rate $\langle Q' \rangle$
+induced by peristalsis in the arterial network $\Lambda_a$; i.e.~the
+net flow induced by traveling waves of arterial wall motion of
+frequency $f$, amplitude $\varepsilon$ and wave lengths $\lambda$
+(\Cref{fig:pvs}D). The corresponding contribution to the periarterial
+flow velocity is defined for each segment $\Lambda_i$ as $\langle Q'_i
+\rangle/A_i$ where $A_i = \pi (R_2^i - R_1^i)$ is the cross-section
+area of the PVS segment and $\langle Q_i' \rangle$ is the estimated
+mean flow rate (see Supplementary Methods~\ref{sec:app:peristalsis}
+for more details). Two wall motion patterns are considered, yielding
+$\hat{u}_a^c$ and $\hat{u}_a^v$ respectively: \emph{cardiac}
+pulsations with $f = 1.0$Hz, $\lambda = 2$ m and $\varepsilon = 1\%$
+(\Cref{fig:pvs}E), and very low-frequency \emph{vasomotion} with $f =
+0.1$Hz, $\lambda = 0.02$m, and $\varepsilon = 10\%$
+(\Cref{fig:pvs}F). No flow induced by peristalsis is considered for
+the perivenous network. In the baseline model, we combine the
+contributions from CSF production and cardiac peristalsis to set
+$\hat{u}$ in~\eqref{eq:multi_transport} as $\hat{u}_a = \hat{u}_a^p +
+\hat{u}_a^c$ in $\Lambda_a$, and $\hat{u}_v = \hat{u}_v^p$ in
+$\Lambda_v$.
 
@@
 
-Model and material parameters are summarized in \Cref{tab:parameters}.
-We assume that the porosity equals the extracellular space volume
-fraction $\phi = \phi_{\rm PAR}$ within the parenchyma and $\phi = 1$
-elsewhere. In influx scenarios, we model the concentration of
-Gadubutrol with effective diffusion coefficient equal to that of
-Gadubutrol in free water in the PVS networks: $\hat{D} = D^{\rm Gad}$;
-weighted by a dispersion factor $R$ in the CSF spaces:
+Model and material parameters are summarized in \Cref{tab:parameters},
+and~\Cref{tab:scenarios} gives an overview of the different in-silico
+scenarios. We assume that the porosity equals the extracellular space
+volume fraction $\phi = \phi_{\rm PAR}$ within the parenchyma and
+$\phi = 1$ elsewhere. The effective diffusion coefficient is set equal
+to that of Gadubutrol in free water in the PVS networks: $\hat{D} =
+D^{\rm Gad}$; weighted by a dispersion factor $R$ in the CSF spaces:
 $D|_{\Omega_{\rm CSF}} = ( 1 + R) D^{\rm Gad}$; and modulated by the
-porosity and tortuousity ($\lambda = 1.78$) in the parenchyma:
+porosity and tortuosity ($\lambda = 1.78$) in the parenchyma:
 $D|_{\Omega_{\rm PAR}} = D^{\rm Gad}_{\rm PAR}$. We set $\xi$
@@
 spaces ($\xi = \xi_{\rm CSF}$) or parenchyma ($\xi = \xi_{\rm EF}$),
-or neither (\Cref{sec:more_maths}). Additionally, we increase the
-permeability $\xi$ ($100\times$) near the leaf nodes of $\Lambda_a$ to
-account for the transition from explicitly resolved PVSs to their
-unresolved continuation in the CSF spaces and parenchyma.
-
+or neither (\Cref{sec:app:more_maths}). Additionally, we increase the
+permeability $\xi$ ($100\times$) \fixme{near} the leaf nodes of
+$\Lambda_a$ to account for the transition from explicitly resolved
+PVSs to their unresolved continuation in the CSF spaces and
+parenchyma.
 \begin{table}
@@
 \end{table}
-
-\subsection*{Initial conditions}
-
-We set the initial concentrations $c_0, \hat{c}_0$ in the
-three-dimensional domain $\Omega$ and networks $\Lambda_a, \Lambda_v$
-to be zero for all influx scenarios. For parenchymal
-clearance scenarios, we set $c_0 = 1, \hat{c}_0 = 0$.
-
-\subsection*{Numerical approximation of multi-dimensional transport and CSF flow}
-
-\mer{@Rami: Could you draft one paragraph here describing the essential bits of the numerical scheme?
-I recommend ``describing with words'' to reach relevant audiences.
-Refer to the Supplementary Methods for details.
-Mention theoretical expected convergence order.}
-
-In this section, we provide a brief overview on the employed numerical discretization.  
-We refer to \Cref{sec:details_numerical_method} for details. 
-Our implementation relies on the FEniCS finite element framework \cite{logg2012automated}; the 3D-1D coupling is handled by its extension FEniCS\textsubscript{ii} \cite{kuchta2020assembly}.
-The numerical scheme for the CSF flow equations utilizes polynomial spaces that preserve the divergence free condition \eqref{eq:divergence_equation} on the discrete level \cite{hong2016robust}.
-This requirement is important for stability in the transport equation, see for example \cite{cesmelioglu2022compatible}.
-The spaces we used for the velocity are continuous along the normal direction; continuity of the tangential component is enforced via interior penalty approaches \cite{hong2016robust}.
-If the exact solution of \eqref{eq:momnetum_equation}--\eqref{eq:divergence_equation} is smooth, then one expects optimal error convergence rates for this approximation.
-
-For the 3D transport problem, we use the interior penalty discontinuous Galerkin (DG) method with weighted averages and upwinding for the convection term \cite{ern2009discontinuous}.
-We remark that the choice of DG in the 3D domain is necessary here since  continuous Galerkin methods even when supplemented with artificial diffusion stabilization terms were unstable in our experiments.
-For the 1D transport problem, we use continuous Galerkin spaces defined over the entire network with additional artificial diffusion silization when the 1D velocity is non--zero.
-DG methods for the coupled 3D--1D problem have been recently analyzed  where convergence of the method for the elliptic case is established \cite{masri2024discontinuous}.
-While the analysis is for DG methods for both the 3D and the 1D problem, we expect similar arguments to hold here.
-However, note that the 3D--1D coupling induces a singularity in the 3D solution and one can only expect suboptimal error rates near the 1D network.
-In subdomains excluding the 1D lines, almost optimal rates can be expected for both DG methods \cite{masri2023discontinuous} and CG methods \cite{koppl2016local}.
-Extension of the convergence analysis to convection--diffusion problems is an interesting future research direction.
-
 \begin{table}
   \begin{center}
-  \begin{tabular}{ll|llllll|ll}
+  \begin{tabular}{ll|llll|lll}
     \toprule
-    & Description & $R_1, R_2$ & $\xi_{\rm PVS-SAS}$ &  $\beta_{\rm efflux}$ & $D$ & $\hat{u}$ & $\mathbf{u}_{SAS}$ & $c_0$ \\
+    & Description & $R_1, R_2$ & $\xi_{\rm PVS-CSF}$ & $D|_{\Omega_{\rm CSF}}$ & $D|_{\Omega_{\rm PAR}}$ & $\mathbf{u}|_{\Omega_{\rm CSF}}$ & $\hat{u}|_{\Lambda_a}$ & $\hat{u}|_{\Lambda_v}$  \\
     \midrule
-    A & Baseline with sheaths & $R_2 = 2 R_1$ & $\xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
-    B1 & Shealth permeability & $R_2 = 2 R_1$ & $0, \xi, 10 \xi, 100 \xi, 1000 \xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
-    B2 & Shealth permeability & $R_2 = 2 R_1$ & $0, \xi, 10 \xi, 100 \xi, 1000 \xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} +  \hat{u}_{\rm cardiac} + \hat{u}_{\rm vaso}$ & $\mathbf{u}_{\rm prod}$ & 0 \\
-    %C & Larger PVS & $R_2 = N R_1$ & -- & -- & --  &  --  & -- &  -- \\
-    C & Higher PVS flow & -- & $\xi$ & -- & --  & $\hat{u}_{\rm prod} +  \hat{u}_{\rm cardiac} + \hat{u}_{\rm vaso}$  & -- & --  \\
-    AA & PVS size & $R_2 = x \beta R_1$ & $\xi$ & -- & -- & -- &  -- & 0 \\ 
-    %C & Larger PVS & $R_2 = N R_1$ & -- & -- & --  &  --  & -- &  -- \\
-    B & PVS flow & -- & $\xi$ & -- & --  & $\hat{u}_{\rm prod} + \hat{u}_{\rm vaso}$  & -- & --  \\
-    C & Clearance & -- & -- & -- & $\alpha D^{\rm gad, \tau, A\beta}$  & $\hat{u}_{\rm prod}$ & -- & 1 \\
-    %E & PVS II & $R_2 = 2 R_1$ & $\xi$, -- & -- & --  &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow\uparrow$  & -- & --  \\
-    %F & PVS III & $R_2 = 2 R_1$ & $\xi$, -- & -- & $10 D_{\rm PVS}$ &  $\hat{u} = \hat{u}_{\rm prod}$  & -- & --  \\
-    %G & Glymphatics & $R_2 = 2 R_1$ & $\xi$, -- & -- & \cite{sykova2008diffusion, valnes2020apparent} &  $\hat{u} = \hat{u}_{\rm prod} + \uparrow$  & $\mathbf{u}_{\rm PAR}$ > 0 & --  \\
+     & baseline & $R_2 = 2 R_1$ & $\xi$ & $(1 + R) D^{\rm Gad}$ & $D^{\rm Gad}_{\rm PAR}$ & $\mathbf{u}_{\rm CSF}$ & $\hat{u}_a^p + \hat{u}_a^c$  & $\hat{u}_v^p$ \\ 
+    & no CSF production & $-$ & $-$ & $-$ & $-$ & $\mathbf{0}$ & $\hat{u}_a^c$ & $0$ \\ 
+    & reduced pulsatility & $-$ & $-$ & $ (1 + 0.1 R) D^{\rm Gad}$ & $-$ & $-$ & $-$  & $-$ \\ 
+    & increased CSF dispersion & $-$ & $-$ & $ (1 + 10 R) D^{\rm Gad}$ & $-$ & $-$ & $-$  & $-$ \\ 
+    & higher PVS flow & $-$ & $-$ & $-$ & $-$ & $-$ & $\hat{u}_a^p + \hat{u}_a^c + \hat{u}_a^v$  & $-$ \\ 
+    & higher PVS exchange & $-$ & 100$\xi$  &  $-$ & $-$ & $-$ & $-$  & $-$ \\ 
+    & dilated PVS & $R_2 = 3 R_1$ & $-$ & $-$ & $-$ & $-$ & $\tilde{u}_a^p + \tilde{u}_a^c$  & $-$ \\ 
     \bottomrule
@@
     \end{center}
-  \caption{Overview of models (-- denotes as immediately above). Each
-    PVS cross-section is modelled as a concentric annulus with inner
-    radius $R_1$ and outer radius $R_2$ and such that $R_2 = \beta
-    R_1$ with $\beta = 2.0$ based on~\cite{eide2024functional}. Let
-    $\xi = \dots$ denote the permeability estimated by Koch et
-    al.~\cite{koch2023estimates}. Model A and its variations aims to
-    quantify the effect of PVS sheaths by varying the permeability
-    $\xi$ between the PVS and its surroundings. We consider a fixed
-    permeability $\xi_{\rm PVS-PAR}$ based
-    on~\cite{koch2023estimates}. \mer{The permeability between the PVS
-      and the SAS $\xi_{\rm PVS-SAS}$ must likely be estimated by a
-      bit of trial-error from the observations of timings of PVS-SAS
-      and SAS from~\cite{eide2024functional}.} \discuss{The
-      extracranial solute efflux permeability $\beta_{\rm efflux}$ is
-      uniformly distributed over the outer SAS boundary with a
-      reasonable rate as baseline\cite{hornkjol2022csf,
-        eide2021clinical, ringstad2024glymphatic}.} At baseline, the
-    effective diffusion coefficients $D^{\rm x}_{\rm SAS} = D^{\rm
-      x}_{\rm PVS}$ of the SAS and PVS represent the free diffusion
-    coefficient of Gadubutrol (${\rm x} = {\rm gad}$) in CSF (water),
-    and $D^{\rm x}_{\rm PAR}$ the effective diffusion coefficient of
-    human cortical tissue, all at body temperature; weighted by a
-    dispersion factor $\alpha$ (see Methods). At baseline, we consider
-    net flow due to CSF-production in the SAS and PVS:
-    $\mathbf{u}_{\rm SAS} > 0$ and $\hat{u} > 0$ (see Methods) but no
-    bulk flow in the parenchyma $\mathbf{u}_{\rm PAR} = 0$. Due to the
-    timescale for human intracranial tracer transport (minutes to
-    hours) compared to typical (cardiac, respiratory, slow vasomotion)
-    human intracranial pulsatility (0.1 -- 1 Hz), we consider net
-    (constant-in-time) flow fields only in subsequent model variations
-    (but see dispersion factor above).  Model B represents
-    perivascular pathway scenarios with more rapid flow in the PVS
-    induced by vasomotion (PVS I)~\cite{gjerde2023directional}.
-    Models C represents a version of the aforementioned scenarios but
-    for studying metabolite clearance (instead of solute influx).}
+  \caption{Overview of models and model variations. ($-$ denotes equal
+    to the corresponding baseline value). Each PVS cross-section is
+    modelled as a concentric annulus with inner radius $R_1$ and outer
+    radius $R_2$ and such that $R_2 = \beta R_1$. Let $\xi$ denote the
+    astrocytic endfeet permeability estimated by Koch et
+    al.~\cite{koch2023estimates}. The diffusion coefficients $D^{\rm
+      Gad}$ represent the (free) diffusion coefficient of Gadubutrol
+    in CSF (water), and $D^{\rm Gad}_{\rm PAR}$ its effective
+    diffusion coefficient in human cortical tissue, all at body
+    temperature~\cite{sykova2008diffusion, valnes2020apparent}. $R$ is
+    the combined cardiac and respiratory dispersion enhancement field
+    (see Methods). $\mathbf{u}_{\rm CSF}$ is the CSF velocity field
+    induced by CSF production in $\Omega_{\rm CSF}$, and $\hat{u}_a^p$
+    and $\hat{u}_v^p$ the axial velocity induced in the periarterial
+    and perivenous networks, respectively, by the corresponding
+    pressure differences. $\hat{u}_a^c$ and $\hat{u}_a^v$ is the
+    periarterial velocity induced by arterial pulse wave wall motion
+    and slow vasomotion, respectively. Last, $\tilde{u}_a^c$ and
+    $\tilde{u}_a^p$ are as $\hat{u}_a^c$ and $\hat{u}_a^p$,
+    respectively, but with dilated PVS.}
 \label{tab:scenarios}
@@
 
-\fi
+\subsection*{Numerical approximation, implementation, verification}
+
+A complete description of the numerical methods used is provided in
+\Cref{sec:details_numerical_method}, while a condensed summary is
+given here. To solve~\eqref{eq:multi_transport} numerically, we use
+the interior penalty discontinuous Galerkin (DG) method with weighted
+averages and upwinding for the convection term in the 3D
+domains~\cite{ern2009discontinuous} -- to ensure numerical stability
+and minimize the artificial presence of negative concentration
+values. For the 1D domain, we use continuous finite element spaces
+defined over each of the networks and stabilized with numerical
+diffusion when $\hat{u}$ is non--zero. Note that the 3D--1D coupling
+in~\eqref{eq:multi_transport} leads to a 3D solution $c$ with low regularity properties. Thus, one can only expect lower-order error convergence near
+$\Lambda_a$ and $\Lambda_v$, see \cite{masri2024discontinuous} for the numerical analysis of DG schemes and \cite{laurino2019derivation} for continuous Galerkin schemes.
+However, away from these 1D networks, almost
+optimal approximation is expected~\cite{masri2023discontinuous,koppl2016local}. To numerically solve the Stokes equations to
+compute the CSF flow velocity and pressure in the ventricular system
+and SAS, we use finite element spaces that preserve the
+incompressibility condition on the discrete
+level~\cite{hong2016robust}. This requirement is important for
+numerical stability when later
+solving~\eqref{eq:multi_transport} %\cite{cesmelioglu2022compatible}.
+ The
+spaces we used for the velocity are continuous along the normal
+direction; continuity of the tangential component is enforced via
+interior penalty approaches~\cite{hong2016robust}. Optimal
+approximation properties are expected for this scheme. These numerical
+methods were implemented using the FEniCS finite element
+framework~\cite{alnaes2015fenics}, and the 3D-1D coupling is handled
+by the extension
+FEniCS\textsubscript{ii}~\cite{kuchta2020assembly}. The correctness
+and accuracy of the numerical solutions were verified by a series of
+numerical verification experiments \fixme{(see \Cref{sec:app:verification})}.
+
+%% The choice this scheme was is necessary here since continuous Galerkin
+%% methods even when supplemented with artificial diffusion stabilization
+%% terms were unstable in our experiments.
+
+\subsection*{Model validation}
+
+As the primary means of model validation, we compare the in-silico
+predictions of tracer enrichment and clearance against glymphatic MRI
+studies (Discussion). In addition, we compare auxiliary model
+quantities (CSF flow rates and pressure differences, dispersion factor
+estimates, and the shapes and sizes of PVSs) with the current
+literature (Supplementary discussion: Extended model validation).
+
+\section*{Acknowledgments}
+
+All authors express their gratitude to Prof.~Jürgen Reichenbach for
+generously sharing the multi-modal human MR data, and to Prof. Erlend
+Hodneland for facilitating this data exchange. M.E.R. acknowledges the
+support of the Foundation Kristian Gerhard Jebsen via the K.~G.~Jebsen
+Center for Brain Fluid Research and the Research Council of Norway via
+grant \#324239 (EMIx). Simulations were performed on the Experimental
+Infrastructure for Exploration of Exascale Computing (eX3), which is
+financially supported by the Research Council of Norway under contract
+270053.
+
+\section*{Author contributions statement}
+
+M.C., M.K, R.M., and M.E.R. conceived and designed the project.
+M.C., M.K, R.M., and M.E.R. contributed to software development. 
+M.C., M.K, and R.M. conducted the experiments.
+M.C. analyzed the results and prepared the figures.
+M.C. M.K. and R.M and M.E.R wrote the manuscript. 
+All authors edited and reviewed the manuscript.
+
+\section*{Competing interests}
+
+The authors declare no competing interests.
 
@@
 
-\iffalse
 \newpage
-\section*{Acknowledgments (not compulsory)}
-
-Acknowledgments should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.
-
-\section*{Author contributions statement}
-
-Must include all authors, identified by initials, for example:
-A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.
-
-\section*{Additional information}
-
-To include, in this order: \textbf{Accession codes} (where applicable); \textbf{Competing interests} (mandatory statement).
-
-The corresponding author is responsible for submitting a \href{http://www.nature.com/srep/policies/index.html#competing}{competing interests statement} on behalf of all authors of the paper.
-This statement must be included in the submitted article file.
-
-\newpage
+\FloatBarrier
 \appendix
@@
 
-\input{background.tex}
-
-\fi
+%\newpage
+%\FloatBarrier
+%\input{background.tex}
 

### 2025-01-23 paper/supplementary_main.tex
Updates from Overleaf
Shuffle parts of the Discussion.

@@
 using finite element methods. The following subsections provide more
-details, model-by-model.
+details, model-by-model. \rami{Add a note on the convection dominated problem and the need for DG}
 
@@
 \end{figure}
+
+\section{Supplementary discussion}
+\label{sec:model_validation}
+
+\subsection{Extended model validation}
+\label{sec:model_validation}
+
+In addition to the comparison of our in-silico predictions of tracer
+enrichment and clearance against glymphatic MRI, we here compare
+auxiliary model quantities against the literature as additional model
+validation.
+
+\paragraph{CSF flow and pressures in the SAS and ventricular system}
+The dynamics of human CSF flow and pressure are better quantified, by
+way of clinical imaging, in-vitro studies, and computational
+modelling, in other areas of the ventricular
+system~\cite{linninger2007cerebrospinal, sweetman2011cerebrospinal,
+  vinje2019respiratory, hornkjol2022csf, causemann2022human,
+  karki2025real, liu2025transmantle}. Linninger et
+al~\cite{linninger2007cerebrospinal} model CSF flow and pressure
+dynamics induced by CSF production and cardiac pulsatility under
+normal and hydrocephalic conditions, and report of very good agreement
+with Cine (phase-contrast) MRI measurements. Our estimates of the
+maximum intracranial pressure difference, 10 Pa from the cardiac
+contribution and 26 mPa from CSF production, is in perfect agreement
+with their maximum transmantle pressure difference of $\sim$10 Pa, and
+also in very good agreement with mean pressure differences of 11.5 Pa
+measured clinically between sensors placed subdurally and in the
+lateral ventricle \cite{vinje2019respiratory}. Liu et
+al~\cite{liu2025transmantle} report of cardiac and respiratory
+pressure differences across the aqueduct of 12.1 $\pm$ 5.7 Pa and 9.5
+$\pm$ 7.2 Pa, respectively; thus our baseline estimate of the
+respiratory contribution (\fixme{1.4 Pa}) may be an
+underestimation. On the other hand, our cardiac- and
+respiratory-driven CSF flow estimates peak at 19.8 cm/s and 4.8 cm/s
+in the caudal direction, respectively, which are higher than
+phase-contrast MRI measurements of cardiac and respiratory CSF flow
+components \cite{takizawa2017characterization,
+  yildiz2017quantifying}. Some variation in CSF flow velocities is
+expected considering that the values from MRI represent averages
+\cite{yildiz2017quantifying} and that the geometry of the CSF spaces
+strongly affects peak velocities\cite{vinje2019respiratory,
+  causemann2022human}. Hornkjøl et al \cite{hornkjol2022csf} model the
+flow dynamics induced by CSF production in the choroid plexus and
+report a peak CSF velocity of 8.9 mm/s in the aqueduct, which is 4.8
+$\times$ higher than our values of 1.85 mm/s. Given that we use the
+same production rate, this deviation again illustrates the impact of
+potential differences in the (aqueduct) geometry on local velocities.
+
+% Dispersion predictions
+\paragraph{Dispersion in the SAS, ventricular system and PVS}
+This pulsatile flow of CSF in the SAS, ventricular system and PVSs
+leads to an increase in effective solute
+diffusivity~\cite{stockman2007effect, hettiarachchi2011effect,
+  asgari2016glymphatic, sharp2019dispersion, ray2021quantitative} via
+a process known as Taylor dispersion \cite{taylor1953dispersion,
+  watson1983diffusion}. Previous estimates of the magnitude of this
+effect in the CSF spaces vary significantly: from an enhancement
+factor of 0.05--1 in periarterial spaces surrounding penetrating
+arteries~\cite{asgari2016glymphatic, troyetsky2021dispersion}, to
+5--100 in the spinal subarachnoid space~\cite{stockman2007effect,
+  hettiarachchi2011effect, sharp2019dispersion}, and up to more than
+$10\,000$ in surface periarterial spaces~\cite{ray2021quantitative,
+  sharp2019dispersion}. The large variability can (at least partly) be
+attributed to methodological differences; e.g. different assumptions
+on the medium, domain width, pressure differences and/or fluid
+velocities, the diversity of CSF flow characteristics, as well as a
+high likelihood of spatial variations. Hornkjøl et
+al~\cite{hornkjol2022csf} consider model variations with constant
+dispersion factors from 1 up to 1000, and indicate that a value of 10
+gives the better agreement with the clinically observed
+enrichment. Our spatially-varying estimates of the dispersion
+enhancement factors $R_c, R_r$ (with $D = (1 + R_c + R_r) D^{\rm
+  Gad}$) range from 0 to 200 for the cardiac contribution $R_c$ and 0
+to 320 for the respiratory contribution $R_r$; and is thus compatible
+within the previously reported spectrum.
+
+% Effect of PVS size and shapes
+\paragraph{Shapes, sizes and structures of the PVS}
+The shapes, sizes and structures of the PVSs likely vary between
+species (e.g.~mice vs.~humans), between spatial compartments
+(e.g.~surface vs.~parenchymal), between vessel types (arteries
+vs.~arterioles vs.~veins), and in
+pathologies~\cite{ichimura1991distribution, foley2012realtime,
+  schain2017cortical, mestre2018flow, bedussi2018paravascular,
+  mestre2022periarteriolar, smets2024perivascular, raicevic2023sizes,
+  vinje2021brain, eide2024functional}. In terms of shape, the PVSs are
+commonly represented as annular (elliptic) cylinders, though it is
+well recognized that this represents an
+idealization~\cite{mestre2018flow, tithof2019hydraulic,
+  vinje2021brain, raicevic2023sizes, boster2024hydraulic,
+  smets2024perivascular}. In terms of sizes, Raicevic et
+al~\cite{raicevic2023sizes} note that the variation in PVS area is
+larger between PVS segments than along a single PVS segment and that
+the PVS area increases with lumen area. In mice, reports of the ratio
+between PVS and lumen area range from
+$\approx$0.35--0.43~\cite{smets2024perivascular} up to
+$\approx$1.12--1.4~\cite{raicevic2023sizes, mestre2018flow}. In
+humans, the PVS may be as wide as the associated surface artery and up
+to 4$\times$ wider in iNPH subjects~\cite{eide2024functional}, which
+would correspond to substantially larger PVS area ratios (3 or
+higher). To reflect the human scale, we here represent each PVS
+segment as an annular cylinder with inner radius $R_1$ and outer
+radius $R_2$ of width and area proportional to that of the
+corresponding blood vessel ($R_2 = 2 R_1$ at baseline, $R_2 = 3 R_1$
+for enlarged PVS). The hydraulic resistance of annular cross-sections
+is $1-6 \times$~\cite{tithof2019hydraulic} larger than more elongated
+cross-sections and thus our estimates of the pressure-induced PVS
+velocities are conservative.

### 2025-01-23 paper/main.tex
Numerical in main text

@@
 junction. While the resulting flow velocities are only about one
-fourth of their cardiac-induced counterparts (see \Cref{fig:csf}H, L),
+fourth of their cardiac-induced counterparts (\Cref{fig:csf}H, L),
 respiratory dispersive mixing reaches a factor of up to 320 due the
@@
 area of the PVS segment and $\langle Q_i' \rangle$ is the estimated
-mean flow rate (see Supplementary Methods~\ref{sec:app:peristalsis}
-for more details). Two wall motion patterns are considered, yielding
-$\hat{u}_a^c$ and $\hat{u}_a^v$ respectively: \emph{cardiac}
-pulsations with $f = 1.0$Hz, $\lambda = 2$ m and $\varepsilon = 1\%$
-(\Cref{fig:pvs}E), and very low-frequency \emph{vasomotion} with $f =
-0.1$Hz, $\lambda = 0.02$m, and $\varepsilon = 10\%$
-(\Cref{fig:pvs}F). No flow induced by peristalsis is considered for
-the perivenous network. In the baseline model, we combine the
-contributions from CSF production and cardiac peristalsis to set
-$\hat{u}$ in~\eqref{eq:multi_transport} as $\hat{u}_a = \hat{u}_a^p +
-\hat{u}_a^c$ in $\Lambda_a$, and $\hat{u}_v = \hat{u}_v^p$ in
-$\Lambda_v$.
+mean flow rate (see~\ref{sec:app:peristalsis} for more details). Two
+wall motion patterns are considered, yielding $\hat{u}_a^c$ and
+$\hat{u}_a^v$ respectively: \emph{cardiac} pulsations with $f =
+1.0$Hz, $\lambda = 2$ m and $\varepsilon = 1\%$ (\Cref{fig:pvs}E), and
+very low-frequency \emph{vasomotion} with $f = 0.1$Hz, $\lambda =
+0.02$m, and $\varepsilon = 10\%$ (\Cref{fig:pvs}F). No flow induced by
+peristalsis is considered for the perivenous network. In the baseline
+model, we combine the contributions from CSF production and cardiac
+peristalsis to set $\hat{u}$ in~\eqref{eq:multi_transport} as
+$\hat{u}_a = \hat{u}_a^p + \hat{u}_a^c$ in $\Lambda_a$, and $\hat{u}_v
+= \hat{u}_v^p$ in $\Lambda_v$.
 
@@
 
-\subsection*{Numerical approximation, implementation, verification}
-
-A complete description of the numerical methods used is provided in
-\Cref{sec:details_numerical_method}, while a condensed summary is
-given here. To solve~\eqref{eq:multi_transport} numerically, we use
-the interior penalty discontinuous Galerkin (DG) method with weighted
-averages and upwinding for the convection term in the 3D
+\subsection*{Numerical approximation, implementation, and verification}
+
+We give a concise summary of the numerical approach here (see
+Supplementary methods for a comprehensive description). To
+solve~\eqref{eq:multi_transport} numerically, we use the interior
+penalty discontinuous Galerkin (DG) method with weighted averages and
+upwinding for the convection term in the 3D
 domains~\cite{ern2009discontinuous} -- to ensure numerical stability
@@
 diffusion when $\hat{u}$ is non--zero. Note that the 3D--1D coupling
-in~\eqref{eq:multi_transport} leads to a 3D solution $c$ with low regularity properties. Thus, one can only expect lower-order error convergence near
-$\Lambda_a$ and $\Lambda_v$, see \cite{masri2024discontinuous} for the numerical analysis of DG schemes and \cite{laurino2019derivation} for continuous Galerkin schemes.
-However, away from these 1D networks, almost
-optimal approximation is expected~\cite{masri2023discontinuous,koppl2016local}. To numerically solve the Stokes equations to
-compute the CSF flow velocity and pressure in the ventricular system
-and SAS, we use finite element spaces that preserve the
-incompressibility condition on the discrete
+in~\eqref{eq:multi_transport} leads to a 3D solution $c$ with low
+regularity properties. Thus, one can only expect lower-order error
+convergence near $\Lambda_a$ and $\Lambda_v$, see the references
+\cite{masri2024discontinuous} and \cite{laurino2019derivation} for the
+numerical analysis of DG and continuous Galerkin schemes,
+respectively. However, away from these 1D networks, almost optimal
+approximation is
+expected~\cite{masri2023discontinuous,koppl2016local}. To numerically
+solve the Stokes equations to compute the CSF flow velocity and
+pressure in the ventricular system and SAS, we use finite element
+spaces that preserve the incompressibility condition on the discrete
 level~\cite{hong2016robust}. This requirement is important for
-numerical stability when later
-solving~\eqref{eq:multi_transport} %\cite{cesmelioglu2022compatible}.
- The
-spaces we used for the velocity are continuous along the normal
+numerical stability when subsequently
+solving~\eqref{eq:multi_transport}\cite{cesmelioglu2022compatible}.
+The spaces we used for the velocity are continuous along the normal
 direction; continuity of the tangential component is enforced via
@@
 and accuracy of the numerical solutions were verified by a series of
-numerical verification experiments \fixme{(see \Cref{sec:app:verification})}.
+numerical verification experiments (\Cref{sec:app:verification}).
 
@@
 estimates, and the shapes and sizes of PVSs) with the current
-literature (Supplementary discussion: Extended model validation).
+literature (\Cref{sec:app:model_validation}).
 

### 2025-01-23 paper/supplementary_main.tex
Numerical in main text

@@
 \subsubsection{Modelling CSF flow via the incompressible Stokes equations}
-\label{sec:csf_flow}
+\label{sec:app:csf_flow}
 The time-independent, incompressible Stokes equations model the flow
@@
 \subsection{Numerical solution of the transport and flow equations}
-\label{sec:details_numerical_method}
+\label{sec:app:details_numerical_method}
 
@@
 \subsection{Numerical verification}
+\label{sec:app:verification}
 
@@
 \section{Supplementary discussion}
-\label{sec:model_validation}
 
 \subsection{Extended model validation}
-\label{sec:model_validation}
+\label{sec:app:model_validation}
 

### 2025-01-23 paper/supplementary_main.tex
Updates in said caption. Spell-check on supplementary.

@@
 Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$.
-$\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivacular networks $\Lambda_a, \Lambda_v$.
+$\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivascular networks $\Lambda_a, \Lambda_v$.
 To summarize,
@@
 
-Finally, we impose a homogogenous Neumann condition at all network end
+Finally, we impose a homogeneous Neumann condition at all network end
 nodes to augment~\eqref{eq:1d_pde}:
@@
 between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
-pointing from $E^1$ to $E^2$. The facet diamter is denoted by $h_F$,
+pointing from $E^1$ to $E^2$. The facet diameter is denoted by $h_F$,
 the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
@@
 
- \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} comlemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0 = 0 $ is the following:
+ \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} complemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0 = 0 $ is the following:
 
@@
 associated with the nearest supply node, and edges between nodes are
-preserved. Next, we compute a minimal, bifurcating and directed tree
-representation of each subnetwork: $\mathcal{T}^1, \mathcal{T}^2,
-\mathcal{T}^3$. Each tree $\mathcal{T}^j = \cup E_n^j$ consists of the
-subset of the nodes from $\Lambda^j$ that have degree 1 (are leaf or
-root nodes) or degree $3$ (are true bifurcation points), and each path
-between nodes with degree $\not = 2$ in $\Lambda^j$ is represented in
-$\mathcal{T}^j$ by an edge $E^j$ with edge length $L$ corresponding to
-the total length of the original path and edge radius $R^j$ as the
-average of the path radii.
+preserved (\Cref{fig:app:peristalsis}B). Next, we compute a minimal,
+bifurcating and directed tree representation of each subnetwork:
+$\mathcal{T}^1, \mathcal{T}^2, \mathcal{T}^3$
+(\Cref{fig:app:peristalsis}C). Each tree $\mathcal{T}^j = \cup E_n^j$
+consists of the subset of the nodes from $\Lambda^j$ that have degree
+1 (are leaf or root nodes) or degree $3$ (are true bifurcation
+points), and each path between nodes with degree $\not = 2$ in
+$\Lambda^j$ is represented in $\mathcal{T}^j$ by an edge $E^j$ with
+edge length $L$ corresponding to the total length of the original path
+and edge radius $R^j$ as the average of the path radii.
 
@@
 the vascular wall motion for each edge $n$ via~\cite[eq.~(5),
-  (34)]{gjerde2023directional}. We next uniquely assign the flow rate
-$\langle Q'_n \rangle$ to each of the branches $\Lambda_i$ that form
-the path $E_n^j$, thus yielding $\langle Q'_i \rangle$ for each
-perivascular segment $\Lambda_i$ (except those ignored above, for
-which we set a flow rate of zero). Finally, we define the mean
-longitudinal perivascular velocity induced by the peristaltic wave
-$\langle \hat{u}^x_i \rangle$ by dividing the flow rate $\langle Q'_i
-\rangle$ by the cross-section area $A_i = \pi ({R_2^i}^2 - {R_1^i}^2)
-= \pi (\beta_i^2 - 1) {R_1^i}^2$.
+  (34)]{gjerde2023directional} (\Cref{fig:app:peristalsis}D). We next
+assign this flow rate $\langle Q'_n \rangle$ to each of the branches
+$\Lambda_i$ that form the path $E_n^j$, thus yielding $\langle Q'_i
+\rangle$ for each perivascular segment $\Lambda_i$ while ensuring that
+mass is conserved (\Cref{fig:app:peristalsis}E). For the segment(s)
+ignored in the separation step (\Cref{fig:app:peristalsis}B), we set a
+flow rate of zero. Finally, we define the mean longitudinal
+perivascular velocity induced by the peristaltic wave $\langle
+\hat{u}^x_i \rangle$ by dividing the flow rate $\langle Q'_i \rangle$
+by the cross-section area $A_i = \pi ({R_2^i}^2 - {R_1^i}^2) = \pi
+(\beta_i^2 - 1) {R_1^i}^2$.
 
@@
 Specifically, we generate a sequence of three meshes, see \Cref{fig:mesh_refinement}, with an increasing number of points and cells, solve all relevant simulation steps (CSF flow and intracranial transport computations) for standard model A, and compare the results with respect to a set of key quantities of interest.
-The mesh refinement employed here is adapative in the sense that the coarser mesh is refined near expected sharp concentration gradients.
+The mesh refinement employed here is adaptive in the sense that the coarser mesh is refined near expected sharp concentration gradients.
 This allows us to better capture dynamics and reduce undershoots, see \Cref{fig:mesh_convergence_concentrations}.
@@
     \end{subfigure}
-    \caption{Minumum and maximum tracer concentrations over the first 24\,h after injection on the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the LowRes, standard and HighRes meshes with a timestep of $dt=120$\,s.}
+    \caption{Minimum and maximum tracer concentrations over the first 24\,h after injection on the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the LowRes, standard and HighRes meshes with a timestep of $dt=120$\,s.}
     \label{fig:mesh_convergence_concentrations}

### 2025-01-23 paper/main.tex
Tweak

@@
 \affil[2]{Brown University, Providence, Rhode Island, USA}
-\affil[3]{K. G. Jebsen Centre for Brain Fluid Research}
+\affil[3]{K. G. Jebsen Centre for Brain Fluid Research, University of Oslo, Norway}
 \affil[*]{meg@simula.no}
@@
 human brain environment such as the role of pulsatility, perivascular
-pathways, structural compartmentalization, or morphological
-pathologies.
+pathways, structural compartmentalization, or morphology.
 

### 2025-01-23 paper/main.tex
A paragraph for the Discussion

@@
 the in-silico concentrations relative to the total amount of tracer.
+
+In conclusion, our findings transfer insights from experimental
+studies and theoretical analysis to the in-silico human setting,
+reconcile seemingly conflicting observations in particular relating to
+directionality of perivascular flow, and integrate different physical
+mechanisms across across spatial and temporal scales. The complete
+simulation pipeline is openly available, including interactive
+visualization of simulation results for all model
+variations~\cite{ZENODO}. Looking ahead, this platform establishes a
+foundation for in-silico studies of molecular movement within the
+human brain environment, such as tailored predictions of intrathecal
+chemotherapy delivery or personalized diagnostics of brain clearance
+capacity.
 

### 2025-01-24 scripts/make_html.py
interactive html

@@ -1,0 +1,94 @@
+from plotting_utils import get_result, read_config
+import pyvista as pv
+import typer
+import numpy as np
+from pathlib import Path
+from tqdm import tqdm
+from extract_vessels import get_tubes
+import k3d.colormaps.paraview_color_maps as pcm
+import k3d
+import matplotlib
+import skimage
+
+hexcolor = lambda c: int(matplotlib.colors.to_hex(c)[1:], base=16)
+compression_level = 6
+
+def make_html(modelname:str):
+    N = 300
+    config = read_config(f"configfiles/{modelname}.yml")
+    dt , T= config["dt"], config["T"]
+    clim = (0,2)
+    times = (np.array([0, 1/3, 2/3, 1,2, 3,4,5,6,8,10,12,15,18,24])*3600).astype(np.int32)
+    meshfile = Path(config["mesh"])
+    skull = pv.read(meshfile.parent / "surfaces/skull.ply")
+    LV = pv.read(meshfile.parent / "surfaces/LV.ply")
+    V34 = pv.read(meshfile.parent / "surfaces/V34.ply")
+    parenchyma = pv.read(meshfile.parent / "surfaces/parenchyma_incl_ventr.ply")
+    art = get_result(modelname, "artery", times).ctp()
+    sas = get_result(modelname, "sas", times)
+    arteries = get_tubes(art)
+    arteries["c"] = arteries["c_0"]
+    for t in times:
+        sas[f"c_{t}"] = sas[f"c_{t}"].clip(min=clim[0]).astype(np.float32)
+    
+    #from IPython import embed; embed()
+        
+    img = pv.create_grid(sas, dimensions=(N, N, N))
+    img = img.sample(sas, progress_bar=True)
+    for t in times:
+        img[f"c_{t}"] = skimage.filters.gaussian(img[f"c_{t}"], sigma=1, truncate=2)
+
+    pl = k3d.plot(
+        camera_rotate_speed=3,
+        camera_zoom_speed=5,
+        background_color=hexcolor("black"),
+        grid_visible=False,
+        camera_auto_fit=True,
+        axes_helper=False,
+        lighting=3,
+        )
+
+    vol_dtype = np.float16
+    cmap = np.array(pcm.Yellow___Gray___Blue, dtype=vol_dtype).reshape(-1,4)[:,:] 
+    cmap[:,0] = np.linspace(-1, 1, cmap.shape[0])[::-1]
[…]

### 2025-01-24 scripts/plot_csf_flow.py
add mean flow at spinal outlet

@@ -8,7 +8,7 @@ from fenics import *
 import matplotlib 
 import k3d.colormaps.paraview_color_maps as pcm
 from compute_dispersion_field import alpha_cardiac, alpha_respiratory
-from plotting_utils import set_plotting_defaults, minmax
+from plotting_utils import set_plotting_defaults, read_config
 from subdomain_ids import CSFID, LVID, V34ID
 import matplotlib.pyplot as plt
 from matplotlib.ticker import PercentFormatter
@@ -16,6 +16,8 @@ import seaborn as sns
 from plot_subdomains import get_camera
 from vtk.util.numpy_support import numpy_to_vtk
 import seaborn as sns
+from pathlib import Path
+from mark_and_refine_mesh import SPINAL_OUTLET_ID
 
 m2mum = 1e6
 
@@ -90,6 +92,18 @@ def plot_csf_flow(dirname: str):
         f.read(sm, "label")
     p = interpolate(p, FunctionSpace(mesh, "DG", v_elem.degree()))
 
+    flow_model = Path(dirname).name
+    flow_config = read_config(f"configfiles/{flow_model}.yml")
+    bm = MeshFunction("size_t", mesh, 2, 0)
+    with XDMFFile(flow_config["mesh"].replace(".xdmf","_facets.xdmf")) as f:
+        f.read(bm, "f")
+
+    ds = Measure("ds", mesh, subdomain_data=bm)
+    n = FacetNormal(mesh)
+    fm_area = assemble(Constant(1)*ds(SPINAL_OUTLET_ID))
+    fm_outflow = assemble(inner(v, n)*ds(SPINAL_OUTLET_ID))
+    fm_mean_velocity = fm_outflow/ fm_area
+    print(fm_mean_velocity)
     topology, cell_types, x = create_vtk_structures(V)
     grid = pv.UnstructuredGrid(topology, cell_types, x)
     grid["v"] = v.vector()[:].reshape(-1, 3)

### 2025-01-24 paper/supplementary_main.tex

@@
 
-We assess the accuracy and numerical convergence of our simulation results by performing a series of experiments with different spatial and temporal resolutions.
-Specifically, we generate a sequence of three meshes, see \Cref{fig:mesh_refinement}, with an increasing number of points and cells, solve all relevant simulation steps (CSF flow and intracranial transport computations) for standard model A, and compare the results with respect to a set of key quantities of interest.
+We assess the numerical accuracy and convergence of our simulation results by performing a series of experiments with different spatial and temporal resolutions.
+Specifically, we generate a sequence of three meshes (\Cref{fig:mesh_refinement}) with an increasing number of computational mesh vertices and cells, solve all relevant simulation steps (CSF flow and intracranial transport computations) for the baseline model on each mesh, and compare the results across meshes with respect to a set of key quantities of interest.
 The mesh refinement employed here is adaptive in the sense that the coarser mesh is refined near expected sharp concentration gradients.
 This allows us to better capture dynamics and reduce undershoots, see \Cref{fig:mesh_convergence_concentrations}.
-Similarly, we investigate the effect of the time step size by solving the intracranial transport model with varying time steps $dt \in \{60, 120, 240 \}$\,s.
+Similarly, we investigate the effect of the time step size by solving the intracranial transport model with different time steps: 1, 2, and 4 minutes.
 
-Considering the mean tracer concentrations in the CSF, parenchyma and arterial and venous PVS domain over the first 24\,h after injection, we observe negligible changes with both mesh and time refinement (\cref{fig:mesh_convergence_concentrations} and
-\ref{fig:time_convergence_concentrations}).
-In addition, we compute the mean and maximum enhancement factor, the maximum CSF pressure and velocity in both the cardiac-driven and CSF production-induced flow fields, and mean concentrations at 3, 6, 12 and 24\,h for all mesh resolutions and time steps (\cref{fig:mesh_convergence} and \cref{fig:time_convergence}).
-While the maximum dispersion factor increases by about 60\,\% from the coarse to the standard mesh, it stabilizes with the next refinement step.
-All other quantities change with less than 10\,\% with mesh refinement, and less than 1\,\% with time refinement.
-We thus conclude that the standard resolution mesh and a time step of $dt=120\,$s offers sufficient accuracy for our simulations.
-In addition, we compute the standard model without allowing tracer transport across the models outer boundary, confirming mass conservation of our numerical scheme (\cref{fig:mass_conservation}).
+Considering the mean tracer concentrations in the CSF, parenchyma and
+arterial and venous PVS domain over the first 24\,h after injection,
+we observe negligible changes with both mesh and time refinement
+(\Cref{fig:mesh_convergence_concentrations} and
+\Cref{fig:time_convergence_concentrations}). As an additional
+verification step, we compute the mean and maximum dispersion
+enhancement factor, the maximum CSF pressure and velocity in both the
+cardiac-driven and CSF production-induced flow fields, and mean
+concentrations at 3, 6, 12 and 24\,h for all mesh resolutions and time
+steps (\Cref{fig:mesh_time_convergence}). While the maximum dispersion factor
+increases by about 60\,\% from the low resolution to the standard mesh, it
+stabilizes with the next refinement step. All other quantities change
+with less than 10\,\% with mesh refinement, and less than 1\,\% with
+time refinement. We thus conclude that the standard resolution mesh
+and a time step of 2 min offer sufficient accuracy for our
+simulations.
 
+Finally, to check for numerical mass conservation, we perform an
+additional simulation not allowing for tracer efflux across the outer
+boundary, and confirm that the total amount of tracer is preserved
+over time after the initial influx phase
+(\Cref{fig:mass_conservation}).
 \begin{figure}
@@
     \includegraphics[width=0.9\linewidth]{figures/mesh_refinement.png}
-    \caption{Illustration of the three different meshes: low resolution, standard resolution, high resolution (from left to right).}
+    \caption{Illustration of the three different meshes; from left to right: low resolution, standard resolution, high resolution.}
     \label{fig:mesh_refinement}
@@
     \end{subfigure}
-    \caption{Minimum and maximum tracer concentrations over the first 24\,h after injection on the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the LowRes, standard and HighRes meshes with a timestep of $dt=120$\,s.}
+    \caption{Minimum and maximum tracer concentrations over the first 24\,h after injection on the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the low resolution (LowRes), standard and high resolution (HighRes) meshes with a timestep of 2 min for the baseline model (Model A).}
     \label{fig:mesh_convergence_concentrations}
 \end{figure}
-
 \begin{figure}
@@
         \includegraphics[trim={0.5cm 1cm 0.05cm 0.8cm}, clip, width = 1.05 \linewidth]{figures/modelALowRes_modelA_modelAHighRes.pdf}
-        \caption{Convergence of key quantities of interest with mesh refinement}
-        \label{fig:mesh_convergence}
+        \caption*{Convergence of key quantities of interest with mesh refinement}
+        %\label{fig:mesh_convergence}
     \end{subfigure}
@@
      \includegraphics[trim={0.5cm 1cm 0.05cm 0.8cm}, clip,width= 1.05 \linewidth]{figures/modelAsmalldt_modelA_modelAlargedt.pdf}
-        \caption{Convergence of key quantities of interest with time refinement}
-        \label{fig:time_convergence}
+        \caption*{Convergence of key quantities of interest with time refinement}
+        %\label{fig:time_convergence}
     \end{subfigure}
-    \caption{minimal ($\rm h_{min}$) and maximal ($\rm h_{max}$) mesh cell sizes (computed as cell circumradius $\times 2$) (A); number of points and tetrahedral cells in each mesh (B); mean cardiac dispersion enhancement factor $R$ (C), maximum cardiac dispersion enhancement factor $R$ (D); maximum pressure in steady CSF production flow (E), maximum pressure in cardiac-driven CSF flow (F); maximum CSF velocity in steady CSF production flow (G); maximum CSF velocity in cardiac-driven CSF flow (H);  mean tracer concentration in the CSF, parenchyma, arterial PVS and venous PVS at 3,6,12 and 24\,h (I, J, K and L).}
+    \caption{For both left and right panels: A: Minimal ($\rm h_{min}$) and maximal ($\rm h_{max}$) mesh cell sizes (computed as cell circumradius $\times 2$); B: number of mesh vertices and tetrahedral cells in each mesh; C: mean cardiac dispersion enhancement factor $R$; D: maximum cardiac dispersion enhancement factor $R$; E: maximum pressure in steady CSF production flow; F: maximum pressure in cardiac-driven CSF flow; G: maximum CSF velocity in steady CSF production flow; H: maximum CSF velocity in cardiac-driven CSF flow; I--L: mean tracer concentration in the CSF, parenchyma, arterial PVS and venous PVS after 3, 6, 12 and 24 hours.}
+        \label{fig:mesh_time_convergence}
 \end{figure}
@@
     \end{subfigure}
-    \caption{Mean tracer concentrations over the first 24\,h after injection on the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the standard resolution mesh for timesteps $dt \in \{60, 120, 240 \}$\,s.}
+    \caption{Mean tracer concentrations after up to 24\,h in the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the standard resolution mesh for timesteps of 1, 2, and 4 minutes (dt of 60, 120, or 240 seconds)}
     \label{fig:time_convergence_concentrations}
 \end{figure}
-
 \begin{figure}
@@
     \includegraphics[width=0.5\linewidth]{figures/modelAMassConservation_total_conc.png}
-    \caption{total tracer content in CSF, parenchyma and arterial and venous PVS for the standard model without tracer outflow, illustrating numerical mass conservation.}
+    \caption{Total tracer content in the CSF, parenchyma, and arterial and venous PVS for a variant of the baseline model without tracer outflow.
+The total amount of tracer is constant after the initial influx phase demonstrating that the numerical scheme conserves mass.}
     \label{fig:mass_conservation}

### 2025-01-24 paper/supplementary_main.tex
Updates from Overleaf

@@
 periarterial network $\Lambda_a$ represented by the (connected)
-centerlines $\Lambda^i_{(a)}$ of the arterial tree, and a perivenous
-network $\Lambda_v$ associated with the centerlines $\Lambda^i_{(v)}$
+centerlines $\Lambda^i_{a}$ of the arterial tree, and a perivenous
+network $\Lambda_v$ associated with the centerlines $\Lambda^i_{v}$
 of the veins. Recall that we represent the vascular domains as the
@@
 graph with the centerlines $\{ \Lambda^i \}$ as oriented edges and the
-connections as nodes $\mathcal{V}$.
+connections as nodes $\mathcal{V}$. \mk{We always keep $R^i_2$, $R^i_1$ fixed in time in A.1.1 and A.1.4, right?}
 
@@
 \end{equation}
-where the function $n^i : \mathcal{V} \rightarrow \{-1, 0, 1\}$
+where the function $n^i : \mathcal{V} \rightarrow \{-1, \mk{0}, 1\}$
 defines an orientation of each vertex relative to the
@@
   n^i(y_{\rm{out}}) = -1, \quad
-  n^i(y) = 0 \quad \foralls y \in 
-  \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} .  
+  \mk{n^i(y) = 0 \quad \foralls y \in 
+  \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} }.  
   \label{eq:def:sign}
 \end{equation}
+\mk{Why do we need the normal on non-IO vertices?}
 
@@
     \label{eq:momentum_equation} \\ 
-    \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF},
+    \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF}
     \label{eq:divergence_equation} .
@@
 a discontinuous Galerkin (DG) finite element method with upwinding in
-space to accurately capture sharp boundary layers.
+space to accurately capture sharp boundary layers. \mk{Comment on Peclet number}
 
@@
   a_h^{\mathrm{up}}(c, v)
-  = - \sum_{E \in \mathcal{T}} \int_E \bm{u} c \nabla v
+  = - \sum_{E \in \mathcal{T}} \int_E c\bm{u} \cdot \nabla v
   + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
@@
 \end{align*}
-Here, $H(\mathrm{div}, \Omega_{\rm{CSF}})$ is the space of $L^2(\Omega_{\rm{CSF}})$ vector fields with $L^2(\Omega_{\rm{CSF}})$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, and $\bm n$ and $\bm \tau$ are the unit outward normal and unit tangent vectors to each facet, respectively.
+Here, $H(\mathrm{div}, \Omega_{\rm{CSF}})$ is the space of $L^2(\Omega_{\rm{CSF}})$ vector fields with $L^2(\Omega_{\rm{CSF}})$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, and $\bm n$ is the unit outward normal vector to each facet.
 Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by
@@
 \begin{align}
-  \hat{Z}_h(\mathcal{I}) &= \{z \in L^2(\Lambda) \, | \, z \vert_{\Lambda_i} \in \mathrm{P}^{0}(\Lambda_i) \, \foralls \Lambda^i \in \mathcal{I}\}, \\
-  \hat{V}_{h, g}(\mathcal{I}) &= \{ v \in C^0(\Lambda) \, | \, v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda_i), \, v(x) = g(x) \, \foralls x \in \mathcal{N} \} .
+  \hat{Z}_h(\mathcal{I}) &= \{z \in L^2(\Lambda) ; \, z \vert_{\Lambda_i} \in \mathrm{P}^{0}(\Lambda_i) \, \foralls \Lambda^i \in \mathcal{I}\}, \\
+  \hat{V}_{h, g}(\mathcal{I}) &= \{ v \in C^0(\Lambda) ; \, v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda_i), \, v(x) = g(x) \, \foralls x \in \mathcal{N} \} .
   %Q_h^{\rm PVS} &= \{ q_h \in C^0(\Lambda),  q_h \vert_{\Lambda^i} \in \mathrm{P}^k(\Lambda_i), q_h \vert_{\Gamma_{\rm Pia}} = p_{\rm CSF} \}
@@
 By definition, the wave number is $k = 2 \pi/\lambda$ and the angular frequency is $\omega = 2 \pi f$.
-Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L$, baseline inner radius $R_1^i$, fixed outer radius $R_2^i$, and outer-to-inner ratio $\beta_i = R_2^i/R_1^i$.
+Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L$, \mk{Should the $L$ here be $L_i$?} baseline inner radius $R_1^i$, fixed outer radius $R_2^i$, and outer-to-inner ratio $\beta_i = R_2^i/R_1^i$.
 These geometric parameters and the assumption of annular cylindrical PVS segments yield hydraulic resistances $\mathcal{R}_{o, i}$ and additional characteristic parameters, see \cite{gjerde2023directional} for the complete definitions and schematics.
@@
     \caption{Total tracer content in the CSF, parenchyma, and arterial and venous PVS for a variant of the baseline model without tracer outflow.
-The total amount of tracer is constant after the initial influx phase demonstrating that the numerical scheme conserves mass.}
+The total amount of tracer is constant after the initial influx phase demonstrating that the numerical scheme conserves mass globally.}
     \label{fig:mass_conservation}

### 2025-01-24 scripts/time_dependent.py
add peclet number

@@ -196,6 +196,18 @@ def run_simulation(configfile: str):
         R = map_dg_on_global(R, parentmesh=mesh)
         Ds *= (1 + R)
 
+    dx_s = Measure("dx", mesh, subdomain_data=vol_subdomains)
+    DG0 = FunctionSpace(mesh, "DG", 0)
+    csf_dofs_dg = get_subdomain_dofs(DG0, vol_subdomains, CSFID)
+
+    # compute peclet number
+    L = 0.05 # characteristic length (along direction of primary transport)
+    Pe = project(L*sqrt(inner(velocity_csf, velocity_csf)) / Ds, DG0,
+                 solver_type="cg",preconditioner_type="jacobi")
+    pemax = Pe.vector().get_local()[csf_dofs_dg].max()
+    pemin = Pe.vector().get_local()[csf_dofs_dg].min()
+    peavg = assemble(Pe*dx_s(CSFID)) / assemble(Constant(1)*dx_s(CSFID))
+    print(f"min Pe: {pemin}, max Pe: {pemax}, avg Pe: {peavg}")
 
     fa = Constant(0.0) 
     fv = Constant(0.0)
@@ -231,7 +243,6 @@ def run_simulation(configfile: str):
 
     eta = Constant(1e3)
 
-    dx_s = Measure("dx", mesh, subdomain_data=vol_subdomains)
     dS = Measure("dS", mesh, subdomain_data=bm)
 
     def a_dg_adv_diff(u,v) :

### 2025-01-24 paper/main.tex

@@
 \newcommand{\draft}[1]{\textcolor{lightgray}{#1}}
-%\newcommand{\brain}{\Omega_{\rm{PAR}}} 
-%\newcommand{\sas}{\Omega_{\rm CSF}}
-%\newcommand{\spinal}{\Gamma_{\rm SAS}}
-%\newcommand{\skull}{\Gamma_{\rm{skull}}}
-%\newcommand{\gin}{g_{\rm{influx}}}
 
@@
 \newcommand{\foralls}{\forall \,}
-
-%\affil[+]{these authors contributed equally to this work}
-
-%\keywords{Keyword1, Keyword2, Keyword3}
-
-%% The problem we are addressing here is that we would like to be able to
-%% understand, integrate and quantify the role of different mechanisms -
-%% including effects associated with individual variations - on molecular
-%% enrichment and clearance of human intracranial spaces. And we think of
-%% this as a step on the way to predict tracer enrichment after
-%% intrathecal injection in a personalized medicine context. Which is
-%% important because of drug delivery/treatment in cancer, diagnostics
-%% for CSF and dementia disorders, and clearance in neurodegenerative
-%% disease.
 
@@
 \maketitle
-% * <john.hammersley@gmail.com> 2015-02-09T12:07:31.197Z:
-%
-%  Click the title above to edit the author information and abstract
-%
+
 \thispagestyle{empty}
@@
 %\mer{MER: Target length: 5000 words (Intro, Results, Discussion) plus <3000 words Methods: 700 words Introduction. 3100 words Results. 1200 words Discussion.}
-
-\iffalse
-\paragraph*{Abbreviations}
-\begin{itemize}
-\item
-  Subarachnoid space (SAS)
-\item
-  Cerebrospinal fluid (CSF)
-\item
-  Perivascular space (PVS), perivascular spaces (PVSs)
-\end{itemize}
-\fi
 
@@
 
-%% \begin{itemize}
-%% \item
-%%   Need better understanding of transport of solutes in intracranial spaces.
-%% \item
-%%   Need to integrate separate observations e.g. on diffusion in
-%%   neuropil, PVS flow, net flow, effect of dispersion
-%% \item
-%%  Need to transfer and integrate insights from animal models,
-%%   computational models and theoretical insights into the human case. 
-%% \item
-%%  Controversies need to be resolved.
-%% \end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Results}
-%% \mer{MER: Target ~3100 words in 6 subsections}
-
-%% \mer{MER: Results #1 is about 500 words}
+
 \subsection*{In-silico predictions of intracranial molecular enrichment and clearance after intrathecal injection}
@@
 
-% \mer{MER: Results #2 is about 600 words}
 \subsection*{Reduced CSF pulsatility strongly shifts intracranial enrichment patterns}
@@
 
-% \mer{MER: Results #3 is about 800 words}
 \subsection*{Perivascular flow shapes and accelerates molecular enrichment}
@@
 scale. For instance, after 4-6 hours, the faster-moving tracer in the
-PVS is clearly visible in the adjacent spaces to the ACA, MCA, and
-other arteries (\Cref{fig:pvs}L). \mer{One more sentence on effect of
-  high PVS flow on tracer enrichment outside the PVS.} In contrast,
+PVS is clearly visible in the space adjacent to the PVS of the ACA, MCA, and other arteries (\Cref{fig:pvs}L). In contrast,
 there are no clear signs of early enrichment surrounding the PVS in
@@
     J) Total amount of tracer in the periarterial spaces as a function of the distance to the nearest network root over time for the baseline (left) and the high PVS flow (right) models;
-    K--M) sagittal (upper) and coronal (lower) views of the simulated tracer concentrations overlayed on the T1-weighted MR image after 4 and 6h for the baseline (K), high PVS flow (L) and high PVS dispersion (M) models (concentrations \discuss{below threshold masked}, pial surface outlined in pink, arachnoid membrane in cyan, and arteries in dark red).}
+    K--M) sagittal (upper) and coronal (lower) views of the simulated tracer concentrations overlayed on the T1-weighted MR image after 4 and 6h for the baseline (K), high PVS flow (L) and high PVS dispersion (M) models (concentration opacity linearly increasing with concentration value, pial surface outlined in pink, arachnoid membrane in cyan, and arteries in dark red).}
 \end{figure}
 
-% \mer{MER: Results #4 is about 500 words}
 \subsection*{Structural versus functional compartmentalization of perivascular spaces}
@@
 tracer to leak out of the PVS and spread within the Sylvian fissure to
-a greater extent. \discuss{With a small delay, more tracer appears via
-  the CSF pathway (\Cref{fig:compartmentalization}F, peak time
-  difference).} At the anterior cerebral artery (ACA-A2), the first tracer arrival in the PVS occurs after 3 hours, with a peak at 5 hours.
+a greater extent. With a small delay, more tracer appears via
+  the CSF pathway (\Cref{fig:compartmentalization}D).
+At the anterior cerebral artery (ACA-A2), the first tracer arrival in the PVS occurs after 3 hours, with a peak at 5 hours.
 The concentration peak in the PVS is tailed by a higher peak in the CSF (\Cref{fig:compartmentalization}E), indicating tracer arrival through a second pathway; i.e., directly through the CSF-filled space outside of the PVS.
@@
 
-% \mer{MER: Target Results 5 300 words}
 \subsection*{Enlarged PVSs delay periarterial and intracranial molecular enrichment}
@@
 within the dilated PVS. Notably, we observe later tracer arrival in
-all MCA segments, with up to \discuss{120} min later arrival in the
+all MCA segments, with up to 96 min later arrival in the
 MCA2s for the high PVS flow scenario (\Cref{fig:enlargement}E, F). For
@@
 accumulation of tracer in the PVS from 3--24 hours and reduced
-enrichment of the parenchyma at 6h, while the mean \discuss{PVS}
-concentration drops slightly (\Cref{fig:enlargement}G,H).
+enrichment of the parenchyma at 6 and 12 h, while the mean PVS
+concentration is initially lower, but later exceeds the baseline model (\Cref{fig:enlargement}G,H).
 \begin{figure}
@@
 \end{figure}
-
-% \mer{MER: Target Results 6 400 words}
-%\subsection*{The perivenous contribution to molecular clearance of CSF tracers}
-
-%\begin{itemize}
-%\item Look at clearance phase 12h-24h-48h-72h.
-%\item Look at concentration in the veins
-%\item Look at concentration in the parenchyma
-%\item Consider low-pressure sink for the perivenous spaces.
-%\end{itemize}
-
-%% \subsection*{Clearance of intra-parenchymal solutes}
-
-%% \begin{itemize}
-%% \item
-%%   Clearance of a molecular concentration in the parenchyma, different
-%%   molecular sizes - ranging from Gadubutrol to amyloid-$\beta$/$\tau$.
-%% \end{itemize}
 
@@
 
-% Variations with disease states
-%cohort variations, individual variations, physiological factor
-%variations. Enlarged PVS reduce transport at the whole cerebral scale.
-
+% Not discussed variations with disease states
+
+% Limitations
 In terms of limitations, our primary constraint modelling-wise is the
@@
 
+% Outlook
 In conclusion, our findings transfer insights from experimental
@@
 recent estimates~\cite{hladky2024regulation}.
-
-%% In addition, we slightly modify the segmentation to ensure correct
-%% connectivity of the ventricular system. In particular, we elongate
-%% the aqueduct and enforce no communication between the ventricles
-%% and the surrounding CSF spaces except for the median aperture.
-%% After morphological smoothing of the segmented image, we extract
-%% surface representations of the outer SAS boundary, the pial
-%% membrane and the components of the ventricular system.
-
-%% Note that the widely reported value of $150$\,ml of intracranial
-%% CSF volume is likely an underestimate. Recent estimates indicate
-%% that a value of $200$ to $300$\,ml is more accurate (see
-%% \cite{hladky2024regulation} and references therein
 
@@
   $\Gamma_v$ respectively.}
-
-%% The resulting vascular networks are further smoothened using a
-%% spline interpolation technique, and represented as graphs with
-%% nodes connected by edges.
 
@@
 \subsection*{CSF flow in the SAS and ventricular system}
-%\label{sec:csf_fluid_vel}
 
@@
 \subsection*{Dispersion in the CSF spaces}
-%\label{sec:csf_dispersion}
-%% MER: I'm commenting out labels as long as there is no numbering
 
@@
 network $\Lambda_v$.
-
-%% \subsection*{Dispersion in the PVS}
-
-%% Similar to the CSF spaces, dispersion due to cardiac and respiratory pulsation has been hypothesized to enhance transport in the perivascular spaces.
-However, recent estimates indicate that the effect of dispersion does not exceed a factor of 2 compared to diffusion alone \cite{asgari2016glymphatic,sharp2019dispersion,bojarskaite2023sleep,asgari2016glymphatic,troyetsky2021dispersion}.
-
-% \cite{sharp2019dispersion}: dispersion in PVS is negligiable in porous PVS, but might be relevant in open PVS (R=220) (theoretical estimates)
-%  \cite{bojarskaite2023sleep} numerical simulations show that the enhancement factor is likely in the range of 10 - 50 \%
-%\cite{asgari2016glymphatic} numerical simulations show increase of 16% to 50%  in distribution lengths
-%\cite{troyetsky2021dispersion}:   maximum enhancement of diffusion of only about 5%
 
@@
 
-%% The choice this scheme was is necessary here since continuous Galerkin
-%% methods even when supplemented with artificial diffusion stabilization
-%% terms were unstable in our experiments.
-
 \subsection*{Model validation}
@@
 financially supported by the Research Council of Norway under contract
-270053.
+270053. M.K. acknowledges support from the RCN grant \#303362.
 
@@
 
-%\newpage
-%\FloatBarrier
-%\input{background.tex}
-
 \end{document}

### 2025-01-26 scripts/figure2.py
update figure 2 labels

@@ -117,21 +117,21 @@ class Figure2_Panels(PanelsSpec):
     n2: Panel = Panel(
             RasterImage("plots/modelA-OnlyDispersion/modelA-OnlyDispersion_overview_1-6-12-24.png",
                          ElemSize(3, hn, units), crop=(1420, 100, 1350, 0)),
-            Location(xstart + compwidth, y3), auto_label=False,
+            Location(xstart + compwidth, y3), auto_label=True,
             content_offset=Location(0,0),
             text=Text("no advection (12 h)", **dict(textargs, y=-0.1, x=1.5))
         )
     n3: Panel = Panel(
             RasterImage("plots/modelA-LowD/modelA-LowD_overview_1-6-12-24.png",
                          ElemSize(3, hn, units), crop=(1420, 100, 1350, 0)),
-            Location(xstart + 2*compwidth, y3), auto_label=False,
+            Location(xstart + 2*compwidth, y3), auto_label=True,
             content_offset=Location(0,0),
             text=Text("low dispersion (12 h)", **dict(textargs, y=-0.1, x=1.5))
         )
     n4: Panel = Panel(
             RasterImage("plots/modelA-HighD/modelA-HighD_overview_1-6-12-24.png",
                          ElemSize(3, hn, units), crop=(1420, 100, 1350, 0)),
-            Location(xstart + 3*compwidth, y3), auto_label=False,
+            Location(xstart + 3*compwidth, y3), auto_label=True,
             content_offset=Location(0,0),
             text=Text("high dispersion (12 h)", **dict(textargs, y=-0.1, x=1.5))
         )

### 2025-01-26 paper/main.tex
Updates from Overleaf

@@
 network ends located within the SAS or up to 6 mm inside the
-parenchyma (\Cref{fig:results1}A--C, \Cref{fig:pvs}A) and includes
+parenchyma (\Cref{fig:results1}A--C, \Cref{fig:pvs_a}A) and includes
 major surface arteries of radius 0.5--1.4 mm. Imposing the pressure
@@
 with an average velocity of 0.08 \textmu m/s (antegrade) and a maximum
-velocity of 0.55 \textmu m/s (\Cref{fig:pvs}B). On the other hand,
-traveling waves of arterial wall motion (\Cref{fig:pvs}C), such as the
-pulse wave or other vasomotion~\cite{vanveluw2020vasomotion,
+velocity of 0.55 \textmu m/s (\Cref{fig:pvs_a}B). On the other hand,
+traveling waves of arterial wall motion (\Cref{fig:pvs_a}C), such as the pulse wave or other vasomotion~\cite{vanveluw2020vasomotion,
   munting2023spontaneous, bojarskaite2023sleep, broggini2024long,hauglund2025norepinephrine},
@@
 to 7.31 \textmu m/s near larger, ventral vessels such as the MCA
-(\Cref{fig:pvs}D). The same theory predicts that strong ultraslow
+(\Cref{fig:pvs_a}D). The same theory predicts that strong ultraslow
 vasomotions, if traveling antegrade at 0.1Hz with a wavelength of 0.02
@@
 13.05 \textmu m/s and maximum velocity 54.44 \textmu m/s
-(\Cref{fig:pvs}E).
+(\Cref{fig:pvs_a}E).
 
@@
 contribution. In both models, tracers were first observed at the basal
-artery after 48 minutes (\Cref{fig:pvs}F), defined by their first-time
-arrival (FTA), the time at which the concentration first exceeds $0.1$
-mmol/l. At all upstream locations along the middle cerebral arteries
+artery after 48 minutes (\Cref{fig:pvs_b}A), defined by their first-time arrival (FTA), the time at which the concentration first exceeds $0.1$ mmol/l.
+At all upstream locations along the middle cerebral arteries
 (MCAs) and anterior cerebral artery (ACA), we found substantially
-reduced FTAs with higher PVS flow (\Cref{fig:pvs}G), up to 2.5 hours
+reduced FTAs with higher PVS flow (\Cref{fig:pvs_a}B), up to 2.5 hours
 earlier in the left M2 segment of the MCA. PVS concentrations peaked
@@
 in the PVS and subsequently spreads into the tissue and SAS
-(\Cref{fig:pvs}H, I), and especially so with higher PVS flow. These
+(\Cref{fig:pvs_a}C, D), and especially so with higher PVS flow. These
 observations also hold on an aggregated level: computing the total
@@
 higher PVS flow, especially at earlier time points (2--9 hours)
-(\Cref{fig:pvs}J). Moreover, differences in tracer transport along the
+(\Cref{fig:pvs_b}E). Moreover, differences in tracer transport along the
 PVS translate to altered enrichment patterns on the whole-organ
 scale. For instance, after 4-6 hours, the faster-moving tracer in the
-PVS is clearly visible in the space adjacent to the PVS of the ACA, MCA, and other arteries (\Cref{fig:pvs}L). In contrast,
+PVS is clearly visible in the space adjacent to the PVS of the ACA, MCA, and other arteries (\Cref{fig:pvs_b}G). In contrast,
 there are no clear signs of early enrichment surrounding the PVS in
-the baseline model (\Cref{fig:pvs}K).
+the baseline model (\Cref{fig:pvs_b}F).
 
@@
 induced only minor changes in the global spreading rate compared to the
-baseline (\Cref{fig:pvs}G, M), thus indicating a negative answer to
+baseline (\Cref{fig:pvs_b}F, H), thus indicating a negative answer to
 this question.
+
 \begin{figure}
     \centering
-    \includegraphics[width=0.95\textwidth]{figures/figure3.png}
-    \caption{\textbf{Perivascular flow shapes and accelerates molecular enrichment.}}
-    \label{fig:pvs}
-\end{figure}
-\begin{figure}
-\ContinuedFloat
-  \caption{\textbf{(cont.)}
+    \includegraphics[width=0.95\textwidth]{figures/figure3a.png}
+    \caption{\textbf{A static pressure gradient and peristaltic pumping are potential drivers of perivascular flow.}
     A) Surface arterial network colored by arterial radius with vessel locations labeled;
@@
     D) Estimated net PVS flow induced by pulse wave peristaltic pumping;
-    E) Estimated net PVS flow induced by vasomotion/slow wave peristaltic pumping;
-    F) Comparison of mean tracer concentration in the PVS and over the outer PVS surface at key locations of the arterial tree, for the baseline (lower) and high PVS flow (upper) models, with labels: ``peak'': time-to-peak (h), ``$\Delta t$'': time difference between peaks (h), and ``FTA'': first-time tracer arrival (h).
-    G) first-time arrival for the main trunks of the periarterial network over the distance from the closest network root node (BA, ICA-R or ICA-L for the baseline, the high PVS flow and the high PVS dispersion models);
-    H, I) 2D slices showing zoom-in on the region surrounding the MCA-R and MCA2-R for the baseline (upper) and high PVS flow (lower) models.
+    E) Estimated net PVS flow induced by vasomotion/slow wave peristaltic pumping.}
+    \label{fig:pvs_a}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.99\textwidth]{figures/figure3b.png}
+    \caption{\textbf{Perivascular flow shapes and accelerates molecular enrichment.}
+    A) Comparison of mean tracer concentration in the PVS and over the outer PVS surface at key locations of the arterial tree, for the baseline (lower) and high PVS flow (upper) models, with labels: ``peak'': time-to-peak (h), ``$\Delta t$'': time difference between peaks (h), and ``FTA'': first-time tracer arrival (h).
+    B) first-time arrival for the main trunks of the periarterial network over the distance from the closest network root node (BA, ICA-R or ICA-L for the baseline, the high PVS flow and the high PVS dispersion models);}
+    \label{fig:pvs_b}
+\end{figure}
+\begin{figure}
+\ContinuedFloat
+\caption{(\textbf{cont.})
+    C, D) 2D slices showing zoom-in on the region surrounding the MCA-R and MCA2-R for the baseline (upper) and high PVS flow (lower) models.
 Innermost circle indicates the artery (blue), the surrounding annulus is the PVS (color represents concentration value), the surroundings are parenchyma and/or CSF spaces, separated by the pia (blue: towards the parenchyma, pink: towards the CSF));
-    J) Total amount of tracer in the periarterial spaces as a function of the distance to the nearest network root over time for the baseline (left) and the high PVS flow (right) models;
-    K--M) sagittal (upper) and coronal (lower) views of the simulated tracer concentrations overlayed on the T1-weighted MR image after 4 and 6h for the baseline (K), high PVS flow (L) and high PVS dispersion (M) models (concentration opacity linearly increasing with concentration value, pial surface outlined in pink, arachnoid membrane in cyan, and arteries in dark red).}
+    E) Total amount of tracer in the periarterial spaces as a function of the distance to the nearest network root over time for the baseline (left) and the high PVS flow (right) models;
+    F--H) sagittal (upper) and coronal (lower) views of the simulated tracer concentrations overlayed on the T1-weighted MR image after 4 and 6h for the baseline (K), high PVS flow (L) and high PVS dispersion (M) models (concentration opacity linearly increasing with concentration value, pial surface outlined in pink, arachnoid membrane in cyan, and arteries in dark red).}
 \end{figure}
@@
 
-From T1-weighted MR images of a human subject~\cite{hodneland2019new},
+From T1-weighted MR images of a 26-year-old, healthy male volunteer~\cite{hodneland2019new},
 we first automatically segment the brain parenchyma and CSF spaces,
@@
 centerline of a blood vessel segment and labeled with its lumen radius
-$R_1^i$ (\Cref{fig:results1}C, \Cref{fig:pvs}A). We also assign an
+$R_1^i$ (\Cref{fig:results1}C, \Cref{fig:pvs_a}A). We also assign an
 outer radius $R_2^i > R_1^i$ to each segment $i$, form the annular
@@
 frequency $f$, amplitude $\varepsilon$ and wave lengths $\lambda$
-(\Cref{fig:pvs}D). The corresponding contribution to the periarterial
+(\Cref{fig:pvs_a}C). The corresponding contribution to the periarterial
 flow velocity is defined for each segment $\Lambda_i$ as $\langle Q'_i
@@
 $\hat{u}_a^v$ respectively: \emph{cardiac} pulsations with $f =
-1.0$Hz, $\lambda = 2$ m and $\varepsilon = 1\%$ (\Cref{fig:pvs}E), and
-very low-frequency \emph{vasomotion} with $f = 0.1$Hz, $\lambda =
-0.02$m, and $\varepsilon = 10\%$ (\Cref{fig:pvs}F). No flow induced by
-peristalsis is considered for the perivenous network. In the baseline
-model, we combine the contributions from CSF production and cardiac
-peristalsis to set $\hat{u}$ in~\eqref{eq:multi_transport} as
+1.0$Hz, $\lambda = 2$ m and $\varepsilon = 1\%$ (\Cref{fig:pvs_a}D), and very low-frequency \emph{vasomotion} with $f = 0.1$Hz, $\lambda =
+0.02$m, and $\varepsilon = 10\%$ (\Cref{fig:pvs_a}E).
+No flow induced by peristalsis is considered for the perivenous network.
+In the baseline model, we combine the contributions from CSF production and cardiac peristalsis to set $\hat{u}$ in~\eqref{eq:multi_transport} as
 $\hat{u}_a = \hat{u}_a^p + \hat{u}_a^c$ in $\Lambda_a$, and $\hat{u}_v

### 2025-01-26 paper/supplementary_main.tex

@@
 using finite element methods. The following subsections provide more
-details, model-by-model. \rami{Add a note on the convection dominated problem and the need for DG}
+details, model-by-model.
 
@@
 these equations using an implicit finite difference scheme in time and
-a discontinuous Galerkin (DG) finite element method with upwinding in
-space to accurately capture sharp boundary layers. \mk{Comment on Peclet number}
+a discontinuous Galerkin (DG) finite element method with upwinding in space to accurately capture sharp boundary layers.
+We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm and taking the increased diffusivity due to dispersion into account).
 
@@
 \end{equation}
-and $\eta$ is a user-specified penalty parameter. To discretize the convection term in~\eqref{eq:3d_pde}, we use
+and $\eta$ is a user-specified penalty parameter (we set $\eta=1000$). To discretize the convection term in~\eqref{eq:3d_pde}, we use
 upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
@@
 $\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$, on the inner ventricular
-interface $\Gamma_{\rm LV}$ by $\mathcal{F}_{\rm LV}$,
 %on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$
@@
 enhance molecular transport~\cite{taylor1953dispersion,
-  watson1983diffusion, asgari2016glymphatic, keith2019dispersion,
+  watson1983diffusion, asgari2016glymphatic, sharp2019dispersion,
   ray2021quantitative, troyetsky2021dispersion}. To account for the
@@
 human CSF spaces, we adapt existing theoretical
-estimates~\cite{watson1983diffusion, keith2019dispersion}. More
+estimates~\cite{watson1983diffusion, sharp2019dispersion}. More
 specifically, we compute spatially-varying dispersion enhancement
@@
   Further, assuming unsteady dispersion, we follow Sharp et
-  al.~\cite{keith2019dispersion} to estimate local enhancement factors
+  al.~\cite{sharp2019dispersion} to estimate local enhancement factors
   $S_c, S_r$ from the non-dimensionalized pressure gradients:
@@
 %% Taking the large spatial differences of pulsatile CSF flow into account, we estimate a spatially varying diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac- and respiration-induced pulsatile CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}.
-These estimates are are derived under simplifying assumptions on the problem geometry, the flow field and concentration profiles \cite{watson1983diffusion, keith2019dispersion}.
+These estimates are are derived under simplifying assumptions on the problem geometry, the flow field and concentration profiles \cite{watson1983diffusion, sharp2019dispersion}.
 Therefore such an estimate is heuristic; however, our computational results qualitatively agree with observed data.
@@
 
-%% Further assuming unsteady dispersion, we follow Sharp et al \cite{keith2019dispersion} and estimate each enhancement factor $R_{\rm cardiac}$ and $R_{\rm resp}$ from the non-dimensionalized pressure gradient
+%% Further assuming unsteady dispersion, we follow Sharp et al \cite{sharp2019dispersion} and estimate each enhancement factor $R_{\rm cardiac}$ and $R_{\rm resp}$ from the non-dimensionalized pressure gradient
 %% $$R_{i} \approx dP_{i}^2 / \alpha_{i}^3, \quad  \text{where} \quad  dP=\frac{|\nabla p^{\rm up}_{i}|}{\omega_{i} \mu / h} \quad \text{for} \quad i = \{\mathrm{cardiac}, \mathrm{resp}\}.$$
@@
 Specifically, we generate a sequence of three meshes (\Cref{fig:mesh_refinement}) with an increasing number of computational mesh vertices and cells, solve all relevant simulation steps (CSF flow and intracranial transport computations) for the baseline model on each mesh, and compare the results across meshes with respect to a set of key quantities of interest.
-The mesh refinement employed here is adaptive in the sense that the coarser mesh is refined near expected sharp concentration gradients.
+The mesh refinement employed here is adaptive in the sense that the coarser mesh is manually refined near expected sharp concentration gradients.
 This allows us to better capture dynamics and reduce undershoots, see \Cref{fig:mesh_convergence_concentrations}.
-Similarly, we investigate the effect of the time step size by solving the intracranial transport model with different time steps: 1, 2, and 4 minutes.
+Similarly, we investigate the effect of the time step size by solving the intracranial transport model on the standard resolution mesh with different time steps: 1, 2, and 4 minutes.
 
@@
 with less than 10\,\% with mesh refinement, and less than 1\,\% with
-time refinement. We thus conclude that the standard resolution mesh
-and a time step of 2 min offer sufficient accuracy for our
-simulations.
+time refinement.
+We thus conclude that the standard resolution mesh and a time step of 2 min offer sufficient accuracy for our simulations, and remark that all reported results are obtained with the standard resolution mesh.
 
@@
 velocities are conservative.
+
+\paragraph{Predictions on overall tracer spreading patterns}
+
+\begin{table}
+\centering
+\begin{tabular}{c|c|c|c|c}
+    \toprule
+   quantity of interest  & Watts et al\cite{watts2019measuring}  & Ringstad et al\cite{ringstad2018brain} & Eide and Ringstand \cite{eide2024functional} & ours (vasomotion model)\\
+   \midrule
+    Spinal transit time &  1 - 3 hours & 20 min &  $13\pm6.3\,$ min & N/A \\
+    Time of peak amount (CSF) & 10 - 15 hours & 6–9 hours & & 2 hours \\
+    Time of peak amount (parenchyma) & 10 - 26 hours & 24 hours & N/A& 12 - 24 hours \\
+    Peak amount (CSF) &  N/A & N/A & N/A & 0.5 mmol \\
+    Peak amount (parenchyma) & N/A & N/A& N/A & 0.09 mmol \\
+    injected tracer quantity &  1.0 mmol & 0.5 mmol & 0.5 mmol & 0.5 mmol (cranially) \\
+    Peak concentration (CSF) & 0.5 mmol/l &  N/A&  N/A & 1.4 mmol/l\\
+    Peak concentration (parenchyma) & 0.1 mmol/l & N/A & N/A & 0.4 mmol/l \\
+    FTA in MCA1 & N/A & N/A &  37.8 ± 47.0 min &  1:48 / 1:24 hours \\
+    FTA in MCA2 & N/A & N/A & 53.1 ± 50.5 min & 2:12 / 1:48 hours \\
+    FTA in ACA2 & N/A & N/A & 48.3 ± 46.1 min & 3:00 / 3:12 hours \\
+    \bottomrule
+\end{tabular}
+\caption{Comparison of literature values and model predictions on overall tracer spreading patters and timing}
+\label{tab:spreading comparison}
+\end{table}

### 2025-01-26 paper/supplementary_main.tex
Updates from Overleaf

@@
 \begin{table}
+\centering
   \small
@@
 pressure differences across the aqueduct of 12.1 $\pm$ 5.7 Pa and 9.5
-$\pm$ 7.2 Pa, respectively; thus our baseline estimate of the
-respiratory contribution (\fixme{1.4 Pa}) may be an
+$\pm$ 7.2 Pa, respectively; thus our baseline estimate of the respiratory contribution 1.4 Pa may be an
 underestimation. On the other hand, our cardiac- and
-respiratory-driven CSF flow estimates peak at 19.8 cm/s and 4.8 cm/s
-in the caudal direction, respectively, which are higher than
-phase-contrast MRI measurements of cardiac and respiratory CSF flow
-components \cite{takizawa2017characterization,
-  yildiz2017quantifying}. Some variation in CSF flow velocities is
-expected considering that the values from MRI represent averages
+respiratory-driven CSF flow estimates peak at 19.8 cm/s and 4.8 cm/s in the caudal direction, respectively, which are higher than phase-contrast MRI measurements of cardiac and respiratory CSF flow
+components \cite{takizawa2017characterization, yildiz2017quantifying}. Some variation in CSF flow velocities is expected considering that the values from MRI represent averages
 \cite{yildiz2017quantifying} and that the geometry of the CSF spaces
@@
 \centering
-\begin{tabular}{c|c|c|c|c}
+\small
+\begin{tabular}{ccccc}
 \toprule
@@
 \midrule
-    Spinal transit time &  1 - 3 hours & 20 min &  $13\pm6.3\,$ min & N/A \\
-    Time of peak amount (CSF) & 10 - 15 hours & 6–9 hours & & 2 hours \\
-    Time of peak amount (parenchyma) & 10 - 26 hours & 24 hours & N/A& 12 - 24 hours \\
+Spinal transit time ($t_{st}$) &  1 - 3 hours & 20 min &  $13\pm6.3\,$ min & N/A \\
+Time of peak amount (CSF) & 10 - 15 hours & 6–9 hours & N/A & 2 hours +$t_{st}$ \\
+Time of peak amount (parenchyma) & 10 - 26 hours & 24 hours & N/A & 12 - 24 hours \\
 Peak amount (CSF) &  N/A & N/A & N/A & 0.5 mmol \\
 Peak amount (parenchyma) & N/A & N/A& N/A & 0.09 mmol \\
-    injected tracer quantity &  1.0 mmol & 0.5 mmol & 0.5 mmol & 0.5 mmol (cranially) \\
+Injected tracer quantity &  1.0 mmol & 0.5 mmol & 0.5 mmol & 0.5 mmol (cranially) \\
 Peak concentration (CSF) & 0.5 mmol/l &  N/A&  N/A & 1.4 mmol/l\\
 Peak concentration (parenchyma) & 0.1 mmol/l & N/A & N/A & 0.4 mmol/l \\
-    FTA in MCA1 & N/A & N/A &  37.8 ± 47.0 min &  1:48 / 1:24 hours \\
-    FTA in MCA2 & N/A & N/A & 53.1 ± 50.5 min & 2:12 / 1:48 hours \\
-    FTA in ACA2 & N/A & N/A & 48.3 ± 46.1 min & 3:00 / 3:12 hours \\
+FTA in MCA1 & N/A & N/A &  37.8 ± 47.0 min &  1:48 / 1:24 hours +$t_{st}$ \\
+FTA in MCA2 & N/A & N/A & 53.1 ± 50.5 min & 2:12 / 1:48 hours +$t_{st}$ \\
+FTA in ACA2 & N/A & N/A & 48.3 ± 46.1 min & 3:00 hours +$t_{st}$ \\
 \bottomrule
 \end{tabular}
-\caption{Comparison of literature values and model predictions on overall tracer spreading patters and timing}
+\caption{Comparison of literature values and model predictions on tracer spreading patterns and timing}
 \label{tab:spreading comparison}

### 2025-01-26 paper/main.tex

@@
 (\Cref{fig:results1}A--B). We next manually adjust the segmentation to
-accurately represent the connections between CSF spaces (aqueduct,
+accurately represent the connections and barriers between CSF spaces (aqueduct,
 median aperture, tentorium cerebelli), smoothen, and finally extract
@@
 ends as leaf nodes. The resulting perivenous network graph consists of
-$\fixme{N}$ edges connected at $\fixme{N}$ inner nodes and with
-$\fixme{N}$ end nodes, and the associated domain is denoted by
-$\Lambda_v$. \discuss{Finally, we label the outer surface of the PVSs
-  associated with $\Lambda_a$ and $\Lambda_v$, by $\Gamma_a$ and
-  $\Gamma_v$ respectively.}
+$24\,829$ edges connected at $23\,881$ inner nodes and with $949$ end nodes, and the associated domain is denoted by $\Lambda_v$.
+Finally, we label the outer surface of the PVSs associated with $\Lambda_a$ and $\Lambda_v$, by $\Gamma_a$ and $\Gamma_v$ respectively.
 

### 2025-01-27 paper/supplementary_main.tex
Updates from Overleaf

@@
 a discontinuous Galerkin (DG) finite element method with upwinding in space to accurately capture sharp boundary layers.
-We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm and taking the increased diffusivity due to dispersion into account).
+We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm, and accounting for the increased diffusivity due to dispersion).
 
@@
 
+We compare our in-silico predictions with clinical data from human subjects, in particular the studies by Ringstad et al~\cite{ringstad2018brain}, Watts et al~\cite{watts2019measuring} and Eide and Ringstad~\cite{eide2024functional}.
+Finding an overall good agreement, we provide a detailed overview on key quantities of interest in \cref{tab:spreading_comparison}.
+We remark that the our simulation begins at the time of first tracer appearance at the craniocervical junction, while the clinical studies report time with respect to intrathecal injection time, leading to an offset, which is captured by the spinal transition time.
+
 \begin{table}
@@
 \small
-\begin{tabular}{ccccc}
+\begin{tabular}{lcccc}
 \toprule
-quantity of interest  & Watts et al\cite{watts2019measuring}  & Ringstad et al\cite{ringstad2018brain} & Eide and Ringstand \cite{eide2024functional} & ours (vasomotion model)\\
+quantity of interest  & Watts et al\cite{watts2019measuring}  & Ringstad et al\cite{ringstad2018brain} & Eide and Ringstad \cite{eide2024functional} & high PVS flow model \\
 \midrule
@@
 \end{tabular}
-\caption{Comparison of literature values and model predictions on tracer spreading patterns and timing}
-\label{tab:spreading comparison}
+\caption{Comparison of clinical studies and model predictions on tracer spreading patterns and arrival times.}
+\label{tab:spreading_comparison}
 \end{table}

### 2025-01-27 paper/main.tex

@@
 regions near large cerebral surface arteries, peaking in the CSF
-spaces between 6 and 24 hours, with parenchymal tracer content still
+spaces between 6 and 9 hours, with parenchymal tracer content still
 increasing after 24 hours, but with large individual
@@
 enrichment pattern and a concentration peak in the SAS of $\approx$0.5
-mmol/L occuring after 10 to 15 hours. Our in-silico enrichment
-patterns agree with these observations, though with an earlier
-\discuss{$\sim$5 + plus spinal transit time)} rather than later peak
-in the CSF spaces. In particular, we note that the substantial
-individual variation observed clinically is comparable to the
-variation in in-silico enrichment patterns associated with reduced CSF
-production or reduced intracranial pulsatility. \mer{One sentence on
-  tracer content comparison.} In terms of perivascular tracer
+mmol/l (relative to a total amount of 1.0 mmol) occuring after 10 to 15 hours. Our in-silico enrichment
+patterns agree with these observations, though with an earlier rather than later peak in the CSF spaces (2 hours after first tracer appearance at the craniocervical junction).
+In particular, we note that the substantial individual variation observed clinically is comparable to the variation in in-silico enrichment patterns associated with reduced CSF production or reduced intracranial pulsatility.
+We find that about 20\% of the tracer reaching the cranium enters the parenchyma, which aligns with the reported peak concentration values of $0.5\,$mmol/l in the SAS and $0.1\,$mmol/l in the parenchyma reported by Watts et al~\cite{watts2019measuring}.
+In terms of perivascular tracer
 enrichment, Eide and Ringstad~\cite{eide2024functional} report of
 first-time tracer appearance in the PVS of the MCA after 37.8 $\pm$
-47.0 (M1), 53.1 $\pm$ 50.5 (M2) and 82.6 $\pm$ 61.5 (M3) min, thus
-with just a 15.3 min delay between M2 and M1. Comparing with the our
-FTAs in the MCA segments, we note that the baseline model predicts a
-more-than 3 hour difference in FTA between the M1 and M2. Instead, the
-high PVS flow scenario predicts a 24 min delay between these
-segments. Moreover, the high PVS flow scenario predicts FTAs of
-84--108 min (M1), 84--146 min (M2) which overlaps with the previous
-ranges \discuss{modulo spinal transit time}. We interpret these
-results as indicative of a net CSF flow in human surface PVS beyond
-what would be induced by CSF production and cardiac peristalsis alone.
+47.0 (M1), 53.1 $\pm$ 50.5 (M2) and 82.6 $\pm$ 61.5 (M3) min after injection, thus with just a 15.3 min delay between M2 and M1.
+Comparing with our FTAs in the MCA segments, we note that the baseline model predicts a more-than 3 hour difference in FTA between the M1 and M2.
+Instead, the high PVS flow scenario predicts a 24 min delay between these segments.
+Moreover, the high PVS flow scenario predicts FTAs of
+84--108 min (M1) and 108--132 min (M2) after tracer appearance at the craniocervical junction, which overlaps with the previous ranges if one assumes a tracer travel time from the point of injection to the craniocervical junction of $13\,$min (as reported by Eide and Ringstad \cite{eide2024functional}).
+We interpret these results as indicative of a net CSF flow in human surface PVS beyond what would be induced by CSF production and cardiac peristalsis alone.
 % M2 - M1 (mean) ~ 15.3 min
@@
 not more than 24 hours. In particular, we note that, we observe tracer
-accumulation at the perivascular network ends after \fixme{24} hours,
+accumulation at the perivascular network ends after 12 hours,
 similar to that previously observed for
@@
 or neither (\Cref{sec:app:more_maths}). Additionally, we increase the
-permeability $\xi$ ($100\times$) \fixme{near} the leaf nodes of
-$\Lambda_a$ to account for the transition from explicitly resolved
+permeability $\xi$ ($100\times$) at the the leaf nodes of
+$\Lambda_a$, linearly decreasing over the corresponding cell, to account for the transition from explicitly resolved
 PVSs to their unresolved continuation in the CSF spaces and

### 2025-01-29 paper/supplementary_main.tex
Updates from Overleaf

--- 2025-01-26 paper/supplementary_main.tex
@@
 inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$ and thus of
-width $R_2^i - R_1^i$. We denote the outer lateral surface of the
+width $R_2^i - R_1^i$, and remark that we interpret $R^i_2$ and $R^i_1$ as temporal averages (fixed in time), as the pulsations considered in \Cref{sec:app:peristalsis} are beyond the temporal resolution of our model.
+We denote the outer lateral surface of the
 periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$,
@@
 graph with the centerlines $\{ \Lambda^i \}$ as oriented edges and the
-connections as nodes $\mathcal{V}$. \mk{We always keep $R^i_2$, $R^i_1$ fixed in time in A.1.1 and A.1.4, right?}
+connections as nodes $\mathcal{V}$.
 
@@
 main text for explicit parameter values while giving the general,
-abstract form here, and to the reference~\cite{masri2024modelling} for
+abstract form here, and to the reference~\cite{masri2024modelling, laurino2019derivation} for
 the derivation and analysis of this 3D-1D model. Recall that we model
@@
 under the assumption that $\xi$ is constant on each $\partial
-\Theta_2^i$, see also~\cite{masri2024discontinuous}.
+\Theta_2^i$.
 %% , since:
@@
 $\mathcal{N} \subset \mathcal{V}$, write $\hat{c}^i =
-\hat{c}|_{\Lambda^i}$, and let $\mathcal{E}(y)$ denote the set of
-(one, two or three) edges in $\{\Lambda^i\}$ sharing the node $y$. To
+\hat{c}|_{\Lambda^i}$, and for $y\in\mathcal{B}$ let $\mathcal{E}(y)$ denote the set of
+(two or three) edges in $\{\Lambda^i\}$ sharing the node $y$. To
 impose continuity of the concentrations, we set that the
@@
   \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n^i(y)
-  = 0
+  = 0.
   \label{eq:flux_conservation}
 \end{equation}
-where the function $n^i : \mathcal{V} \rightarrow \{-1, \mk{0}, 1\}$
-defines an orientation of each vertex relative to the
-edges. Specifically, for an edge $\Lambda^i$ from $y_{\rm in}$ to $y_{\rm out}$
+Here the (normal ``vector") function $n^i$ takes the values in $\left\{-1, 1\right\}$ and 
+defines an orientation of vertices of edge $\Lambda^i$. Specifically, for an edge $\Lambda^i$ oriented from $y_{\rm in}$ to $y_{\rm out}$ we set 
 \begin{equation}
-  n^i(y_{\rm in}) = 1, \quad
-  n^i(y_{\rm{out}}) = -1, \quad
-  \mk{n^i(y) = 0 \quad \foralls y \in 
-  \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} }.  
+  n^i(y_{\rm in}) = 1,\quad
+  n^i(y_{\rm{out}}) = -1. %\quad
+  %\mk{n^i(y) = 0 \quad \foralls y \in 
+  %\mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} }.  
 \label{eq:def:sign}
 \end{equation}
-\mk{Why do we need the normal on non-IO vertices?}
+%\mk{Why do we need the normal on non-IO vertices?}
 
@@
   \begin{alignat}{2}
-    \mu \nabla \bm u \cdot \bm{n} - p \bm n &= -R_0 ( \bm u \cdot \bm n ) \bm n
+( \mu \nabla \bm u \cdot \bm{n} - p \bm n) \cdot \bm n  &= -R_0 ( \bm u \cdot \bm n ), \quad \bm u \cdot \bm t = 0
     && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
@@
     %\bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
-    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm \tau = 0 \quad && \text{ on } \Gamma_{\rm{LV}} ,  
+    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm t = 0 \quad && \text{ on } \Gamma_{\rm{LV}} ,  
   \end{alignat}
@@
 \end{subequations}%
-where $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the (dynamic) CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$.
+where $\bm n$ and $\bm t$ denote the unit outward normal and tangent vectors to the boundary respectively,  $\mu$ is the (dynamic) CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$.
 We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:app:dispersion}.
@@
 conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
-these equations using an implicit finite difference scheme in time and
-a discontinuous Galerkin (DG) finite element method with upwinding in space to accurately capture sharp boundary layers.
-We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm and taking the increased diffusivity due to dispersion into account).
+these equations using an implicit finite difference scheme in time, 
+a discontinuous Galerkin (DG) finite element method with upwinding in space for the 3D domain to accurately capture sharp boundary layers, and a continuous Galerkin method for the 1D networks.
+We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm, and accounting for the increased diffusivity due to dispersion).
 
@@
   \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
+  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
+  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right) 
    + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
-  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
-  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right).
+  .
   \label{eq:dg_sip}
@@
 \begin{equation}
-  \gamma_{D,F} = \frac{\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
+  \gamma_{D,F} = \frac{2\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
 \end{equation}
@@
  where $h_{\Lambda^i}$ is the mesh--size of $\Lambda^i$.
+Note that the condition \eqref{eq:flux_conservation} is enforced weakly in the above formulation.
 
@@
 \end{align*}
-Here, $H(\mathrm{div}, \Omega_{\rm{CSF}})$ is the space of $L^2(\Omega_{\rm{CSF}})$ vector fields with $L^2(\Omega_{\rm{CSF}})$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, and $\bm n$ is the unit outward normal vector to each facet.
+Here, $H(\mathrm{div}, \Omega_{\rm{CSF}})$ is the space of $L^2(\Omega_{\rm{CSF}})$ vector fields with $L^2(\Omega_{\rm{CSF}})$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, $g$ is a given constant, and $\bm n$ is the unit outward normal vector to each facet.
 Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by
-\begin{equation}
-\bm v_n = (\bm v \cdot \bm n) \bm n, \quad \bm v_t = \bm v - \bm v_n. 
-\end{equation}
+$$ \bm v_n = (\bm v \cdot \bm n) \bm n, \quad \bm v_t = \bm v - \bm v_n. 
+$$ 
 Since $\bm V_{h,g} \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm
-  v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$. Continuity in the
+  v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$, the interior facets to $\Omega_{\mathrm{CSF}}$. Continuity in the
 tangential component is enforced weakly via interior penalization. For
-convenience, we collect all facets exterior to the CSF space but not
-on the (Neumann) boundary $\Gamma_{\rm AM-U}$ in the set
-$\mathcal{F}_{e} = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
-\cup \mathcal{F}_{\rm AM-L} \cup \mathcal{F}_{\rm SSAS}$, where facets
+convenience, we collect all facets exterior to the CSF space  $\Gamma_{\rm AM-U}$ in the set
+$$\mathcal{F}_{e} = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
+\cup \mathcal{F}_{\rm AM-L}\cup \mathcal{F}_{\rm AM-U}  \cup \mathcal{F}_{\rm SSAS}, $$ 
+where facets
 lying on the pial interface $\Gamma_{\rm pia}$ are denoted by
 $\mathcal{F}_{\rm pia}$, on the lower and upper outer (arachnoid) boundary
-$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$, on the inner ventricular
+$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$ respectivley, 
 %on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$
-and on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
-$\mathcal{F}_{\rm SSAS}$. Now, define the form 
+ on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
+$\mathcal{F}_{\rm SSAS}$, and on the surface of the lateral ventricles by $\mathcal{F}_{\rm LV}$. Now, define the form 
 \begin{multline}
@@
   - \sum_{F \in \mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}} \left (
-  \int_{F} \mu \{\nabla \bm u\} \cdot \bm n_F \cdot [\bm v_t] 
+  \int_{F} \mu \{\nabla \bm u\}  \bm n_F \cdot [\bm v_t] 
   - %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
-  \int_{F} \mu \{\nabla \bm v\}\cdot \bm n_F \cdot [\bm u_t]
+  \int_{F} \mu \{\nabla \bm v\}  \bm n_F \cdot [\bm u_t]
   + %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
@@
 We set the penalty parameter for the tangential continuity to be $\sigma = 20$.
-Our finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_{h,g} \times Q_h$ with $g  = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}$  such that
+The finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_{h,g} \times Q_h$ with $g  = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}$  such that
 \begin{subequations}
@@
   \begin{equation}
-    \alpha^2_i = \frac{h^2 \omega_i \rho}{\mu} \qquad i \in \{c, r\},
+    \alpha^2_i = \frac{h_{\rm SAS}^2 \omega_i \rho}{\mu} \qquad i \in \{c, r\},
   \end{equation}
-  with CSF density $\rho = 10^3$ kg/m$^3$, a mean CSF space width $h =
+  with CSF density $\rho = 10^3$ kg/m$^3$, a mean CSF space width $h_{\rm SAS} =
   1.5$mm, and CSF viscosity $\mu$ given in~\Cref{tab:parameters}.  For
@@
   \begin{equation}
-    S_i(x) = \frac{1}{\alpha_i^3} \frac{1}{\omega_i \mu/h} | \nabla p_i^1(x) |
+    S_i(x) = \frac{1}{\alpha_i^3} \frac{1}{\omega_i \mu/h_{\rm SAS}} | \nabla p_i^1(x) |
     \qquad i \in \{c, r\}.
@@
 By definition, the wave number is $k = 2 \pi/\lambda$ and the angular frequency is $\omega = 2 \pi f$.
-Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L$, \mk{Should the $L$ here be $L_i$?} baseline inner radius $R_1^i$, fixed outer radius $R_2^i$, and outer-to-inner ratio $\beta_i = R_2^i/R_1^i$.
+Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L_i$, baseline inner radius $R_1^i$, fixed outer radius $R_2^i$, and outer-to-inner ratio $\beta_i = R_2^i/R_1^i$.
 These geometric parameters and the assumption of annular cylindrical PVS segments yield hydraulic resistances $\mathcal{R}_{o, i}$ and additional characteristic parameters, see \cite{gjerde2023directional} for the complete definitions and schematics.
-Since the analytical estimate is derived under the assumption that $k L \approx \mathcal{O} (1)$~\cite{gjerde2023directional} and has been verified against numerical simulations for $k L$ of the order $10^{-1}-10^2$~\cite[Table I]{gjerde2023directional}, we consider it applicable for the wave lengths and vascular network data considered here in which $k L$ range from 0.15 to 30 for the strong vasomotion and 0.0015 to 0.30 for the cardiac waves (\Cref{tab:pvs:parameters}).
+Since the analytical estimate is derived under the assumption that $k L_i \approx \mathcal{O} (1)$~\cite{gjerde2023directional} and has been verified against numerical simulations for $k L_i$ of the order $10^{-1}-10^2$~\cite[Table I]{gjerde2023directional}, we consider it applicable for the wave lengths and vascular network data considered here in which $k L_i$ range from 0.15 to 30 for the strong vasomotion and 0.0015 to 0.30 for the cardiac waves (\Cref{tab:pvs:parameters}).
 \begin{table}
@@
 Specifically, we generate a sequence of three meshes (\Cref{fig:mesh_refinement}) with an increasing number of computational mesh vertices and cells, solve all relevant simulation steps (CSF flow and intracranial transport computations) for the baseline model on each mesh, and compare the results across meshes with respect to a set of key quantities of interest.
-The mesh refinement employed here is adaptive in the sense that the coarser mesh is manually refined near expected sharp concentration gradients.
+The mesh refinement employed here is localized near expected sharp concentration gradients.
 This allows us to better capture dynamics and reduce undershoots, see \Cref{fig:mesh_convergence_concentrations}.
@@
 
-\paragraph{Predictions on overall tracer spreading patterns}
+%\paragraph{Predictions on overall tracer spreading patterns}
 
-\begin{table}
-\centering
-\small
-\begin{tabular}{ccccc}
-\toprule
-quantity of interest  & Watts et al\cite{watts2019measuring}  & Ringstad et al\cite{ringstad2018brain} & Eide and Ringstand \cite{eide2024functional} & ours (vasomotion model)\\
-\midrule
-Spinal transit time ($t_{st}$) &  1 - 3 hours & 20 min &  $13\pm6.3\,$ min & N/A \\
-Time of peak amount (CSF) & 10 - 15 hours & 6–9 hours & N/A & 2 hours +$t_{st}$ \\
-Time of peak amount (parenchyma) & 10 - 26 hours & 24 hours & N/A & 12 - 24 hours \\
-Peak amount (CSF) &  N/A & N/A & N/A & 0.5 mmol \\
-Peak amount (parenchyma) & N/A & N/A& N/A & 0.09 mmol \\
-Injected tracer quantity &  1.0 mmol & 0.5 mmol & 0.5 mmol & 0.5 mmol (cranially) \\
-Peak concentration (CSF) & 0.5 mmol/l &  N/A&  N/A & 1.4 mmol/l\\
-Peak concentration (parenchyma) & 0.1 mmol/l & N/A & N/A & 0.4 mmol/l \\
-FTA in MCA1 & N/A & N/A &  37.8 ± 47.0 min &  1:48 / 1:24 hours +$t_{st}$ \\
-FTA in MCA2 & N/A & N/A & 53.1 ± 50.5 min & 2:12 / 1:48 hours +$t_{st}$ \\
-FTA in ACA2 & N/A & N/A & 48.3 ± 46.1 min & 3:00 hours +$t_{st}$ \\
-\bottomrule
-\end{tabular}
-\caption{Comparison of literature values and model predictions on tracer spreading patterns and timing}
-\label{tab:spreading comparison}
-\end{table}
+%We compare our in-silico predictions with clinical data from human subjects, in particular the studies by Ringstad et al~\cite{ringstad2018brain}, Watts et al~\cite{watts2019measuring} and Eide and Ringstad~\cite{eide2024functional}.
+Finding an overall good agreement, we provide a detailed overview on key quantities of interest in %FTA in MCA2 & N/A & N/A & 53.1 ± 50.5 min & 2:12 / 1:48 hours +$t_{st}$ \\
+%FTA in ACA2 & N/A & N/A & 48.3 ± 46.1 min & 3:00 hours +$t_{st}$ \\
+%\bottomrule
+%\end{tabular}
+%\caption{Comparison of clinical studies and model predictions on tracer spreading patterns %and arrival times.}
+%\label{tab:spreading_comparison}
+%\end{table}

### 2025-01-29 paper/main.tex

@@
   The mechanisms of intracranial solute transport are fundamental to
-  human brain health and function, with alterations often linked to
+  human brain health, with alterations often linked to
   disease and functional impairment, and with distinct opportunities
@@
   across scales, between species and from different modalities. Here,
-  we combine geometric model reduction and mixed-dimensional
-  modelling, multi-modal magnetic resonance images, and high
-  performance computing to construct and explore a high-fidelity
-  in-silico model of human intracranial molecular enrichment and
-  clearance. This model predicts the temporal and spatial spreading of
-  a solute within an image-derived geometric representation of the
+  we combine mixed-dimensional modelling, multi-modal magnetic 
+  resonance images, and high performance computing to construct and 
+  explore a high-fidelity in-silico model of human intracranial molecular enrichment. 
+  This model predicts the temporo-spatial spreading of a solute
+  within an image-derived geometric representation of the
   subarachnoid space, ventricular system and brain parenchyma,
-  including branching networks of surface perivascular spaces
+  including networks of surface perivascular spaces
   (PVSs). Our findings highlight the significant impact of
@@
   demonstrate that low-frequency vasomotion induces moderate CSF flow
-  rates in surface PVS networks that substantially enhance tracer
-  enrichment, and that impaired enrichment is a natural direct
+  in surface PVS networks that substantially enhances tracer
+  enrichment, and that impaired enrichment is a direct
   consequence of enlarged PVSs. This openly available technology
@@
   intracranial pulsatility, CSF production, and efflux, and for
-  exploring questions relating to drug delivery and clearance in the
-  human brain environment.
+  exploring drug delivery and clearance in the human brain.
 \end{abstract}
@@
 field induced by CSF production at the periarterial network ends
-induces ultraslow steady CSF flow of variable direction in these PVSs,
+induces slow steady CSF flow of variable direction in these PVSs,
 with an average velocity of 0.08 \textmu m/s (antegrade) and a maximum
@@
     E) Total amount of tracer in the periarterial spaces as a function of the distance to the nearest network root over time for the baseline (left) and the high PVS flow (right) models;
-    F--H) sagittal (upper) and coronal (lower) views of the simulated tracer concentrations overlayed on the T1-weighted MR image after 4 and 6h for the baseline (K), high PVS flow (L) and high PVS dispersion (M) models (concentration opacity linearly increasing with concentration value, pial surface outlined in pink, arachnoid membrane in cyan, and arteries in dark red).}
+    F--H) sagittal (upper) and coronal (lower) views of the simulated tracer concentrations overlayed on the T1-weighted MR image after 4 and 6h for the baseline (K), high PVS flow (L) and high PVS dispersion (M) models (concentration opacity linearly increasing with concentration value, pial surface outlined in pink, arachnoid membrane in cyan, and arteries in dark red).
+An interactive visualization of the baseline and the high PVS flow model can be found \href{https://baseline-pvs-flow.surge.sh/}{here}, and \href{https://high-pvs-flow.surge.sh/}{here}.}
 \end{figure}
@@
       Total molecular volume & $m_{\rm tot}$ & $0.5$ & mmol & \cite{eide2024functional} \\
-      Molecular influx, time-to-peak & $T_{\rm max}$ & $1$ & h & \fixme{$\ast$} \\
+      Molecular influx, time-to-peak & $T_{\rm max}$ & $1$ & h & $\ast$ \\
       CSF production rate & $u_{\mathrm{in}}$ & $0.63$  & l / day & \cite{nilsson1992circadian} \\
@@
 financially supported by the Research Council of Norway under contract
-270053. M.K. acknowledges support from the RCN grant \#303362.
+270053. 
+M.C. acknowledges support from the RCN grant \#324239 (EMIx).
+M.K. acknowledges support from the RCN grant \#303362.
+The figures 1A, 1D, 2C, 3C, 5A and 6A were created in Biorender.
 

### 2025-01-29 paper/stokes.tex

\documentclass[11pt, dvipsnames]{amsart}
\usepackage[utf8]{inputenc} 
\newcommand{\R}{\mathbb{R}}

\usepackage{bm}
\usepackage{caption}
\usepackage{subcaption}
% \captionsetup[subfigure]{justification=justified,singlelinecheck=false}
\usepackage{tikz}
\usepackage{newtxmath} % MER: Prettier math
\usepackage{cleveref} % MER: Allows \Cref (but use Section X.Y not Subsection X.Y)
\usepackage{booktabs} % MER: Prettier table

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{float}
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{placeins}
\usetikzlibrary{quotes}
\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,positioning,fit,petri}
\newtheorem{lemma}{Lemma}
\begin{document}

The time-independent, incompressible Stokes equations model the flow
of an incompressible Newtonian fluid at low Reynolds numbers, and read
as follows: over the domain $\Omega_{\rm CSF} \subset \R^3$, find the
velocity vector field $\bm u : \Omega_{\rm CSF} \rightarrow \R^3$ and
the pressure field $p : \Omega_{\rm CSF} \rightarrow \R$ such that
\begin{subequations}
  \begin{alignat}{2}
    - \mu \Delta \bm u  + \nabla p &=  0 \quad && \text{ in } \Omega_{\rm CSF},
    \label{eq:momentum_equation} \\ 
    \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF}
    \label{eq:divergence_equation} .
  \end{alignat}
  \label{eq:stokes}%
\end{subequations}%
In addition, we impose the following boundary conditions to model flow
induced by CSF production in the choroid plexus with $\Gamma_{\rm AM-U}$ as the main efflux site:
\begin{subequations}
  \begin{alignat}{2}
   \bm n^T (\mu \nabla \bm u \bm{n} - p \bm n ) &= -R_0 ( \bm u \cdot \bm n ), \quad \bm u \cdot \bm \tau = 0  
    && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
    \bm u &= 0 && \text{ on } \Gamma_{\mathrm{AM-L}}\cup\Gamma_{\rm{pia}}\cup\Gamma_{\rm{SSAS}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm \tau = 0 \quad && \text{ on } \Gamma_{\rm{LV}} ,  
  \end{alignat}
  \label{eq:stokes_bcs}%
\end{subequations}%
where $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the (dynamic) CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$. %  We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:app:dispersion}.

We consider an H(div)-based finite element approximation of the incompressible Stokes equations~\eqref{eq:stokes} defined over $\Omega_{\rm CSF}$ with the boundary conditions~\eqref{eq:stokes_bcs}. Following\cite{hong2016robust}, we approximate the velocity field $\bm u$ and the pressure field $p$ with the following finite element spaces:
\begin{align*}
  \bm V_{h,g} &= \{ \bm v  \in H(\mathrm{div}, \Omega_{\mathrm{CSF}}): \;
  \bm v \vert_{E} \in \mathrm{BDM}^2(E), \; E \in \mathcal{T}_{\rm CSF}; \; \\ & \quad 
  \bm v \cdot \bm n = 0
  \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\rm AM-L} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n =  g %\frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} 
  \text{ on } \Gamma_{\mathrm{LV}} \}  \\ 
  Q_h  &= \{q \in L^2(\Omega_{\rm CSF}): \; q \vert_E \in \mathrm{P}^1(E),
  \; E \in \mathcal{T}_{\rm CSF} \}. 
\end{align*}
Here, $H(\mathrm{div}, \Omega_{\rm{CSF}})$ is the space of $L^2(\Omega_{\rm{CSF}})$ vector fields with $L^2(\Omega_{\rm{CSF}})$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, $g$ is a given constant, and $\bm n$ is the unit outward normal vector to each facet. Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by 
\begin{equation}
\bm v_n = (\bm v \cdot \bm n) \bm n, \quad \bm v_t = \bm v - \bm v_n. 
\end{equation}
Since $\bm V_{h,g} \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm
  v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$. Continuity in the
tangential component is enforced weakly via interior penalization. For
convenience, we collect all facets exterior to the CSF space in  the set
$$\mathcal{F}_{e} = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
\cup \mathcal{F}_{\rm AM-L}\cup \mathcal{F}_{\rm AM-U}  \cup \mathcal{F}_{\rm SSAS}, $$ where facets
lying on the pial interface $\Gamma_{\rm pia}$ are denoted by
$\mathcal{F}_{\rm pia}$, on the lower and upper outer (arachnoid) boundary
$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$, on the inner ventricular
%on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$
and on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
$\mathcal{F}_{\rm SSAS}$ and $\mathcal{F}_{\rm LV}$ respectively. Now, define the form 
\begin{multline}
  \mathcal{A}_h(\bm u, \bm v)
  = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
 \\  - \sum_{F \in \mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}} \left (
  \int_{F} \mu \{\nabla \bm u\}  \bm n_F \cdot [\bm v_t] 
  - %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
  \int_{F} \mu \{\nabla \bm v\} \bm n_F \cdot [\bm u_t]
  + %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
\int_{F}   \frac{\sigma \mu}{h_F} [\bm u_t ] \cdot [\bm v_t] \right ),   
\end{multline}
where on exterior facets the average and jump operators take the one-sided values. We set the penalty parameter for the tangential continuity to be $\sigma = 20$. Our finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_{h,g} \times Q_h$ with $g  = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}$  such that 
\begin{subequations}
\label{eq:discrete_stokes}
\begin{align}
  \mathcal{A}_h(\bm u_h, \bm v_h)
  + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0 \quad \forall \bm v_h \in \bm V_{h,0} \\ 
\int_{\Omega_{\rm CSF}} \nabla \cdot \bm u_h \, q_h  &= 0 \quad \forall q_h \in Q_h.
\end{align}
\end{subequations}

\begin{lemma}[Consistency]
Let $\bm u \in H^2(\Omega)$ be the true solution of  \eqref{eq:stokes}. Then, \eqref{eq:discrete_stokes} holds with $\bm u$ replacing $\bm u_h$. 
\end{lemma}

\begin{proof}
 Multiply \eqref{eq:momentum_equation} by $\bm v_h \in \bm V_{h,0}$ and integrate by parts locally on each element. 
 \begin{align}
     \sum_{E \in \mathcal{T}} \int_E  (\mu \nabla \bm u : \nabla \bm v - \nabla \cdot \bm v_h p_h ) -  \sum_{E \in \mathcal{T}}  \int_{\partial E}  ( \mu \nabla u \bm n_E  - p \bm n_E) \cdot \bm v  = 0. 
  \end{align} 
The second term denoted here by $T$  can be written as follows.
\begin{align} \nonumber
T = -\sum_{F \in  \mathcal{F}_{i, \rm CSF}}  \int_{F}  \left( \{ \mu \nabla \bm u\} \bm n_F \cdot [\bm v]  -  \{ p \} \bm n_F \cdot [\bm v]  \right) \mathrm{d}s   -  \sum_{F \in \mathcal{F}_e}   \int_{F} ( \mu \nabla u \bm n   - p \bm n) \cdot \bm v \mathrm{d}s
\end{align}
Since $V_h $ is H-div conforming, $[\bm v ] = [\bm v_t] $ on $F \in \mathcal{F}_{i,\rm CSF}$. Therefore, the first term above, denoted by $T_1$, is written as 
\begin{align}
T_1 & = -  \sum_{F \in  \mathcal{F}_{i, \rm CSF}}  \int_{F}  \left( \{ \mu \nabla \bm u\} \bm n_F \cdot [\bm v_t]  -  \{ p \} \bm n_F \cdot [\bm v_t]  \right) \mathrm{d}s  \\ 
& = - \sum_{F \in  \mathcal{F}_{i, \rm CSF}}  \int_{F}  \{ \mu \nabla \bm u\} \bm n_F \cdot [\bm v_t].  \nonumber 
\end{align}
For the second term, we have that for $F \in \mathcal{F}_e$: 
\begin{align*}
\int_F (\mu \nabla \bm u \bm n - p \bm n) \cdot \bm v & = \int_F (\bm n^T (\mu \nabla \bm u \bm n - p \bm n)\bm n  +\bm t^T (\mu \nabla \bm u \bm n - p \bm n)\bm t ) \cdot (\bm v_n + \bm v_t) \\ 
&  = \int_F (\bm n^T (\mu \nabla \bm u \bm n - p \bm n)\bm n \cdot \bm v_n   +\bm t^T (\mu \nabla \bm u \bm n - p \bm n)\bm t ) \cdot  \bm v_t)
\end{align*}
 On $F \in \mathcal{F}_e \backslash \mathcal{F}_{\rm{AM-U}}$, the first term above is zero since $\bm v \cdot \bm n = 0 $ by the definition of $\bm V_{h,0}$.  For $F \in \mathcal{F}_{\rm{AM-U}}$, we use the boundary condition and write 
 \begin{align}
     \sum_{F \in \mathcal{F}_{\rm{AM-U}}} \int_F (\bm n^T (\mu \nabla \bm u \bm n - p \bm n)\bm n \cdot \bm v_n   &  = \sum_{F \in \mathcal{F}_{\rm{AM-U}}} \int_F - R_0 (\bm u \cdot \bm n)  \bm n \cdot \bm v_n  \\ 
     & = \sum_{F \in \mathcal{F}_{\rm{AM-U}}} \int_F - R_0 (\bm u \cdot \bm n)  \bm v \cdot \bm n  
 \end{align}
 Collecting the above computations, we see that 
\begin{multline}
 \nonumber
 \sum_{E \in \mathcal{T}} \int_E  (\mu \nabla \bm u : \nabla \bm v - \nabla \cdot \bm v_h p_h )   - \sum_{F \in  \mathcal{F}_{i, \rm CSF}}  \int_{F}  \{ \mu \nabla \bm u\} \bm n_F \cdot [\bm v_t] \\ - \sum_{F \in \mathcal{F}_{e}} \int_F \bm t^T (\mu \nabla \bm u \bm n - p \bm n)\bm t ) \cdot  \bm v_t) \
 + \sum_{F \in \mathcal{F}_{\rm{AM-U}}} \int_F  R_0 (\bm u \cdot \bm n)  \bm v \cdot \bm n    \nonumber = 0. 
\end{multline}
Now, since $\bm n^T (\nabla \bm u \bm n - p \bm n) \bm n \cdot \bm v_t = 0 $, we arrive at 
\begin{multline}
 \nonumber
 \sum_{E \in \mathcal{T}} \int_E  (\mu \nabla \bm u : \nabla \bm v - \nabla \cdot \bm v_h p_h )   - \sum_{F \in  \mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}}  \int_{F}  \{ \mu \nabla \bm u\} \bm n_F \cdot [\bm v_t] \\ 
 + \sum_{F \in \mathcal{F}_{\rm{AM-U}}} \int_F  R_0 (\bm u \cdot \bm n)  \bm v \cdot \bm n    \nonumber = 0. 
\end{multline}
The result is concluded by noting that $[\bm u \cdot \bm \tau ] = 0$ on  $\mathcal{F}_e \cup \mathcal{F}_{i,\mathrm{CSF}}$ from the associated boundary conditions and from the regularity of $\bm u$.s 
\end{proof} 
\bibliographystyle{plain}
\bibliography{references}
\end{document}

### 2025-01-29 paper/supplementary_main.tex
Updates from Overleaf

@@
 $\mathcal{F}_{\rm pia}$, on the lower and upper outer (arachnoid) boundary
-$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$ respectivley, 
+$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$ respectively, 
 %on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$

### 2025-01-30 paper/main.tex
Updates from Overleaf

@@
 \affil[1]{Department of Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway}
-\affil[2]{Brown University, Providence, Rhode Island, USA}
+\affil[2]{Division of Applied Mathematics, Brown University, Providence, Rhode Island, USA}
 \affil[3]{K. G. Jebsen Centre for Brain Fluid Research, University of Oslo, Norway}
@@
   demonstrate that low-frequency vasomotion induces moderate CSF flow
-  in surface PVS networks that substantially enhances tracer
-  enrichment, and that impaired enrichment is a direct
+  in surface PVS networks which substantially enhances tracer
+  enrichment, and that impaired enrichment is a direct natural
   consequence of enlarged PVSs. This openly available technology
@@
 perivascular enrichment, but that fluid flow induced by low-frequency
-vasomotion in surface periarterial spaces (on the order of 13 \textmu
+vasomotion in surface periarterial spaces (on the order of 10 \textmu
 m/s) is sufficient, even in the absence of a structural
@@
 effect of pulse wave pulsations in surface periarterial
-spaces~\cite{mestre2018flow, gjerde2023directional}
-(\Cref{fig:results1}D). Mathematically, this model is represented by a
+spaces~\cite{mestre2018flow, gjerde2023directional}. Mathematically, this model is represented by a
 mixed-dimensional system of coupled time-dependent partial
@@
 Simulating a glymphatic MRI protocol~\cite{ringstad2017glymphatic,
-  ringstad2018brain, eide2024functional}, we then predict the
+  ringstad2018brain, eide2024functional}(\Cref{fig:results1}D), we then predict the
 spreading of $0.5$ mmol intrathecally injected Gadobutrol after its
@@
 covers most of the brain surface (with the exception of some posterior
-regions) and has penetrated several centimeters into the parenchymal
+regions) and has penetrated substantially into the parenchymal
 tissue. This pattern is reflected in the mean tracer concentrations in
@@
 is more pronounced -- with up to a three-hour difference in time of
-arrival -- with PVS flow rates of approximately 10 \textmu m/s. Our
+arrival -- with mean PVS flow rates of approximately 10 \textmu m/s. Our
 models of CSF flow in perivascular networks, based on first principles
@@
 Comparing with our FTAs in the MCA segments, we note that the baseline model predicts a more-than 3 hour difference in FTA between the M1 and M2.
-Instead, the high PVS flow scenario predicts a 24 min delay between these segments.
+In contrast, the high PVS flow scenario predicts a 24 min delay between these segments.
 Moreover, the high PVS flow scenario predicts FTAs of
-84--108 min (M1) and 108--132 min (M2) after tracer appearance at the craniocervical junction, which overlaps with the previous ranges if one assumes a tracer travel time from the point of injection to the craniocervical junction of $13\,$min (as reported by Eide and Ringstad \cite{eide2024functional}).
+84--108 min (M1) and 108--132 min (M2) after tracer appearance at the craniocervical junction.
+Assuming a tracer travel time from the point of injection to the craniocervical junction of $13\,$min (as reported by Eide and Ringstad \cite{eide2024functional}), our predictions fall at the upper end of the clinical ranges.
+This can partly be attributed to the limited caudal extend of our PVS network, with the lowest point located approximately $4\,$cm above the craniocervical junction due to imaging constrains.
 We interpret these results as indicative of a net CSF flow in human surface PVS beyond what would be induced by CSF production and cardiac peristalsis alone.
@@
 permeability $\xi$ ($100\times$) at the the leaf nodes of
-$\Lambda_a$, linearly decreasing over the corresponding cell, to account for the transition from explicitly resolved
+$\Lambda_a$, linearly decreasing to the baseline permeability at the next network node ($\approx0.4\,$mm), to account for the transition from explicitly resolved
 PVSs to their unresolved continuation in the CSF spaces and

### 2025-01-30 paper/supplementary_main.tex
Updates from Overleaf

@@
   \begin{alignat}{2}
-( \mu \nabla \bm u \cdot \bm{n} - p \bm n) \cdot \bm n  &= -R_0 ( \bm u \cdot \bm n ), \quad \bm u \cdot \bm t = 0
+    ( \mu \nabla \bm u \cdot \bm{n} - p \bm n) \cdot \bm n  &= -R_0 ( \bm u \cdot \bm n ), 
+    \quad \bm u \cdot \bm t = 0
     && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
-    \bm u &= 0 && \text{ on } \Gamma_{\mathrm{AM-L}}\cup\Gamma_{\rm{pia}}\cup\Gamma_{\rm{SSAS}}, \\
+    \bm u &= 0 
+    && \text{ on } \Gamma_{\mathrm{AM-L}}\cup\Gamma_{\rm{pia}}\cup\Gamma_{\rm{SSAS}}, \\
     %\bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
     %\bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
-    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm t = 0 \quad && \text{ on } \Gamma_{\rm{LV}} ,  
+    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, 
+    \quad \quad \bm u \cdot \bm t = 0 \quad 
+    && \text{ on } \Gamma_{\rm{LV}} ,  
   \end{alignat}
@@
 $\mathcal{F}_{\rm pia}$, on the lower and upper outer (arachnoid) boundary
-$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$ respectively, 
+$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$, respectively, 
 %on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$

### 2025-01-30 paper/main.tex
Tweaks in the acknowledgments

@@
 generously sharing the multi-modal human MR data, and to Prof. Erlend
-Hodneland for facilitating this data exchange. M.E.R. acknowledges the
-support of the Foundation Kristian Gerhard Jebsen via the K.~G.~Jebsen
-Center for Brain Fluid Research and the Research Council of Norway via
-grant \#324239 (EMIx). Simulations were performed on the Experimental
-Infrastructure for Exploration of Exascale Computing (eX3), which is
-financially supported by the Research Council of Norway under contract
-270053. 
-M.C. acknowledges support from the RCN grant \#324239 (EMIx).
-M.K. acknowledges support from the RCN grant \#303362.
-The figures 1A, 1D, 2C, 3C, 5A and 6A were created in Biorender.
+Hodneland for facilitating this data exchange. M.E.R.~acknowledges
+support from Stiftelsen Kristian Gerhard Jebsen via the K.~G.~Jebsen
+Centre for Brain Fluid Research. M.E.R.~and M.C.~acknowledge support
+from the Research Council of Norway (RCN) via grant \#324239
+(EMIx). Simulations were performed on the Experimental Infrastructure
+for Exploration of Exascale Computing (eX3), which is financially
+supported by the RCN under contract 270053. M.K.~acknowledges support
+from the RCN grant \#303362. The figures 1A, 1D, 2C, 3C, 5A and 6A
+were created in Biorender.
 

### 2025-01-30 paper/main.tex
Remove line breaks in abstract for convenient copy-pasting.
Add Rami acknowledgement of support from EMIx.
Fix figure references

@@
 \begin{abstract}
-  The mechanisms of intracranial solute transport are fundamental to
-  human brain health, with alterations often linked to
-  disease and functional impairment, and with distinct opportunities
-  for personalized diagnostics and treatment. However, our
-  understanding of these mechanisms and their interplay remains
-  incomplete, in part due to the complexity of integrating insights
-  across scales, between species and from different modalities. Here,
-  we combine mixed-dimensional modelling, multi-modal magnetic 
-  resonance images, and high performance computing to construct and 
-  explore a high-fidelity in-silico model of human intracranial molecular enrichment. 
-  This model predicts the temporo-spatial spreading of a solute
-  within an image-derived geometric representation of the
-  subarachnoid space, ventricular system and brain parenchyma,
-  including networks of surface perivascular spaces
-  (PVSs). Our findings highlight the significant impact of
-  cerebrospinal fluid (CSF) production and intracranial pulsatility on
-  molecular enrichment following intrathecal tracer injection. We
-  demonstrate that low-frequency vasomotion induces moderate CSF flow
-  in surface PVS networks which substantially enhances tracer
-  enrichment, and that impaired enrichment is a direct natural
-  consequence of enlarged PVSs. This openly available technology
-  platform thus provides an opportunity for integrating separate
-  observations on diffusion in neuropil, vascular dynamics,
-  intracranial pulsatility, CSF production, and efflux, and for
-  exploring drug delivery and clearance in the human brain.
+  The mechanisms of intracranial solute transport are fundamental to human brain health, with alterations often linked to disease and functional impairment, and with distinct opportunities for personalized diagnostics and treatment.
+However, our understanding of these mechanisms and their interplay remains incomplete, in part due to the complexity of integrating insights across scales, between species and from different modalities.
+Here, we combine mixed-dimensional modelling, multi-modal magnetic resonance images, and high performance computing to construct and explore a high-fidelity in-silico model of human intracranial molecular enrichment.
+This model predicts the temporo-spatial spreading of a solute within an image-derived geometric representation of the subarachnoid space, ventricular system and brain parenchyma, including networks of surface perivascular spaces (PVSs).
+Our findings highlight the significant impact of cerebrospinal fluid (CSF) production and intracranial pulsatility on molecular enrichment following intrathecal tracer injection.
+We demonstrate that low-frequency vasomotion induces moderate CSF flow in surface PVS networks which substantially enhances tracer enrichment, and that impaired enrichment is a direct natural consequence of enlarged PVSs.
+This openly available technology platform thus provides an opportunity for integrating separate observations on diffusion in neuropil, vascular dynamics, intracranial pulsatility, CSF production, and efflux, and for exploring drug delivery and clearance in the human brain.
 \end{abstract}
@@
 field induced by CSF production at the periarterial network ends
-induces slow steady CSF flow of variable direction in these PVSs,
-with an average velocity of 0.08 \textmu m/s (antegrade) and a maximum
+induces slow steady CSF flow of variable direction in these PVSs, with
+an average velocity of 0.08 \textmu m/s (antegrade) and a maximum
 velocity of 0.55 \textmu m/s (\Cref{fig:pvs_a}B). On the other hand,
-traveling waves of arterial wall motion (\Cref{fig:pvs_a}C), such as the pulse wave or other vasomotion~\cite{vanveluw2020vasomotion,
-  munting2023spontaneous, bojarskaite2023sleep, broggini2024long,hauglund2025norepinephrine},
-also induce net directional flow in the
-PVS~\cite{kedarasetti2020arterial, kedarasetti2020functional,
-  coenen2021lubrication, gjerde2023directional, nozaleda2024arterial}
--- of magnitude and direction depending on the amplitude, frequency,
-and length of the waves and the characteristics of the perivascular
-network~\cite{gjerde2023directional}. Applying a
-semi-analytic model of the net flow induced by peristalsis in
-perivascular networks~\cite{gjerde2023directional} (see Methods), we
-estimate that the cardiac pulse wave alone, traveling at a frequency
-of 1 Hz with a wavelength of 2.0 m and a 1\% wall
-displacement~\cite{jung2021novel}, will induce mainly antegrade PVS
-flow with an average net velocity of 0.92 \textmu m/s while reaching up
-to 7.31 \textmu m/s near larger, ventral vessels such as the MCA
-(\Cref{fig:pvs_a}D). The same theory predicts that strong ultraslow
-vasomotions, if traveling antegrade at 0.1Hz with a wavelength of 0.02
-m and a 10\% displacement~\cite{broggini2024long}, will induce both
-retrograde and antegrade net PVS flow with an average velocity of
-13.05 \textmu m/s and maximum velocity 54.44 \textmu m/s
-(\Cref{fig:pvs_a}E).
+traveling waves of arterial wall motion (\Cref{fig:pvs_a}C), such as
+the pulse wave or other vasomotion~\cite{vanveluw2020vasomotion,
+  munting2023spontaneous, bojarskaite2023sleep,
+  broggini2024long,hauglund2025norepinephrine}, also induce net
+directional flow in the PVS~\cite{kedarasetti2020arterial,
+  kedarasetti2020functional, coenen2021lubrication,
+  gjerde2023directional, nozaleda2024arterial} -- of magnitude and
+direction depending on the amplitude, frequency, and length of the
+waves and the characteristics of the perivascular
+network~\cite{gjerde2023directional}. Applying a semi-analytic model
+of the net flow induced by peristalsis in perivascular
+networks~\cite{gjerde2023directional} (see Methods), we estimate that
+the cardiac pulse wave alone, traveling at a frequency of 1 Hz with a
+wavelength of 2.0 m and a 1\% wall displacement~\cite{jung2021novel},
+will induce mainly antegrade PVS flow with an average net velocity of
+0.92 \textmu m/s while reaching up to 7.31 \textmu m/s near larger,
+ventral vessels such as the MCA (\Cref{fig:pvs_a}D). The same theory
+predicts that strong ultraslow vasomotions, if traveling antegrade at
+0.1Hz with a wavelength of 0.02 m and a 10\%
+displacement~\cite{broggini2024long}, will induce both retrograde and
+antegrade net PVS flow with an average velocity of 13.05 \textmu m/s
+and maximum velocity 54.44 \textmu m/s (\Cref{fig:pvs_a}E).
 
@@
 (MCAs) and anterior cerebral artery (ACA), we found substantially
-reduced FTAs with higher PVS flow (\Cref{fig:pvs_a}B), up to 2.5 hours
+reduced FTAs with higher PVS flow (\Cref{fig:pvs_b}B), up to 2.5 hours
 earlier in the left M2 segment of the MCA. PVS concentrations peaked
@@
 in the PVS and subsequently spreads into the tissue and SAS
-(\Cref{fig:pvs_a}C, D), and especially so with higher PVS flow. These
+(\Cref{fig:pvs_b}C, D), and especially so with higher PVS flow. These
 observations also hold on an aggregated level: computing the total
@@
 support from Stiftelsen Kristian Gerhard Jebsen via the K.~G.~Jebsen
-Centre for Brain Fluid Research. M.E.R.~and M.C.~acknowledge support
-from the Research Council of Norway (RCN) via grant \#324239
+Centre for Brain Fluid Research. M.C., R.M., and M.E.R.~acknowledge
+support from the Research Council of Norway (RCN) via grant \#324239
 (EMIx). Simulations were performed on the Experimental Infrastructure

### 2025-02-25 paper/main.tex
Updates from Overleaf

@@
     \centering
-    \includegraphics[width=0.95\textwidth]{figures/figure1.png}
-     \caption{A) Illustration of the spatial domains of the model: CSF-filled space (ventricles and SAS) $\Omega_{\rm CSF}$, parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and the veins $\Lambda_{\rm vein}$ and their interfaces and boundaries (see Methods); B) 3D rendering of the actual geometry used in this study parenchyma, CSF (clipped, blue), arterial network (red), venous network (blue)); C) arterial network (red) and venous network (blue); D) illustration of the modelled clinical tracer injection procedure; E) time of simulated first tracer arrival after intrathecal injection (concentration exceeds $0.1\,$mol/l); F) time of first tracer arrival in the arterial PVS (top) and venous PVS (bottom); G) sagittal, coronal and axial view of computed tracer concentrations over T1-image 1, 6, 12 and 24\,h after injection (low concentrations transparent, pial surface in pink, arachnoid membrane in cyan, and arteries in dark red); H) total amount of tracer in each cranial compartment over the first 24\,h; I) average tracer concentration in each compartment over the first 24\,h.
+    \includegraphics[width=0.93\textwidth]{figures/figure1.png}
+     \caption{
+     \textbf{Our in-silico model predicts tracer transport in PVS, SAS and parenchyma following intrathecal injection.}\\
+     A) Illustration of the spatial domains of the model: CSF-filled space (ventricles and SAS) $\Omega_{\rm CSF}$, parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and the veins $\Lambda_{\rm vein}$ and their interfaces and boundaries (see Methods); B) 3D rendering of the actual geometry used in this study parenchyma, CSF (clipped, blue), arterial network (red), venous network (blue)); C) arterial network (red) and venous network (blue); D) illustration of the modelled clinical tracer injection procedure; E) time of simulated first tracer arrival after intrathecal injection (concentration exceeds $0.1\,$mol/l); F) time of first tracer arrival in the arterial PVS (top) and venous PVS (bottom); G) sagittal, coronal and axial view of computed tracer concentrations over T1-image 1, 6, 12 and 24\,h after injection (low concentrations transparent, pial surface in pink, arachnoid membrane in cyan, and arteries in dark red); H) total amount of tracer in each cranial compartment over the first 24\,h; I) average tracer concentration in each compartment over the first 24\,h.
 }
@@
 would translate from idealized geometries to human vascular networks
-and moreover evaluate their integrated effect in the context of
+and moreover, evaluate their integrated effect in the context of
 intracranial solute transport.
@@
 networks~\cite{gjerde2023directional} (see Methods), we estimate that
-the cardiac pulse wave alone, traveling at a frequency of 1 Hz with a
+the cardiac pulse wave alone, travelling at a frequency of 1 Hz with a
 wavelength of 2.0 m and a 1\% wall displacement~\cite{jung2021novel},
@@
 0.92 \textmu m/s while reaching up to 7.31 \textmu m/s near larger,
-ventral vessels such as the MCA (\Cref{fig:pvs_a}D). The same theory
-predicts that strong ultraslow vasomotions, if traveling antegrade at
+ventral vessels such as the MCA (\Cref{fig:pvs_a}D,E). The same theory
+predicts that strong ultraslow vasomotions, if travelling antegrade at
 0.1Hz with a wavelength of 0.02 m and a 10\%
@@
 antegrade net PVS flow with an average velocity of 13.05 \textmu m/s
-and maximum velocity 54.44 \textmu m/s (\Cref{fig:pvs_a}E).
+and maximum velocity 54.44 \textmu m/s (\Cref{fig:pvs_a}D,F).
 
@@
     \caption{\textbf{A static pressure gradient and peristaltic pumping are potential drivers of perivascular flow.}
-    A) Surface arterial network colored by arterial radius with vessel locations labeled;
-    B) Net PVS flow induced by CSF production (cf.~\Cref{fig:csf}D) -- histograms show the relative frequencies of local periarterial velocities.
-Positive values indicate antegrade PVS flow, while negative values indicate retrograde PVS flow;
+    A) Surface arterial network coloured by arterial radius with vessel locations labelled;
+    B) Net PVS flow induced by CSF production (cf.~\Cref{fig:csf}
+    -- histograms show the relative frequencies of local periarterial velocities. Positive values indicate antegrade PVS flow, while negative values indicate retrograde PVS flow;
     C) Illustration of the concept of peristaltic pumping;
-    D) Estimated net PVS flow induced by pulse wave peristaltic pumping;
-    E) Estimated net PVS flow induced by vasomotion/slow wave peristaltic pumping.}
+    D) histograms of estimated net PVS flow induced by pulse wave or vasomotion peristaltic pumping;
+    E) Estimated net PVS flow induced by pulse wave peristaltic pumping.
+    F) Estimated net PVS flow induced by vasomotion/slow wave peristaltic pumping.}
     \label{fig:pvs_a}

### 2025-02-25 paper/main.tex
Updates from Overleaf

@@
     \centering
-    \includegraphics[width=0.93\textwidth]{figures/figure1.png}
+    \includegraphics[width=0.98\textwidth]{figures/figure1.png}
      \caption{
      \textbf{Our in-silico model predicts tracer transport in PVS, SAS and parenchyma following intrathecal injection.}\\
-     A) Illustration of the spatial domains of the model: CSF-filled space (ventricles and SAS) $\Omega_{\rm CSF}$, parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and the veins $\Lambda_{\rm vein}$ and their interfaces and boundaries (see Methods); B) 3D rendering of the actual geometry used in this study parenchyma, CSF (clipped, blue), arterial network (red), venous network (blue)); C) arterial network (red) and venous network (blue); D) illustration of the modelled clinical tracer injection procedure; E) time of simulated first tracer arrival after intrathecal injection (concentration exceeds $0.1\,$mol/l); F) time of first tracer arrival in the arterial PVS (top) and venous PVS (bottom); G) sagittal, coronal and axial view of computed tracer concentrations over T1-image 1, 6, 12 and 24\,h after injection (low concentrations transparent, pial surface in pink, arachnoid membrane in cyan, and arteries in dark red); H) total amount of tracer in each cranial compartment over the first 24\,h; I) average tracer concentration in each compartment over the first 24\,h.
-}
+     A) Illustration of the spatial domains of the model: CSF-filled space (ventricles and SAS) $\Omega_{\rm CSF}$, parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and the veins $\Lambda_{\rm vein}$ and their interfaces and boundaries (see Methods);
+     B) 3D rendering of the actual geometry used in this study parenchyma, CSF (clipped, blue), arterial network (red), venous network (blue)); 
+     C) illustration of the modelled clinical tracer injection procedure; 
+     D) simulated tracer concentration after 1, 6, 12, and 24 hours (opacity increases linearly from 0 to $2\,$mol/l); 
+     E) sagittal, coronal and axial view of computed tracer concentrations over T1-image after 1, 6, 12 and 24\,h (low concentrations transparent, pial surface in pink, arachnoid membrane in cyan, and arteries in dark red);
+     F) total amount of tracer in each cranial compartment over the first 24\,h; 
+     G) average tracer concentration in each compartment over the first 24\,h. }
      \label{fig:results1}
@@
 embedded networks of topologically one-dimensional (1D) curves
-(\Cref{fig:results1}A--C). We consider a solute concentration field,
+(\Cref{fig:results1}A--B). We consider a solute concentration field,
 varying in space and time, in the 3D domains and in the PVS networks,
@@
 Simulating a glymphatic MRI protocol~\cite{ringstad2017glymphatic,
-  ringstad2018brain, eide2024functional}(\Cref{fig:results1}D), we then predict the
+  ringstad2018brain, eide2024functional}(\Cref{fig:results1}C), we then predict the
 spreading of $0.5$ mmol intrathecally injected Gadobutrol after its
 appearance at the craniocervical junction
-(\Cref{fig:results1}E--G). After one hour, the in-silico
+(\Cref{fig:results1}D--E). After one hour, the in-silico
 tracer moves upwards in the SAS frontally of the brainstem, and
@@
 concentration peak ($2.9\,$mmol/l after 4 hours)
-(\Cref{fig:results1}I). While the mean concentrations in these
+(\Cref{fig:results1}G). While the mean concentrations in these
 compartments drop soon after peaking, the parenchymal tissue

### 2025-02-25 paper/main.tex
update Fig1 caption and references

@@
     \centering
-    \includegraphics[width=0.98\textwidth]{figures/figure1.png}
+    \includegraphics[width=1\textwidth]{figures/figure1.png}
      \caption{
@@
      E) sagittal, coronal and axial view of computed tracer concentrations over T1-image after 1, 6, 12 and 24\,h (low concentrations transparent, pial surface in pink, arachnoid membrane in cyan, and arteries in dark red);
-     F) total amount of tracer in each cranial compartment over the first 24\,h; 
-     G) average tracer concentration in each compartment over the first 24\,h. }
+     F) average tracer concentration in each compartment over the first 24\,h. 
+     G) total amount of tracer in each cranial compartment over the first 24\,h; }
      \label{fig:results1}
@@
 concentration peak ($2.9\,$mmol/l after 4 hours)
-(\Cref{fig:results1}G). While the mean concentrations in these
+(\Cref{fig:results1}F). While the mean concentrations in these
 compartments drop soon after peaking, the parenchymal tissue
@@
 venous PVS (8\%) and the arterial PVS
-(5\%) (\Cref{fig:results1}H).
+(5\%) (\Cref{fig:results1}G).
 
@@
 network ends located within the SAS or up to 6 mm inside the
-parenchyma (\Cref{fig:results1}A--C, \Cref{fig:pvs_a}A) and includes
+parenchyma (\Cref{fig:results1}A--B, \Cref{fig:pvs_a}A) and includes
 major surface arteries of radius 0.5--1.4 mm. Imposing the pressure
@@
 centerline of a blood vessel segment and labeled with its lumen radius
-$R_1^i$ (\Cref{fig:results1}C, \Cref{fig:pvs_a}A). We also assign an
+$R_1^i$ (\Cref{fig:pvs_a}A). We also assign an
 outer radius $R_2^i > R_1^i$ to each segment $i$, form the annular
@@
 a mixed-dimensional transport model~\cite{masri2024modelling,laurino2019derivation} over a
-timescale of minutes to hours (\Cref{fig:results1}A--C,
+timescale of minutes to hours (\Cref{fig:results1}A--B,
 \Cref{fig:csf}A--C). Specifically, for $t > 0$, we solve for a

### 2025-02-25 paper/supplementary_main.tex
update Fig1 caption and references

@@
 edge-disjoint subnetworks $\Lambda^1, \Lambda^2, \Lambda^3$, one for
-each of the supply nodes (\Cref{fig:results1}A, C,
+each of the supply nodes (\Cref{fig:results1}A--B,
 \Cref{fig:app:peristalsis}). Each node is assigned to the subnetwork

### 2025-02-27 paper/main.tex
Split supplementary into separate file with independent references...

@@
 The PVSs are recognized across species as critical pathways for solute
-transport in and around the brain\cite{}, and thus as potential
-targets for enhancing brain drug delivery and metabolic waste
-clearance\cite{}. However, whether CSF flows more rapidly in PVSs and
-what the forces and mechanisms required to drive such flow are, remain
-as key points of debate~\cite{bohr2022glymphatic,
-  van2024caa}. Motivated by experimental observations in animal
+transport in and around the brain, and thus as potential targets for
+enhancing brain drug delivery and metabolic waste clearance. However,
+whether CSF flows more rapidly in PVSs and what the forces and
+mechanisms required to drive such flow are, remain as key points of
+debate~\cite{bohr2022glymphatic, van2024caa}. Motivated by
+experimental observations in animal
 models~\cite{iliff2012paravascular, iliff2013cerebral, mestre2018flow,
@@
 networks~\cite{gjerde2023directional} (see Methods), we estimate that
-the cardiac pulse wave alone, travelling at a frequency of 1 Hz with a
+the cardiac pulse wave alone, traveling at a frequency of 1 Hz with a
 wavelength of 2.0 m and a 1\% wall displacement~\cite{jung2021novel},
@@
 ventral vessels such as the MCA (\Cref{fig:pvs_a}D,E). The same theory
-predicts that strong ultraslow vasomotions, if travelling antegrade at
+predicts that strong ultraslow vasomotions, if traveling antegrade at
 0.1Hz with a wavelength of 0.02 m and a 10\%
@@
     \caption{\textbf{A static pressure gradient and peristaltic pumping are potential drivers of perivascular flow.}
-    A) Surface arterial network coloured by arterial radius with vessel locations labelled;
+    A) Surface arterial network colored by arterial radius with vessel locations labeled;
     B) Net PVS flow induced by CSF production (cf.~\Cref{fig:csf}
@@
 complete description of this mathematical model and simulation
-scenarios, see \Cref{sec:app:more_maths}. The initial concentrations
+scenarios, see the Supplementary information S1. The initial concentrations
 in the three-dimensional domain $\Omega$ and networks $\Lambda_a,
@@
 (\Cref{tab:parameters}, \Cref{fig:csf}A--E, see also
-\Cref{sec:app:more_maths}). We do not model bulk flow within the brain
+the Supplementary information (S1)). We do not model bulk flow within the brain
 parenchyma, except within the perivascular network segments extending
@@
 dispersion \cite{taylor1953dispersion, watson1983diffusion,
-  sharp2019dispersion} (see \Cref{sec:app:dispersion} for a complete
-description). Combining these effects, we then set $R = R_c + R_r$ and
-define the effective diffusion coefficient $D = (1 + R) D^{\rm Gad}$
-in $\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport} in the baseline model.
+  sharp2019dispersion} (see Supplementary information S1.3 for a
+complete description). Combining these effects, we then set $R = R_c +
+R_r$ and define the effective diffusion coefficient $D = (1 + R)
+D^{\rm Gad}$ in $\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport} in
+the baseline model.
 
@@
 flow velocity in the periarterial network (see
-also~\Cref{sec:app:hydraulic_network}). We repeat this procedure to
+also Supplementary information S1.1.4). We repeat this procedure to
 compute a corresponding velocity $\hat{u}^p_v$ in the perivenous
@@
 area of the PVS segment and $\langle Q_i' \rangle$ is the estimated
-mean flow rate (see~\ref{sec:app:peristalsis} for more details). Two
+mean flow rate (see Supplementary information S1.4 for more details). Two
 wall motion patterns are considered, yielding $\hat{u}_a^c$ and
@@
 spaces ($\xi = \xi_{\rm CSF}$) or parenchyma ($\xi = \xi_{\rm EF}$),
-or neither (\Cref{sec:app:more_maths}). Additionally, we increase the
-permeability $\xi$ ($100\times$) at the the leaf nodes of
-$\Lambda_a$, linearly decreasing to the baseline permeability at the next network node ($\approx0.4\,$mm), to account for the transition from explicitly resolved
-PVSs to their unresolved continuation in the CSF spaces and
-parenchyma.
+or neither (see also Supplementary information S1). Additionally, we
+increase the permeability $\xi$ ($100\times$) at the the leaf nodes of
+$\Lambda_a$, linearly decreasing to the baseline permeability at the
+next network node ($\approx0.4\,$mm), to account for the transition
+from explicitly resolved PVSs to their unresolved continuation in the
+CSF spaces and parenchyma.
 \begin{table}
@@
 and accuracy of the numerical solutions were verified by a series of
-numerical verification experiments (\Cref{sec:app:verification}).
+numerical verification experiments (see Supplementary information S1.5).
 
@@
 estimates, and the shapes and sizes of PVSs) with the current
-literature (\Cref{sec:app:model_validation}).
+literature (see Supplementary information S2.1).
 
@@
 
-\newpage
-\FloatBarrier
-\appendix
-\input{supplementary_main.tex}
+%% \newpage
+%% \FloatBarrier
+%% \appendix
+%% \input{supplementary_main.tex}
 

### 2025-02-27 paper/supplementary.tex
Split supplementary into separate file with independent references...

--- 2025-02-25 paper/main.tex
@@
 \usepackage{bm}
-\usepackage{caption}
 \usepackage{subcaption}
@@
 
+\renewcommand\thefigure{S\arabic{figure}}    
+\renewcommand\thesection{S\arabic{section}}
+
 \captionsetup[subfigure]{justification=justified,singlelinecheck=false}
 
-\title{In-silico molecular enrichment and clearance of the human intracranial space}
+\title{Supplementary information: In-silico molecular enrichment and clearance of the human intracranial space}
 \author[1]{Marius Causemann}
@@
 
-\newcommand{\rami}[1]{\textcolor{blue}{#1}}
-\newcommand{\mer}[1]{\textcolor{magenta}{#1}}
-\newcommand{\mar}[1]{\textcolor{violet}{#1}}
-\newcommand{\mk}[1]{\textcolor{orange}{#1}}
-\newcommand{\discuss}[1]{\textcolor{red}{#1}}
-\newcommand{\fixme}[1]{\textcolor{red}{#1}}
-\newcommand{\draft}[1]{\textcolor{lightgray}{#1}}
+\begin{abstract}
+  Supplementary methods and discussion accomanying the main text of
+  \emph{In-silico molecular enrichment and clearance of the human
+  intracranial space}.
+\end{abstract}
 
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\foralls}{\forall \,}
-
-\begin{abstract}
-  The mechanisms of intracranial solute transport are fundamental to human brain health, with alterations often linked to disease and functional impairment, and with distinct opportunities for personalized diagnostics and treatment.
-However, our understanding of these mechanisms and their interplay remains incomplete, in part due to the complexity of integrating insights across scales, between species and from different modalities.
-Here, we combine mixed-dimensional modelling, multi-modal magnetic resonance images, and high performance computing to construct and explore a high-fidelity in-silico model of human intracranial molecular enrichment.
-This model predicts the temporo-spatial spreading of a solute within an image-derived geometric representation of the subarachnoid space, ventricular system and brain parenchyma, including networks of surface perivascular spaces (PVSs).
-Our findings highlight the significant impact of cerebrospinal fluid (CSF) production and intracranial pulsatility on molecular enrichment following intrathecal tracer injection.
-We demonstrate that low-frequency vasomotion induces moderate CSF flow in surface PVS networks which substantially enhances tracer enrichment, and that impaired enrichment is a direct natural consequence of enlarged PVSs.
-This openly available technology platform thus provides an opportunity for integrating separate observations on diffusion in neuropil, vascular dynamics, intracranial pulsatility, CSF production, and efflux, and for exploring drug delivery and clearance in the human brain.
-\end{abstract}
 \begin{document}
@@
 
-\thispagestyle{empty}
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\foralls}{\forall \,}
 
-%\mer{MER: Target length: 5000 words (Intro, Results, Discussion) plus <3000 words Methods: 700 words Introduction. 3100 words Results. 1200 words Discussion.}
-
-\section*{Introduction}
-
-% What is the big picture problem and why is it important? 
-The mechanisms underlying molecular transport within the intracranial
-space are fundamental to human brain health and
-function. Neurodegenerative diseases such as Alzheimer's, Parkinson's,
-and Huntington's disease are all associated with abnormal accumulation
-of protein aggregates together with alterations in transport and
-clearance characteristics~\cite{rasmussen2018glymphatic,
-  harrison2020impaired, eide2023plasma, liu2024glymphatic}. Moreover,
-sleep and conversely sleep-deprivation play a definite yet enigmatic
-role in modulating molecular enrichment and
-clearance~\cite{xie2013sleep, eide2021sleep, eide2022altered,
-  miao2024brain, hauglund2025norepinephrine}. In the last decade,
-established theories have been challenged by new findings revealing a
-greater degree of molecular movement and
-exchange~\cite{iliff2012paravascular, ringstad2017glymphatic,
-  louveau2017understanding, proulx2021cerebrospinal,
-  bohr2022glymphatic}, as well as substantial variability in
-enrichment and clearance between individuals and between patient
-cohorts~\cite{ringstad2018brain, eide2021direction, eide2021impaired,
-  eide2022altered}. These observations provide distinct opportunities
-(and challenges) for personalized medicine e.g.~for tailored
-intrathecal delivery of chemotherapy~\cite{lohela2022glymphatic} and
-for early diagnostics of impaired brain
-clearance~\cite{eide2021clinical, van2024human}. In spite of their
-importance, our understanding of these mechanisms is incomplete with
-open challenges and significant debate -- in part relating to the
-translation of knowledge between scales, species, experimental
-protocols, clinical cohorts, and individuals.
-
-Perivascular pathways along the brain surface and within the brain
-parenchyma have long been hypothesized to serve a designated role in
-this context~\cite{rennels1985evidence, zhang1990interrelationships,
-  ichimura1991distribution, carare2008solutes, iliff2012paravascular,
-  foley2012realtime, hannocks2018molecular, van2024caa}. Recently, Eide and
-Ringstad~\cite{eide2024functional} and Yamamoto et
-al~\cite{yamamoto2024perivascular} demonstrated that perivascular
-spaces (PVSs) define preferential pathways for molecular transport in
-humans, with delayed periarterial enrichment in dementia
-subtypes~\cite{eide2024functional}. Perivascular flow of cerebrospinal
-fluid (CSF) clearly contributes to this transport, and is inherently
-associated with vascular pulsations~\cite{hadaczek2006perivascular,
-  iliff2013cerebral, bedussi2018paravascular, mestre2018flow,
-  boster2023artificial, hirschler2024region}. Fluid mechanics
-considerations point at intracranial pressure differences and shorter
-wavelength vascular wall pulsations as drivers of directional net flow
-and convection in the PVS~\cite{bilston2003arterial, rey2018pulsatile,
-  daversin2020mechanisms, kedarasetti2020functional,
-  gjerde2023directional, nozaleda2024arterial}, while longer waves
-such as the pulse wave primarily contribute to oscillatory flow and
-dispersion~\cite{asgari2016glymphatic, sharp2019dispersion, thomas2019fluid,
-  kedarasetti2020arterial, troyetsky2021dispersion, 
-  martinac2021phase}. In the bigger picture, CSF is produced by the
-choroid plexus~\cite{damkier2013cerebrospinal,
-  steffensen2018cotransporter, liu2020direct}, pulsates through the
-ventricular system, cisterns, and subarachnoid space (SAS) in
-synchrony with cardiac, respiratory, and neural
-waves~\cite{greitz1993pulsatile, wagshul2011pulsating,
-  sweetman2011cerebrospinal, fultz2019coupled, vinje2019respiratory,
-  eide2021direction, causemann2022human, williams2023neural,
-  zimmermann2023total}, and drains via the dural sinuses, meningeal
-lymphatics, cranial nerves, or other efflux
-pathways~\cite{proulx2021cerebrospinal}. However, how these
-physiological factors and physical mechanisms integrate to enhance or
-impair human intracranial molecular transport over larger spatial
-scales and longer time scales remain unknown. A related key question
-is to what extent surface PVSs are separated from the SAS by
-structural barriers~\cite{zhang1990interrelationships,
-  weller2005microscopic, bedussi2017paravascular,
-  pizzo2018intrathecal, mestre2022periarteriolar,
-  mollgard2023mesothelium, smets2024perivascular, eide2024functional},
-and in turn to what extent such structural compartmentalization is a
-prerequisite for effective perivascular transport.
-
-% What are our research questions and how do we answer them
-In this study, by leveraging geometric model reduction and
-mixed-dimensional modelling~\cite{masri2024modelling}, structural
-magnetic resonance (MR) images~\cite{hodneland2019new}, and high
-performance computing, we introduce an integrated computational model
-of intracranial molecular enrichment and clearance. Focusing on the
-interplay between perivascular pathways, pulsatility and CSF flow
-dynamics, the model predicts the temporal evolution and spatial
-distribution of a solute concentration within a detailed geometric
-representation of the human SAS and ventricular system, networks of
-surface PVSs and the brain parenchyma. In terms of transport dynamics,
-we account for heterogeneous diffusion, dispersive mixing induced by
-cardiac and respiratory pulsatility in the CSF spaces and PVSs,
-convective fluid flow driven by CSF production and peristaltic
-pumping, as well as solute exchange and clearance across
-semi-permeable membranes. Comparing with glymphatic MRI
-studies~\cite{ringstad2017glymphatic, ringstad2018brain,
-  watts2019measuring}, the in-silico predictions accurately represent
-molecular enrichment patterns, timing and intercompartmental
-distributions. This open platform~\cite{ZENODO} thus provides a
-technological opportunity for qualitatively and quantitatively
-exploring key open questions relating to molecular movement within the
-human brain environment such as the role of pulsatility, perivascular
-pathways, structural compartmentalization, or morphology.
-
-% Main findings and outlook
-By exploring this high-dimensional parameter space, we propose that
-the balance between CSF production and intracranial pulsatility is key
-to shaping the large-scale features of intracranial enrichment
-patterns, with the potential to span a wide range of individual and
-cohort variability. Moreover, we predict that CSF production, cardiac-
-and respiratory pulsatility is not sufficient to explain early
-perivascular enrichment, but that fluid flow induced by low-frequency
-vasomotion in surface periarterial spaces (on the order of 10 \textmu
-m/s) is sufficient, even in the absence of a structural
-compartmentalization of the PVS. Conversely, enlarged PVSs in the SAS
-will cause a substantial reduction in cardiac- and vasomotion-driven
-flow velocities, strongly delay perivascular transport, and thus
-impair intracranial enrichment. These findings transfer, reconcile,
-integrate and extend insights from clinical, experimental, and
-theoretical studies, and lay a new foundation for in-silico studies of
-personalized intrathecal drug delivery and brain clearance.
-
-\section*{Results}
-
-\subsection*{In-silico predictions of intracranial molecular enrichment and clearance after intrathecal injection}
-\begin{figure}
-    \centering
-    \includegraphics[width=1\textwidth]{figures/figure1.png}
-     \caption{
-     \textbf{Our in-silico model predicts tracer transport in PVS, SAS and parenchyma following intrathecal injection.}\\
-     A) Illustration of the spatial domains of the model: CSF-filled space (ventricles and SAS) $\Omega_{\rm CSF}$, parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and the veins $\Lambda_{\rm vein}$ and their interfaces and boundaries (see Methods);
-     B) 3D rendering of the actual geometry used in this study parenchyma, CSF (clipped, blue), arterial network (red), venous network (blue)); 
-     C) illustration of the modelled clinical tracer injection procedure; 
-     D) simulated tracer concentration after 1, 6, 12, and 24 hours (opacity increases linearly from 0 to $2\,$mol/l); 
-     E) sagittal, coronal and axial view of computed tracer concentrations over T1-image after 1, 6, 12 and 24\,h (low concentrations transparent, pial surface in pink, arachnoid membrane in cyan, and arteries in dark red);
-     F) average tracer concentration in each compartment over the first 24\,h. 
-     G) total amount of tracer in each cranial compartment over the first 24\,h; }
-     \label{fig:results1}
-\end{figure}
-Using previously published multi-modal magnetic resonance imaging
-(MRI) data~\cite{hodneland2019new, deistung2009tof,
-  schweser2012quantitative, reichenbach2012future,
-  deistung2017overview}, we construct a multiscale computational
-representation of the human intracranial compartments consisting of
-the CSF spaces and brain parenchyma as three-dimensional (3D) domains
-and with the PVSs surrounding major surface arteries and veins as
-embedded networks of topologically one-dimensional (1D) curves
-(\Cref{fig:results1}A--B). We consider a solute concentration field,
-varying in space and time, in the 3D domains and in the PVS networks,
-and assume that the solute can cross between these compartments
-through semi-permeable membranes. As the drivers and modes of
-intracranial transport are under substantial
-debate\cite{smith2019going, proulx2021cerebrospinal,
-  bohr2022glymphatic, hladky2022glymphatic, betsholtz2024advances},
-our first target is to establish a baseline model accounting for a
-reasonably conservative set of mechanisms and their integrated effect
-over a timescale of several minutes to a few days. To this end, we
-assume that the solute will (i) diffuse within all compartments, with
-diffusivity depending on the effective properties of the relevant
-medium\cite{sykova2008diffusion}; (ii) experience significant
-dispersive effects due to the pulsatile flow of CSF induced by the
-cardiac and respiratory cycles\cite{vinje2019respiratory,
-  sharp2019dispersion, ray2021quantitative, troyetsky2021dispersion};
-and (iii) be convected by a (small) net flow of CSF resulting from
-production in the choroid plexus with CSF efflux across the upper
-convexity\cite{hornkjol2022csf} and from the peristaltic pumping
-effect of pulse wave pulsations in surface periarterial
-spaces~\cite{mestre2018flow, gjerde2023directional}. Mathematically, this model is represented by a
-mixed-dimensional system of coupled time-dependent partial
-differential equations\cite{masri2024modelling}, which we solve
-numerically with high accuracy using a mass-conserving finite element
-scheme and the FEniCS finite element software\cite{alnaes2015fenics,
-  kuchta2020assembly} (see Methods). The computational framework and
-associated software are all openly available~\cite{ZENODO}.
-
-Simulating a glymphatic MRI protocol~\cite{ringstad2017glymphatic,
-  ringstad2018brain, eide2024functional}(\Cref{fig:results1}C), we then predict the
-spreading of $0.5$ mmol intrathecally injected Gadobutrol after its
-appearance at the craniocervical junction
-(\Cref{fig:results1}D--E). After one hour, the in-silico
-tracer moves upwards in the SAS frontally of the brainstem, and
-quickly reaches the supratentorial regions. Here, it spreads both
-posterior through the quadrigeminal cistern and the longitudinal
-fissure, and anteriorly through the outer SAS, reaching the top of the
-cerebral cortex after around 12 hours. After 24 hours, the tracer
-covers most of the brain surface (with the exception of some posterior
-regions) and has penetrated substantially into the parenchymal
-tissue. This pattern is reflected in the mean tracer concentrations in
-each compartment, where the CSF space reaches its peak of
-$1.4\,$mmol/l after 2 hours, followed by the arterial PVS
-concentration peak ($2.9\,$mmol/l after 4 hours)
-(\Cref{fig:results1}F). While the mean concentrations in these
-compartments drop soon after peaking, the parenchymal tissue
-slowly enriches with tracer over the first 24 hours, with a final
-value of $0.3$ mmol/l. Overall, about 40\% of the total amount of
-tracer remains in the cranium 24 hours post-injection, with the largest
-share in the CSF (53\%), followed by the parenchyma (35\%), the
-venous PVS (8\%) and the arterial PVS
-(5\%) (\Cref{fig:results1}G).
-
-\subsection*{Reduced CSF pulsatility strongly shifts intracranial enrichment patterns}
-\begin{figure}%[h!]
-\centering 
-\includegraphics[width = 0.95\textwidth]{figures/figure2.png}
-\caption{\textbf{The balance between CSF pulsatility and production shapes molecular enrichment patterns.}
-A) CSF spaces: the third and fourth ventricle (V3\&4), lateral ventricles (LV) and SAS; 
-B) Boundaries and interfaces: the spinal subarachnoid space (SSAS), upper and lower arachnoid membrane (AM-U and AM-L), pia (Pia) and lateral ventricles (LV); 
-C) Schematic of the CSF flow model; 
-D) CSF flow induced by CSF production with outflow allowed across the
-upper convexity (AM-U); 
-}
-\label{fig:csf}
-\end{figure}
-\begin{figure}
-\ContinuedFloat
-\caption{(\textbf{cont.})
-E) Histograms of the production-induced CSF velocities in the SAS and third and fourth ventricle (V3 \& V4) in terms of the relative frequency of the mean flow speed in each computational cell weighted by its volume;
-F) CSF flow induced at peak systolic blood inflow; 
-G) Cardiac-induced dispersion factor $R_c$;
-H) CSF flow induced at peak respiratory expansion; 
-I) Respiration-induced dispersion factor $R_r$, $D = (1 + R_c + R_r) D^{\rm Gad}$; J--M): Histograms of the cardiac-induced flow velocities (J),
-cardiac-induced dispersion factors (K),
-respiratory flow velocities (L), respiratory dispersion factor (M);
-N--Q): sagittal, coronal and axial planes of tracer concentration after 12h for different model variations: baseline (N), no CSF production (O), low dispersion (P), and high dispersion (Q).
-}
-\end{figure}
-
-The enrichment patterns observed clinically after intrathecal
-injection of contrast differ substantially between subjects and
-between pathological conditions~\cite{ringstad2018brain,
-  eide2021direction, eide2021impaired, eide2022altered}. These
-neurological conditions are also associated with alterations in the
-pulsatile flow of CSF in the ventricular system and
-SAS~\cite{eide2021direction}. Clearly, key physiological factors such
-as cardiac pulsatility and respiration easily differ between
-individuals and between both pathological and physiological states. We
-therefore next asked to what extent -- and how -- variations in CSF
-pulsatility would affect the intracranial enrichment characteristics.
-
-For the CSF dynamics in our baseline model, we combine the net
-contribution to CSF flow induced by CSF production with the integrated
-dispersive effects of cardiac and respiratory pulsatile CSF flow
-(\Cref{fig:csf}A--C). Imposing a constant CSF production of
-$400\,$ml/day across the surface of the lateral ventricles, while
-allowing for efflux across the upper convexity, yields a total CSF
-pressure drop of $26\,$mPa (0.00020 mmHg) (\Cref{fig:csf}D) with a
-maximum flow velocity of $1.85\,$mm/s (in the aqueduct) and a mean
-velocity in the SAS of $2.6$\textmu m/s (\Cref{fig:csf}E). To examine the
-dispersive effects of pulsatile CSF flow, we employ the theory of
-shear-augmented dispersion together with computational fluid dynamics
-to determine a dispersion factor $R$ enhancing the effective solute
-diffusivity (see Methods). For the cardiac contribution, we compute
-CSF pressure and flow fields at peak systolic blood inflow,
-corresponding to a reduction in CSF space volume at a total rate of
-$6\,$ml/s~\cite{baledent2014imaging, causemann2022human} in the SAS
-and $0.31\,$ml/s across the lateral ventricle
-surface\cite{vinje2019respiratory}. This scenario sets up a pressure
-drop of $10\,$Pa ($0.075$ mmHg) between the lateral ventricles and the
-spinal SAS and a maximum flow velocity of $19.8\,$cm/s
-(\Cref{fig:csf}F, J). Assuming a cardiac frequency of $1\,$Hz, we
-infer that this cardiac-induced pulsatile CSF flow increases the
-effective diffusion by more than two orders of magnitude in the
-aqueduct and near the cisterna magna, but has little effect ($R <
-1$) in most of the SAS (\Cref{fig:csf}G, K). For the respiratory
-contribution, we employ the same methodology, but with a total rate of
-$1\,$ml/s \cite{gutierrez2022effect} in the SAS and $0.121\,$ml/s
-\cite{liu2024using} in the lateral ventricles, yielding a respiratory
-peak flow volume of $1.121\,$ ml/s at the craniocervical
-junction. While the resulting flow velocities are only about one
-fourth of their cardiac-induced counterparts (\Cref{fig:csf}H, L),
-respiratory dispersive mixing reaches a factor of up to 320 due the
-lower respiratory frequency of $0.25\,$Hz (\Cref{fig:csf}I, M).
-
-Now, to examine how CSF pulsatility affects the enrichment patterns,
-we consider three variations of the baseline: (i) no CSF production,
-(ii) reduced pulsatility and thus decreased dispersion ($0.1 \times
-R$), and (iii) higher pulsatility with increased dispersion ($10
-\times R$). Without CSF production, transport is considerably delayed
-(\Cref{fig:csf}N, O). Even after $12$ hours, tracer remains in the
-subtentorial regions around the cerebellum, the brain stem and in the
-surrounding CSF. Interestingly, the lack of CSF production instead
-allows tracer to travel upwards through the ventricular system,
-reaching the third ventricle after around $12$ hours. On the other
-hand, if the CSF pulsatility is reduced, we observe rapid transport
-towards the upper convexity of the cranium, as in the baseline model,
-but the tracer spreads exclusively within the anterior regions
-(\Cref{fig:csf}P). This feature can be attributed to the CSF flow
-bifurcation posterior to the ambient cistern (\Cref{fig:csf}D):
-without sufficient diffusion, the tracer is unable to cross into the
-posterior SAS. Indeed, with higher dispersion, the tracer moves
-through the quadrigeminal cistern and further upwards into the
-longitudinal fissure, with also enrichment of the cerebellum
-(\Cref{fig:csf}Q).
-
-\subsection*{Perivascular flow shapes and accelerates molecular enrichment}
-\label{sec:pvs_flow_results}
-
-The PVSs are recognized across species as critical pathways for solute
-transport in and around the brain\cite{}, and thus as potential
-targets for enhancing brain drug delivery and metabolic waste
-clearance\cite{}. However, whether CSF flows more rapidly in PVSs and
-what the forces and mechanisms required to drive such flow are, remain
-as key points of debate~\cite{bohr2022glymphatic,
-  van2024caa}. Motivated by experimental observations in animal
-models~\cite{iliff2012paravascular, iliff2013cerebral, mestre2018flow,
-  bedussi2018paravascular}, there is now a remarkable body of
-literature on modelling perivascular fluid flow and
-transport~\cite{bilston2003arterial, asgari2016glymphatic,
-  rey2018pulsatile, daversin2020mechanisms, sharp2019dispersion,
-  thomas2019fluid, kedarasetti2020functional, kedarasetti2020arterial,
-  troyetsky2021dispersion, martinac2021phase, gjerde2023directional,
-  nozaleda2024arterial}. Here, we ask how these proposed mechanisms
-would translate from idealized geometries to human vascular networks
-and moreover, evaluate their integrated effect in the context of
-intracranial solute transport.
-
-Our periarterial network extends from the internal carotid arteries
-and basilar artery through up to 18 bifurcations to reach upstream
-network ends located within the SAS or up to 6 mm inside the
-parenchyma (\Cref{fig:results1}A--B, \Cref{fig:pvs_a}A) and includes
-major surface arteries of radius 0.5--1.4 mm. Imposing the pressure
-field induced by CSF production at the periarterial network ends
-induces slow steady CSF flow of variable direction in these PVSs, with
-an average velocity of 0.08 \textmu m/s (antegrade) and a maximum
-velocity of 0.55 \textmu m/s (\Cref{fig:pvs_a}B). On the other hand,
-traveling waves of arterial wall motion (\Cref{fig:pvs_a}C), such as
-the pulse wave or other vasomotion~\cite{vanveluw2020vasomotion,
-  munting2023spontaneous, bojarskaite2023sleep,
-  broggini2024long,hauglund2025norepinephrine}, also induce net
-directional flow in the PVS~\cite{kedarasetti2020arterial,
-  kedarasetti2020functional, coenen2021lubrication,
-  gjerde2023directional, nozaleda2024arterial} -- of magnitude and
-direction depending on the amplitude, frequency, and length of the
-waves and the characteristics of the perivascular
-network~\cite{gjerde2023directional}. Applying a semi-analytic model
-of the net flow induced by peristalsis in perivascular
-networks~\cite{gjerde2023directional} (see Methods), we estimate that
-the cardiac pulse wave alone, travelling at a frequency of 1 Hz with a
-wavelength of 2.0 m and a 1\% wall displacement~\cite{jung2021novel},
-will induce mainly antegrade PVS flow with an average net velocity of
-0.92 \textmu m/s while reaching up to 7.31 \textmu m/s near larger,
-ventral vessels such as the MCA (\Cref{fig:pvs_a}D,E). The same theory
-predicts that strong ultraslow vasomotions, if travelling antegrade at
-0.1Hz with a wavelength of 0.02 m and a 10\%
-displacement~\cite{broggini2024long}, will induce both retrograde and
-antegrade net PVS flow with an average velocity of 13.05 \textmu m/s
-and maximum velocity 54.44 \textmu m/s (\Cref{fig:pvs_a}D,F).
-
-We study the effect of such rapid PVS flow on intracranial molecular
-enrichment by comparing the more conservative baseline model and a
-high PVS flow model, where the former still includes the net flow
-contributions from CSF production and the cardiac pulse wave, while
-the latter additionally includes the ultraslow vasomotion
-contribution. In both models, tracers were first observed at the basal
-artery after 48 minutes (\Cref{fig:pvs_b}A), defined by their first-time arrival (FTA), the time at which the concentration first exceeds $0.1$ mmol/l.
-At all upstream locations along the middle cerebral arteries
-(MCAs) and anterior cerebral artery (ACA), we found substantially
-reduced FTAs with higher PVS flow (\Cref{fig:pvs_b}B), up to 2.5 hours
-earlier in the left M2 segment of the MCA. PVS concentrations peaked
-before the surrounding tissue (0:48h vs 1:12h at the MCA and 1:00h vs
-1:48h at MCA2) with high PVS flow, while these peaks occurred nearly
-simultaneously in the baseline model. The earlier appearance and the
-delay in time-to-peak between the PVS and surrounding tissue clearly
-indicates the directionality of tracer enrichment -- it first arrives
-in the PVS and subsequently spreads into the tissue and SAS
-(\Cref{fig:pvs_b}C, D), and especially so with higher PVS flow. These
-observations also hold on an aggregated level: computing the total
-amount of tracer in the PVS as a function of the distance to the
-arterial network roots, we find accelerated tracer transport with
-higher PVS flow, especially at earlier time points (2--9 hours)
-(\Cref{fig:pvs_b}E). Moreover, differences in tracer transport along the
-PVS translate to altered enrichment patterns on the whole-organ
-scale. For instance, after 4-6 hours, the faster-moving tracer in the
-PVS is clearly visible in the space adjacent to the PVS of the ACA, MCA, and other arteries (\Cref{fig:pvs_b}G). In contrast,
-there are no clear signs of early enrichment surrounding the PVS in
-the baseline model (\Cref{fig:pvs_b}F).
-
-A natural question is whether early PVS enrichment could be the result
-of increased dispersion rather than net flow in the
-PVS\cite{asgari2016glymphatic,sharp2019dispersion,bojarskaite2023sleep,asgari2016glymphatic,troyetsky2021dispersion}. Interestingly,
-even increasing the dispersion factor by $100 \times$ in the PVS,
-induced only minor changes in the global spreading rate compared to the
-baseline (\Cref{fig:pvs_b}F, H), thus indicating a negative answer to
-this question.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\textwidth]{figures/figure3a.png}
-    \caption{\textbf{A static pressure gradient and peristaltic pumping are potential drivers of perivascular flow.}
-    A) Surface arterial network coloured by arterial radius with vessel locations labelled;
-    B) Net PVS flow induced by CSF production (cf.~\Cref{fig:csf}
-    -- histograms show the relative frequencies of local periarterial velocities. Positive values indicate antegrade PVS flow, while negative values indicate retrograde PVS flow;
-    C) Illustration of the concept of peristaltic pumping;
-    D) histograms of estimated net PVS flow induced by pulse wave or vasomotion peristaltic pumping;
-    E) Estimated net PVS flow induced by pulse wave peristaltic pumping.
-    F) Estimated net PVS flow induced by vasomotion/slow wave peristaltic pumping.}
-    \label{fig:pvs_a}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.99\textwidth]{figures/figure3b.png}
-    \caption{\textbf{Perivascular flow shapes and accelerates molecular enrichment.}
-    A) Comparison of mean tracer concentration in the PVS and over the outer PVS surface at key locations of the arterial tree, for the baseline (lower) and high PVS flow (upper) models, with labels: ``peak'': time-to-peak (h), ``$\Delta t$'': time difference between peaks (h), and ``FTA'': first-time tracer arrival (h).
-    B) first-time arrival for the main trunks of the periarterial network over the distance from the closest network root node (BA, ICA-R or ICA-L for the baseline, the high PVS flow and the high PVS dispersion models);}
-    \label{fig:pvs_b}
-\end{figure}
-\begin{figure}
-\ContinuedFloat
-\caption{(\textbf{cont.})
-    C, D) 2D slices showing zoom-in on the region surrounding the MCA-R and MCA2-R for the baseline (upper) and high PVS flow (lower) models.
-Innermost circle indicates the artery (blue), the surrounding annulus is the PVS (color represents concentration value), the surroundings are parenchyma and/or CSF spaces, separated by the pia (blue: towards the parenchyma, pink: towards the CSF));
-    E) Total amount of tracer in the periarterial spaces as a function of the distance to the nearest network root over time for the baseline (left) and the high PVS flow (right) models;
-    F--H) sagittal (upper) and coronal (lower) views of the simulated tracer concentrations overlayed on the T1-weighted MR image after 4 and 6h for the baseline (K), high PVS flow (L) and high PVS dispersion (M) models (concentration opacity linearly increasing with concentration value, pial surface outlined in pink, arachnoid membrane in cyan, and arteries in dark red).
-An interactive visualization of the baseline and the high PVS flow model can be found \href{https://baseline-pvs-flow.surge.sh/}{here}, and \href{https://high-pvs-flow.surge.sh/}{here}.}
-\end{figure}
-
-\subsection*{Structural versus functional compartmentalization of perivascular spaces}
-
-Human and rodent observations indicate that tracers concentrate in perivascular spaces surrounding the pial and subarachnoid vasculature \cite{zhang1990interrelationships,zhang1992directional, bedussi2017paravascular, mestre2018flow, bedussi2018paravascular, eide2024functional}.
-However, it remains unclear whether such enrichment patterns necessitate a structural barrier, such as a membrane with limited permeability, or if the patterns could result from enhanced flow or mixing in these areas alone.
-We therefore next investigate how the permeability of the interface between the PVS and the surrounding CSF affects tracer enrichment around the major arterial trunks (\Cref{fig:compartmentalization}A--B).
-To this end, we compare models with high and low  permeabilities, representing a highly permeable PVS-CSF interface (functional compartmentalization) and a less permeable interface (structural compartmentalization), respectively.
-For the low permeability model, we set the permeability to $3.8 \cdot 10^{-7}\,$m/s, consistent with previous estimates for the endfoot sheath surrounding penetrating arterioles \cite{koch2023estimates}, which we consider to be a lower bound for the surface PVS-CSF permeability.
-For the high permeability model, we increase this permeability by a factor of 100.
-For both scenarios, we consider the high PVS flow regime as examined in the previous section.
-
-At the basal artery (BA), tracer appears within one hour in the surrounding CSF in both models (\Cref{fig:compartmentalization}C, E).
-With a more permeable PVS-CSF interface, tracer quickly crosses from the CSF into the PVS, resulting in similar peak concentrations of 25\,mmol/l in the PVS and at its outer surface, while values up to 50\,mmol/l are attained in the vicinity.
-In contrast, the low permeability model exhibits substantially lower perivascular tracer enrichment, with a maximum concentration of 10\,mmol/l after approximately two hours.
-%The high permeability model allows for more tracer crossing into the PVS, leading to higher concentrations further downstream.
-Notably, crossover into the parenchyma is also observed, most clearly at 4 hours.
-For the middle cerebral arteries (MCA-R and MCA-L), tracer appears
-first in the PVS in both models (\Cref{fig:compartmentalization}D, E),
-though with higher concentrations in the high permeability model (up
-to 24 mmol/l). We also observe that the higher permeability allows the
-tracer to leak out of the PVS and spread within the Sylvian fissure to
-a greater extent. With a small delay, more tracer appears via
-  the CSF pathway (\Cref{fig:compartmentalization}D).
-At the anterior cerebral artery (ACA-A2), the first tracer arrival in the PVS occurs after 3 hours, with a peak at 5 hours.
-The concentration peak in the PVS is tailed by a higher peak in the CSF (\Cref{fig:compartmentalization}E), indicating tracer arrival through a second pathway; i.e., directly through the CSF-filled space outside of the PVS.
-Similar patterns are observed for ACA-A3.
-
-Summarizing and quantifying these observations
-(\Cref{fig:compartmentalization}F), we find that neither the time of
-first arrival nor the time-to-peak are substantially affected by the
-PVS-CSF permeability at any of the periarterial segments considered
-(BA, MCAs, ACAs). However, the concentration difference between the
-PVS and its surrounding CSF quickly decreases with higher
-permeability, as does the time lag between the concentration peaks in
-these domains. Thus, fast transport along the PVS is not contingent
-upon a structural compartmentalization of the PVS, whereas a sharp
-concentration gradient between PVS and CSF is unlikely without a
-restricting barrier.
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\textwidth]{figures/figure4.png}
-    \caption{\textbf{Does early perivascular enrichment rely on a structural compartmentalization of the PVS?}
-    A) The PVS-CSF membrane permeability regulates the exchange of solutes between the PVS and surrounding CSF spaces; 
-    B) The arterial network with highlighted regions of interest; 
-    C) 2D slices showing zoom-in to the region surrounding the basal artery after 1, 2, 3, and 4 hours for a low (above) and high (below) PVS-CSF permeability model.
-Innermost circle indicates the artery (blue), the surrounding annulus is the PVS (color represents concentration value), the surroundings are parenchyma and/or CSF spaces, separated by the pia (blue: towards the parenchyma, pink: towards the CSF));
-    D) As in (C) for regions surrounding (from upper to lower) the ACA-A3, ACA-A2, MCA-R and MCA-L segments;}
-    \label{fig:compartmentalization}
-\end{figure}
-\begin{figure}
-\ContinuedFloat
-\caption{(\textbf{cont.})
-E) Comparison of mean tracer concentration in the PVS (dashed) and over the
-outer PVS surface (dotted) for the BA, MCA-R, MCA-L, ACA-2, and ACA3 segments over time for the high (purple) and low permeability (yellow) models;
-F) Effect of varying the PVS-CSF permeability on first-time-of-arrival (FTA), time difference ($\Delta t$) between concentration peaks in the PVS and surrounding tissue, time of PVS peak concentration, PVS peak concentration and difference in peak concentration between PVS and surrounding tissue.}
-\end{figure}
-
-\subsection*{Enlarged PVSs delay periarterial and intracranial molecular enrichment}
-
-Enlarged PVSs are associated with impaired periarterial tracer
-transport and more generally with cognitive decline as well as
-certain subtypes of
-dementia~\cite{bown2022physiology,eide2024functional}. A key
-question is thus whether enlarged PVSs alone leads to delayed perivascular
-and intracranial tracer enrichment. To address this question from a
-fluid and transport dynamics perspective, we ask our in-silico model
-to predict the integrated effect of perivascular dilation on PVS flow
-velocities and tracer enrichment patterns (\Cref{fig:enlargement}A).
-
-Again we compute CSF production-, cardiac-, and vasomotion-induced net PVS flow velocities, but now in a perivascular geometry with an increased outer radius ($R^{\rm dilated}_2 = 3 R_1$) compared to the previous ($R^{\rm control}_2 = 2 R_1$), which corresponds to a $2.67 \times$ increase in PVS cross-section area.
-%% \begin{align*}
-%%   A_1 = \pi (R_2^2 - R_1^2) = \pi (2^2 R_1^2 - R_1^2) = 3 \pi R_1^2 \\
-%%   A_2 = \pi (\hat{R}_2^2 - R_1^2) = \pi (3^2 R_1^2 - R_1^2) = 8 \pi R_1^2 \\
-%%   A_2 / A_1 = 8/3 = 2.76
-%% \end{align*}
-The enlargement yields notable and non-uniform changes in the flow
-dynamics. On the one hand, the pressure-induced flow velocities
-increase from a mean of 0.08 \textmu m/s to 0.33 \textmu m/s due to
-the reduced effective resistance of the PVS. However, the peristaltic
-pumping becomes considerably less effective: the cardiac and
-vasomotion-induced net CSF flow velocities drop from a mean of 0.92
-\textmu m/s to 0.17 \textmu m/s and from 13.05 \textmu m/s to 2.33
-\textmu m/s, respectively (\Cref{fig:enlargement}B--D). In total, the
-net CSF velocity reduces, which in turn alters the tracer enrichment
-within the dilated PVS. Notably, we observe later tracer arrival in
-all MCA segments, with up to 96 min later arrival in the
-MCA2s for the high PVS flow scenario (\Cref{fig:enlargement}E, F). For
-the baseline model, which omits the vasomotion-induced net PVS flow,
-the effect is less pronounced in the MCA but clearly persists and is
-evident in the MCA2s (\Cref{fig:enlargement}F). Finally, we note that
-the larger volume of the dilated PVS results in a greater total
-accumulation of tracer in the PVS from 3--24 hours and reduced
-enrichment of the parenchyma at 6 and 12 h, while the mean PVS
-concentration is initially lower, but later exceeds the baseline model (\Cref{fig:enlargement}G,H).
-\begin{figure}
-    \centering
-    \includegraphics[width=1 \textwidth]{figures/figure5.png}
-\caption{\textbf{PVS enlargement reduces net PVS flow and delays tracer transport}
-A) Modelling dilated PVS by extending the outer PVS boundaries;
-B) mean and max PVS flow velocities with normal/control and dilated PVS;
-C--D) 3D rendering of vasomotion-induced PVS flow in the normal (C) and dilated (D) periarterial networks; 
-E) 2D slices showing zoom-in to the region surrounding the MCA-R after 1, 2, 3, and 4\,h with normal (lower) and dilated (upper) PVS.
-Innermost circle indicates the artery (blue), the surrounding annulus is the PVS (color represents concentration value), the surroundings are parenchyma and/or CSF spaces, separated by the pia (blue: towards the parenchyma, pink: towards the CSF));
-F) First-time arrival for different branches of the arterial tree for the baseline model (with pressure-driven and cardiac-induced net PVS flow) and the high PVS flow model (also including vasomotion-induced net PVS flow) for normal and dilated PVS;
-G) total tracer content in the CSF, parenchyma and arterial and venous PVS at 1,3,6,12 and 24\,h for the four models from F;
-H) mean concentration in the CSF, parenchyma and arterial and venous PVS over time for the four models from F.
-}
-    \label{fig:enlargement}
-\end{figure}
-
-\section*{Discussion}
-
-%% Summary of findings
-We have presented a high-fidelity in-silico model of molecular
-enrichment and clearance in human intracranial spaces, enabling
-tailored predictions of the influence of CSF space and (peri)vascular
-morphology, physiological factors such as cardiac and respiratory
-pulsatility, as well as of pathological conditions such as enlarged
-PVSs. A key observation is that the balance between CSF production and
-intracranial pulsatility significantly shapes the large-scale features
-of molecular enrichment, with e.g.~ventricular tracer reflux after
-intrathecal injection in the absence of CSF production, and preference
-towards anterior brain regions with reduced pulsatility. Indeed, the
-simulated distribution patterns cover a substantial range of the
-individual variations observed clinically \cite{ringstad2018brain}. In
-terms of perivascular pathways, we find that even moderate CSF flow on
-the order of 0.1--1.0 \textmu m/s in the surface PVSs results in
-earlier tracer enrichment around major cerebral arteries. This effect
-is more pronounced -- with up to a three-hour difference in time of
-arrival -- with mean PVS flow rates of approximately 10 \textmu m/s. Our
-models of CSF flow in perivascular networks, based on first principles
-and asymptotic analysis, predict net flow of such magnitudes and both
-antegrade and retrograde flow within the perivascular network,
-occurring with or without pressure differences at the network
-ends. The PVS may thus function as highways facilitating rapid
-transport, even in the absence of structural barriers, though a sharp
-concentration difference between the PVS and surrounding tissues is
-unlikely without such barriers. Finally, dilated PVS will cause a
-substantial reduction in cardiac- and vasomotion-driven flow
-velocities, which in turn impedes perivascular transport.
-
-%% Comparison with others
-
-% Tracer enrichment predictions
-While molecular enrichment of the brain and surrounding CSF spaces
-have been extensively studied in animal models, especially in murine
-models in connection with studies of the glymphatic system, there are
-fewer reports of contrast enrichment in human subjects over 0--24h. We
-therefore mainly compare our in-silico predictions with the series of
-papers by Eide and colleagues~\cite{ringstad2017glymphatic,
-  ringstad2018brain, eide2021sleep, eide2024functional} and Watts et
-al~\cite{watts2019measuring}. Ringstad et al.~\cite{ringstad2018brain}
-report of tracer enrichment in a centripetal pattern, primarily in
-regions near large cerebral surface arteries, peaking in the CSF
-spaces between 6 and 9 hours, with parenchymal tracer content still
-increasing after 24 hours, but with large individual
-differences. Watts et al~\cite{watts2019measuring} report of a similar
-enrichment pattern and a concentration peak in the SAS of $\approx$0.5
-mmol/l (relative to a total amount of 1.0 mmol) occuring after 10 to 15 hours. Our in-silico enrichment
-patterns agree with these observations, though with an earlier rather than later peak in the CSF spaces (2 hours after first tracer appearance at the craniocervical junction).
-In particular, we note that the substantial individual variation observed clinically is comparable to the variation in in-silico enrichment patterns associated with reduced CSF production or reduced intracranial pulsatility.
-We find that about 20\% of the tracer reaching the cranium enters the parenchyma, which aligns with the reported peak concentration values of $0.5\,$mmol/l in the SAS and $0.1\,$mmol/l in the parenchyma reported by Watts et al~\cite{watts2019measuring}.
-In terms of perivascular tracer
-enrichment, Eide and Ringstad~\cite{eide2024functional} report of
-first-time tracer appearance in the PVS of the MCA after 37.8 $\pm$
-47.0 (M1), 53.1 $\pm$ 50.5 (M2) and 82.6 $\pm$ 61.5 (M3) min after injection, thus with just a 15.3 min delay between M2 and M1.
-Comparing with our FTAs in the MCA segments, we note that the baseline model predicts a more-than 3 hour difference in FTA between the M1 and M2.
-In contrast, the high PVS flow scenario predicts a 24 min delay between these segments.
-Moreover, the high PVS flow scenario predicts FTAs of
-84--108 min (M1) and 108--132 min (M2) after tracer appearance at the craniocervical junction.
-Assuming a tracer travel time from the point of injection to the craniocervical junction of $13\,$min (as reported by Eide and Ringstad \cite{eide2024functional}), our predictions fall at the upper end of the clinical ranges.
-This can partly be attributed to the limited caudal extend of our PVS network, with the lowest point located approximately $4\,$cm above the craniocervical junction due to imaging constrains.
-We interpret these results as indicative of a net CSF flow in human surface PVS beyond what would be induced by CSF production and cardiac peristalsis alone.
-% M2 - M1 (mean) ~ 15.3 min
-
-% PVS flow predictions. 
-The existence, magnitude, and directionality of flow in surface and
-parenchymal PVSs have been the subject of active debate for
-decades~\cite{rennels1985evidence, bilston2003arterial,
-  hadaczek2006perivascular, carare2008solutes, iliff2012paravascular,
-  iliff2013cerebral, bakker2016lymphatic, mestre2018flow,
-  bedussi2018paravascular, rey2018pulsatile, thomas2019fluid,
-  daversin2020mechanisms, kedarasetti2020functional,
-  martinac2021phase, vanveluw2020vasomotion, bohr2022glymphatic,
-  boster2023artificial, gjerde2023directional,
-  nozaleda2024arterial}. In mice, Mestre et al~\cite{mestre2018flow}
-observe CSF flow in surface perivascular regions surrounding branches
-of the MCA with average flow speeds of 18.7 \textmu m/s and net flow
-in the antegrade direction, but also regions with retrograde
-flow. Similarly, Bedussi et al.~\cite{bedussi2018paravascular} report
-of oscillatory flow, with an average (net) CSF velocity of 17 $\pm$ 2
-\textmu m/s in the antegrade direction. Using physics-informed neural
-networks, Boster et al~\cite{boster2023artificial} estimate PVS
-velocities of $12.75 \pm 6.25$ \textmu m/s. Less is known about the
-magnitude and directionality of CSF flow in surface PVSs in
-humans. Here, assuming that human surface PVS follow the major surface
-vessels and are CSF-filled but otherwise open regions of width
-proportional to the vessel radius, our estimates indicate that the CSF
-pressure differences due to CSF production, cardiac peristaltic
-pumping and low-frequency vasomotion all yield spatially-varying net
-CSF flow with both antegrade and retrograde PVS segments. This
-observation is thus in agreement with the experimental
-observations~\cite{mestre2018flow, bedussi2018paravascular}, and
-inherently supports the original notion that antegrade and retrograde
-solute transport along PVSs may coexist~\cite{rennels1985evidence}. In
-terms of magnitude, the contribution from CSF production and cardiac
-wall motion are small (0.08 \textmu m/s and 0.92 \textmu m/s on
-average), but intriguingly, our estimates of the contribution from
-vasomotion (13.05 \textmu m/s) are comparable to the experimental
-reports. We do note that this estimate is based on the presence of
-rhythmic waves of vasomotion at a frequency of 0.1Hz, a wavelength of
-20 mm, and a wall displacement of 10\% -- in the antegrade direction,
-all values currently associated with significant
-uncertainty~\cite{gokina1996electrical, vanveluw2020vasomotion,
-  broggini2024long}.
-
-% Compartmentalization of the PVS
-To what extent are surface PVSs in communication with the surrounding
-SAS and to what extent are these structurally separated compartments?
-Originally, Weller and coauthors~\cite{zhang1990interrelationships,
-  zhang1992directional, weller2005microscopic} identified thin sheaths
-of pial cells surrounding human surface arteries and penetrating
-arterioles (but not veins or venules). On the other hand, Bakker and
-coauthors report that the subarachnoid space, the cisterns, ventricles
-and penetrating periarteriolar spaces form a continuous CSF
-space~\cite{bedussi2017paravascular}. Thorne and colleagues study the
-architectures of rat perivascular spaces in detail, emphasizing the
-presence of openings (stomata) on the interface between the
-vasculature and the CSF within the SAS~\cite{pizzo2018intrathecal,
-  hannocks2018molecular}. Further, Mestre et
-al~\cite{mestre2022periarteriolar} study the properties of pial
-perivascular spaces in mice, and report of pial cells forming sheaths
-for larger surface arteries and partially cover smaller surface
-arteries, with higher coverage in ventral SAS regions. Our results
-demonstrate that the PVSs may act as rapid transport pathways even
-with only a partial barrier between the PVS and surrounding CSF, thus
-contributing to quantifying the properties of the human PVS-CSF
-interface~\cite{eide2024functional}.
-
-% Variations with sleep, exercise
-Lifestyle factors such as sleep~\cite{xie2013sleep, miao2024brain,
-  bojarskaite2023sleep, eide2021sleep, vinje2023human,
-  hauglund2025norepinephrine, larsen2024sleep},
-exercise~\cite{holstein2018voluntary, yildiz2022immediate}, and
-alcohol intake~\cite{lundgaard2018beneficial} affect molecular
-enrichment and clearance of the brain. In particular, pial arteries
-display higher-amplitude low-frequency vasomotion during NREM and
-intermediate sleep states, while REM sleep is associated with reduced
-PVS width~\cite{bojarskaite2023sleep}. Peristaltic pumping is more
-effective under both of these configurations, resulting in higher
-estimated net PVS flow velocities~\cite{gjerde2023directional}. Even
-higher PVS velocities would amplify the PVS flow effects simulated
-here, with earlier arrival times along the PVSs, clear enrichment in
-adjacent spaces, and accelerated molecular enrichment
-overall. Moreover, sleep is linked with low-frequency oscillations in
-human CSF flow~\cite{fultz2019coupled}. As we have shown here, the
-level of dispersion in the CSF spaces strongly shapes molecular
-enrichment patterns. While we, at baseline, account for the dispersion
-induced by cardiac and respiratory pulsatility, the high dispersion
-scenarios considered can be viewed as representative of additional
-pulsatility induced by e.g.~neural waves~\cite{fultz2019coupled,
-  williams2023neural} or other respiration patterns depending on
-activity level. More broadly, we consider this as a key strength of
-the current in-silico framework: allowing for integrating and
-exploring the mechanistic macroscale implications of changes in
-pulsatility at smaller spatial or temporal scales.
-
-% Not discussed variations with disease states
-
-% Limitations
-In terms of limitations, our primary constraint modelling-wise is the
-geometric representation of the vascular (and therefore also
-perivascular) networks. Concretely, we account for no vessels below
-the original MRI resolution limits, and thus only a part of the
-cerebral surface vasculature and effectively not the cortical
-vessels. In addition, and again due to lack of resolution, there is
-uncertainty associated with the location of the vessel segments
-relative to the CSF spaces and parenchyma, with some vessels
-surrounded by both domains and some vessels completely entering the
-parenchyma before returning to the SAS. Higher resolution MRI data
-(T1w and e.g. time-of-flight, QSM or other modalities) in more
-subjects would immediately allow for more detailed studies of this
-aspect. As a result of the lack of an accurate representation of the
-cortical and subcortical white matter vasculature, we have considered
-a conservative transport model within the parenchyma, modelling
-extracellular diffusion alone. To compensate, we have focused on
-reporting detailed predictions in the CSF spaces and surface PVS after
-not more than 24 hours. In particular, we note that, we observe tracer
-accumulation at the perivascular network ends after 12 hours,
-similar to that previously observed for
-microspheres~\cite{bedussi2018paravascular, mestre2018flow}, which
-here can be interpreted as an artifact of the incomplete network
-creating up a barrier for tracer movement. Previously, we have found
-that enrichment within the parenchyma is well-represented by $\approx
-3\times$ enhanced diffusion augmented by local clearance e.g.~across
-the blood-brain barrier or diffusion-convection with a complex pattern
-of spatially-varying convective velocities of 1-10 \textmu
-m/min~\cite{vinje2023human}. We will incorporate such transport
-mechanisms within the parenchyma in future work. Finally, we also note
-that we simulate the intracranial distribution of a tracer
-concentration appearing at the craniocervical junction rather than as
-an intrathecal injection. We therefore overestimate the amount of
-tracer entering into the intracranial spaces; however, due to the
-linearity of the mathematical model, it is valid to directly interpret
-the in-silico concentrations relative to the total amount of tracer.
-
-% Outlook
-In conclusion, our findings transfer insights from experimental
-studies and theoretical analysis to the in-silico human setting,
-reconcile seemingly conflicting observations in particular relating to
-directionality of perivascular flow, and integrate different physical
-mechanisms across across spatial and temporal scales. The complete
-simulation pipeline is openly available, including interactive
-visualization of simulation results for all model
-variations~\cite{ZENODO}. Looking ahead, this platform establishes a
-foundation for in-silico studies of molecular movement within the
-human brain environment, such as tailored predictions of intrathecal
-chemotherapy delivery or personalized diagnostics of brain clearance
-capacity.
-
-\section*{Methods}
-
-\subsection*{Intracranial compartments: ventricular system, SAS, and brain parenchyma}
-
-From T1-weighted MR images of a 26-year-old, healthy male volunteer~\cite{hodneland2019new},
-we first automatically segment the brain parenchyma and CSF spaces,
-including the ventricular system and SAS, using
-Synthseg~\cite{billot2023robust,billot2023synthseg}
-(\Cref{fig:results1}A--B). We next manually adjust the segmentation to
-accurately represent the connections and barriers between CSF spaces (aqueduct,
-median aperture, tentorium cerebelli), smoothen, and finally extract
-surface representations of the outer (arachnoid) boundary, the pial
-membrane and other interfaces. Conforming to these surface and
-interface representations, we generate a tetrahedral mesh $\Omega$
-representing the complete intracranial volume as the union of the
-parenchyma $\Omega_{\rm PAR}$ and CSF spaces $\Omega_{\rm CSF}$, using
-fTetWild~\cite{hu2020fast} (\Cref{fig:csf}A--B). The resulting
-computational mesh consists of $233\,592$ vertices and $1\,290\,131$
-mesh cells, which vary between $0.22$ mm and $8.9$ mm in mesh cell
-diameter. The parenchyma has a total volume of $1318$ ml, whereas the
-ventricles and the outer SAS contribute $32$ ml and $329$ ml,
-respectively, to a total CSF volume of $361$ml, in agreement with
-recent estimates~\cite{hladky2024regulation}.
-
-\subsection*{Periarterial and perivenous spaces}
-
-To represent surface networks of periarterial and perivenous spaces,
-we use Kiminaro~\cite{william_silversmith_2021_5539913} to separately
-skeletonize time-of-flight angiography (ToF) and quantitative
-susceptibility mapping (QSM) images from the same human
-subject~\cite{hodneland2019new}. This technique yields networks of
-one-dimensional curves, each curve $\Lambda^i$ indicating the
-centerline of a blood vessel segment and labeled with its lumen radius
-$R_1^i$ (\Cref{fig:pvs_a}A). We also assign an
-outer radius $R_2^i > R_1^i$ to each segment $i$, form the annular
-cylinder ensheathing $\Lambda^i$ and define the union of these as the
-PVS. The periarterial network graph consists of $12\,708$ edges
-connected at $12\,562$ inner nodes and with $147$ end nodes, and the
-associated domain is denoted by $\Lambda_a$. We identify three of the
-end nodes -- corresponding to the two internal carotid arteries (ICAs)
-and the basilar artery (BA) -- as root nodes, and designate the other
-ends as leaf nodes. The resulting perivenous network graph consists of
-$24\,829$ edges connected at $23\,881$ inner nodes and with $949$ end nodes, and the associated domain is denoted by $\Lambda_v$.
-Finally, we label the outer surface of the PVSs associated with $\Lambda_a$ and $\Lambda_v$, by $\Gamma_a$ and $\Gamma_v$ respectively.
-
-\subsection*{Molecular transport equations}
-
-We model diffusion, convection, and exchange of a molecular solute in
-the PVS networks $\Lambda_a$ and $\Lambda_v$, in the CSF spaces
-$\Omega_{\rm CSF}$ and in the brain parenchyma $\Omega_{\rm PAR}$ via
-a mixed-dimensional transport model~\cite{masri2024modelling,laurino2019derivation} over a
-timescale of minutes to hours (\Cref{fig:results1}A--B,
-\Cref{fig:csf}A--C). Specifically, for $t > 0$, we solve for a
-concentration $c = c(x, t)$ in the 3D intracranial compartments ($x
-\in \Omega_{\rm CSF}, x \in \Omega_{\rm PAR}$) and for a cross-section
-averaged concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial
-and perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$) such that
-the following equations hold:
-\begin{subequations}
-\begin{alignat}{2}
-  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma & = 0 && \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}},
-  \label{eq:multi_transport_3d}
-  \\ 
-  \partial_t (A  \hat c) - \partial_s(\hat D A \partial_s ( \hat c)) +\partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c})  &= 0 && \quad \quad \mathrm{in} \quad  \Lambda_a, \Lambda_v .
-  \label{eq:multi_transport_1d}
- \end{alignat}%
-\label{eq:multi_transport}%
-\end{subequations}%
-In~\eqref{eq:multi_transport}, $0 < \phi \leqslant 1$ is the fluid
-volume fraction, $D$ and $\hat{D}$ are effective diffusion
-coefficients, and $\bm u$ and $\hat{u}$ are fluid (CSF/ISF)
-velocities, all defined in the respective compartments; $P = 2 \pi
-R_2$ is the perimeter and $A = \pi (R_2^2 - R_1^2)$ the area of the
-PVS cross-sections, $\delta_{\Gamma}$ is a Dirac term concentrated on
-the interfaces $\Gamma_a, \Gamma_v$, the overlined $\overline{c}$
-denotes the average of $c$ over cross-sections of $\Gamma_a,
-\Gamma_v$, and $\xi$ is a permeability allowing for transfer/exchange
-over the interfaces $\Gamma_a, \Gamma_v$ between the periarterial and
-perivenous networks and their surroundings.  Moreover, we model the
-membrane between the parenchyma and the CSF spaces as a semi-permeable
-membrane with permeability coefficient $\beta_{\rm pia}$. For a
-complete description of this mathematical model and simulation
-scenarios, see \Cref{sec:app:more_maths}. The initial concentrations
-in the three-dimensional domain $\Omega$ and networks $\Lambda_a,
-\Lambda_v$ are all set to zero.
-
-\subsection*{Intracranial influx after intrathecal injection}
-To represent molecular influx into the intracranial compartments
-following intrathecal injection, without modelling the spinal
-compartment explicitly, we prescribe a time-dependent molecular influx
-over the interface towards the spinal CSF compartment with the condition
-\begin{equation*}
-  (D \nabla c - \bm u c ) \cdot \bm{n} = g_{\mathrm{influx}},  \quad \mathrm{on}  \quad \Gamma_{\mathrm{SSAS}}.
-\end{equation*}
-Here we set
-\begin{equation*}
-  g_\mathrm{influx}(t) = \frac{m_{\rm tot}}{T_{\rm max}^2 |\Gamma_{\rm SSAS}|} \max(0, T_{\rm max} - |t - T_{\rm max}|), 
-\end{equation*}
-which expresses a hat-shaped influx function with a peak at $T_{\rm max}$ and a total tracer injection of $m_{\rm tot}$ (\Cref{tab:parameters}).
-
-\subsection*{Molecular clearance from the intracranial space and perivascular network ends}
-
-We model molecular clearance, into the meningeal lymphatics or other
-pathways, across the upper, outer (arachnoid) boundary
-(\Cref{fig:results1}A, \Cref{fig:csf}C), proportional to the
-concentration in the SAS with a rate constant $\beta_{\rm exit}$. We
-set a no-flux condition at the end nodes of the perivascular networks,
-not allowing for direct molecular clearance there.
-
-\subsection*{CSF flow in the SAS and ventricular system}
-
-We model CSF as an incompressible Newtonian fluid at low Reynolds and
-Womersley numbers via the Stokes equations with viscosity $\mu$, and
-account for the convective contribution of the flow induced by CSF
-production in the choroid plexus. Specifically, we compute the
-velocity $\bm u_{\rm CSF}$ and associated pressure $p_{\rm CSF}$ in
-the SAS and ventricular system induced by a constant production at a
-rate of $u_{\rm in}$ across the lateral ventricle walls and with
-efflux across the upper, outer (arachnoid) boundary with efflux
-resistance $R_0$; and then set $\bm u = \bm u_{\rm CSF}$ in
-$\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport}
-(\Cref{tab:parameters}, \Cref{fig:csf}A--E, see also
-\Cref{sec:app:more_maths}). We do not model bulk flow within the brain
-parenchyma, except within the perivascular network segments extending
-below the pial surface, and thus set $\bm u = 0$ in $\Omega_{\rm PAR}$
-in~\eqref{eq:multi_transport}.
-
-\subsection*{Dispersion in the CSF spaces}
-
-The pulsatile flow of CSF in the SAS and ventricular system,
-associated with the cardiac and respiratory cycles, substantially
-enhances molecular diffusion through
-dispersion~\cite{taylor1953dispersion, watson1983diffusion,
-  asgari2016glymphatic, sharp2019dispersion, ray2021quantitative,
-  troyetsky2021dispersion}. To account for these dispersive effects,
-while bridging from the pulsatile flow time scale of seconds to a
-molecular transport time scale of hours, we compute cardiac and
-respiratory dispersion factors $R_c$ and $R_r$, both
-spatially-varying, in $\Omega_{\rm CSF}$ (\Cref{fig:csf}A) by
-combining computational estimates of the respective peak CSF pressure
-gradients with theoretical estimates for shear-augmented (Taylor)
-dispersion \cite{taylor1953dispersion, watson1983diffusion,
-  sharp2019dispersion} (see \Cref{sec:app:dispersion} for a complete
-description). Combining these effects, we then set $R = R_c + R_r$ and
-define the effective diffusion coefficient $D = (1 + R) D^{\rm Gad}$
-in $\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport} in the baseline model.
-
-\subsection*{Perivascular fluid flow induced by CSF pressure differences}
-
-To estimate the contribution also to perivascular fluid flow from CSF
-production, we impose the fluid pressure $p_{\rm CSF}$ induced by CSF
-production, computed in the SAS and ventricular system, at the end
-nodes of the periarterial network located within the SAS. For the
-network end nodes located within the parenchyma, we compute a harmonic
-extension of the CSF pressure field (by solving a Laplace equation in
-$\Omega_{\rm PAR}$ with $p_{\rm CSF}$ as the boundary value), and set
-the corresponding value at these end nodes. We then numerically solve
-a system of hydraulic network equations to compute $\hat{u}^p_a$
-defined over $\Lambda_a$; i.e., the pressure-difference induced CSF
-flow velocity in the periarterial network (see
-also~\Cref{sec:app:hydraulic_network}). We repeat this procedure to
-compute a corresponding velocity $\hat{u}^p_v$ in the perivenous
-network $\Lambda_v$.
-
-\subsection*{Perivascular fluid flow induced by arterial wall motion}
-
-Using the theoretical framework introduced by Gjerde et
-al.~\cite{gjerde2023directional}, we also compute analytic estimates
-for the time-average perivascular flow rate $\langle Q' \rangle$
-induced by peristalsis in the arterial network $\Lambda_a$; i.e.~the
-net flow induced by traveling waves of arterial wall motion of
-frequency $f$, amplitude $\varepsilon$ and wave lengths $\lambda$
-(\Cref{fig:pvs_a}C). The corresponding contribution to the periarterial
-flow velocity is defined for each segment $\Lambda_i$ as $\langle Q'_i
-\rangle/A_i$ where $A_i = \pi (R_2^i - R_1^i)$ is the cross-section
-area of the PVS segment and $\langle Q_i' \rangle$ is the estimated
-mean flow rate (see~\ref{sec:app:peristalsis} for more details). Two
-wall motion patterns are considered, yielding $\hat{u}_a^c$ and
-$\hat{u}_a^v$ respectively: \emph{cardiac} pulsations with $f =
-1.0$Hz, $\lambda = 2$ m and $\varepsilon = 1\%$ (\Cref{fig:pvs_a}D), and very low-frequency \emph{vasomotion} with $f = 0.1$Hz, $\lambda =
-0.02$m, and $\varepsilon = 10\%$ (\Cref{fig:pvs_a}E).
-No flow induced by peristalsis is considered for the perivenous network.
-In the baseline model, we combine the contributions from CSF production and cardiac peristalsis to set $\hat{u}$ in~\eqref{eq:multi_transport} as
-$\hat{u}_a = \hat{u}_a^p + \hat{u}_a^c$ in $\Lambda_a$, and $\hat{u}_v
-= \hat{u}_v^p$ in $\Lambda_v$.
-
-\subsection*{Model and material parameters}
-
-Model and material parameters are summarized in \Cref{tab:parameters},
-and~\Cref{tab:scenarios} gives an overview of the different in-silico
-scenarios. We assume that the porosity equals the extracellular space
-volume fraction $\phi = \phi_{\rm PAR}$ within the parenchyma and
-$\phi = 1$ elsewhere. The effective diffusion coefficient is set equal
-to that of Gadubutrol in free water in the PVS networks: $\hat{D} =
-D^{\rm Gad}$; weighted by a dispersion factor $R$ in the CSF spaces:
-$D|_{\Omega_{\rm CSF}} = ( 1 + R) D^{\rm Gad}$; and modulated by the
-porosity and tortuosity ($\lambda = 1.78$) in the parenchyma:
-$D|_{\Omega_{\rm PAR}} = D^{\rm Gad}_{\rm PAR}$. We set $\xi$
-depending on whether the PVS segment is fully embedded within the CSF
-spaces ($\xi = \xi_{\rm CSF}$) or parenchyma ($\xi = \xi_{\rm EF}$),
-or neither (\Cref{sec:app:more_maths}). Additionally, we increase the
-permeability $\xi$ ($100\times$) at the the leaf nodes of
-$\Lambda_a$, linearly decreasing to the baseline permeability at the next network node ($\approx0.4\,$mm), to account for the transition from explicitly resolved
-PVSs to their unresolved continuation in the CSF spaces and
-parenchyma.
-\begin{table}
-  \begin{center}
-    \begin{tabular}{ll|ccc}
-      \toprule
-      Parameter& Symbol & Value & Unit& Reference\\
-      \midrule
-      Extracellular space volume fraction & $\phi_{\rm PAR}$ & 0.2 & -- & \cite{nicholson1981ion} \\
-      Free diffusion coefficient, Gadubutrol & $D^{\rm Gad}$ & $3.8 \times 10^{-4}$& $\text{mm}^2/\text{s}$ & \cite{valnes2020apparent}\\
-      Parenchymal diffusion coefficient, Gadubutrol & $D^{\rm Gad}_{\rm PAR}$ & $1.2 \times 10^{-4}$ & $\text{mm}^2/\text{s}$  & \cite{hornkjol2022csf} \\
-      Permeability coefficient, pia & $\beta_{\rm pia}$ & $2.6 \cdot 10^{-8}$ & m/s & \cite{riseth2025twocompartment} \\
-      Permeability coefficient, endfoot sheath & $\xi_{\rm EF}$ & $3.8\cdot 10^{-7}$  & m/s & \cite{koch2023estimates} \\
-      Molecular efflux rate & $\beta_{\rm exit}$ & $10^{-4}$ & mm$^2$/s & \cite{hornkjol2022csf} \\
-      Total molecular volume & $m_{\rm tot}$ & $0.5$ & mmol & \cite{eide2024functional} \\
-      Molecular influx, time-to-peak & $T_{\rm max}$ & $1$ & h & $\ast$ \\
-      CSF production rate & $u_{\mathrm{in}}$ & $0.63$  & l / day & \cite{nilsson1992circadian} \\
-      CSF viscosity & $\mu$ & $0.7$ & mPa$ \cdot $s & \cite{bloomfield1998effects} \\ 
-      CSF outflow resistance & $R_0$ & $10^{-5}$  & Pa/(mm$\cdot$s) & \cite{hornkjol2022csf} \\ 
-      Dispersion factor, CSF spaces & $R$ & $\dagger$ & -- & $\ast$ \\
-      \bottomrule
-    \end{tabular}
-  \end{center}
-  \caption{Model parameters. $\dagger$/$\ast$ denotes computed/estimated within this work.}
-  \label{tab:parameters}
-\end{table}
-\begin{table}
-  \begin{center}
-  \begin{tabular}{ll|llll|lll}
-    \toprule
-    & Description & $R_1, R_2$ & $\xi_{\rm PVS-CSF}$ & $D|_{\Omega_{\rm CSF}}$ & $D|_{\Omega_{\rm PAR}}$ & $\mathbf{u}|_{\Omega_{\rm CSF}}$ & $\hat{u}|_{\Lambda_a}$ & $\hat{u}|_{\Lambda_v}$  \\
-    \midrule
-     & baseline & $R_2 = 2 R_1$ & $\xi$ & $(1 + R) D^{\rm Gad}$ & $D^{\rm Gad}_{\rm PAR}$ & $\mathbf{u}_{\rm CSF}$ & $\hat{u}_a^p + \hat{u}_a^c$  & $\hat{u}_v^p$ \\ 
-    & no CSF production & $-$ & $-$ & $-$ & $-$ & $\mathbf{0}$ & $\hat{u}_a^c$ & $0$ \\ 
-    & reduced pulsatility & $-$ & $-$ & $ (1 + 0.1 R) D^{\rm Gad}$ & $-$ & $-$ & $-$  & $-$ \\ 
-    & increased CSF dispersion & $-$ & $-$ & $ (1 + 10 R) D^{\rm Gad}$ & $-$ & $-$ & $-$  & $-$ \\ 
-    & higher PVS flow & $-$ & $-$ & $-$ & $-$ & $-$ & $\hat{u}_a^p + \hat{u}_a^c + \hat{u}_a^v$  & $-$ \\ 
-    & higher PVS exchange & $-$ & 100$\xi$  &  $-$ & $-$ & $-$ & $-$  & $-$ \\ 
-    & dilated PVS & $R_2 = 3 R_1$ & $-$ & $-$ & $-$ & $-$ & $\tilde{u}_a^p + \tilde{u}_a^c$  & $-$ \\ 
-    \bottomrule
-    \end{tabular}
-    \end{center}
-  \caption{Overview of models and model variations. ($-$ denotes equal
-    to the corresponding baseline value). Each PVS cross-section is
-    modelled as a concentric annulus with inner radius $R_1$ and outer
-    radius $R_2$ and such that $R_2 = \beta R_1$. Let $\xi$ denote the
-    astrocytic endfeet permeability estimated by Koch et
-    al.~\cite{koch2023estimates}. The diffusion coefficients $D^{\rm
-      Gad}$ represent the (free) diffusion coefficient of Gadubutrol
-    in CSF (water), and $D^{\rm Gad}_{\rm PAR}$ its effective
-    diffusion coefficient in human cortical tissue, all at body
-    temperature~\cite{sykova2008diffusion, valnes2020apparent}. $R$ is
-    the combined cardiac and respiratory dispersion enhancement field
-    (see Methods). $\mathbf{u}_{\rm CSF}$ is the CSF velocity field
-    induced by CSF production in $\Omega_{\rm CSF}$, and $\hat{u}_a^p$
-    and $\hat{u}_v^p$ the axial velocity induced in the periarterial
-    and perivenous networks, respectively, by the corresponding
-    pressure differences. $\hat{u}_a^c$ and $\hat{u}_a^v$ is the
-    periarterial velocity induced by arterial pulse wave wall motion
-    and slow vasomotion, respectively. Last, $\tilde{u}_a^c$ and
-    $\tilde{u}_a^p$ are as $\hat{u}_a^c$ and $\hat{u}_a^p$,
-    respectively, but with dilated PVS.}
-\label{tab:scenarios}
-\end{table}
-
-\subsection*{Numerical approximation, implementation, and verification}
-
-We give a concise summary of the numerical approach here (see
-Supplementary methods for a comprehensive description). To
-solve~\eqref{eq:multi_transport} numerically, we use the interior
-penalty discontinuous Galerkin (DG) method with weighted averages and
-upwinding for the convection term in the 3D
-domains~\cite{ern2009discontinuous} -- to ensure numerical stability
-and minimize the artificial presence of negative concentration
-values. For the 1D domain, we use continuous finite element spaces
-defined over each of the networks and stabilized with numerical
-diffusion when $\hat{u}$ is non--zero. Note that the 3D--1D coupling
-in~\eqref{eq:multi_transport} leads to a 3D solution $c$ with low
-regularity properties. Thus, one can only expect lower-order error
-convergence near $\Lambda_a$ and $\Lambda_v$, see the references
-\cite{masri2024discontinuous} and \cite{laurino2019derivation} for the
-numerical analysis of DG and continuous Galerkin schemes,
-respectively. However, away from these 1D networks, almost optimal
-approximation is
-expected~\cite{masri2023discontinuous,koppl2016local}. To numerically
-solve the Stokes equations to compute the CSF flow velocity and
-pressure in the ventricular system and SAS, we use finite element
-spaces that preserve the incompressibility condition on the discrete
-level~\cite{hong2016robust}. This requirement is important for
-numerical stability when subsequently
-solving~\eqref{eq:multi_transport}\cite{cesmelioglu2022compatible}.
-The spaces we used for the velocity are continuous along the normal
-direction; continuity of the tangential component is enforced via
-interior penalty approaches~\cite{hong2016robust}. Optimal
-approximation properties are expected for this scheme. These numerical
-methods were implemented using the FEniCS finite element
-framework~\cite{alnaes2015fenics}, and the 3D-1D coupling is handled
-by the extension
-FEniCS\textsubscript{ii}~\cite{kuchta2020assembly}. The correctness
-and accuracy of the numerical solutions were verified by a series of
-numerical verification experiments (\Cref{sec:app:verification}).
-
-\subsection*{Model validation}
-
-As the primary means of model validation, we compare the in-silico
-predictions of tracer enrichment and clearance against glymphatic MRI
-studies (Discussion). In addition, we compare auxiliary model
-quantities (CSF flow rates and pressure differences, dispersion factor
-estimates, and the shapes and sizes of PVSs) with the current
-literature (\Cref{sec:app:model_validation}).
-
-\section*{Acknowledgments}
-
-All authors express their gratitude to Prof.~Jürgen Reichenbach for
-generously sharing the multi-modal human MR data, and to Prof. Erlend
-Hodneland for facilitating this data exchange. M.E.R.~acknowledges
-support from Stiftelsen Kristian Gerhard Jebsen via the K.~G.~Jebsen
-Centre for Brain Fluid Research. M.C., R.M., and M.E.R.~acknowledge
-support from the Research Council of Norway (RCN) via grant \#324239
-(EMIx). Simulations were performed on the Experimental Infrastructure
-for Exploration of Exascale Computing (eX3), which is financially
-supported by the RCN under contract 270053. M.K.~acknowledges support
-from the RCN grant \#303362. The figures 1A, 1D, 2C, 3C, 5A and 6A
-were created in Biorender.
-
-\section*{Author contributions statement}
-
-M.C., M.K, R.M., and M.E.R. conceived and designed the project.
-M.C., M.K, R.M., and M.E.R. contributed to software development. 
-M.C., M.K, and R.M. conducted the experiments.
-M.C. analyzed the results and prepared the figures.
-M.C. M.K. and R.M and M.E.R wrote the manuscript. 
-All authors edited and reviewed the manuscript.
-
-\section*{Competing interests}
-
-The authors declare no competing interests.
+\input{supplementary_main.tex}
 
@@
 
-\newpage
-\FloatBarrier
-\appendix
-\input{supplementary_main.tex}
-
 \end{document}

### 2025-02-27 paper/supplementary_main.tex
Split supplementary into separate file with independent references...

@@
 \subset \R^3$ with $\Omega = \Omega_{\rm PAR} \cup \Omega_{\rm CSF}$
-(\Cref{fig:results1}A, \Cref{fig:csf}A--C). The coordinates in these
+(Fig.~1A, Fig.~2A--C). The coordinates in these
 3D domains is denoted by $x$. The interface between the parenchyma and
@@
 
-In this section, we describe the mathematical
-model~\eqref{eq:multi_transport}, the associated definitions,
-interface and boundary conditions in further detail. We refer to the
-main text for explicit parameter values while giving the general,
-abstract form here, and to the reference~\cite{masri2024modelling, laurino2019derivation} for
-the derivation and analysis of this 3D-1D model. Recall that we model
-a concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
+In this section, we describe the mathematical model (eq.~(1)), the
+associated definitions, interface and boundary conditions in further
+detail. We refer to the main text for explicit parameter values while
+giving the general, abstract form here, and to the
+reference~\cite{masri2024modelling, laurino2019derivation} for the
+derivation and analysis of this 3D-1D model. Recall that we model a
+concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
 defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
@@
 specifically, we compute spatially-varying dispersion enhancement
-fields $R_c$ and $R_r$ (\Cref{fig:csf}G, I), associated with the
+fields $R_c$ and $R_r$ (Fig.~2G, I), associated with the
 cardiac and respiratory cycles respectively, via the algorithm
@@
   to the Stokes flow induced by the peak volumetric reduction of the
-  CSF space in the respective cycle (\Cref{fig:csf}F, H). More
+  CSF space in the respective cycle (Fig.~2F, H). More
   precisely, we numerically solve the incompressible Stokes
@@
   with CSF density $\rho = 10^3$ kg/m$^3$, a mean CSF space width $h_{\rm SAS} =
-  1.5$mm, and CSF viscosity $\mu$ given in~\Cref{tab:parameters}.  For
+  1.5$mm, and CSF viscosity $\mu$ given in~Table.~1.  For
   the cardiac cycle, we consider an angular frequency $\omega_c = 2
@@
     Relative PVS size $\beta$ & $\beta = \beta_i = R_2^i / R_1^i$ & 2 & $\approx$\cite{eide2024functional} \\
-    Wave frequency $f$ & Traveling wave frequency of vascular motion & $0.1-1.0$ Hz & \discuss{\cite{gjerde2023directional}} \\
-    Wave length $\lambda$ & Traveling wave length of vascular motion & $0.02-2$ mm & \discuss{\cite{broggini2024long, gjerde2023directional}} \\
-    Wave amplitude & Relative amplitude of inner wall motion & $1-10\%$ & \discuss{\cite{gjerde2023directional}} \\
+    Wave frequency $f$ & Traveling wave frequency of vascular motion & $0.1-1.0$ Hz & \cite{gjerde2023directional} \\
+    Wave length $\lambda$ & Traveling wave length of vascular motion & $0.02-2$ mm & \cite{broggini2024long, gjerde2023directional} \\
+    Wave amplitude & Relative amplitude of inner wall motion & $1-10\%$ & \cite{gjerde2023directional} \\
     \midrule
@@
     Vasomotion amplitude $\varepsilon_v$ & Wall displacement amplitude of vasomotion wave & 10\%  & \cite{broggini2024long, gjerde2023directional}\\
-    %Vascular radii $\{r_{o_i}\}$ & Vascular network radii data (avg $\pm$ std, $\min$, $\max$) & $X$, $X$, $X$  mm & \cite{hodneland2019new} \\
-    %Vascular branch lengths $\{L_{i}\}$ & Path length between vasc. junctions (avg $\pm$ std, $\min$, $\max$) & $X \pm Y$, $X$, $X$  mm & \discuss{--} \\
     \bottomrule
@@
 edge-disjoint subnetworks $\Lambda^1, \Lambda^2, \Lambda^3$, one for
-each of the supply nodes (\Cref{fig:results1}A--B,
+each of the supply nodes (Fig.~1A--B,
 \Cref{fig:app:peristalsis}). Each node is assigned to the subnetwork

### 2025-02-27 paper/main.tex
Updates in Figure 1 caption. @Marius, please check numbers and unit again cf. discussion on mol/l vs mmol/l

@@
      \caption{
-     \textbf{Our in-silico model predicts tracer transport in PVS, SAS and parenchyma following intrathecal injection.}\\
-     A) Illustration of the spatial domains of the model: CSF-filled space (ventricles and SAS) $\Omega_{\rm CSF}$, parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and the veins $\Lambda_{\rm vein}$ and their interfaces and boundaries (see Methods);
-     B) 3D rendering of the actual geometry used in this study parenchyma, CSF (clipped, blue), arterial network (red), venous network (blue)); 
-     C) illustration of the modelled clinical tracer injection procedure; 
-     D) simulated tracer concentration after 1, 6, 12, and 24 hours (opacity increases linearly from 0 to $2\,$mol/l); 
-     E) sagittal, coronal and axial view of computed tracer concentrations over T1-image after 1, 6, 12 and 24\,h (low concentrations transparent, pial surface in pink, arachnoid membrane in cyan, and arteries in dark red);
+     \textbf{In-silico modelling of molecular enrichment and clearance of the PVS, SAS and parenchyma.}
+     A) Illustration of the model geometry including the CSF-filled spaces (ventricles and SAS) $\Omega_{\rm CSF}$, the parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and veins $\Lambda_{\rm vein}$, as well as their interfaces and boundaries;
+     B) 3D rendering of the computational geometry: the parenchyma, CSF (clipped, blue), arterial network (red), and venous network (blue); 
+     C) illustration of model representation of the intrathecal tracer injection; 
+     D) in-silico predictions of tracer concentration after 1, 6, 12, and 24 hours (opacity increasing linearly from 0 to 2 mmol/l); 
+     E) sagittal, coronal and axial view of in-silico tracer concentrations overlayed on T1-weighted MR image after 1, 6, 12 and 24\,h (low concentrations transparent, pial surface in pink, arachnoid membrane in cyan, and arteries in dark red);
      F) average tracer concentration in each compartment over the first 24\,h. 
-     G) total amount of tracer in each cranial compartment over the first 24\,h; }
+     G) total amount of tracer in each compartment over the first 24\,h.}
      \label{fig:results1}

### 2025-02-27 paper/main.tex
Fix punctuation.

--- 2025-02-27 paper/main.tex
@@
      \caption{
-     \textbf{Our in-silico model predicts tracer transport in PVS, SAS and parenchyma following intrathecal injection.}\\
-     A) Illustration of the spatial domains of the model: CSF-filled space (ventricles and SAS) $\Omega_{\rm CSF}$, parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and the veins $\Lambda_{\rm vein}$ and their interfaces and boundaries (see Methods);
-     B) 3D rendering of the actual geometry used in this study parenchyma, CSF (clipped, blue), arterial network (red), venous network (blue)); 
-     C) illustration of the modelled clinical tracer injection procedure; 
-     D) simulated tracer concentration after 1, 6, 12, and 24 hours (opacity increases linearly from 0 to $2\,$mol/l); 
-     E) sagittal, coronal and axial view of computed tracer concentrations over T1-image after 1, 6, 12 and 24\,h (low concentrations transparent, pial surface in pink, arachnoid membrane in cyan, and arteries in dark red);
-     F) average tracer concentration in each compartment over the first 24\,h. 
-     G) total amount of tracer in each cranial compartment over the first 24\,h; }
+     \textbf{In-silico modelling of molecular enrichment and clearance of the PVS, SAS and parenchyma.}
+     A) Illustration of the model geometry including the CSF-filled spaces (ventricles and SAS) $\Omega_{\rm CSF}$, the parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and veins $\Lambda_{\rm vein}$, as well as their interfaces and boundaries;
+     B) 3D rendering of the computational geometry: the parenchyma, CSF (clipped, blue), arterial network (red), and venous network (blue); 
+     C) illustration of model representation of the intrathecal tracer injection; 
+     D) in-silico predictions of tracer concentration after 1, 6, 12, and 24 hours (opacity increasing linearly from 0 to 2 mmol/l); 
+     E) sagittal, coronal and axial view of in-silico tracer concentrations overlayed on T1-weighted MR image after 1, 6, 12 and 24\,h (low concentrations transparent, pial surface in pink, arachnoid membrane in cyan, and arteries in dark red);
+     F) average tracer concentration in each compartment over the first 24\,h; 
+     G) total amount of tracer in each compartment over the first 24\,h.}
      \label{fig:results1}

### 2025-03-20 paper/main.tex
Updates from Overleaf

@@
 \usepackage{placeins}
+\usepackage{lineno}
 \usetikzlibrary{quotes}
@@
 
+\linenumbers
+
 \flushbottom
+
 \maketitle
@@
 here can be interpreted as an artifact of the incomplete network
-creating up a barrier for tracer movement. Previously, we have found
+creating a barrier for tracer movement. Previously, we have found
 that enrichment within the parenchyma is well-represented by $\approx
@@
 directionality of perivascular flow, and integrate different physical
-mechanisms across across spatial and temporal scales. The complete
+mechanisms across spatial and temporal scales. The complete
 simulation pipeline is openly available, including interactive

### 2025-06-03 paper/main.tex
Updates in response to reviewer's comments on enlarged PVS.

@@
 Our findings highlight the significant impact of cerebrospinal fluid (CSF) production and intracranial pulsatility on molecular enrichment following intrathecal tracer injection.
-We demonstrate that low-frequency vasomotion induces moderate CSF flow in surface PVS networks which substantially enhances tracer enrichment, and that impaired enrichment is a direct natural consequence of enlarged PVSs.
+We demonstrate that low-frequency vasomotion induces moderate CSF flow in surface PVS networks which substantially enhances tracer enrichment, and that impaired enrichment is a direct natural consequence of enlarged surface PVSs.
 This openly available technology platform thus provides an opportunity for integrating separate observations on diffusion in neuropil, vascular dynamics, intracranial pulsatility, CSF production, and efflux, and for exploring drug delivery and clearance in the human brain.
@@
 
-\subsection*{Enlarged PVSs delay periarterial and intracranial molecular enrichment}
-
-Enlarged PVSs are associated with impaired periarterial tracer
-transport and more generally with cognitive decline as well as
-certain subtypes of
-dementia~\cite{bown2022physiology,eide2024functional}. A key
-question is thus whether enlarged PVSs alone leads to delayed perivascular
-and intracranial tracer enrichment. To address this question from a
-fluid and transport dynamics perspective, we ask our in-silico model
-to predict the integrated effect of perivascular dilation on PVS flow
-velocities and tracer enrichment patterns (\Cref{fig:enlargement}A).
+\subsection*{Enlarged PVSs in the SAS delay periarterial and intracranial molecular enrichment}
+
+Idiopathic normal pressure hydrocephalus (iNPH), a dementia subtype,
+is associated with enlarged PVS in the SAS and impaired periarterial
+and parenchymal tracer transport~\cite{eide2024functional}. More
+generally, enlarged PVSs are associated with a range of pathophysiogical
+conditions and cognitive decline~\cite{bown2022physiology}. A key
+question is thus whether enlarged PVSs alone leads to delayed
+perivascular and intracranial tracer enrichment. To address this
+question from a fluid and transport dynamics perspective, we ask our
+in-silico model to predict the integrated effect of perivascular
+dilation on PVS flow velocities and tracer enrichment patterns
+(\Cref{fig:enlargement}A).
 
@@
     \includegraphics[width=1 \textwidth]{figures/figure5.png}
-\caption{\textbf{PVS enlargement reduces net PVS flow and delays tracer transport}
-A) Modelling dilated PVS by extending the outer PVS boundaries;
+\caption{\textbf{PVS enlargement in the SAS reduces net PVS flow and delays tracer transport}
+A) Modelling dilated PVS in the SAS by extending the outer PVS boundaries;
 B) mean and max PVS flow velocities with normal/control and dilated PVS;
@@
 pulsatility, as well as of pathological conditions such as enlarged
-PVSs. A key observation is that the balance between CSF production and
-intracranial pulsatility significantly shapes the large-scale features
-of molecular enrichment, with e.g.~ventricular tracer reflux after
-intrathecal injection in the absence of CSF production, and preference
-towards anterior brain regions with reduced pulsatility. Indeed, the
-simulated distribution patterns cover a substantial range of the
-individual variations observed clinically \cite{ringstad2018brain}. In
-terms of perivascular pathways, we find that even moderate CSF flow on
-the order of 0.1--1.0 \textmu m/s in the surface PVSs results in
-earlier tracer enrichment around major cerebral arteries. This effect
-is more pronounced -- with up to a three-hour difference in time of
-arrival -- with mean PVS flow rates of approximately 10 \textmu m/s. Our
-models of CSF flow in perivascular networks, based on first principles
-and asymptotic analysis, predict net flow of such magnitudes and both
-antegrade and retrograde flow within the perivascular network,
-occurring with or without pressure differences at the network
-ends. The PVS may thus function as highways facilitating rapid
-transport, even in the absence of structural barriers, though a sharp
-concentration difference between the PVS and surrounding tissues is
-unlikely without such barriers. Finally, dilated PVS will cause a
-substantial reduction in cardiac- and vasomotion-driven flow
-velocities, which in turn impedes perivascular transport.
+PVSs in the SAS. A key observation is that the balance between CSF
+production and intracranial pulsatility significantly shapes the
+large-scale features of molecular enrichment, with e.g.~ventricular
+tracer reflux after intrathecal injection in the absence of CSF
+production, and preference towards anterior brain regions with reduced
+pulsatility. Indeed, the simulated distribution patterns cover a
+substantial range of the individual variations observed clinically
+\cite{ringstad2018brain}. In terms of perivascular pathways, we find
+that even moderate CSF flow on the order of 0.1--1.0 \textmu m/s in
+the surface PVSs results in earlier tracer enrichment around major
+cerebral arteries. This effect is more pronounced -- with up to a
+three-hour difference in time of arrival -- with mean PVS flow rates
+of approximately 10 \textmu m/s. Our models of CSF flow in
+perivascular networks, based on first principles and asymptotic
+analysis, predict net flow of such magnitudes and both antegrade and
+retrograde flow within the perivascular network, occurring with or
+without pressure differences at the network ends. The PVS may thus
+function as highways facilitating rapid transport, even in the absence
+of structural barriers, though a sharp concentration difference
+between the PVS and surrounding tissues is unlikely without such
+barriers. Finally, dilated PVS in the SAS will cause a substantial reduction in
+cardiac- and vasomotion-driven flow velocities, which in turn impedes
+perivascular transport.
 
@@
     & higher PVS exchange & $-$ & 100$\xi$  &  $-$ & $-$ & $-$ & $-$  & $-$ \\ 
-    & dilated PVS & $R_2 = 3 R_1$ & $-$ & $-$ & $-$ & $-$ & $\tilde{u}_a^p + \tilde{u}_a^c$  & $-$ \\ 
+    & dilated PVS in the SAS & $R_2 = 3 R_1$ & $-$ & $-$ & $-$ & $-$ & $\tilde{u}_a^p + \tilde{u}_a^c$  & $-$ \\ 
     \bottomrule

### 2025-06-04 paper/main.tex
Address point about PVS permeability.

@@
 For the high permeability model, we increase this permeability by a factor of 100.
-For both scenarios, we consider the high PVS flow regime as examined in the previous section.
+We emphasize that, for both scenarios, we consider the high PVS flow regime as examined in the previous section.
 
@@
 contributing to quantifying the properties of the human PVS-CSF
-interface~\cite{eide2024functional}.
+interface~\cite{eide2024functional}. In this regard, we emphasize that
+we report on the isolated effect of variations in permeability on
+molecular transport, separated from any effects on fluid flow in the
+PVS. Clearly, one may expect that the presence and/or properties of
+such a barrier would affect also PVS flow. However, considering the
+uncertainty associated with PVS flow characteristics in the human SAS
+in general and the lack of computational models describing net PVS
+flow in the presence of fluid exchange, this remains challenging to
+quantify.
 
@@
 frequency $f$, amplitude $\varepsilon$ and wave lengths $\lambda$
-(\Cref{fig:pvs_a}C). The corresponding contribution to the periarterial
-flow velocity is defined for each segment $\Lambda_i$ as $\langle Q'_i
-\rangle/A_i$ where $A_i = \pi (R_2^i - R_1^i)$ is the cross-section
-area of the PVS segment and $\langle Q_i' \rangle$ is the estimated
-mean flow rate (see Supplementary information S1.4 for more details). Two
-wall motion patterns are considered, yielding $\hat{u}_a^c$ and
-$\hat{u}_a^v$ respectively: \emph{cardiac} pulsations with $f =
-1.0$Hz, $\lambda = 2$ m and $\varepsilon = 1\%$ (\Cref{fig:pvs_a}D), and very low-frequency \emph{vasomotion} with $f = 0.1$Hz, $\lambda =
-0.02$m, and $\varepsilon = 10\%$ (\Cref{fig:pvs_a}E).
-No flow induced by peristalsis is considered for the perivenous network.
-In the baseline model, we combine the contributions from CSF production and cardiac peristalsis to set $\hat{u}$ in~\eqref{eq:multi_transport} as
-$\hat{u}_a = \hat{u}_a^p + \hat{u}_a^c$ in $\Lambda_a$, and $\hat{u}_v
-= \hat{u}_v^p$ in $\Lambda_v$.
+(\Cref{fig:pvs_a}C). The corresponding contribution to the
+periarterial flow velocity is defined for each segment $\Lambda_i$ as
+$\langle Q'_i \rangle/A_i$ where $A_i = \pi (R_2^i - R_1^i)$ is the
+cross-section area of the PVS segment and $\langle Q_i' \rangle$ is
+the estimated mean flow rate (see Supplementary information S1.4 for
+more details). Two wall motion patterns are considered, yielding
+$\hat{u}_a^c$ and $\hat{u}_a^v$ respectively: \emph{cardiac}
+pulsations with $f = 1.0$Hz, $\lambda = 2$ m and $\varepsilon = 1\%$
+(\Cref{fig:pvs_a}D), and very low-frequency \emph{vasomotion} with $f
+= 0.1$Hz, $\lambda = 0.02$m, and $\varepsilon = 10\%$
+(\Cref{fig:pvs_a}E). No flow induced by peristalsis is considered for
+the perivenous network. In the baseline model, we combine the
+contributions from CSF production and cardiac peristalsis to set
+$\hat{u}$ in~\eqref{eq:multi_transport} as $\hat{u}_a = \hat{u}_a^p +
+\hat{u}_a^c$ in $\Lambda_a$, and $\hat{u}_v = \hat{u}_v^p$ in
+$\Lambda_v$. We remark that none of these models account for fluid
+exchange across the lateral outer interface of the PVSs (over
+$\Gamma_a, \Gamma_v$), while fluid entry and exit is allowed at the
+end nodes of the PVS networks.
 
@@
     $\tilde{u}_a^p$ are as $\hat{u}_a^c$ and $\hat{u}_a^p$,
-    respectively, but with dilated PVS.}
+    respectively, but with dilated PVS. }
 \label{tab:scenarios}

### 2025-06-04 paper/main.tex
Address overstatement comment.

@@
 (and challenges) for personalized medicine e.g.~for tailored
-intrathecal delivery of chemotherapy~\cite{lohela2022glymphatic} and
-for early diagnostics of impaired brain
-clearance~\cite{eide2021clinical, van2024human}. In spite of their
-importance, our understanding of these mechanisms is incomplete with
-open challenges and significant debate -- in part relating to the
-translation of knowledge between scales, species, experimental
-protocols, clinical cohorts, and individuals.
+intrathecal delivery of chemotherapies\cite{lohela2022glymphatic}
+such as methodextrate in acute lymphoblastic leukemia
+patients\cite{kadan2009comparison}, and for early diagnostics of
+impaired brain clearance\cite{eide2021clinical, van2024human}. In
+spite of their importance, our understanding of these mechanisms is
+incomplete with open challenges and significant debate -- in part
+relating to the translation of knowledge between scales, species,
+experimental protocols, clinical cohorts, and individuals.
 

### 2025-06-05 paper/main.tex
Updates

@@
     A) Surface arterial network colored by arterial radius with vessel locations labeled;
-    B) Net PVS flow induced by CSF production (cf.~\Cref{fig:csf}
-    -- histograms show the relative frequencies of local periarterial velocities. Positive values indicate antegrade PVS flow, while negative values indicate retrograde PVS flow;
+    B) Net PVS flow induced by CSF production alone (see also ~\Cref{fig:csf}D).
+    Histograms show the relative frequencies of local periarterial velocities. Positive values indicate antegrade PVS flow, while negative values indicate retrograde PVS flow;
     C) Illustration of the concept of peristaltic pumping;
     D) histograms of estimated net PVS flow induced by pulse wave or vasomotion peristaltic pumping;
-    E) Estimated net PVS flow induced by pulse wave peristaltic pumping.
-    F) Estimated net PVS flow induced by vasomotion/slow wave peristaltic pumping.}
+    E) Estimated net PVS flow induced by pulse wave peristaltic pumping alone.
+    F) Estimated net PVS flow induced by vasomotion/slow wave peristaltic pumping alone.}
     \label{fig:pvs_a}

### 2025-06-05 paper/supplementary.tex
Updates from Overleaf
Also note about vasomotion antegrade in supplementary.

@@
 \begin{abstract}
-  Supplementary methods and discussion accomanying the main text of
+  Supplementary methods and discussion accompanying the main text of
   \emph{In-silico molecular enrichment and clearance of the human
@@
 
+\linenumbers
 \flushbottom
@@
 
-\input{supplementary_main.tex}
+\section{Supplementary methods}
+
+These sections provide more detailed descriptions of the mathematical models and numerical approximations considered.
+
+\subsection{Mixed-dimensional transport and flow equations}
+\label{sec:app:more_maths}
+
+\subsubsection{Notation}
+
+In terms of geometrical domains, we consider the parenchyma
+$\Omega_{\rm PAR} \subset \R^3$ and CSF spaces $\Omega_{\rm CSF}
+\subset \R^3$ with $\Omega = \Omega_{\rm PAR} \cup \Omega_{\rm CSF}$
+(Fig.~1A, Fig.~2A--C). The coordinates in these
+3D domains is denoted by $x$. The interface between the parenchyma and
+CSF spaces is given by $\partial \Omega_{\rm PAR} \cap \partial
+\Omega_{\rm CSF}$ and we separate it as two parts: the surface of the
+lateral ventricles $\Gamma_{\rm LV}$ and the remaining pial interface
+$\Gamma_{\rm pia}$ (thus $\partial \Omega_{\rm CSF} \bigcap \partial
+\Omega_{\rm PAR} = \Gamma_{\rm LV} \cup \Gamma_{\rm pia}$). The
+remaining, outer, boundary of $\Omega_{\rm CSF}$ is again separated
+into two parts: $\Gamma_{\rm SSAS}$ represents the lower interface
+towards the spinal subarachnoid space (SSAS), while $\Gamma_{\rm AM}$
+is the outer interface towards the arachnoid and dura
+membranes. $\Gamma_{\rm AM}$ is further subdivided into its lower and
+upper parts: $\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$. The boundary
+towards the spinal cord is denoted by $\Gamma_{\rm{SC}}$ and given by
+$\Gamma_{\rm{SC}} = \partial \Omega_{\rm PAR} \setminus \partial
+\Omega_{\rm CSF}$.
+
+In addition, we consider two sets of perivascular networks: a
+periarterial network $\Lambda_a$ represented by the (connected)
+centerlines $\Lambda^i_{a}$ of the arterial tree, and a perivenous
+network $\Lambda_v$ associated with the centerlines $\Lambda^i_{v}$
+of the veins. Recall that we represent the vascular domains as the
+union of cylindrical vessels of radius $R_1^i$ surrounding the
+centerlines $\Lambda^i$. Moreover, we consider the surrounding
+perivascular spaces as the union of annular cylinders $\Omega^i$ of
+inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$ and thus of
+width $R_2^i - R_1^i$, and remark that we interpret $R^i_2$ and $R^i_1$ as temporal averages (fixed in time), as the pulsations considered in \Cref{sec:app:peristalsis} are beyond the temporal resolution of our model.
+We denote the outer lateral surface of the
+periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$,
+respectively. We omit the subscript $a$ or $v$ when referring to any
+such network, vessel segment or perivascular outer surface. We assume
+that $\Lambda$ is parametrized by the coordinate $s$, and with a minor
+abuse of notation, simply write $s \in \Lambda$ to represent the point
+${\bm \lambda}(s)$ on $\Lambda$ corresponding to $s$. Finally, we view
+each perivascular network both as a geometric domain and as a directed
+graph with the centerlines $\{ \Lambda^i \}$ as oriented edges and the
+connections as nodes $\mathcal{V}$.
+
+\subsubsection{Solute transport and exchange in intracranial domains (3D) and perivascular networks (1D)}
+\label{sec:app:transport_eqs}
+
+In this section, we describe the mathematical model (eq.~(1)), the
+associated definitions, interface and boundary conditions in further
+detail. We refer to the main text for explicit parameter values while
+giving the general, abstract form here, and to the
+reference~\cite{masri2024modelling, laurino2019derivation} for the
+derivation and analysis of this 3D-1D model. Recall that we model a
+concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
+defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
+each of these domains, $c$ satisfies
+%\mer{MER: Rami asks if it should be a $\phi$ in the advective term here.
+We can include it, or drop it, but in the latter case, perhaps we should clarify that u is an average/superficial/Darcy velocity.}
+\begin{equation}
+  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma
+  %= f
+   = 0
+  \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
+  \label{eq:3d_pde}
+\end{equation}
+In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$).
+In the parenchyma, $\phi$ represents the extracellular space volume fraction, and thus $c$ here generally represents the intrinsic (in contrast to the superficial) concentration~\cite{sykova2008diffusion}.
+Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$.
+$\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivascular networks $\Lambda_a, \Lambda_v$.
+To summarize,
+\begin{equation*}
+\phi =  \begin{cases}
+  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
+  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
+  \end{cases}, \; 
+D = \begin{cases}
+  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
+  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
+  \quad \bm u  = \begin{cases}
+  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
+  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
+\end{cases}, \; 
+\xi = \begin{cases}
+  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
+  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
+\end{cases} . 
+\end{equation*}
+In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp.
+$|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp.
+$\Omega_{\rm{CSF}}$).
+If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.
+
+Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by
+\begin{equation*}
+  \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta_2^i(s)} c 
+\end{equation*}
+where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$.
+Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the term $\delta_\Gamma$ is concentrated on the outer lateral surfaces of the PVSs, and defined in terms of its action on any sufficiently smooth function $v : \Omega \rightarrow \R$
+%and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$
+such that
+\begin{equation*}
+  \langle \xi (\overline{c} - \hat c ) \delta_\Gamma, v \rangle =
+  \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
+\end{equation*}
+under the assumption that $\xi$ is constant on each $\partial
+\Theta_2^i$.
+%% , since:
+%% \begin{multline}
+%%   \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
+%%   \int_{\Gamma} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
+%%   \sum_i \int_{\Gamma^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = 
+%%   \sum_i \int_{\Lambda^i} \int_{\partial \Theta_2^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \\
+%%   \sum_i \int_{\Lambda^i} \left (  \xi (\overline{c} - \hat c ) \int_{\partial \Theta_2^i} \delta_\Gamma (v) \right ) =
+%%   \sum_i \int_{\Lambda^i} P \xi (\overline{c} - \hat c ) \overline{v} =
+%%   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v} .
+%% \end{multline}
+%% \mer{MER: The previous was:
+%% \begin{equation}
+%%   \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
+%% \end{equation}
+%% }
+
+On the interface between the brain and the CSF spaces
+$\Gamma_{\mathrm{pia}} \cup \Gamma_{\rm LV}$, we prescribe the
+following interface conditions, which represent a semi--permeable
+interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and
+$c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
+%% \begin{subequations}
+%% \begin{alignat}{2}
+%% (-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} &= - (- D (\phi \nabla c_{\rm{CSF}}) + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n} &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\
+%% (-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,
+%% \end{alignat} 
+%% \end{subequations}
+\begin{subequations}
+  \begin{alignat}{2}
+  (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n}
+  &= - (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm CSF}} \cdot \bm{n}
+  &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
+  (- D \nabla (\phi  c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
+\end{alignat} 
+  \label{eq:3d_interface}%
+\end{subequations}%
+where $\bm{n} = \bm{n}_{\rm PAR}$ is the normal vector field defined
+over the interface, outward-pointing when viewed from $\Omega_{\rm
+  PAR}$, and $\beta_{\mathrm{pia}} \geqslant 0$ is a permeability
+constant.
+
+We supplement~\eqref{eq:3d_pde} and~\eqref{eq:3d_interface} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
+\begin{subequations}
+\begin{alignat}{2}
+(D \nabla (\phi c) - \bm u c ) \cdot \bm{n} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \Gamma_{\rm SSAS},  \\ 
+% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\Gamma_{\rm SAS}} &= 0 &&  \quad  \mathrm{on} \quad \Gamma_{\rm SAS}^{\mathrm{out}}  \\ 
+(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = \beta_{\rm exit} c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-U},  \\ 
+(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = 0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-L} \cup \Gamma_{\rm SC}  .
+\end{alignat} 
+\label{eq:3d_bcs}
+\end{subequations}
+where again $\bm n$ denotes the outward-pointing boundary normal.
+
+The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
+perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
+in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
+space averaged over each perivascular cross-section and is governed by 
+\begin{equation}
+  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s \hat c  ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
+  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
+\label{eq:1d_pde}
+\end{equation}
+where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
+of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
+PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
+associated PVS $\Omega^i$, the area of the cross-section
+$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
+and $\hat{u}$ is a convective velocity representing an (average) CSF
+flow velocity in the axial direction of the perivascular spaces.
+
+To complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
+internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
+boundary conditions at the end nodes. To this end, define the set of
+internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
+that are connected to two or more edges and the set of end nodes as
+$\mathcal{N} \subset \mathcal{V}$, write $\hat{c}^i =
+\hat{c}|_{\Lambda^i}$, and for $y\in\mathcal{B}$ let $\mathcal{E}(y)$ denote the set of
+(two or three) edges in $\{\Lambda^i\}$ sharing the node $y$. To
+impose continuity of the concentrations, we set that the
+concentrations when viewed from each edge must match at nodes:
+\begin{equation}
+  \hat{c}^i (y) = \hat{c}^j(y)
+  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
+  \qquad \foralls y \in \mathcal{B} .
+  \label{eq:conc_continuity}
+\end{equation}
+Moreover, to ensure mass conservation, we set that the flux
+going in and out at nodes should add to zero, or more precisely, that
+for each $y \in \mathcal{B}$:
+\begin{equation}
+  \sum_{\Lambda^i \in \mathcal{E}(y)}
+  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n^i(y)
+  = 0.
+  \label{eq:flux_conservation}
+\end{equation}
+Here the (normal ``vector") function $n^i$ takes the values in $\left\{-1, 1\right\}$ and 
+defines an orientation of vertices of edge $\Lambda^i$. Specifically, for an edge $\Lambda^i$ oriented from $y_{\rm in}$ to $y_{\rm out}$ we set 
+\begin{equation}
+  n^i(y_{\rm in}) = 1,\quad
+  n^i(y_{\rm{out}}) = -1. %\quad
+  %\mk{n^i(y) = 0 \quad \foralls y \in 
+  %\mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} }.  
+\label{eq:def:sign}
+\end{equation}
+%\mk{Why do we need the normal on non-IO vertices?}
+
+Finally, we impose a homogeneous Neumann condition at all network end
+nodes to augment~\eqref{eq:1d_pde}:
+\begin{equation}
+    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{N}.
+    \label{eq:1d_bcs}
+\end{equation}
+This no-flux condition states that the network end nodes represent
+barriers for molecular efflux into or out of the perivascular network,
+and thus that all solute exchange between the PVSs and their
+surroundings takes place via the lateral outer PVS surface and is
+regulated by the exchange parameter $\xi$ cf.~\eqref{eq:3d_pde}
+and~\eqref{eq:1d_pde}. Larger particles have been observed to
+accumulate within the PVS as the surface arteries penetrate into the
+brain parenchyma~\cite{mestre2018flow, bedussi2018paravascular},
+and~\eqref{eq:1d_bcs} is appropriate to represent such
+behavior. However, to represent a continuously extending PVSs also
+along penetrating vessels, this condition should be revisited.
+
+\subsubsection{Modelling CSF flow via the incompressible Stokes equations}
+\label{sec:app:csf_flow}
+The time-independent, incompressible Stokes equations model the flow
+of an incompressible Newtonian fluid at low Reynolds numbers, and read
+as follows: over the domain $\Omega_{\rm CSF} \subset \R^3$, find the
+velocity vector field $\bm u : \Omega_{\rm CSF} \rightarrow \R^3$ and
+the pressure field $p : \Omega_{\rm CSF} \rightarrow \R$ such that
+\begin{subequations}
+  \begin{alignat}{2}
+    - \mu \Delta \bm u  + \nabla p &=  0 \quad && \text{ in } \Omega_{\rm CSF},
+    \label{eq:momentum_equation} \\ 
+    \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF}
+    \label{eq:divergence_equation} .
+  \end{alignat}
+  \label{eq:stokes}%
+\end{subequations}%
+In addition, we impose the following boundary conditions to model flow
+induced by CSF production in the choroid plexus with $\Gamma_{\rm AM-U}$ as the main efflux site:
+\begin{subequations}
+  \begin{alignat}{2}
+    ( \mu \nabla \bm u \cdot \bm{n} - p \bm n) \cdot \bm n  &= -R_0 ( \bm u \cdot \bm n ), 
+    \quad \bm u \cdot \bm t = 0
+    && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
+    \bm u &= 0 
+    && \text{ on } \Gamma_{\mathrm{AM-L}}\cup\Gamma_{\rm{pia}}\cup\Gamma_{\rm{SSAS}}, \\
+    %\bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
+    %\bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
+    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, 
+    \quad \quad \bm u \cdot \bm t = 0 \quad 
+    && \text{ on } \Gamma_{\rm{LV}} ,  
+  \end{alignat}
+  \label{eq:stokes_bcs}%
+\end{subequations}%
+where $\bm n$ and $\bm t$ denote the unit outward normal and tangent vectors to the boundary respectively,  $\mu$ is the (dynamic) CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$.
+We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:app:dispersion}.
+
+%% We consider non-zero convection in the CSF space $u_{\rm CSF} \not = 0$ in \eqref{eq:multi_transport_3d} resulting from steady state CSF production.
+Recall that $\Omega_{\mathrm{CSF}}$ represents the CSF space and that $\Gamma_{\mathrm{skull^{u/l}}}$ and $\Gamma_{\mathrm{pia}}$, $\Gamma_{\mathrm{LV}}$ represent the boundaries facing the upper and lower parts of the dura, the pia and the lateral ventricles, respectively.
+Assuming CSF production in the choroid plexus located at the lateral ventricle wall, we solve the  steady state Stokes equations for the velocity and the pressure $(\bm u_{\mathrm{CSF}}, p_{\mathrm{CSF}})$ in the CSF domain:
+
+\subsubsection{Steady flow in perivascular networks induced by pressure differences}
+\label{sec:app:hydraulic_network}
+
+To model steady flow induced by pressure differences between end nodes
+in a perivascular network $\Lambda$, with edges $\{\Lambda^i\}_i$,
+internal nodes $\mathcal{B}$ and end nodes $\mathcal{N}$, we consider
+the following system of hydraulic network
+equations~\cite{tithof2022network, daversin2022geometrically,
+  gjerde2024directional}. The unknowns are the PVS flux $\hat{q} :
+\Lambda \rightarrow \R$ and pressure $\hat{p} : \Lambda
+\rightarrow \R$, which represent the fluid flux and the average
+pressure across cross-sections of the PVSs, respectively. These are
+defined for each $\Lambda^i$ by solving:
+\begin{subequations}
+\begin{alignat}{2}
+  \hat{q} + \frac{A \kappa}{\mu} \partial_{s} \hat{p} & = 0 &&
+  \quad \text{ in } \Lambda^i , \\ 
+  - \partial_s \hat{q} & = 0 &&
+  \quad \text{ in } \Lambda^i ,  
+\end{alignat} 
+\label{eq:1d_flow}%
+\end{subequations}%
+where $A$ is the area of the PVS cross-sections, $\mu$ is the CSF viscosity as before, and $\kappa$ is derived from an assumption of Poiseuille flow in the annular cross-section of the PVS as~\cite{tithof2022network}:
+\begin{equation}
+\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
+\end{equation}
+The PVS flux $\hat{q}$ and the associated average PVS velocity in the
+axial direction $\hat{u}$ are directly related by the cross-sectional area
+\begin{equation}
+  \hat{q}^i = A^i \hat{u}^i \qquad \foralls \Lambda^i.
+\end{equation}
+
+In addition, to complete~\eqref{eq:1d_flow}, we impose continuity of
+the fluid pressure and conservation of flux at each internal node. Write
+$\hat{p}^i = \hat{p}|_{\Lambda^i}$. These conditions then read as: for
+each $y \in \mathcal{B}$ with connected edges $\mathcal{E}(y)$:
+\begin{subequations}
+\begin{align}
+  \hat{p}^i(y) &= \hat{p}^j(y) \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y), \\
+  \sum_{\Lambda^i \in \mathcal{E}(y)} \hat{q}^i n^i(y) &= 0  .
+\end{align}
+\label{eq:1d_flow:bifurcations}
+\end{subequations}
+where $n^i$ is as defined by~\eqref{eq:def:sign}. Finally, to drive this flow, we impose given fluid pressures $p_0$ at the
+network end nodes:
+\begin{equation}
+  \hat{p}(y) = p_0(y) \qquad y \in \mathcal{N} .
+  \label{eq:1d_flow:bcs}
+\end{equation}
+In the case where modelling flow in the perivascular networks
+$\Lambda_a, \Lambda_v$ induced by CSF production, we impose the fluid
+pressure $p_{\rm CSF}$ computed in the CSF spaces $\Omega_{\rm CSF}$
+(and its harmonic extensions into $\Omega_{\rm PAR}$) as the given
+fluid pressures $p_0$. Note that we here consider steady PVS flow in a
+non-moving domain; the effect of domain motion is addressed
+in~\Cref{sec:app:peristalsis}.
+
+\subsection{Numerical solution of the transport and flow equations}
+\label{sec:app:details_numerical_method}
+
+We solve the models presented in~\Cref{sec:app:more_maths} numerically
+using finite element methods. The following subsections provide more
+details, model-by-model.
+
+\subsubsection{Computational mesh entities: definitions and common notation}
+
+We consider a mesh $\mathcal{T} = \{ E \}$ of $\Omega = \Omega_{\rm
+  PAR} \cup \Omega_{\rm CSF}$, consisting of tetrahedral mesh cells
+$E$, and conforming to the domains $\Omega_{\rm PAR}$ and $\Omega_{\rm
+  CSF}$ and to the CSF-brain interface $\Gamma_{\rm pia} \cup
+\Gamma_{\rm LV}$. We denote the restriction of $\mathcal{T}$ to
+$\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ by $\mathcal{T}_{\rm PAR}$
+and $\mathcal{T}_{\rm CSF}$, respectively. The collection of all
+interior facets (i.e.~triangular faces of the tetrahedral mesh cells)
+in $\mathcal{T}_{\rm PAR}$ and $\mathcal{T}_{\rm CSF}$ are denoted by
+$\mathcal{F}_{i, \rm PAR}$ and $\mathcal{F}_{i, \rm CSF}$,
+respectively. We define the union of facets interior both to
+$\mathcal{T}_{ \rm PAR}$ and $\mathcal{T}_{ \rm CSF}$ as
+$\mathcal{F}_i = \mathcal{F}_{i, \rm PAR} \cup \mathcal{F}_{i, \rm
+  CSF}$.
+
+%Note in particular that $\mathcal{F}_i \cap \mathcal{F}_{pia}} =
+%  \emptyset$.
+
+%The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is
+%denoted by $\mathcal{E}_{h}$.
+
+\subsubsection{Finite element solution of coupled 3D-1D solute transport equations }
+
+We consider the system of coupled 3D-1D solute transport equations
+given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
+conditions~\eqref{eq:3d_interface}, the bifurcation
+conditions~\eqref{eq:conc_continuity}
+and~\eqref{eq:flux_conservation}, and the boundary
+conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
+these equations using an implicit finite difference scheme in time, 
+a discontinuous Galerkin (DG) finite element method with upwinding in space for the 3D domain to accurately capture sharp boundary layers, and a continuous Galerkin method for the 1D networks.
+We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm, and accounting for the increased diffusivity due to dispersion).
+
+We first consider the discretization of~\eqref{eq:3d_pde} and
+introduce the discrete space
+\begin{equation}
+  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
+  V_h = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^1(E), \; E \in \mathcal{T}\} ,
+\end{equation}
+where $L^2(\Omega)$ is the space of square-integrable functions
+defined over $\Omega$ and $\mathrm{P}^1(E)$ denotes the space of
+polynomials of total degree $\leqslant 1$ defined over the tetrahedra
+$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
+symmetric weighted interior penalty DG formulation, referring
+to~\cite{ern2009discontinuous} and \cite[Section
+  4.5.2.3]{di2011mathematical} for details on this
+method. Recall that $\phi$ is constant in each domain
+  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
+  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.
+Define for $c, v \in V_h^k$:
+\begin{equation}
+  a_h (c, v)
+  =
+  \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
+  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
+  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right) 
+   + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
+  .
+  \label{eq:dg_sip}
+\end{equation}
+In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
+between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
+pointing from $E^1$ to $E^2$. The facet diameter is denoted by $h_F$,
+the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
+unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
+are defined as:
+\begin{equation}
+  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
+  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
+  \quad \text{where} \quad \kappa_i = (D \phi) |_{E^i}.  
+\end{equation}
+The parameter $\gamma_{D,F}$ is the harmonic mean of the
+porosity-weighted diffusion coefficient given by
+\begin{equation}
+  \gamma_{D,F} = \frac{2\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
+\end{equation}
+and $\eta$ is a user-specified penalty parameter (we set $\eta=1000$). To discretize the convection term in~\eqref{eq:3d_pde}, we use
+upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
+references therein. For $c, v \in V_h^k$, define
+\begin{equation}
+  a_h^{\mathrm{up}}(c, v)
+  = - \sum_{E \in \mathcal{T}} \int_E c\bm{u} \cdot \nabla v
+  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
+  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
+\end{equation}
+Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h$ such that for all $v \in V_h$:
+\begin{equation} \label{eq:3d_dg_disc}
+  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
+  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
+  + \int_{\Gamma_{\rm pia} \cup \Gamma_{\rm LV}} \beta_{\rm pia}[c_h^n][v] 
+  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
+  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
+  = 
+  %\int_{\Omega} f^n v +
+  \int_{\Gamma_{\rm SSAS}} g_{\rm influx} v
+  %, \quad \forall v \in V_h^k
+  .
+\end{equation}
+
+To discretize~\eqref{eq:1d_pde} with the bifurcation
+conditions~\eqref{eq:conc_continuity}
+and~\eqref{eq:flux_conservation}, and the boundary
+conditions~\eqref{eq:1d_bcs}, we use the space of continuous piecewise
+linear polynomials defined over $\Lambda$:
+\begin{equation}
+  \hat V_h
+  = \{ v \in C^0(\Lambda), \; v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda^i) \}, 
+\end{equation}
+where $\mathrm{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$.
+The discrete formulation then reads: for $n = 1,2,\ldots, $ find $\hat c_h^n \in \hat V_h$ such that for all $\hat v \in \hat V_h^k$:
+\begin{align}\label{eq:1d_dg_disc}
+\int_{\Lambda}  \frac{1}{\tau}(A  \hat c_h^n - A \hat c_h^{n-1}) \hat v  + \int_{\Lambda} \hat D A \partial_s \hat c_h^n  \partial_s \hat v^n - \int_{\Lambda} A \hat u \hat c_h^n \partial_s \hat v   +  \int_{\Lambda} \xi P (\hat c_h^n - \overline{c_h^n}) \hat v +a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) = 0.
+\end{align}
+ In the above, if $\hat u$ is non--zero on $\Lambda^i$, then the artificial diffusion $a_h^{\mathrm{stab}}$ stabilization term is nonzero  and is given below, for more details see for  \cite[Section 12.6]{quarteroni2009numerical}.
+ \begin{align}
+    a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) =  \frac{1}{\|\hat u\|_{L^2(\Lambda^i)}} 
+ h_{\Lambda^i} \int_{\Lambda^i} A (\hat u \partial_s \hat c_h  ) (\hat u \partial_s \hat v),  
+ \end{align}
+ where $h_{\Lambda^i}$ is the mesh--size of $\Lambda^i$.
+Note that the condition \eqref{eq:flux_conservation} is enforced weakly in the above formulation.
+
+ \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} complemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0 = 0 $ is the following:
+
+ For $n = 1,2, \ldots $, find  $c_h^n \in V_h$ and $\hat c^n_h \in \hat V_h$ such that the coupled equations \eqref{eq:3d_dg_disc} and \eqref{eq:1d_dg_disc} hold for all $v \in V_h$ and for all $\hat v \in \hat V_h$.
+\subsubsection{Finite element solution of the incompressible Stokes equations}
+
+We consider an H(div)-based finite element approximation of the incompressible Stokes equations~\eqref{eq:stokes} defined over $\Omega_{\rm CSF}$ with the boundary conditions~\eqref{eq:stokes_bcs}.
+Following\cite{hong2016robust}, we approximate the velocity field $\bm u$ and the pressure field $p$ with the following finite element spaces:
+\begin{align*}
+  \bm V_{h,g} &= \{ \bm v  \in H(\mathrm{div}, \Omega_{\mathrm{CSF}}): \;
+  \bm v \vert_{E} \in \mathrm{BDM}^2(E), \; E \in \mathcal{T}_{\rm CSF}; \;
+  \bm v \cdot \bm n = 0
+  \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\rm AM-L} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n =  g %\frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} 
+  \text{ on } \Gamma_{\mathrm{LV}} \}  \\ 
+  Q_h  &= \{q \in L^2(\Omega_{\rm CSF}): \; q \vert_E \in \mathrm{P}^1(E),
+  \; E \in \mathcal{T}_{\rm CSF} \}. 
+\end{align*}
+Here, $H(\mathrm{div}, \Omega_{\rm{CSF}})$ is the space of $L^2(\Omega_{\rm{CSF}})$ vector fields with $L^2(\Omega_{\rm{CSF}})$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, $g$ is a given constant, and $\bm n$ is the unit outward normal vector to each facet.
+Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by
+$$ \bm v_n = (\bm v \cdot \bm n) \bm n, \quad \bm v_t = \bm v - \bm v_n. 
+$$ 
+Since $\bm V_{h,g} \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm
+  v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$, the interior facets to $\Omega_{\mathrm{CSF}}$. Continuity in the
+tangential component is enforced weakly via interior penalization. For
+convenience, we collect all facets exterior to the CSF space  $\Gamma_{\rm AM-U}$ in the set
+$$\mathcal{F}_{e} = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
+\cup \mathcal{F}_{\rm AM-L}\cup \mathcal{F}_{\rm AM-U}  \cup \mathcal{F}_{\rm SSAS}, $$ 
+where facets
+lying on the pial interface $\Gamma_{\rm pia}$ are denoted by
+$\mathcal{F}_{\rm pia}$, on the lower and upper outer (arachnoid) boundary
+$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$, respectively, 
+%on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$
+ on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
+$\mathcal{F}_{\rm SSAS}$, and on the surface of the lateral ventricles by $\mathcal{F}_{\rm LV}$. Now, define the form 
+\begin{multline}
+  \mathcal{A}_h(\bm u, \bm v)
+  = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
+ + \sum_{F \in \mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}} \left (
+  -\int_{F} \mu \{\nabla \bm u\}  \bm n_F \cdot [\bm v_t] 
+  - %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
+  \int_{F} \mu \{\nabla \bm v\}  \bm n_F \cdot [\bm u_t]
+  + %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
+\int_{F}   \frac{\sigma \mu}{h_F} [\bm u_t ] \cdot [\bm v_t] \right ),   
+\end{multline}
+where on exterior facets the average and jump operators take the one-sided values.
+We set the penalty parameter for the tangential continuity to be $\sigma = 20$.
+The finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_{h,g} \times Q_h$ with $g  = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}$  such that
+\begin{subequations}
+\begin{align}
+  \mathcal{A}_h(\bm u_h, \bm v_h)
+  + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
+  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0 \qquad \foralls \bm v_h \in \bm V_{h,0} \\ 
+\int_{\Omega_{\rm CSF}} \nabla \cdot \bm u_h \, q_h  &= 0 \qquad \foralls q_h \in Q_h.
+\end{align}
+\end{subequations}
+
+\subsubsection{Finite element solution of the perivascular network equations}
+
+We consider meshes $\mathcal{I}_a, \mathcal{I}_v$ representing a
+conforming subdivision of each of the perivascular networks
+$\Lambda_a$, $\Lambda_v$. Relative to each mesh $\mathcal{I}$, we define the
+space of (discontinuous) piecewise constants $\hat{Z}_h(\mathcal{I})$ and define
+spaces of continuous piecewise linears $\hat{V}_{h, g}(\mathcal{I})$ with
+prescribed boundary node values (on $\mathcal{N}$) given by $g$:
+\begin{align}
+  \hat{Z}_h(\mathcal{I}) &= \{z \in L^2(\Lambda) ; \, z \vert_{\Lambda_i} \in \mathrm{P}^{0}(\Lambda_i) \, \foralls \Lambda^i \in \mathcal{I}\}, \\
+  \hat{V}_{h, g}(\mathcal{I}) &= \{ v \in C^0(\Lambda) ; \, v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda_i), \, v(x) = g(x) \, \foralls x \in \mathcal{N} \} .
+  %Q_h^{\rm PVS} &= \{ q_h \in C^0(\Lambda),  q_h \vert_{\Lambda^i} \in \mathrm{P}^k(\Lambda_i), q_h \vert_{\Gamma_{\rm Pia}} = p_{\rm CSF} \}
+\end{align}
+To discretize~\eqref{eq:1d_flow} with the bifurcation
+conditions~\eqref{eq:1d_flow:bifurcations} and boundary
+conditions~\eqref{eq:1d_flow:bcs}, we use the space of continuous
+functions $\hat{Z}_h(\mathcal{I})$ to enforce the continuity of the pressure at
+bifurcation points, while the conservation of flux is enforced
+(weakly) through the variational formulation.
+The discrete variational form of the equations then reads: Find $(\hat{q}, \hat{p}) \in \hat{Z}_h(\mathcal{I}) \times \hat{V}_{h, p_0}(\mathcal{I})$ such that
+\begin{subequations}
+\begin{align}
+  \langle \hat{q}, z \rangle +  \langle z, \frac{A \kappa}{\mu} \partial_s \hat{p}  \rangle &= 0 \quad \forall z \in \hat{Z}_h(\mathcal{I}) \\
+  \langle \hat{q}, \partial_s w \rangle &= 0 \quad  \forall w \in \hat{V}_{h, 0}(\mathcal{I}).
+\end{align}
+\end{subequations}%
+where $\langle \cdot, \cdot \rangle$ denotes the $L^2(\mathcal{I})$-inner product and defined segment-wise.
+For a stability and convergence analysis of this discrete model, we refer to the reference\cite{gjerde2024directional}.
+
+\subsection{Estimating dispersion factors from pulsatile CSF flow}
+\label{sec:app:dispersion}
+
+The cardiac ($\sim$1Hz) and respiratory ($\sim$0.25Hz) cycles induce
+pulsatile flow of CSF in the ventricular system and in the cranial and
+spinal SAS. Pulsatile flow leads to dispersion which in turn may
+enhance molecular transport~\cite{taylor1953dispersion,
+  watson1983diffusion, asgari2016glymphatic, sharp2019dispersion,
+  ray2021quantitative, troyetsky2021dispersion}. To account for the
+dispersive effects over a longer time scale (hours to days), and in
+the absence of measurements or estimates of dispersion coefficients in
+human CSF spaces, we adapt existing theoretical
+estimates~\cite{watson1983diffusion, sharp2019dispersion}. More
+specifically, we compute spatially-varying dispersion enhancement
+fields $R_c$ and $R_r$ (Fig.~2G, I), associated with the
+cardiac and respiratory cycles respectively, via the algorithm
+presented below. These fields then contribute to the diffusion
+coefficient in $\Omega_{\rm CSF}$ in~\eqref{eq:3d_pde} as $D = (1 +
+R_c + R_r) D^{\rm Gad}$.
+\begin{enumerate}[label=\roman*)]
+\item
+  To account for viscous forces, we compute spatially-varying CSF
+  pressure fields $p^0_c, p^0_r$ in $\Omega_{\rm CSF}$ corresponding
+  to the Stokes flow induced by the peak volumetric reduction of the
+  CSF space in the respective cycle (Fig.~2F, H). More
+  precisely, we numerically solve the incompressible Stokes
+  equations~\eqref{eq:stokes} equipped with the following boundary
+  conditions mimicking a dilation of the brain parenchyma with the
+  spinal SAS as the only route for CSF efflux:
+  \begin{equation}\label{eq:pulsatility:bcs}
+    \bm u = 0 \text{ on } \Gamma_{\rm Pia}; \quad 
+    \mu \nabla \bm u \cdot \bm n - p \bm n = 0 \text{ on } \Gamma_{\rm SSAS}; \quad 
+    \bm u = \frac{u^{\rm in}_{\rm LV} \cdot \bm n}{|\Gamma_{\rm LV}|} \text{ on } \Gamma_{\rm LV}; \quad
+    \bm u = \frac{u^{\rm in}_{\rm AM} \cdot \bm n}{|\Gamma_{\rm skull}|} \text{ on } \Gamma_{\rm AM}.
+\end{equation}
+  In the cardiac cycle case, we set $u^{\rm in}_{\rm AM} = 6\,$ml/s \cite{causemann2022human,baledent2014imaging} and $u^{\rm in}_{\rm LV}=0.31\,$ml/s \cite{vinje2019respiratory} to solve for $p^0_c$.
+In the respiratory cycle case, we set $u^{\rm{in}}_{\rm{AM}} = 1\,$ml/s \cite{gutierrez2022effect} and $u^{\rm in}_{\rm{LV}}=0.121\,$ml/s \cite{liu2024using} to solve for $p^0_r$.
+\item
+  We also estimate the Womersley numbers $\alpha_c, \alpha_r$
+  associated with the cardiac and respiratory flow patterns,
+  respectively, by the definition
+  \begin{equation}
+    \alpha^2_i = \frac{h_{\rm SAS}^2 \omega_i \rho}{\mu} \qquad i \in \{c, r\},
+  \end{equation}
+  with CSF density $\rho = 10^3$ kg/m$^3$, a mean CSF space width $h_{\rm SAS} =
+  1.5$mm, and CSF viscosity $\mu$ given in~Table.~1.  For
+  the cardiac cycle, we consider an angular frequency $\omega_c = 2
+  \pi$, while for the respiratory cycle, we set $\omega_r = 0.5
+  \pi$. The resulting (square) Womersley numbers are $\alpha_c^2 = 20.2$
+  for the cardiac cycle and $\alpha_r^2 = 5.05$ for the respiratory
+  cycle.
+\item
+  To account for inertial forces in addition to the viscous forces, we
+  use the Womersley numbers $\alpha_c, \alpha_r$ to calculate upscaled
+  pressure fields $p^1_c, p^1_r$ from $p^0_c, p^0_r$ as:
+  \begin{equation}
+    p^1_i(x) = \left ( 1 + \frac{\alpha_i^2}{8} \right ) p_i^0(x)
+    \qquad \text{for } x \in \Omega_{\rm CSF}, \qquad i \in \{c, r\}.
+  \end{equation}
+  Note that this scaling is based on theoretical considerations on the
+  ratio of oscillatory flow to steady flow impedances in a
+  tube~\cite[Chap.~4.3.]{van1998cardiovascular}.
+\item
+  Further, assuming unsteady dispersion, we follow Sharp et
+  al.~\cite{sharp2019dispersion} to estimate local enhancement factors
+  $S_c, S_r$ from the non-dimensionalized pressure gradients:
+  \begin{equation}
+    S_i(x) = \frac{1}{\alpha_i^3} \frac{1}{\omega_i \mu/h_{\rm SAS}} | \nabla p_i^1(x) |
+    \qquad i \in \{c, r\}.
+  \end{equation}
+\item
+  Finally, we define the cardiac and respiratory dispersion
+  enhancement factors $R_c$ and $R_r$ by smoothing $S_c$ and $S_r$,
+  respectively, to account for the non-local nature of
+  dispersion. Specifically, for $i \in \{c, r\}$, we define $R_i :
+  \Omega_{\rm CSF} \rightarrow \R^{+}$ by solving a heuristic weighted
+  Helmholtz problem over $\Omega_{\rm CSF}$ with $S_i$ as the
+  right-hand side:
+  \begin{equation}
+    - 10^{-4} \Delta R_i + R_i
+    = S_i \quad \text{ on } \Omega_{\rm CSF}
+    \qquad \text{ and } \quad
+    \nabla R_i \cdot \bm n = 0, \quad \text{ on } \partial \Omega_{\rm CSF} .
+\end{equation}
+\end{enumerate}
+Considering the uncertainty associated with the validity of
+simplifying assumptions, the resulting estimates of the cardiac and
+respiratory dispersion factors $R_c$ and $R_r$ 
+should be viewed as heuristic rather than absolute.
+
+%% \mar{To examine cardiac-driven dispersion, we compute a steady CSF flow field at peak systolic blood inflow.
+We mimic the resulting reduction of CSF space volume by imposing an inflow of $6\,$ml/s \cite{baledent2014imaging, causemann2022human} and $0.31\,$ml/s \cite{vinje2019respiratory} across the upper arachnoid mater ($\Gamma_{\rm AM-U}$) and the lateral ventricle ($\Gamma_{\rm LV}$) surface, respectively.
+We allow free outflow into the spinal SAS ($\Gamma_{\rm SSAS}$).
+In addition, we upscale the obtained pressure field to account for the lack of pulsatile inertial effects in our model.}
+
+%% % respiratory pulsatility
+%% \mar{Employing the same methodology, we estimate the respiratory contribution to dispersive mixing by computing the CSF flow field driven by an inflow of $0.121\,$ml/s \cite{liu2024using} and $1\,$ml/s \cite{gutierrez2022effect} across the lateral ventricular wall and the arachnoid mater, respectively.}
+
+%% Taking the large spatial differences of pulsatile CSF flow into account, we estimate a spatially varying diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac- and respiration-induced pulsatile CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}.
+These estimates are are derived under simplifying assumptions on the problem geometry, the flow field and concentration profiles \cite{watson1983diffusion, sharp2019dispersion}.
+Therefore such an estimate is heuristic; however, our computational results qualitatively agree with observed data.
+
+%% Similar to the case of CSF flow driven by production introduced in \Cref{sec:csf_flow}, we employ a Stokes flow model to predict CSF velocities corresponding to the peak cardiac and respiratory blood inflow rates.
+In particular, we enforce a uniformly distributed inflow of $u^{\rm in}_{\rm AM}$ and $u^{\rm in}_{\rm LV}$ across the arachnoid mater ($\Gamma_{\rm AM}$) and lateral ventricular surfaces ($\Gamma_{\rm LV}$) respectively, mimicking the effect of reduced CSF volume due to blood inflow.
+We set $u^{\rm in}_{\rm AM} = 6\,$ml/s \cite{causemann2022human,baledent2014imaging} and $u^{\rm in}_{\rm LV}=0.31\,$ml/s \cite{vinje2019respiratory} in the cardiac case, and $u^{\rm{in}}_{\rm{AM}} = 1\,$ml/s \cite{gutierrez2022effect} and $u^{\rm in}_{\rm{LV}}=0.121\,$ml/s \cite{liu2024using} in the respiratory case, respectively.
+%% We prescribe a no-slip boundary condition on the pial surface $\Gamma_{\rm Pia}$ and zero-traction on the spinal boundary $\Gamma_{\rm SSAS}$, making the spinal compartment the only outflow route of CSF.
+Thus, we compute the pulsatile flow fields by augmenting the Stokes flow model (\eqref{eq:momentum_equation} and \ref{eq:divergence_equation}) with the following boundary conditions:
+%% \begin{equation}\label{eq:cardiac_flow_bcs}
+%%     \bm u = 0 \text{ on } \Gamma_{\rm Pia}; \quad 
+%%     \mu \nabla \bm u \cdot \bm n - p \bm n = 0 \text{ on } \Gamma_{\rm SSAS}; \quad 
+%%     \bm u = \frac{u^{\rm in}_{\rm LV} \cdot \bm n}{|\Gamma_{\rm LV}|} \text{ on } \Gamma_{\rm LV}; \quad
+%%     \bm u = \frac{u^{\rm in}_{\rm AM} \cdot \bm n}{|\Gamma_{\rm skull}|} \text{ on } \Gamma_{\rm AM}.
+%% \end{equation}
+
+%% The obtained pressure fields $p_{\rm cardiac}$ and $p_{\rm resp}$ only account for viscous forces, albeit cardiac and respiratory induced CSF flow is known to be unsteady and substantially impacted by transient inertial forces.
+Assuming an oscillatory flow with either the heartbeat frequency $\omega_{\rm cardiac} = 2 \pi$ or respiratory frequency $\omega_{\rm resp} = 0.5 \pi$, a mean gap width of the CSF-filled spaces of $2h=3$\,mm and CSF density $\rho=10^3\,\text{kg/m}^3$, the squared Womersley number $\alpha_i^2 = \frac{h^2 \omega_{i} \rho}{\mu}$ becomes $\alpha_{\rm cardiac}^2 = 20.2$ and $\alpha_{\rm resp}^2=5.05$, which is unsteady.
+Based on theoretical considerations on the ratio of oscillatory flow to steady flow impedances in a tube, we account for inertial forces in the pulsatile flow by upscaling the computed viscous pressure field by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter.
+4.3).
+Thus, the upscaled pressure fields become $p^{\rm up}_{i}= (1 + \alpha_{i}^2 / 8)p_i \text{ for } i = \{\mathrm{cardiac}, \mathrm{resp}\}$.
+
+%% Further assuming unsteady dispersion, we follow Sharp et al \cite{sharp2019dispersion} and estimate each enhancement factor $R_{\rm cardiac}$ and $R_{\rm resp}$ from the non-dimensionalized pressure gradient
+%% $$R_{i} \approx dP_{i}^2 / \alpha_{i}^3, \quad  \text{where} \quad  dP=\frac{|\nabla p^{\rm up}_{i}|}{\omega_{i} \mu / h} \quad \text{for} \quad i = \{\mathrm{cardiac}, \mathrm{resp}\}.$$
+%% Finally, we account for the non-local nature of the dispersion mechanism and smooth the resulting field by solving a Helmholtz problem with $dP+i^2/\alpha_i^3$ as the right-hand side
+%% \begin{equation}
+%% - 10^{-4} \Delta R_i + R_i = dP_i^2/\alpha_i^3 \quad \text{ on } \Omega_{\rm CSF} \qquad \text{ and } \quad \nabla R_i \cdot \bm n = 0, \quad \text{ on } \partial \Omega_{\rm CSF}
+%% \end{equation}
+%% to obtain the locally averaged, spatially varying diffusion enhancement factor $R_i$.
+The diffusion enhancement factor $R$ in $D^{\mathrm{eff}} = D^{\mathrm{Gad}}( 1 +R)$ is then estimated by $R \approx R_{\rm{cardiac}} + R_{\rm{resp}}$.
+
+% %\subsubsection{Cardiac Dispersion Factor Estimation}
+% %\label{sec:dispersion_details}
+
+% To compute a spatially varying estimate of the diffusion enhancement factor $R$, we consider the Stokes flow problem (\cref{eq:momnetum_equation} and \cref{eq:divergence_equation}) equipped with the boundary conditions \cref{eq:cardiac_flow_bcs} describing cardiac induced CSF flow at peak arterial pulsation.
+We use the numerical method described in \cref{sec:details_numerical_method} to solve the Stokes system and compute the non-dimensionalized local pressure difference $dP=\frac{|\nabla p| \rho}{\omega \mu / h}$ pointwise.
+Here, $\mu$ is the CSF viscosity, $\omega = 2 \pi$ is the oscillatory heart beat frequency, $2h=3$\,mm the assumed gap width and $\rho=10^3\,\text{kg/m}^3$ the density of the CSF.
+
+% In addition, we account for inertial forces in the pulsatile flow by upscaling the obtained pressure gradient by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter.
+4.3), where $\alpha^2 = \frac{h^2 \omega \rho}{\mu} = 20.2$ is the square of the Womersley number for an oscillatory flow.
+
+% Finally, we smooth the resulting field by solving a Helmholtz problem with $dP^2/alpha^3$ as the right-hand side
+
+% \begin{equation}
+% - 10^{-4} \Delta R + R = dP^2/\alpha^3 \quad \text{ on } \Omega_{\rm CSF} \qquad \text{ and } \quad \nabla R \cdot \bm n = 0 \quad \text{ on } \partial \Omega_{\rm CSF}
+% \end{equation}
+
+% to obtain the locally averaged, spatially varying diffusion enhancement factor $R$.
+
+\subsection{Estimating net perivascular flow induced by peristaltic waves}
+\label{sec:app:peristalsis}
+
+We use the theoretical framework previously introduced by Gjerde et al.~\cite{gjerde2023directional} to compute an analytic estimate of the time-averaged (or \emph{net}) flow rates $\langle Q'_i \rangle$ (mm$^3$/s) induced by peristaltic pumping in a perivascular network $\Lambda = \cup_i \Lambda_i$.
+The motion of the (inner) vascular wall is assumed to be described by a periodic traveling (peristaltic) wave of relative amplitude $\varepsilon$, wave length $\lambda$ (mm) and frequency $f$, acting normal to the wall and traveling in the antegrade direction (with the blood flow).
+By definition, the wave number is $k = 2 \pi/\lambda$ and the angular frequency is $\omega = 2 \pi f$.
+Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L_i$, baseline inner radius $R_1^i$, fixed outer radius $R_2^i$, and outer-to-inner ratio $\beta_i = R_2^i/R_1^i$.
+These geometric parameters and the assumption of annular cylindrical PVS segments yield hydraulic resistances $\mathcal{R}_{o, i}$ and additional characteristic parameters, see \cite{gjerde2023directional} for the complete definitions and schematics.
+Since the analytical estimate is derived under the assumption that $k L_i \approx \mathcal{O} (1)$~\cite{gjerde2023directional} and has been verified against numerical simulations for $k L_i$ of the order $10^{-1}-10^2$~\cite[Table I]{gjerde2023directional}, we consider it applicable for the wave lengths and vascular network data considered here in which $k L_i$ range from 0.15 to 30 for the strong vasomotion and 0.0015 to 0.30 for the cardiac waves (\Cref{tab:pvs:parameters}).
+\begin{table}
+\centering
+  \small
+  \begin{tabular}{llrc}
+    \toprule
+    Parameter & Description & Value(s)  & Ref.\\ 
+    \midrule
+    Relative PVS size $\beta$ & $\beta = \beta_i = R_2^i / R_1^i$ & 2 & $\approx$\cite{eide2024functional} \\
+    Wave frequency $f$ & Traveling wave frequency of vascular motion & $0.1-1.0$ Hz & \cite{gjerde2023directional} \\
+    Wave length $\lambda$ & Traveling wave length of vascular motion & $0.02-2$ mm & \cite{broggini2024long, gjerde2023directional} \\
+    Wave amplitude & Relative amplitude of inner wall motion & $1-10\%$ & \cite{gjerde2023directional} \\
+    \midrule
+    Cardiac frequency $f_c$ & Frequency of human cardiac pulse wave  & $1.0$ Hz & \cite{jung2021novel} \\
+    Cardiac wave length $\lambda_c$ & Wave length of human cardiac pulse wave & $2.0$m & \\
+    Cardiac amplitude $\varepsilon_c$ & Wall displacement amplitude of cardiac pulse wave & 1\%  & \\
+    \midrule
+    Vasomotion frequency $f_v$ & Frequency of slow vasomotion wave  & $0.1$ Hz & \cite{jung2021novel} \\
+    Vasomotion wave length $\lambda_v$ & Wave length of vasomotion wave & $0.02$ m & \cite{broggini2024long, gjerde2023directional} \\
+    Vasomotion amplitude $\varepsilon_v$ & Wall displacement amplitude of vasomotion wave & 10\%  & \cite{broggini2024long, gjerde2023directional}\\
+    \bottomrule
+  \end{tabular}
+  \caption{Perivascular flow induced by vascular wall motion: overview of parameters.}
+  \label{tab:pvs:parameters}
+\end{table}
+% Vascular radii r_o (min, max, avg, std) (mm):  ....
+% Vascular branch lengths L (avg, std, min, max) (mm): ...
+
+\input{tikz/fig_pvs_flow}
+
+This theoretical formalism~\cite{gjerde2023directional} is defined
+relative to a network in the form of a directed, bifurcating tree with
+a single supply node/root $i_0$. To extend to a network of cerebral
+arteries with multiple supply nodes (such as the basilar and two
+internal carotid arteries in the current data
+set~\cite{hodneland2019new}), we separate the network $\Lambda$ into
+edge-disjoint subnetworks $\Lambda^1, \Lambda^2, \Lambda^3$, one for
+each of the supply nodes (Fig.~1A--B,
+\Cref{fig:app:peristalsis}). Each node is assigned to the subnetwork
+associated with the nearest supply node, and edges between nodes are
+preserved (\Cref{fig:app:peristalsis}B). Next, we compute a minimal,
+bifurcating and directed tree representation of each subnetwork:
+$\mathcal{T}^1, \mathcal{T}^2, \mathcal{T}^3$
+(\Cref{fig:app:peristalsis}C). Each tree $\mathcal{T}^j = \cup E_n^j$
+consists of the subset of the nodes from $\Lambda^j$ that have degree
+1 (are leaf or root nodes) or degree $3$ (are true bifurcation
+points), and each path between nodes with degree $\not = 2$ in
+$\Lambda^j$ is represented in $\mathcal{T}^j$ by an edge $E^j$ with
+edge length $L$ corresponding to the total length of the original path
+and edge radius $R^j$ as the average of the path radii.
+
+For each subtree $\mathcal{T}^j = \cup_n E_n^j$, we compute the
+time-averaged downstream flow rate $\langle Q'_n \rangle$ induced by
+the vascular wall motion for each edge $n$ via~\cite[eq.~(5),
+  (34)]{gjerde2023directional} (\Cref{fig:app:peristalsis}D). We next
+assign this flow rate $\langle Q'_n \rangle$ to each of the branches
+$\Lambda_i$ that form the path $E_n^j$, thus yielding $\langle Q'_i
+\rangle$ for each perivascular segment $\Lambda_i$ while ensuring that
+mass is conserved (\Cref{fig:app:peristalsis}E). For the segment(s)
+ignored in the separation step (\Cref{fig:app:peristalsis}B), we set a
+flow rate of zero. Finally, we define the mean longitudinal
+perivascular velocity induced by the peristaltic wave $\langle
+\hat{u}^x_i \rangle$ by dividing the flow rate $\langle Q'_i \rangle$
+by the cross-section area $A_i = \pi ({R_2^i}^2 - {R_1^i}^2) = \pi
+(\beta_i^2 - 1) {R_1^i}^2$.
+
+\subsection{Numerical verification}
+\label{sec:app:verification}
+
+We assess the numerical accuracy and convergence of our simulation results by performing a series of experiments with different spatial and temporal resolutions.
+Specifically, we generate a sequence of three meshes (\Cref{fig:mesh_refinement}) with an increasing number of computational mesh vertices and cells, solve all relevant simulation steps (CSF flow and intracranial transport computations) for the baseline model on each mesh, and compare the results across meshes with respect to a set of key quantities of interest.
+The mesh refinement employed here is localized near expected sharp concentration gradients.
+This allows us to better capture dynamics and reduce undershoots, see \Cref{fig:mesh_convergence_concentrations}.
+Similarly, we investigate the effect of the time step size by solving the intracranial transport model on the standard resolution mesh with different time steps: 1, 2, and 4 minutes.
+
+Considering the mean tracer concentrations in the CSF, parenchyma and
+arterial and venous PVS domain over the first 24\,h after injection,
+we observe negligible changes with both mesh and time refinement
+(\Cref{fig:mesh_convergence_concentrations} and
+\Cref{fig:time_convergence_concentrations}). As an additional
+verification step, we compute the mean and maximum dispersion
+enhancement factor, the maximum CSF pressure and velocity in both the
+cardiac-driven and CSF production-induced flow fields, and mean
+concentrations at 3, 6, 12 and 24\,h for all mesh resolutions and time
+steps (\Cref{fig:mesh_time_convergence}). While the maximum dispersion factor
+increases by about 60\,\% from the low resolution to the standard mesh, it
+stabilizes with the next refinement step. All other quantities change
+with less than 10\,\% with mesh refinement, and less than 1\,\% with
+time refinement.
+We thus conclude that the standard resolution mesh and a time step of 2 min offer sufficient accuracy for our simulations, and remark that all reported results are obtained with the standard resolution mesh.
+
+Finally, to check for numerical mass conservation, we perform an
+additional simulation not allowing for tracer efflux across the outer
+boundary, and confirm that the total amount of tracer is preserved
+over time after the initial influx phase
+(\Cref{fig:mass_conservation}).
+\begin{figure}
+    \centering
+\includegraphics[width=0.9\linewidth]{figures/mesh_refinement.png}
+    \caption{Illustration of the three different meshes; from left to right: low resolution, standard resolution, high resolution.}
+    \label{fig:mesh_refinement}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \centering
+        \includegraphics[width = 1 \linewidth]{figures/csf_minmax.png}
+        \caption{Minimum and maximum tracer concentration in the CSF under mesh refinement}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.45\textwidth}
+        \centering
+     \includegraphics[width= 1 \linewidth]{figures/par_minmax.png}
+        \caption{Minimum and maximum tracer concentration in the parenchyma under mesh refinement}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.45\textwidth}
+        \centering
+        \includegraphics[width = 1 \linewidth]{figures/art_minmax.png}
+        \caption{Minimum and maximum tracer concentration in PVSs around arteries under mesh refinement}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.45\textwidth}
+        \centering
+     \includegraphics[width= 1 \linewidth]{figures/ven_minmax.png}
+         \caption{Minimum and maximum tracer concentration in PVSs around veins under mesh refinement}
+    \end{subfigure}
+    \caption{Minimum and maximum tracer concentrations over the first 24\,h after injection on the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the low resolution (LowRes), standard and high resolution (HighRes) meshes with a timestep of 2 min for the baseline model (Model A).}
+    \label{fig:mesh_convergence_concentrations}
+\end{figure}
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.49\textwidth}
+        \centering
+        \includegraphics[trim={0.5cm 1cm 0.05cm 0.8cm}, clip, width = 1.05 \linewidth]{figures/modelALowRes_modelA_modelAHighRes.pdf}
+        \caption*{Convergence of key quantities of interest with mesh refinement}
+        %\label{fig:mesh_convergence}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.49\textwidth}
+        \centering
+     \includegraphics[trim={0.5cm 1cm 0.05cm 0.8cm}, clip,width= 1.05 \linewidth]{figures/modelAsmalldt_modelA_modelAlargedt.pdf}
+        \caption*{Convergence of key quantities of interest with time refinement}
+        %\label{fig:time_convergence}
+    \end{subfigure}
+    \caption{For both left and right panels: A: Minimal ($\rm h_{min}$) and maximal ($\rm h_{max}$) mesh cell sizes (computed as cell circumradius $\times 2$); B: number of mesh vertices and tetrahedral cells in each mesh; C: mean cardiac dispersion enhancement factor $R$; D: maximum cardiac dispersion enhancement factor $R$; E: maximum pressure in steady CSF production flow; F: maximum pressure in cardiac-driven CSF flow; G: maximum CSF velocity in steady CSF production flow; H: maximum CSF velocity in cardiac-driven CSF flow; I--L: mean tracer concentration in the CSF, parenchyma, arterial PVS and venous PVS after 3, 6, 12 and 24 hours.}
+        \label{fig:mesh_time_convergence}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \centering
+        \includegraphics[width = 1 \linewidth]{figures/time_refinement_par_csf_mean.png}
+        \caption{Mean tracer concentration in CSF and parenchyma under time step refinement}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.45\textwidth}
+        \centering
+     \includegraphics[width= 1\linewidth]{figures/time_refinement_art_ven_mean.png}
+         \caption{Mean tracer concentration in PVSs around (veins and) arteries under time step refinement}
+    \end{subfigure}
+    \caption{Mean tracer concentrations after up to 24\,h in the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the standard resolution mesh for timesteps of 1, 2, and 4 minutes (dt of 60, 120, or 240 seconds)}
+    \label{fig:time_convergence_concentrations}
+\end{figure}
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\linewidth]{figures/modelAMassConservation_total_conc.png}
+    \caption{Total tracer content in the CSF, parenchyma, and arterial and venous PVS for a variant of the baseline model without tracer outflow.
+The total amount of tracer is constant after the initial influx phase demonstrating that the numerical scheme conserves mass globally.}
+    \label{fig:mass_conservation}
+\end{figure}
+
+\section{Supplementary discussion}
+
+\subsection{Extended model validation}
+\label{sec:app:model_validation}
+
+In addition to the comparison of our in-silico predictions of tracer
+enrichment and clearance against glymphatic MRI, we here compare
+auxiliary model quantities against the literature as additional model
+validation.
+
+\paragraph{CSF flow and pressures in the SAS and ventricular system}
+The dynamics of human CSF flow and pressure are better quantified, by
+way of clinical imaging, in-vitro studies, and computational
+modelling, in other areas of the ventricular
+system~\cite{linninger2007cerebrospinal, sweetman2011cerebrospinal,
+  vinje2019respiratory, hornkjol2022csf, causemann2022human,
+  karki2025real, liu2025transmantle}. Linninger et
+al~\cite{linninger2007cerebrospinal} model CSF flow and pressure
+dynamics induced by CSF production and cardiac pulsatility under
+normal and hydrocephalic conditions, and report of very good agreement
+with Cine (phase-contrast) MRI measurements. Our estimates of the
+maximum intracranial pressure difference, 10 Pa from the cardiac
+contribution and 26 mPa from CSF production, is in perfect agreement
+with their maximum transmantle pressure difference of $\sim$10 Pa, and
+also in very good agreement with mean pressure differences of 11.5 Pa
+measured clinically between sensors placed subdurally and in the
+lateral ventricle \cite{vinje2019respiratory}. Liu et
+al~\cite{liu2025transmantle} report of cardiac and respiratory
+pressure differences across the aqueduct of 12.1 $\pm$ 5.7 Pa and 9.5
+$\pm$ 7.2 Pa, respectively; thus our baseline estimate of the respiratory contribution 1.4 Pa may be an
+underestimation. On the other hand, our cardiac- and
+respiratory-driven CSF flow estimates peak at 19.8 cm/s and 4.8 cm/s in the caudal direction, respectively, which are higher than phase-contrast MRI measurements of cardiac and respiratory CSF flow
+components \cite{takizawa2017characterization, yildiz2017quantifying}. Some variation in CSF flow velocities is expected considering that the values from MRI represent averages
+\cite{yildiz2017quantifying} and that the geometry of the CSF spaces
+strongly affects peak velocities\cite{vinje2019respiratory,
+  causemann2022human}. Hornkjøl et al \cite{hornkjol2022csf} model the
+flow dynamics induced by CSF production in the choroid plexus and
+report a peak CSF velocity of 8.9 mm/s in the aqueduct, which is 4.8
+$\times$ higher than our values of 1.85 mm/s. Given that we use the
+same production rate, this deviation again illustrates the impact of
+potential differences in the (aqueduct) geometry on local velocities.
+
+% Dispersion predictions
+\paragraph{Dispersion in the SAS, ventricular system and PVS}
+This pulsatile flow of CSF in the SAS, ventricular system and PVSs
+leads to an increase in effective solute
+diffusivity~\cite{stockman2007effect, hettiarachchi2011effect,
+  asgari2016glymphatic, sharp2019dispersion, ray2021quantitative} via
+a process known as Taylor dispersion \cite{taylor1953dispersion,
+  watson1983diffusion}. Previous estimates of the magnitude of this
+effect in the CSF spaces vary significantly: from an enhancement
+factor of 0.05--1 in periarterial spaces surrounding penetrating
+arteries~\cite{asgari2016glymphatic, troyetsky2021dispersion}, to
+5--100 in the spinal subarachnoid space~\cite{stockman2007effect,
+  hettiarachchi2011effect, sharp2019dispersion}, and up to more than
+$10\,000$ in surface periarterial spaces~\cite{ray2021quantitative,
+  sharp2019dispersion}. The large variability can (at least partly) be
+attributed to methodological differences; e.g. different assumptions
+on the medium, domain width, pressure differences and/or fluid
+velocities, the diversity of CSF flow characteristics, as well as a
+high likelihood of spatial variations. Hornkjøl et
+al~\cite{hornkjol2022csf} consider model variations with constant
+dispersion factors from 1 up to 1000, and indicate that a value of 10
+gives the better agreement with the clinically observed
+enrichment. Our spatially-varying estimates of the dispersion
+enhancement factors $R_c, R_r$ (with $D = (1 + R_c + R_r) D^{\rm
+  Gad}$) range from 0 to 200 for the cardiac contribution $R_c$ and 0
+to 320 for the respiratory contribution $R_r$; and is thus compatible
+within the previously reported spectrum.
+
+% Effect of PVS size and shapes
+\paragraph{Shapes, sizes and structures of the PVS}
+The shapes, sizes and structures of the PVSs likely vary between
+species (e.g.~mice vs.~humans), between spatial compartments
+(e.g.~surface vs.~parenchymal), between vessel types (arteries
+vs.~arterioles vs.~veins), and in
+pathologies~\cite{ichimura1991distribution, foley2012realtime,
+  schain2017cortical, mestre2018flow, bedussi2018paravascular,
+  mestre2022periarteriolar, smets2024perivascular, raicevic2023sizes,
+  vinje2021brain, eide2024functional}. In terms of shape, the PVSs are
+commonly represented as annular (elliptic) cylinders, though it is
+well recognized that this represents an
+idealization~\cite{mestre2018flow, tithof2019hydraulic,
+  vinje2021brain, raicevic2023sizes, boster2024hydraulic,
+  smets2024perivascular}. In terms of sizes, Raicevic et
+al~\cite{raicevic2023sizes} note that the variation in PVS area is
+larger between PVS segments than along a single PVS segment and that
+the PVS area increases with lumen area. In mice, reports of the ratio
+between PVS and lumen area range from
+$\approx$0.35--0.43~\cite{smets2024perivascular} up to
+$\approx$1.12--1.4~\cite{raicevic2023sizes, mestre2018flow}. In
+humans, the PVS may be as wide as the associated surface artery and up
+to 4$\times$ wider in iNPH subjects~\cite{eide2024functional}, which
+would correspond to substantially larger PVS area ratios (3 or
+higher). To reflect the human scale, we here represent each PVS
+segment as an annular cylinder with inner radius $R_1$ and outer
+radius $R_2$ of width and area proportional to that of the
+corresponding blood vessel ($R_2 = 2 R_1$ at baseline, $R_2 = 3 R_1$
+for enlarged PVS). The hydraulic resistance of annular cross-sections
+is $1-6 \times$~\cite{tithof2019hydraulic} larger than more elongated
+cross-sections and thus our estimates of the pressure-induced PVS
+velocities are conservative.
+
+%\paragraph{Predictions on overall tracer spreading patterns}
+
+%We compare our in-silico predictions with clinical data from human subjects, in particular the studies by Ringstad et al~\cite{ringstad2018brain}, Watts et al~\cite{watts2019measuring} and Eide and Ringstad~\cite{eide2024functional}.
+Finding an overall good agreement, we provide a detailed overview on key quantities of interest in %FTA in MCA2 & N/A & N/A & 53.1 ± 50.5 min & 2:12 / 1:48 hours +$t_{st}$ \\
+%FTA in ACA2 & N/A & N/A & 48.3 ± 46.1 min & 3:00 hours +$t_{st}$ \\
+%\bottomrule
+%\end{tabular}
+%\caption{Comparison of clinical studies and model predictions on tracer spreading patterns %and arrival times.}
+%\label{tab:spreading_comparison}
+%\end{table}
 

### 2025-06-05 paper/main.tex
Updates from Overleaf

--- 2025-06-04 paper/main.tex
@@
 Our findings highlight the significant impact of cerebrospinal fluid (CSF) production and intracranial pulsatility on molecular enrichment following intrathecal tracer injection.
-We demonstrate that low-frequency vasomotion induces moderate CSF flow in surface PVS networks which substantially enhances tracer enrichment, and that impaired enrichment is a direct natural consequence of enlarged surface PVSs.
+We demonstrate that low-frequency vasomotion induces moderate CSF flow in surface PVS networks which substantially enhances tracer enrichment, and that impaired enrichment is a direct natural consequence of enlarged PVSs.
 This openly available technology platform thus provides an opportunity for integrating separate observations on diffusion in neuropil, vascular dynamics, intracranial pulsatility, CSF production, and efflux, and for exploring drug delivery and clearance in the human brain.
@@
 performance computing, we introduce an integrated computational model
-of intracranial molecular enrichment and clearance. Focusing on the
+of \mar{intracranial molecular enrichment and clearance}. Focusing on the
 interplay between perivascular pathways, pulsatility and CSF flow
@@
 
-\subsection*{In-silico predictions of intracranial molecular enrichment and clearance after intrathecal injection}
+\subsection*{In-silico predictions of intracranial \mar{molecular enrichment and clearance} after intrathecal injection}
 \begin{figure}
@@
      \caption{
-     \textbf{In-silico modelling of molecular enrichment and clearance of the PVS, SAS and parenchyma.}
+     \textbf{In-silico modelling of \mar{molecular enrichment and clearance} of the PVS, SAS and parenchyma.}
      A) Illustration of the model geometry including the CSF-filled spaces (ventricles and SAS) $\Omega_{\rm CSF}$, the parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and veins $\Lambda_{\rm vein}$, as well as their interfaces and boundaries;
@@
 maximum flow velocity of $1.85\,$mm/s (in the aqueduct) and a mean
-velocity in the SAS of $2.6$\textmu m/s (\Cref{fig:csf}E). To examine the
-dispersive effects of pulsatile CSF flow, we employ the theory of
+velocity in the SAS of $2.6$\textmu m/s (\Cref{fig:csf}E). 
+To examine the dispersive effects of pulsatile CSF flow, we employ the theory of
 shear-augmented dispersion together with computational fluid dynamics
 to determine a dispersion factor $R$ enhancing the effective solute
-diffusivity (see Methods). For the cardiac contribution, we compute
-CSF pressure and flow fields at peak systolic blood inflow,
-corresponding to a reduction in CSF space volume at a total rate of
-$6\,$ml/s~\cite{baledent2014imaging, causemann2022human} in the SAS
-and $0.31\,$ml/s across the lateral ventricle
-surface\cite{vinje2019respiratory}. This scenario sets up a pressure
-drop of $10\,$Pa ($0.075$ mmHg) between the lateral ventricles and the
-spinal SAS and a maximum flow velocity of $19.8\,$cm/s
-(\Cref{fig:csf}F, J). Assuming a cardiac frequency of $1\,$Hz, we
-infer that this cardiac-induced pulsatile CSF flow increases the
+diffusivity (see Supplementary information S1.3 for a complete description). 
+To estimate the cardiac contribution, we compute the steady state
+CSF pressure and flow fields at peak systolic blood inflow. Modelling the reduction 
+in CSF space volume corresponding to peak systolic conditions, we assume a
+total inflow rate of $6\,$ml/s~\cite{baledent2014imaging, causemann2022human} 
+in the SAS across its outer surface and $0.31\,$ml/s across the lateral ventricle surface\cite{vinje2019respiratory}. This scenario sets up a pressure drop of 
+$10\,$Pa ($0.075$ mmHg) between the lateral ventricles and the spinal 
+SAS and a maximum flow velocity of $19.8\,$cm/s (\Cref{fig:csf}F, J). 
+Assuming a cardiac frequency of $1\,$Hz, we infer that this cardiac-induced 
+pulsatile CSF flow increases the
 effective diffusion by more than two orders of magnitude in the
@@
 For the high permeability model, we increase this permeability by a factor of 100.
-We emphasize that, for both scenarios, we consider the high PVS flow regime as examined in the previous section.
+For both scenarios, we consider the high PVS flow regime as examined in the previous section.
 
@@
 
-\subsection*{Enlarged PVSs in the SAS delay periarterial and intracranial molecular enrichment}
-
-Idiopathic normal pressure hydrocephalus (iNPH), a dementia subtype,
-is associated with enlarged PVS in the SAS and impaired periarterial
-and parenchymal tracer transport~\cite{eide2024functional}. More
-generally, enlarged PVSs are associated with a range of pathophysiogical
-conditions and cognitive decline~\cite{bown2022physiology}. A key
-question is thus whether enlarged PVSs alone leads to delayed
-perivascular and intracranial tracer enrichment. To address this
-question from a fluid and transport dynamics perspective, we ask our
-in-silico model to predict the integrated effect of perivascular
-dilation on PVS flow velocities and tracer enrichment patterns
-(\Cref{fig:enlargement}A).
+\subsection*{Enlarged PVSs delay periarterial and intracranial molecular enrichment}
+
+Enlarged PVSs are associated with impaired periarterial tracer
+transport and more generally with cognitive decline as well as
+certain subtypes of
+dementia~\cite{bown2022physiology,eide2024functional}. A key
+question is thus whether enlarged PVSs alone leads to delayed perivascular
+and intracranial tracer enrichment. To address this question from a
+fluid and transport dynamics perspective, we ask our in-silico model
+to predict the integrated effect of perivascular dilation on PVS flow
+velocities and tracer enrichment patterns (\Cref{fig:enlargement}A).
 
@@
     \includegraphics[width=1 \textwidth]{figures/figure5.png}
-\caption{\textbf{PVS enlargement in the SAS reduces net PVS flow and delays tracer transport}
-A) Modelling dilated PVS in the SAS by extending the outer PVS boundaries;
+\caption{\textbf{PVS enlargement reduces net PVS flow and delays tracer transport}
+A) Modelling dilated PVS by extending the outer PVS boundaries;
 B) mean and max PVS flow velocities with normal/control and dilated PVS;
@@
 %% Summary of findings
-We have presented a high-fidelity in-silico model of molecular
-enrichment and clearance in human intracranial spaces, enabling
+We have presented a high-fidelity in-silico model of \mar{molecular
+enrichment and clearance} in human intracranial spaces, enabling
 tailored predictions of the influence of CSF space and (peri)vascular
@@
 pulsatility, as well as of pathological conditions such as enlarged
-PVSs in the SAS. A key observation is that the balance between CSF
-production and intracranial pulsatility significantly shapes the
-large-scale features of molecular enrichment, with e.g.~ventricular
-tracer reflux after intrathecal injection in the absence of CSF
-production, and preference towards anterior brain regions with reduced
-pulsatility. Indeed, the simulated distribution patterns cover a
-substantial range of the individual variations observed clinically
-\cite{ringstad2018brain}. In terms of perivascular pathways, we find
-that even moderate CSF flow on the order of 0.1--1.0 \textmu m/s in
-the surface PVSs results in earlier tracer enrichment around major
-cerebral arteries. This effect is more pronounced -- with up to a
-three-hour difference in time of arrival -- with mean PVS flow rates
-of approximately 10 \textmu m/s. Our models of CSF flow in
-perivascular networks, based on first principles and asymptotic
-analysis, predict net flow of such magnitudes and both antegrade and
-retrograde flow within the perivascular network, occurring with or
-without pressure differences at the network ends. The PVS may thus
-function as highways facilitating rapid transport, even in the absence
-of structural barriers, though a sharp concentration difference
-between the PVS and surrounding tissues is unlikely without such
-barriers. Finally, dilated PVS in the SAS will cause a substantial reduction in
-cardiac- and vasomotion-driven flow velocities, which in turn impedes
-perivascular transport.
+PVSs. A key observation is that the balance between CSF production and
+intracranial pulsatility significantly shapes the large-scale features
+of molecular enrichment, with e.g.~ventricular tracer reflux after
+intrathecal injection in the absence of CSF production, and preference
+towards anterior brain regions with reduced pulsatility. Indeed, the
+simulated distribution patterns cover a substantial range of the
+individual variations observed clinically \cite{ringstad2018brain}. In
+terms of perivascular pathways, we find that even moderate CSF flow on
+the order of 0.1--1.0 \textmu m/s in the surface PVSs results in
+earlier tracer enrichment around major cerebral arteries. This effect
+is more pronounced -- with up to a three-hour difference in time of
+arrival -- with mean PVS flow rates of approximately 10 \textmu m/s. Our
+models of CSF flow in perivascular networks, based on first principles
+and asymptotic analysis, predict net flow of such magnitudes and both
+antegrade and retrograde flow within the perivascular network,
+occurring with or without pressure differences at the network
+ends. The PVS may thus function as highways facilitating rapid
+transport, even in the absence of structural barriers, though a sharp
+concentration difference between the PVS and surrounding tissues is
+unlikely without such barriers. Finally, dilated PVS will cause a
+substantial reduction in cardiac- and vasomotion-driven flow
+velocities, which in turn impedes perivascular transport.
 
@@
 contributing to quantifying the properties of the human PVS-CSF
-interface~\cite{eide2024functional}. In this regard, we emphasize that
-we report on the isolated effect of variations in permeability on
-molecular transport, separated from any effects on fluid flow in the
-PVS. Clearly, one may expect that the presence and/or properties of
-such a barrier would affect also PVS flow. However, considering the
-uncertainty associated with PVS flow characteristics in the human SAS
-in general and the lack of computational models describing net PVS
-flow in the presence of fluid exchange, this remains challenging to
-quantify.
+interface~\cite{eide2024functional}.
 
@@
 frequency $f$, amplitude $\varepsilon$ and wave lengths $\lambda$
-(\Cref{fig:pvs_a}C). The corresponding contribution to the
-periarterial flow velocity is defined for each segment $\Lambda_i$ as
-$\langle Q'_i \rangle/A_i$ where $A_i = \pi (R_2^i - R_1^i)$ is the
-cross-section area of the PVS segment and $\langle Q_i' \rangle$ is
-the estimated mean flow rate (see Supplementary information S1.4 for
-more details). Two wall motion patterns are considered, yielding
-$\hat{u}_a^c$ and $\hat{u}_a^v$ respectively: \emph{cardiac}
-pulsations with $f = 1.0$Hz, $\lambda = 2$ m and $\varepsilon = 1\%$
-(\Cref{fig:pvs_a}D), and very low-frequency \emph{vasomotion} with $f
-= 0.1$Hz, $\lambda = 0.02$m, and $\varepsilon = 10\%$
-(\Cref{fig:pvs_a}E). No flow induced by peristalsis is considered for
-the perivenous network. In the baseline model, we combine the
-contributions from CSF production and cardiac peristalsis to set
-$\hat{u}$ in~\eqref{eq:multi_transport} as $\hat{u}_a = \hat{u}_a^p +
-\hat{u}_a^c$ in $\Lambda_a$, and $\hat{u}_v = \hat{u}_v^p$ in
-$\Lambda_v$. We remark that none of these models account for fluid
-exchange across the lateral outer interface of the PVSs (over
-$\Gamma_a, \Gamma_v$), while fluid entry and exit is allowed at the
-end nodes of the PVS networks.
+(\Cref{fig:pvs_a}C). The corresponding contribution to the periarterial
+flow velocity is defined for each segment $\Lambda_i$ as $\langle Q'_i
+\rangle/A_i$ where $A_i = \pi (R_2^i - R_1^i)$ is the cross-section
+area of the PVS segment and $\langle Q_i' \rangle$ is the estimated
+mean flow rate (see Supplementary information S1.4 for more details). Two
+wall motion patterns are considered, yielding $\hat{u}_a^c$ and
+$\hat{u}_a^v$ respectively: \emph{cardiac} pulsations with $f =
+1.0$Hz, $\lambda = 2$ m and $\varepsilon = 1\%$ (\Cref{fig:pvs_a}D), and very low-frequency \emph{vasomotion} with $f = 0.1$Hz, $\lambda =
+0.02$m, and $\varepsilon = 10\%$ (\Cref{fig:pvs_a}E).
+No flow induced by peristalsis is considered for the perivenous network.
+In the baseline model, we combine the contributions from CSF production and cardiac peristalsis to set $\hat{u}$ in~\eqref{eq:multi_transport} as
+$\hat{u}_a = \hat{u}_a^p + \hat{u}_a^c$ in $\Lambda_a$, and $\hat{u}_v
+= \hat{u}_v^p$ in $\Lambda_v$.
 
@@
     & higher PVS exchange & $-$ & 100$\xi$  &  $-$ & $-$ & $-$ & $-$  & $-$ \\ 
-    & dilated PVS in the SAS & $R_2 = 3 R_1$ & $-$ & $-$ & $-$ & $-$ & $\tilde{u}_a^p + \tilde{u}_a^c$  & $-$ \\ 
+    & dilated PVS & $R_2 = 3 R_1$ & $-$ & $-$ & $-$ & $-$ & $\tilde{u}_a^p + \tilde{u}_a^c$  & $-$ \\ 
     \bottomrule
@@
     $\tilde{u}_a^p$ are as $\hat{u}_a^c$ and $\hat{u}_a^p$,
-    respectively, but with dilated PVS. }
+    respectively, but with dilated PVS.}
 \label{tab:scenarios}

### 2025-06-05 paper/supplementary_main.tex
Updates from Overleaf

@@
   (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n}
-  &= - (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm CSF}} \cdot \bm{n}
+  &= (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm CSF}} \cdot \bm{n}
   &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
@@
   = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
-  - \sum_{F \in \mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}} \left (
-  \int_{F} \mu \{\nabla \bm u\}  \bm n_F \cdot [\bm v_t] 
+ + \sum_{F \in \mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}} \left (
+  -\int_{F} \mu \{\nabla \bm u\}  \bm n_F \cdot [\bm v_t] 
   - %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}

### 2025-06-05 paper/main.tex
Updates from Overleaf

--- 2025-06-05 paper/main.tex
@@
 20 mm, and a wall displacement of 10\% -- in the antegrade direction,
-all values currently associated with significant
+all values currently associated with major
 uncertainty~\cite{gokina1996electrical, vanveluw2020vasomotion,
-  broggini2024long}.
+  munting2023spontaneous, broggini2024long}. In particular, Broggini
+et al\cite{broggini2024long} report of near-0.1Hz vasomotion waves
+traveling both antegrade and retrograde along pial arterioles in
+mice, while Munting et al\cite{munting2023spontaneous} report of
+vessel diameter changes at similar frequencies mainly (but not
+entirely) in the retrograde direction. To the best of our knowledge,
+less is known about vasomotion wave characteristics in the human
+cerebral vasculature, though Gokina et al\cite{gokina1996electrical}
+report of spontaneous vasomotion also in human pial arteries ex-vivo.
 

### 2025-06-05 paper/main.tex
Fix subsection title.
Add thank you.
Move Methods and revamp transport description. WIP.

@@
 
-\subsection*{Molecular transport equations}
-
-We model diffusion, convection, and exchange of a molecular solute in
-the PVS networks $\Lambda_a$ and $\Lambda_v$, in the CSF spaces
-$\Omega_{\rm CSF}$ and in the brain parenchyma $\Omega_{\rm PAR}$ via
-a mixed-dimensional transport model~\cite{masri2024modelling,laurino2019derivation} over a
-timescale of minutes to hours (\Cref{fig:results1}A--B,
-\Cref{fig:csf}A--C). Specifically, for $t > 0$, we solve for a
-concentration $c = c(x, t)$ in the 3D intracranial compartments ($x
-\in \Omega_{\rm CSF}, x \in \Omega_{\rm PAR}$) and for a cross-section
-averaged concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial
-and perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$) such that
-the following equations hold:
-\begin{subequations}
-\begin{alignat}{2}
-  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma & = 0 && \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}},
-  \label{eq:multi_transport_3d}
-  \\ 
-  \partial_t (A  \hat c) - \partial_s(\hat D A \partial_s ( \hat c)) +\partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c})  &= 0 && \quad \quad \mathrm{in} \quad  \Lambda_a, \Lambda_v .
-  \label{eq:multi_transport_1d}
- \end{alignat}%
-\label{eq:multi_transport}%
-\end{subequations}%
-In~\eqref{eq:multi_transport}, $0 < \phi \leqslant 1$ is the fluid
-volume fraction, $D$ and $\hat{D}$ are effective diffusion
-coefficients, and $\bm u$ and $\hat{u}$ are fluid (CSF/ISF)
-velocities, all defined in the respective compartments; $P = 2 \pi
-R_2$ is the perimeter and $A = \pi (R_2^2 - R_1^2)$ the area of the
-PVS cross-sections, $\delta_{\Gamma}$ is a Dirac term concentrated on
-the interfaces $\Gamma_a, \Gamma_v$, the overlined $\overline{c}$
-denotes the average of $c$ over cross-sections of $\Gamma_a,
-\Gamma_v$, and $\xi$ is a permeability allowing for transfer/exchange
-over the interfaces $\Gamma_a, \Gamma_v$ between the periarterial and
-perivenous networks and their surroundings.  Moreover, we model the
-membrane between the parenchyma and the CSF spaces as a semi-permeable
-membrane with permeability coefficient $\beta_{\rm pia}$. For a
-complete description of this mathematical model and simulation
-scenarios, see the Supplementary information S1. The initial concentrations
-in the three-dimensional domain $\Omega$ and networks $\Lambda_a,
-\Lambda_v$ are all set to zero.
-
-\subsection*{Intracranial influx after intrathecal injection}
-To represent molecular influx into the intracranial compartments
-following intrathecal injection, without modelling the spinal
-compartment explicitly, we prescribe a time-dependent molecular influx
-over the interface towards the spinal CSF compartment with the condition
-\begin{equation*}
-  (D \nabla c - \bm u c ) \cdot \bm{n} = g_{\mathrm{influx}},  \quad \mathrm{on}  \quad \Gamma_{\mathrm{SSAS}}.
-\end{equation*}
-Here we set
-\begin{equation*}
-  g_\mathrm{influx}(t) = \frac{m_{\rm tot}}{T_{\rm max}^2 |\Gamma_{\rm SSAS}|} \max(0, T_{\rm max} - |t - T_{\rm max}|), 
-\end{equation*}
-which expresses a hat-shaped influx function with a peak at $T_{\rm max}$ and a total tracer injection of $m_{\rm tot}$ (\Cref{tab:parameters}).
-
-\subsection*{Molecular clearance from the intracranial space and perivascular network ends}
-
-We model molecular clearance, into the meningeal lymphatics or other
-pathways, across the upper, outer (arachnoid) boundary
-(\Cref{fig:results1}A, \Cref{fig:csf}C), proportional to the
-concentration in the SAS with a rate constant $\beta_{\rm exit}$. We
-set a no-flux condition at the end nodes of the perivascular networks,
-not allowing for direct molecular clearance there.
-
 \subsection*{CSF flow in the SAS and ventricular system}
@@
 end nodes of the PVS networks.
+
+\subsection*{Molecular transport equations}
+
+We represent the evolution and distribution of a solute in the CSF
+spaces $\Omega_{\rm CSF}$, brain parenchyma $\Omega_{\rm PAR}$, and
+periarterial and perivenous networks $\Lambda_a \cup \Lambda_v$ via a
+mixed-dimensional transport model~\cite{masri2024modelling} over a timescale
+of minutes to hours (\Cref{fig:results1}A--B, \Cref{fig:csf}A--C). The
+current model accounts for solute exchange between the CSF spaces,
+brain parenchyma, and the PVSs, diffusion within all compartments,
+dispersion and convection in the CSF spaces, and convection within the
+PVS. Specifically, for $t > 0$, we solve for a concentration field
+$c(x, t)$ in the 3D intracranial compartments (for $x \in \Omega_{\rm
+  CSF} \cup \Omega_{\rm PAR}$) and for a cross section-averaged
+concentration field $\hat{c}(s, t)$ in the perivascular networks (for
+$s \in \Lambda_a, \Lambda_v$) satisfying the following equations:
+\begin{subequations}
+\begin{alignat}{2}
+  \partial_t c - \nabla \cdot D (1 + R) \nabla c  + \nabla \cdot (\bm u_{\rm CSF} c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma & = 0 && \quad \quad \mathrm{in} \quad \Omega_{\rm CSF},
+  \label{eq:multi_transport_3d_CSF}
+  \\ 
+  \partial_t (\phi c) - \nabla \cdot D^{\ast} \nabla (\phi c)  + \nabla \cdot (\bm u_{\rm PAR} c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma & = 0 && \quad \quad \mathrm{in} \quad \Omega_{\rm{PAR}},
+  \label{eq:multi_transport_3d_PAR}
+  \\ 
+  \partial_t (A  \hat c) - \partial_s(\hat D A \partial_s ( \hat c)) +\partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c})  &= 0 && \quad \quad \mathrm{in} \quad  \Lambda_a, \Lambda_v .
+  \label{eq:multi_transport_1d}
+ \end{alignat}%
+\label{eq:multi_transport}%
+\end{subequations}%
+In~\eqref{eq:multi_transport_3d_CSF}, $D$ is the (free) diffusion
+coefficient of the solute, $R$ is a dispersion coefficient, $\bm
+u_{\rm CSF}$ is a CSF velocity, $\xi$ is a permeability allowing for
+transfer/exchange across the interfaces $\Gamma_a, \Gamma_v$ between
+the periarterial and perivenous networks and their surroundings, the
+overlined $\overline{c}$ denotes the average of $c$ over
+cross-sections of $\Gamma_a, \Gamma_v$, and $\delta_{\Gamma}$ is a
+Dirac term concentrated on $\Gamma_a, \Gamma_v$. Additionally, in
+\eqref{eq:multi_transport_3d_PAR}, $\phi$ is the brain extracellular
+volume fraction\footnote{We set $\phi = 1.0$ in $\Omega_{\rm CSF}$ and
+therefore omit $\phi$ in~\eqref{eq:multi_transport_3d_CSF}.},
+$D^{\ast}$ is the effective diffusion coefficient in the parenchyma,
+and $\bm u_{\rm PAR}$ is an interstitial fluid velocity. In this
+study, we set $\bm u_{\rm PAR} = 0$. In~\eqref{eq:multi_transport_1d},
+$A = \pi (R_2^2 - R_1^2)$ is the area and $P = 2 \pi R_2$ is the
+perimeter of the PVS cross-sections, $\hat{D}$ is the effective
+diffusion coefficient in the PVSs, and $\hat{u}$ is the fluid velocity
+(in the axial direction) in the PVS. Further, we model the membrane
+between the parenchyma and the CSF spaces as a semi-permeable membrane
+with permeability coefficient $\beta_{\rm pia}$. For an extended description of the mathematical model and simulation
+scenarios, see the Supplementary information S1. The initial
+concentrations in the three-dimensional domain $\Omega$ and networks
+$\Lambda_a, \Lambda_v$ are all set to zero.
+
+%% \subsection*{Molecular transport equations}
+
+%% We model diffusion, convection, and exchange of a molecular solute in
+%% the PVS networks $\Lambda_a$ and $\Lambda_v$, in the CSF spaces
+%% $\Omega_{\rm CSF}$ and in the brain parenchyma $\Omega_{\rm PAR}$ via
+%% a mixed-dimensional transport model~\cite{masri2024modelling,laurino2019derivation} over a
+%% timescale of minutes to hours (\Cref{fig:results1}A--B,
+%% \Cref{fig:csf}A--C). Specifically, for $t > 0$, we solve for a
+%% concentration $c = c(x, t)$ in the 3D intracranial compartments ($x
+%% \in \Omega_{\rm CSF}, x \in \Omega_{\rm PAR}$) and for a cross-section
+%% averaged concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial
+%% and perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$) such that
+%% the following equations hold:
+%% \begin{subequations}
+%% \begin{alignat}{2}
+%%   \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma & = 0 && \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}},
+%%   \label{eq:multi_transport_3d}
+%%   \\ 
+%%   \partial_t (A  \hat c) - \partial_s(\hat D A \partial_s ( \hat c)) +\partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c})  &= 0 && \quad \quad \mathrm{in} \quad  \Lambda_a, \Lambda_v .
+%%   \label{eq:multi_transport_1d}
+%%  \end{alignat}%
+%% \label{eq:multi_transport}%
+%% \end{subequations}%
+%% In~\eqref{eq:multi_transport}, $0 < \phi \leqslant 1$ is the fluid
+%% volume fraction, $D$ and $\hat{D}$ are effective diffusion
+%% coefficients, and $\bm u$ and $\hat{u}$ are fluid (CSF/ISF)
+%% velocities, all defined in the respective compartments; $P = 2 \pi
+%% R_2$ is the perimeter and $A = \pi (R_2^2 - R_1^2)$ the area of the
+%% PVS cross-sections, $\delta_{\Gamma}$ is a Dirac term concentrated on
+%% the interfaces $\Gamma_a, \Gamma_v$, the overlined $\overline{c}$
+%% denotes the average of $c$ over cross-sections of $\Gamma_a,
+%% \Gamma_v$, and $\xi$ is a permeability allowing for transfer/exchange
+%% over the interfaces $\Gamma_a, \Gamma_v$ between the periarterial and
+%% perivenous networks and their surroundings.  Moreover, we model the
+%% membrane between the parenchyma and the CSF spaces as a semi-permeable
+%% membrane with permeability coefficient $\beta_{\rm pia}$. For a
+%% complete description of this mathematical model and simulation
+%% scenarios, see the Supplementary information S1. The initial concentrations
+%% in the three-dimensional domain $\Omega$ and networks $\Lambda_a,
+%% \Lambda_v$ are all set to zero.
+
+\subsection*{Intracranial influx after intrathecal injection}
+To represent molecular influx into the intracranial compartments
+following intrathecal injection, without modelling the spinal
+compartment explicitly, we prescribe a time-dependent molecular influx
+over the interface towards the spinal CSF compartment with the condition
+\begin{equation*}
+  (D \nabla c - \bm u c ) \cdot \bm{n} = g_{\mathrm{influx}},  \quad \mathrm{on}  \quad \Gamma_{\mathrm{SSAS}}.
+\end{equation*}
+Here we set
+\begin{equation*}
+  g_\mathrm{influx}(t) = \frac{m_{\rm tot}}{T_{\rm max}^2 |\Gamma_{\rm SSAS}|} \max(0, T_{\rm max} - |t - T_{\rm max}|), 
+\end{equation*}
+which expresses a hat-shaped influx function with a peak at $T_{\rm max}$ and a total tracer injection of $m_{\rm tot}$ (\Cref{tab:parameters}).
+
+\subsection*{Molecular clearance from the intracranial space and perivascular network ends}
+
+We model molecular clearance, into the meningeal lymphatics or other
+pathways, across the upper, outer (arachnoid) boundary
+(\Cref{fig:results1}A, \Cref{fig:csf}C), proportional to the
+concentration in the SAS with a rate constant $\beta_{\rm exit}$. We
+set a no-flux condition at the end nodes of the perivascular networks,
+not allowing for direct molecular clearance there.
 
@@
 
-\subsection*{Model validation}
+\subsection*{Comparison with literature results}
 
@@
 generously sharing the multi-modal human MR data, and to Prof. Erlend
-Hodneland for facilitating this data exchange. M.E.R.~acknowledges
-support from Stiftelsen Kristian Gerhard Jebsen via the K.~G.~Jebsen
-Centre for Brain Fluid Research. M.C., R.M., and M.E.R.~acknowledge
-support from the Research Council of Norway (RCN) via grant \#324239
-(EMIx). Simulations were performed on the Experimental Infrastructure
-for Exploration of Exascale Computing (eX3), which is financially
-supported by the RCN under contract 270053. M.K.~acknowledges support
-from the RCN grant \#303362. The figures 1A, 1D, 2C, 3C, 5A and 6A
-were created in Biorender.
+Hodneland for facilitating this data exchange. The authors also thank
+Profs.~Per Kristian Eide, Geir Ringstad and Rune Enger for valuable
+input and discussions. M.E.R.~acknowledges support from Stiftelsen
+Kristian Gerhard Jebsen via the K.~G.~Jebsen Centre for Brain Fluid
+Research. M.C., R.M., and M.E.R.~acknowledge support from the Research
+Council of Norway (RCN) via grant \#324239 (EMIx). Simulations were
+performed on the Experimental Infrastructure for Exploration of
+Exascale Computing (eX3), which is financially supported by the RCN
+under contract 270053. M.K.~acknowledges support from the RCN grant
+\#303362. The figures 1A, 1D, 2C, 3C, 5A and 6A were created in
+Biorender.
 
@@
 M.C. analyzed the results and prepared the figures.
-M.C. M.K. and R.M and M.E.R wrote the manuscript. 
+M.C., M.K., R.M and M.E.R wrote the manuscript. 
 All authors edited and reviewed the manuscript.

### 2025-06-06 paper/supplementary_r0.tex
Updates from Overleaf

--- 2025-06-05 paper/supplementary.tex
@@
 We use the theoretical framework previously introduced by Gjerde et al.~\cite{gjerde2023directional} to compute an analytic estimate of the time-averaged (or \emph{net}) flow rates $\langle Q'_i \rangle$ (mm$^3$/s) induced by peristaltic pumping in a perivascular network $\Lambda = \cup_i \Lambda_i$.
-The motion of the (inner) vascular wall is assumed to be described by a periodic traveling (peristaltic) wave of relative amplitude $\varepsilon$, wave length $\lambda$ (mm) and frequency $f$, acting normal to the wall and traveling in the antegrade direction (with the blood flow).
+The motion of the (inner) vascular wall is described by a periodic traveling (peristaltic) wave of relative amplitude $\varepsilon$, wave length $\lambda$ (mm) and frequency $f$, acting normal to the wall.
 By definition, the wave number is $k = 2 \pi/\lambda$ and the angular frequency is $\omega = 2 \pi f$.

### 2025-06-06 scripts/make_video.py
minor change

@@ -11,11 +11,13 @@ import seaborn as sns
 
 def make_movie(modelname:str):
     config = read_config(f"configfiles/{modelname}.yml")
-    dt , T= config["dt"], config["T"]
+    dt , T= config["dt"]*config["output_frequency"], config["T"]
     #cmap = sns.blend_palette(["lightsteelblue", "orange", "darkviolet"],as_cmap=True)
     cmap = from_k3d(pcm.Yellow___Gray___Blue).reversed()
     clim = (0,2)
-    times = np.arange(0, T + dt, dt*config["output_frequency"])
+    N = 300
+    times = np.arange(0, T + dt, dt)
+    print(times)
     scalarbar_args = dict(color="white", title="concentration (mmol/l)", 
                           vertical=False, height=0.08, width=0.6,
                           position_x=0.2, position_y=0.05, 
@@ -31,8 +33,12 @@ def make_movie(modelname:str):
     for t in times:
         sas[f"c_{t}"] = np.where(sas[f"c_{t}"] > clim[1], 
                                  clim[1], sas[f"c_{t}"])
-    img = pv.create_grid(sas, dimensions=(300, 300, 300))
+    img = pv.create_grid(sas, dimensions=(N, N, N))
     img = img.sample(sas, progress_bar=True)
+    print("interpolating...")
+    for t in tqdm(times[:-1]):
+        img[f"c_{int(t + dt/2)}"] = 0.5 *(img[f"c_{t}"] + img[f"c_{t + dt}"])
+        arteries[f"c_{int(t + dt/2)}"] = 0.5 *(arteries[f"c_{t}"] + arteries[f"c_{t + dt}"])
 
     pl = pv.Plotter(off_screen=True, window_size=(1200, 1200))
     pl.set_background("black")
@@ -47,12 +53,15 @@ def make_movie(modelname:str):
     pl.add_mesh(arteries, scalars="c", cmap=cmap, 
                 clim=clim, show_scalar_bar=False,)
     pl.camera_position = 'xz'
-    pl.camera.zoom = 1.4
+    pl.camera.focal_point = np.array(skull.center) - np.array([0,0,0.011])
+    pl.camera.zoom(1.4)
+
     pl.show(auto_close=False) 
     pl.write_frame() 
 
     # generate frames
-    for t in tqdm(list(times[1:]) + [times[-1]]*24*10):
+    frametimes = np.arange(0, T + dt/2, dt/2, dtype=int)
[…]

### 2025-06-06 scripts/plot_conc_at_label.py
change output times

@@ -12,7 +12,7 @@ def plot_conc_at_label(model:str):
     config = read_config(f"configfiles/{model}.yml")
 
     dt , T= config["dt"], config["T"]
-    times = np.arange(0, T + dt, 3*dt*config["output_frequency"])
+    times = np.arange(0, T + dt, 2*dt*config["output_frequency"])
     set_plotting_defaults()
 
     artlabelsets = [["BA",],#["ICA-L", "ICA-R"], 

### 2025-06-06 paper/old/supplementary_main.tex

--- 2025-06-06 paper/supplementary_r0.tex
@@
-\documentclass[fleqn,10pt]{wlscirep}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{bm}
-\usepackage{subcaption}
-% \captionsetup[subfigure]{justification=justified,singlelinecheck=false}
-\usepackage{tikz}
-\usepackage{newtxmath} % MER: Prettier math
-\usepackage{cleveref} % MER: Allows \Cref (but use Section X.Y not Subsection X.Y)
-\usepackage{booktabs} % MER: Prettier tables
-
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{tikz}
-\usepackage{float}
-\usepackage{amsmath}
-\usepackage{siunitx}
-\usepackage{placeins}
-\usepackage{lineno}
-\usetikzlibrary{quotes}
-\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,positioning,fit,petri}
-
-\graphicspath{{figures/}}
-
-\renewcommand\thefigure{S\arabic{figure}}    
-\renewcommand\thesection{S\arabic{section}}
-
-\captionsetup[subfigure]{justification=justified,singlelinecheck=false}
-
-\title{Supplementary information: In-silico molecular enrichment and clearance of the human intracranial space}
-\author[1]{Marius Causemann}
-\author[1]{Miroslav Kuchta}
-\author[2]{Rami Masri}
-\author[1,3,*]{Marie E. Rognes }
-\affil[1]{Department of Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway}
-\affil[2]{Division of Applied Mathematics, Brown University, Providence, Rhode Island, USA}
-\affil[3]{K. G. Jebsen Centre for Brain Fluid Research, University of Oslo, Norway}
-\affil[*]{meg@simula.no}
-
-\begin{abstract}
-  Supplementary methods and discussion accompanying the main text of
-  \emph{In-silico molecular enrichment and clearance of the human
-  intracranial space}.
-\end{abstract}
-
-\begin{document}
-
-\linenumbers
-\flushbottom
-\maketitle
-
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\foralls}{\forall \,}
-
 \section{Supplementary methods}
@@
 %\end{table}
-
-\bibliography{references}
-
-\end{document}

### 2025-06-07 paper/main.tex
More shuffling
revision from Marius I

@@
 performance computing, we introduce an integrated computational model
-of intracranial molecular enrichment and clearance. Focusing on the
+of intracranial molecular transport. Focusing on the
 interplay between perivascular pathways, pulsatility and CSF flow
@@
 
-\subsection*{In-silico predictions of intracranial molecular enrichment and clearance after intrathecal injection}
+\subsection*{In-silico predictions of intracranial transport after intrathecal injection}
 \begin{figure}
@@
      \caption{
-     \textbf{In-silico modelling of molecular enrichment and clearance of the PVS, SAS and parenchyma.}
+     \textbf{In-silico modelling of transport of the PVS, SAS and parenchyma.}
      A) Illustration of the model geometry including the CSF-filled spaces (ventricles and SAS) $\Omega_{\rm CSF}$, the parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and veins $\Lambda_{\rm vein}$, as well as their interfaces and boundaries;
@@
 Simulating a glymphatic MRI protocol~\cite{ringstad2017glymphatic,
-  ringstad2018brain, eide2024functional}(\Cref{fig:results1}C), we then predict the
-spreading of $0.5$ mmol intrathecally injected Gadobutrol after its
-appearance at the craniocervical junction
-(\Cref{fig:results1}D--E). After one hour, the in-silico
+ringstad2018brain, eide2024functional}(\Cref{fig:results1}C), we then predict the spreading of $0.5$ mmol intrathecally injected Gadobutrol after its appearance at the craniocervical junction, represented by tracer inflow across the spinal SAS boundary over the first two hours (\Cref{fig:results1}D--E).
+After one hour, the in-silico
 tracer moves upwards in the SAS frontally of the brainstem, and
@@
 
-The enrichment patterns observed clinically after intrathecal
+The tracer enrichment patterns observed clinically after intrathecal
 injection of contrast differ substantially between subjects and
@@
 dispersive effects of cardiac and respiratory pulsatile CSF flow
-(\Cref{fig:csf}A--C). Imposing a constant CSF production of
-$400\,$ml/day across the surface of the lateral ventricles, while
-allowing for efflux across the upper convexity, yields a total CSF
-pressure drop of $26\,$mPa (0.00020 mmHg) (\Cref{fig:csf}D) with a
+(\Cref{fig:csf}A--C). Beginning with the advective, net flow flield driven by 
+CSF production, we impose a constant CSF inflow of $400\,$ml/day across the surface
+of the lateral ventricles. Simultaneously allowing for efflux across the upper 
+convexity yields a total CSF pressure drop of $26\,$mPa (0.00020 mmHg) (\Cref{fig:csf}D) with a
 maximum flow velocity of $1.85\,$mm/s (in the aqueduct) and a mean
-velocity in the SAS of $2.6$\textmu m/s (\Cref{fig:csf}E). To examine the
+velocity in the SAS of $2.6$\textmu m/s (\Cref{fig:csf}E). Next, to examine the
 dispersive effects of pulsatile CSF flow, we employ the theory of
@@
 to determine a dispersion factor $R$ enhancing the effective solute
-diffusivity (see Methods). For the cardiac contribution, we compute
-CSF pressure and flow fields at peak systolic blood inflow,
-corresponding to a reduction in CSF space volume at a total rate of
-$6\,$ml/s~\cite{baledent2014imaging, causemann2022human} in the SAS
-and $0.31\,$ml/s across the lateral ventricle
-surface\cite{vinje2019respiratory}. This scenario sets up a pressure
+diffusivity (see Supplementary information S1.3 for a complete description). 
+To estimate the cardiac contribution, we compute the steady state
+CSF pressure and flow fields at peak systolic blood inflow.
+Modelling the reduction in CSF space volume corresponding to peak systolic
+conditions, we assume a total inflow rate of $6\,$ml/s~\cite{baledent2014imaging, causemann2022human}
+in the SAS across its outer surface and $0.31\,$ml/s 
+across the lateral ventricle surface\cite{vinje2019respiratory}. 
+This scenario sets up a pressure
 drop of $10\,$Pa ($0.075$ mmHg) between the lateral ventricles and the
@@
 
-Now, to examine how CSF pulsatility affects the enrichment patterns,
+Now, to examine how CSF pulsatility affects the tracer enrichment patterns,
 we consider three variations of the baseline: (i) no CSF production,
@@
 parenchyma (\Cref{fig:results1}A--B, \Cref{fig:pvs_a}A) and includes
-major surface arteries of radius 0.5--1.4 mm. Imposing the pressure
-field induced by CSF production at the periarterial network ends
-induces slow steady CSF flow of variable direction in these PVSs, with
+major surface arteries of radius 0.5--1.4 mm.
+Due to imaging limitiations, the network primarily covers the supratentorial regions, with the lowest network point located approximately $4\,$cm above the craniocervical junction.
+Imposing the pressure field induced by CSF production at the periarterial network ends induces slow steady CSF flow of variable direction in these PVSs, with
 an average velocity of 0.08 \textmu m/s (antegrade) and a maximum
@@
     C) 2D slices showing zoom-in to the region surrounding the basal artery after 1, 2, 3, and 4 hours for a low (above) and high (below) PVS-CSF permeability model.
-Innermost circle indicates the artery (blue), the surrounding annulus is the PVS (color represents concentration value), the surroundings are parenchyma and/or CSF spaces, separated by the pia (blue: towards the parenchyma, pink: towards the CSF));
+Innermost circle indicates the artery (blue), the surrounding annulus is the PVS (color represents concentration value), the surroundings are parenchyma and/or CSF spaces, separated by the pia (blue: towards the parenchyma, pink: towards the CSF).
+The cutting planes are normal to the blood vessels;
     D) As in (C) for regions surrounding (from upper to lower) the ACA-A3, ACA-A2, MCA-R and MCA-L segments;}
@@
 E) 2D slices showing zoom-in to the region surrounding the MCA-R after 1, 2, 3, and 4\,h with normal (lower) and dilated (upper) PVS.
-Innermost circle indicates the artery (blue), the surrounding annulus is the PVS (color represents concentration value), the surroundings are parenchyma and/or CSF spaces, separated by the pia (blue: towards the parenchyma, pink: towards the CSF));
+Innermost circle indicates the artery (blue), the surrounding annulus is the PVS (color represents concentration value), the surroundings are parenchyma and/or CSF spaces, separated by the pia (blue: towards the parenchyma, pink: towards the CSF).The cutting plane is normal to the blood vessel;
 F) First-time arrival for different branches of the arterial tree for the baseline model (with pressure-driven and cardiac-induced net PVS flow) and the high PVS flow model (also including vasomotion-induced net PVS flow) for normal and dilated PVS;
@@
 We find that about 20\% of the tracer reaching the cranium enters the parenchyma, which aligns with the reported peak concentration values of $0.5\,$mmol/l in the SAS and $0.1\,$mmol/l in the parenchyma reported by Watts et al~\cite{watts2019measuring}.
-In terms of perivascular tracer
-enrichment, Eide and Ringstad~\cite{eide2024functional} report of
-first-time tracer appearance in the PVS of the MCA after 37.8 $\pm$
-47.0 (M1), 53.1 $\pm$ 50.5 (M2) and 82.6 $\pm$ 61.5 (M3) min after injection, thus with just a 15.3 min delay between M2 and M1.
-Comparing with our FTAs in the MCA segments, we note that the baseline model predicts a more-than 3 hour difference in FTA between the M1 and M2.
-In contrast, the high PVS flow scenario predicts a 24 min delay between these segments.
-Moreover, the high PVS flow scenario predicts FTAs of
-84--108 min (M1) and 108--132 min (M2) after tracer appearance at the craniocervical junction.
-Assuming a tracer travel time from the point of injection to the craniocervical junction of $13\,$min (as reported by Eide and Ringstad \cite{eide2024functional}), our predictions fall at the upper end of the clinical ranges.
-This can partly be attributed to the limited caudal extend of our PVS network, with the lowest point located approximately $4\,$cm above the craniocervical junction due to imaging constrains.
-We interpret these results as indicative of a net CSF flow in human surface PVS beyond what would be induced by CSF production and cardiac peristalsis alone.
-% M2 - M1 (mean) ~ 15.3 min
+
+Our simulation results highlight that reproducing clinically observed tracer timings requires flow drivers beyond CSF production and cardiac peristalsis.
+To illustrate, we compared our findings to tracer arrival times reported by Eide and Ringstad~\cite{eide2024functional}, who observed a delay of just 15.3 min between the M1 and M2 segments of the MCA.
+Our baseline model, accounting only for CSF production and cardiac-driven flow, predicted a physiologically unrealistic delay of over three hours.
+In sharp contrast, our 'high PVS flow' scenario, which incorporates the effects of low-frequency vasomotion, yielded a much more consistent delay of 24 minutes.
+Furthermore, after adjusting for the 13 min travel time from injection to the craniocervical junction~\cite{eide2024functional}, the 'high PVS flow' model's FTAs (84––108 min for M1) aligns well with the upper end of the reported clinical range (37.8 $\pm$ 47.0), particularly when considering potentially delayed transport due to the limited caudal extend of our PVS network.
+Therefore, we interpret these results as strong indication that PVS flow driven by CSF production and cardiac peristalsis alone is insufficient to explain rapid perivascular transport observed in humans.
 
@@
 
+For an extended description of the mathematical model and simulation
+scenarios, see the Supplementary information S1.
+
 \subsection*{Intracranial compartments: ventricular system, SAS, and brain parenchyma}
@@
 between the parenchyma and the CSF spaces as a semi-permeable membrane
-with permeability coefficient $\beta_{\rm pia}$. For an extended description of the mathematical model and simulation
-scenarios, see the Supplementary information S1. The initial
+with permeability coefficient $\beta_{\rm pia}$. The initial
 concentrations in the three-dimensional domain $\Omega$ and networks
@@
 
+\subsection*{Molecular clearance from the intracranial space and perivascular network ends}
+
+We model molecular clearance, into the meningeal lymphatics or other
+pathways, across the upper, outer (arachnoid) boundary
+(\Cref{fig:results1}A, \Cref{fig:csf}C), proportional to the
+concentration in the SAS with a rate constant $\beta_{\rm exit}$. We
+set a no-flux condition at the end nodes of the perivascular networks,
+not allowing for direct molecular clearance there.
+
 \subsection*{Intracranial influx after intrathecal injection}
@@
 
-\subsection*{Molecular clearance from the intracranial space and perivascular network ends}
-
-We model molecular clearance, into the meningeal lymphatics or other
-pathways, across the upper, outer (arachnoid) boundary
-(\Cref{fig:results1}A, \Cref{fig:csf}C), proportional to the
-concentration in the SAS with a rate constant $\beta_{\rm exit}$. We
-set a no-flux condition at the end nodes of the perivascular networks,
-not allowing for direct molecular clearance there.
-
 \subsection*{Model and material parameters}
@@
 spaces ($\xi = \xi_{\rm CSF}$) or parenchyma ($\xi = \xi_{\rm EF}$),
-or neither (see also Supplementary information S1). Additionally, we
-increase the permeability $\xi$ ($100\times$) at the the leaf nodes of
-$\Lambda_a$, linearly decreasing to the baseline permeability at the
-next network node ($\approx0.4\,$mm), to account for the transition
-from explicitly resolved PVSs to their unresolved continuation in the
-CSF spaces and parenchyma.
+or neither (see also Supplementary information S1).
+To account for the unresolved continuation of the perivascular network beyond the imaging resolution limit, we implemented a permeability gradient to avoid artificial tracer buildup.
+At the terminal 'leaf nodes' of $\Lambda_a$, permeability was increased by a factor of 100 to simulate tracer exchange with the surrounding CSF/tissue, before linearly decreasing to the baseline value at the adjacent network node.
 \begin{table}
@@
       Molecular efflux rate & $\beta_{\rm exit}$ & $10^{-4}$ & mm$^2$/s & \cite{hornkjol2022csf} \\
-      Total molecular volume & $m_{\rm tot}$ & $0.5$ & mmol & \cite{eide2024functional} \\
+      Total amount of injected tracer & $m_{\rm tot}$ & $0.5$ & mmol & \cite{eide2024functional} \\
       Molecular influx, time-to-peak & $T_{\rm max}$ & $1$ & h & $\ast$ \\

### 2025-06-09 paper/main.tex
revision I changes by Marius  II

@@
 
-\subsection*{Perivascular flow shapes and accelerates molecular enrichment}
+\subsection*{Perivascular flow shapes and accelerates molecular transport}
 \label{sec:pvs_flow_results}
@@
 We study the effect of such rapid PVS flow on intracranial molecular
-enrichment by comparing the more conservative baseline model and a
+transport by comparing the more conservative baseline model and a
 high PVS flow model, where the former still includes the net flow
@@
     \includegraphics[width=0.99\textwidth]{figures/figure3b.png}
-    \caption{\textbf{Perivascular flow shapes and accelerates molecular enrichment.}
+    \caption{\textbf{Perivascular flow shapes and accelerates molecular transport.}
     A) Comparison of mean tracer concentration in the PVS and over the outer PVS surface at key locations of the arterial tree, for the baseline (lower) and high PVS flow (upper) models, with labels: ``peak'': time-to-peak (h), ``$\Delta t$'': time difference between peaks (h), and ``FTA'': first-time tracer arrival (h).
@@
 pulsatility at smaller spatial or temporal scales.
+This approach of modeling pulsatility through an effective dispersion factor is a simplification that does not capture the full complexity of transient fluid dynamics, for instance caused by non-uniform wall motion and ependymal cilia within the ventricular system~\cite{siyahhan2014flow}.
+However, it is further supported by recent clinical observations.
+High-resolution MRI measurements by Hirschler et al.~\cite{hirschler2024region} revealed that CSF mobility in regions with high pulsatility is approximately tenfold greater than the self-diffusion of water, an enhancement factor consistent with the values computed in our model.
 

### 2025-06-09 paper/main.tex
revision I from Marius III

@@
 we consider three variations of the baseline: (i) no CSF production,
-(ii) reduced pulsatility and thus decreased dispersion ($0.1 \times
-R$), and (iii) higher pulsatility with increased dispersion ($10
-\times R$). Without CSF production, transport is considerably delayed
+(ii) reduced pulsatility and thus decreased dispersion, and (iii) higher pulsatility with increased dispersion.
+We computed the reduced and increased pulsatility scenarios by halving and doubling the inflow rates associated with both cardiac and respiratory pulsations, which reduced the dispersion enhancement factors by factor of 0.25 and 4, respectively.
+Note that behaviour directly reflects the nature of our model: The Stokes flow is linear in its inflow boundary conditions, and the diffusion enhancement factor quadratic in the predicted pressure gradient.
+Without CSF production, transport is considerably delayed
 (\Cref{fig:csf}N, O). Even after $12$ hours, tracer remains in the

### 2025-06-09 paper/supplementary.tex
revision I from Marius III

@@
 $10\,000$ in surface periarterial spaces~\cite{ray2021quantitative,
-  sharp2019dispersion}. The large variability can (at least partly) be
+  sharp2019dispersion}.
+Additionally, high-resolution MRI measurements by Hirschler et al.~\cite{hirschler2024region} revealed that CSF mobility in regions with high pulsatility is approximately tenfold greater than the self-diffusion of water.
+The large variability can (at least partly) be
 attributed to methodological differences; e.g. different assumptions
@@
 within the previously reported spectrum.
+
+Investigating the link between cardiac pulsatility and the resulting diffusion enhancement factor, we compute for $R$ different levels of cardiac pulsatility.
+We reduce the inflow boundary conditions of $6\,$ml/s (across the SAS outer surface) and $0.31\,$ml/s (across the lateral ventricle surface) by $25\,\%$, $50\,\%$, $75\,\%$ and $100\,\%$ to model lower cardiac pulsatility, reaveling that $R$ scales quadratically with the reduction of inflow (\cref{fig:R_comparison}).
+This behaviour directly reflects the nature of our model - the Stokes flow is linear in its boundary conditions, and the diffusion enhancement factor quadratic in the predicted pressure gradient - and simultaneously underscores the substantial role of pulsatility in shaping intracranial transport.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=1.0\linewidth]{paper/figures/cardiac_R_comparison.png}
+    \caption{Comparison of cardicac-driven diffusion enhancement factors for different levels of cardiac pulsatility.
+From left to right: The diffusion enhancement factor $R_c$ for $25\,\%$, $50\,\%$, $75\,\%$ and $100\,\%$ of the inflow boundary conditions of $6\,$ml/s and $0.31\,$ml/s across the SAS outer surface and the lateral ventricle surface, respectively.}
+    \label{fig:R_comparison}
+\end{figure}
 

### 2025-06-09 scripts/figure2.py
change low/high dispersion models

@@ -126,14 +126,14 @@ class Figure2_Panels(PanelsSpec):
                          ElemSize(3, hn, units), crop=(1420, 100, 1350, 0)),
             Location(xstart + 2*compwidth, y3), auto_label=True,
             content_offset=Location(0,0),
-            text=Text("low dispersion (12 h)", **dict(textargs, y=-0.1, x=1.5))
+            text=Text("low pulsatility (12 h)", **dict(textargs, y=-0.1, x=1.5))
         )
     n4: Panel = Panel(
             RasterImage("plots/modelA-HighD/modelA-HighD_overview_1-6-12-24.png",
                          ElemSize(3, hn, units), crop=(1420, 100, 1350, 0)),
             Location(xstart + 3*compwidth, y3), auto_label=True,
             content_offset=Location(0,0),
-            text=Text("high dispersion (12 h)", **dict(textargs, y=-0.1, x=1.5))
+            text=Text("high pulsatility (12 h)", **dict(textargs, y=-0.1, x=1.5))
         )
     n5: Panel = Panel(
             RasterImage("plots/modelA-OnlyDispersion/modelA-OnlyDispersion_overview_1-6-12-24.png",

### 2025-06-09 scripts/pubfig.py
remove white space

@@ -250,7 +250,7 @@ class RasterImage:
         import numpy as np
         file = Path(file)
         if crop is not None:
-            tmpfile = Path.cwd() / ".cropped" / (f" {np.random.randint(1e6)}_" + file.name)
+            tmpfile = Path.cwd() / ".cropped" / (f"{np.random.randint(1e6)}_" + file.name)
             print(tmpfile)
             makedirs(tmpfile.parent, exist_ok=True)
             with Image.open(file) as img:

### 2025-06-10 paper/main.tex
fix beta_exit
fix table 1 units

@@
       Permeability coefficient, endfoot sheath & $\xi_{\rm EF}$ & $3.8\cdot 10^{-7}$  & m/s & \cite{koch2023estimates} \\
-      Molecular efflux rate & $\beta_{\rm exit}$ & $10^{-4}$ & mm$^2$/s & \cite{hornkjol2022csf} \\
+      Molecular efflux rate & $\beta_{\rm exit}$ & $8\cdot 10^{-7}$ & m/s & \cite{hornkjol2022csf} \\
       Total amount of injected tracer & $m_{\rm tot}$ & $0.5$ & mmol & \cite{eide2024functional} \\
       Molecular influx, time-to-peak & $T_{\rm max}$ & $1$ & h & $\ast$ \\
-      CSF production rate & $u_{\mathrm{in}}$ & $0.63$  & l / day & \cite{nilsson1992circadian} \\
+      CSF production rate & $u_{\mathrm{in}}$ & $0.4$  & l / day & \cite{nilsson1992circadian} \\
       CSF viscosity & $\mu$ & $0.7$ & mPa$ \cdot $s & \cite{bloomfield1998effects} \\ 
-      CSF outflow resistance & $R_0$ & $10^{-5}$  & Pa/(mm$\cdot$s) & \cite{hornkjol2022csf} \\ 
+      CSF outflow resistance & $R_0$ & $10^{4}$  & Pa$\cdot$s/m & \cite{hornkjol2022csf} \\ 
       Dispersion factor, CSF spaces & $R$ & $\dagger$ & -- & $\ast$ \\

### 2025-06-10 paper/main.tex
remove hornkjol ref
Updates in Methods

@@
 
-For an extended description of the mathematical model and simulation
-scenarios, see the Supplementary information S1.
+This section gives a condensed summary of the in-silico modelling
+framework. For a complete description of the mathematical models and
+simulation scenarios, we refer to the Supplementary information S1.
 
@@
 
-\subsection*{CSF flow in the SAS and ventricular system}
+\subsection*{Steady flow in the CSF spaces induced by CSF production}
 
 We model CSF as an incompressible Newtonian fluid at low Reynolds and
-Womersley numbers via the Stokes equations with viscosity $\mu$, and
-account for the convective contribution of the flow induced by CSF
-production in the choroid plexus. Specifically, we compute the
-velocity $\bm u_{\rm CSF}$ and associated pressure $p_{\rm CSF}$ in
-the SAS and ventricular system induced by a constant production at a
-rate of $u_{\rm in}$ across the lateral ventricle walls and with
-efflux across the upper, outer (arachnoid) boundary with efflux
-resistance $R_0$; and then set $\bm u = \bm u_{\rm CSF}$ in
-$\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport}
-(\Cref{tab:parameters}, \Cref{fig:csf}A--E, see also
-the Supplementary information (S1)). We do not model bulk flow within the brain
-parenchyma, except within the perivascular network segments extending
-below the pial surface, and thus set $\bm u = 0$ in $\Omega_{\rm PAR}$
-in~\eqref{eq:multi_transport}.
-
-\subsection*{Dispersion in the CSF spaces}
+Womersley numbers via the Stokes equations with viscosity $\mu$ (see
+\Cref{tab:parameters} for parameter values). To account for steady
+flow induced by CSF production, we compute the CSF velocity field $\bm
+u_{\rm CSF}$ and associated pressure field $p_{\rm CSF}$ in the SAS
+and ventricular system $\Omega_{\rm CSF}$ that result from a steady
+CSF production at a rate of $u_{\rm in}$ across the lateral ventricle
+walls (\Cref{fig:csf}A--E). We allow for CSF efflux across the upper,
+outer (arachnoid) boundary with efflux resistance $R_0$.
+
+%then set $\bm u = \bm u_{\rm CSF}$ in $\Omega_{\rm CSF}$
+%in~\eqref{eq:multi_transport}
+
+%% We do not model bulk flow within the brain parenchyma, except within
+%% the perivascular network segments extending below the pial surface,
+%% and thus set $\bm u = 0$ in $\Omega_{\rm PAR}$
+%% in~\eqref{eq:multi_transport}.
+
+\subsection*{Perivascular fluid flow induced by CSF production}
+
+To estimate the contribution also to perivascular fluid flow from CSF
+production, we impose the fluid pressure $p_{\rm CSF}$ induced by CSF
+production, computed in the SAS and ventricular system, at the end
+nodes of the periarterial network. For the end nodes located within
+the SAS ($v \in \Omega_{\rm CSF}$), the value $p_{\rm CSF}(v)$ is used
+directly, while for the end nodes located within the parenchyma ($v \in
+\Omega_{\rm PAR}$), we compute and apply a harmonic extension
+$E(p_{\rm CSF})(v)$. We then numerically solve a system of hydraulic
+network equations to compute $\hat{u}^p_a$ defined over $\Lambda_a$;
+i.e., the CSF production-induced velocity field in the periarterial
+network. We refer to Supplementary information S1.1.4 \emph{Steady
+flow in perivascular networks induced by pressure differences} for
+details. We repeat this procedure to compute a corresponding velocity
+field $\hat{u}^p_v$ in the perivenous network $\Lambda_v$.
+
+\subsection*{Dispersion in the CSF spaces induced by pulsatile CSF flow}
 
@@
 while bridging from the pulsatile flow time scale of seconds to a
-molecular transport time scale of hours, we compute cardiac and
-respiratory dispersion factors $R_c$ and $R_r$, both
-spatially-varying, in $\Omega_{\rm CSF}$ (\Cref{fig:csf}A) by
-combining computational estimates of the respective peak CSF pressure
-gradients with theoretical estimates for shear-augmented (Taylor)
-dispersion \cite{taylor1953dispersion, watson1983diffusion,
-  sharp2019dispersion} (see Supplementary information S1.3 for a
-complete description). Combining these effects, we then set $R = R_c +
-R_r$ and define the effective diffusion coefficient $D = (1 + R)
-D^{\rm Gad}$ in $\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport} in
-the baseline model.
-
-\subsection*{Perivascular fluid flow induced by CSF pressure differences}
-
-To estimate the contribution also to perivascular fluid flow from CSF
-production, we impose the fluid pressure $p_{\rm CSF}$ induced by CSF
-production, computed in the SAS and ventricular system, at the end
-nodes of the periarterial network located within the SAS. For the
-network end nodes located within the parenchyma, we compute a harmonic
-extension of the CSF pressure field (by solving a Laplace equation in
-$\Omega_{\rm PAR}$ with $p_{\rm CSF}$ as the boundary value), and set
-the corresponding value at these end nodes. We then numerically solve
-a system of hydraulic network equations to compute $\hat{u}^p_a$
-defined over $\Lambda_a$; i.e., the pressure-difference induced CSF
-flow velocity in the periarterial network (see
-also Supplementary information S1.1.4). We repeat this procedure to
-compute a corresponding velocity $\hat{u}^p_v$ in the perivenous
-network $\Lambda_v$.
+molecular transport time scale of hours, we estimate cardiac- and
+respiratory dispersion factors ($R_c$ and $R_r$) in $\Omega_{\rm CSF}$
+(\Cref{fig:csf}G, I). To compute these spatially-varying factors, we
+combine computational estimates of peak CSF pressure gradients with
+theoretical estimates for shear-augmented (Taylor) dispersion
+\cite{taylor1953dispersion, watson1983diffusion, sharp2019dispersion}
+(see Supplementary information S1.3 Estimating dispersion factors from
+pulsatile CSF flow for a complete description). Combining the cardiac
+and respiratory effects, we set $R = R_c + R_r$ in $\Omega_{\rm CSF}$.
+
+%and define the effective diffusion coefficient $D = (1 + R) D^{\rm
+%Gad}$ in $\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport} in the
+%baseline model.
 
@@
       Permeability coefficient, endfoot sheath & $\xi_{\rm EF}$ & $3.8\cdot 10^{-7}$  & m/s & \cite{koch2023estimates} \\
-      Molecular efflux rate & $\beta_{\rm exit}$ & $8\cdot 10^{-7}$ & m/s & \cite{hornkjol2022csf} \\
+      Molecular efflux rate & $\beta_{\rm exit}$ & $8\cdot 10^{-7}$ & m/s & $\ast$ \\
       Total amount of injected tracer & $m_{\rm tot}$ & $0.5$ & mmol & \cite{eide2024functional} \\
@@
       CSF viscosity & $\mu$ & $0.7$ & mPa$ \cdot $s & \cite{bloomfield1998effects} \\ 
-      CSF outflow resistance & $R_0$ & $10^{4}$  & Pa$\cdot$s/m & \cite{hornkjol2022csf} \\ 
+      CSF outflow resistance & $R_0$ & $10^{4}$  & Pa$\cdot$s/m & $\ast$ \\ 
       Dispersion factor, CSF spaces & $R$ & $\dagger$ & -- & $\ast$ \\

### 2025-06-10 paper/main.tex
More work on Methods

@@
 cortical and subcortical white matter vasculature, we have considered
-a conservative transport model within the parenchyma, modelling
-extracellular diffusion alone. To compensate, we have focused on
-reporting detailed predictions in the CSF spaces and surface PVS after
-not more than 24 hours. In particular, we note that, we observe tracer
-accumulation at the perivascular network ends after 12 hours,
-similar to that previously observed for
-microspheres~\cite{bedussi2018paravascular, mestre2018flow}, which
-here can be interpreted as an artifact of the incomplete network
-creating a barrier for tracer movement. Previously, we have found
-that enrichment within the parenchyma is well-represented by $\approx
-3\times$ enhanced diffusion augmented by local clearance e.g.~across
-the blood-brain barrier or diffusion-convection with a complex pattern
-of spatially-varying convective velocities of 1-10 \textmu
+a basic transport model within the parenchyma, modelling extracellular
+diffusion alone. To compensate, we have focused on reporting detailed
+predictions in the CSF spaces and surface PVS after not more than 24
+hours. In particular, we note that, we observe tracer accumulation at
+the perivascular network ends after 12 hours, similar to that
+previously observed for microspheres~\cite{bedussi2018paravascular,
+  mestre2018flow}, which here can be interpreted as an artifact of the
+incomplete network creating a barrier for tracer movement. Previously,
+by inferring transport parameters within the human brain from
+glymphatic MRI data, we have found that molecular transport within the
+parenchyma is well-represented by $\approx 3\times$ enhanced diffusion
+augmented by local clearance e.g.~across the blood-brain barrier, or
+by diffusion-convection with a complex pattern of spatially-varying
+interstitial fluid flow on the order of 1-10 \textmu
 m/min~\cite{vinje2023human}. We will incorporate such transport
@@
 (\Cref{fig:csf}G, I). To compute these spatially-varying factors, we
-combine computational estimates of peak CSF pressure gradients with
-theoretical estimates for shear-augmented (Taylor) dispersion
-\cite{taylor1953dispersion, watson1983diffusion, sharp2019dispersion}
-(see Supplementary information S1.3 Estimating dispersion factors from
-pulsatile CSF flow for a complete description). Combining the cardiac
-and respiratory effects, we set $R = R_c + R_r$ in $\Omega_{\rm CSF}$.
+combine computational estimates of peak CSF pressure gradients
+(\Cref{fig:csf}F, H) with theoretical estimates for shear-augmented
+(Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion,
+  sharp2019dispersion} (see Supplementary information S1.3 \emph{Estimating
+dispersion factors from pulsatile CSF flow} for a complete
+description). Combining the cardiac and respiratory effects, we set $R
+= R_c + R_r$ in $\Omega_{\rm CSF}$.
 
@@
 cross-section area of the PVS segment and $\langle Q_i' \rangle$ is
-the estimated mean flow rate (see Supplementary information S1.4 for
-more details). Two wall motion patterns are considered, yielding
-$\hat{u}_a^c$ and $\hat{u}_a^v$ respectively: \emph{cardiac}
-pulsations with $f = 1.0$Hz, $\lambda = 2$ m and $\varepsilon = 1\%$
-(\Cref{fig:pvs_a}D), and very low-frequency \emph{vasomotion} with $f
-= 0.1$Hz, $\lambda = 0.02$m, and $\varepsilon = 10\%$
+the estimated mean flow rate (see Supplementary information S1.4
+\emph{Estimating net perivascular flow induced by peristaltic waves}
+for more details). Two wall motion patterns are considered:
+\emph{cardiac} pulsations with $f = 1.0$Hz, $\lambda = 2$ m and
+$\varepsilon = 1\%$ yielding $\hat{u}_a^c$ (\Cref{fig:pvs_a}D), and
+very low-frequency \emph{vasomotion} with $f = 0.1$Hz, $\lambda =
+0.02$m, and $\varepsilon = 10\%$ yielding $\hat{u}_a^v$
 (\Cref{fig:pvs_a}E). No flow induced by peristalsis is considered for
@@
 contributions from CSF production and cardiac peristalsis to set
-$\hat{u}$ in~\eqref{eq:multi_transport} as $\hat{u}_a = \hat{u}_a^p +
-\hat{u}_a^c$ in $\Lambda_a$, and $\hat{u}_v = \hat{u}_v^p$ in
-$\Lambda_v$. We remark that none of these models account for fluid
-exchange across the lateral outer interface of the PVSs (over
-$\Gamma_a, \Gamma_v$), while fluid entry and exit is allowed at the
-end nodes of the PVS networks.
+$\hat{u}$ as $\hat{u}_a = \hat{u}_a^p + \hat{u}_a^c$ in $\Lambda_a$,
+and $\hat{u}_v = \hat{u}_v^p$ in $\Lambda_v$. We remark that none of
+these models account for fluid exchange across the lateral outer
+interface of the PVSs (over $\Gamma_a, \Gamma_v$), while fluid entry
+and exit is allowed at the end nodes of the PVS networks.
 

### 2025-06-10 paper/main.tex
More updates in Methods.

@@
 dispersive effects of cardiac and respiratory pulsatile CSF flow
-(\Cref{fig:csf}A--C). Beginning with the advective, net flow flield driven by 
+(\Cref{fig:csf}A--C). Beginning with the advective, net flow field driven by 
 CSF production, we impose a constant CSF inflow of $400\,$ml/day across the surface
@@
 major surface arteries of radius 0.5--1.4 mm.
-Due to imaging limitiations, the network primarily covers the supratentorial regions, with the lowest network point located approximately $4\,$cm above the craniocervical junction.
+Due to imaging limitations, the network primarily covers the supratentorial regions, with the lowest network point located approximately $4\,$cm above the craniocervical junction.
 Imposing the pressure field induced by CSF production at the periarterial network ends induces slow steady CSF flow of variable direction in these PVSs, with
@@
 In~\eqref{eq:multi_transport_3d_CSF}, $D$ is the (free) diffusion
-coefficient of the solute, $R$ is a dispersion coefficient, $\bm
-u_{\rm CSF}$ is a CSF velocity, $\xi$ is a permeability allowing for
-transfer/exchange across the interfaces $\Gamma_a, \Gamma_v$ between
-the periarterial and perivenous networks and their surroundings, the
-overlined $\overline{c}$ denotes the average of $c$ over
-cross-sections of $\Gamma_a, \Gamma_v$, and $\delta_{\Gamma}$ is a
-Dirac term concentrated on $\Gamma_a, \Gamma_v$. Additionally, in
-\eqref{eq:multi_transport_3d_PAR}, $\phi$ is the brain extracellular
-volume fraction\footnote{We set $\phi = 1.0$ in $\Omega_{\rm CSF}$ and
-therefore omit $\phi$ in~\eqref{eq:multi_transport_3d_CSF}.},
-$D^{\ast}$ is the effective diffusion coefficient in the parenchyma,
-and $\bm u_{\rm PAR}$ is an interstitial fluid velocity. In this
-study, we set $\bm u_{\rm PAR} = 0$. In~\eqref{eq:multi_transport_1d},
-$A = \pi (R_2^2 - R_1^2)$ is the area and $P = 2 \pi R_2$ is the
-perimeter of the PVS cross-sections, $\hat{D}$ is the effective
-diffusion coefficient in the PVSs, and $\hat{u}$ is the fluid velocity
-(in the axial direction) in the PVS. Further, we model the membrane
-between the parenchyma and the CSF spaces as a semi-permeable membrane
-with permeability coefficient $\beta_{\rm pia}$. The initial
-concentrations in the three-dimensional domain $\Omega$ and networks
-$\Lambda_a, \Lambda_v$ are all set to zero.
+coefficient of the solute, $R$ is the dispersion coefficient in the
+CSF spaces, $\bm u_{\rm CSF}$ is the steady CSF velocity, $\xi$ is a
+permeability allowing for transfer/exchange across the interfaces
+$\Gamma_a, \Gamma_v$ between the periarterial and perivenous networks
+and their surroundings, the overlined $\overline{c}$ denotes the
+average of $c$ over cross-sections of $\Gamma_a, \Gamma_v$, and
+$\delta_{\Gamma}$ is a Dirac term concentrated on $\Gamma_a,
+\Gamma_v$. Additionally, in \eqref{eq:multi_transport_3d_PAR}, $\phi$
+is the brain extracellular volume fraction\footnote{We set $\phi =
+1.0$ in $\Omega_{\rm CSF}$ and therefore omit $\phi$
+in~\eqref{eq:multi_transport_3d_CSF}.}, $D^{\ast}$ is the effective
+diffusion coefficient in the parenchyma, and $\bm u_{\rm PAR}$ is an
+interstitial fluid velocity. In this study, we set $\bm u_{\rm PAR} =
+0$. In~\eqref{eq:multi_transport_1d}, $A = \pi (R_2^2 - R_1^2)$ is the
+area and $P = 2 \pi R_2$ is the perimeter of the PVS cross-sections,
+$\hat{D}$ is the effective diffusion coefficient in the PVSs, and
+$\hat{u}$ is the fluid velocity (in the axial direction) in the PVS
+defined for the periarterial and perivenous spaces separately
+($\hat{u}|_{\Lambda_a} = \hat{u}_a$, $\hat{u}|_{\Lambda_v} =
+\hat{u}_v$). Further, we model the membrane between the parenchyma and
+the CSF spaces as a semi-permeable membrane with permeability
+coefficient $\beta_{\rm pia}$, allowing for a jump in $c$ there. The
+initial concentrations in $\Omega$ and networks $\Lambda_a, \Lambda_v$
+are all set to zero.
 
@@
 and~\Cref{tab:scenarios} gives an overview of the different in-silico
-scenarios. We assume that the porosity equals the extracellular space
-volume fraction $\phi = \phi_{\rm PAR}$ within the parenchyma and
-$\phi = 1$ elsewhere. The effective diffusion coefficient is set equal
-to that of Gadubutrol in free water in the PVS networks: $\hat{D} =
-D^{\rm Gad}$; weighted by a dispersion factor $R$ in the CSF spaces:
-$D|_{\Omega_{\rm CSF}} = ( 1 + R) D^{\rm Gad}$; and modulated by the
-porosity and tortuosity ($\lambda = 1.78$) in the parenchyma:
-$D|_{\Omega_{\rm PAR}} = D^{\rm Gad}_{\rm PAR}$. We set $\xi$
-depending on whether the PVS segment is fully embedded within the CSF
-spaces ($\xi = \xi_{\rm CSF}$) or parenchyma ($\xi = \xi_{\rm EF}$),
-or neither (see also Supplementary information S1).
-To account for the unresolved continuation of the perivascular network beyond the imaging resolution limit, we implemented a permeability gradient to avoid artificial tracer buildup.
-At the terminal 'leaf nodes' of $\Lambda_a$, permeability was increased by a factor of 100 to simulate tracer exchange with the surrounding CSF/tissue, before linearly decreasing to the baseline value at the adjacent network node.
+scenarios.
+%
+%We assume that the porosity equals the extracellular space
+%volume fraction $\phi = \phi_{\rm PAR}$ within the parenchyma and
+%$\phi = 1$ elsewhere.
+%
+The effective diffusion coefficient is set equal to that of Gadubutrol
+in water in the PVS networks: $\hat{D} = D \equiv D^{\rm Gad}$; weighted by
+the dispersion factor $R$ in the CSF spaces: $(1 + R) D$;
+and modulated by the porosity and tortuosity ($\lambda = 1.78$) in the
+parenchyma $D^{\ast} = D^{\rm Gad}_{\rm PAR}$. We set
+$\xi$ depending on whether the PVS segment is fully embedded within
+the CSF spaces ($\xi = \xi_{\rm CSF}$) or parenchyma ($\xi = \xi_{\rm
+  EF}$), or neither (see also Supplementary information S1). To
+account for the unresolved continuation of the perivascular network
+beyond the imaging resolution limit, we implemented a permeability
+gradient to avoid artificial tracer buildup. At the terminal leaf
+nodes of $\Lambda_a$, permeability was increased by a factor of 100
+to simulate tracer exchange with the surrounding CSF/tissue, and
+linearly decreasing to the baseline value at the adjacent network
+node.
 \begin{table}
@@
       \midrule
-      Extracellular space volume fraction & $\phi_{\rm PAR}$ & 0.2 & -- & \cite{nicholson1981ion} \\
+      %Extracellular space volume fraction & $\phi_{\rm PAR}$ & 0.2 & -- & \cite{nicholson1981ion} \\
+      Extracellular space volume fraction & $\phi$ & 0.2 & -- & \cite{nicholson1981ion} \\
       Free diffusion coefficient, Gadubutrol & $D^{\rm Gad}$ & $3.8 \times 10^{-4}$& $\text{mm}^2/\text{s}$ & \cite{valnes2020apparent}\\

### 2025-06-11 paper/main.tex
Address minor points from Reviewer #3

@@
   miao2024brain, hauglund2025norepinephrine}. In the last decade,
-established theories have been challenged by new findings revealing a
-greater degree of molecular movement and
-exchange~\cite{iliff2012paravascular, ringstad2017glymphatic,
-  louveau2017understanding, proulx2021cerebrospinal,
-  bohr2022glymphatic}, as well as substantial variability in
-enrichment and clearance between individuals and between patient
+established theories have been challenged by new findings on molecular
+movement and exchange~\cite{iliff2012paravascular,
+  ringstad2017glymphatic, louveau2017understanding,
+  proulx2021cerebrospinal, bohr2022glymphatic}, including substantial
+variability between individuals and between patient
 cohorts~\cite{ringstad2018brain, eide2021direction, eide2021impaired,
@@
 (and challenges) for personalized medicine e.g.~for tailored
-intrathecal delivery of chemotherapies\cite{lohela2022glymphatic}
-such as methodextrate in acute lymphoblastic leukemia
+intrathecal delivery of chemotherapies\cite{lohela2022glymphatic} such
+as methodextrate in acute lymphoblastic leukemia
 patients\cite{kadan2009comparison}, and for early diagnostics of
@@
 ventricular system, cisterns, and subarachnoid space (SAS) in
-synchrony with cardiac, respiratory, and neural
+association with cardiac, respiratory, and neural
 waves~\cite{greitz1993pulsatile, wagshul2011pulsating,
@@
 analysis, predict net flow of such magnitudes and both antegrade and
-retrograde flow within the perivascular network, occurring with or
-without pressure differences at the network ends. The PVS may thus
+retrograde flow within the perivascular network. The PVS may thus
 function as highways facilitating rapid transport, even in the absence

### 2025-06-11 paper/main.tex
Almost all updates

@@
 Here we set
-\begin{equation*}
+\begin{equation}
   g_\mathrm{influx}(t) = \frac{m_{\rm tot}}{T_{\rm max}^2 |\Gamma_{\rm SSAS}|} \max(0, T_{\rm max} - |t - T_{\rm max}|), 
-\end{equation*}
-which expresses a hat-shaped influx function with a peak at $T_{\rm max}$ and a total tracer injection of $m_{\rm tot}$ (\Cref{tab:parameters}).
+  \label{eq:g}
+\end{equation}
+which expresses a hat-shaped influx function with a peak at $T_{\max}$ and a total tracer injection of $m_{\rm tot}$ (\Cref{tab:parameters}).
+We here consider a fixed $T_{\max} = 1.0$ h in~\eqref{eq:g}, and do not consider the effect of e.g.~reduced pulsatility on spinal transit time.
 

### 2025-06-11 paper/main.tex
Finally Table 2 updated.

@@
 (ii) reduced pulsatility and thus decreased dispersion, and (iii) higher pulsatility with increased dispersion.
-We computed the reduced and increased pulsatility scenarios by halving and doubling the inflow rates associated with both cardiac and respiratory pulsations, which reduced the dispersion enhancement factors by factor of 0.25 and 4, respectively.
+We computed the reduced and increased pulsatility scenarios by halving and doubling the inflow rates associated with both cardiac and respiratory pulsations, which reduced the dispersion factors by factor of 0.25 and 4, respectively.
 Note that behaviour directly reflects the nature of our model: The Stokes flow is linear in its inflow boundary conditions, and the diffusion enhancement factor quadratic in the predicted pressure gradient.
@@
 walls (\Cref{fig:csf}A--E). We allow for CSF efflux across the upper,
-outer (arachnoid) boundary with efflux resistance $R_0$.
+outer (arachnoid) boundary with efflux resistance $R_{\rm CSF, 0}$.
 
@@
 $\delta_{\Gamma}$ is a Dirac term concentrated on $\Gamma_a,
-\Gamma_v$. Additionally, in \eqref{eq:multi_transport_3d_PAR}, $\phi$
+\Gamma_v$. Additionally, in \eqref{eq:multi_transport_3d_PAR}, $\phi = \phi_{\rm ECS}$
 is the brain extracellular volume fraction\footnote{We set $\phi =
@@
 \begin{equation*}
-  (D \nabla c - \bm u c ) \cdot \bm{n} = g_{\mathrm{influx}},  \quad \mathrm{on}  \quad \Gamma_{\mathrm{SSAS}}.
+  (D \nabla c - \bm u_{\rm CSF} c ) \cdot \bm{n} = g_{\mathrm{influx}},  \quad \mathrm{on}  \quad \Gamma_{\mathrm{SSAS}}.
 \end{equation*}
@@
 The effective diffusion coefficient is set equal to that of Gadubutrol
-in water in the PVS networks: $\hat{D} = D \equiv D^{\rm Gad}$; weighted by
-the dispersion factor $R$ in the CSF spaces: $(1 + R) D$;
+in water in the PVS networks $\hat{D} = D \equiv D^{\rm Gad}$; is 
+weighted by the dispersion factor $R$ in the CSF spaces: $(1 + R) D$;
 and modulated by the porosity and tortuosity ($\lambda = 1.78$) in the
-parenchyma $D^{\ast} = D^{\rm Gad}_{\rm PAR}$. We set
-$\xi$ depending on whether the PVS segment is fully embedded within
-the CSF spaces ($\xi = \xi_{\rm CSF}$) or parenchyma ($\xi = \xi_{\rm
-  EF}$), or neither (see also Supplementary information S1). To
-account for the unresolved continuation of the perivascular network
-beyond the imaging resolution limit, we implemented a permeability
-gradient to avoid artificial tracer buildup. At the terminal leaf
-nodes of $\Lambda_a$, permeability was increased by a factor of 100
-to simulate tracer exchange with the surrounding CSF/tissue, and
-linearly decreasing to the baseline value at the adjacent network
-node.
+parenchyma $D^{\ast} = D^{\rm Gad}_{\rm PAR}$. We set $\xi$ depending
+on whether the PVS segment is fully embedded within the CSF spaces
+($\xi = \xi_{\rm PVS}$) or parenchyma ($\xi = \xi_{\rm EF}$), or
+neither (see also Supplementary information S1). To account for the
+unresolved continuation of the perivascular network beyond the imaging
+resolution limit, we implemented a permeability gradient to avoid
+artificial tracer buildup. At the terminal leaf nodes of $\Lambda_a$,
+permeability was increased by a factor of 100 to simulate tracer
+exchange with the surrounding CSF/tissue, and linearly decreasing to
+the baseline value at the adjacent network node.
 \begin{table}
@@
       %Extracellular space volume fraction & $\phi_{\rm PAR}$ & 0.2 & -- & \cite{nicholson1981ion} \\
-      Extracellular space volume fraction & $\phi$ & 0.2 & -- & \cite{nicholson1981ion} \\
-      Free diffusion coefficient, Gadubutrol & $D^{\rm Gad}$ & $3.8 \times 10^{-4}$& $\text{mm}^2/\text{s}$ & \cite{valnes2020apparent}\\
-      Parenchymal diffusion coefficient, Gadubutrol & $D^{\rm Gad}_{\rm PAR}$ & $1.2 \times 10^{-4}$ & $\text{mm}^2/\text{s}$  & \cite{hornkjol2022csf} \\
+      Extracellular space volume fraction & $\phi_{\rm ECS}$ & 0.2 & -- & \cite{nicholson1981ion} \\
+      Free diffusion coefficient, Gadubutrol & $D^{\rm Gad}$ & $3.8 \cdot 10^{-4}$& $\text{mm}^2/\text{s}$ & \cite{valnes2020apparent}\\
+      Parenchymal diffusion coefficient, Gadubutrol & $D^{\rm Gad}_{\rm PAR}$ & $1.2 \cdot 10^{-4}$ & $\text{mm}^2/\text{s}$  & \cite{hornkjol2022csf} \\
+      %Permeability coefficient, endfoot sheath & $\xi_{\rm EF}$ & $3.8\cdot 10^{-7}$  & m/s & \cite{koch2023estimates} \\
+      Permeability coefficient, endfoot sheath & $\xi_{\rm EF}$ & $3.8\cdot 10^{-7}$  & m/s & \cite{koch2023estimates} \\
+      Permeability coefficient, PVS--CSF & $\xi_{\rm PVS}$ & $3.8\cdot 10^{-7}$  & m/s & \cite{koch2023estimates} \\
       Permeability coefficient, pia & $\beta_{\rm pia}$ & $2.6 \cdot 10^{-8}$ & m/s & \cite{riseth2025twocompartment} \\
-      Permeability coefficient, endfoot sheath & $\xi_{\rm EF}$ & $3.8\cdot 10^{-7}$  & m/s & \cite{koch2023estimates} \\
-      Molecular efflux rate & $\beta_{\rm exit}$ & $8\cdot 10^{-7}$ & m/s & $\ast$ \\
+      Molecular efflux rate & $\beta_{\rm exit}$ & $8.0 \cdot 10^{-7}$ & m/s & $\ast$ \\
       Total amount of injected tracer & $m_{\rm tot}$ & $0.5$ & mmol & \cite{eide2024functional} \\
@@
       CSF viscosity & $\mu$ & $0.7$ & mPa$ \cdot $s & \cite{bloomfield1998effects} \\ 
-      CSF outflow resistance & $R_0$ & $10^{4}$  & Pa$\cdot$s/m & $\ast$ \\ 
+      CSF outflow resistance & $R_{\rm CSF, 0}$ & $10^{4}$  & Pa$\cdot$s/m & $\ast$ \\ 
       Dispersion factor, CSF spaces & $R$ & $\dagger$ & -- & $\ast$ \\
@@
   \end{center}
-  \caption{Model parameters. $\dagger$/$\ast$ denotes computed/estimated within this work.}
+  \caption{Model parameters. $\dagger$/$\ast$ denotes
+    computed/estimated within this work. The diffusion coefficients
+    $D^{\rm Gad}$ represent the diffusion coefficient of Gadubutrol in
+    CSF (water), and $D^{\rm Gad}_{\rm PAR}$ its effective diffusion
+    coefficient in human cortical tissue, both at body
+    temperature~\cite{sykova2008diffusion, valnes2020apparent,
+      hornkjol2022csf}. We use the astrocytic endfeet permeability
+    $\xi_{\rm EF}$ estimated by Koch et al.~\cite{koch2023estimates},
+    and we set $\xi_{\rm PVS} = \xi_{\rm EF}$ as a baseline.}
   \label{tab:parameters}
 \end{table}
+\newcommand{\blue}[1]{\textcolor{blue}{#1}}
 \begin{table}
   \begin{center}
-  \begin{tabular}{ll|llll|lll}
+  \begin{tabular}{ll|lll|lll}
     \toprule
-    & Description & $R_1, R_2$ & $\xi_{\rm PVS-CSF}$ & $D|_{\Omega_{\rm CSF}}$ & $D|_{\Omega_{\rm PAR}}$ & $\mathbf{u}|_{\Omega_{\rm CSF}}$ & $\hat{u}|_{\Lambda_a}$ & $\hat{u}|_{\Lambda_v}$  \\
+    No. & Model description & $\bm u_{\rm CSF}$ & $\hat{u}|_{\Lambda_a}$ & $\hat{u}|_{\Lambda_v}$ & $R$ & $\xi_{\rm PVS}$ & $R_2$ \\
     \midrule
-     & baseline & $R_2 = 2 R_1$ & $\xi$ & $(1 + R) D^{\rm Gad}$ & $D^{\rm Gad}_{\rm PAR}$ & $\mathbf{u}_{\rm CSF}$ & $\hat{u}_a^p + \hat{u}_a^c$  & $\hat{u}_v^p$ \\ 
-    & no CSF production & $-$ & $-$ & $-$ & $-$ & $\mathbf{0}$ & $\hat{u}_a^c$ & $0$ \\ 
-    & reduced pulsatility & $-$ & $-$ & $ (1 + 0.1 R) D^{\rm Gad}$ & $-$ & $-$ & $-$  & $-$ \\ 
-    & increased CSF dispersion & $-$ & $-$ & $ (1 + 10 R) D^{\rm Gad}$ & $-$ & $-$ & $-$  & $-$ \\ 
-    & higher PVS flow & $-$ & $-$ & $-$ & $-$ & $-$ & $\hat{u}_a^p + \hat{u}_a^c + \hat{u}_a^v$  & $-$ \\ 
-    & higher PVS exchange & $-$ & 100$\xi$  &  $-$ & $-$ & $-$ & $-$  & $-$ \\ 
-    & dilated PVS in the SAS & $R_2 = 3 R_1$ & $-$ & $-$ & $-$ & $-$ & $\tilde{u}_a^p + \tilde{u}_a^c$  & $-$ \\ 
+    0 & baseline & $\bm u_{\rm CSF}$ & $\hat{u}_a^p + \hat{u}_a^c$ & $\hat{u}_v^p$                & $R_c + R_r$ & $\xi_{\rm EF}$ & $2 R_1$   \\ 
+    1 & no CSF production & \blue{$\bm 0$} & \blue{$\hat{u}_a^c$} & \blue{0}                    & $R_c + R_r$ & $\xi_{\rm EF}$ & $2 R_1$   \\ 
+    2 & reduced CSF pulsatility & $\bm u_{\rm CSF}$ & $\hat{u}_a^p + \hat{u}_a^c$ & $\hat{u}_v^p$  & \blue{$ R_{\rm red}$} & $\xi_{\rm EF}$ & $2 R_1$  \\ 
+    3 & increased CSF pulsatility & $\bm u_{\rm CSF}$ & $\hat{u}_a^p + \hat{u}_a^c$ & $\hat{u}_v^p$  & \blue{$R_{\rm inc}$} & $\xi_{\rm EF}$ & $2 R_1$  \\ 
+    4 & higher PVS flow & $\bm u_{\rm CSF}$ & \blue{$\hat{u}_a^p + \hat{u}_a^c + \hat{u}_a^v$} & $\hat{u}_v^p$  & $R_c + R_r$ & $\xi_{\rm EF}$ & $2 R_1$  \\ 
+    5 & higher PVS exchange & $\bm u_{\rm CSF}$ & \blue{$\hat{u}_a^p + \hat{u}_a^c + \hat{u}_a^v$} & $\hat{u}_v^p$ & $R_c + R_r$ & \blue{100 $\xi_{\rm EF}$} & $2 R_1$   \\ 
+    6 & dilated PVS in the SAS & $\bm u_{\rm CSF}$ & \blue{$\tilde{u}_a^p + \tilde{u}_a^c + \tilde{u}_a^v$} & \blue{$\tilde{u}_v^p$} & $R_c + R_r$ & $\xi_{\rm EF}$ & \blue{$3 R_1$}\\ 
     \bottomrule
@@
     \end{center}
-  \caption{Overview of models and model variations. ($-$ denotes equal
-    to the corresponding baseline value). Each PVS cross-section is
-    modelled as a concentric annulus with inner radius $R_1$ and outer
-    radius $R_2$ and such that $R_2 = \beta R_1$. Let $\xi$ denote the
-    astrocytic endfeet permeability estimated by Koch et
-    al.~\cite{koch2023estimates}. The diffusion coefficients $D^{\rm
-      Gad}$ represent the (free) diffusion coefficient of Gadubutrol
-    in CSF (water), and $D^{\rm Gad}_{\rm PAR}$ its effective
-    diffusion coefficient in human cortical tissue, all at body
-    temperature~\cite{sykova2008diffusion, valnes2020apparent}. $R$ is
-    the combined cardiac and respiratory dispersion enhancement field
-    (see Methods). $\mathbf{u}_{\rm CSF}$ is the CSF velocity field
-    induced by CSF production in $\Omega_{\rm CSF}$, and $\hat{u}_a^p$
-    and $\hat{u}_v^p$ the axial velocity induced in the periarterial
-    and perivenous networks, respectively, by the corresponding
-    pressure differences. $\hat{u}_a^c$ and $\hat{u}_a^v$ is the
-    periarterial velocity induced by arterial pulse wave wall motion
-    and slow vasomotion, respectively. Last, $\tilde{u}_a^c$ and
-    $\tilde{u}_a^p$ are as $\hat{u}_a^c$ and $\hat{u}_a^p$,
-    respectively, but with dilated PVS. }
+  \caption{Overview of model variations, with color highlighting
+    changes from the baseline model. $\bm u_{\rm CSF}$ is the CSF
+    velocity field induced by CSF production in $\Omega_{\rm CSF}$,
+    and $\hat{u}_a^p$ and $\hat{u}_v^p$ are the axial velocities
+    induced in the periarterial and perivenous networks, respectively,
+    by the corresponding CSF pressure $p_{\rm CSF}$. Moreover,
+    $\hat{u}_a^c$ and $\hat{u}_a^v$ are the periarterial velocities
+    induced by arterial pulse wave wall motion and slow vasomotion,
+    respectively. In the model variation with no CSF production, $\bm
+    u_{\rm CSF} = 0$, $p_{\rm CSF} = 0$, and thus $\hat{u}_a^p =
+    \hat{u}_v^p = 0$. $R_c + R_r$ is the combined cardiac and
+    respiratory dispersion factor in the CSF spaces, with model
+    variations $R_{\rm red}$ and $R_{\rm inc}$ corresponding to
+    reduced pulsatility and increased pulsatility, respectively.  Each
+    PVS cross-section is modelled as a concentric annulus with inner
+    radius $R_1$ and outer radius $R_2 > R_1$. Finally,
+    $\tilde{u}_a^p$, $\tilde{u}_v^p$, $\tilde{u}_a^c$ and
+    $\tilde{u}_a^v$ are as $\hat{u}_a^p$, $\hat{u}_v^p$, $\hat{u}_a^c$
+    and $\hat{u}_a^v$, respectively, but with a dilated PVS ($R_2 = 3
+    R_1$).}
 \label{tab:scenarios}

### 2025-06-17 paper/main.tex
change R_red to to R_c^red etc
revision I by Marius III

@@
 pulsatility at smaller spatial or temporal scales.
-This approach of modeling pulsatility through an effective dispersion factor is a simplification that does not capture the full complexity of transient fluid dynamics, for instance caused by non-uniform wall motion and ependymal cilia within the ventricular system~\cite{siyahhan2014flow}.
-However, it is further supported by recent clinical observations.
-High-resolution MRI measurements by Hirschler et al.~\cite{hirschler2024region} revealed that CSF mobility in regions with high pulsatility is approximately tenfold greater than the self-diffusion of water, an enhancement factor consistent with the values computed in our model.
+This approach of modeling pulsatility through an effective dispersion 
+factor is a simplification that does not capture the full complexity
+of transient fluid dynamics, such as non-uniform wall motion and ependymal
+cilia within the ventricular system~\cite{siyahhan2014flow}, dynamic CSF
+flow at the craniocervical junction, or the effect of microanatomical
+features such as subarachnoid trabeculae on flow resitance and dispersion
+in the SAS. However, our approach is supported by recent clinical 
+observations. High-resolution MRI measurements by Hirschler et
+al.~\cite{hirschler2024region} revealed that CSF mobility in regions
+with high pulsatility is approximately tenfold greater than the 
+self-diffusion of water, an enhancement factor consistent with 
+the values computed in our model.
 
@@
     1 & no CSF production & \blue{$\bm 0$} & \blue{$\hat{u}_a^c$} & \blue{0}                    & $R_c + R_r$ & $\xi_{\rm EF}$ & $2 R_1$   \\ 
-    2 & reduced CSF pulsatility & $\bm u_{\rm CSF}$ & $\hat{u}_a^p + \hat{u}_a^c$ & $\hat{u}_v^p$  & \blue{$ R_{\rm red}$} & $\xi_{\rm EF}$ & $2 R_1$  \\ 
-    3 & increased CSF pulsatility & $\bm u_{\rm CSF}$ & $\hat{u}_a^p + \hat{u}_a^c$ & $\hat{u}_v^p$  & \blue{$R_{\rm inc}$} & $\xi_{\rm EF}$ & $2 R_1$  \\ 
+    2 & reduced CSF pulsatility & $\bm u_{\rm CSF}$ & $\hat{u}_a^p + \hat{u}_a^c$ & $\hat{u}_v^p$  & \blue{$ R_c^{\rm red} + R_r^{\rm red}$} & $\xi_{\rm EF}$ & $2 R_1$  \\ 
+    3 & increased CSF pulsatility & $\bm u_{\rm CSF}$ & $\hat{u}_a^p + \hat{u}_a^c$ & $\hat{u}_v^p$  & \blue{$ R_c^{\rm inc} + R_r^{\rm inc}$} & $\xi_{\rm EF}$ & $2 R_1$  \\ 
     4 & higher PVS flow & $\bm u_{\rm CSF}$ & \blue{$\hat{u}_a^p + \hat{u}_a^c + \hat{u}_a^v$} & $\hat{u}_v^p$  & $R_c + R_r$ & $\xi_{\rm EF}$ & $2 R_1$  \\ 
@@
     respiratory dispersion factor in the CSF spaces, with model
-    variations $R_{\rm red}$ and $R_{\rm inc}$ corresponding to
+    variations $R_{r/c}^{\rm red}$ and $R_{r/c}^{\rm inc}$ corresponding to
     reduced pulsatility and increased pulsatility, respectively.  Each
@@
 framework~\cite{alnaes2015fenics}, and the 3D-1D coupling is handled
-by the extension
-FEniCS\textsubscript{ii}~\cite{kuchta2020assembly}. The correctness
-and accuracy of the numerical solutions were verified by a series of
+by the extension FEniCS\textsubscript{ii}~\cite{kuchta2020assembly}.
+The correctness and accuracy of the numerical solutions were verified by a series of
 numerical verification experiments (see Supplementary information S1.5).
+The majority of the simulation time was spend on the Stokes flow and transport model simulations.
+Each Stokes flow simulation utilizes a hybrid parallel approach with 16 MPI processes and 4 threads each, requiring approximately 300 GB of RAM and completing in 23 minutes.
+The transport model is limited to serial execution by FEniCS\textsubscript{ii}, but employes 16 threads for the linear system factorization, taking 22 minutes and consuming about 50 GB of memory on a AMD EPYC 9684X machine.
 

### 2025-06-17 paper/supplementary.tex
Fix link
revision I by Marius III

@@
 These sections provide more detailed descriptions of the mathematical models and numerical approximations considered.
-
-\subsection{Mixed-dimensional transport and flow equations}
+We solve the models presented in Methods numerically
+using finite element methods. The following subsections provide more
+details, model-by-model.
+
+\subsection{Nomenclature and notation} We begin by including \Cref{table:geometry_notation}, \Cref{table:variable}, and \Cref{table:mesh}, which summarize the common notation used for the geometry, model variables and parameters, and finite element discretization, respectively.
+\begin{table}[h!] 
+\centering
+\caption{Geometrical notation, refer to (Fig.~1A, Fig.~2A--C)}
+\label{table:geometry_notation}
+\begin{tabular}{ll}
+\hline
+\textbf{Symbol} & \textbf{Description} \\
+\hline
+$\Omega$ & Full intracranial domain: $\Omega = \Omega_{\mathrm{PAR}} \cup \Omega_{\mathrm{CSF}}$ \\
+$\Omega_{\mathrm{PAR}}$ & Brain parenchyma domain \\
+$\Omega_{\mathrm{CSF}}$ & Cerebrospinal‐fluid (CSF) domain \\
+$\partial\Omega_{\mathrm{PAR}}\cap\partial\Omega_{\mathrm{CSF}}$ & Parenchyma–CSF interface: $\partial \Omega_{\rm CSF} \bigcap \partial
+\Omega_{\rm PAR} = \Gamma_{\rm LV} \cup \Gamma_{\rm pia}$ \\
+$\Gamma_{\mathrm{LV}}$ & Lateral‐ventricle surface (part of parenchyma–CSF interface) \\
+$\Gamma_{\mathrm{pia}}$ & Pial surface (remainder of parenchyma–CSF interface) \\
+$\Gamma_{\mathrm{SSAS}}$ & Lower CSF boundary toward spinal subarachnoid space \\
+$\Gamma_{\mathrm{AM}}$ & Outer CSF boundary toward arachnoid/dura \\
+$\Gamma_{\mathrm{AM\text{-}L}}$, $\Gamma_{\mathrm{AM\text{-}U}}$ & Lower/upper parts of $\Gamma_{\mathrm{AM}}$ \\
+$\Gamma_{\mathrm{SC}}$ & Parenchyma–spinal‐cord boundary: $\partial\Omega_{\mathrm{PAR}}\setminus\partial\Omega_{\mathrm{CSF}}$ \\
+$\Lambda_{a},\Lambda_{v}$ & Arterial/venous PVS centerline networks \\
+$\Lambda_{i}$ & Individual PVS centerline segment ($\Lambda_a^i$, $\Lambda_v^i$ corresponding to arterial and venous segments) \\
+$\Omega_{i}$ & Perivascular space around $\Lambda_{i}$ \\
+$R_{1}^{i},\,R_{2}^{i}$ & Inner/outer radii of the annular PVS $\Omega_{i}$ \\
+$\Gamma_{a},\,\Gamma_{v}$ & Outer lateral surfaces of periarterial/perivenous PVSs \\
+$P$ & Perimeter of the PVS cross‐section \\
+$\mathcal V$ & Graph nodes of the PVS network \\
+$\mathcal B\subset  \mathcal V$ & Internal (bifurcation) nodes \\
+$\mathcal N\subset  \mathcal V$ & End (leaf) nodes \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[htb!]
+\centering 
+\caption{Variable and parameter notation}
+\label{table:variable}
+\begin{tabular}{ll}
+\hline
+\textbf{Symbol} & \textbf{Description} \\
+\hline
+$\phi$ & Fluid volume fraction (porosity) \\
+$D$ & Effective diffusion coefficient ($D_{\mathrm{PAR}} = D^*$ or $D_{\mathrm{CSF}} = D(1+R)$ (see eq. (1)) \\
+$\mathbf{u}$ & Convective velocity ($\mathbf{u}_{\mathrm{PAR}} = 0$ or $\mathbf{u}_{\mathrm{CSF}}$ computed from steady state flow) \\
+$c$ & Solute concentration in 3D domains \\
+$\bar c$ & Lateral average of $c$ on the PVS boundary \\
+$\xi$ & Permeability coefficient between the periarterial and perivenous networks
+and their surroundings\\
+$\delta_{\Gamma}$ & Dirac term on the PVS outer surfaces \\
+$\beta_{\mathrm{pia}}$ & Permeability of pial/ventricle interface \\
+$\beta_{\mathrm{exit}}$ & Efflux rate on $\Gamma_{\mathrm{AM\text{-}U}}$ \\
+$g_{\mathrm{influx}}$ & Influx on $\Gamma_{\mathrm{SSAS}}$ \\
+$A(s)$ & Cross‐sectional area of PVS at $s$ \\
+$\hat D$ & 1D effective diffusion in PVS \\
+$\hat u$ & 1D convective velocity in PVS \\
+$p$ & Pressure in CSF (Stokes) or PVS (network) \\
+$q$ & Volumetric flux in PVS network \\
+$\kappa$ & Hydraulic conductivity in annular PVS \\
+$\mu$ & Dynamic viscosity of CSF \\
+$R_{0}$ & Resistance parameter for CSF efflux \\
+$u_{\mathrm{in}}$ & Imposed fluid influx on $\Gamma_{\mathrm{LV}}$ \\
+$\omega,\;\rho$ & Angular frequency, CSF density (in dispersion estimates) \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[h!]
+\centering
+\caption{Common mesh and finite‐element notation}
+\label{table:mesh}
+\begin{tabular}{ll}
+\hline
+\textbf{Symbol} & \textbf{Description} \\
+\hline
+$\mathcal T$ & Tetrahedral mesh of $\Omega$ \\
+$\mathcal{T}_{\mathrm{PAR}},\,T_{\mathrm{CSF}}$ & Restrictions of $\mathcal T$ to $\Omega_{\mathrm{PAR}},\Omega_{\mathrm{CSF}}$ \\
+$E$ & Tetrahedral cell in $\mathcal T$ \\
+$\mathcal F_{i}$ & Interior facets of $\mathcal T$ \\
+$\mathcal F_{i,\mathrm{PAR}},\, \mathcal F_{i,\mathrm{CSF}}$ & Interior facets in $\Omega_{\mathrm{PAR}}$, $\Omega_{\mathrm{CSF}}$ \\
+$\mathcal F_{e}$ & Exterior facets of CSF mesh ($\mathcal F_{\mathrm{pia}}\cup \mathcal F_{\mathrm{LV}}\cup \mathcal F_{\mathrm{AM\text{-}L}}\cup \mathcal \mathcal F_{\mathrm{AM\text{-}U}}\cup \mathcal F_{\mathrm{SSAS}}$) \\
+$\mathcal F _{\mathrm{pia}},\,\mathcal F_{\mathrm{LV}},\, \mathcal F_{\mathrm{AM\text{-}L}},\,\mathcal F_{\mathrm{AM\text{-}U}},\, \mathcal F_{\mathrm{SSAS}}$ & Facets on corresponding boundaries \\
+$V_{h}$ & Discontinuous Galerkin space on $\Omega$ of degree $1$\\
+$\widehat V_{h}$ & Continuous Galerkin space on $\Lambda$ of degree $1$.  \\
+% $\widehat V_{h,0}$, $\widehat V_{h,g}$ & Continuous Galerkin space on network with homogeneous or prescribed BCs \\
+$\mathit{BDM}_{2}$ & Brezzi–Douglas–Marini space of degree 2 \\
+$Q_{h}$ & Pressure space on $\mathcal T_{\mathrm{CSF}}$  \\
+$I_{a},I_{v}$ & 1D meshes of arterial/venous networks $\Lambda_{a},\Lambda_{v}$ \\
+$\widehat V_{h}(I)$ & Space of piecewise constants ($P^0$) on network mesh $I$ \\
+\hline
+\end{tabular}
+\end{table}
+
+\subsection{Flow and mixed-dimensional models: governing equations and corresponding finite element methods }
 \label{sec:app:more_maths}
 
-\subsubsection{Notation}
+\subsubsection{Notation pertaining to the geometry and variables}
 
@@
 
-\subsubsection{Solute transport and exchange in intracranial domains (3D) and perivascular networks (1D)}
-\label{sec:app:transport_eqs}
-
-In this section, we describe the mathematical model (eq.~(1)), the
-associated definitions, interface and boundary conditions in further
-detail. We refer to the main text for explicit parameter values while
-giving the general, abstract form here, and to the
-reference~\cite{masri2024modelling, laurino2019derivation} for the
-derivation and analysis of this 3D-1D model. Recall that we model a
-concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
-defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
-each of these domains, $c$ satisfies
-%\mer{MER: Rami asks if it should be a $\phi$ in the advective term here.
-We can include it, or drop it, but in the latter case, perhaps we should clarify that u is an average/superficial/Darcy velocity.}
-\begin{equation}
-  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma
-  %= f
-   = 0
-  \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
-  \label{eq:3d_pde}
-\end{equation}
-In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$).
-In the parenchyma, $\phi$ represents the extracellular space volume fraction, and thus $c$ here generally represents the intrinsic (in contrast to the superficial) concentration~\cite{sykova2008diffusion}.
-Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$.
-$\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivascular networks $\Lambda_a, \Lambda_v$.
-To summarize,
-\begin{equation*}
-\phi =  \begin{cases}
-  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
-  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
-  \end{cases}, \; 
-D = \begin{cases}
-  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
-  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
-  \quad \bm u  = \begin{cases}
-  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
-  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
-\end{cases}, \; 
-\xi = \begin{cases}
-  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
-  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
-\end{cases} . 
-\end{equation*}
-In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp.
-$|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp.
-$\Omega_{\rm{CSF}}$).
-If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.
-
-Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by
-\begin{equation*}
-  \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta_2^i(s)} c 
-\end{equation*}
-where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$.
-Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the term $\delta_\Gamma$ is concentrated on the outer lateral surfaces of the PVSs, and defined in terms of its action on any sufficiently smooth function $v : \Omega \rightarrow \R$
-%and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$
-such that
-\begin{equation*}
-  \langle \xi (\overline{c} - \hat c ) \delta_\Gamma, v \rangle =
-  \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
-\end{equation*}
-under the assumption that $\xi$ is constant on each $\partial
-\Theta_2^i$.
-%% , since:
-%% \begin{multline}
-%%   \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
-%%   \int_{\Gamma} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
-%%   \sum_i \int_{\Gamma^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = 
-%%   \sum_i \int_{\Lambda^i} \int_{\partial \Theta_2^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \\
-%%   \sum_i \int_{\Lambda^i} \left (  \xi (\overline{c} - \hat c ) \int_{\partial \Theta_2^i} \delta_\Gamma (v) \right ) =
-%%   \sum_i \int_{\Lambda^i} P \xi (\overline{c} - \hat c ) \overline{v} =
-%%   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v} .
-%% \end{multline}
-%% \mer{MER: The previous was:
-%% \begin{equation}
-%%   \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
-%% \end{equation}
-%% }
-
-On the interface between the brain and the CSF spaces
-$\Gamma_{\mathrm{pia}} \cup \Gamma_{\rm LV}$, we prescribe the
-following interface conditions, which represent a semi--permeable
-interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and
-$c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
-%% \begin{subequations}
-%% \begin{alignat}{2}
-%% (-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} &= - (- D (\phi \nabla c_{\rm{CSF}}) + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n} &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\
-%% (-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,
-%% \end{alignat} 
-%% \end{subequations}
-\begin{subequations}
-  \begin{alignat}{2}
-  (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n}
-  &= - (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm CSF}} \cdot \bm{n}
-  &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
-  (- D \nabla (\phi  c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
-\end{alignat} 
-  \label{eq:3d_interface}%
-\end{subequations}%
-where $\bm{n} = \bm{n}_{\rm PAR}$ is the normal vector field defined
-over the interface, outward-pointing when viewed from $\Omega_{\rm
-  PAR}$, and $\beta_{\mathrm{pia}} \geqslant 0$ is a permeability
-constant.
-
-We supplement~\eqref{eq:3d_pde} and~\eqref{eq:3d_interface} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
-\begin{subequations}
-\begin{alignat}{2}
-(D \nabla (\phi c) - \bm u c ) \cdot \bm{n} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \Gamma_{\rm SSAS},  \\ 
-% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\Gamma_{\rm SAS}} &= 0 &&  \quad  \mathrm{on} \quad \Gamma_{\rm SAS}^{\mathrm{out}}  \\ 
-(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = \beta_{\rm exit} c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-U},  \\ 
-(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = 0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-L} \cup \Gamma_{\rm SC}  .
-\end{alignat} 
-\label{eq:3d_bcs}
-\end{subequations}
-where again $\bm n$ denotes the outward-pointing boundary normal.
-
-The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
-perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
-in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
-space averaged over each perivascular cross-section and is governed by 
-\begin{equation}
-  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s \hat c  ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
-  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
-\label{eq:1d_pde}
-\end{equation}
-where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
-of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
-PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
-associated PVS $\Omega^i$, the area of the cross-section
-$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
-and $\hat{u}$ is a convective velocity representing an (average) CSF
-flow velocity in the axial direction of the perivascular spaces.
-
-To complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
-internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
-boundary conditions at the end nodes. To this end, define the set of
-internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
-that are connected to two or more edges and the set of end nodes as
-$\mathcal{N} \subset \mathcal{V}$, write $\hat{c}^i =
-\hat{c}|_{\Lambda^i}$, and for $y\in\mathcal{B}$ let $\mathcal{E}(y)$ denote the set of
-(two or three) edges in $\{\Lambda^i\}$ sharing the node $y$. To
-impose continuity of the concentrations, we set that the
-concentrations when viewed from each edge must match at nodes:
-\begin{equation}
-  \hat{c}^i (y) = \hat{c}^j(y)
-  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
-  \qquad \foralls y \in \mathcal{B} .
-  \label{eq:conc_continuity}
-\end{equation}
-Moreover, to ensure mass conservation, we set that the flux
-going in and out at nodes should add to zero, or more precisely, that
-for each $y \in \mathcal{B}$:
-\begin{equation}
-  \sum_{\Lambda^i \in \mathcal{E}(y)}
-  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n^i(y)
-  = 0.
-  \label{eq:flux_conservation}
-\end{equation}
-Here the (normal ``vector") function $n^i$ takes the values in $\left\{-1, 1\right\}$ and 
-defines an orientation of vertices of edge $\Lambda^i$. Specifically, for an edge $\Lambda^i$ oriented from $y_{\rm in}$ to $y_{\rm out}$ we set 
-\begin{equation}
-  n^i(y_{\rm in}) = 1,\quad
-  n^i(y_{\rm{out}}) = -1. %\quad
-  %\mk{n^i(y) = 0 \quad \foralls y \in 
-  %\mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} }.  
-\label{eq:def:sign}
-\end{equation}
-%\mk{Why do we need the normal on non-IO vertices?}
-
-Finally, we impose a homogeneous Neumann condition at all network end
-nodes to augment~\eqref{eq:1d_pde}:
-\begin{equation}
-    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{N}.
-    \label{eq:1d_bcs}
-\end{equation}
-This no-flux condition states that the network end nodes represent
-barriers for molecular efflux into or out of the perivascular network,
-and thus that all solute exchange between the PVSs and their
-surroundings takes place via the lateral outer PVS surface and is
-regulated by the exchange parameter $\xi$ cf.~\eqref{eq:3d_pde}
-and~\eqref{eq:1d_pde}. Larger particles have been observed to
-accumulate within the PVS as the surface arteries penetrate into the
-brain parenchyma~\cite{mestre2018flow, bedussi2018paravascular},
-and~\eqref{eq:1d_bcs} is appropriate to represent such
-behavior. However, to represent a continuously extending PVSs also
-along penetrating vessels, this condition should be revisited.
-
-\subsubsection{Modelling CSF flow via the incompressible Stokes equations}
+\subsubsection{Computational mesh entities: definitions and common notation}
+
+We consider a mesh $\mathcal{T} = \{ E \}$ of $\Omega = \Omega_{\rm
+  PAR} \cup \Omega_{\rm CSF}$, consisting of tetrahedral mesh cells
+$E$, and conforming to the domains $\Omega_{\rm PAR}$ and $\Omega_{\rm
+  CSF}$ and to the CSF-brain interface $\Gamma_{\rm pia} \cup
+\Gamma_{\rm LV}$. We denote the restriction of $\mathcal{T}$ to
+$\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ by $\mathcal{T}_{\rm PAR}$
+and $\mathcal{T}_{\rm CSF}$, respectively. The collection of all
+interior facets (i.e.~triangular faces of the tetrahedral mesh cells)
+in $\mathcal{T}_{\rm PAR}$ and $\mathcal{T}_{\rm CSF}$ are denoted by
+$\mathcal{F}_{i, \rm PAR}$ and $\mathcal{F}_{i, \rm CSF}$,
+respectively. We define the union of facets interior both to
+$\mathcal{T}_{ \rm PAR}$ and $\mathcal{T}_{ \rm CSF}$ as
+$\mathcal{F}_i = \mathcal{F}_{i, \rm PAR} \cup \mathcal{F}_{i, \rm
+  CSF}$.
+
+\subsubsection{Modelling and simulating steady state CSF flow via the incompressible Stokes equations}
 \label{sec:app:csf_flow}
@@
   \begin{alignat}{2}
-    ( \mu \nabla \bm u \cdot \bm{n} - p \bm n) \cdot \bm n  &= -R_0 ( \bm u \cdot \bm n ), 
+    ( \mu \nabla \bm u \cdot \bm{n} - p \bm n) \cdot \bm n  &= -R_{\rm CSF, 0} ( \bm u \cdot \bm n ), 
     \quad \bm u \cdot \bm t = 0
@@
 \end{subequations}%
-where $\bm n$ and $\bm t$ denote the unit outward normal and tangent vectors to the boundary respectively,  $\mu$ is the (dynamic) CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$.
+where $\bm n$ and $\bm t$ denote the unit outward normal and tangent vectors to the boundary respectively,  $\mu$ is the (dynamic) CSF viscosity, $R_{\rm CSF, 0} > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$.
 We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:app:dispersion}.
 
-%% We consider non-zero convection in the CSF space $u_{\rm CSF} \not = 0$ in \eqref{eq:multi_transport_3d} resulting from steady state CSF production.
-Recall that $\Omega_{\mathrm{CSF}}$ represents the CSF space and that $\Gamma_{\mathrm{skull^{u/l}}}$ and $\Gamma_{\mathrm{pia}}$, $\Gamma_{\mathrm{LV}}$ represent the boundaries facing the upper and lower parts of the dura, the pia and the lateral ventricles, respectively.
-Assuming CSF production in the choroid plexus located at the lateral ventricle wall, we solve the  steady state Stokes equations for the velocity and the pressure $(\bm u_{\mathrm{CSF}}, p_{\mathrm{CSF}})$ in the CSF domain:
-
-\subsubsection{Steady flow in perivascular networks induced by pressure differences}
-\label{sec:app:hydraulic_network}
-
-To model steady flow induced by pressure differences between end nodes
-in a perivascular network $\Lambda$, with edges $\{\Lambda^i\}_i$,
-internal nodes $\mathcal{B}$ and end nodes $\mathcal{N}$, we consider
-the following system of hydraulic network
-equations~\cite{tithof2022network, daversin2022geometrically,
-  gjerde2024directional}. The unknowns are the PVS flux $\hat{q} :
-\Lambda \rightarrow \R$ and pressure $\hat{p} : \Lambda
-\rightarrow \R$, which represent the fluid flux and the average
-pressure across cross-sections of the PVSs, respectively. These are
-defined for each $\Lambda^i$ by solving:
-\begin{subequations}
-\begin{alignat}{2}
-  \hat{q} + \frac{A \kappa}{\mu} \partial_{s} \hat{p} & = 0 &&
-  \quad \text{ in } \Lambda^i , \\ 
-  - \partial_s \hat{q} & = 0 &&
-  \quad \text{ in } \Lambda^i ,  
-\end{alignat} 
-\label{eq:1d_flow}%
-\end{subequations}%
-where $A$ is the area of the PVS cross-sections, $\mu$ is the CSF viscosity as before, and $\kappa$ is derived from an assumption of Poiseuille flow in the annular cross-section of the PVS as~\cite{tithof2022network}:
-\begin{equation}
-\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
-\end{equation}
-The PVS flux $\hat{q}$ and the associated average PVS velocity in the
-axial direction $\hat{u}$ are directly related by the cross-sectional area
-\begin{equation}
-  \hat{q}^i = A^i \hat{u}^i \qquad \foralls \Lambda^i.
-\end{equation}
-
-In addition, to complete~\eqref{eq:1d_flow}, we impose continuity of
-the fluid pressure and conservation of flux at each internal node. Write
-$\hat{p}^i = \hat{p}|_{\Lambda^i}$. These conditions then read as: for
-each $y \in \mathcal{B}$ with connected edges $\mathcal{E}(y)$:
-\begin{subequations}
-\begin{align}
-  \hat{p}^i(y) &= \hat{p}^j(y) \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y), \\
-  \sum_{\Lambda^i \in \mathcal{E}(y)} \hat{q}^i n^i(y) &= 0  .
-\end{align}
-\label{eq:1d_flow:bifurcations}
-\end{subequations}
-where $n^i$ is as defined by~\eqref{eq:def:sign}. Finally, to drive this flow, we impose given fluid pressures $p_0$ at the
-network end nodes:
-\begin{equation}
-  \hat{p}(y) = p_0(y) \qquad y \in \mathcal{N} .
-  \label{eq:1d_flow:bcs}
-\end{equation}
-In the case where modelling flow in the perivascular networks
-$\Lambda_a, \Lambda_v$ induced by CSF production, we impose the fluid
-pressure $p_{\rm CSF}$ computed in the CSF spaces $\Omega_{\rm CSF}$
-(and its harmonic extensions into $\Omega_{\rm PAR}$) as the given
-fluid pressures $p_0$. Note that we here consider steady PVS flow in a
-non-moving domain; the effect of domain motion is addressed
-in~\Cref{sec:app:peristalsis}.
-
-\subsection{Numerical solution of the transport and flow equations}
-\label{sec:app:details_numerical_method}
-
-We solve the models presented in~\Cref{sec:app:more_maths} numerically
-using finite element methods. The following subsections provide more
-details, model-by-model.
-
-\subsubsection{Computational mesh entities: definitions and common notation}
-
-We consider a mesh $\mathcal{T} = \{ E \}$ of $\Omega = \Omega_{\rm
-  PAR} \cup \Omega_{\rm CSF}$, consisting of tetrahedral mesh cells
-$E$, and conforming to the domains $\Omega_{\rm PAR}$ and $\Omega_{\rm
-  CSF}$ and to the CSF-brain interface $\Gamma_{\rm pia} \cup
-\Gamma_{\rm LV}$. We denote the restriction of $\mathcal{T}$ to
-$\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ by $\mathcal{T}_{\rm PAR}$
-and $\mathcal{T}_{\rm CSF}$, respectively. The collection of all
-interior facets (i.e.~triangular faces of the tetrahedral mesh cells)
-in $\mathcal{T}_{\rm PAR}$ and $\mathcal{T}_{\rm CSF}$ are denoted by
-$\mathcal{F}_{i, \rm PAR}$ and $\mathcal{F}_{i, \rm CSF}$,
-respectively. We define the union of facets interior both to
-$\mathcal{T}_{ \rm PAR}$ and $\mathcal{T}_{ \rm CSF}$ as
-$\mathcal{F}_i = \mathcal{F}_{i, \rm PAR} \cup \mathcal{F}_{i, \rm
-  CSF}$.
-
-%Note in particular that $\mathcal{F}_i \cap \mathcal{F}_{pia}} =
-%  \emptyset$.
-
-%The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is
-%denoted by $\mathcal{E}_{h}$.
-
-\subsubsection{Finite element solution of coupled 3D-1D solute transport equations }
-
-We consider the system of coupled 3D-1D solute transport equations
-given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
-conditions~\eqref{eq:3d_interface}, the bifurcation
-conditions~\eqref{eq:conc_continuity}
-and~\eqref{eq:flux_conservation}, and the boundary
-conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
-these equations using an implicit finite difference scheme in time, 
-a discontinuous Galerkin (DG) finite element method with upwinding in space for the 3D domain to accurately capture sharp boundary layers, and a continuous Galerkin method for the 1D networks.
-We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm, and accounting for the increased diffusivity due to dispersion).
-
-We first consider the discretization of~\eqref{eq:3d_pde} and
-introduce the discrete space
-\begin{equation}
-  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
-  V_h = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^1(E), \; E \in \mathcal{T}\} ,
-\end{equation}
-where $L^2(\Omega)$ is the space of square-integrable functions
-defined over $\Omega$ and $\mathrm{P}^1(E)$ denotes the space of
-polynomials of total degree $\leqslant 1$ defined over the tetrahedra
-$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
-symmetric weighted interior penalty DG formulation, referring
-to~\cite{ern2009discontinuous} and \cite[Section
-  4.5.2.3]{di2011mathematical} for details on this
-method. Recall that $\phi$ is constant in each domain
-  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
-  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.
-Define for $c, v \in V_h^k$:
-\begin{equation}
-  a_h (c, v)
-  =
-  \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
-  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
-  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right) 
-   + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
-  .
-  \label{eq:dg_sip}
-\end{equation}
-In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
-between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
-pointing from $E^1$ to $E^2$. The facet diameter is denoted by $h_F$,
-the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
-unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
-are defined as:
-\begin{equation}
-  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
-  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
-  \quad \text{where} \quad \kappa_i = (D \phi) |_{E^i}.  
-\end{equation}
-The parameter $\gamma_{D,F}$ is the harmonic mean of the
-porosity-weighted diffusion coefficient given by
-\begin{equation}
-  \gamma_{D,F} = \frac{2\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
-\end{equation}
-and $\eta$ is a user-specified penalty parameter (we set $\eta=1000$). To discretize the convection term in~\eqref{eq:3d_pde}, we use
-upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
-references therein. For $c, v \in V_h^k$, define
-\begin{equation}
-  a_h^{\mathrm{up}}(c, v)
-  = - \sum_{E \in \mathcal{T}} \int_E c\bm{u} \cdot \nabla v
-  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
-  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
-\end{equation}
-Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h$ such that for all $v \in V_h$:
-\begin{equation} \label{eq:3d_dg_disc}
-  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
-  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
-  + \int_{\Gamma_{\rm pia} \cup \Gamma_{\rm LV}} \beta_{\rm pia}[c_h^n][v] 
-  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
-  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
-  = 
-  %\int_{\Omega} f^n v +
-  \int_{\Gamma_{\rm SSAS}} g_{\rm influx} v
-  %, \quad \forall v \in V_h^k
-  .
-\end{equation}
-
-To discretize~\eqref{eq:1d_pde} with the bifurcation
-conditions~\eqref{eq:conc_continuity}
-and~\eqref{eq:flux_conservation}, and the boundary
-conditions~\eqref{eq:1d_bcs}, we use the space of continuous piecewise
-linear polynomials defined over $\Lambda$:
-\begin{equation}
-  \hat V_h
-  = \{ v \in C^0(\Lambda), \; v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda^i) \}, 
-\end{equation}
-where $\mathrm{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$.
-The discrete formulation then reads: for $n = 1,2,\ldots, $ find $\hat c_h^n \in \hat V_h$ such that for all $\hat v \in \hat V_h^k$:
-\begin{align}\label{eq:1d_dg_disc}
-\int_{\Lambda}  \frac{1}{\tau}(A  \hat c_h^n - A \hat c_h^{n-1}) \hat v  + \int_{\Lambda} \hat D A \partial_s \hat c_h^n  \partial_s \hat v^n - \int_{\Lambda} A \hat u \hat c_h^n \partial_s \hat v   +  \int_{\Lambda} \xi P (\hat c_h^n - \overline{c_h^n}) \hat v +a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) = 0.
-\end{align}
- In the above, if $\hat u$ is non--zero on $\Lambda^i$, then the artificial diffusion $a_h^{\mathrm{stab}}$ stabilization term is nonzero  and is given below, for more details see for  \cite[Section 12.6]{quarteroni2009numerical}.
- \begin{align}
-    a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) =  \frac{1}{\|\hat u\|_{L^2(\Lambda^i)}} 
- h_{\Lambda^i} \int_{\Lambda^i} A (\hat u \partial_s \hat c_h  ) (\hat u \partial_s \hat v),  
- \end{align}
- where $h_{\Lambda^i}$ is the mesh--size of $\Lambda^i$.
-Note that the condition \eqref{eq:flux_conservation} is enforced weakly in the above formulation.
-
- \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} complemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0 = 0 $ is the following:
-
- For $n = 1,2, \ldots $, find  $c_h^n \in V_h$ and $\hat c^n_h \in \hat V_h$ such that the coupled equations \eqref{eq:3d_dg_disc} and \eqref{eq:1d_dg_disc} hold for all $v \in V_h$ and for all $\hat v \in \hat V_h$.
-\subsubsection{Finite element solution of the incompressible Stokes equations}
-
+\textit{Finite element solution of the incompressible Stokes equations}.
 We consider an H(div)-based finite element approximation of the incompressible Stokes equations~\eqref{eq:stokes} defined over $\Omega_{\rm CSF}$ with the boundary conditions~\eqref{eq:stokes_bcs}.
@@
   \mathcal{A}_h(\bm u_h, \bm v_h)
-  + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
+  + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_{\rm CSF, 0} (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
   - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0 \qquad \foralls \bm v_h \in \bm V_{h,0} \\ 
@@
 
-\subsubsection{Finite element solution of the perivascular network equations}
-
+\subsubsection{Modeling and simulating steady flow in perivascular networks induced by pressure differences}
+\label{sec:app:hydraulic_network}
+
+To model steady flow induced by pressure differences between end nodes
+in a perivascular network $\Lambda$, with edges $\{\Lambda^i\}_i$,
+internal nodes $\mathcal{B}$ and end nodes $\mathcal{N}$, we consider
+the following system of hydraulic network
+equations~\cite{tithof2022network, daversin2022geometrically,
+  gjerde2024directional}. The unknowns are the PVS flux $\hat{q} :
+\Lambda \rightarrow \R$ and pressure $\hat{p} : \Lambda
+\rightarrow \R$, which represent the fluid flux and the average
+pressure across cross-sections of the PVSs, respectively. These are
+defined for each $\Lambda^i$ by solving:
+\begin{subequations}
+\begin{alignat}{2}
+  \hat{q} + \frac{A \kappa}{\mu} \partial_{s} \hat{p} & = 0 &&
+  \quad \text{ in } \Lambda^i , \\ 
+  - \partial_s \hat{q} & = 0 &&
+  \quad \text{ in } \Lambda^i ,  
+\end{alignat} 
+\label{eq:1d_flow}%
+\end{subequations}%
+where $A$ is the area of the PVS cross-sections, $\mu$ is the CSF viscosity as before, and $\kappa$ is derived from an assumption of Poiseuille flow in the annular cross-section of the PVS as~\cite{tithof2022network}:
+\begin{equation}
+\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
+\end{equation}
+The PVS flux $\hat{q}$ and the associated average PVS velocity in the
+axial direction $\hat{u}$ are directly related by the cross-sectional area
+\begin{equation}
+  \hat{q}^i = A^i \hat{u}^i \qquad \foralls \Lambda^i.
+\end{equation}
+
+In addition, to complete~\eqref{eq:1d_flow}, we impose continuity of
+the fluid pressure and conservation of flux at each internal node. Write
+$\hat{p}^i = \hat{p}|_{\Lambda^i}$. These conditions then read as: for
+each $y \in \mathcal{B}$ with connected edges $\mathcal{E}(y)$:
+\begin{subequations}
+\begin{align}
+  \hat{p}^i(y) &= \hat{p}^j(y) \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y), \\
+  \sum_{\Lambda^i \in \mathcal{E}(y)} \hat{q}^i n^i(y) &= 0  .
+\end{align}
+\label{eq:1d_flow:bifurcations}
+\end{subequations}
+Here the (normal ``vector") function $n^i$ takes the values in $\left\{-1, 1\right\}$ and 
+defines an orientation of vertices of edge $\Lambda^i$. Specifically, for an edge $\Lambda^i$ oriented from $y_{\rm in}$ to $y_{\rm out}$ we set 
+\begin{equation}
+  n^i(y_{\rm in}) = 1,\quad
+  n^i(y_{\rm{out}}) = -1. %\quad
+  %\mk{n^i(y) = 0 \quad \foralls y \in 
+  %\mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} }.  
+\label{eq:def:sign}
+\end{equation}
+
+Finally, to drive this flow, we impose given fluid pressures $p_0$ at the
+network end nodes:
+\begin{equation}
+  \hat{p}(y) = p_0(y) \qquad y \in \mathcal{N} .
+  \label{eq:1d_flow:bcs}
+\end{equation}
+When modelling flow in the perivascular networks
+$\Lambda_a, \Lambda_v$ induced by CSF production, we impose the fluid
+pressure $p_{\rm CSF}$ computed in the CSF spaces $\Omega_{\rm CSF}$
+(and its harmonic extensions into $\Omega_{\rm PAR}$) as the given
+fluid pressures $p_0$. Note that we here consider steady PVS flow in a
+non-moving domain; the effect of domain motion is addressed
+in~\Cref{sec:app:peristalsis}.
+
+\textit{Finite element solution of the perivascular network equations}. 
 We consider meshes $\mathcal{I}_a, \mathcal{I}_v$ representing a
@@
 \end{subequations}%
-where $\langle \cdot, \cdot \rangle$ denotes the $L^2(\mathcal{I})$-inner product and defined segment-wise.
+where $\langle \cdot, \cdot \rangle$ denotes the $L^2(\mathcal{I})$-inner product and is defined segment-wise.
 For a stability and convergence analysis of this discrete model, we refer to the reference\cite{gjerde2024directional}.
+
+\subsubsection{Modeling and simulating solute transport and exchange in intracranial domains (3D) and perivascular networks (1D)}
+\label{sec:app:transport_eqs}
+
+In this section, we describe the mathematical model (eq.~(1)), the
+associated definitions, interface and boundary conditions in further
+detail. We refer to the main text for explicit parameter values while
+giving the general, abstract form here, and to the
+reference~\cite{masri2024modelling, laurino2019derivation} for the
+derivation and analysis of this 3D-1D model. Recall that we model a
+concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
+defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
+each of these domains, $c$ satisfies
+%\mer{MER: Rami asks if it should be a $\phi$ in the advective term here.
+We can include it, or drop it, but in the latter case, perhaps we should clarify that u is an average/superficial/Darcy velocity.}
+\begin{equation}
+  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma
+  %= f
+   = 0
+  \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
+  \label{eq:3d_pde}
+\end{equation}
+In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$).
+In the parenchyma, $\phi$ represents the extracellular space volume fraction, and thus $c$ here generally represents the intrinsic (in contrast to the superficial) concentration~\cite{sykova2008diffusion}.
+Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$.
+$\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivascular networks $\Lambda_a, \Lambda_v$.
+To summarize,
+\begin{equation*}
+\phi =  \begin{cases}
+  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
+  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
+  \end{cases}, \; 
+D = \begin{cases}
+  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
+  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
+  \quad \bm u  = \begin{cases}
+  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
+  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
+\end{cases}, \; 
+\xi = \begin{cases}
+  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
+  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
+\end{cases} . 
+\end{equation*}
+In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp.
+$|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp.
+$\Omega_{\rm{CSF}}$).
+If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.
+
+Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by
+\begin{equation*}
+  \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta_2^i(s)} c 
+\end{equation*}
+where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$.
+Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the term $\delta_\Gamma$ is concentrated on the outer lateral surfaces of the PVSs, and defined in terms of its action on any sufficiently smooth function $v : \Omega \rightarrow \R$
+%and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$
+such that
+\begin{equation*}
+  \langle \xi (\overline{c} - \hat c ) \delta_\Gamma, v \rangle =
+  \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
+\end{equation*}
+under the assumption that $\xi$ is constant on each $\partial
+\Theta_2^i$.
+%% , since:
+%% \begin{multline}
+%%   \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
+%%   \int_{\Gamma} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
+%%   \sum_i \int_{\Gamma^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = 
+%%   \sum_i \int_{\Lambda^i} \int_{\partial \Theta_2^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \\
+%%   \sum_i \int_{\Lambda^i} \left (  \xi (\overline{c} - \hat c ) \int_{\partial \Theta_2^i} \delta_\Gamma (v) \right ) =
+%%   \sum_i \int_{\Lambda^i} P \xi (\overline{c} - \hat c ) \overline{v} =
+%%   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v} .
+%% \end{multline}
+%% \mer{MER: The previous was:
+%% \begin{equation}
+%%   \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
+%% \end{equation}
+%% }
+
+On the interface between the brain and the CSF spaces
+$\Gamma_{\mathrm{pia}} \cup \Gamma_{\rm LV}$, we prescribe the
+following interface conditions, which represent a semi--permeable
+interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and
+$c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
+%% \begin{subequations}
+%% \begin{alignat}{2}
+%% (-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} &= - (- D (\phi \nabla c_{\rm{CSF}}) + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n} &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\
+%% (-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,
+%% \end{alignat} 
+%% \end{subequations}
+\begin{subequations}
+  \begin{alignat}{2}
+  (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n}
+  &= - (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm CSF}} \cdot \bm{n}
+  &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
+  (- D \nabla (\phi  c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
+\end{alignat} 
+  \label{eq:3d_interface}%
+\end{subequations}%
+where $\bm{n} = \bm{n}_{\rm PAR}$ is the normal vector field defined
+over the interface, outward-pointing when viewed from $\Omega_{\rm
+  PAR}$, and $\beta_{\mathrm{pia}} \geqslant 0$ is a permeability
+constant.
+
+We supplement~\eqref{eq:3d_pde} and~\eqref{eq:3d_interface} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
+\begin{subequations}
+\begin{alignat}{2}
+(D \nabla (\phi c) - \bm u c ) \cdot \bm{n} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \Gamma_{\rm SSAS},  \\ 
+% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\Gamma_{\rm SAS}} &= 0 &&  \quad  \mathrm{on} \quad \Gamma_{\rm SAS}^{\mathrm{out}}  \\ 
+(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = \beta_{\rm exit} c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-U},  \\ 
+(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = 0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-L} \cup \Gamma_{\rm SC}  .
+\end{alignat} 
+\label{eq:3d_bcs}
+\end{subequations}
+where again $\bm n$ denotes the outward-pointing boundary normal.
+
+The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
+perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
+in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
+space averaged over each perivascular cross-section and is governed by 
+\begin{equation}
+  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s \hat c  ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
+  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
+\label{eq:1d_pde}
+\end{equation}
+where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
+of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
+PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
+associated PVS $\Omega^i$, the area of the cross-section
+$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
+and $\hat{u}$ is a convective velocity representing an (average) CSF
+flow velocity in the axial direction of the perivascular spaces.
+
+To complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
+internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
+boundary conditions at the end nodes. To this end, define the set of
+internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
+that are connected to two or more edges and the set of end nodes as
+$\mathcal{N} \subset \mathcal{V}$, write $\hat{c}^i =
+\hat{c}|_{\Lambda^i}$, and for $y\in\mathcal{B}$ let $\mathcal{E}(y)$ denote the set of
+(two or three) edges in $\{\Lambda^i\}$ sharing the node $y$. To
+impose continuity of the concentrations, we set that the
+concentrations when viewed from each edge must match at nodes:
+\begin{equation}
+  \hat{c}^i (y) = \hat{c}^j(y)
+  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
+  \qquad \foralls y \in \mathcal{B} .
+  \label{eq:conc_continuity}
+\end{equation}
+Moreover, to ensure mass conservation, we set that the flux
+going in and out at nodes should add to zero, or more precisely, that
+for each $y \in \mathcal{B}$:
+\begin{equation}
+  \sum_{\Lambda^i \in \mathcal{E}(y)}
+  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n^i(y)
+  = 0.
+  \label{eq:flux_conservation}
+\end{equation}
+Here the normal $n^i$ is as defined in ~\eqref{eq:def:sign}.
+%Here the (normal ``vector") function $n^i$ takes the values %in $\left\{-1, 1\right\}$ and 
+%defines an orientation of vertices of edge $\Lambda^i$. %Specifically, for an edge $\Lambda^i$ oriented from $y_{\rm %in}$ to $y_{\rm out}$ we set 
+%\begin{equation}
+%  n^i(y_{\rm in}) = 1,\quad
+%  n^i(y_{\rm{out}}) = -1. %\quad
+%  %\mk{n^i(y) = 0 \quad \foralls y \in 
+%  %\mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} }.  
+%\label{eq:def:sign}
+%\end{equation}
+%\mk{Why do we need the normal on non-IO vertices?}
+
+Finally, we impose a homogeneous Neumann condition at all network end
+nodes to augment~\eqref{eq:1d_pde}:
+\begin{equation}
+    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{N}.
+    \label{eq:1d_bcs}
+\end{equation}
+This no-flux condition states that the network end nodes represent
+barriers for molecular efflux into or out of the perivascular network,
+and thus that all solute exchange between the PVSs and their
+surroundings takes place via the lateral outer PVS surface and is
+regulated by the exchange parameter $\xi$ cf.~\eqref{eq:3d_pde}
+and~\eqref{eq:1d_pde}. Larger particles have been observed to
+accumulate within the PVS as the surface arteries penetrate into the
+brain parenchyma~\cite{mestre2018flow, bedussi2018paravascular},
+and~\eqref{eq:1d_bcs} is appropriate to represent such
+behavior. However, to represent a continuously extending PVSs also
+along penetrating vessels, this condition should be revisited.
+
+%% We consider non-zero convection in the CSF space $u_{\rm CSF} \not = 0$ in \eqref{eq:multi_transport_3d} resulting from steady state CSF production.
+Recall that $\Omega_{\mathrm{CSF}}$ represents the CSF space and that $\Gamma_{\mathrm{skull^{u/l}}}$ and $\Gamma_{\mathrm{pia}}$, $\Gamma_{\mathrm{LV}}$ represent the boundaries facing the upper and lower parts of the dura, the pia and the lateral ventricles, respectively.
+Assuming CSF production in the choroid plexus located at the lateral ventricle wall, we solve the  steady state Stokes equations for the velocity and the pressure $(\bm u_{\mathrm{CSF}}, p_{\mathrm{CSF}})$ in the CSF domain:
+
+%\subsection{Numerical solution of the transport and flow equations}
+%\label{sec:app:details_numerical_method}
+%Note in particular that $\mathcal{F}_i \cap \mathcal{F}_{pia}} =
+%  \emptyset$.
+
+%The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is
+%denoted by $\mathcal{E}_{h}$.
+
+\textit{Finite element solution of coupled 3D-1D solute transport equations}. 
+We consider the system of coupled 3D-1D solute transport equations
+given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
+conditions~\eqref{eq:3d_interface}, the bifurcation
+conditions~\eqref{eq:conc_continuity}
+and~\eqref{eq:flux_conservation}, and the boundary
+conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
+these equations using an implicit finite difference scheme in time, 
+a discontinuous Galerkin (DG) finite element method with upwinding in space for the 3D domain to accurately capture sharp boundary layers, and a continuous Galerkin method for the 1D networks.
+We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm, and accounting for the increased diffusivity due to dispersion).
+
+We first consider the discretization of~\eqref{eq:3d_pde} and
+introduce the discrete space
+\begin{equation}
+  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
+  V_h = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^1(E), \; E \in \mathcal{T}\} ,
+\end{equation}
+where $L^2(\Omega)$ is the space of square-integrable functions
+defined over $\Omega$ and $\mathrm{P}^1(E)$ denotes the space of
+polynomials of total degree $\leqslant 1$ defined over the tetrahedra
+$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
+symmetric weighted interior penalty DG formulation, referring
+to~\cite{ern2009discontinuous} and \cite[Section
+  4.5.2.3]{di2011mathematical} for details on this
+method. Recall that $\phi$ is constant in each domain
+  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
+  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.
+Define for $c, v \in V_h^k$:
+\begin{equation}
+  a_h (c, v)
+  =
+  \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
+  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
+  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right) 
+   + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
+  .
+  \label{eq:dg_sip}
+\end{equation}
+Here, for each facet $F \in \mathcal{F}^i$ shared
+between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
+pointing from $E^1$ to $E^2$. The facet diameter is denoted by $h_F$,
+the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
+unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
+are defined as:
+\begin{equation}
+  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
+  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
+  \quad \text{where} \quad \kappa_i = (D \phi) |_{E^i}.  
+\end{equation}
+In~\eqref{eq:dg_sip} the parameter $\gamma_{D,F}$ is the harmonic mean of the
+porosity-weighted diffusion coefficient given by
+\begin{equation}
+  \gamma_{D,F} = \frac{2\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
+\end{equation}
+and $\eta$ is a user-specified penalty parameter (we set $\eta=1000$). To discretize the convection term in~\eqref{eq:3d_pde}, we use
+upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
+references therein. For $c, v \in V_h^k$, define
+\begin{equation}
+  a_h^{\mathrm{up}}(c, v)
+  = - \sum_{E \in \mathcal{T}} \int_E c\bm{u} \cdot \nabla v
+  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
+  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
+\end{equation}
+Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h$ such that for all $v \in V_h$:
+\begin{equation} \label{eq:3d_dg_disc}
+  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
+  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
+  + \int_{\Gamma_{\rm pia} \cup \Gamma_{\rm LV}} \beta_{\rm pia}[c_h^n][v] 
+  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
+  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
+  = 
+  %\int_{\Omega} f^n v +
+  \int_{\Gamma_{\rm SSAS}} g_{\rm influx} v
+  %, \quad \forall v \in V_h^k
+  .
+\end{equation}
+
+To discretize~\eqref{eq:1d_pde} with the bifurcation
+conditions~\eqref{eq:conc_continuity}
+and~\eqref{eq:flux_conservation}, and the boundary
+conditions~\eqref{eq:1d_bcs}, we use the space of continuous piecewise
+linear polynomials defined over $\Lambda$:
+\begin{equation}
+  \hat V_h
+  = \{ v \in C^0(\Lambda), \; v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda^i) \}, 
+\end{equation}
+where $\mathrm{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$.
+The discrete formulation then reads: for $n = 1,2,\ldots, $ find $\hat c_h^n \in \hat V_h$ such that for all $\hat v \in \hat V_h^k$:
+\begin{align}\label{eq:1d_dg_disc}
+\int_{\Lambda}  \frac{1}{\tau}(A  \hat c_h^n - A \hat c_h^{n-1}) \hat v  + \int_{\Lambda} \hat D A \partial_s \hat c_h^n  \partial_s \hat v^n - \int_{\Lambda} A \hat u \hat c_h^n \partial_s \hat v   +  \int_{\Lambda} \xi P (\hat c_h^n - \overline{c_h^n}) \hat v +a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) = 0.
+\end{align}
+ In the above, if $\hat u$ is non--zero on $\Lambda^i$, then the artificial diffusion $a_h^{\mathrm{stab}}$ stabilization term is nonzero  and is given below, for more details see for  \cite[Section 12.6]{quarteroni2009numerical}.
+ \begin{align}
+    a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) =  \frac{1}{\|\hat u\|_{L^2(\Lambda^i)}} 
+ h_{\Lambda^i} \int_{\Lambda^i} A (\hat u \partial_s \hat c_h  ) (\hat u \partial_s \hat v),  
+ \end{align}
+ where $h_{\Lambda^i}$ is the mesh--size of $\Lambda^i$.
+Note that the condition \eqref{eq:flux_conservation} is enforced weakly in the above formulation.
+
+ \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} complemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0 = 0 $ is the following:
+
+ For $n = 1,2, \ldots $, find  $c_h^n \in V_h$ and $\hat c^n_h \in \hat V_h$ such that the coupled equations \eqref{eq:3d_dg_disc} and \eqref{eq:1d_dg_disc} hold for all $v \in V_h$ and for all $\hat v \in \hat V_h$.
 
@@
     \centering
-    \includegraphics[width=1.0\linewidth]{paper/figures/cardiac_R_comparison.png}
+    \includegraphics[width=1.0\linewidth]{figures/cardiac_R_comparison.png}
     \caption{Comparison of cardicac-driven diffusion enhancement factors for different levels of cardiac pulsatility.

### 2025-06-20 paper/main.tex
Updates from Overleaf

@@
      B) 3D rendering of the computational geometry: the parenchyma, CSF (clipped, blue), arterial network (red), and venous network (blue); 
-     C) illustration of model representation of the intrathecal tracer injection; 
+     C) illustration of the intrathecal tracer injection protocol;  
      D) in-silico predictions of tracer concentration after 1, 6, 12, and 24 hours (opacity increasing linearly from 0 to 2 mmol/l); 
@@
 enrichment pattern and a concentration peak in the SAS of $\approx$0.5
-mmol/l (relative to a total amount of 1.0 mmol) occuring after 10 to 15 hours. Our in-silico enrichment
+mmol/l (relative to a total amount of 1.0 mmol) occuring after 10 to 15 hours. 
+Our in-silico enrichment
 patterns agree with these observations, though with an earlier rather than later peak in the CSF spaces (2 hours after first tracer appearance at the craniocervical junction).
@@
 Our simulation results highlight that reproducing clinically observed tracer timings requires flow drivers beyond CSF production and cardiac peristalsis.
-To illustrate, we compared our findings to tracer arrival times reported by Eide and Ringstad~\cite{eide2024functional}, who observed a delay of just 15.3 min between the M1 and M2 segments of the MCA.
+Similar to the CSF spaces, dispersion due to cardiac and respiratory pulsation has been hypothesized to enhance transport in the perivascular spaces.
+However, recent estimates indicate that the effect of dispersion does not exceed a factor of 2 compared to diffusion alone \cite{asgari2016glymphatic,sharp2019dispersion,bojarskaite2023sleep,asgari2016glymphatic,troyetsky2021dispersion}.
+This conclusion is supported by our own model, where a 100-fold increase in the PVS dispersion coefficient did not substantially accelerate transport (\cref{fig:pvs_b}H).
+To illustrate the need for a stronger advective driver, we compared our findings to tracer arrival times reported by Eide and Ringstad~\cite{eide2024functional}, who observed a delay of just 15.3 min between the M1 and M2 segments of the MCA.
 Our baseline model, accounting only for CSF production and cardiac-driven flow, predicted a physiologically unrealistic delay of over three hours.
@@
 concentration appearing at the craniocervical junction rather than as
-an intrathecal injection. We therefore overestimate the amount of
+an intrathecal injection. We thus neglect absorption of both CSF and tracer in the spinal compartment, and therefore overestimate the amount of
 tracer entering into the intracranial spaces; however, due to the

### 2025-06-20 paper/supplementary.tex
Updates from Overleaf

@@
 $\Lambda_{i}$ & Individual PVS centerline segment ($\Lambda_a^i$, $\Lambda_v^i$ corresponding to arterial and venous segments) \\
+$L_i$ & Length of a PVS segment $\Lambda_i$ \\
 $\Omega_{i}$ & Perivascular space around $\Lambda_{i}$ \\
@@
 $\phi$ & Fluid volume fraction (porosity) \\
-$D$ & Effective diffusion coefficient ($D_{\mathrm{PAR}} = D^*$ or $D_{\mathrm{CSF}} = D(1+R)$ (see eq. (1)) \\
+$D$ & Effective diffusion coefficient ($D_{\mathrm{PAR}} = D^*$ or $D_{\mathrm{CSF}} = D(1+R)$ \\
 $\mathbf{u}$ & Convective velocity ($\mathbf{u}_{\mathrm{PAR}} = 0$ or $\mathbf{u}_{\mathrm{CSF}}$ computed from steady state flow) \\
@@
 $p$ & Pressure in CSF (Stokes) or PVS (network) \\
-$q$ & Volumetric flux in PVS network \\
+$\hat{q}$ & Volumetric flux in PVS network \\
 $\kappa$ & Hydraulic conductivity in annular PVS \\
 $\mu$ & Dynamic viscosity of CSF \\
-$R_{0}$ & Resistance parameter for CSF efflux \\
+$R_{\rm CSF, 0}$ & Resistance parameter for CSF efflux \\
 $u_{\mathrm{in}}$ & Imposed fluid influx on $\Gamma_{\mathrm{LV}}$ \\
-$\omega,\;\rho$ & Angular frequency, CSF density (in dispersion estimates) \\
+$\omega,\;\rho$ & Angular frequency, CSF density (for dispersion estimates) \\
+$\alpha_c,\, \alpha_r$ & Womersley number (cardiac, respiratory) \\
+$h_{\rm SAS}$ & Mean width of the subarachnoid space (for dispersion estimates) \\
+$R_c,\,R_r$ & Spatially-varying dispersion enhancement factors (cardiac, respiratory) \\
+
 \hline
@@
 \hline
-$\mathcal T$ & Tetrahedral mesh of $\Omega$ \\
-$\mathcal{T}_{\mathrm{PAR}},\,T_{\mathrm{CSF}}$ & Restrictions of $\mathcal T$ to $\Omega_{\mathrm{PAR}},\Omega_{\mathrm{CSF}}$ \\
+$\mathcal{T}$ & Tetrahedral mesh of $\Omega$ \\
+$\mathcal{T}_{\mathrm{PAR}},\,\mathcal{T}_{\mathrm{CSF}}$ & Restrictions of $\mathcal T$ to $\Omega_{\mathrm{PAR}},\Omega_{\mathrm{CSF}}$ \\
 $E$ & Tetrahedral cell in $\mathcal T$ \\
@@
 % $\widehat V_{h,0}$, $\widehat V_{h,g}$ & Continuous Galerkin space on network with homogeneous or prescribed BCs \\
-$\mathit{BDM}_{2}$ & Brezzi–Douglas–Marini space of degree 2 \\
+$\mathrm{BDM}_{2}$ & Brezzi–Douglas–Marini space of degree 2 \\
 $Q_{h}$ & Pressure space on $\mathcal T_{\mathrm{CSF}}$  \\
-$I_{a},I_{v}$ & 1D meshes of arterial/venous networks $\Lambda_{a},\Lambda_{v}$ \\
-$\widehat V_{h}(I)$ & Space of piecewise constants ($P^0$) on network mesh $I$ \\
+$\mathcal{I}_a, \mathcal{I}_v$ & 1D meshes of arterial/venous networks $\Lambda_{a},\Lambda_{v}$ \\
+$\widehat{Z}_{h}(\mathcal{I})$ & Space of piecewise constants ($P^0$) on network mesh $I$ \\
+$\tau$ & Time step size \\
 \hline
@@
 
-\subsection{Flow and mixed-dimensional models: governing equations and corresponding finite element methods }
-\label{sec:app:more_maths}
-
-\subsubsection{Notation pertaining to the geometry and variables}
+\subsection{Notation pertaining to the geometry, variables and computational domains}
 
@@
 
-\subsubsection{Computational mesh entities: definitions and common notation}
+% \subsection{Computational mesh entities: definitions and common notation}
 
@@
 
-\subsubsection{Modelling and simulating steady state CSF flow via the incompressible Stokes equations}
+%\subsection{Flow and mixed-dimensional models: governing equations and corresponding finite element methods }
+%\label{sec:app:more_maths} 
+\subsection{Modelling and simulating steady state CSF flow via the incompressible Stokes equations}
 \label{sec:app:csf_flow}
@@
 
-\subsubsection{Modeling and simulating steady flow in perivascular networks induced by pressure differences}
+\subsection{Modeling and simulating steady flow in perivascular networks induced by pressure differences}
 \label{sec:app:hydraulic_network}
@@
 $\Lambda_a$, $\Lambda_v$. Relative to each mesh $\mathcal{I}$, we define the
-space of (discontinuous) piecewise constants $\hat{Z}_h(\mathcal{I})$ and define
+space of (discontinuous) piecewise constants $\widehat{Z}_h(\mathcal{I})$ and define
 spaces of continuous piecewise linears $\hat{V}_{h, g}(\mathcal{I})$ with
@@
 For a stability and convergence analysis of this discrete model, we refer to the reference\cite{gjerde2024directional}.
-
-\subsubsection{Modeling and simulating solute transport and exchange in intracranial domains (3D) and perivascular networks (1D)}
-\label{sec:app:transport_eqs}
-
-In this section, we describe the mathematical model (eq.~(1)), the
-associated definitions, interface and boundary conditions in further
-detail. We refer to the main text for explicit parameter values while
-giving the general, abstract form here, and to the
-reference~\cite{masri2024modelling, laurino2019derivation} for the
-derivation and analysis of this 3D-1D model. Recall that we model a
-concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
-defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
-each of these domains, $c$ satisfies
-%\mer{MER: Rami asks if it should be a $\phi$ in the advective term here.
-We can include it, or drop it, but in the latter case, perhaps we should clarify that u is an average/superficial/Darcy velocity.}
-\begin{equation}
-  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma
-  %= f
-   = 0
-  \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
-  \label{eq:3d_pde}
-\end{equation}
-In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$).
-In the parenchyma, $\phi$ represents the extracellular space volume fraction, and thus $c$ here generally represents the intrinsic (in contrast to the superficial) concentration~\cite{sykova2008diffusion}.
-Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$.
-$\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivascular networks $\Lambda_a, \Lambda_v$.
-To summarize,
-\begin{equation*}
-\phi =  \begin{cases}
-  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
-  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
-  \end{cases}, \; 
-D = \begin{cases}
-  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
-  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
-  \quad \bm u  = \begin{cases}
-  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
-  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
-\end{cases}, \; 
-\xi = \begin{cases}
-  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
-  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
-\end{cases} . 
-\end{equation*}
-In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp.
-$|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp.
-$\Omega_{\rm{CSF}}$).
-If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.
-
-Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by
-\begin{equation*}
-  \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta_2^i(s)} c 
-\end{equation*}
-where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$.
-Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the term $\delta_\Gamma$ is concentrated on the outer lateral surfaces of the PVSs, and defined in terms of its action on any sufficiently smooth function $v : \Omega \rightarrow \R$
-%and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$
-such that
-\begin{equation*}
-  \langle \xi (\overline{c} - \hat c ) \delta_\Gamma, v \rangle =
-  \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
-\end{equation*}
-under the assumption that $\xi$ is constant on each $\partial
-\Theta_2^i$.
-%% , since:
-%% \begin{multline}
-%%   \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
-%%   \int_{\Gamma} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
-%%   \sum_i \int_{\Gamma^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = 
-%%   \sum_i \int_{\Lambda^i} \int_{\partial \Theta_2^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \\
-%%   \sum_i \int_{\Lambda^i} \left (  \xi (\overline{c} - \hat c ) \int_{\partial \Theta_2^i} \delta_\Gamma (v) \right ) =
-%%   \sum_i \int_{\Lambda^i} P \xi (\overline{c} - \hat c ) \overline{v} =
-%%   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v} .
-%% \end{multline}
-%% \mer{MER: The previous was:
-%% \begin{equation}
-%%   \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
-%% \end{equation}
-%% }
-
-On the interface between the brain and the CSF spaces
-$\Gamma_{\mathrm{pia}} \cup \Gamma_{\rm LV}$, we prescribe the
-following interface conditions, which represent a semi--permeable
-interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and
-$c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
-%% \begin{subequations}
-%% \begin{alignat}{2}
-%% (-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} &= - (- D (\phi \nabla c_{\rm{CSF}}) + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n} &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\
-%% (-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,
-%% \end{alignat} 
-%% \end{subequations}
-\begin{subequations}
-  \begin{alignat}{2}
-  (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n}
-  &= - (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm CSF}} \cdot \bm{n}
-  &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
-  (- D \nabla (\phi  c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
-\end{alignat} 
-  \label{eq:3d_interface}%
-\end{subequations}%
-where $\bm{n} = \bm{n}_{\rm PAR}$ is the normal vector field defined
-over the interface, outward-pointing when viewed from $\Omega_{\rm
-  PAR}$, and $\beta_{\mathrm{pia}} \geqslant 0$ is a permeability
-constant.
-
-We supplement~\eqref{eq:3d_pde} and~\eqref{eq:3d_interface} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
-\begin{subequations}
-\begin{alignat}{2}
-(D \nabla (\phi c) - \bm u c ) \cdot \bm{n} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \Gamma_{\rm SSAS},  \\ 
-% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\Gamma_{\rm SAS}} &= 0 &&  \quad  \mathrm{on} \quad \Gamma_{\rm SAS}^{\mathrm{out}}  \\ 
-(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = \beta_{\rm exit} c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-U},  \\ 
-(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = 0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-L} \cup \Gamma_{\rm SC}  .
-\end{alignat} 
-\label{eq:3d_bcs}
-\end{subequations}
-where again $\bm n$ denotes the outward-pointing boundary normal.
-
-The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
-perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
-in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
-space averaged over each perivascular cross-section and is governed by 
-\begin{equation}
-  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s \hat c  ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
-  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
-\label{eq:1d_pde}
-\end{equation}
-where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
-of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
-PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
-associated PVS $\Omega^i$, the area of the cross-section
-$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
-and $\hat{u}$ is a convective velocity representing an (average) CSF
-flow velocity in the axial direction of the perivascular spaces.
-
-To complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
-internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
-boundary conditions at the end nodes. To this end, define the set of
-internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
-that are connected to two or more edges and the set of end nodes as
-$\mathcal{N} \subset \mathcal{V}$, write $\hat{c}^i =
-\hat{c}|_{\Lambda^i}$, and for $y\in\mathcal{B}$ let $\mathcal{E}(y)$ denote the set of
-(two or three) edges in $\{\Lambda^i\}$ sharing the node $y$. To
-impose continuity of the concentrations, we set that the
-concentrations when viewed from each edge must match at nodes:
-\begin{equation}
-  \hat{c}^i (y) = \hat{c}^j(y)
-  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
-  \qquad \foralls y \in \mathcal{B} .
-  \label{eq:conc_continuity}
-\end{equation}
-Moreover, to ensure mass conservation, we set that the flux
-going in and out at nodes should add to zero, or more precisely, that
-for each $y \in \mathcal{B}$:
-\begin{equation}
-  \sum_{\Lambda^i \in \mathcal{E}(y)}
-  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n^i(y)
-  = 0.
-  \label{eq:flux_conservation}
-\end{equation}
-Here the normal $n^i$ is as defined in ~\eqref{eq:def:sign}.
-%Here the (normal ``vector") function $n^i$ takes the values %in $\left\{-1, 1\right\}$ and 
-%defines an orientation of vertices of edge $\Lambda^i$. %Specifically, for an edge $\Lambda^i$ oriented from $y_{\rm %in}$ to $y_{\rm out}$ we set 
-%\begin{equation}
-%  n^i(y_{\rm in}) = 1,\quad
-%  n^i(y_{\rm{out}}) = -1. %\quad
-%  %\mk{n^i(y) = 0 \quad \foralls y \in 
-%  %\mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} }.  
-%\label{eq:def:sign}
-%\end{equation}
-%\mk{Why do we need the normal on non-IO vertices?}
-
-Finally, we impose a homogeneous Neumann condition at all network end
-nodes to augment~\eqref{eq:1d_pde}:
-\begin{equation}
-    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{N}.
-    \label{eq:1d_bcs}
-\end{equation}
-This no-flux condition states that the network end nodes represent
-barriers for molecular efflux into or out of the perivascular network,
-and thus that all solute exchange between the PVSs and their
-surroundings takes place via the lateral outer PVS surface and is
-regulated by the exchange parameter $\xi$ cf.~\eqref{eq:3d_pde}
-and~\eqref{eq:1d_pde}. Larger particles have been observed to
-accumulate within the PVS as the surface arteries penetrate into the
-brain parenchyma~\cite{mestre2018flow, bedussi2018paravascular},
-and~\eqref{eq:1d_bcs} is appropriate to represent such
-behavior. However, to represent a continuously extending PVSs also
-along penetrating vessels, this condition should be revisited.
-
-%% We consider non-zero convection in the CSF space $u_{\rm CSF} \not = 0$ in \eqref{eq:multi_transport_3d} resulting from steady state CSF production.
-Recall that $\Omega_{\mathrm{CSF}}$ represents the CSF space and that $\Gamma_{\mathrm{skull^{u/l}}}$ and $\Gamma_{\mathrm{pia}}$, $\Gamma_{\mathrm{LV}}$ represent the boundaries facing the upper and lower parts of the dura, the pia and the lateral ventricles, respectively.
-Assuming CSF production in the choroid plexus located at the lateral ventricle wall, we solve the  steady state Stokes equations for the velocity and the pressure $(\bm u_{\mathrm{CSF}}, p_{\mathrm{CSF}})$ in the CSF domain:
-
-%\subsection{Numerical solution of the transport and flow equations}
-%\label{sec:app:details_numerical_method}
-%Note in particular that $\mathcal{F}_i \cap \mathcal{F}_{pia}} =
-%  \emptyset$.
-
-%The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is
-%denoted by $\mathcal{E}_{h}$.
-
-\textit{Finite element solution of coupled 3D-1D solute transport equations}. 
-We consider the system of coupled 3D-1D solute transport equations
-given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
-conditions~\eqref{eq:3d_interface}, the bifurcation
-conditions~\eqref{eq:conc_continuity}
-and~\eqref{eq:flux_conservation}, and the boundary
-conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
-these equations using an implicit finite difference scheme in time, 
-a discontinuous Galerkin (DG) finite element method with upwinding in space for the 3D domain to accurately capture sharp boundary layers, and a continuous Galerkin method for the 1D networks.
-We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm, and accounting for the increased diffusivity due to dispersion).
-
-We first consider the discretization of~\eqref{eq:3d_pde} and
-introduce the discrete space
-\begin{equation}
-  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
-  V_h = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^1(E), \; E \in \mathcal{T}\} ,
-\end{equation}
-where $L^2(\Omega)$ is the space of square-integrable functions
-defined over $\Omega$ and $\mathrm{P}^1(E)$ denotes the space of
-polynomials of total degree $\leqslant 1$ defined over the tetrahedra
-$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
-symmetric weighted interior penalty DG formulation, referring
-to~\cite{ern2009discontinuous} and \cite[Section
-  4.5.2.3]{di2011mathematical} for details on this
-method. Recall that $\phi$ is constant in each domain
-  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
-  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.
-Define for $c, v \in V_h^k$:
-\begin{equation}
-  a_h (c, v)
-  =
-  \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
-  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
-  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right) 
-   + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
-  .
-  \label{eq:dg_sip}
-\end{equation}
-Here, for each facet $F \in \mathcal{F}^i$ shared
-between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
-pointing from $E^1$ to $E^2$. The facet diameter is denoted by $h_F$,
-the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
-unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
-are defined as:
-\begin{equation}
-  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
-  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
-  \quad \text{where} \quad \kappa_i = (D \phi) |_{E^i}.  
-\end{equation}
-In~\eqref{eq:dg_sip} the parameter $\gamma_{D,F}$ is the harmonic mean of the
-porosity-weighted diffusion coefficient given by
-\begin{equation}
-  \gamma_{D,F} = \frac{2\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
-\end{equation}
-and $\eta$ is a user-specified penalty parameter (we set $\eta=1000$). To discretize the convection term in~\eqref{eq:3d_pde}, we use
-upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
-references therein. For $c, v \in V_h^k$, define
-\begin{equation}
-  a_h^{\mathrm{up}}(c, v)
-  = - \sum_{E \in \mathcal{T}} \int_E c\bm{u} \cdot \nabla v
-  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
-  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
-\end{equation}
-Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h$ such that for all $v \in V_h$:
-\begin{equation} \label{eq:3d_dg_disc}
-  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
-  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
-  + \int_{\Gamma_{\rm pia} \cup \Gamma_{\rm LV}} \beta_{\rm pia}[c_h^n][v] 
-  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
-  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
-  = 
-  %\int_{\Omega} f^n v +
-  \int_{\Gamma_{\rm SSAS}} g_{\rm influx} v
-  %, \quad \forall v \in V_h^k
-  .
-\end{equation}
-
-To discretize~\eqref{eq:1d_pde} with the bifurcation
-conditions~\eqref{eq:conc_continuity}
-and~\eqref{eq:flux_conservation}, and the boundary
-conditions~\eqref{eq:1d_bcs}, we use the space of continuous piecewise
-linear polynomials defined over $\Lambda$:
-\begin{equation}
-  \hat V_h
-  = \{ v \in C^0(\Lambda), \; v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda^i) \}, 
-\end{equation}
-where $\mathrm{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$.
-The discrete formulation then reads: for $n = 1,2,\ldots, $ find $\hat c_h^n \in \hat V_h$ such that for all $\hat v \in \hat V_h^k$:
-\begin{align}\label{eq:1d_dg_disc}
-\int_{\Lambda}  \frac{1}{\tau}(A  \hat c_h^n - A \hat c_h^{n-1}) \hat v  + \int_{\Lambda} \hat D A \partial_s \hat c_h^n  \partial_s \hat v^n - \int_{\Lambda} A \hat u \hat c_h^n \partial_s \hat v   +  \int_{\Lambda} \xi P (\hat c_h^n - \overline{c_h^n}) \hat v +a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) = 0.
-\end{align}
- In the above, if $\hat u$ is non--zero on $\Lambda^i$, then the artificial diffusion $a_h^{\mathrm{stab}}$ stabilization term is nonzero  and is given below, for more details see for  \cite[Section 12.6]{quarteroni2009numerical}.
- \begin{align}
-    a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) =  \frac{1}{\|\hat u\|_{L^2(\Lambda^i)}} 
- h_{\Lambda^i} \int_{\Lambda^i} A (\hat u \partial_s \hat c_h  ) (\hat u \partial_s \hat v),  
- \end{align}
- where $h_{\Lambda^i}$ is the mesh--size of $\Lambda^i$.
-Note that the condition \eqref{eq:flux_conservation} is enforced weakly in the above formulation.
-
- \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} complemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0 = 0 $ is the following:
-
- For $n = 1,2, \ldots $, find  $c_h^n \in V_h$ and $\hat c^n_h \in \hat V_h$ such that the coupled equations \eqref{eq:3d_dg_disc} and \eqref{eq:1d_dg_disc} hold for all $v \in V_h$ and for all $\hat v \in \hat V_h$.
 
@@
 
-%% \mar{To examine cardiac-driven dispersion, we compute a steady CSF flow field at peak systolic blood inflow.
-We mimic the resulting reduction of CSF space volume by imposing an inflow of $6\,$ml/s \cite{baledent2014imaging, causemann2022human} and $0.31\,$ml/s \cite{vinje2019respiratory} across the upper arachnoid mater ($\Gamma_{\rm AM-U}$) and the lateral ventricle ($\Gamma_{\rm LV}$) surface, respectively.
-We allow free outflow into the spinal SAS ($\Gamma_{\rm SSAS}$).
-In addition, we upscale the obtained pressure field to account for the lack of pulsatile inertial effects in our model.}
-
-%% % respiratory pulsatility
-%% \mar{Employing the same methodology, we estimate the respiratory contribution to dispersive mixing by computing the CSF flow field driven by an inflow of $0.121\,$ml/s \cite{liu2024using} and $1\,$ml/s \cite{gutierrez2022effect} across the lateral ventricular wall and the arachnoid mater, respectively.}
-
-%% Taking the large spatial differences of pulsatile CSF flow into account, we estimate a spatially varying diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac- and respiration-induced pulsatile CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}.
-These estimates are are derived under simplifying assumptions on the problem geometry, the flow field and concentration profiles \cite{watson1983diffusion, sharp2019dispersion}.
-Therefore such an estimate is heuristic; however, our computational results qualitatively agree with observed data.
-
-%% Similar to the case of CSF flow driven by production introduced in \Cref{sec:csf_flow}, we employ a Stokes flow model to predict CSF velocities corresponding to the peak cardiac and respiratory blood inflow rates.
-In particular, we enforce a uniformly distributed inflow of $u^{\rm in}_{\rm AM}$ and $u^{\rm in}_{\rm LV}$ across the arachnoid mater ($\Gamma_{\rm AM}$) and lateral ventricular surfaces ($\Gamma_{\rm LV}$) respectively, mimicking the effect of reduced CSF volume due to blood inflow.
-We set $u^{\rm in}_{\rm AM} = 6\,$ml/s \cite{causemann2022human,baledent2014imaging} and $u^{\rm in}_{\rm LV}=0.31\,$ml/s \cite{vinje2019respiratory} in the cardiac case, and $u^{\rm{in}}_{\rm{AM}} = 1\,$ml/s \cite{gutierrez2022effect} and $u^{\rm in}_{\rm{LV}}=0.121\,$ml/s \cite{liu2024using} in the respiratory case, respectively.
-%% We prescribe a no-slip boundary condition on the pial surface $\Gamma_{\rm Pia}$ and zero-traction on the spinal boundary $\Gamma_{\rm SSAS}$, making the spinal compartment the only outflow route of CSF.
-Thus, we compute the pulsatile flow fields by augmenting the Stokes flow model (\eqref{eq:momentum_equation} and \ref{eq:divergence_equation}) with the following boundary conditions:
-%% \begin{equation}\label{eq:cardiac_flow_bcs}
-%%     \bm u = 0 \text{ on } \Gamma_{\rm Pia}; \quad 
-%%     \mu \nabla \bm u \cdot \bm n - p \bm n = 0 \text{ on } \Gamma_{\rm SSAS}; \quad 
-%%     \bm u = \frac{u^{\rm in}_{\rm LV} \cdot \bm n}{|\Gamma_{\rm LV}|} \text{ on } \Gamma_{\rm LV}; \quad
-%%     \bm u = \frac{u^{\rm in}_{\rm AM} \cdot \bm n}{|\Gamma_{\rm skull}|} \text{ on } \Gamma_{\rm AM}.
-%% \end{equation}
-
-%% The obtained pressure fields $p_{\rm cardiac}$ and $p_{\rm resp}$ only account for viscous forces, albeit cardiac and respiratory induced CSF flow is known to be unsteady and substantially impacted by transient inertial forces.
-Assuming an oscillatory flow with either the heartbeat frequency $\omega_{\rm cardiac} = 2 \pi$ or respiratory frequency $\omega_{\rm resp} = 0.5 \pi$, a mean gap width of the CSF-filled spaces of $2h=3$\,mm and CSF density $\rho=10^3\,\text{kg/m}^3$, the squared Womersley number $\alpha_i^2 = \frac{h^2 \omega_{i} \rho}{\mu}$ becomes $\alpha_{\rm cardiac}^2 = 20.2$ and $\alpha_{\rm resp}^2=5.05$, which is unsteady.
-Based on theoretical considerations on the ratio of oscillatory flow to steady flow impedances in a tube, we account for inertial forces in the pulsatile flow by upscaling the computed viscous pressure field by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter.
-4.3).
-Thus, the upscaled pressure fields become $p^{\rm up}_{i}= (1 + \alpha_{i}^2 / 8)p_i \text{ for } i = \{\mathrm{cardiac}, \mathrm{resp}\}$.
-
-%% Further assuming unsteady dispersion, we follow Sharp et al \cite{sharp2019dispersion} and estimate each enhancement factor $R_{\rm cardiac}$ and $R_{\rm resp}$ from the non-dimensionalized pressure gradient
-%% $$R_{i} \approx dP_{i}^2 / \alpha_{i}^3, \quad  \text{where} \quad  dP=\frac{|\nabla p^{\rm up}_{i}|}{\omega_{i} \mu / h} \quad \text{for} \quad i = \{\mathrm{cardiac}, \mathrm{resp}\}.$$
-%% Finally, we account for the non-local nature of the dispersion mechanism and smooth the resulting field by solving a Helmholtz problem with $dP+i^2/\alpha_i^3$ as the right-hand side
-%% \begin{equation}
-%% - 10^{-4} \Delta R_i + R_i = dP_i^2/\alpha_i^3 \quad \text{ on } \Omega_{\rm CSF} \qquad \text{ and } \quad \nabla R_i \cdot \bm n = 0, \quad \text{ on } \partial \Omega_{\rm CSF}
-%% \end{equation}
-%% to obtain the locally averaged, spatially varying diffusion enhancement factor $R_i$.
-The diffusion enhancement factor $R$ in $D^{\mathrm{eff}} = D^{\mathrm{Gad}}( 1 +R)$ is then estimated by $R \approx R_{\rm{cardiac}} + R_{\rm{resp}}$.
-
-% %\subsubsection{Cardiac Dispersion Factor Estimation}
-% %\label{sec:dispersion_details}
-
-% To compute a spatially varying estimate of the diffusion enhancement factor $R$, we consider the Stokes flow problem (\cref{eq:momnetum_equation} and \cref{eq:divergence_equation}) equipped with the boundary conditions \cref{eq:cardiac_flow_bcs} describing cardiac induced CSF flow at peak arterial pulsation.
-We use the numerical method described in \cref{sec:details_numerical_method} to solve the Stokes system and compute the non-dimensionalized local pressure difference $dP=\frac{|\nabla p| \rho}{\omega \mu / h}$ pointwise.
-Here, $\mu$ is the CSF viscosity, $\omega = 2 \pi$ is the oscillatory heart beat frequency, $2h=3$\,mm the assumed gap width and $\rho=10^3\,\text{kg/m}^3$ the density of the CSF.
-
-% In addition, we account for inertial forces in the pulsatile flow by upscaling the obtained pressure gradient by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter.
-4.3), where $\alpha^2 = \frac{h^2 \omega \rho}{\mu} = 20.2$ is the square of the Womersley number for an oscillatory flow.
-
-% Finally, we smooth the resulting field by solving a Helmholtz problem with $dP^2/alpha^3$ as the right-hand side
-
-% \begin{equation}
-% - 10^{-4} \Delta R + R = dP^2/\alpha^3 \quad \text{ on } \Omega_{\rm CSF} \qquad \text{ and } \quad \nabla R \cdot \bm n = 0 \quad \text{ on } \partial \Omega_{\rm CSF}
-% \end{equation}
-
-% to obtain the locally averaged, spatially varying diffusion enhancement factor $R$.
-
 \subsection{Estimating net perivascular flow induced by peristaltic waves}
@@
 Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L_i$, baseline inner radius $R_1^i$, fixed outer radius $R_2^i$, and outer-to-inner ratio $\beta_i = R_2^i/R_1^i$.
-These geometric parameters and the assumption of annular cylindrical PVS segments yield hydraulic resistances $\mathcal{R}_{o, i}$ and additional characteristic parameters, see \cite{gjerde2023directional} for the complete definitions and schematics.
+These geometric parameters and the assumption of annular cylindrical PVS segments yield hydraulic resistances and additional characteristic parameters, see \cite{gjerde2023directional} for the complete definitions and schematics.
 Since the analytical estimate is derived under the assumption that $k L_i \approx \mathcal{O} (1)$~\cite{gjerde2023directional} and has been verified against numerical simulations for $k L_i$ of the order $10^{-1}-10^2$~\cite[Table I]{gjerde2023directional}, we consider it applicable for the wave lengths and vascular network data considered here in which $k L_i$ range from 0.15 to 30 for the strong vasomotion and 0.0015 to 0.30 for the cardiac waves (\Cref{tab:pvs:parameters}).
@@
 (\beta_i^2 - 1) {R_1^i}^2$.
+
+\subsection{Modeling and simulating solute transport and exchange in intracranial domains (3D) and perivascular networks (1D)}
+\label{sec:app:transport_eqs}
+
+In this section, we describe the mathematical model (eq.~(1)), the
+associated definitions, interface and boundary conditions in further
+detail. We refer to the main text for explicit parameter values while
+giving the general, abstract form here, and to the
+reference~\cite{masri2024modelling, laurino2019derivation} for the
+derivation and analysis of this 3D-1D model. Recall that we model a
+concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
+defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
+each of these domains, $c$ satisfies
+%\mer{MER: Rami asks if it should be a $\phi$ in the advective term here.
+We can include it, or drop it, but in the latter case, perhaps we should clarify that u is an average/superficial/Darcy velocity.}
+\begin{equation}
+  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma
+  %= f
+   = 0
+  \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
+  \label{eq:3d_pde}
+\end{equation}
+In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$).
+In the parenchyma, $\phi$ represents the extracellular space volume fraction, and thus $c$ here generally represents the intrinsic (in contrast to the superficial) concentration~\cite{sykova2008diffusion}.
+Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$.
+$\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivascular networks $\Lambda_a, \Lambda_v$.
+To summarize,
+\begin{equation*}
+\phi =  \begin{cases}
+  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
+  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
+  \end{cases}, \; 
+D = \begin{cases}
+  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
+  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
+  \quad \bm u  = \begin{cases}
+  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
+  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
+\end{cases}, \; 
+\xi = \begin{cases}
+  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
+  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
+\end{cases} . 
+\end{equation*}
+In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp.
+$|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp.
+$\Omega_{\rm{CSF}}$).
+If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.
+
+Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by
+\begin{equation*}
+  \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta_2^i(s)} c 
+\end{equation*}
+where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$.
+Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the term $\delta_\Gamma$ is concentrated on the outer lateral surfaces of the PVSs, and defined in terms of its action on any sufficiently smooth function $v : \Omega \rightarrow \R$
+%and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$
+such that
+\begin{equation*}
+  \langle \xi (\overline{c} - \hat c ) \delta_\Gamma, v \rangle =
+  \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
+\end{equation*}
+under the assumption that $\xi$ is constant on each $\partial
+\Theta_2^i$.
+%% , since:
+%% \begin{multline}
+%%   \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
+%%   \int_{\Gamma} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
+%%   \sum_i \int_{\Gamma^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = 
+%%   \sum_i \int_{\Lambda^i} \int_{\partial \Theta_2^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \\
+%%   \sum_i \int_{\Lambda^i} \left (  \xi (\overline{c} - \hat c ) \int_{\partial \Theta_2^i} \delta_\Gamma (v) \right ) =
+%%   \sum_i \int_{\Lambda^i} P \xi (\overline{c} - \hat c ) \overline{v} =
+%%   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v} .
+%% \end{multline}
+%% \mer{MER: The previous was:
+%% \begin{equation}
+%%   \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
+%% \end{equation}
+%% }
+
+On the interface between the brain and the CSF spaces
+$\Gamma_{\mathrm{pia}} \cup \Gamma_{\rm LV}$, we prescribe the
+following interface conditions, which represent a semi--permeable
+interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and
+$c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
+%% \begin{subequations}
+%% \begin{alignat}{2}
+%% (-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} &= - (- D (\phi \nabla c_{\rm{CSF}}) + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n} &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\
+%% (-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,
+%% \end{alignat} 
+%% \end{subequations}
+\begin{subequations}
+  \begin{alignat}{2}
+  (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n}
+  &= - (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm CSF}} \cdot \bm{n}
+  &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
+  (- D \nabla (\phi  c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
+\end{alignat} 
+  \label{eq:3d_interface}%
+\end{subequations}%
+where $\bm{n} = \bm{n}_{\rm PAR}$ is the normal vector field defined
+over the interface, outward-pointing when viewed from $\Omega_{\rm
+  PAR}$, and $\beta_{\mathrm{pia}} \geqslant 0$ is a permeability
+constant.
+
+We supplement~\eqref{eq:3d_pde} and~\eqref{eq:3d_interface} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
+\begin{subequations}
+\begin{alignat}{2}
+(D \nabla (\phi c) - \bm u c ) \cdot \bm{n} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \Gamma_{\rm SSAS},  \\ 
+% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\Gamma_{\rm SAS}} &= 0 &&  \quad  \mathrm{on} \quad \Gamma_{\rm SAS}^{\mathrm{out}}  \\ 
+(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = \beta_{\rm exit} c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-U},  \\ 
+(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = 0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-L} \cup \Gamma_{\rm SC}  .
+\end{alignat} 
+\label{eq:3d_bcs}
+\end{subequations}
+where again $\bm n$ denotes the outward-pointing boundary normal.
+
+The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
+perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
+in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
+space averaged over each perivascular cross-section and is governed by 
+\begin{equation}
+  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s \hat c  ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
+  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
+\label{eq:1d_pde}
+\end{equation}
+where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
+of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
+PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
+associated PVS $\Omega^i$, the area of the cross-section
+$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
+and $\hat{u}$ is a convective velocity representing an (average) CSF
+flow velocity in the axial direction of the perivascular spaces.
+
+To complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
+internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
+boundary conditions at the end nodes. To this end, define the set of
+internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
+that are connected to two or more edges and the set of end nodes as
+$\mathcal{N} \subset \mathcal{V}$, write $\hat{c}^i =
+\hat{c}|_{\Lambda^i}$, and for $y\in\mathcal{B}$ let $\mathcal{E}(y)$ denote the set of
+(two or three) edges in $\{\Lambda^i\}$ sharing the node $y$. To
+impose continuity of the concentrations, we set that the
+concentrations when viewed from each edge must match at nodes:
+\begin{equation}
+  \hat{c}^i (y) = \hat{c}^j(y)
+  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
+  \qquad \foralls y \in \mathcal{B} .
+  \label{eq:conc_continuity}
+\end{equation}
+Moreover, to ensure mass conservation, we set that the flux
+going in and out at nodes should add to zero, or more precisely, that
+for each $y \in \mathcal{B}$:
+\begin{equation}
+  \sum_{\Lambda^i \in \mathcal{E}(y)}
+  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n^i(y)
+  = 0.
+  \label{eq:flux_conservation}
+\end{equation}
+Here the normal $n^i$ is as defined in ~\eqref{eq:def:sign}.
+%Here the (normal ``vector") function $n^i$ takes the values %in $\left\{-1, 1\right\}$ and 
+%defines an orientation of vertices of edge $\Lambda^i$. %Specifically, for an edge $\Lambda^i$ oriented from $y_{\rm %in}$ to $y_{\rm out}$ we set 
+%\begin{equation}
+%  n^i(y_{\rm in}) = 1,\quad
+%  n^i(y_{\rm{out}}) = -1. %\quad
+%  %\mk{n^i(y) = 0 \quad \foralls y \in 
+%  %\mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} }.  
+%\label{eq:def:sign}
+%\end{equation}
+%\mk{Why do we need the normal on non-IO vertices?}
+
+Finally, we impose a homogeneous Neumann condition at all network end
+nodes to augment~\eqref{eq:1d_pde}:
+\begin{equation}
+    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{N}.
+    \label{eq:1d_bcs}
+\end{equation}
+This no-flux condition states that the network end nodes represent
+barriers for molecular efflux into or out of the perivascular network,
+and thus that all solute exchange between the PVSs and their
+surroundings takes place via the lateral outer PVS surface and is
+regulated by the exchange parameter $\xi$ cf.~\eqref{eq:3d_pde}
+and~\eqref{eq:1d_pde}. Larger particles have been observed to
+accumulate within the PVS as the surface arteries penetrate into the
+brain parenchyma~\cite{mestre2018flow, bedussi2018paravascular},
+and~\eqref{eq:1d_bcs} is appropriate to represent such
+behavior. However, to represent a continuously extending PVSs also
+along penetrating vessels, this condition should be revisited.
+
+%% We consider non-zero convection in the CSF space $u_{\rm CSF} \not = 0$ in \eqref{eq:multi_transport_3d} resulting from steady state CSF production.
+Recall that $\Omega_{\mathrm{CSF}}$ represents the CSF space and that $\Gamma_{\mathrm{skull^{u/l}}}$ and $\Gamma_{\mathrm{pia}}$, $\Gamma_{\mathrm{LV}}$ represent the boundaries facing the upper and lower parts of the dura, the pia and the lateral ventricles, respectively.
+Assuming CSF production in the choroid plexus located at the lateral ventricle wall, we solve the  steady state Stokes equations for the velocity and the pressure $(\bm u_{\mathrm{CSF}}, p_{\mathrm{CSF}})$ in the CSF domain:
+
+%\subsection{Numerical solution of the transport and flow equations}
+%\label{sec:app:details_numerical_method}
+%Note in particular that $\mathcal{F}_i \cap \mathcal{F}_{pia}} =
+%  \emptyset$.
+
+%The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is
+%denoted by $\mathcal{E}_{h}$.
+
+\textit{Finite element solution of coupled 3D-1D solute transport equations}. 
+We consider the system of coupled 3D-1D solute transport equations
+given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
+conditions~\eqref{eq:3d_interface}, the bifurcation
+conditions~\eqref{eq:conc_continuity}
+and~\eqref{eq:flux_conservation}, and the boundary
+conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
+these equations using an implicit finite difference scheme in time, 
+a discontinuous Galerkin (DG) finite element method with upwinding in space for the 3D domain to accurately capture sharp boundary layers, and a continuous Galerkin method for the 1D networks.
+We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm, and accounting for the increased diffusivity due to dispersion).
+
+We first consider the discretization of~\eqref{eq:3d_pde} and
+introduce the discrete space
+\begin{equation}
+  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
+  V_h = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^1(E), \; E \in \mathcal{T}\} ,
+\end{equation}
+where $L^2(\Omega)$ is the space of square-integrable functions
+defined over $\Omega$ and $\mathrm{P}^1(E)$ denotes the space of
+polynomials of total degree $\leqslant 1$ defined over the tetrahedra
+$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
+symmetric weighted interior penalty DG formulation, referring
+to~\cite{ern2009discontinuous} and \cite[Section
+  4.5.2.3]{di2011mathematical} for details on this
+method. Recall that $\phi$ is constant in each domain
+  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
+  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.
+Define for $c, v \in V_h^k$:
+\begin{equation}
+  a_h (c, v)
+  =
+  \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
+  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
+  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right) 
+   + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
+  .
+  \label{eq:dg_sip}
+\end{equation}
+Here, for each facet $F \in \mathcal{F}^i$ shared
+between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
+pointing from $E^1$ to $E^2$. The facet diameter is denoted by $h_F$,
+the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
+unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
+are defined as:
+\begin{equation}
+  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
+  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
+  \quad \text{where} \quad \kappa_i = (D \phi) |_{E^i}.  
+\end{equation}
+In~\eqref{eq:dg_sip} the parameter $\gamma_{D,F}$ is the harmonic mean of the
+porosity-weighted diffusion coefficient given by
+\begin{equation}
+  \gamma_{D,F} = \frac{2\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
+\end{equation}
+and $\eta$ is a user-specified penalty parameter (we set $\eta=1000$). To discretize the convection term in~\eqref{eq:3d_pde}, we use
+upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
+references therein. For $c, v \in V_h^k$, define
+\begin{equation}
+  a_h^{\mathrm{up}}(c, v)
+  = - \sum_{E \in \mathcal{T}} \int_E c\bm{u} \cdot \nabla v
+  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
+  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
+\end{equation}
+Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h$ such that for all $v \in V_h$:
+\begin{equation} \label{eq:3d_dg_disc}
+  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
+  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
+  + \int_{\Gamma_{\rm pia} \cup \Gamma_{\rm LV}} \beta_{\rm pia}[c_h^n][v] 
+  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
+  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
+  = 
+  %\int_{\Omega} f^n v +
+  \int_{\Gamma_{\rm SSAS}} g_{\rm influx} v
+  %, \quad \forall v \in V_h^k
+  .
+\end{equation}
+
+To discretize~\eqref{eq:1d_pde} with the bifurcation
+conditions~\eqref{eq:conc_continuity}
+and~\eqref{eq:flux_conservation}, and the boundary
+conditions~\eqref{eq:1d_bcs}, we use the space of continuous piecewise
+linear polynomials defined over $\Lambda$:
+\begin{equation}
+  \hat V_h
+  = \{ v \in C^0(\Lambda), \; v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda^i) \}, 
+\end{equation}
+where $\mathrm{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$.
+The discrete formulation then reads: for $n = 1,2,\ldots, $ find $\hat c_h^n \in \hat V_h$ such that for all $\hat v \in \hat V_h^k$:
+\begin{align}\label{eq:1d_dg_disc}
+\int_{\Lambda}  \frac{1}{\tau}(A  \hat c_h^n - A \hat c_h^{n-1}) \hat v  + \int_{\Lambda} \hat D A \partial_s \hat c_h^n  \partial_s \hat v^n - \int_{\Lambda} A \hat u \hat c_h^n \partial_s \hat v   +  \int_{\Lambda} \xi P (\hat c_h^n - \overline{c_h^n}) \hat v +a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) = 0.
+\end{align}
+ In the above, if $\hat u$ is non--zero on $\Lambda^i$, then the artificial diffusion $a_h^{\mathrm{stab}}$ stabilization term is nonzero  and is given below, for more details see for  \cite[Section 12.6]{quarteroni2009numerical}.
+ \begin{align}
+    a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) =  \frac{1}{\|\hat u\|_{L^2(\Lambda^i)}} 
+ h_{\Lambda^i} \int_{\Lambda^i} A (\hat u \partial_s \hat c_h  ) (\hat u \partial_s \hat v),  
+ \end{align}
+ where $h_{\Lambda^i}$ is the mesh--size of $\Lambda^i$.
+Note that the condition \eqref{eq:flux_conservation} is enforced weakly in the above formulation.
+
+ \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} complemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0 = 0 $ is the following:
+
+ For $n = 1,2, \ldots $, find  $c_h^n \in V_h$ and $\hat c^n_h \in \hat V_h$ such that the coupled equations \eqref{eq:3d_dg_disc} and \eqref{eq:1d_dg_disc} hold for all $v \in V_h$ and for all $\hat v \in \hat V_h$.
+
+%% \mar{To examine cardiac-driven dispersion, we compute a steady CSF flow field at peak systolic blood inflow.
+We mimic the resulting reduction of CSF space volume by imposing an inflow of $6\,$ml/s \cite{baledent2014imaging, causemann2022human} and $0.31\,$ml/s \cite{vinje2019respiratory} across the upper arachnoid mater ($\Gamma_{\rm AM-U}$) and the lateral ventricle ($\Gamma_{\rm LV}$) surface, respectively.
+We allow free outflow into the spinal SAS ($\Gamma_{\rm SSAS}$).
+In addition, we upscale the obtained pressure field to account for the lack of pulsatile inertial effects in our model.}
+
+%% % respiratory pulsatility
+%% \mar{Employing the same methodology, we estimate the respiratory contribution to dispersive mixing by computing the CSF flow field driven by an inflow of $0.121\,$ml/s \cite{liu2024using} and $1\,$ml/s \cite{gutierrez2022effect} across the lateral ventricular wall and the arachnoid mater, respectively.}
+
+%% Taking the large spatial differences of pulsatile CSF flow into account, we estimate a spatially varying diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac- and respiration-induced pulsatile CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}.
+These estimates are are derived under simplifying assumptions on the problem geometry, the flow field and concentration profiles \cite{watson1983diffusion, sharp2019dispersion}.
+Therefore such an estimate is heuristic; however, our computational results qualitatively agree with observed data.
+
+%% Similar to the case of CSF flow driven by production introduced in \Cref{sec:csf_flow}, we employ a Stokes flow model to predict CSF velocities corresponding to the peak cardiac and respiratory blood inflow rates.
+In particular, we enforce a uniformly distributed inflow of $u^{\rm in}_{\rm AM}$ and $u^{\rm in}_{\rm LV}$ across the arachnoid mater ($\Gamma_{\rm AM}$) and lateral ventricular surfaces ($\Gamma_{\rm LV}$) respectively, mimicking the effect of reduced CSF volume due to blood inflow.
+We set $u^{\rm in}_{\rm AM} = 6\,$ml/s \cite{causemann2022human,baledent2014imaging} and $u^{\rm in}_{\rm LV}=0.31\,$ml/s \cite{vinje2019respiratory} in the cardiac case, and $u^{\rm{in}}_{\rm{AM}} = 1\,$ml/s \cite{gutierrez2022effect} and $u^{\rm in}_{\rm{LV}}=0.121\,$ml/s \cite{liu2024using} in the respiratory case, respectively.
+%% We prescribe a no-slip boundary condition on the pial surface $\Gamma_{\rm Pia}$ and zero-traction on the spinal boundary $\Gamma_{\rm SSAS}$, making the spinal compartment the only outflow route of CSF.
+Thus, we compute the pulsatile flow fields by augmenting the Stokes flow model (\eqref{eq:momentum_equation} and \ref{eq:divergence_equation}) with the following boundary conditions:
+%% \begin{equation}\label{eq:cardiac_flow_bcs}
+%%     \bm u = 0 \text{ on } \Gamma_{\rm Pia}; \quad 
+%%     \mu \nabla \bm u \cdot \bm n - p \bm n = 0 \text{ on } \Gamma_{\rm SSAS}; \quad 
+%%     \bm u = \frac{u^{\rm in}_{\rm LV} \cdot \bm n}{|\Gamma_{\rm LV}|} \text{ on } \Gamma_{\rm LV}; \quad
+%%     \bm u = \frac{u^{\rm in}_{\rm AM} \cdot \bm n}{|\Gamma_{\rm skull}|} \text{ on } \Gamma_{\rm AM}.
+%% \end{equation}
+
+%% The obtained pressure fields $p_{\rm cardiac}$ and $p_{\rm resp}$ only account for viscous forces, albeit cardiac and respiratory induced CSF flow is known to be unsteady and substantially impacted by transient inertial forces.
+Assuming an oscillatory flow with either the heartbeat frequency $\omega_{\rm cardiac} = 2 \pi$ or respiratory frequency $\omega_{\rm resp} = 0.5 \pi$, a mean gap width of the CSF-filled spaces of $2h=3$\,mm and CSF density $\rho=10^3\,\text{kg/m}^3$, the squared Womersley number $\alpha_i^2 = \frac{h^2 \omega_{i} \rho}{\mu}$ becomes $\alpha_{\rm cardiac}^2 = 20.2$ and $\alpha_{\rm resp}^2=5.05$, which is unsteady.
+Based on theoretical considerations on the ratio of oscillatory flow to steady flow impedances in a tube, we account for inertial forces in the pulsatile flow by upscaling the computed viscous pressure field by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter.
+4.3).
+Thus, the upscaled pressure fields become $p^{\rm up}_{i}= (1 + \alpha_{i}^2 / 8)p_i \text{ for } i = \{\mathrm{cardiac}, \mathrm{resp}\}$.
+
+%% Further assuming unsteady dispersion, we follow Sharp et al \cite{sharp2019dispersion} and estimate each enhancement factor $R_{\rm cardiac}$ and $R_{\rm resp}$ from the non-dimensionalized pressure gradient
+%% $$R_{i} \approx dP_{i}^2 / \alpha_{i}^3, \quad  \text{where} \quad  dP=\frac{|\nabla p^{\rm up}_{i}|}{\omega_{i} \mu / h} \quad \text{for} \quad i = \{\mathrm{cardiac}, \mathrm{resp}\}.$$
+%% Finally, we account for the non-local nature of the dispersion mechanism and smooth the resulting field by solving a Helmholtz problem with $dP+i^2/\alpha_i^3$ as the right-hand side
+%% \begin{equation}
+%% - 10^{-4} \Delta R_i + R_i = dP_i^2/\alpha_i^3 \quad \text{ on } \Omega_{\rm CSF} \qquad \text{ and } \quad \nabla R_i \cdot \bm n = 0, \quad \text{ on } \partial \Omega_{\rm CSF}
+%% \end{equation}
+%% to obtain the locally averaged, spatially varying diffusion enhancement factor $R_i$.
+The diffusion enhancement factor $R$ in $D^{\mathrm{eff}} = D^{\mathrm{Gad}}( 1 +R)$ is then estimated by $R \approx R_{\rm{cardiac}} + R_{\rm{resp}}$.
+
+% %\subsubsection{Cardiac Dispersion Factor Estimation}
+% %\label{sec:dispersion_details}
+
+% To compute a spatially varying estimate of the diffusion enhancement factor $R$, we consider the Stokes flow problem (\cref{eq:momnetum_equation} and \cref{eq:divergence_equation}) equipped with the boundary conditions \cref{eq:cardiac_flow_bcs} describing cardiac induced CSF flow at peak arterial pulsation.
+We use the numerical method described in \cref{sec:details_numerical_method} to solve the Stokes system and compute the non-dimensionalized local pressure difference $dP=\frac{|\nabla p| \rho}{\omega \mu / h}$ pointwise.
+Here, $\mu$ is the CSF viscosity, $\omega = 2 \pi$ is the oscillatory heart beat frequency, $2h=3$\,mm the assumed gap width and $\rho=10^3\,\text{kg/m}^3$ the density of the CSF.
+
+% In addition, we account for inertial forces in the pulsatile flow by upscaling the obtained pressure gradient by a factor of $1 + \alpha^2 / 8$ (see \cite{van1998cardiovascular}, chapter.
+4.3), where $\alpha^2 = \frac{h^2 \omega \rho}{\mu} = 20.2$ is the square of the Womersley number for an oscillatory flow.
+
+% Finally, we smooth the resulting field by solving a Helmholtz problem with $dP^2/alpha^3$ as the right-hand side
+
+% \begin{equation}
+% - 10^{-4} \Delta R + R = dP^2/\alpha^3 \quad \text{ on } \Omega_{\rm CSF} \qquad \text{ and } \quad \nabla R \cdot \bm n = 0 \quad \text{ on } \partial \Omega_{\rm CSF}
+% \end{equation}
+
+% to obtain the locally averaged, spatially varying diffusion enhancement factor $R$.
 

### 2025-06-20 paper/main.tex
Spell-check and PVSAS

@@
 performance computing, we introduce an integrated computational model
-of intracranial molecular transport. Focusing on the
-interplay between perivascular pathways, pulsatility and CSF flow
-dynamics, the model predicts the temporal evolution and spatial
-distribution of a solute concentration within a detailed geometric
-representation of the human SAS and ventricular system, networks of
-surface PVSs and the brain parenchyma. In terms of transport dynamics,
-we account for heterogeneous diffusion, dispersive mixing induced by
-cardiac and respiratory pulsatility in the CSF spaces and PVSs,
-convective fluid flow driven by CSF production and peristaltic
-pumping, as well as solute exchange and clearance across
-semi-permeable membranes. Comparing with glymphatic MRI
+of intracranial molecular transport. Focusing on the interplay between
+perivascular pathways, pulsatility and CSF flow dynamics, the model
+predicts the temporal evolution and spatial distribution of a solute
+concentration within a detailed geometric representation of the human
+SAS and ventricular system, in networks of surface PVSs (i.e.~those
+surrounding arteries and veins in the SAS), and in the brain
+parenchyma. In terms of transport dynamics, we account for
+heterogeneous diffusion, dispersive mixing induced by cardiac and
+respiratory pulsatility in the CSF spaces and PVSs, convective fluid
+flow driven by CSF production and peristaltic pumping, as well as
+solute exchange and clearance across semi-permeable
+membranes. Comparing with glymphatic MRI
 studies~\cite{ringstad2017glymphatic, ringstad2018brain,
-  watts2019measuring}, the in-silico predictions accurately represent
-molecular enrichment patterns, timing and intercompartmental
-distributions. This open platform~\cite{ZENODO} thus provides a
-technological opportunity for qualitatively and quantitatively
-exploring key open questions relating to molecular movement within the
-human brain environment such as the role of pulsatility, perivascular
-pathways, structural compartmentalization, or morphology.
+  watts2019measuring, eide2024functional}, the in-silico predictions
+accurately represent molecular enrichment patterns, timing and
+intercompartmental distributions. This open platform~\cite{ZENODO}
+thus provides a technological opportunity for qualitatively and
+quantitatively exploring key open questions relating to molecular
+movement within the human brain environment such as the role of
+pulsatility, perivascular pathways, structural compartmentalization,
+or morphology.
 
@@
     A) Surface arterial network colored by arterial radius with vessel locations labeled;
-    B) Net PVS flow induced by CSF production alone (see also ~\Cref{fig:csf}D).
+    B) Net PVS flow induced by CSF production alone (see also~\Cref{fig:csf}D).
     Histograms show the relative frequencies of local periarterial velocities. Positive values indicate antegrade PVS flow, while negative values indicate retrograde PVS flow;
@@
 mmol/l (relative to a total amount of 1.0 mmol) occuring after 10 to 15 hours. 
-Our in-silico enrichment
-patterns agree with these observations, though with an earlier rather than later peak in the CSF spaces (2 hours after first tracer appearance at the craniocervical junction).
+Our in-silico enrichment patterns agree with these observations, though with an earlier rather than later peak in the CSF spaces (2 hours after first tracer appearance at the craniocervical junction).
 In particular, we note that the substantial individual variation observed clinically is comparable to the variation in in-silico enrichment patterns associated with reduced CSF production or reduced intracranial pulsatility.
@@
 
-Our simulation results highlight that reproducing clinically observed tracer timings requires flow drivers beyond CSF production and cardiac peristalsis.
-Similar to the CSF spaces, dispersion due to cardiac and respiratory pulsation has been hypothesized to enhance transport in the perivascular spaces.
-However, recent estimates indicate that the effect of dispersion does not exceed a factor of 2 compared to diffusion alone \cite{asgari2016glymphatic,sharp2019dispersion,bojarskaite2023sleep,asgari2016glymphatic,troyetsky2021dispersion}.
-This conclusion is supported by our own model, where a 100-fold increase in the PVS dispersion coefficient did not substantially accelerate transport (\cref{fig:pvs_b}H).
-To illustrate the need for a stronger advective driver, we compared our findings to tracer arrival times reported by Eide and Ringstad~\cite{eide2024functional}, who observed a delay of just 15.3 min between the M1 and M2 segments of the MCA.
-Our baseline model, accounting only for CSF production and cardiac-driven flow, predicted a physiologically unrealistic delay of over three hours.
-In sharp contrast, our 'high PVS flow' scenario, which incorporates the effects of low-frequency vasomotion, yielded a much more consistent delay of 24 minutes.
-Furthermore, after adjusting for the 13 min travel time from injection to the craniocervical junction~\cite{eide2024functional}, the 'high PVS flow' model's FTAs (84––108 min for M1) aligns well with the upper end of the reported clinical range (37.8 $\pm$ 47.0), particularly when considering potentially delayed transport due to the limited caudal extend of our PVS network.
-Therefore, we interpret these results as strong indication that PVS flow driven by CSF production and cardiac peristalsis alone is insufficient to explain rapid perivascular transport observed in humans.
+Importantly, our simulation results indicate that the tracer
+enrichment observed clinically relies on the presence of one or more
+flow drivers in addition to CSF production and cardiac
+peristalsis. Dispersion due to cardiac and respiratory pulsations has
+been hypothesized to enhance transport in the perivascular spaces as
+well as in the CSF spaces. However, recent estimates indicate that the
+effect of dispersion in the PVS does not exceed a factor of 2 compared
+to diffusion alone
+\cite{asgari2016glymphatic,sharp2019dispersion,bojarskaite2023sleep,asgari2016glymphatic,troyetsky2021dispersion}. Our
+model demonstrates that not even a 100-fold increase in the PVS
+dispersion coefficient accelerates transport substantially
+(\Cref{fig:pvs_b}H). To illustrate the need for a stronger advective
+driver, we compared our findings to tracer arrival times reported by
+Eide and Ringstad~\cite{eide2024functional}, who observed a delay of
+just 15.3 min between the M1 and M2 segments of the MCA. First, we
+note that the spaces studied by Eide and
+Ringstad~\cite{eide2024functional}, labeled therein as the PVSAS, are
+naturally compared to the parts of our surface PVS networks embedded
+within the SAS. Our baseline model, accounting only for CSF production
+and cardiac-driven flow, predicted a physiologically unrealistic delay
+of over three hours. In contrast, our high PVS flow scenario, which
+incorporates the effects of low-frequency vasomotion, yielded a more
+consistent delay of 24 minutes. Furthermore, after adjusting for the
+13 min travel time from injection to the craniocervical
+junction~\cite{eide2024functional}, the high PVS flow model's FTAs
+(84--108 min for M1) aligns well with the upper end of the reported
+clinical range (37.8 $\pm$ 47.0), particularly when considering
+potentially delayed transport due to the limited caudal extent of our
+PVS network. All in all, we interpret these results as strongly
+indicative of that PVS flow driven by CSF production and cardiac
+peristalsis alone is insufficient to explain the rapid perivascular
+transport observed in humans.
 
@@
 flow at the craniocervical junction, or the effect of microanatomical
-features such as subarachnoid trabeculae on flow resitance and dispersion
+features such as subarachnoid trabeculae on flow resistance and dispersion
 in the SAS. However, our approach is supported by recent clinical 
@@
 Each Stokes flow simulation utilizes a hybrid parallel approach with 16 MPI processes and 4 threads each, requiring approximately 300 GB of RAM and completing in 23 minutes.
-The transport model is limited to serial execution by FEniCS\textsubscript{ii}, but employes 16 threads for the linear system factorization, taking 22 minutes and consuming about 50 GB of memory on a AMD EPYC 9684X machine.
+The transport model is limited to serial execution by FEniCS\textsubscript{ii}, but employs 16 threads for the linear system factorization, taking 22 minutes and consuming about 50 GB of memory on a AMD EPYC 9684X machine.
 

### 2025-06-20 paper/main.tex
Attempt to answer Reviewer #1

@@
 a basic transport model within the parenchyma, modelling extracellular
-diffusion alone. To compensate, we have focused on reporting detailed
+diffusion alone and zero bulk interstitial fluid flow. To compensate, we have focused on reporting detailed
 predictions in the CSF spaces and surface PVS after not more than 24

### 2025-06-20 paper/main.tex
More work on reviewer comments

--- 2025-06-20 paper/main.tex
@@
 a basic transport model within the parenchyma, modelling extracellular
-diffusion alone. To compensate, we have focused on reporting detailed
-predictions in the CSF spaces and surface PVS after not more than 24
-hours. In particular, we note that, we observe tracer accumulation at
-the perivascular network ends after 12 hours, similar to that
-previously observed for microspheres~\cite{bedussi2018paravascular,
-  mestre2018flow}, which here can be interpreted as an artifact of the
-incomplete network creating a barrier for tracer movement. Previously,
-by inferring transport parameters within the human brain from
-glymphatic MRI data, we have found that molecular transport within the
-parenchyma is well-represented by $\approx 3\times$ enhanced diffusion
-augmented by local clearance e.g.~across the blood-brain barrier, or
-by diffusion-convection with a complex pattern of spatially-varying
+diffusion alone and no bulk interstitial fluid flow. To compensate, we
+have focused on reporting detailed predictions in the CSF spaces and
+surface PVS after not more than 24 hours. In particular, we note that,
+we observe tracer accumulation at the perivascular network ends after
+12 hours, similar to that previously observed for
+microspheres~\cite{bedussi2018paravascular, mestre2018flow}, which
+here can be interpreted as an artifact of the incomplete network
+creating a barrier for tracer movement. Previously, by inferring
+transport parameters within the human brain from glymphatic MRI data,
+we have found that molecular transport within the parenchyma is
+well-represented by $\approx 3\times$ enhanced diffusion augmented by
+local clearance e.g.~across the blood-brain barrier, or by
+diffusion-convection with a complex pattern of spatially-varying
 interstitial fluid flow on the order of 1-10 \textmu
@@
 concentration appearing at the craniocervical junction rather than as
-an intrathecal injection. We thus neglect absorption of both CSF and tracer in the spinal compartment, and therefore overestimate the amount of
-tracer entering into the intracranial spaces; however, due to the
-linearity of the mathematical model, it is valid to directly interpret
-the in-silico concentrations relative to the total amount of tracer.
+an intrathecal injection. We thus neglect absorption of both CSF and
+tracer in the spinal compartment, and therefore overestimate the
+amount of tracer entering into the intracranial spaces; however, due
+to the linearity of the mathematical model, it is valid to directly
+interpret the in-silico concentrations relative to the total amount of
+tracer.
 
@@
 directionality of perivascular flow, and integrate different physical
-mechanisms across spatial and temporal scales. The complete
-simulation pipeline is openly available, including interactive
-visualization of simulation results for all model
-variations~\cite{ZENODO}. Looking ahead, this platform establishes a
-foundation for in-silico studies of molecular movement within the
-human brain environment, such as tailored predictions of intrathecal
-chemotherapy delivery or personalized diagnostics of brain clearance
-capacity.
+mechanisms across spatial and temporal scales. The complete simulation
+pipeline is openly available, including interactive visualization of
+simulation results for all model variations~\cite{ZENODO}. Future work
+will focus on further model validation including comparisons between
+in-silico predictions and glymphatic MRI for a number of subjects
+across several patient cohorts. Looking ahead, this platform
+establishes a foundation for in-silico studies of molecular movement
+within the human brain environment, such as tailored predictions of
+intrathecal chemotherapy delivery or personalized diagnostics of brain
+clearance capacity.
 

### 2025-06-20 paper/main.tex
Add data availability statement.

@@
 
+\section*{Data availability}
+
+The software developed to generate and analyze the numerical results
+of this study are openly available on Zenodo with the identifier doi:
+10.5281/zenodo.14749163. In addition to simulation and analysis
+software, this archive also includes surface triangulations used for
+geometry representation, computational meshes (3D meshes and 1D
+networks), segmentations, sample simulation outputs from the baseline
+and high PVS flow models, and animations.
+
 \bibliography{references}

### 2025-06-20 paper/supplementary.tex
Fix typos.

@@
 Investigating the link between cardiac pulsatility and the resulting diffusion enhancement factor, we compute for $R$ different levels of cardiac pulsatility.
-We reduce the inflow boundary conditions of $6\,$ml/s (across the SAS outer surface) and $0.31\,$ml/s (across the lateral ventricle surface) by $25\,\%$, $50\,\%$, $75\,\%$ and $100\,\%$ to model lower cardiac pulsatility, reaveling that $R$ scales quadratically with the reduction of inflow (\cref{fig:R_comparison}).
+We reduce the inflow boundary conditions of $6\,$ml/s (across the SAS outer surface) and $0.31\,$ml/s (across the lateral ventricle surface) by $25\,\%$, $50\,\%$, $75\,\%$ and $100\,\%$ to model lower cardiac pulsatility, revealing that $R$ scales quadratically with the reduction of inflow (\cref{fig:R_comparison}).
 This behaviour directly reflects the nature of our model - the Stokes flow is linear in its boundary conditions, and the diffusion enhancement factor quadratic in the predicted pressure gradient - and simultaneously underscores the substantial role of pulsatility in shaping intracranial transport.
@@
     \includegraphics[width=1.0\linewidth]{figures/cardiac_R_comparison.png}
-    \caption{Comparison of cardicac-driven diffusion enhancement factors for different levels of cardiac pulsatility.
+    \caption{Comparison of cardiac-driven diffusion enhancement factors for different levels of cardiac pulsatility.
 From left to right: The diffusion enhancement factor $R_c$ for $25\,\%$, $50\,\%$, $75\,\%$ and $100\,\%$ of the inflow boundary conditions of $6\,$ml/s and $0.31\,$ml/s across the SAS outer surface and the lateral ventricle surface, respectively.}

### 2025-06-24 paper/main.tex
Fix typo.
Updates from Overleaf

@@
 The software developed to generate and analyze the numerical results
-of this study are openly available on Zenodo with the identifier doi:
+of this study is openly available on Zenodo with the identifier doi:
 10.5281/zenodo.14749163. In addition to simulation and analysis

### 2025-06-24 paper/main.tex
weaken intro sentence

@@
 flow velocities, strongly delay perivascular transport, and thus
-impair intracranial enrichment. These findings transfer, reconcile,
-integrate and extend insights from clinical, experimental, and
-theoretical studies, and lay a new foundation for in-silico studies of
-personalized intrathecal drug delivery and brain clearance.
+impair intracranial enrichment.
+These findings help to integrate and extend insights from clinical, experimental, and theoretical studies, and provide a framework for future in-silico investigations of personalized intrathecal drug delivery and brain solute transport.
 
@@
 respiratory dispersion factors ($R_c$ and $R_r$) in $\Omega_{\rm CSF}$
-(\Cref{fig:csf}G, I). To compute these spatially-varying factors, we
-combine computational estimates of peak CSF pressure gradients
+(\Cref{fig:csf}G, I). To compute these spatially-varying factors, we rely on a simplified model that combines computational estimates of peak CSF pressure gradients
 (\Cref{fig:csf}F, H) with theoretical estimates for shear-augmented
@@
 by the extension FEniCS\textsubscript{ii}~\cite{kuchta2020assembly}.
-The correctness and accuracy of the numerical solutions were verified by a series of
-numerical verification experiments (see Supplementary information S1.5).
-The majority of the simulation time was spend on the Stokes flow and transport model simulations.
+The correctness and accuracy of the numerical solutions were verified
+by a series of numerical verification experiments (see Supplementary
+information S1.5).
+
+The majority of the simulation time was spent on the Stokes flow and transport model simulations.
 Each Stokes flow simulation utilizes a hybrid parallel approach with 16 MPI processes and 4 threads each, requiring approximately 300 GB of RAM and completing in 23 minutes.

### 2025-06-26 paper/main.tex
Updates from Overleaf

@@
 impair intracranial enrichment.
-These findings help to integrate and extend insights from clinical, experimental, and theoretical studies, and provide a framework for future in-silico investigations of personalized intrathecal drug delivery and brain solute transport.
+These findings transfer, reconcile, integrate and extend insights from clinical, experimental, and theoretical studies, and provide a framework for future in-silico studies of personalized intrathecal drug delivery and brain clearance.
 

### 2025-06-26 paper/supplementary.tex
Update title

@@
 
-\title{Supplementary information: In-silico molecular enrichment and clearance of the human intracranial space}
+\title{Supplementary information for In-silico molecular transport via perivascular networks in the human intracranial space}
 \author[1]{Marius Causemann}
@@
   Supplementary methods and discussion accompanying the main text of
-  \emph{In-silico molecular enrichment and clearance of the human
-  intracranial space}.
+  \emph{In-silico molecular transport via perivascular networks in the human intracranial space}.
 \end{abstract}

### 2025-06-26 paper/supplementary.tex
Update title

--- 2025-06-20 paper/supplementary.tex
@@
 
-\title{Supplementary information: In-silico molecular enrichment and clearance of the human intracranial space}
+\title{\textcolor{blue}{Supplementary information for In-silico molecular transport via perivascular networks in the human intracranial space}}
 \author[1]{Marius Causemann}
@@
   Supplementary methods and discussion accompanying the main text of
-  \emph{In-silico molecular enrichment and clearance of the human
-  intracranial space}.
+  \emph{\textcolor{blue}{In-silico molecular transport via perivascular networks in the human intracranial space}}.
 \end{abstract}

### 2025-06-26 paper/main.tex
Update title
Updates after reading the response.

@@
 
-\title{In-silico molecular enrichment and clearance of the human intracranial space}
+\title{\textcolor{blue}{In-silico molecular transport via perivascular networks in the human intracranial space}}
 \author[1]{Marius Causemann}
@@
      \caption{
-     \textbf{In-silico modelling of transport of the PVS, SAS and parenchyma.}
+     \textbf{In-silico modelling of tracer transport in the PVS, SAS and parenchyma.}
      A) Illustration of the model geometry including the CSF-filled spaces (ventricles and SAS) $\Omega_{\rm CSF}$, the parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and veins $\Lambda_{\rm vein}$, as well as their interfaces and boundaries;
@@
 (ii) reduced pulsatility and thus decreased dispersion, and (iii) higher pulsatility with increased dispersion.
-We computed the reduced and increased pulsatility scenarios by halving and doubling the inflow rates associated with both cardiac and respiratory pulsations, which reduced the dispersion factors by factor of 0.25 and 4, respectively.
-Note that behaviour directly reflects the nature of our model: The Stokes flow is linear in its inflow boundary conditions, and the diffusion enhancement factor quadratic in the predicted pressure gradient.
+We computed the reduced and increased pulsatility scenarios by halving and doubling the inflow rates associated with both cardiac and respiratory pulsations, which reduced the dispersion factors by 0.25$\times$ and 4$\times$, respectively.
+Note that behaviour directly reflects the nature of our model: Stokes flow is linear in its inflow boundary conditions, and the diffusion enhancement factor is quadratic in the predicted pressure gradient.
 Without CSF production, transport is considerably delayed

### 2025-06-26 paper/supplementary.tex
Updates in the supplementary

@@
 \renewcommand\thesection{S\arabic{section}}    
+\renewcommand\thetable{S\arabic{table}}
 
@@
 
-These sections provide more detailed descriptions of the mathematical models and numerical approximations considered.
-We solve the models presented in Methods numerically
-using finite element methods. The following subsections provide more
-details, model-by-model.
-
-\subsection{Nomenclature and notation} We begin by including \Cref{table:geometry_notation}, \Cref{table:variable}, and \Cref{table:mesh}, which summarize the common notation used for the geometry, model variables and parameters, and finite element discretization, respectively.
-\begin{table}[h!] 
-\centering
-\caption{Geometrical notation, refer to (Fig.~1A, Fig.~2A--C)}
-\label{table:geometry_notation}
-\begin{tabular}{ll}
-\hline
-\textbf{Symbol} & \textbf{Description} \\
-\hline
-$\Omega$ & Full intracranial domain: $\Omega = \Omega_{\mathrm{PAR}} \cup \Omega_{\mathrm{CSF}}$ \\
-$\Omega_{\mathrm{PAR}}$ & Brain parenchyma domain \\
-$\Omega_{\mathrm{CSF}}$ & Cerebrospinal‐fluid (CSF) domain \\
-$\partial\Omega_{\mathrm{PAR}}\cap\partial\Omega_{\mathrm{CSF}}$ & Parenchyma–CSF interface: $\partial \Omega_{\rm CSF} \bigcap \partial
-\Omega_{\rm PAR} = \Gamma_{\rm LV} \cup \Gamma_{\rm pia}$ \\
-$\Gamma_{\mathrm{LV}}$ & Lateral‐ventricle surface (part of parenchyma–CSF interface) \\
-$\Gamma_{\mathrm{pia}}$ & Pial surface (remainder of parenchyma–CSF interface) \\
-$\Gamma_{\mathrm{SSAS}}$ & Lower CSF boundary toward spinal subarachnoid space \\
-$\Gamma_{\mathrm{AM}}$ & Outer CSF boundary toward arachnoid/dura \\
-$\Gamma_{\mathrm{AM\text{-}L}}$, $\Gamma_{\mathrm{AM\text{-}U}}$ & Lower/upper parts of $\Gamma_{\mathrm{AM}}$ \\
-$\Gamma_{\mathrm{SC}}$ & Parenchyma–spinal‐cord boundary: $\partial\Omega_{\mathrm{PAR}}\setminus\partial\Omega_{\mathrm{CSF}}$ \\
-$\Lambda_{a},\Lambda_{v}$ & Arterial/venous PVS centerline networks \\
-$\Lambda_{i}$ & Individual PVS centerline segment ($\Lambda_a^i$, $\Lambda_v^i$ corresponding to arterial and venous segments) \\
-$L_i$ & Length of a PVS segment $\Lambda_i$ \\
-$\Omega_{i}$ & Perivascular space around $\Lambda_{i}$ \\
-$R_{1}^{i},\,R_{2}^{i}$ & Inner/outer radii of the annular PVS $\Omega_{i}$ \\
-$\Gamma_{a},\,\Gamma_{v}$ & Outer lateral surfaces of periarterial/perivenous PVSs \\
-$P$ & Perimeter of the PVS cross‐section \\
-$\mathcal V$ & Graph nodes of the PVS network \\
-$\mathcal B\subset  \mathcal V$ & Internal (bifurcation) nodes \\
-$\mathcal N\subset  \mathcal V$ & End (leaf) nodes \\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{table}[htb!]
-\centering 
-\caption{Variable and parameter notation}
-\label{table:variable}
-\begin{tabular}{ll}
-\hline
-\textbf{Symbol} & \textbf{Description} \\
-\hline
-$\phi$ & Fluid volume fraction (porosity) \\
-$D$ & Effective diffusion coefficient ($D_{\mathrm{PAR}} = D^*$ or $D_{\mathrm{CSF}} = D(1+R)$ \\
-$\mathbf{u}$ & Convective velocity ($\mathbf{u}_{\mathrm{PAR}} = 0$ or $\mathbf{u}_{\mathrm{CSF}}$ computed from steady state flow) \\
-$c$ & Solute concentration in 3D domains \\
-$\bar c$ & Lateral average of $c$ on the PVS boundary \\
-$\xi$ & Permeability coefficient between the periarterial and perivenous networks
-and their surroundings\\
-$\delta_{\Gamma}$ & Dirac term on the PVS outer surfaces \\
-$\beta_{\mathrm{pia}}$ & Permeability of pial/ventricle interface \\
-$\beta_{\mathrm{exit}}$ & Efflux rate on $\Gamma_{\mathrm{AM\text{-}U}}$ \\
-$g_{\mathrm{influx}}$ & Influx on $\Gamma_{\mathrm{SSAS}}$ \\
-$A(s)$ & Cross‐sectional area of PVS at $s$ \\
-$\hat D$ & 1D effective diffusion in PVS \\
-$\hat u$ & 1D convective velocity in PVS \\
-$p$ & Pressure in CSF (Stokes) or PVS (network) \\
-$\hat{q}$ & Volumetric flux in PVS network \\
-$\kappa$ & Hydraulic conductivity in annular PVS \\
-$\mu$ & Dynamic viscosity of CSF \\
-$R_{\rm CSF, 0}$ & Resistance parameter for CSF efflux \\
-$u_{\mathrm{in}}$ & Imposed fluid influx on $\Gamma_{\mathrm{LV}}$ \\
-$\omega,\;\rho$ & Angular frequency, CSF density (for dispersion estimates) \\
-$\alpha_c,\, \alpha_r$ & Womersley number (cardiac, respiratory) \\
-$h_{\rm SAS}$ & Mean width of the subarachnoid space (for dispersion estimates) \\
-$R_c,\,R_r$ & Spatially-varying dispersion enhancement factors (cardiac, respiratory) \\
-
-\hline
-\end{tabular}
-\end{table}
-
-\begin{table}[h!]
-\centering
-\caption{Common mesh and finite‐element notation}
-\label{table:mesh}
-\begin{tabular}{ll}
-\hline
-\textbf{Symbol} & \textbf{Description} \\
-\hline
-$\mathcal{T}$ & Tetrahedral mesh of $\Omega$ \\
-$\mathcal{T}_{\mathrm{PAR}},\,\mathcal{T}_{\mathrm{CSF}}$ & Restrictions of $\mathcal T$ to $\Omega_{\mathrm{PAR}},\Omega_{\mathrm{CSF}}$ \\
-$E$ & Tetrahedral cell in $\mathcal T$ \\
-$\mathcal F_{i}$ & Interior facets of $\mathcal T$ \\
-$\mathcal F_{i,\mathrm{PAR}},\, \mathcal F_{i,\mathrm{CSF}}$ & Interior facets in $\Omega_{\mathrm{PAR}}$, $\Omega_{\mathrm{CSF}}$ \\
-$\mathcal F_{e}$ & Exterior facets of CSF mesh ($\mathcal F_{\mathrm{pia}}\cup \mathcal F_{\mathrm{LV}}\cup \mathcal F_{\mathrm{AM\text{-}L}}\cup \mathcal \mathcal F_{\mathrm{AM\text{-}U}}\cup \mathcal F_{\mathrm{SSAS}}$) \\
-$\mathcal F _{\mathrm{pia}},\,\mathcal F_{\mathrm{LV}},\, \mathcal F_{\mathrm{AM\text{-}L}},\,\mathcal F_{\mathrm{AM\text{-}U}},\, \mathcal F_{\mathrm{SSAS}}$ & Facets on corresponding boundaries \\
-$V_{h}$ & Discontinuous Galerkin space on $\Omega$ of degree $1$\\
-$\widehat V_{h}$ & Continuous Galerkin space on $\Lambda$ of degree $1$.  \\
-% $\widehat V_{h,0}$, $\widehat V_{h,g}$ & Continuous Galerkin space on network with homogeneous or prescribed BCs \\
-$\mathrm{BDM}_{2}$ & Brezzi–Douglas–Marini space of degree 2 \\
-$Q_{h}$ & Pressure space on $\mathcal T_{\mathrm{CSF}}$  \\
-$\mathcal{I}_a, \mathcal{I}_v$ & 1D meshes of arterial/venous networks $\Lambda_{a},\Lambda_{v}$ \\
-$\widehat{Z}_{h}(\mathcal{I})$ & Space of piecewise constants ($P^0$) on network mesh $I$ \\
-$\tau$ & Time step size \\
-\hline
-\end{tabular}
-\end{table}
+These sections provide more detailed descriptions of the mathematical
+models and numerical approximations considered.
 
@@
 centerlines $\Lambda^i_{a}$ of the arterial tree, and a perivenous
-network $\Lambda_v$ associated with the centerlines $\Lambda^i_{v}$
-of the veins. Recall that we represent the vascular domains as the
-union of cylindrical vessels of radius $R_1^i$ surrounding the
-centerlines $\Lambda^i$. Moreover, we consider the surrounding
-perivascular spaces as the union of annular cylinders $\Omega^i$ of
-inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$ and thus of
-width $R_2^i - R_1^i$, and remark that we interpret $R^i_2$ and $R^i_1$ as temporal averages (fixed in time), as the pulsations considered in \Cref{sec:app:peristalsis} are beyond the temporal resolution of our model.
-We denote the outer lateral surface of the
-periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$,
-respectively. We omit the subscript $a$ or $v$ when referring to any
-such network, vessel segment or perivascular outer surface. We assume
-that $\Lambda$ is parametrized by the coordinate $s$, and with a minor
-abuse of notation, simply write $s \in \Lambda$ to represent the point
-${\bm \lambda}(s)$ on $\Lambda$ corresponding to $s$. Finally, we view
+network $\Lambda_v$ associated with the centerlines $\Lambda^i_{v}$ of
+the veins. Recall that we represent the vascular domains as the union
+of cylindrical vessels of radius $R_1^i$ surrounding the centerlines
+$\Lambda^i$. Moreover, we consider the surrounding perivascular spaces
+as the union of annular cylinders $\Omega^i$ of inner radius $R_1^i$
+and outer radius $R_2^i > R_1^i$ and thus of width $R_2^i - R_1^i$,
+and remark that we interpret $R^i_2$ and $R^i_1$ as temporal averages
+(fixed in time), as the pulsations considered in
+\Cref{sec:app:peristalsis} are beyond the temporal resolution of our
+model. We denote the outer lateral surface of the periarterial and
+perivenous spaces by $\Gamma_a$ and $\Gamma_v$, respectively. We omit
+the subscript $a$ or $v$ when referring to any such network, vessel
+segment or perivascular outer surface. We assume that $\Lambda$ is
+parametrized by the coordinate $s$, and with a minor abuse of
+notation, simply write $s \in \Lambda$ to represent the point ${\bm
+  \lambda}(s)$ on $\Lambda$ corresponding to $s$. Finally, we view
 each perivascular network both as a geometric domain and as a directed
 graph with the centerlines $\{ \Lambda^i \}$ as oriented edges and the
-connections as nodes $\mathcal{V}$.
+connections as nodes $\mathcal{V}$. The notation used for the
+geometry, model variables and parameters is summarized in
+Table~\ref{table:geometry_notation} and \ref{table:variable}.
+\begin{table}[h!] 
+\centering
+\begin{tabular}{ll}
+\toprule
+\textbf{Symbol} & \textbf{Description} \\
+\midrule
+$\Omega$ & Full intracranial domain: $\Omega = \Omega_{\mathrm{PAR}} \cup \Omega_{\mathrm{CSF}}$ \\
+$\Omega_{\rm PAR}$ & Brain parenchyma domain \\
+$\Omega_{\mathrm{CSF}}$ & Cerebrospinal fluid (CSF) domain \\
+$\partial\Omega_{\mathrm{PAR}}\cap\partial\Omega_{\mathrm{CSF}}$ & Parenchyma–CSF interface: $\partial \Omega_{\rm CSF} \bigcap \partial
+\Omega_{\rm PAR} = \Gamma_{\rm LV} \cup \Gamma_{\rm pia}$ \\
+$\Gamma_{\mathrm{LV}}$ & Lateral ventricle surface (part of parenchyma–CSF interface) \\
+$\Gamma_{\mathrm{pia}}$ & Pial surface (remainder of parenchyma–CSF interface) \\
+$\Gamma_{\mathrm{SSAS}}$ & Lower CSF boundary toward spinal subarachnoid space \\
+$\Gamma_{\mathrm{AM}}$ & Outer CSF boundary toward arachnoid/dura \\
+$\Gamma_{\mathrm{AM\text{-}L}}$, $\Gamma_{\mathrm{AM\text{-}U}}$ & Lower/upper parts of $\Gamma_{\mathrm{AM}}$ \\
+$\Gamma_{\mathrm{SC}}$ & Parenchyma–spinal‐cord boundary: $\partial\Omega_{\mathrm{PAR}}\setminus\partial\Omega_{\mathrm{CSF}}$ \\
+$\Lambda_{a},\Lambda_{v}$ & Arterial/venous PVS networks \\
+$\Lambda^{i}$ & Individual PVS centerline segment \\
+$L_i$ & Length of a PVS segment $\Lambda^i$ \\
+$R_{1}^{i},\,R_{2}^{i}$ & Inner/outer radii of the annular PVS \\
+$\Gamma_{a},\,\Gamma_{v}$ & Outer lateral surfaces of periarterial/perivenous PVSs \\
+$A^i$ & Area of the cross‐section for PVS segment $\Lambda^i$ \\
+$P^i$ & Perimeter of the cross‐section for PVS segment $\Lambda^i$ \\
+$\mathcal V$ & Graph nodes of the PVS network \\
+$\mathcal B\subset  \mathcal V$ & Internal (bifurcation) nodes \\
+$\mathcal N\subset  \mathcal V$ & End (leaf) nodes \\
+\bottomrule
+\end{tabular}
+\caption{Geometrical notation, refer to (Fig.~1A, Fig.~2A--C)}
+\label{table:geometry_notation}
+\end{table}
+\begin{table}[htb!]
+\centering 
+\begin{tabular}{ll}
+\toprule
+\textbf{Symbol} & \textbf{Description} \\
+\midrule
+$\phi$ & Fluid volume fraction (porosity) \\
+$D$ & Effective diffusion coefficient ($D_{\mathrm{PAR}} = D^*$ or $D_{\mathrm{CSF}} = D(1+R)$ \\
+$\bm u$ & Convective velocity (${\bm u}_{\mathrm{PAR}} = 0$ or ${\bm u}_{\mathrm{CSF}}$ computed from steady state flow) \\
+$c$ & Solute concentration in 3D domains \\
+$\overline{c}$ & Lateral average of $c$ on the PVS boundary \\
+$\xi$ & Permeability coefficient between the periarterial and perivenous networks
+and their surroundings\\
+$\delta_{\Gamma}$ & Dirac term on the PVS outer surfaces \\
+$\beta_{\mathrm{pia}}$ & Permeability of pial/ventricle interface \\
+$\beta_{\mathrm{exit}}$ & Efflux rate on $\Gamma_{\mathrm{AM\text{-}U}}$ \\
+$g_{\mathrm{influx}}$ & Influx on $\Gamma_{\mathrm{SSAS}}$ \\
+$\hat D$ & Effective diffusion coefficient in the PVS networks \\
+$\hat u$ & Convective velocity in PVS networks \\
+$p$, $\hat{p}$ & Pressures in the CSF spaces and PVS networks \\
+$\hat{q}$ & Volumetric flux in PVS network \\
+$\kappa$ & Hydraulic conductivity in annular PVS \\
+$\mu$ & Dynamic viscosity of CSF \\
+$R_{\rm CSF, 0}$ & Resistance parameter for CSF efflux \\
+$u_{\mathrm{in}}$ & Imposed fluid influx on $\Gamma_{\mathrm{LV}}$ \\
+$\omega,\;\rho$ & Angular frequency, CSF density (for dispersion estimates) \\
+$\alpha_c,\, \alpha_r$ & Womersley number (cardiac, respiratory) \\
+$h_{\rm SAS}$ & Mean width of the subarachnoid space (for dispersion estimates) \\
+$R_c,\,R_r$ & Spatially-varying dispersion enhancement factors (cardiac, respiratory) \\
+\bottomrule
+\end{tabular}
+\caption{Variable and parameter notation}
+\label{table:variable}
+\end{table}
+
+%% \begin{table}[h!]
+%% \centering
+%% \caption{Common mesh and finite element notation}
+%% \label{table:mesh}
+%% \begin{tabular}{ll}
+%% \hline
+%% \textbf{Symbol} & \textbf{Description} \\
+%% \hline
+%% $\mathcal{T}$ & Tetrahedral mesh of $\Omega$ \\
+%% $\mathcal{T}_{\mathrm{PAR}},\,\mathcal{T}_{\mathrm{CSF}}$ & Restrictions of $\mathcal T$ to $\Omega_{\mathrm{PAR}},\Omega_{\mathrm{CSF}}$ \\
+%% $E$ & Tetrahedral cell in $\mathcal T$ \\
+%% $\mathcal F_{i}$ & Interior facets of $\mathcal T$ \\
+%% $\mathcal F_{i,\mathrm{PAR}},\, \mathcal F_{i,\mathrm{CSF}}$ & Interior facets in $\Omega_{\mathrm{PAR}}$, $\Omega_{\mathrm{CSF}}$ \\
+%% $\mathcal F_{e}$ & Exterior facets of CSF mesh ($\mathcal F_{\mathrm{pia}}\cup \mathcal F_{\mathrm{LV}}\cup \mathcal F_{\mathrm{AM\text{-}L}}\cup \mathcal \mathcal F_{\mathrm{AM\text{-}U}}\cup \mathcal F_{\mathrm{SSAS}}$) \\
+%% $\mathcal F _{\mathrm{pia}},\,\mathcal F_{\mathrm{LV}},\, \mathcal F_{\mathrm{AM\text{-}L}},\,\mathcal F_{\mathrm{AM\text{-}U}},\, \mathcal F_{\mathrm{SSAS}}$ & Facets on corresponding boundaries \\
+%% $V_{h}(\mathcal{T})$ & Space of piecewise linear polynomials on $\mathcal{T}$\\
+%% $Q_{h}(\mathcal{T})$ & Pressure space on $\mathcal T_{\mathrm{CSF}}$  \\
+%% $\mathcal{I}_a, \mathcal{I}_v$ & 1D meshes of arterial/venous networks $\Lambda_{a},\Lambda_{v}$ \\
+%% $\widehat V_{h}(\mathcal{I})$ & Space of continuous piecewise linear polynomials on $\mathcal{I}$, $\mathcal{I}$ omitted when superfluous \\
+%% $\widehat{Z}_{h}(\mathcal{I})$ & Space of piecewise constants ($P^0$) on $\mathcal{I}$, , $\mathcal{I}$ omitted when superfluous \\% $\widehat V_{h,0}$, $\widehat V_{h,g}$ & Continuous Galerkin space on network with homogeneous or prescribed BCs
+%% $\mathrm{BDM}_{2}$ & Brezzi–Douglas–Marini space of degree 2 \\
+%% $\tau$ & Time step size \\
+%% \hline
+%% \end{tabular}
+%% \end{table}
 
@@
 
-\textit{Finite element solution of the incompressible Stokes equations}.
+\subsubsection{Finite element solution of the incompressible Stokes equations}
 We consider an H(div)-based finite element approximation of the incompressible Stokes equations~\eqref{eq:stokes} defined over $\Omega_{\rm CSF}$ with the boundary conditions~\eqref{eq:stokes_bcs}.
@@
 
-\textit{Finite element solution of the perivascular network equations}. 
+\subsubsection{Finite element solution of the perivascular network equations} 
 We consider meshes $\mathcal{I}_a, \mathcal{I}_v$ representing a
@@
 \begin{align}
-  \hat{Z}_h(\mathcal{I}) &= \{z \in L^2(\Lambda) ; \, z \vert_{\Lambda_i} \in \mathrm{P}^{0}(\Lambda_i) \, \foralls \Lambda^i \in \mathcal{I}\}, \\
-  \hat{V}_{h, g}(\mathcal{I}) &= \{ v \in C^0(\Lambda) ; \, v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda_i), \, v(x) = g(x) \, \foralls x \in \mathcal{N} \} .
-  %Q_h^{\rm PVS} &= \{ q_h \in C^0(\Lambda),  q_h \vert_{\Lambda^i} \in \mathrm{P}^k(\Lambda_i), q_h \vert_{\Gamma_{\rm Pia}} = p_{\rm CSF} \}
+  \hat{Z}_h(\mathcal{I}) &= \{z \in L^2(\Lambda) ; \, z \vert_{\Lambda^i} \in \mathrm{P}^{0}(\Lambda^i) \, \foralls \Lambda^i \in \mathcal{I}\}, \\
+  \hat{V}_{h, g}(\mathcal{I}) &= \{ v \in C^0(\Lambda) ; \, v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda^i), \, v(x) = g(x) \, \foralls x \in \mathcal{N} \} .
+  %Q_h^{\rm PVS} &= \{ q_h \in C^0(\Lambda),  q_h \vert_{\Lambda^i} \in \mathrm{P}^k(\Lambda^i), q_h \vert_{\Gamma_{\rm Pia}} = p_{\rm CSF} \}
 \end{align}
@@
 where $\langle \cdot, \cdot \rangle$ denotes the $L^2(\mathcal{I})$-inner product and is defined segment-wise.
-For a stability and convergence analysis of this discrete model, we refer to the reference\cite{gjerde2024directional}.
+For a stability and convergence analysis of this discrete model, we refer to previously presented results\cite{gjerde2024directional}.
 
@@
 
-We use the theoretical framework previously introduced by Gjerde et al.~\cite{gjerde2023directional} to compute an analytic estimate of the time-averaged (or \emph{net}) flow rates $\langle Q'_i \rangle$ (mm$^3$/s) induced by peristaltic pumping in a perivascular network $\Lambda = \cup_i \Lambda_i$.
+We use the theoretical framework previously introduced by Gjerde et al.~\cite{gjerde2023directional} to compute an analytic estimate of the time-averaged (or \emph{net}) flow rates $\langle Q'_i \rangle$ (mm$^3$/s) induced by peristaltic pumping in a perivascular network $\Lambda = \cup_i \Lambda^i$.
 The motion of the (inner) vascular wall is assumed to be described by a periodic traveling (peristaltic) wave of relative amplitude $\varepsilon$, wave length $\lambda$ (mm) and frequency $f$, acting normal to the wall and traveling in the antegrade direction (with the blood flow).
 By definition, the wave number is $k = 2 \pi/\lambda$ and the angular frequency is $\omega = 2 \pi f$.
-Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L_i$, baseline inner radius $R_1^i$, fixed outer radius $R_2^i$, and outer-to-inner ratio $\beta_i = R_2^i/R_1^i$.
+Each PVS segment $\Lambda^i$ has length $L_i$ with wave-relative length $\ell_i = k L_i$, baseline inner radius $R_1^i$, fixed outer radius $R_2^i$, and outer-to-inner ratio $\beta_i = R_2^i/R_1^i$.
 These geometric parameters and the assumption of annular cylindrical PVS segments yield hydraulic resistances and additional characteristic parameters, see \cite{gjerde2023directional} for the complete definitions and schematics.
@@
 assign this flow rate $\langle Q'_n \rangle$ to each of the branches
-$\Lambda_i$ that form the path $E_n^j$, thus yielding $\langle Q'_i
-\rangle$ for each perivascular segment $\Lambda_i$ while ensuring that
+$\Lambda^i$ that form the path $E_n^j$, thus yielding $\langle Q'_i
+\rangle$ for each perivascular segment $\Lambda^i$ while ensuring that
 mass is conserved (\Cref{fig:app:peristalsis}E). For the segment(s)
@@
 
-\textit{Finite element solution of coupled 3D-1D solute transport equations}. 
+\subsubsection{Finite element solution of coupled 3D-1D solute transport equations}
 We consider the system of coupled 3D-1D solute transport equations
@@
 
- \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} complemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0 = 0 $ is the following:
-
+ \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} complemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0 = 0 $ is the following.
 For $n = 1,2, \ldots $, find  $c_h^n \in V_h$ and $\hat c^n_h \in \hat V_h$ such that the coupled equations \eqref{eq:3d_dg_disc} and \eqref{eq:1d_dg_disc} hold for all $v \in V_h$ and for all $\hat v \in \hat V_h$.
@@
 
-\subsection{Extended model validation}
+\subsection{Extended comparison with literature}
 \label{sec:app:model_validation}
@@
 Investigating the link between cardiac pulsatility and the resulting diffusion enhancement factor, we compute for $R$ different levels of cardiac pulsatility.
-We reduce the inflow boundary conditions of $6\,$ml/s (across the SAS outer surface) and $0.31\,$ml/s (across the lateral ventricle surface) by $25\,\%$, $50\,\%$, $75\,\%$ and $100\,\%$ to model lower cardiac pulsatility, revealing that $R$ scales quadratically with the reduction of inflow (\cref{fig:R_comparison}).
+We reduce the inflow boundary conditions of $6\,$ml/s (across the SAS outer surface) and $0.31\,$ml/s (across the lateral ventricle surface) by $25\,\%$, $50\,\%$, $75\,\%$ and $100\,\%$ to model lower cardiac pulsatility, revealing that $R$ scales quadratically with the reduction of inflow (\Cref{fig:R_comparison}).
 This behaviour directly reflects the nature of our model - the Stokes flow is linear in its boundary conditions, and the diffusion enhancement factor quadratic in the predicted pressure gradient - and simultaneously underscores the substantial role of pulsatility in shaping intracranial transport.

### 2025-06-26 paper/main.tex
Minor edits in the main text after reading through the diff.
Reduced -> changed

@@
 
-\title{\textcolor{blue}{In-silico molecular transport via perivascular networks in the human intracranial space}}
+\title{\textcolor{black}{In-silico molecular transport via perivascular networks in the human intracranial space}}
 \author[1]{Marius Causemann}
@@
 (ii) reduced pulsatility and thus decreased dispersion, and (iii) higher pulsatility with increased dispersion.
-We computed the reduced and increased pulsatility scenarios by halving and doubling the inflow rates associated with both cardiac and respiratory pulsations, which reduced the dispersion factors by 0.25$\times$ and 4$\times$, respectively.
+We computed the reduced and increased pulsatility scenarios by halving and doubling the inflow rates associated with both cardiac and respiratory pulsations, which changed the dispersion factors by 0.25$\times$ and 4$\times$, respectively.
 Note that behaviour directly reflects the nature of our model: Stokes flow is linear in its inflow boundary conditions, and the diffusion enhancement factor is quadratic in the predicted pressure gradient.
@@
   \caption{Model parameters. $\dagger$/$\ast$ denotes
-    computed/estimated within this work. The diffusion coefficients
+    computed/estimated within this work. The diffusion coefficient
     $D^{\rm Gad}$ represent the diffusion coefficient of Gadubutrol in

### 2025-06-26 paper/main.tex
Work towards complying with journal style.

@@
 \Gamma_v$. Additionally, in \eqref{eq:multi_transport_3d_PAR}, $\phi = \phi_{\rm ECS}$
-is the brain extracellular volume fraction\footnote{We set $\phi =
-1.0$ in $\Omega_{\rm CSF}$ and therefore omit $\phi$
-in~\eqref{eq:multi_transport_3d_CSF}.}, $D^{\ast}$ is the effective
+is the brain extracellular volume fraction, $D^{\ast}$ is the effective
 diffusion coefficient in the parenchyma, and $\bm u_{\rm PAR}$ is an
@@
 
+\section*{Data availability}
+
+Numerical simulation results generated as part of this study are
+openly available on Zenodo with the identifier doi:
+10.5281/zenodo.14749163. This archive includes surface triangulations
+used for geometry representation, computational meshes (3D meshes and
+1D networks), segmentations, sample simulation outputs from the
+baseline and high PVS flow models, and animations.
+
+\section*{Code availability}
+
+The software developed to generate and analyze the numerical results
+of this study is openly available on Zenodo with the identifier doi:
+10.5281/zenodo.14749163. In addition to simulation and analysis
+software, this archive also includes surface triangulations used for
+geometry representation, computational meshes (3D meshes and 1D
+networks), segmentations, sample simulation outputs from the baseline
+and high PVS flow models, and animations.
+
+\bibliography{references}
+
 \section*{Acknowledgments}
@@
 
-\section*{Data availability}
-
-The software developed to generate and analyze the numerical results
-of this study is openly available on Zenodo with the identifier doi:
-10.5281/zenodo.14749163. In addition to simulation and analysis
-software, this archive also includes surface triangulations used for
-geometry representation, computational meshes (3D meshes and 1D
-networks), segmentations, sample simulation outputs from the baseline
-and high PVS flow models, and animations.
-
-\bibliography{references}
-
 %% \newpage

### 2025-06-27 paper/main.tex
change molecular enrichment with mol transport or tracer enrichment

@@
 However, our understanding of these mechanisms and their interplay remains incomplete, in part due to the complexity of integrating insights across scales, between species and from different modalities.
-Here, we combine mixed-dimensional modelling, multi-modal magnetic resonance images, and high performance computing to construct and explore a high-fidelity in-silico model of human intracranial molecular enrichment.
+Here, we combine mixed-dimensional modelling, multi-modal magnetic resonance images, and high performance computing to construct and explore a high-fidelity in-silico model of human intracranial molecular transport.
 This model predicts the temporo-spatial spreading of a solute within an image-derived geometric representation of the subarachnoid space, ventricular system and brain parenchyma, including networks of surface perivascular spaces (PVSs).
-Our findings highlight the significant impact of cerebrospinal fluid (CSF) production and intracranial pulsatility on molecular enrichment following intrathecal tracer injection.
+Our findings highlight the significant impact of cerebrospinal fluid (CSF) production and intracranial pulsatility on molecular transport following intrathecal tracer injection.
 We demonstrate that low-frequency vasomotion induces moderate CSF flow in surface PVS networks which substantially enhances tracer enrichment, and that impaired enrichment is a direct natural consequence of enlarged surface PVSs.
@@
 sleep and conversely sleep-deprivation play a definite yet enigmatic
-role in modulating molecular enrichment and
+role in modulating molecular transport and
 clearance~\cite{xie2013sleep, eide2021sleep, eide2022altered,
@@
   watts2019measuring, eide2024functional}, the in-silico predictions
-accurately represent molecular enrichment patterns, timing and
+accurately represent tracer enrichment patterns, timing and
 intercompartmental distributions. This open platform~\cite{ZENODO}
@@
 \includegraphics[width = 0.95\textwidth]{figures/figure2.png}
-\caption{\textbf{The balance between CSF pulsatility and production shapes molecular enrichment patterns.}
+\caption{\textbf{The balance between CSF pulsatility and production shapes tracer enrichment patterns.}
 A) CSF spaces: the third and fourth ventricle (V3\&4), lateral ventricles (LV) and SAS; 
@@
 
-\subsection*{Enlarged PVSs in the SAS delay periarterial and intracranial molecular enrichment}
+\subsection*{Enlarged PVSs in the SAS delay periarterial and intracranial tracer enrichment}
 
@@
 production and intracranial pulsatility significantly shapes the
-large-scale features of molecular enrichment, with e.g.~ventricular
+large-scale features of tracer enrichment, with e.g.~ventricular
 tracer reflux after intrathecal injection in the absence of CSF
@@
 % Tracer enrichment predictions
-While molecular enrichment of the brain and surrounding CSF spaces
+While tracer enrichment of the brain and surrounding CSF spaces
 have been extensively studied in animal models, especially in murine
@@
 here, with earlier arrival times along the PVSs, clear enrichment in
-adjacent spaces, and accelerated molecular enrichment
+adjacent spaces, and accelerated tracer enrichment
 overall. Moreover, sleep is linked with low-frequency oscillations in

### 2025-06-27 paper/main.tex
change molecular enrichment with mol transport or tracer enrichment

@@
 We have presented a high-fidelity in-silico model of molecular
-enrichment and clearance in human intracranial spaces, enabling
+transport and clearance in human intracranial spaces, enabling
 tailored predictions of the influence of CSF space and (peri)vascular
@@
 alcohol intake~\cite{lundgaard2018beneficial} affect molecular
-enrichment and clearance of the brain. In particular, pial arteries
+transport and clearance of the brain. In particular, pial arteries
 display higher-amplitude low-frequency vasomotion during NREM and
@@
 level of dispersion in the CSF spaces strongly shapes molecular
-enrichment patterns. While we, at baseline, account for the dispersion
+transport patterns. While we, at baseline, account for the dispersion
 induced by cardiac and respiratory pulsatility, the high dispersion

### 2025-12-03 scripts/evaluate_pvs_flow.py
white and black plotting

@@ -23,12 +23,12 @@ m2mm = 1e3
 m2mum = 1e6
 Pa2mPa = 1e3
 
-def plot_pvs_velocity(mesh, uh_mag, artery_radii, filename):
+def plot_pvs_velocity(mesh, uh_mag, artery_radii, filename, col="white"):
     CG1 = FunctionSpace(mesh, "CG", 1)
     bar_args=dict(title="velocity (μm/s)", vertical=False,
-                    height=0.07, width=0.6, position_x=0.2,
-                    position_y=0.0, title_font_size=28, font_family="arial",
-                    label_font_size=28, fmt="%.0f")
+                    height=0.07, width=0.6, position_x=0.2, color=col,
+                    position_y=0.0, title_font_size=32, font_family="arial",
+                    label_font_size=32, fmt="%.0f")
     topology, cell_types, x = create_vtk_structures(CG1)
     grid = pv.UnstructuredGrid(topology, cell_types, x)
     grid["u"] = uh_mag.vector()[:] *m2mum
@@ -71,8 +71,8 @@ def compute_pvs_flow(pvs_flow_file):
     plot_dir = f"plots/{'/'.join(modelstr)}"
     os.makedirs(plot_dir, exist_ok=True)
 
-    plot_pvs_velocity(mesh, uh_mag, pvs_radii,
-                      plot_dir + "/vel3D.png")
+    for col in ["white", "black"]: plot_pvs_velocity(mesh, uh_mag, pvs_radii,
+                      plot_dir + f"/vel3D_{col}.png", col=col)
 
     segments, segids ,_ = color_branches(mesh)
     dxs = Measure("dx", mesh, subdomain_data=segments)

### 2025-12-03 scripts/label_arteries.py

@@ -30,7 +30,7 @@ def get_tangent(v, n):
 
 if __name__=="__main__":
    filename = "plots/meshplots/arteries.png"
-   filename_labeled = "plots/meshplots/labeled_arteries.png"
+   filename_labeled = "plots/meshplots/labeled_arteries"
    art = pv.read("mesh/networks/arteries_tubes.vtk").triangulate()
 
    art["radius"] *= 1e3
@@ -47,79 +47,82 @@ if __name__=="__main__":
 
    ex = art.extract_points(np.unique(idx))
    #from IPython import embed; embed()
-   bar_args=dict(title="radius (mm)", vertical=False,
-                     height=0.07, width=0.6, position_x=0.2,
-                     position_y=0.0, title_font_size=28,
-                     label_font_size=28, fmt="%.1f")
-   pl = pv.Plotter(off_screen=True, window_size=window_size)
-   pl.add_mesh(art, cmap=cmap, scalar_bar_args=bar_args)
-   pl.add_mesh(ex, show_scalar_bar=False, color=col)
-   pl.camera_position = 'xz'
-   #pl.camera.roll += 5
-   pl.camera.azimuth += 165
-   pl.camera.elevation += 10
-   pl.camera.zoom(1.85)
-   pl.camera.focal_point = np.array(pl.camera.focal_point) + np.array([0,0,-0.008])
+   for textcol in ["white", "black"]:
 
-   import vtk
-   coordinate = vtk.vtkCoordinate()
-   coordinate.SetCoordinateSystemToWorld()
+      bar_args=dict(title="radius (mm)", vertical=False,color=textcol,
+                        height=0.07, width=0.6, position_x=0.2,
+                        position_y=0.0, title_font_size=28,
+                        label_font_size=28, fmt="%.1f")
+      
+      pl = pv.Plotter(off_screen=True, window_size=window_size)
+      pl.add_mesh(art, cmap=cmap, scalar_bar_args=bar_args)
+      pl.add_mesh(ex, show_scalar_bar=False, color=col)
+      pl.camera_position = 'xz'
+      #pl.camera.roll += 5
+      pl.camera.azimuth += 165
+      pl.camera.elevation += 10
+      pl.camera.zoom(1.85)
+      pl.camera.focal_point = np.array(pl.camera.focal_point) + np.array([0,0,-0.008])
 
-   df = pd.DataFrame(pointlabels, columns=["label", "coord"])
+      import vtk
[…]

### 2025-12-03 paper/main.tex

@@
 This model predicts the temporo-spatial spreading of a solute within an image-derived geometric representation of the subarachnoid space, ventricular system and brain parenchyma, including networks of surface perivascular spaces (PVSs).
-Our findings highlight the significant impact of cerebrospinal fluid (CSF) production and intracranial pulsatility on molecular transport following intrathecal tracer injection.
+Our findings highlight the significant impact of cerebrospinal fluid (CSF) production and intracranial pulsatility on tracer enrichment following intrathecal tracer injection.
 We demonstrate that low-frequency vasomotion induces moderate CSF flow in surface PVS networks which substantially enhances tracer enrichment, and that impaired enrichment is a direct natural consequence of enlarged surface PVSs.
@@
 flow velocities, strongly delay perivascular transport, and thus
-impair intracranial enrichment.
-These findings transfer, reconcile, integrate and extend insights from clinical, experimental, and theoretical studies, and provide a framework for future in-silico studies of personalized intrathecal drug delivery and brain clearance.
+impair intracranial enrichment. These findings transfer, reconcile, 
+integrate and extend insights from clinical, experimental, and theoretical 
+studies, and provide a framework for future in-silico studies of personalized
+ intrathecal drug delivery and brain clearance.
 
